### 2017-08-30 memos/toy_analytic_survey_20170728.tex
initial commit

%\documentclass[12pt,preprint]{aastex}
%\documentclass[preprint2,12pt]{aastex}
\documentclass{emulateapj}
%\usepackage{apjfonts}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}

\shortauthors{Name et al.}
\shorttitle{Best transiting planets}

\begin{document}

% ------------------------------------------------------------------------
% New commands
%
\def\ltsima{$\; \buildrel < \over \sim \;$}
\def\lsim{\lower.5ex\hbox{\ltsima}}
\def\gtsima{$\; \buildrel > \over \sim \;$}
\def\gsim{\lower.5ex\hbox{\gtsima}}
\def\tess{{\it TESS} }
\def \teff {T_{\rm eff}}
\def \phir {\Phi_{\rm R}}
\def \fov {24$^{\circ}$}
\def \pixsz {21.1''}
\def \aeff {69.1 cm$^2$ }    
\def \epd {105 mm}

% -------------------------------------------------------------------------
%

\bibliographystyle{apj}

\title{ A Toy Analytic Transit Survey }

\author{
  LGB, KM, JNW
}

% \journalinfo{Draft version}
\slugcomment{Memo for internal use}

%\altaffiltext{1}{Princeton University}

\begin{abstract}

We derive the signal to noise distribution of detected planets in a toy transit 
survey. We give corresponding expressions for the number of detected planets. 
We then discuss the effect of various errors that one makes by ignoring 
binarity when deriving occurrence rates.

\end{abstract}

\keywords{planets and satellites:\ detection}

\section*{Statement of problem}

Ah, transiting planets! We learn so much by studying them. But how much can we 
actually learn, and how much is messed up by binarity?

Imagine the following transit survey:
\begin{itemize}
\item Take a magnitude-limited sample of ``stars'' of a single spectral class. 
For instance, G2V dwarfs.  In other words, choose all ``points'' on-sky with 
$m_\lambda < m_{\lambda,{\rm lim}}$, or equivalently with $F_\lambda > 
F_{\lambda,{\rm lim}}$, and also with colors that make you think they are this 
spectral class.
\item The true population of ``points'' (hereafter, systems) contains both 
single and double star systems. Single star systems have luminosity in the 
observed bandpass $L_1$, radii $R_1$, and effective temperature $T_{\rm eff,1}$.
Double star systems have luminosity in the observed bandpass $L_d = 
(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of the 
secondary to the primary. The ratio of the two number densities in a 
volume-limited sample is the binary fraction\footnote{The binary fraction is 
equivalent to the multiplicity fraction if there are no triple, quadruple, 
$\ldots$ systems.}.
\item The true population of planets around these stars is as follows:
	\subitem A fraction $\Gamma_{t,s}$ of stars in single star systems 
	have a planet of radius $R_p$, with orbital period $P$.
	\subitem A fraction $\Gamma_{t,d}$ of each star in a double star 
	system has a planet of radius $R_p$, with orbital period $P$. For instance, 
	if $\Gamma_{t,s} = \Gamma_{t,d} = 0.1$, on average each double 
	system contributes 0.2 planets, and each single system 0.1 planets.
    Any astrophysical difference in planet formation between singles and 
    binaries is captured by these two terms.
\end{itemize}

A signal to noise limited transit survey is then conducted. All planets with 
${\rm S/N} > {\rm (S/N)_{min}}$ are detected. The rest are not.
Note that while the S/N threshold does imply a minimum stellar flux about which 
planets of a given radius and period can be detected, this minimum flux is 
generally different from the minimum used to define the original magnitude cut 
for a sample [see Batalha et al, 2010].

Consider the following set of questions:
\begin{enumerate}
\item How many single and double star systems, respectively, are in the sample?
Correspondingly, how many stars are in the sample?

\item How many planets are in the sample? (Orbiting single stars, and orbiting 
double stars respectively).

\item What is the true occurrence rate?

\item How many planets are detected?

\item What occurrence rate does astronomer A, who has never heard of binary 
star systems, derive for planets of radius $R_p$ and period $P$?

\item What occurrence rate does astronomer B, who accounts for the ``2 for 1'' 
effect of binarity (\textit{i.e.} that the sample actually has more stars than 
astronomer A thought) derive?

\item What about astronomer C, who accounts for ``2 for 1'' \textit{and} 
misclassification due to diluted radii? In other words, astronomer C has lots 
of Keck time, and did high resolution imaging followup on every candidate, and 
correctly classifies the planetary radii.

\item What about astronomer D, who additionally notes the importance of 
completeness?
\end{enumerate}

\section{How many stars are in the sample?}

Let $N_s$ be the number of single star systems, and $N_d$ the number of double 
star systems. Then the total number of stars in the sample can be written
\begin{equation}
N_{\rm stars} = N_s + 2N_d.
\end{equation}

In a magnitude-limited sample in which stars are uniformly distributed in 
volume, the number of stars will be the number density times the volume.
If the volume is taken to be a sphere over which the number density is uniform,
\begin{equation}
N_i = n_{i} \frac{4\pi}{3} d_{{\rm max},i}^3,
\label{eq:number_systems}
\end{equation}
for $i\in{\rm \{single,double\} } \equiv { \{s,d\} }$, and
\begin{equation}
\frac{n_{d}}{n_{s}} = {\rm binary\ fraction} \equiv {\rm BF}
\end{equation}
by definition. The absolute normalization of the number density is a measured 
quantity, as is the binary fraction. For G2V dwarfs, the latter is $\approx 
0.45$ [Duchene \& Kraus, 2013]. The former is given by [Bovy 2017].

$d_{{\rm max},i}$ in Eq.~\ref{eq:number_systems} is the maximum distance 
corresponding to the given magnitude limit:
\begin{equation}
d_{{\rm max},i} = \left(\frac{L_{i}}{4 \pi F_{{\rm lim}}}\right)^{1/2},
\label{eq:d_max}
\end{equation}
where the limiting flux in the bandpass $F_{{\rm lim}}$ can also be 
stated in terms of the limiting magnitude $m_{{\rm lim}}$,
\begin{equation}
m_{{\rm lim}} = m_{0} - \frac{5}{2} \log_{10} \left(\frac{F_{{\rm 
lim}}}{F_{0}}\right),
\end{equation}
for $m_{0}$ a zero-point magnitude and $F_{0}$ its corresponding flux (as 
always, everything is implicitly written in a defined bandpass).

In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$, and as a 
consequence the maximum distance to which binary stars will be selected is 
greater than that of single stars, simply as a consequence of imposing a 
magnitude cut.
The ratio of double to single systems is
\begin{align}
\frac{N_d}{N_s} &= 
	\frac{n_d}{n_s} \left(\frac{d_{{\rm max}, d} }{d_{{\rm max}, s}}\right)^3 \\
&= {\rm BF} \times (1+\gamma_R)^{3/2}.
\end{align}
In the nominal case of twin binaries ($\gamma_R = 1$), with a binary fraction 
${\rm BF} = 0.5$, there are 
$\sqrt{2}$ more binary systems than single systems in the sample.
Correspondingly, there are $2\sqrt{2}$ more stars in binary systems than stars 
in single systems.

As a comment on Eq.~\ref{eq:number_systems}, if 
we wished to write a stellar number density profile that accounted for the 
vertical structure of the Milky Way, we might choose a profile either 
$\propto \exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from 
the galactic midplane and $H$ a scale-height. Both density profiles would lead 
closed form analytic solutions.

\section{How many planets are in the sample?}

The number of planets in the sample is
\begin{align}
N_{\rm planets} =& N_{\rm planets\ in\ single\ star\ systems}  +  \\
				  &\quad\quad N_{\rm planets\ in\ double\ star\ systems} 
				  \nonumber \\
			   =& \Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d.
\end{align}

The factor of 2 accounts for the fact that there are twice as many stars in 
double star systems.

\section{What is the true occurrence rate?}
\label{sec:true_rate}

The ``true occurrence rate'' is the average number of planets per star. Thus

\begin{align}
\Gamma_t &= \frac{N_{\rm planets}}{N_{\rm stars}} \\
\Gamma_t &= \frac{\Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d}{N_s + 2N_d}.
\label{eq:true_occ}
\end{align}

\section{How many planets are detected?}
The total number of planet detections is the sum of the number of planets 
detected in single star systems $N_{{\rm det},s}$ and the number of planets 
detected in double star systems $N_{{\rm det},d}$.
These can be expressed individually. 
The former is
\begin{equation}
N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,{\rm S/N} > {\rm (S/N)_{min}}},
\label{eq:N_det_s}
\end{equation}
where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
star systems of the sample, and $f_{s,({\rm S/N} > {\rm (S/N)_{min}})}$ is the 
fraction of these planets that have signal to noise greater than the minimum 
detection threshold.
\begin{equation}
N_{{\rm det},d} = 2 N_d \Gamma_{t,d} f_{d,{\rm S/N} > {\rm (S/N)_{min}}},
\label{eq:N_det_d}
\end{equation}
where now $2 N_d \Gamma_{t,d}$ is the number of planets in the double star 
systems of the sample, and the completeness term must account for any 
differences in the signal to noise distribution that come from stellar 
binarity.

As a note of warning, Eqs.~\ref{eq:N_det_s} and~\ref{eq:N_det_d} ignore the 
geometric transit probability, $\approx R_\star / a$. This means we are 
assuming that every existing planet in this universe transits. This assumption 
can easily be rectified, if desired.

\subsection{Analytic completeness}
The only terms we have yet to compute are the completeness terms, $f_{i,{\rm 
S/N} > {\rm (S/N)_{min}}}$ for $i\in{\rm \{single,double\} }$. 
We proceed as follows.

The signal ${\rm S}$ for a box-car train transiting planet is
\begin{align}
{\rm S} &= \delta \mathcal{D} \\
&= \left(\frac{R_p}{R_\star}\right)^2 \mathcal{D},
\end{align}
for $R_p$ the planet's radius, $R_\star$ that of its host star, and 
$\mathcal{D}$ the dilution parameter defined as
\begin{equation}
  \mathcal{D} = 
  \Bigg\{\begin{array}{lr}
  L_1 / L_d, & {\rm\ if\ binary\ and\ target\ primary}\\
  \gamma_R L_1 / L_d, & {\rm\ if\ binary\ and\ target\ 
  secondary}\\
  1, & {\rm\ if\ single},
  \end{array}
  \label{eq:dilution}
\end{equation}
where $L_1, L_d,$ and $\gamma_R$ were defined in the opening monograph.

Assuming the only source of noise is Poissonian counting noise, the noise ${\rm 
N}$ can be written
\begin{equation}
{\rm N} = \frac{1}{\sqrt{N_\gamma}},
\end{equation}
for $N_\gamma$ the number of photons received by the detector. This noise model 
is a useful simplification -- see [Howell 2006, pg 75] for the full 
CCD equation.
The number of received photons can be written
\begin{equation}
N_\gamma = F^{\rm N}_\gamma A N_{\rm tra} T_{\rm dur},
\end{equation}
for $F^{\rm N}_\gamma$ the photon number flux from the system [${\rm 
ph\,cm^{-2}\,s^{-1}}$], 
$A$ the detector area, $T_{\rm dur}$ the transit duration, and $N_{\rm tra} $ 
the number of transits observed, which is multiplied in assuming the transits 
are ``phase-folded''.

Thus the signal to noise ratio can be written
\begin{equation}
{\rm S/N} = \delta \mathcal{D} \sqrt{F^{\rm N}_\gamma A N_{\rm tra} T_{\rm 
dur}}.
\label{eq:snr_ivory_tower}
\end{equation}

In passing, given the parameters that define a survey and planet type, 
Eq.~\ref{eq:snr_ivory_tower} would need to be re-expressed with 
$N_{\rm tra}$ roughly the ratio of the observing baseline to the planet period, 
and $T_{\rm dur}$ a function of $R_\star, P, a$, and impact parameter $b$, and 
then perhaps averaged over $b$. We leave them as-is for subsequent 
development.

The interesting term in Eq.~\ref{eq:snr_ivory_tower} that changes between stars 
of the same binarity class in our idealized sample is the square root of 
$F^{\rm N}_\gamma$.
This is the term that leads to a distribution of signal to noises for different 
stars.
The completenesses $f_{i,{\rm S/N} > {\rm (S/N)_{min}}}$ can be directly 
expressed in terms of those probability density functions:
\begin{equation}
f_{i,{\rm S/N} > {\rm (S/N)_{min}}} = 
	\int_{{\rm (S/N)_{min}}}^{\infty} 
		{\rm d}\left(\frac{{\rm S}}{{\rm N}}\right)_i \ 
		{\rm prob}\left(\frac{{\rm S}}{{\rm N}}\right)_i.
\label{eq:completeness_long}
\end{equation}
We keep the subscript $i$ because the signal to noise distributions are 
different for the cases of single star systems ($i=s$) and double star systems 
($i=d$).
Notably:
\begin{itemize}
	\item The dilution differs (Eq.~\ref{eq:dilution}).
	\item The photon number flux from the system differs.
\end{itemize}

To simplify notation, we let $x_i \equiv {\rm (S/N)}_i$, and rewrite 
Eq.~\ref{eq:completeness_long} as
\begin{equation}
f_{i,x > x_{\rm min}} = 
\int_{{\rm x_{\rm min}}}^{\infty} 
{\rm d}x_i \ 
{\rm prob}(x_i).
\label{eq:completeness_short}
\end{equation}

\subsection{Deriving ${\rm prob}(x_i)$}
We want expressions for the probability density function of the observed signal 
to noise ratio, ${\rm prob}(x_i)$, for both the single and binary 
system case.

First, note that a star placed uniformly in the volume of the search space will 
have a probability density function for its distance $r$ from the origin of
\begin{equation}
{\rm prob} (r) = \frac{3 r^2}{d_{\rm max}^3},
\end{equation}
where the appropriate maximum distances should be substituted per 
Eq.~\ref{eq:d_max}.
Noting the transformation rule for probability density functions, we can 
evaluate the probability of a star having a observed number flux $F^{\rm 
N}_{i,\gamma}$ in the bandpass,
\begin{align}
{\rm prob} (F^{\rm N}_{i,\gamma}) 
&= {\rm prob}(r(F^{\rm N}_{i,\gamma}))
	\left| \frac{{\rm d} r}{{\rm d} F^{\rm N}_{i,\gamma}} \right| \\
&= \frac{3}{2 d_{\rm max}^3} c_i^{3/2} (F^{\rm N}_{i,\gamma})^{-5/2},
\label{eq:pdf_observed_flux}
\end{align}
where in the latter equality we have written a ``number luminosity'' $c_i$ 
(units of inverse time) defined for $i\in{\rm \{single,double\} }$ as
\begin{equation}
c_i = 
\Bigg\{\begin{array}{lr}
R_1^2 F^{\rm N}_{s1,\gamma}, & {\rm\ if\ single}\\
R_1^2 F^{\rm N}_{s1,\gamma} + R_2^2 F^{\rm N}_{s2,\gamma} & {\rm\ if\ double}.
\end{array}
\label{eq:c_i}
\end{equation}
In Eq.~\ref{eq:c_i}, $F^{\rm N}_{s1,\gamma}$ and $F^{\rm N}_{s2,\gamma}$ are 
the photon number fluxes at the surfaces of the stars. To derive 
Eq.~\ref{eq:pdf_observed_flux}, we simply scaled these by the distance:
\begin{equation}
F^{\rm N}_{i,\gamma} = \frac{c_i}{r^2}.
\label{eq:flux_at_detector}
\end{equation}

The surface photon number fluxes $F^{\rm N}_{si,\gamma}$ in Eq.~\ref{eq:c_i} 
are usually evaluated numerically, by convolving the wavelength-specific photon 
flux density of a star with the dimensionless spectral response function of the 
instrument. In other words,
\begin{equation}
F^{\rm N}_{s,\gamma} = \int F_\lambda T_\lambda \, {\rm d}\lambda.
\label{eq:pfd}
\end{equation}
The wavelength-specific photon flux density $F_\lambda$ [${\rm 
ph\,cm^{-2}\,s^{-1}\,\AA^{-1}}$] could be from Pickles' library, or could be 
a blackbody function. The transmission function is, up to factor of order 
unity, a step function between two wavelengths $\lambda_{\rm min}$ and 
$\lambda_{\rm max}$.
If we assume a blackbody source, Eq.~\ref{eq:pfd} becomes
\begin{equation}
F^{\rm N}_{s,\gamma} = 8\pi c \left(\frac{k T}{h c}\right)^3 
\int_{hc/(\lambda_{\rm max} kT)}^{hc/(\lambda_{\rm min} kT)}
\frac{u^2}{e^u - 1} \,{\rm d} u,
\end{equation}
which can be evaluated numerically\footnote{It may help in the numerics to note 
that infinite series representations of this type of dimensionless integral 
exist and converge quickly. For instance, one can show that
\begin{equation}
\int_{0}^{a} \frac{u^3}{e^u -1} \,{\rm d}u = \sum_{n=1}^{\infty} \frac{6}{n^4} 
- \frac{e^{-an}}{n^4} (6 + 6an + 3(an)^2 + (an)^3).
\nonumber
\end{equation}
A similar expression exists for the similar integral in the text. 
[Michels 1968] explains an analogous derivation.
}.

The importance of the functional form of $F^{\rm N}_{s,\gamma}$ is that it is 
to first order only a function of the blackbody temperature and the bandpass 
wavelength bounds.
Thus in the most general case $c_i(R_1, R_2, T_1, T_2, \lambda_{\rm min}, 
\lambda_{\rm max})$ and nothing else.
The only random variable involved in the flux being received at the detector is 
$r$, so we can indeed write the flux received at the detector as in 
Eq.~\ref{eq:flux_at_detector}.

We can finally write out the probability density functions for the signal to 
noise ratios in the single and double-star cases by using the transformation 
rule for pdfs, and applying Eq.~\ref{eq:snr_ivory_tower}.
For single stars,
\begin{equation}
{\rm prob}(x_s) = \frac{3}{d_{\rm max, s}^3} c_s^{3/2} \delta^3 \left( A T_{\rm 
dur} N_{\rm tra}\right)^{3/2} x_s^{-4}.
\label{eq:prob_xs}
\end{equation}
Analogously for double stars,
\begin{equation}
{\rm prob}(x_d) = \frac{3}{d_{\rm max, d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 
\left( A T_{\rm dur} N_{\rm tra}\right)^{3/2} x_d^{-4}.
\label{eq:prob_xd}
\end{equation}

\subsection{Number of detected planets}

Performing the integrals of Eq.~\ref{eq:completeness_short}, we get:

\begin{align}
N_{{\rm det},s} &= N_s \Gamma_{t,s} f_{s,{\rm S/N} > {\rm (S/N)_{min}}} \\
&= N_s \Gamma_{t,s} \frac{1}{d_{\rm max,s}^3} c_s^{3/2} \delta^3 \left(A 
T_{\rm dur} N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3},
\end{align}
and
\begin{align}
N_{{\rm det},d} &= 2 N_d \Gamma_{t,d} f_{d,{\rm S/N} > {\rm (S/N)_{min}}} \\
&= 2 N_d \Gamma_{t,d} 
\frac{1}{d_{\rm max,d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 \left(A T_{\rm dur} 
N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3}.
\end{align}

Formally, the $f_{s,{\rm S/N} > {\rm (S/N)_{min}}}$ term should really be 
written as ${\rm min}(1, \ldots)$, where $(\ldots)$ is the given expression. 
This ensures that the fraction of planets above the signal to noise threshold 
is less than 1.

The number of detected planets $N_{\rm det}$ is the sum of the two previously 
written equations, and can be written
\begin{align}
N_{\rm det} = 
&\left( \frac{\delta}{x_{\rm min}} \right)^3 (A T_{\rm dur} N_{\rm tra})^{3/2} 
\nonumber\\
&\quad \times
\left[ N_s \Gamma_{t,s} \frac{c_s^{3/2}}{d_{\rm max,s}^3}  +
	   2 N_d \Gamma_{t,d} \frac{c_d^{3/2}}{d_{\rm max,d}^3}  \right].
\label{eq:N_det}
\end{align}
All terms can be input to a computer, and checked against a Monte Carlo 
simulation if desired.

\section{Astronomer A ignores binarity}
Astronomer A has never heard of binary star systems. What occurrence rate does 
he derive for planets of radius $R_p$ and period $P$?

The {\it total} occurrence rate (number of detections divided number of 
``stars'') for Astronomer A would be $N_{\rm det}/(N_s + N_d)$. However, even 
though Astronomer A does not know about binaries, the radii he derives for any 
planets in binary systems are too small, by a factor $\sqrt{\mathcal{D}}$.
The question is what occurrence rate is derived for planets of radius $R_p$ and 
period $P$.
The answer is thus
\begin{equation}
\Gamma_{\rm A,\ planets\ of\ R_p} = \frac{N_{\rm det,s}}{N_s + N_d}.
\end{equation}

This astronomer will also think there is a second population of planets, with 
radius $R_p \sqrt{\mathcal{D}}$, and will thus rush to Nature claiming to also
have derived a second occurrence rate,
\begin{equation}
\Gamma_{\rm A,\ planets\ of\ R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm det,d}}{N_s 
+ N_d}.
\end{equation}

\section{Astronomer B counts host stars correctly}
Astronomer B can somehow account correctly for the ``2 for 1'' 
effect of binarity, \textit{i.e.} that the sample actually has more stars than 
astronomer A thought.

By the same token as above,
\begin{equation}
\Gamma_{\rm B,\ planets\ of\ R_p} = \frac{N_{\rm det,s}}{N_s + 2N_d},
\end{equation}
and
\begin{equation}
\Gamma_{\rm B,\ planets\ of\ R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm det,d}}{N_s 
	+ 2N_d}.
\end{equation}

\section{Astronomer C counts host stars correctly and figures out diluted radii}
Astronomer C did high resolution imaging followup on every candidate, and 
correctly classifies the planetary radii.
Thus, she also knows which planets are in binary systems, and which are in 
single star systems.

She knows that the purported population of planets with radii $R_p 
\sqrt{\mathcal{D}}$ does not exist. All detected planets from this survey have 
radii $R_p$. She computes an occurrence rate
\begin{align}
\Gamma_{\rm C,\ planets\ of\ R_p} &= 
\frac{N_{\rm det,s} + N_{\rm det,d}}{N_s + 2N_d} \\
&= \frac{N_{\rm det}}{N_s + 2N_d},
\end{align}

the closest yet to the true rate (Sec.~\ref{sec:true_rate}).

\section{Astronomer D counts host stars correctly, figures out diluted radii, 
and accounts for completeness}

Astronomer D, just like Astronomer C, knows which detections were around 
binaries and the associated radius correction.

Astronomer D notes that the number of detected planets is not equal to the 
actual number of planets in the sample.
They must account for completeness.
They do injection recovery, and derive independent estimates for their 
completeness functions about single stars, $f_{s,x>x_{\rm min}} \equiv f_s$ and 
about double star systems, $f_{d,x>x_{\rm min}} \equiv f_d$.
With this knowledge in hand, they compute the respective single and binary 
occurrence rates
\begin{equation}
\Gamma_{t,s} = \frac{N_{\rm det,s}}{N_s f_s},
\end{equation}
\begin{equation}
\Gamma_{t,d} = \frac{N_{\rm det,d}}{2 N_d f_d}.
\end{equation}

With these in hand, they derive the overall occurrence rate
\begin{align}
\Gamma_{\rm D,\ planets\ of\ R_p} 
&= \frac{N_{\rm det,s}/f_s + N_{\rm det,d}/f_d  }{N_s + 2 N_d} \\
&= \frac{\Gamma_{t,s} N_s + 2\Gamma_{t,d} N_d}{N_s + 2 N_d}.
\end{align}

Although we admit it's been a bit of a slog, it happens that Astronomer D's 
occurrence rate is the true occurrence rate (cf. Eq.~\ref{eq:true_occ}).

All it takes is $\approx$ a full semester at Keck, a system-by-system analysis, 
and perfect understanding of the completeness of the detection efficiency for 
single and double star systems.

\section{Numerical verification}

As a check on the preceding analytic development, we implemented a Monte Carlo 
simulation of this idealized transit survey.
To run the survey, we defined the instrument specifications (detector area and 
transmission function), the stellar population (binary fraction, total number 
density of a given stellar class, binary light ratio, fixed stellar 
properties), the planet population (fixed planet radius, period, and occurrence 
rate about single and binary stars), and finally the survey parameters 
(observing baseline, minimum SNR for ``detection'').
We then randomly drew star positions, randomly assigned planets to stars in 
single and binary systems, and computed the resulting signal to noise 
(Eq.~\ref{eq:snr_ivory_tower}) with which 
the transits would be observed.
As in the preceding analytics\footnote{TODO could be to generalize away from 
this}, we assumed ``twin'' binaries (same stellar radii, same effective 
temperature, and dilution does not depend on which stellar binary is the 
``target'').

The results are shown in Fig.~\ref{fig:snr_dist_analytic_v_numeric}, and 
indicate that the analytic probability distribution functions 
Eqs.~\ref{eq:prob_xs},~\ref{eq:prob_xd}, and the number of 
detections (Eq.~\ref{eq:N_det}) are correct.

A point evident in Fig.~\ref{fig:snr_dist_analytic_v_numeric} is that, for 
fixed planet parameters, and fixed stellar parameters ($R_\star, L_\star$, and 
distance $r$) the SNR distribution for planets in binaries is poorer than that 
of planets in single star systems.
We can see analytically that this can be written only as function of the 
binary light ratio:
\begin{equation}
\frac{{\rm prob}(x_d)}{{\rm prob}(x_s)} =(1 + \gamma_R)^{-1}.
\end{equation}
Deriving this simple form requires noting that the ratios of the 
bandpass-specific number luminosities (Eq.~\ref{eq:c_i}) is equal to the ratio 
of the bandpass-specific energy luminosities (otherwise a term with $c_s/c_d$ 
must be included).

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=0.4]{figures/snr_distribution.pdf}
	\end{center}
	\caption{Comparison of analytic and numeric probability 
	density functions of the SNR in an idealized transit survey.
	The analytic lines are Eqs.~\ref{eq:prob_xs} and~\ref{eq:prob_xd} for the 
	planet populations orbiting single and binary stars. The underlying stepped 
	histogram is output from Monte Carlo simulations. Poisson noise leads to 
	a small deviation at the faint and bright limits, but the numerics and 
	analytics otherwise agree.
		 }
	\label{fig:snr_dist_analytic_v_numeric}
\end{figure}

\section{Representative numbers for a few cases}

\subsection{Twin binaries: if we ignore binarity, for what fraction of 
detections do we misclassify the radii?}

Ignoring binarity, we will detect $N_{\rm det,s}$ planets around single stars, 
and $N_{\rm det,d}$ planets around double stars. The latter set will be assumed 
to have radii $R_p \sqrt{\mathcal{D}}$.
The fraction of detections with misclassified radii can then be written
\begin{equation}
\frac{N_{\rm det,d}}{N_{\rm det,s} + N_{\rm det,d}} = \frac{1}{1+\alpha},
\end{equation}
for
\begin{equation}
\alpha \equiv \frac{1}{2({\rm BF})} (1+\gamma_R)^{3/2} 
\frac{\Gamma_{t,s}}{\Gamma_{t,d}}.
\end{equation}

For the nominal G2V dwarf case of ${\rm BF}=0.45$, twin binaries with equal 
occurrence rates this produces a misclassification rate of $24\%$, in agreement 
with Fig.~\ref{fig:snr_dist_analytic_v_numeric}.

\subsection{Twin binaries: if we ignore binarity, how wrong is our occurrence 
rate for planets of radius $R_p$?}

This is almost simply asking ``what is the relative difference between the 
occurrence rates derived by Astronomers D and A for planets of radius $R_p$?''
However, in the more realistic case, Astronomer A also has derived a 
completeness, which we assume is the same as for Astronomer D in the single 
star case.
So Astronomer A now misclassifies planetary radii, and miscounts the total 
number of stars, but knows his completeness for single stars.
Astronomer D corrects all these errors.

For brevity, write $\Gamma_{\rm A,\ planets\ of\ R_p} = \Gamma_{\rm A,R_p}$, 
and similarly for ${\rm D}$.
Then the relative difference between the two occurrence rates is

\begin{align}
\left|\frac{\Gamma_{\rm A,R_p} - \Gamma_{\rm D,R_p}}
		   {\Gamma_{\rm A,R_p}} \right| &=
\left| 1 - \frac{\Gamma_{\rm D,R_p}}{\Gamma_{\rm A,R_p}}\right| \\
&=
\left| 1 - \left(\frac{\Gamma_{t,s}N_s + 2\Gamma_{t,d}N_d}{N_s + 2N_d}  \cdot 
\frac{N_s + N_d}{\Gamma_{t,s} N_s }\right)\right| \\
&= 
\left|
1 - \frac{(1 + 2 \beta \Gamma_{t,d}/\Gamma_{t,s})(1 + \beta)}{(1+2\beta)}
\right|,
\end{align}
for
\begin{equation}
\beta \equiv N_d/N_s = {\rm BF} \times (1+\gamma_R)^{3/2}.
\end{equation}

For the nominal G2V dwarf case of ${\rm BF}=0.45$ with twin binaries ($\gamma_R 
= 1$) and $\Gamma_{t,d} = \Gamma_{t,s}$ this gives a relative error of $127\%$.
For instance, in the numerical simulation corresponding to 
Fig.~\ref{fig:snr_dist_analytic_v_numeric}, Astronomer A finds $\Gamma_{\rm 
A,R_p} = 0.22$, while Astronomer 
D derives the true (input) occurrence rate of $\Gamma_{\rm D,R_p} = 0.5$.

\newpage

%\begin{thebibliography}{}

%\end{thebibliography}

\end{document}

### 2017-08-30 memos/toy_analytic_survey_fixed_lightratio.tex
initial commit

--- 2017-08-30 memos/toy_analytic_survey_20170728.tex
@@
 
-\title{ A Toy Analytic Transit Survey }
+\title{ A Toy Analytic Transit Survey: Fixed Light Ratio Binaries }
 
@@
 
-We derive the signal to noise distribution of detected planets in a toy transit 
-survey. We give corresponding expressions for the number of detected planets. 
-We then discuss the effect of various errors that one makes by ignoring 
-binarity when deriving occurrence rates.
+We discuss various errors one can make by ignoring binarity when deriving 
+occurrence rates.
+In this memo, we assume the stars in any binary system have fixed light ratio.
 
@@
 
-Imagine the following transit survey:
+Imagine the following survey, similar to that outlined by [Pepper et al, 2003]:
+\begin{enumerate}
+\item You are going to observe the entire sky for a duration $T_{\rm obs}$, 
+with a detector of area $A$, and known bandpass. Your detector is photon-noise 
+limited.
+%
+\item You are interested only in detecting planets of radius $R_p$, and orbital 
+period $P$. For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
+%
+\item You only want to detect them around stars of radius $R_1$, and luminosity 
+$L_1$. For instance, G2V dwarfs.
+%
+\item Since your detections will be S/N limited, you want only to observe stars 
+for which your target can be detected with ${\rm S/N} > {\rm (S/N)_{min}}$.
+%
+\item For a photon-noise limited survey, the signal to noise limit is 
+equivalent to a magnitude (flux) limit. So you select all the points on the sky 
+above a flux limit, \textit{i.e.} with flux $F_{\rm pt} > F_{\rm min}$, for 
+$F_{\rm min}$ defined by your target planet and stellar types, and your survey 
+specifications.
+%
+\item You carry out a transit survey, and detect many transiting planets.
+\end{enumerate}
+
+You now wish to derive an occurrence rate for planets of radius $R_p$ and 
+orbital period $P$.
+Assume your universe is a universe in which:
 \begin{itemize}
-\item Take a magnitude-limited sample of ``stars'' of a single spectral class. 
-For instance, G2V dwarfs.  In other words, choose all ``points'' on-sky with 
-$m_\lambda < m_{\lambda,{\rm lim}}$, or equivalently with $F_\lambda > 
-F_{\lambda,{\rm lim}}$, and also with colors that make you think they are this 
-spectral class.
-\item The true population of ``points'' (hereafter, systems) contains both 
-single and double star systems. Single star systems have luminosity in the 
-observed bandpass $L_1$, radii $R_1$, and effective temperature $T_{\rm eff,1}$.
+\item The true population of ``points'' (stellar systems, all unresolved) 
+	comprises both single and double star systems. Single star systems have 
+	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
+	temperature $T_{\rm eff,1}$.
 	Double star systems have luminosity in the observed bandpass $L_d = 
-(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of the 
-secondary to the primary. The ratio of the two number densities in a 
-volume-limited sample is the binary fraction\footnote{The binary fraction is 
-equivalent to the multiplicity fraction if there are no triple, quadruple, 
-$\ldots$ systems.}.
+	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
+	the secondary to the primary. 
+  In this memo, $\gamma_R$ is assumed to be a constant across the population of
+  star systems.
+  The ratio of the two number densities in a 
+	volume-limited sample is the binary fraction\footnote{The binary fraction 
+	is equivalent to the multiplicity fraction if there are no triple, 
+	quadruple, $\ldots$ systems.}.
 \item The true population of planets around these stars is as follows:
@@
     Any astrophysical difference in planet formation between singles and 
-    binaries is captured by these two terms.
+    binaries is captured by these two terms.\footnote{We'll see that it's 
+    an extreme limit to assume the primary and secondary of binary systems host 
+    planets at equal rates if the secondary is non-identical to the primary.}
 \end{itemize}
-
-A signal to noise limited transit survey is then conducted. All planets with 
-${\rm S/N} > {\rm (S/N)_{min}}$ are detected. The rest are not.
-Note that while the S/N threshold does imply a minimum stellar flux about which 
-planets of a given radius and period can be detected, this minimum flux is 
-generally different from the minimum used to define the original magnitude cut 
-for a sample [see Batalha et al, 2010].
 
@@
 \item What about astronomer C, who accounts for ``2 for 1'' \textit{and} 
-misclassification due to diluted radii? In other words, astronomer C has lots 
-of Keck time, and did high resolution imaging followup on every candidate, and 
-correctly classifies the planetary radii.
-
-\item What about astronomer D, who additionally notes the importance of 
-completeness?
+misclassification due to diluted radii? In other words, astronomer C did a 
+combination of high resolution imaging and RV followup on every candidate, and
+correctly classifies the planetary radii in every case.
 \end{enumerate}
@@
 magnitude cut.
-The ratio of double to single systems is
+The ratio of double to single systems, assuming $\gamma_R$ is a constant for 
+the entire population, is
 \begin{align}
@@
 
-\section{How many planets are in the sample?}
+\subsection{Actually finding $F_{\rm lim}$}
 
-The number of planets in the sample is
-\begin{align}
-N_{\rm planets} =& N_{\rm planets\ in\ single\ star\ systems}  +  \\
-				  &\quad\quad N_{\rm planets\ in\ double\ star\ systems} 
-				  \nonumber \\
-			   =& \Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d.
-\end{align}
-
-The factor of 2 accounts for the fact that there are twice as many stars in 
-double star systems.
-
-\section{What is the true occurrence rate?}
-\label{sec:true_rate}
-
-The ``true occurrence rate'' is the average number of planets per star. Thus
-
-\begin{align}
-\Gamma_t &= \frac{N_{\rm planets}}{N_{\rm stars}} \\
-\Gamma_t &= \frac{\Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d}{N_s + 2N_d}.
-\label{eq:true_occ}
-\end{align}
-
-\section{How many planets are detected?}
-The total number of planet detections is the sum of the number of planets 
-detected in single star systems $N_{{\rm det},s}$ and the number of planets 
-detected in double star systems $N_{{\rm det},d}$.
-These can be expressed individually. 
-The former is
-\begin{equation}
-N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,{\rm S/N} > {\rm (S/N)_{min}}},
-\label{eq:N_det_s}
-\end{equation}
-where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
-star systems of the sample, and $f_{s,({\rm S/N} > {\rm (S/N)_{min}})}$ is the 
-fraction of these planets that have signal to noise greater than the minimum 
-detection threshold.
-\begin{equation}
-N_{{\rm det},d} = 2 N_d \Gamma_{t,d} f_{d,{\rm S/N} > {\rm (S/N)_{min}}},
-\label{eq:N_det_d}
-\end{equation}
-where now $2 N_d \Gamma_{t,d}$ is the number of planets in the double star 
-systems of the sample, and the completeness term must account for any 
-differences in the signal to noise distribution that come from stellar 
-binarity.
-
-As a note of warning, Eqs.~\ref{eq:N_det_s} and~\ref{eq:N_det_d} ignore the 
-geometric transit probability, $\approx R_\star / a$. This means we are 
-assuming that every existing planet in this universe transits. This assumption 
-can easily be rectified, if desired.
-
-\subsection{Analytic completeness}
-The only terms we have yet to compute are the completeness terms, $f_{i,{\rm 
-S/N} > {\rm (S/N)_{min}}}$ for $i\in{\rm \{single,double\} }$. 
-We proceed as follows.
+To express $F_{\rm lim}$ in terms of the survey's defining parameters, we need 
+to convert between ${\rm (S/N)_{lim}}$ and $F_{\rm lim}$. We proceed as follows.
 
@@
 
-The interesting term in Eq.~\ref{eq:snr_ivory_tower} that changes between stars 
-of the same binarity class in our idealized sample is the square root of 
-$F^{\rm N}_\gamma$.
-This is the term that leads to a distribution of signal to noises for different 
-stars.
-The completenesses $f_{i,{\rm S/N} > {\rm (S/N)_{min}}}$ can be directly 
-expressed in terms of those probability density functions:
+Using Eq.~\ref{eq:snr_ivory_tower}, we can write the minimum number flux of 
+photons required for a detection at threshold as
 \begin{equation}
-f_{i,{\rm S/N} > {\rm (S/N)_{min}}} = 
-	\int_{{\rm (S/N)_{min}}}^{\infty} 
-		{\rm d}\left(\frac{{\rm S}}{{\rm N}}\right)_i \ 
-		{\rm prob}\left(\frac{{\rm S}}{{\rm N}}\right)_i.
-\label{eq:completeness_long}
+F_{\rm lim}^N = \left[ \left({\rm \frac{S}{N}}\right)_{\rm min} \frac{1}{\delta 
+\mathcal{D}} \right]^2 \frac{1}{A N_{\rm tra} T_{\rm dur}}.
 \end{equation}
-We keep the subscript $i$ because the signal to noise distributions are 
-different for the cases of single star systems ($i=s$) and double star systems 
-($i=d$).
-Notably:
-\begin{itemize}
-	\item The dilution differs (Eq.~\ref{eq:dilution}).
-	\item The photon number flux from the system differs.
-\end{itemize}
+To convert this to $F_{\rm lim}$, multiply by the average photon energy in the 
+bandpass.
 
-To simplify notation, we let $x_i \equiv {\rm (S/N)}_i$, and rewrite 
-Eq.~\ref{eq:completeness_long} as
-\begin{equation}
-f_{i,x > x_{\rm min}} = 
-\int_{{\rm x_{\rm min}}}^{\infty} 
-{\rm d}x_i \ 
-{\rm prob}(x_i).
-\label{eq:completeness_short}
-\end{equation}
+\section{How many planets are in the sample?}
 
-\subsection{Deriving ${\rm prob}(x_i)$}
-We want expressions for the probability density function of the observed signal 
-to noise ratio, ${\rm prob}(x_i)$, for both the single and binary 
-system case.
+The number of planets in the sample is
+\begin{align}
+N_{\rm planets} =& N_{\rm planets\ in\ single\ star\ systems}  +  \\
+				  &\quad\quad N_{\rm planets\ in\ double\ star\ systems} 
+				  \nonumber \\
+			   =& \Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d.
+\end{align}
 
-First, note that a star placed uniformly in the volume of the search space will 
-have a probability density function for its distance $r$ from the origin of
-\begin{equation}
-{\rm prob} (r) = \frac{3 r^2}{d_{\rm max}^3},
-\end{equation}
-where the appropriate maximum distances should be substituted per 
-Eq.~\ref{eq:d_max}.
-Noting the transformation rule for probability density functions, we can 
-evaluate the probability of a star having a observed number flux $F^{\rm 
-N}_{i,\gamma}$ in the bandpass,
-\begin{align}
-{\rm prob} (F^{\rm N}_{i,\gamma}) 
-&= {\rm prob}(r(F^{\rm N}_{i,\gamma}))
-	\left| \frac{{\rm d} r}{{\rm d} F^{\rm N}_{i,\gamma}} \right| \\
-&= \frac{3}{2 d_{\rm max}^3} c_i^{3/2} (F^{\rm N}_{i,\gamma})^{-5/2},
-\label{eq:pdf_observed_flux}
-\end{align}
-where in the latter equality we have written a ``number luminosity'' $c_i$ 
-(units of inverse time) defined for $i\in{\rm \{single,double\} }$ as
-\begin{equation}
-c_i = 
-\Bigg\{\begin{array}{lr}
-R_1^2 F^{\rm N}_{s1,\gamma}, & {\rm\ if\ single}\\
-R_1^2 F^{\rm N}_{s1,\gamma} + R_2^2 F^{\rm N}_{s2,\gamma} & {\rm\ if\ double}.
-\end{array}
-\label{eq:c_i}
-\end{equation}
-In Eq.~\ref{eq:c_i}, $F^{\rm N}_{s1,\gamma}$ and $F^{\rm N}_{s2,\gamma}$ are 
-the photon number fluxes at the surfaces of the stars. To derive 
-Eq.~\ref{eq:pdf_observed_flux}, we simply scaled these by the distance:
-\begin{equation}
-F^{\rm N}_{i,\gamma} = \frac{c_i}{r^2}.
-\label{eq:flux_at_detector}
-\end{equation}
+The factor of 2 accounts for the fact that there are twice as many stars in 
+double star systems.
 
-The surface photon number fluxes $F^{\rm N}_{si,\gamma}$ in Eq.~\ref{eq:c_i} 
-are usually evaluated numerically, by convolving the wavelength-specific photon 
-flux density of a star with the dimensionless spectral response function of the 
-instrument. In other words,
-\begin{equation}
-F^{\rm N}_{s,\gamma} = \int F_\lambda T_\lambda \, {\rm d}\lambda.
-\label{eq:pfd}
-\end{equation}
-The wavelength-specific photon flux density $F_\lambda$ [${\rm 
-ph\,cm^{-2}\,s^{-1}\,\AA^{-1}}$] could be from Pickles' library, or could be 
-a blackbody function. The transmission function is, up to factor of order 
-unity, a step function between two wavelengths $\lambda_{\rm min}$ and 
-$\lambda_{\rm max}$.
-If we assume a blackbody source, Eq.~\ref{eq:pfd} becomes
-\begin{equation}
-F^{\rm N}_{s,\gamma} = 8\pi c \left(\frac{k T}{h c}\right)^3 
-\int_{hc/(\lambda_{\rm max} kT)}^{hc/(\lambda_{\rm min} kT)}
-\frac{u^2}{e^u - 1} \,{\rm d} u,
-\end{equation}
-which can be evaluated numerically\footnote{It may help in the numerics to note 
-that infinite series representations of this type of dimensionless integral 
-exist and converge quickly. For instance, one can show that
-\begin{equation}
-\int_{0}^{a} \frac{u^3}{e^u -1} \,{\rm d}u = \sum_{n=1}^{\infty} \frac{6}{n^4} 
-- \frac{e^{-an}}{n^4} (6 + 6an + 3(an)^2 + (an)^3).
-\nonumber
-\end{equation}
-A similar expression exists for the similar integral in the text. 
-[Michels 1968] explains an analogous derivation.
-}.
+\section{What is the true occurrence rate?}
+\label{sec:true_rate}
 
-The importance of the functional form of $F^{\rm N}_{s,\gamma}$ is that it is 
-to first order only a function of the blackbody temperature and the bandpass 
-wavelength bounds.
-Thus in the most general case $c_i(R_1, R_2, T_1, T_2, \lambda_{\rm min}, 
-\lambda_{\rm max})$ and nothing else.
-The only random variable involved in the flux being received at the detector is 
-$r$, so we can indeed write the flux received at the detector as in 
-Eq.~\ref{eq:flux_at_detector}.
-
-We can finally write out the probability density functions for the signal to 
-noise ratios in the single and double-star cases by using the transformation 
-rule for pdfs, and applying Eq.~\ref{eq:snr_ivory_tower}.
-For single stars,
-\begin{equation}
-{\rm prob}(x_s) = \frac{3}{d_{\rm max, s}^3} c_s^{3/2} \delta^3 \left( A T_{\rm 
-dur} N_{\rm tra}\right)^{3/2} x_s^{-4}.
-\label{eq:prob_xs}
-\end{equation}
-Analogously for double stars,
-\begin{equation}
-{\rm prob}(x_d) = \frac{3}{d_{\rm max, d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 
-\left( A T_{\rm dur} N_{\rm tra}\right)^{3/2} x_d^{-4}.
-\label{eq:prob_xd}
-\end{equation}
-
-\subsection{Number of detected planets}
-
-Performing the integrals of Eq.~\ref{eq:completeness_short}, we get:
+The ``true occurrence rate'' is the average number of planets per star. Thus
 
 \begin{align}
-N_{{\rm det},s} &= N_s \Gamma_{t,s} f_{s,{\rm S/N} > {\rm (S/N)_{min}}} \\
-&= N_s \Gamma_{t,s} \frac{1}{d_{\rm max,s}^3} c_s^{3/2} \delta^3 \left(A 
-T_{\rm dur} N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3},
-\end{align}
-and
-\begin{align}
-N_{{\rm det},d} &= 2 N_d \Gamma_{t,d} f_{d,{\rm S/N} > {\rm (S/N)_{min}}} \\
-&= 2 N_d \Gamma_{t,d} 
-\frac{1}{d_{\rm max,d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 \left(A T_{\rm dur} 
-N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3}.
+\Gamma_t &= \frac{N_{\rm planets}}{N_{\rm stars}} \\
+\Gamma_t &= \frac{\Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d}{N_s + 2N_d}.
+\label{eq:true_occ}
 \end{align}
 
-Formally, the $f_{s,{\rm S/N} > {\rm (S/N)_{min}}}$ term should really be 
-written as ${\rm min}(1, \ldots)$, where $(\ldots)$ is the given expression. 
-This ensures that the fraction of planets above the signal to noise threshold 
-is less than 1.
+\section{How many planets are detected?}
+The total number of planet detections is the sum of the number of planets 
+detected in single star systems $N_{{\rm det},s}$ and the number of planets 
+detected in double star systems $N_{{\rm det},d}$.
+These can be expressed individually. 
+Since we selected stars for which there was enough light to make detections, 
+there is no need to compute the ${\rm S/N}$ distribution of 
+``threshold-crossing events''.
 
+The number of planets detected in single star systems is
+\begin{equation}
+N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,{\rm geom}},
+\label{eq:N_det_s}
+\end{equation}
+where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
+star systems of the sample, and $f_{s,{\rm geom}}$ is the geometric transit 
+probability.
+The number of planets detected in double star systems is
+\begin{equation}
+N_{{\rm det},d} = 2 N_d \Gamma_{t,d} f_{d,{\rm geom}},
+\label{eq:N_det_d}
+\end{equation}
+where now $2 N_d \Gamma_{t,d}$ is the number of planets in the double star 
+systems of the sample.
 The number of detected planets $N_{\rm det}$ is the sum of the two previously 
-written equations, and can be written
-\begin{align}
-N_{\rm det} = 
-&\left( \frac{\delta}{x_{\rm min}} \right)^3 (A T_{\rm dur} N_{\rm tra})^{3/2} 
-\nonumber\\
-&\quad \times
-\left[ N_s \Gamma_{t,s} \frac{c_s^{3/2}}{d_{\rm max,s}^3}  +
-	   2 N_d \Gamma_{t,d} \frac{c_d^{3/2}}{d_{\rm max,d}^3}  \right].
-\label{eq:N_det}
-\end{align}
-All terms can be input to a computer, and checked against a Monte Carlo 
-simulation if desired.
+written equations.
 
@@
 
-The {\it total} occurrence rate (number of detections divided number of 
-``stars'') for Astronomer A would be $N_{\rm det}/(N_s + N_d)$. However, even 
-though Astronomer A does not know about binaries, the radii he derives for any 
-planets in binary systems are too small, by a factor $\sqrt{\mathcal{D}}$.
-The question is what occurrence rate is derived for planets of radius $R_p$ and 
-period $P$.
-The answer is thus
+The {\it total} occurrence rate (number of planets divided number of 
+``stars'') for Astronomer A would be $(N_{\rm det}/f_{s,{\rm geom}})/(N_s + 
+N_d)$. However, even though Astronomer A does not know about binaries, the 
+radii he derives for any planets in binary systems are too small, by a factor 
+$\sqrt{\mathcal{D}}$. The question asks what occurrence rate is derived for 
+planets of radius $R_p$ and period $P$.
+The answer is
+\begin{align}
+\Gamma_{\rm A,\ planets\ of\ R_p} &= \frac{N_{\rm det,s}/f_{s,{\rm geom}}}{N_s 
++ N_d}
+\label{eq:A_0} \\
+&= \frac{\Gamma_{t,s} N_s}{N_s + N_d}
+\end{align}
+
+This astronomer will also think there is a second population of planets, with 
+radius $R_p \sqrt{\mathcal{D}}$ (a constant number for $\gamma_R = {\rm 
+constant}$). He will then also claim
+have derived a second occurrence rate,
 \begin{equation}
-\Gamma_{\rm A,\ planets\ of\ R_p} = \frac{N_{\rm det,s}}{N_s + N_d}.
+\Gamma_{\rm A,\ planets\ of\ R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm 
+det,d}/f_{d,{\rm geom}}}{N_s + N_d},
 \end{equation}
 
-This astronomer will also think there is a second population of planets, with 
-radius $R_p \sqrt{\mathcal{D}}$, and will thus rush to Nature claiming to also
-have derived a second occurrence rate,
-\begin{equation}
-\Gamma_{\rm A,\ planets\ of\ R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm det,d}}{N_s 
-+ N_d}.
-\end{equation}
+where at least for the twin binary case the geometric completeness term is the 
+same as for Eq.~\ref{eq:A_0}.
 
@@
 \begin{equation}
-\Gamma_{\rm B,\ planets\ of\ R_p} = \frac{N_{\rm det,s}}{N_s + 2N_d},
+\Gamma_{\rm B,\ planets\ of\ R_p} = \frac{N_{\rm det,s}/f_{s,{\rm geom}}}{N_s + 
+2N_d},
 \end{equation}
@@
 \begin{equation}
-\Gamma_{\rm B,\ planets\ of\ R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm det,d}}{N_s 
+\Gamma_{\rm B,\ planets\ of\ R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm 
+det,d}/f_{d,{\rm geom}}}{N_s 
 	+ 2N_d}.
@@
 single star systems.
+Since there are no completeness corrections, she has everything she needs.
 
@@
 \Gamma_{\rm C,\ planets\ of\ R_p} &= 
-\frac{N_{\rm det,s} + N_{\rm det,d}}{N_s + 2N_d} \\
-&= \frac{N_{\rm det}}{N_s + 2N_d},
-\end{align}
-
-the closest yet to the true rate (Sec.~\ref{sec:true_rate}).
-
-\section{Astronomer D counts host stars correctly, figures out diluted radii, 
-and accounts for completeness}
-
-Astronomer D, just like Astronomer C, knows which detections were around 
-binaries and the associated radius correction.
-
-Astronomer D notes that the number of detected planets is not equal to the 
-actual number of planets in the sample.
-They must account for completeness.
-They do injection recovery, and derive independent estimates for their 
-completeness functions about single stars, $f_{s,x>x_{\rm min}} \equiv f_s$ and 
-about double star systems, $f_{d,x>x_{\rm min}} \equiv f_d$.
-With this knowledge in hand, they compute the respective single and binary 
-occurrence rates
-\begin{equation}
-\Gamma_{t,s} = \frac{N_{\rm det,s}}{N_s f_s},
-\end{equation}
-\begin{equation}
-\Gamma_{t,d} = \frac{N_{\rm det,d}}{2 N_d f_d}.
-\end{equation}
-
-With these in hand, they derive the overall occurrence rate
-\begin{align}
-\Gamma_{\rm D,\ planets\ of\ R_p} 
-&= \frac{N_{\rm det,s}/f_s + N_{\rm det,d}/f_d  }{N_s + 2 N_d} \\
+\frac{N_{\rm det,s}/f_{s,{\rm geom}} + N_{\rm det,d}/f_{d,{\rm geom}}}{N_s + 
+2N_d} \\
 &= \frac{\Gamma_{t,s} N_s + 2\Gamma_{t,d} N_d}{N_s + 2 N_d}.
@@
 
-Although we admit it's been a bit of a slog, it happens that Astronomer D's 
-occurrence rate is the true occurrence rate (cf. Eq.~\ref{eq:true_occ}).
-
-All it takes is $\approx$ a full semester at Keck, a system-by-system analysis, 
+With $\approx$ a full semester at Keck, a system-by-system analysis, 
 and perfect understanding of the completeness of the detection efficiency for 
-single and double star systems.
+single and double star systems, Astronomer C has found the true occurrence rate 
+(cf. Eq.~\ref{eq:true_occ}).
 
@@
 
-As a check on the preceding analytic development, we implemented a Monte Carlo 
+As a check on the preceding analytic discussion, we implemented a Monte Carlo 
 simulation of this idealized transit survey.
@@
 the transits would be observed.
-As in the preceding analytics\footnote{TODO could be to generalize away from 
-this}, we assumed ``twin'' binaries (same stellar radii, same effective 
-temperature, and dilution does not depend on which stellar binary is the 
-``target'').
+As in the preceding analytics, we assumed ``twin'' binaries (same stellar 
+radii, same effective temperature, and dilution does not depend on which 
+stellar binary is the ``target'').
 
 The results are shown in Fig.~\ref{fig:snr_dist_analytic_v_numeric}, and 
-indicate that the analytic probability distribution functions 
-Eqs.~\ref{eq:prob_xs},~\ref{eq:prob_xd}, and the number of 
-detections (Eq.~\ref{eq:N_det}) are correct.
+indicate that the analytic probability distribution functions derived in an 
+earlier version of this memo are correct.
 
@@
 fixed planet parameters, and fixed stellar parameters ($R_\star, L_\star$, and 
-distance $r$) the SNR distribution for planets in binaries is poorer than that 
-of planets in single star systems.
+distance $r$), and a constant $\gamma_R$ across the population, the SNR 
+distribution for planets in binaries is poorer than that of planets in single 
+star systems.
 We can see analytically that this can be written only as function of the 
@@
 Deriving this simple form requires noting that the ratios of the 
-bandpass-specific number luminosities (Eq.~\ref{eq:c_i}) is equal to the ratio 
+bandpass-specific number luminosities is equal to the ratio 
 of the bandpass-specific energy luminosities (otherwise a term with $c_s/c_d$ 
@@
 	density functions of the SNR in an idealized transit survey.
-	The analytic lines are Eqs.~\ref{eq:prob_xs} and~\ref{eq:prob_xd} for the 
-	planet populations orbiting single and binary stars. The underlying stepped 
-	histogram is output from Monte Carlo simulations. Poisson noise leads to 
-	a small deviation at the faint and bright limits, but the numerics and 
-	analytics otherwise agree.
+	The analytic lines are from a previous version of this memo 
+	(\texttt{20170728}) for the planet populations orbiting single and binary 
+	stars. The underlying stepped histogram is output from Monte Carlo 
+	simulations. Poisson noise leads to a small deviation at the faint and 
+	bright limits, but the numerics and analytics otherwise agree.
 		 }
@@
 D derives the true (input) occurrence rate of $\Gamma_{\rm D,R_p} = 0.5$.
+For the more ``realistic'' case of $\gamma_R = 0.1$, $\Gamma_{t,d} = 
+\Gamma_{t,s}$, the relative error is $52\%$. Not as bad, and this latter number 
+should be looked at skeptically because the ``equal occurrence rates for 
+primary and secondary'' assumption may be unlikely with this light ratio as a 
+median.
 

### 2017-08-30 memos/toy_analytic_survey_vary_lightratio.tex
initial commit

%\documentclass[12pt,preprint]{aastex}
%\documentclass[preprint2,12pt]{aastex}
\documentclass{emulateapj}
%\usepackage{apjfonts}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{comment}

\shortauthors{Name et al.}
\shorttitle{Best transiting planets}

\begin{document}

% ------------------------------------------------------------------------
% New commands
%
\def\ltsima{$\; \buildrel < \over \sim \;$}
\def\lsim{\lower.5ex\hbox{\ltsima}}
\def\gtsima{$\; \buildrel > \over \sim \;$}
\def\gsim{\lower.5ex\hbox{\gtsima}}
\def\tess{{\it TESS} }
\def \teff {T_{\rm eff}}
\def \phir {\Phi_{\rm R}}
\def \fov {24$^{\circ}$}
\def \pixsz {21.1''}
\def \aeff {69.1 cm$^2$ }    
\def \epd {105 mm}

% -------------------------------------------------------------------------
%

\bibliographystyle{apj}

\title{ A Toy Analytic Transit Survey: Varying the Light Ratio of Binaries }

\author{
  LGB, KM, JNW
}

% \journalinfo{Draft version}
\slugcomment{Memo for internal use}

%\altaffiltext{1}{Princeton University}

\begin{abstract}

In this memo, we assume the stars in the population of binary systems have 
varying light ratios.

\end{abstract}

\keywords{planets and satellites:\ detection}

\section*{Statement of problem}

Ah, transiting planets! We learn so much by studying them. But how much can we 
actually learn, and how much is messed up by binarity?

Imagine the following survey, similar to that outlined by [Pepper et al, 2003]:
\begin{itemize}
\item You are going to observe the entire sky for a duration $T_{\rm obs}$, 
with a detector of area $A$, and known bandpass. Your detector is photon-noise 
limited.
%
\item You are interested only in detecting planets of radius $R_p$, and orbital 
period $P$. For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
%
\item You only want to detect them around stars of radius $R_1$, and luminosity 
$L_1$. For instance, G2V dwarfs.
%
\item Since your detections will be S/N limited, you want only to observe stars 
for which your target can be detected with ${\rm S/N} > {\rm (S/N)_{min}}$.
%
\item For a photon-noise limited survey, the signal to noise limit is 
equivalent to a magnitude (flux) limit. So you select all the points on the sky 
above a flux limit, \textit{i.e.} with flux $F_{\rm pt} > F_{\rm min}$, for 
$F_{\rm min}$ defined by your target planet and stellar types, and your survey 
specifications.
%
\item You carry out a transit survey, and detect many transiting planets.
\end{itemize}

You now wish to derive an occurrence rate for planets of radius $R_p$ and 
orbital period $P$.
Assume your universe is a universe in which
the true population of ``points'' (stellar systems, all unresolved) 
comprises single and double star systems. 
Single star systems have 
luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
temperature $T_{\rm eff,1}$.
Double star systems have luminosity in the observed bandpass $L_d = 
(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
the secondary to the primary.

In this memo $\gamma_R$ varies across the population of star systems.
It does so because the underlying mass ratio varies, and $\gamma_R = q^\alpha$. 
Since we are interested in solar type binaries, we take the distribution of the 
mass ratio $q=M_2/M_1$ by coarsely approximating [Rhagavan et al 2010, Fig 16]:
\begin{equation}
{\rm prob}(q) =
\Bigg\{\begin{array}{lr}
c_q & 0.1 < q \leq 1  \\
0 & {\rm otherwise},
\end{array}
\label{eq:mass_ratio}
\end{equation}
for $c_q = 1/9$ to normalize the distribution to 1.
The mass-luminosity relation is that given by [Salaris \& Cassisi (2005), 
Sec 5.7],
\begin{equation}
L \propto
\Bigg\{\begin{array}{lr}
M^{4.5} & 0.5 < M_\star/M_\odot \leq 2.0  \\
M^{2.6} & 0.2 < M_\star/M_\odot \leq 0.5.
\end{array}
\label{eq:mass_luminosity}
\end{equation}
For $M_1 = M_\odot$, this gives $\alpha = 4.5$ for $0.5 < q \leq 1$, and 
otherwise $\alpha = 2.6$. 
For analytic convenience this distribution may also be approximated as 
a constant, \textit{e.g.,} $\alpha = 3$, in the subsequent development.

The true population of planets around these stars is then as follows:
a fraction $\Gamma_{t,s}$ of stars in single star systems 
	have a planet of radius $R_p$, with orbital period $P$.
A fraction $\Gamma_{t,d}$ of each star in a double star 
	system has a planet of radius $R_p$, with orbital period $P$. For instance, 
	if $\Gamma_{t,s} = \Gamma_{t,d} = 0.1$, on average each double 
	system contributes 0.2 planets, and each single system 0.1 planets.

There are alternative choices for the planet population to be assumed in binary 
star systems.
We have taken the limit that says ``any planet around the secondary is as 
likely as around the primary''.
The opposite extreme says ``there are only planets around the primaries, not the 
secondaries''.
An intermediate case might be allowing a third occurrence rate, for instance 
splitting $\Gamma_{t,d} \rightarrow (\Gamma_{t,d1}, \Gamma_{t,d2})$ where the 
latter terms represent the occurrence rate around the primary of a double star 
system, and the occurrence rate around the secondary of a double star system.
Another intermediate would be to prescribe $\Gamma_{t}(M_\star)$ -- in other 
words to allow the true occurrence rate to vary as a function of stellar mass.

Consider the same set of questions as from the first memo.
\begin{enumerate}
\item How many single and double star systems, respectively, are in the sample?
Correspondingly, how many stars are in the sample?

\item How many planets are in the sample? (Orbiting single stars, and orbiting 
double stars respectively).

\item What is the true occurrence rate?

\item How many planets are detected?

\item What occurrence rate does astronomer A, who has never heard of binary 
star systems, derive for planets of radius $R_p$ and period $P$?

\item What occurrence rate does astronomer B, who accounts for the ``2 for 1'' 
effect of binarity (\textit{i.e.} that the sample actually has more stars than 
astronomer A thought) derive?

\item What about astronomer C, who accounts for ``2 for 1'' \textit{and} 
misclassification due to diluted radii? In other words, astronomer C did a 
combination of high resolution imaging and RV followup on every candidate, and
correctly classifies the planetary radii in every case.
\end{enumerate}

\section{How many stars are in the sample?}
\label{sec:how_many_stars}

Let $N_s$ be the number of single star systems, and $N_d$ the number of double 
star systems. $N_s$ is the same as in the first memo.

Now
\begin{equation}
N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3(L_1^N, \gamma_R, F_{\rm 
lim}^N)
\end{equation}
for
\begin{equation}
d_{\rm max,d}(L_1^N, \gamma_R, F_{\rm lim}^N) =
d_{\rm max,s}(L_1^N, F_{\rm lim}^N)\times (1+\gamma_R)^{1/2}.
\end{equation}
Since $q$ is a random variable, $\gamma_R$ is a random variable, and $d_{\rm 
max,d}$ is a random variable. Since $N_d$ is a function of $d_{\rm max,d}$, 
this means the number of double star systems is a random variable.
The mean number of double star systems in the sample is
\begin{equation}
\langle N_d \rangle = \int_{0}^{\infty} N_d\, {\rm prob}(N_d) \,{\rm d}N_d.
\end{equation}
By applying the chain rule for probability density functions, the distribution 
${\rm prob}(N_d)$ can be written 
\begin{equation}
{\rm prob}(N_d) = {\rm prob}(q(\gamma_R)) 
	  \left| \frac{{\rm d}q}{{\rm d}\gamma_R}  \right|
	  \left| \frac{{\rm d}\gamma_R}{{\rm d}d_{\rm max,d}}  \right|				
	  \left| \frac{{\rm d}d_{\rm max,d}}{{\rm d}N_d}  \right|.
\end{equation}
Doing some algebra, and assuming $\gamma_R = q^3$, this can be shown to be
\begin{align}
{\rm prob}(N_d) = \frac{2}{81} &N_d^{-1/3} \left(\frac{3}{4\pi 
n_d}\right)^{2/3} d_{\rm max,s}^{-2/3} \ \times \nonumber\\
&\left[ \left(\frac{3 N_d}{4\pi n_d}\right)^{2/3} - d_{\rm 
max,s}^2 \right]^{-2/3}
\end{align}
over the interval
\begin{equation}
{\rm lower\ bound} = \frac{4 \pi n_d}{3} (\sqrt{0.1^3 + 1} d_{\rm max,s})^3
\end{equation}
\begin{equation}
{\rm upper\ bound} = \frac{4 \pi n_d}{3} (\sqrt{2}d_{\rm max,s})^3,
\end{equation}
and outside the stated interval ${\rm prob}(N_d) = 0$.

While a closed analytic expression for $\langle N_d \rangle$ does exist, it is 
messy, and it does not yield much intuition. Instead, summarizing the important 
points and supporting them numerically:
\begin{itemize}
	\item In a volume-limited sample of binary star systems in which the 
	primary mass is fixed, and the mass ratio is drawn from a bounded uniform 
	distribution, the distribution of $\gamma_R$ will be biased towards low 
	values ($\gamma_R \approx 0.1$). This is shown in 
	Fig.~\ref{fig:gammaR_distribn_vol_limited}.
	\item In a magnitude-limited sample of binary star systems in which the 
	primary mass is fixed, and the mass ratio of the \textit{population} is 
	drawn from a bounded uniform distribution, the observed distribution of 
	mass ratios will be biased towards high values. You will see more twins,
  because they are detectable out to a greater distance. This is shown in 
	Fig.~\ref{fig:q_distribn_mag_limited}.
	\item In a magnitude-limited sample of binary star systems in which the 
	primary mass is fixed, the distribution of $\gamma_R$ will be biased 
	towards low values ($\approx 0.1$), but less so than in a 
	volume-limited sample. This is shown in 
	Fig.~\ref{fig:gammaR_distribn_mag_limited}.
\end{itemize}

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/gammaR_distribn_vol_limited.pdf}
	\end{center}
	\caption{The distribution of the luminosity ratio for a volume limited 
	sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; total 
	number density from Bovy 2017; $M(L)$ relation used is 
	Eq.~\ref{eq:mass_luminosity}; mass ratios are drawn from 
	Eq.~\ref{eq:mass_ratio}.}
	\label{fig:gammaR_distribn_vol_limited}
\end{figure}
\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/q_distribn_mag_limited.pdf}
	\end{center}
	\caption{The distribution of the mass ratio for a magnitude limited 
		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
		total 
		number density from Bovy 2017; $M(L)$ relation used is 
		Eq.~\ref{eq:mass_luminosity}; mass ratios are drawn from 
		Eq.~\ref{eq:mass_ratio} -- a \textit{uniform} distribution in a 
		volume-limited sample!}
	\label{fig:q_distribn_mag_limited}
\end{figure}
\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/gammaR_distribn_mag_limited.pdf}
	\end{center}
	\caption{The distribution of the luminosity ratio for a magnitude limited 
		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
		total 
		number density from Bovy 2017; $M(L)$ relation used is 
		Eq.~\ref{eq:mass_luminosity}; mass ratios are drawn from 
		Eq.~\ref{eq:mass_ratio}.}
	\label{fig:gammaR_distribn_mag_limited}
\end{figure}

\paragraph{How many stars are in the sample?}

We return to the original question: how many stars?
\begin{equation}
\langle N_{\rm stars} \rangle = N_s + 2 \langle N_d \rangle,
\end{equation}
as before.
Since we are not giving analytic expressions for either of the $\langle \ldots 
\rangle$ terms, we just note that in the case of a population with a single 
$\gamma_R$ value, $N_d/N_s = {\rm BF} \times (1 + \gamma_R)^{3/2}$. A lower
bound for this ratio will always be the binary fraction.
For a single run of a Monte Carlo code, where ${\rm mean}(\gamma_R) = 0.30$, 
${\rm median}(\gamma_R) = 0.16$, and the distribution was that given in 
Fig.~\ref{fig:gammaR_distribn_mag_limited}, the observed ratio was $N_d/N_s = 
0.62$.
This corresponds to a single-valued population with $\gamma_R \sim 0.23$, a
number between the mean and median of the true distribution.
It means $\sim 1.2$ stars in binary star systems for every star in a single 
star system from this sample.

We also note in passing that $N_d/N_s = 0.62$ is a smaller ratio than for the 
case of a fixed $\gamma_R$ population with $\gamma_R = 1$, which gave $N_d/N_s 
= 1.27$. The difference is that in the latter population, the volume of 
searchable stars is greater. Based on the fixed-$\gamma_R$ population scaling 
law $N_d / N_s \propto (1+\gamma_R)^{3/2}$ (derived in the first memo), we 
would expect the ratio of single stars to be $\approx (2/1.1)^{3/2} = 2.45$ 
times smaller, which is roughly (though not exactly) what we observe.

\section{How many planets are in the sample?}

The mean number of planets in the sample is
\begin{align}
\langle N_{\rm planets} \rangle
			   =& N_{\rm planets\ in\ single\ star\ systems}  +  \\
				  &\quad\quad N_{\rm planets\ in\ double\ star\ systems} 
				  \nonumber \\
			   =& \Gamma_{t,s} N_s + 2 \Gamma_{t,d} \langle N_d \rangle.
\end{align}

The factor of 2 accounts for the fact that there are twice as many stars in 
double star systems. We are assuming planets can be around either 
component.

\section{What is the true occurrence rate?}
\label{sec:true_rate}

The ``true occurrence rate'' is the average number of planets per star. Thus

\begin{align}
\Gamma_t &= \frac{\langle N_{\rm planets}\rangle}{\langle N_{\rm stars} \rangle} \\
\Gamma_t &= \frac{\Gamma_{t,s} N_s + 2 \Gamma_{t,d} \langle N_d\rangle}{N_s +
2\langle N_d\rangle}.
\label{eq:true_occ}
\end{align}

\section{How many planets are detected?}
The total number of planet detections is the sum of the number of planets 
detected in single star systems $N_{{\rm det},s}$ and the number of planets 
detected in double star systems $N_{{\rm det},d}$.
The latter of these is a random variable.
Since we selected stars for which there was enough light to make detections, 
there is no need to compute the ${\rm S/N}$ distribution of 
``threshold-crossing events''.

The number of planets detected in single star systems is
\begin{equation}
N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,{\rm geom}},
\label{eq:N_det_s}
\end{equation}
where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
star systems of the sample, and $f_{s,{\rm geom}}$ is the geometric transit 
probability.
The mean number of planets detected in double star systems is
\begin{equation}
\langle N_{{\rm det},d}\rangle = 2 \langle N_d\rangle  \Gamma_{t,d} f_{d,{\rm geom}},
\label{eq:N_det_d}
\end{equation}
where now $2 \langle N_d\rangle  \Gamma_{t,d}$ is the mean number of planets in
the double star systems of the sample.
The mean number of detected planets $\langle N_{\rm det} \rangle$ is the sum of
the two equations above.

\section{Astronomer A ignores binarity}
Astronomer A has never heard of binary star systems. What occurrence rate does 
he derive for planets of radius $R_p$ and period $P$?

The {\it total} occurrence rate (number of planets divided number of 
``stars'') for Astronomer A would be $(N_{\rm det}/f_{s,{\rm geom}})/(N_s + 
N_d)$. However, even though Astronomer A does not know about binaries, the 
radii he derives for any planets in binary systems are too small, by a factor 
$\sqrt{\mathcal{D}}$. The question asks what occurrence rate is derived for 
planets of radius $R_p$ and period $P$.
The answer is
\begin{align}
\Gamma_{\rm A,\ planets\ of\ R_p} &= \frac{N_{\rm det,s}/f_{s,{\rm geom}}}{N_s 
+ N_d}
\label{eq:A_0} \\
&= \frac{\Gamma_{t,s} N_s}{N_s + N_d}.
\end{align}

This astronomer will also think there is a second population of planets, with 
radius $R_p \sqrt{\mathcal{D}}$ (a constant number for $\gamma_R = {\rm 
constant}$). He will then also claim
have derived a second occurrence rate,
\begin{equation}
\Gamma_{\rm A,\ planets\ of\ R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm 
det,d}/f_{d,{\rm geom}}}{N_s + N_d},
\end{equation}

where at least for the twin binary case the geometric completeness term is the 
same as for Eq.~\ref{eq:A_0}.

\section{Astronomer B counts host stars correctly}
Astronomer B can somehow account correctly for the ``2 for 1'' 
effect of binarity, \textit{i.e.} that the sample actually has more stars than 
astronomer A thought.

By the same token as above,
\begin{equation}
\Gamma_{\rm B,\ planets\ of\ R_p} = \frac{N_{\rm det,s}/f_{s,{\rm geom}}}{N_s + 
2N_d},
\end{equation}
and
\begin{equation}
\Gamma_{\rm B,\ planets\ of\ R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm 
det,d}/f_{d,{\rm geom}}}{N_s 
	+ 2N_d}.
\end{equation}

\section{Astronomer C counts host stars correctly and figures out diluted radii}
Astronomer C knows which planets are in binary systems, and which are in 
single star systems.
She knows that the purported population of planets with radii $R_p 
\sqrt{\mathcal{D}}$ does not exist. All detected planets from this survey have 
radii $R_p$. She computes an occurrence rate
\begin{align}
\Gamma_{\rm C,\ planets\ of\ R_p} &= 
\frac{N_{\rm det,s}/f_{s,{\rm geom}} + N_{\rm det,d}/f_{d,{\rm geom}}}{N_s + 
2N_d} \\
&= \frac{\Gamma_{t,s} N_s + 2\Gamma_{t,d} N_d}{N_s + 2 N_d}.
\end{align}

With $\approx$ a full semester at Keck, a system-by-system analysis, 
and perfect understanding of the completeness of the detection efficiency for 
single and double star systems, Astronomer C has found the true occurrence rate 
(cf. Eq.~\ref{eq:true_occ}).

\section{Representative numbers for a few cases}

We have not given analytic expressions for the number of double star systems 
$N_d$ because even with simplified $M(L)$ relations, they are unwieldy.
However, we can ask similar numerical questions as from the first memo, and 
compare the answers.

\subsection{If we ignore binarity, for what fraction of 
	detections do we misclassify the radii? How much?}

Ignoring binarity, we will detect $N_{\rm det,s}$ planets around single stars, 
and $N_{\rm det,d}$ planets around double stars. The latter set will be assumed 
to have radii $R_p \sqrt{\mathcal{D}}$, where $\mathcal{D}$ is the vector of 
dilutions appropriate for each binary star system.
The fraction of detections with misclassified radii can then be written
\begin{equation}
\frac{N_{\rm det,d}}{N_{\rm det,s} + N_{\rm det,d}} = \frac{1}{1+\alpha},
\end{equation}
for
\begin{equation}
\alpha \equiv 
\frac{2 N_d \Gamma_{t,d} f_{d,{\rm geom}}}{N_s \Gamma_{t,s} f_{s,{\rm geom}}}
\end{equation}

For the nominal G2V dwarf case of ${\rm BF}=0.45$, and the same distributions 
used to generate 
Figs.~\ref{fig:gammaR_distribn_vol_limited}~--~\ref{fig:gammaR_distribn_mag_limited},
we found $N_d/N_s=0.62$.
We are assuming $\Gamma_{t,d} / \Gamma_{t,s} = 1$.
If we maintained the (incorrect) assumption that $f_{d,{\rm geom}} / f_{s,{\rm 
geom}} = 1$, this would give a misclassification rate of 45\%.

This latter assumption is wrong because although we are letting the occurrence 
rate be the same across binary and single systems, our binary systems have 
secondaries with varying masses. Thus they have varying radii.
Taking [Demircan \& Kahraman 1991]'s empirical fit to eclipsing binary data,
\begin{equation}
R = 1.06 M^{0.945}
\label{eq:mass_radius}
\end{equation}
for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 1991's 
stated bounds).
For a zero eccentricity system, $f_{\rm geom}=R_\star/a$.
Since $f_{d,{\rm geom}} = 0.5\times (f_{d1,{\rm geom}} + f_{d2,{\rm geom}})$, 
\textit{i.e.} it is the average of the transit probabilities for both the 
primary and secondary (d1 and d2), we can write (dropping the ``geom'' 
subscript)
\begin{align}
\frac{f_d}{f_s} &= \frac{1}{2} \left(1 + \left(\frac{M_1}{M_{d2}}\right)^{1/3} 
\frac{R_{d2}}{R_1}\right) \\
&= \frac{1}{2} \left( 1 + q^{-1/3 + 0.945}\right).
\end{align}

Computing it all out in a numerical simulation, the only thing that changes 
analytically is that
\begin{equation}
N_d f_d \rightarrow \sum_{i=1}^{N_d} f_{d,i},
\end{equation}
so when considering the ratio $f_d/f_s$, we are in fact interested in
\begin{equation}
\frac{\langle f_d \rangle}{f_s} = \frac{\frac{1}{N_d} \sum_{i=1}^{N_d} 
f_{d,i}}{f_s},
\end{equation}
where $f_{d,i}$ is the vector of geometric transit probabilities across all 
systems.
For the same numerical realization with $N_d/N_s = 0.62$, and using the mass 
radius relation given by Eq.~\ref{eq:mass_radius}, I get $\langle f_d 
\rangle / f_s = 0.84$ (smaller secondaries lead to a smaller average transit 
probability in double star systems), for a radius misclassification rate of 
$0.51$.

\subsection{Twin binaries: if we ignore binarity, how wrong is our occurrence 
	rate for planets of radius $R_p$?}

For brevity, write $\Gamma_{\rm A,\ planets\ of\ R_p} = \Gamma_{\rm A,R_p}$, 
and similarly for ${\rm C}$.
Then the relative difference between the two occurrence rates is

\begin{align}
\left|\frac{\Gamma_{\rm A,R_p} - \Gamma_{\rm C,R_p}}
{\Gamma_{\rm A,R_p}} \right| 
&=
\left| 1 - \left(
\frac{\Gamma_{t,s}N_s + 2\Gamma_{t,d}N_d}{\Gamma_{t,s} N_s }  
\cdot 
\frac{N_s + N_d}{N_s + 2N_d }
\right) \right|  \\
&=
\left| 1 - \left(
\left[ 1 + \frac{2\beta}{\chi} \right]
\cdot 
\frac{1 + \beta}{1 + 2\beta }
\right) \right|
\end{align}
for
\begin{equation}
\beta \equiv N_d/N_s, \quad \chi \equiv \Gamma_{t,s}/\Gamma_{t,d}.
\end{equation}

Again, our nominal case has $\chi = 1$, so this simplifies to
\begin{equation}
\left|\frac{\Gamma_{\rm A,R_p} - \Gamma_{\rm C,R_p}}
{\Gamma_{\rm A,R_p}} \right| = \beta.
\end{equation}

For our nominal case, this means a relative error of $62\%$.
As discussed in Sec.~\ref{sec:how_many_stars}, this is less than our earlier 
(twin binary model) case because of the smaller volume of searchable binary 
stars.

\newpage

%\begin{thebibliography}{}

%\end{thebibliography}

\end{document}

### 2017-08-30 memos/toy_analytic_survey_vary_lightratio_170803.tex
initial commit

--- 2017-08-30 memos/toy_analytic_survey_vary_lightratio.tex
@@
 for $c_q = 1/9$ to normalize the distribution to 1.
-The mass-luminosity relation is that given by [Salaris \& Cassisi (2005), 
-Sec 5.7],
-\begin{equation}
-L \propto
-\Bigg\{\begin{array}{lr}
-M^{4.5} & 0.5 < M_\star/M_\odot \leq 2.0  \\
-M^{2.6} & 0.2 < M_\star/M_\odot \leq 0.5.
-\end{array}
-\label{eq:mass_luminosity}
-\end{equation}
-For $M_1 = M_\odot$, this gives $\alpha = 4.5$ for $0.5 < q \leq 1$, and 
-otherwise $\alpha = 2.6$. 
-For analytic convenience this distribution may also be approximated as 
-a constant, \textit{e.g.,} $\alpha = 3$, in the subsequent development.
+The mass-luminosity relation can, for analytic convenience, be approximated as 
+$L = M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
+While we use this in subsequent analytic development, for numerics we fit a 
+lines to mass-luminosity data collected by Torres et al. [2009] for 
+dwarfs above M, and Benedict et al. [2016] for dwarfs below M.
+To convert Benedict et al. [2016]'s reported $M_V$ values to absolute 
+luminosities, we interpolated over E. Mamajek's 
+table\footnote{
+\texttt{pas.rochester.edu/\textasciitilde 
+emamajek/EEM\_dwarf\_UBVIJHK\_colors\_Teff.txt},
+downloaded 2017.08.02}.
+In log-log space, we let the intersection point of the lines float, and make 
+various cuts on the data as indicated in Fig~\ref{fig:mass_luminosity}. The 
+fit parameters are available in a footnote\footnote{
+	m lo: 1.8818719873988132~--~
+	c lo: -0.9799647314108376~--~
+	m hi: 5.1540712426599882~--~
+	c hi: 0.0127626185389781~--~
+	M at merge: 0.4972991257826812~--~
+	L at merge: 0.0281260412126928.
+	}.
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=0.9]{figures/mass_luminosity.pdf}
+	\end{center}
+	\caption{Empirical fit to main sequence dwarf mass luminosity data from 
+	Torres et al. [2009] and Benedict et al. [2016]. The ``low fit'' is a least 
+	squares fit to data from $0.1-0.5M_\odot$, and the ``high fit'' is to data 
+	above that, and below the Kraft break ($1.2M_\odot$).
+	The $L(M)$ relation taken for subsequent numerics is the maximum 
+	of the two fits.
+	}
+	\label{fig:mass_luminosity}
+\end{figure}
 
@@
 
+In passing, the bias in ${\rm prob}(\gamma_R)$ towards low luminosity ratios 
+can be seen analytically. If we assume $\gamma_R = q^\alpha$, then
+\begin{equation}
+{\rm prob}(\gamma_R) = \frac{1}{9 \alpha} \ \gamma_R^{\frac{1-\alpha}{\alpha}}
+\quad {\rm for\ } (0.1)^\alpha < \gamma_R < 1,
+\end{equation}
+and otherwise zero. For instance if $\alpha = 3$, ${\rm prob}(\gamma_R) 
+\propto \gamma_R^{-2/3}$ and the domain extended from $1$ to 
+$10^{-3}$, where the probability distribution peaks.
+
 \begin{figure}[!t]
@@
 	sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; total 
-	number density from Bovy 2017; $M(L)$ relation used is 
-	Eq.~\ref{eq:mass_luminosity}; mass ratios are drawn from 
+	number density from Bovy 2017; $M(L)$ relation from 
+	Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
 	Eq.~\ref{eq:mass_ratio}.}
@@
 		total 
-		number density from Bovy 2017; $M(L)$ relation used is 
-		Eq.~\ref{eq:mass_luminosity}; mass ratios are drawn from 
+		number density from Bovy 2017; $M(L)$ relation from 
+		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
 		Eq.~\ref{eq:mass_ratio} -- a \textit{uniform} distribution in a 
@@
 		total 
-		number density from Bovy 2017; $M(L)$ relation used is 
-		Eq.~\ref{eq:mass_luminosity}; mass ratios are drawn from 
+		number density from Bovy 2017; $M(L)$ relation from 
+		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
 		Eq.~\ref{eq:mass_ratio}.}
@@
 Fig.~\ref{fig:gammaR_distribn_mag_limited}, the observed ratio was $N_d/N_s = 
-0.62$.
+0.59$.
 This corresponds to a single-valued population with $\gamma_R \sim 0.23$, a
@@
 
-We also note in passing that $N_d/N_s = 0.62$ is a smaller ratio than for the 
+We also note in passing that $N_d/N_s = 0.59$ is a smaller ratio than for the 
 case of a fixed $\gamma_R$ population with $\gamma_R = 1$, which gave $N_d/N_s 
@@
 Figs.~\ref{fig:gammaR_distribn_vol_limited}~--~\ref{fig:gammaR_distribn_mag_limited},
-we found $N_d/N_s=0.62$.
+we found $N_d/N_s=0.59$.
 We are assuming $\Gamma_{t,d} / \Gamma_{t,s} = 1$.
@@
 systems.
-For the same numerical realization with $N_d/N_s = 0.62$, and using the mass 
+For the same numerical realization with $N_d/N_s = 0.59$, and using the mass 
 radius relation given by Eq.~\ref{eq:mass_radius}, I get $\langle f_d 
@@
 
-For our nominal case, this means a relative error of $62\%$.
+For our nominal case, this means a relative error of $59\%$.
 As discussed in Sec.~\ref{sec:how_many_stars}, this is less than our earlier 

### 2017-08-30 memos/toy_analytic_surveys_170804.tex
initial commit

%\documentclass[12pt,preprint]{aastex}
%\documentclass[preprint2,12pt]{aastex}
\documentclass{emulateapj}
%\usepackage{apjfonts}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{todonotes}

\shortauthors{LGB, KM, JNW}
\shorttitle{Toy transit surveys}

\begin{document}

% ------------------------------------------------------------------------
% New commands
%
\def\ltsima{$\; \buildrel < \over \sim \;$}
\def\lsim{\lower.5ex\hbox{\ltsima}}
\def\gtsima{$\; \buildrel > \over \sim \;$}
\def\gsim{\lower.5ex\hbox{\gtsima}}
\def\tess{{\it TESS} }
\def \teff {T_{\rm eff}}
\def \phir {\Phi_{\rm R}}
\def \fov {24$^{\circ}$}
\def \pixsz {21.1''}
\def \aeff {69.1 cm$^2$ }    
\def \epd {105 mm}

% -------------------------------------------------------------------------
%

\bibliographystyle{apj}

\title{ Analytic Transit Surveys }

\author{
  LGB, KM, JNW
}

% \journalinfo{Draft version}
\slugcomment{Memo for internal use}

%\altaffiltext{1}{Princeton University}

\begin{abstract}

We describe a few different toy transit surveys.

First, we derive the signal to noise distribution of planets detected in a 
survey where all the stars and planets are the same (except some stars are twin 
binary systems). We give expressions for the number of detected planets, and 
for possible errors one can make when deriving occurrence rates.

Then, we allow the properties of the secondary in binary systems to vary. A 
distribution of light ratios introduces additional biases, which we discuss.
We reassess the errors related to occurrence rates.

\end{abstract}

\keywords{planets and satellites:\ detection}

\section{Model \#1: fixed stars, fixed planets, fixed light-ratio binaries}
\label{sec:model_1}

Ah, transiting planets! We learn so much by studying them. But how much can we 
actually learn, and how much is messed up by binarity?

Imagine the following transit survey:
\begin{enumerate}
\item You are going to observe the entire sky for a duration $T_{\rm obs}$, 
with a detector of area $A$, and known bandpass. Your detector is photon-noise 
limited.
%
\item You are interested only in detecting planets of radius $R_p$, and orbital 
period $P$. For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
%
\item You are only interested in detecting them around stars of radius $R_1$, 
and luminosity $L_1$. For instance, G2V dwarfs.
%
\item You can only detect your `nominal planet' when you observe signals with
${\rm S/N} > {\rm (S/N)_{min}}$.
For a photon-noise limited survey, this minimum signal to noise ratio is 
equivalent to a minimum flux, $F_{\rm min}$.
%
\item However, you opt to observe all the ``points'' on-sky with apparent 
magnitude $m < m_{\rm lim}$, or equivalently with energy flux in your bandpass 
greater than some limit $F > F_{\rm lim}$.
(You will not be able to detect planets around the faintest of 
these, and will need to perform a completeness correction to later derive 
occurrence rates.)
Generally $F_{\rm lim} < F_{\rm min}$, \textit{e.g.}, whenever the original 
magnitude cut defines a ``source catalog'' from which ``searchable stars'' are 
selected, as in the KIC [Batalha et al, 2010, Brown et al 2011] and the TIC 
[Stassun et al 2017].
\end{enumerate}

You get funding, and perform the S/N-limited survey. All planets with 
${\rm S/N} > {\rm (S/N)_{min}}$ are detected. The rest are not.
You now wish to derive an occurrence rate for planets of radius $R_p$ and 
orbital period $P$.
Assume your universe is a universe in which:
\begin{itemize}
	\item The true population of ``points'' (stellar systems, all unresolved) 
	comprises both single and double star systems. Single star systems have 
	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
	temperature $T_{\rm eff,1}$.
	Double star systems have luminosity in the observed bandpass $L_d = 
	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
	the secondary to the primary. 
	In this section, $\gamma_R$ is a constant across the population 
	of star systems.
	The ratio of the two number densities in a 
	volume-limited sample is the binary fraction\footnote{The binary fraction 
		is equivalent to the multiplicity fraction if there are no triple, 
		quadruple, $\ldots$ systems.}.
	\item The true population of planets around these stars is as follows:
	\subitem A fraction $\Gamma_{t,s}$ of stars in single star systems 
	have a planet of radius $R_p$, with orbital period $P$.
	\subitem A fraction $\Gamma_{t,d}$ of each star in a double star 
	system has a planet of radius $R_p$, with orbital period $P$. For instance, 
	if $\Gamma_{t,s} = \Gamma_{t,d} = 0.1$, on average each double 
	system contributes 0.2 planets, and each single system 0.1 planets.
	Any astrophysical difference in planet formation between singles and 
	binaries is captured by these two terms.
	Note that we are taking an limit by assuming the primary and secondary of 
	a binary system host planets at equal rates. If the secondary is 
	non-identical to the primary, it's clearly wrong.	
\end{itemize}

For simplicity, one might wish to imagine the observer correctly 
pre-selects all of the ``searchable stars'', in the style of [Pepper et al, 
2003].
In other words, they would begin by knowing ${\rm (S/N)_{min}}$, and then only 
observe stars for which they would have adequate flux to achieve it -- setting 
$F_{\rm lim} = F_{\rm min}$.
This simplification would naively allow the observer to ignore 
completeness effects when deriving occurrence rates, since their completeness 
is 100\%.

We do not begin this way because in a stellar population of both singles and 
binaries, the above approach ignores necessary incompleteness for 
binary systems.
If we set the limiting flux to give 100\% completeness for single stars, we 
will unwittingly select binaries near that flux limit, for which dilution 
will push transit signals below the detection threshold.
This is because a flux limit maps onto three different maximum 
detection distances for a population of single stars and twin binaries:
\begin{enumerate}
\item $d_{{\rm max},s}$: the maximum distance out to which single stars are
  selected. This can be chosen to coincide with the maximum distance out to
  which planets are detectable around single stars.
%
\item $d_{{\rm max},d}^\star$: the maximum distance out to which double stars
  are selected. This is different from $d_{{\rm max},d}^{\rm p}$, the maximum
  distance out to which planets are detectable around double stars. For a 
  population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
  L^{-1/2} d$,
  \begin{equation}
    d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
    (1 + \gamma_R)^{-1} d_{{\rm max},d}^\star.
  \end{equation}
  For $\gamma_R = 1$, this means only 1 in 8 selected binary stars can yield
  detectable planets.
\end{enumerate}
We develop tools to address this and related completeness effects from the 
outset.

Consider the following questions:
\begin{enumerate}
\item How many single and double star systems, respectively, are in the sample?
Correspondingly, how many stars are in the sample?

\item How many planets are in the sample? (Orbiting single stars, and orbiting 
double stars respectively).

\item What is the true occurrence rate?

\item How many planets are detected?

\item What occurrence rate does astronomer A, who has never heard of binary 
star systems, or completeness, derive for planets of radius $R_p$ and period 
$P$?

\item What occurrence rate does astronomer B, who accounts for the ``2 for 1'' 
effect of binarity (\textit{i.e.} that the sample actually has more stars than 
astronomer A thought) derive? (Astronomer B neglects completeness).

\item What about astronomer C, who accounts for ``2 for 1'' \textit{and} 
misclassification due to diluted radii? In other words, astronomer C did a 
combination of high resolution imaging and RV followup on every candidate, and
correctly classifies the planetary radii in every case.

\item What about astronomer D, who additionally notes the importance of 
completeness?
\end{enumerate}

\subsection{How many stars are in the sample?}

Let $N_s$ be the number of single star systems, and $N_d$ the number of double 
star systems. Then the total number of stars in the sample is
\begin{equation}
N_{\rm stars} = N_s + 2N_d.
\end{equation}

In a magnitude-limited sample in which stars are uniformly distributed in 
volume, the number of stars will be the number density times the volume.
If the volume is taken to be a sphere over which the number density is uniform,
\begin{equation}
N_i = n_{i} \frac{4\pi}{3} d_{{\rm max},i}^3,
\label{eq:number_systems}
\end{equation}
for $i\in{\rm \{single,double\} } \equiv { \{s,d\} }$, and
\begin{equation}
\frac{n_{d}}{n_{s}} = {\rm binary\ fraction} \equiv {\rm BF}
\end{equation}
by definition. The absolute normalization of the number density is a measured 
quantity, as is the binary fraction. For G2V dwarfs, the latter is $\approx 
0.45$ [Duchene \& Kraus, 2013]. The former is given by [Bovy 2017].

$d_{{\rm max},i}$ in Eq.~\ref{eq:number_systems} is the maximum distance 
corresponding to the given magnitude limit:
\begin{equation}
d_{{\rm max},i} = \left(\frac{L_{i}}{4 \pi F_{{\rm min}}}\right)^{1/2},
\label{eq:d_max}
\end{equation}
where the limiting flux in the bandpass $F_{{\rm min}}$ can also be 
stated in terms of the limiting magnitude $m_{{\rm min}}$,
\begin{equation}
m_{{\rm min}} = m_{0} - \frac{5}{2} \log_{10} \left(\frac{F_{{\rm 
min}}}{F_{0}}\right),
\end{equation}
for $m_{0}$ a zero-point magnitude and $F_{0}$ its corresponding flux (as 
always, everything is implicitly written in a defined bandpass).

In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$, and as a 
consequence the maximum distance to which binary stars will be selected is 
greater than that of single stars, simply as a consequence of imposing a 
magnitude cut.
The ratio of double to single systems is
\begin{align}
\frac{N_d}{N_s} &= 
	\frac{n_d}{n_s} \left(\frac{d_{{\rm max}, d} }{d_{{\rm max}, s}}\right)^3 \\
&= {\rm BF} \times (1+\gamma_R)^{3/2}.
\end{align}
In the nominal case of twin binaries ($\gamma_R = 1$), with a binary fraction 
${\rm BF} = 0.5$, there are 
$\sqrt{2}$ more binary systems than single systems in the sample.
Correspondingly, there are $2\sqrt{2}$ more stars in binary systems than stars 
in single systems.

As a comment on Eq.~\ref{eq:number_systems}, if 
we wished to write a stellar number density profile that accounted for the 
vertical structure of the Milky Way, we might choose a profile either 
$\propto \exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from 
the galactic midplane and $H$ a scale-height. Both density profiles would lead 
closed form analytic solutions.

\subsection{How many planets are in the sample?}

The number of planets in the sample is
\begin{align}
N_{\rm planets} =& N_{\rm planets\ in\ single\ star\ systems}  +  \\
				  &\quad\quad N_{\rm planets\ in\ double\ star\ systems} 
				  \nonumber \\
			   =& \Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d.
\end{align}

The factor of 2 accounts for the fact that there are twice as many stars in 
double star systems.

\subsection{What is the true occurrence rate?}
\label{sec:true_rate}

The ``true occurrence rate'' is the average number of planets per star. Thus

\begin{align}
\Gamma_t &= \frac{N_{\rm planets}}{N_{\rm stars}} \\
\Gamma_t &= \frac{\Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d}{N_s + 2N_d}.
\label{eq:true_occ}
\end{align}

\subsection{How many planets are detected?}
The total number of planet detections is the sum of the number of planets 
detected in single star systems $N_{{\rm det},s}$ and the number of planets 
detected in double star systems $N_{{\rm det},d}$.
These can be expressed individually. 
The former is
\begin{equation}
N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,{\rm geom}} f_{s,{\rm S/N} > {\rm 
(S/N)_{min}}},
\label{eq:N_det_s}
\end{equation}
where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
star systems of the sample, $f_{s,{\rm geom}}\equiv f_{s,g}$ is the geometric 
transit probability, and $f_{s,({\rm S/N} > {\rm (S/N)_{min}})} \equiv f_{s,c}$ 
is the fraction of these transiting planets that are observed with signal to 
noise greater than the minimum detection threshold (the completeness). 
Analogously,
\begin{equation}
N_{{\rm det},d} = 2 N_d \Gamma_{t,d} f_{d,g} f_{d,c},
\label{eq:N_det_d}
\end{equation}
where now $2 N_d \Gamma_{t,d}$ is the number of planets in the double star 
systems of the sample, the geometric transit probability is the same (if we 
have twin binaries -- not if we consider a distribution of secondaries) and the 
completeness term must account for any differences in the signal to noise 
distribution that come from stellar binarity.

\subsubsection{Analytic completeness}
Since the geometric transit probability is known, the only terms we have yet to 
compute are the completeness terms, $f_{i,c}$ for $i\in{\rm \{single,double\} 
}$. We proceed as follows.

The signal ${\rm S}$ for a box-car train transiting planet is
\begin{align}
{\rm S} &= \delta \mathcal{D} \\
&= \left(\frac{R_p}{R_\star}\right)^2 \mathcal{D},
\end{align}
for $R_p$ the planet's radius, $R_\star$ that of its host star, and 
$\mathcal{D}$ the dilution parameter defined as
\begin{align}
  \mathcal{D} &= 
  \Bigg\{\begin{array}{lr}
  L_1 / L_d, & {\rm\ if\ binary\ and\ target\ primary}\\
  \gamma_R L_1 / L_d, & {\rm\ if\ binary\ and\ target\ 
  secondary}\\
  1, & {\rm\ if\ single},
  \end{array} \nonumber\\
  &=
    \Bigg\{\begin{array}{lr}
    (1+\gamma_R)^{-1}, & {\rm\ if\ binary\ and\ target\ primary}\\
    (1 + \gamma_R^{-1})^{-1}, & {\rm\ if\ binary\ and\ target\ 
    	secondary}\\
    1, & {\rm\ if\ single},
    \end{array} 
  \label{eq:dilution}
\end{align}
where $L_1, L_d,$ and $\gamma_R$ were defined in the opening monograph.

Assuming the only source of noise is Poissonian counting noise, the noise ${\rm 
N}$ can be written
\begin{equation}
{\rm N} = \frac{1}{\sqrt{N_\gamma}},
\end{equation}
for $N_\gamma$ the number of photons received by the detector. This noise model 
is a useful simplification -- see [Howell 2006, pg 75] for the full 
CCD equation.
The number of received photons can be written
\begin{equation}
N_\gamma = F^{\rm N}_\gamma A N_{\rm tra} T_{\rm dur},
\end{equation}
for $F^{\rm N}_\gamma$ the photon number flux from the system [${\rm 
ph\,cm^{-2}\,s^{-1}}$], 
$A$ the detector area, $T_{\rm dur}$ the transit duration, and $N_{\rm tra} $ 
the number of transits observed, which is multiplied in assuming the transits 
are ``phase-folded''.

Thus the signal to noise ratio can be written
\begin{equation}
{\rm S/N} = \delta \mathcal{D} \sqrt{F^{\rm N}_\gamma A N_{\rm tra} T_{\rm 
dur}}.
\label{eq:snr_ivory_tower}
\end{equation}

This means we can write the minimum number flux of photons required for a 
detection at threshold as
\begin{equation}
F_{\rm lim}^N = \left[ \left({\rm \frac{S}{N}}\right)_{\rm min} \frac{1}{\delta 
	\mathcal{D}} \right]^2 \frac{1}{A N_{\rm tra} T_{\rm dur}}.
\end{equation}
To convert this to $F_{\rm lim}$, multiply by the average photon energy in the 
bandpass.

In passing, given the parameters that define a survey and planet type, 
Eq.~\ref{eq:snr_ivory_tower} would need to be re-expressed with 
$N_{\rm tra}$ roughly the ratio of the observing baseline to the planet period, 
and $T_{\rm dur}$ a function of $R_\star, P, a$, and impact parameter $b$, and 
then perhaps averaged over $b$. We leave them as-is for subsequent 
development.

The interesting term in Eq.~\ref{eq:snr_ivory_tower} that changes between stars 
of the same binarity class in our idealized sample is the square root of 
$F^{\rm N}_\gamma$.
This is the term that leads to a distribution of signal to noises for different 
stars.
The completenesses $f_{i,c}$ can be directly 
expressed in terms of those probability density functions:
\begin{equation}
f_{i,c} = 
	\int_{{\rm (S/N)_{min}}}^{\infty} 
		{\rm d}\left(\frac{{\rm S}}{{\rm N}}\right)_i \ 
		{\rm prob}\left(\frac{{\rm S}}{{\rm N}}\right)_i.
\label{eq:completeness_long}
\end{equation}
We keep the subscript $i$ because the signal to noise distributions are 
different for the cases of single star systems ($i=s$) and double star systems 
($i=d$).
Notably:
\begin{itemize}
	\item The dilution differs (Eq.~\ref{eq:dilution}).
	\item The photon number flux from the system differs.
\end{itemize}

To simplify notation, we let $x_i \equiv {\rm (S/N)}_i$, and rewrite 
Eq.~\ref{eq:completeness_long} as
\begin{equation}
f_{i,c} = 
\int_{{\rm x_{\rm min}}}^{\infty} 
{\rm d}x_i \ 
{\rm prob}(x_i).
\label{eq:completeness_short}
\end{equation}

\subsubsection{Deriving ${\rm prob}(x_i)$}
We want expressions for the probability density function of the observed signal 
to noise ratio, ${\rm prob}(x_i)$, for both the single and binary 
system case.

First, note that a star placed uniformly in the volume of the search space will 
have a probability density function for its distance $r$ from the origin of
\begin{equation}
{\rm prob} (r) = \frac{3 r^2}{d_{\rm max}^3},
\end{equation}
where the appropriate maximum distances should be substituted per 
Eq.~\ref{eq:d_max}.
Noting the transformation rule for probability density functions, we can 
evaluate the probability of a star having a observed number flux $F^{\rm 
N}_{i,\gamma}$ in the bandpass,
\begin{align}
{\rm prob} (F^{\rm N}_{i,\gamma}) 
&= {\rm prob}(r(F^{\rm N}_{i,\gamma}))
	\left| \frac{{\rm d} r}{{\rm d} F^{\rm N}_{i,\gamma}} \right| \\
&= \frac{3}{2 d_{\rm max}^3} c_i^{3/2} (F^{\rm N}_{i,\gamma})^{-5/2},
\label{eq:pdf_observed_flux}
\end{align}
where in the latter equality we have written a ``number luminosity'' $c_i$ 
(units of inverse time) defined for $i\in{\rm \{single,double\} }$ as
\begin{equation}
c_i = 
\Bigg\{\begin{array}{lr}
R_1^2 F^{\rm N}_{s1,\gamma}, & {\rm\ if\ single}\\
R_1^2 F^{\rm N}_{s1,\gamma} + R_2^2 F^{\rm N}_{s2,\gamma} & {\rm\ if\ double}.
\end{array}
\label{eq:c_i}
\end{equation}
In Eq.~\ref{eq:c_i}, $F^{\rm N}_{s1,\gamma}$ and $F^{\rm N}_{s2,\gamma}$ are 
the photon number fluxes at the surfaces of the stars. To derive 
Eq.~\ref{eq:pdf_observed_flux}, we simply scaled these by the distance:
\begin{equation}
F^{\rm N}_{i,\gamma} = \frac{c_i}{r^2}.
\label{eq:flux_at_detector}
\end{equation}

The surface photon number fluxes $F^{\rm N}_{si,\gamma}$ in Eq.~\ref{eq:c_i} 
are usually evaluated numerically, by convolving the wavelength-specific photon 
flux density of a star with the dimensionless spectral response function of the 
instrument. In other words,
\begin{equation}
F^{\rm N}_{s,\gamma} = \int F_\lambda T_\lambda \, {\rm d}\lambda.
\label{eq:pfd}
\end{equation}
The wavelength-specific photon flux density $F_\lambda$ [${\rm 
ph\,cm^{-2}\,s^{-1}\,\AA^{-1}}$] could be from Pickles' library, or could be 
a blackbody function. The transmission function is, up to factor of order 
unity, a step function between two wavelengths $\lambda_{\rm min}$ and 
$\lambda_{\rm max}$.
If we assume a blackbody source, Eq.~\ref{eq:pfd} becomes
\begin{equation}
F^{\rm N}_{s,\gamma} = 8\pi c \left(\frac{k T}{h c}\right)^3 
\int_{hc/(\lambda_{\rm max} kT)}^{hc/(\lambda_{\rm min} kT)}
\frac{u^2}{e^u - 1} \,{\rm d} u,
\end{equation}
which can be evaluated numerically\footnote{It may help in the numerics to note 
that infinite series representations of this type of dimensionless integral 
exist and converge quickly. For instance, one can show that
\begin{equation}
\int_{0}^{a} \frac{u^3}{e^u -1} \,{\rm d}u = \sum_{n=1}^{\infty} \frac{6}{n^4} 
- \frac{e^{-an}}{n^4} (6 + 6an + 3(an)^2 + (an)^3).
\nonumber
\end{equation}
A similar expression exists for the similar integral in the text. 
[Michels 1968] explains an analogous derivation.
}.

The importance of the functional form of $F^{\rm N}_{s,\gamma}$ is that it is 
to first order only a function of the blackbody temperature and the bandpass 
wavelength bounds.
Thus in the most general case $c_i(R_1, R_2, T_1, T_2, \lambda_{\rm min}, 
\lambda_{\rm max})$ and nothing else.
The only random variable involved in the flux being received at the detector is 
$r$, so we can indeed write the flux received at the detector as in 
Eq.~\ref{eq:flux_at_detector}.

We can finally write out the probability density functions for the signal to 
noise ratios in the single and double-star cases by using the transformation 
rule for pdfs, and applying Eq.~\ref{eq:snr_ivory_tower}.
For single stars,
\begin{equation}
{\rm prob}(x_s) = \frac{3}{d_{\rm max, s}^3} c_s^{3/2} \delta^3 \left( A T_{\rm 
dur} N_{\rm tra}\right)^{3/2} x_s^{-4}.
\label{eq:prob_xs}
\end{equation}
Analogously for double stars,
\begin{equation}
{\rm prob}(x_d) = \frac{3}{d_{\rm max, d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 
\left( A T_{\rm dur} N_{\rm tra}\right)^{3/2} x_d^{-4}.
\label{eq:prob_xd}
\end{equation}

\subsubsection{Number of detected planets}

Performing the integrals of Eq.~\ref{eq:completeness_short}, we get:

\begin{align}
N_{{\rm det},s} &= N_s \Gamma_{t,s} f_{s,g} f_{s,c} \\
&= N_s \Gamma_{t,s} \frac{R_\star}{a} \frac{1}{d_{\rm max,s}^3} c_s^{3/2} 
\delta^3 \left(A 
T_{\rm dur} N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3},
\end{align}
and
\begin{align}
N_{{\rm det},d} &= 2 N_d \Gamma_{t,d} f_{d,g} f_{d,c} \\
&= 2 N_d \Gamma_{t,d} \frac{R_\star}{a}
\frac{1}{d_{\rm max,d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 \left(A T_{\rm dur} 
N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3}.
\end{align}

Formally, the $f_{i,c}$ terms should be written as ${\rm min}(1, \ldots)$, 
where $(\ldots)$ is the given expression. This ensures that the fraction of 
planets above the signal to noise threshold is less than 1.
Note that we assumed twin binaries in the geometric transit probabilities.
Interestingly, the ratio of the two completeness fractions is
\begin{equation}
\frac{f_{d,c}}{f_{s,c}} = \left(\frac{d_{\rm max,s}}{d_{\rm max,d}}\right)^3 
\left(\frac{c_d}{c_s}\right)^{3/2} \mathcal{D}^3 = (1+\gamma_R)^{-3},
\end{equation}
which, for the $\gamma_R = 1$ case, gives 
$1/8$. (We got the same number from simple scaling laws at the outset, but the 
above machinery will be useful when generalizing in Sec.~\ref{sec:model_2}.)

The number of detected planets $N_{\rm det}$ is the sum of the two appropriate 
equations above, and can be written
\begin{align}
N_{\rm det} = 
&\left( \frac{\delta}{x_{\rm min}} \right)^3 (A T_{\rm dur} N_{\rm tra})^{3/2}
\frac{R_\star}{a}
\nonumber\\
&\quad \times
\left[ N_s \Gamma_{t,s} \frac{c_s^{3/2}}{d_{\rm max,s}^3}  +
	   2 N_d \Gamma_{t,d} \frac{c_d^{3/2}}{d_{\rm max,d}^3}  \right].
\label{eq:N_det}
\end{align}
All terms can be input to a computer, and checked against a Monte Carlo 
simulation if desired.

\subsection{Astronomer A ignores binarity}
Astronomer A has never heard of binary star systems. 
Nor has he heard of completeness corrections.
He knows about geometric transit probabilities.
What occurrence rate does he derive for planets of radius $R_p$ and period $P$?

The {\it total} occurrence rate (number of planets divided by number of 
``stars'') for Astronomer A would be $N_{\rm det}/(N_s + N_d)$, up to the 
geometric correction. 
However, even though Astronomer A does not know about binaries, the radii he 
derives for any planets in binary systems are too small, by a factor 
$\sqrt{\mathcal{D}}$ (for twin binaries, they are all $R_p/\sqrt{2}$).
Astronomer A wants an occurrence rate for planets of radius $R_p$ and period 
$P$.
The answer is thus
\begin{equation}
\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/f_{s,g}}{N_s + N_d}.
\end{equation}

This astronomer will also think there is a second population of planets, with 
radius $R_p \sqrt{\mathcal{D}}$, and will thus rush to Nature claiming to also
have derived a second occurrence rate,
\begin{equation}
\Gamma_{{\rm A}, R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm det,d}/f_{s,g}}{N_s 
+ N_d}.
\end{equation}
Note in the second case that this astronomer thinks these are single stars, and 
generally will miscompute the geometric transit probability.

\subsection{Astronomer B counts host stars correctly}
Astronomer B can somehow account correctly for the ``2 for 1'' 
effect of binarity, \textit{i.e.} that the sample actually has more stars than 
astronomer A thought.

By the same token as above,
\begin{equation}
\Gamma_{{\rm B},R_p} = \frac{N_{\rm det,s}/f_{s,g}}{N_s + 2N_d},
\end{equation}
and
\begin{equation}
\Gamma_{{\rm B},R_p \sqrt{\mathcal{D}}} = 
		\frac{N_{\rm det,d}/f_{d,g}}{N_s + 2N_d}.
\end{equation}

\subsection{Astronomer C counts host stars correctly and figures out diluted 
radii}
Astronomer C did high resolution imaging followup on every candidate, and 
correctly classifies the planetary radii.
Thus, she also knows which planets are in binary systems, and which are in 
single star systems.

She knows that the purported population of planets with radii $R_p 
\sqrt{\mathcal{D}}$ does not exist. All detected planets from this survey have 
radii $R_p$. She computes an occurrence rate
\begin{equation}
\Gamma_{{\rm C},R_p} = 
\frac{N_{\rm det,s}/f_{s,g} + N_{\rm det,d}/f_{d,g}}{N_s + 2N_d}
\end{equation}

the closest yet to the true rate (Sec.~\ref{sec:true_rate}).

\subsection{Astronomer D counts host stars correctly, figures out diluted 
radii, 
and accounts for completeness}

Astronomer D, knows which detections were around 
binaries and the associated radius correction.
They do injection recovery, and derive correct estimates for their 
completeness functions about single stars $f_{s,c}$, and 
about double star systems, $f_{d,c}$.
With this knowledge in hand, they compute the respective single and binary 
occurrence rates
\begin{equation}
\Gamma_{t,s} = \frac{N_{\rm det,s}}{N_s f_{s,g} f_{s,c}},
\end{equation}
\begin{equation}
\Gamma_{t,d} = \frac{N_{\rm det,d}}{2 N_d f_{d,g} f_{d,c}}.
\end{equation}

With these in hand, they derive the overall occurrence rate
\begin{align}
\Gamma_{{\rm D}, R_p} 
&= \frac{N_{\rm det,s}/(f_{s,g}f_{s,c}) + N_{\rm det,d}/(f_{d,g}f_{d,c})  }{N_s 
+ 2 N_d} \\
&= \frac{\Gamma_{t,s} N_s + 2\Gamma_{t,d} N_d}{N_s + 2 N_d}.
\end{align}

Although we admit it's been a bit of a slog, it happens that Astronomer D's 
occurrence rate is the true occurrence rate (cf. Eq.~\ref{eq:true_occ}).

All it takes is $\approx$ a full semester at Keck, a system-by-system analysis, 
and perfect understanding of the completeness of the detection efficiency for 
single and double star systems.

\subsection{Numerical verification}

To check the preceding analytic development, we implemented a Monte Carlo 
simulation of this idealized transit survey.
To run the survey, we defined the instrument specifications (detector area and 
transmission function), the stellar population (binary fraction, total number 
density of a given stellar class, fixed stellar 
properties), the planet population (fixed planet radius, period, and occurrence 
rate about single and binary stars), and finally the survey parameters 
(observing baseline, minimum SNR for ``detection'').
We then randomly drew star positions, randomly assigned planets to stars in 
single and binary systems, and computed the resulting signal to noise 
(Eq.~\ref{eq:snr_ivory_tower}) with which 
the transits would be observed.
As in the preceding analytics, we assumed ``twin'' binaries (same stellar 
radii, same effective 
temperature, and dilution does not depend on which stellar binary is the 
``target'').

The results are shown in Fig.~\ref{fig:snr_dist_analytic_v_numeric}, and 
indicate that the analytic probability distribution functions 
Eqs.~\ref{eq:prob_xs},~\ref{eq:prob_xd}, and the number of 
detections (Eq.~\ref{eq:N_det}) are correct.

A point evident in Fig.~\ref{fig:snr_dist_analytic_v_numeric} is that, for 
fixed planet parameters, and fixed stellar parameters ($R_\star, L_\star$, and 
distance $r$) the SNR distribution for planets in binaries is poorer than that 
of planets in single star systems.
We can see analytically that this simply due to dilution:
\begin{equation}
\frac{{\rm prob}(x_d)}{{\rm prob}(x_s)} =(1 + \gamma_R)^{-1} = \mathcal{D}.
\end{equation}
Deriving this simple form requires noting that the ratios of the 
bandpass-specific number luminosities (Eq.~\ref{eq:c_i}) is equal to the ratio 
of the bandpass-specific energy luminosities (otherwise a term with $c_s/c_d$ 
must be included).

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=0.4]{figures/snr_distribution.pdf}
	\end{center}
	\caption{Comparison of analytic and numeric probability 
	density functions of the SNR in an idealized transit survey.
	The analytic lines are Eqs.~\ref{eq:prob_xs} and~\ref{eq:prob_xd} for the 
	planet populations orbiting single and binary stars. The underlying stepped 
	histogram is output from Monte Carlo simulations. Poisson noise leads to 
	a small deviation at the faint and bright limits, but the numerics and 
	analytics otherwise agree.
		 }
	\label{fig:snr_dist_analytic_v_numeric}
\end{figure}

\subsection{Representative numbers for a few cases}

\subsubsection{Twin binaries: if we ignore binarity, for what fraction of 
detections do we misclassify the radii?}

Ignoring binarity, we will detect $N_{\rm det,s}$ planets around single stars, 
and $N_{\rm det,d}$ planets around double stars. The latter set will be assumed 
to have radii $R_p \sqrt{\mathcal{D}}$.
The fraction of detections with misclassified radii can then be written
\begin{equation}
\frac{N_{\rm det,d}}{N_{\rm det,s} + N_{\rm det,d}} = \frac{1}{1+\alpha},
\end{equation}
for
\begin{equation}
\alpha \equiv \frac{1}{2({\rm BF})} (1+\gamma_R)^{3/2} 
\frac{\Gamma_{t,s}}{\Gamma_{t,d}}.
\end{equation}

For the nominal G2V dwarf case of ${\rm BF}=0.45$, twin binaries with equal 
occurrence rates this produces a misclassification rate of $24\%$, in agreement 
with Fig.~\ref{fig:snr_dist_analytic_v_numeric}.

\subsubsection{Twin binaries: if we ignore binarity, how wrong is our
occurrence rate for planets of radius $R_p$?}

This is almost simply asking ``what is the relative difference between the 
occurrence rates derived by Astronomers D and A for planets of radius $R_p$?''
However, in the more realistic case, Astronomer A also has derived a 
completeness, which we assume is the same as for Astronomer D in the single 
star case.
So Astronomer A now misclassifies planetary radii, and miscounts the total 
number of stars, but knows his completeness for single stars.
Astronomer D corrects all these errors.
Then
\begin{equation}
\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d},
\end{equation}
and the relative difference between the two occurrence rates is

\begin{align}
\left|\frac{\Gamma_{{\rm A},R_p} - \Gamma_{{\rm D},R_p}}
		   {\Gamma_{{\rm A},R_p}} \right| &=
\left| 1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}}\right| \\
&=
\left| 1 - \left(\frac{\Gamma_{t,s}N_s + 2\Gamma_{t,d}N_d}{N_s + 2N_d}  \cdot 
\frac{N_s + N_d}{\Gamma_{t,s} N_s }\right)\right| \\
&= 
\left|
1 - \frac{(1 + 2 \beta \Gamma_{t,d}/\Gamma_{t,s})(1 + \beta)}{(1+2\beta)}
\right|,
\end{align}
for
\begin{equation}
\beta \equiv N_d/N_s = {\rm BF} \times (1+\gamma_R)^{3/2}.
\end{equation}

For the nominal G2V dwarf case of ${\rm BF}=0.45$ with twin binaries ($\gamma_R 
= 1$) and $\Gamma_{t,d} = \Gamma_{t,s}$ this gives a relative error of $127\%$.
For instance, in the numerical simulation corresponding to 
Fig.~\ref{fig:snr_dist_analytic_v_numeric}, Astronomer A finds $\Gamma_{\rm 
A,R_p} = 0.22$, while Astronomer 
D derives the true (input) occurrence rate of $\Gamma_{\rm D,R_p} = 0.5$.

\section{Model \#2: fixed stars, fixed planets, varying light-ratio binaries}
\label{sec:model_2}

Same story as Model \#1, except now $\gamma_R$ varies across the population 
of star systems.
It does so because the underlying mass ratio varies. 
Since we are interested in solar type binaries, we take the distribution of the 
mass ratio $q=M_2/M_1$ by approximating [Rhagavan et al 2010, Fig 16]:
\begin{equation}
{\rm prob}(q) =
\Bigg\{\begin{array}{lr}
c_q & 0.1 < q \leq 1  \\
0 & {\rm otherwise},
\end{array}
\label{eq:mass_ratio}
\end{equation}
for $c_q = 1/9$ to normalize the distribution to 1.
The mass-luminosity relation can, for analytic convenience, be approximated as 
$L = M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
While we use this in subsequent analytic development, for numerics we fit a 
lines to mass-luminosity data collected by Torres et al. [2009] for 
dwarfs above M, and Benedict et al. [2016] for dwarfs below M.
To convert Benedict et al. [2016]'s reported $M_V$ values to absolute 
luminosities, we interpolated over E. Mamajek's 
table\footnote{
	\texttt{pas.rochester.edu/\textasciitilde 
		emamajek/EEM\_dwarf\_UBVIJHK\_colors\_Teff.txt},
	downloaded 2017.08.02}.
In log-log space, we let the intersection point of the lines float, and make 
various cuts on the data as indicated in Fig~\ref{fig:mass_luminosity}. The 
fit parameters are available in a footnote\footnote{
	m lo: 1.8818719873988132~--~
	c lo: -0.9799647314108376~--~
	m hi: 5.1540712426599882~--~
	c hi: 0.0127626185389781~--~
	M at merge: 0.4972991257826812~--~
	L at merge: 0.0281260412126928.
}.
\begin{figure}
	\begin{center}
		\includegraphics[scale=0.9]{figures/mass_luminosity.pdf}
	\end{center}
	\caption{Empirical fit to main sequence dwarf mass luminosity data from 
		Torres et al. [2009] and Benedict et al. [2016]. The ``low fit'' is a 
		least 
		squares fit to data from $0.1-0.5M_\odot$, and the ``high fit'' is to 
		data 
		above that, and below the Kraft break ($1.2M_\odot$).
		The $L(M)$ relation taken for subsequent numerics is the maximum 
		of the two fits.
	}
	\label{fig:mass_luminosity}
\end{figure}

The other changes between Model \#2 and Model \#1 are:
\begin{enumerate}
\item We allow a third occurrence rate, splitting $\Gamma_{t,d} \rightarrow 
(\Gamma_{t,d1}, \Gamma_{t,d2})$ where the latter terms represent the occurrence 
rate around the primary of a double star system, and the occurrence rate around 
the secondary of a double star system. For the case of \textit{e.g.}, $q=0.1$, 
this seems more relevant.
We can then take the limit that says ``any planet around the secondary is as 
likely as around the primary'', or go to the opposite extreme of ``there are 
only planets around the primaries, not the secondaries''.
%
\item Our binary systems have secondaries with varying masses. Thus they have 
varying radii. Taking [Demircan \& Kahraman 1991]'s empirical fit to eclipsing 
binary data,
\begin{equation}
R = 1.06 M^{0.945}
\label{eq:mass_radius}
\end{equation}
for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 1991's 
stated bounds), and for each in solar units. This only affects the transit 
probabilities about the secondaries of binary star systems.
\end{enumerate}

We then ask the same questions as for Model \#1.

\subsection{How many stars are in the sample?}

Let $N_s$ be the number of single star systems, and $N_d$ the number of double 
star systems. $N_s$ is the same as in Sec.~\ref{sec:model_1}.

Now
\begin{equation}
N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3(L_1^N, \gamma_R, F_{\rm 
	lim}^N)
\end{equation}
for
\begin{equation}
d_{\rm max,d}(L_1^N, \gamma_R, F_{\rm lim}^N) =
d_{\rm max,s}(L_1^N, F_{\rm lim}^N)\times (1+\gamma_R)^{1/2}.
\end{equation}
Since $q$ is a random variable, $\gamma_R$ is a random variable, and $d_{\rm 
	max,d}$ is a random variable. Since $N_d$ is a function of $d_{\rm max,d}$, 
the number of double star systems becomes a random variable.
The expected (mean) number of double star systems in the sample is
\begin{equation}
\langle N_d \rangle = \int_{0}^{\infty} N_d\, {\rm prob}(N_d) \,{\rm d}N_d.
\end{equation}
By applying the chain rule for probability density functions, the distribution 
${\rm prob}(N_d)$ can be written 
\begin{equation}
{\rm prob}(N_d) = {\rm prob}(q(\gamma_R)) 
\left| \frac{{\rm d}q}{{\rm d}\gamma_R}  \right|
\left| \frac{{\rm d}\gamma_R}{{\rm d}d_{\rm max,d}}  \right|				
\left| \frac{{\rm d}d_{\rm max,d}}{{\rm d}N_d}  \right|.
\end{equation}
Doing some algebra, and assuming $\gamma_R = q^3$, this can be shown to be
\begin{align}
{\rm prob}(N_d) = \frac{2}{81} &N_d^{-1/3} \left(\frac{3}{4\pi 
	n_d}\right)^{2/3} d_{\rm max,s}^{-2/3} \ \times \nonumber\\
&\left[ \left(\frac{3 N_d}{4\pi n_d}\right)^{2/3} - d_{\rm 
	max,s}^2 \right]^{-2/3}
\end{align}
over the interval
\begin{equation}
{\rm lower\ bound} = \frac{4 \pi n_d}{3} (\sqrt{0.1^3 + 1} d_{\rm max,s})^3
\end{equation}
\begin{equation}
{\rm upper\ bound} = \frac{4 \pi n_d}{3} (\sqrt{2}d_{\rm max,s})^3,
\end{equation}
and outside the stated interval ${\rm prob}(N_d) = 0$.

While a closed analytic expression for $\langle N_d \rangle$ does exist, it is 
messy, and it does not yield much intuition. Instead, summarizing the important 
points and supporting them numerically:
\begin{itemize}
	\item In a volume-limited sample of binary star systems in which the 
	primary mass is fixed, and the mass ratio is drawn from a bounded uniform 
	distribution, the distribution of $\gamma_R$ will be biased towards low 
	values ($\gamma_R \approx 0.1$). This is shown in 
	Fig.~\ref{fig:gammaR_distribn_vol_limited}.
	\item In a magnitude-limited sample of binary star systems in which the 
	primary mass is fixed, and the mass ratio of the \textit{population} is 
	drawn from a bounded uniform distribution, the observed distribution of 
	mass ratios will be biased towards high values. You will see more twins,
	because they are detectable out to a greater distance\footnote{This is a 
	scarcy systematic w.r.t. the claimed intrinsic excess of twin binaries.}. 
	This is shown in 
	Fig.~\ref{fig:q_distribn_mag_limited}.
	\item In a magnitude-limited sample of binary star systems in which the 
	primary mass is fixed, the distribution of $\gamma_R$ will be biased 
	towards low values ($\approx 0.1$), but less so than in a 
	volume-limited sample. This is shown in 
	Fig.~\ref{fig:gammaR_distribn_mag_limited}.
\end{itemize}

In passing, the bias in ${\rm prob}(\gamma_R)$ towards low luminosity ratios 
can be seen analytically. If we assume $\gamma_R = q^\alpha$, then
\begin{equation}
{\rm prob}(\gamma_R) = \frac{1}{9 \alpha} \ \gamma_R^{\frac{1-\alpha}{\alpha}}
\quad {\rm for\ } (0.1)^\alpha < \gamma_R < 1,
\end{equation}
and otherwise zero. For instance if $\alpha = 3$, ${\rm prob}(\gamma_R) 
\propto \gamma_R^{-2/3}$ and the domain extends from $1$ to 
$10^{-3}$, where the probability distribution peaks.

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/gammaR_distribn_vol_limited.pdf}
	\end{center}
	\caption{The distribution of the luminosity ratio for a volume limited 
		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
		total 
		number density from Bovy 2017; $M(L)$ relation from 
		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
		Eq.~\ref{eq:mass_ratio}.}
	\label{fig:gammaR_distribn_vol_limited}
\end{figure}
\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/q_distribn_mag_limited.pdf}
	\end{center}
	\caption{The distribution of the mass ratio for a magnitude limited 
		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
		total 
		number density from Bovy 2017; $M(L)$ relation from 
		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
		Eq.~\ref{eq:mass_ratio} -- a \textit{uniform} distribution in a 
		volume-limited sample!}
	\label{fig:q_distribn_mag_limited}
\end{figure}
\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/gammaR_distribn_mag_limited.pdf}
	\end{center}
	\caption{The distribution of the luminosity ratio for a magnitude limited 
		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
		total 
		number density from Bovy 2017; $M(L)$ relation from 
		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
		Eq.~\ref{eq:mass_ratio}.}
	\label{fig:gammaR_distribn_mag_limited}
\end{figure}

\paragraph{How many stars are in the sample?}

We return to the original question: how many stars?
\begin{equation}
\langle N_{\rm stars} \rangle = N_s + 2 \langle N_d \rangle,
\end{equation}
as before.
Since we are not giving analytic expressions for either of the $\langle \ldots 
\rangle$ terms, we just note that in the case of a population with a single 
$\gamma_R$ value, $N_d/N_s = {\rm BF} \times (1 + \gamma_R)^{3/2}$. A lower
bound for this ratio will always be the binary fraction.
For a single run of a Monte Carlo code, where ${\rm mean}(\gamma_R) = 0.28$, 
${\rm median}(\gamma_R) = 0.13$, and the distribution was that given in 
Fig.~\ref{fig:gammaR_distribn_mag_limited}, the observed ratio was $N_d/N_s = 
0.59$.
This corresponds to a single-valued population with $\gamma_R \sim 0.20$, a
number between the mean and median of the true distribution.
It means $\sim 1.2$ stars in binary star systems for every star in a single 
star system from this sample.

We also note in passing that $N_d/N_s = 0.59$ is a smaller ratio than for the 
case of a fixed $\gamma_R$ population with $\gamma_R = 1$, which gave $N_d/N_s 
= 1.27$. The difference is that in the latter population, the volume of 
searchable stars is greater. Based on the fixed-$\gamma_R$ population scaling 
law $N_d / N_s \propto (1+\gamma_R)^{3/2}$, we 
would expect the ratio of single stars to be $\sim (2/1.1)^{3/2} = 2.4$ 
times smaller, which is roughly (though not exactly) what we observe.

\subsection{How many planets are in the sample?}

The mean number of planets in the sample is
\begin{align}
\langle N_{\rm planets} \rangle
=& N_{\rm planets\ in\ single\ star\ systems}  +  \\
&\quad\quad N_{\rm planets\ in\ double\ star\ systems} 
\nonumber \\
=& \Gamma_{t,s} N_s + (\Gamma_{t,d1} + \Gamma_{t,d2}) \langle N_d \rangle.
\end{align}

What was formerly a factor of 2 has now been split into the true occurrence 
rates about the primaries and secondaries of double star systems.

\subsection{What is the true occurrence rate?}

The ``true occurrence rate'' is the average number of planets per star. Thus

\begin{align}
\Gamma_t &= \frac{\langle N_{\rm planets}\rangle}{\langle N_{\rm stars} 
\rangle} \\
\Gamma_t &= \frac{\Gamma_{t,s} N_s + (\Gamma_{t,d1} + \Gamma_{t,d2}) \langle 
N_d\rangle}{N_s +	2\langle N_d\rangle}.
\label{eq:true_occ_2}
\end{align}

\subsection{How many planets are detected?}

The total number of planet detections is the sum of the number of planets 
detected in single star systems $N_{{\rm det},s}$ and the number of planets 
detected in double star systems $N_{{\rm det},d}$.
The latter of these is a random variable, and can be further split into the 
primary and secondary contributions $N_{{\rm det},d1}$, $N_{{\rm det},d2}$.

The number of planets detected in single star systems is
\begin{equation}
N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,g} f_{s,c},
\end{equation}
where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
star systems of the sample, $f_{s,g}$ is the geometric probability of the 
planets transiting, and $f_{s,c}$ is the fraction of transiting planets around 
single star systems that are detected (the completeness).

The mean number of planets detected in double star systems is
\begin{equation}
\langle N_{{\rm det},d}\rangle = \langle N_d\rangle  
(\Gamma_{t,d1} f_{d1,g} f_{d1,c} + 
\Gamma_{t,d2} f_{d2,g} f_{d2,c} ).
\label{eq:N_det_d_2}
\end{equation}
Now $\langle N_d\rangle  \Gamma_{t,d1}$ is the mean number of planets orbiting 
the primaries of double star systems, ditto the corresponding expression for 
the secondaries.
For this population, $f_{d1,g} = f_{s,g}$, but $f_{d1,g} \neq f_{d2,g}$.
The completeness fractions differ between the primary and secondary because 
the dilution as a function of the light ratio, $\mathcal{D}(\gamma_R)$, has 
different behavior for the two cases.
We show this in Fig.~\ref{fig:rad_vs_gammaR}.
Specifically, a planet orbiting the secondary will have worse dilution than a 
planet orbiting the primary, and consequently lower completeness.

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.9]{figures/observed_radii_vs_gammaR.pdf}
	\end{center}
	\caption{Dilution $\mathcal{D}$ and observed to true radius ratio 
	$\sqrt{\mathcal{D}}$ plotted against binary light ratio $\gamma_R$. This 
	illustrates	Eq.~\ref{eq:dilution}. The lower panel's horizontal line is at 
	$1/\sqrt{2}$.
	The implication is that if a detected planet orbits the secondary and you do 
	not know it, you cannot measure $R_p$ to better than $\approx 29\%$ 
	accuracy.
	Conversely, if it orbits the primary and you do not know it, you will get 
	the radius wrong by at most $\approx 29\%$.
	}
	\label{fig:rad_vs_gammaR}
\end{figure}

We could probably derive analytic expressions for all the terms between 
brackets in Eq.~\ref{eq:N_det_d_2}, completeness included.
I'm not yet convinced we need to -- we should get cancellations for
most of the questions we want to ask.

\subsubsection{Analytic completeness}
\texttt{todo, if necessary}
\subsubsection{Deriving ${\rm prob}(x_i)$}
\texttt{todo, if necessary}
\subsubsection{Number of detected planets}
\texttt{todo, if necessary}

\subsection{Astronomer A ignores binarity}

Just as in Sec.~\ref{sec:model_1}, Astronomer A
\begin{itemize}
\item has never heard about binary star systems.
\item has never heard about completeness corrections.
\item knows about geometric transit probabilities.
\end{itemize}

What occurrence rate does he derive for planets of radius $R_p$ and period $P$?

The answer is the same as in Sec.~\ref{sec:model_1},
\begin{equation}
\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/f_{s,g}}{N_s + N_d}.
\end{equation}

However, now rather than planets detected in binary systems being perceived as 
a second population of planets with fixed radius $R_p \sqrt{\mathcal{D}}$, 
there will be an apparent spectrum of diluted radii, since $\mathcal{D}$ 
varies by system.
Writing the number of detections in double systems as a function of observed 
radius, $N_{\rm det,d}(R_p^{\rm obs})$, Astronomer A would derive an 
occurrence rate for this separate population of
\begin{equation}
\Gamma_{\rm A}(R_p^{\rm obs}) = \frac{N_{\rm det,d}(R_p^{\rm 
obs})/f_{s,g}}{N_s + N_d}.
\end{equation}
Note that this astronomer thinks these planets orbit single stars, 
and so miscomputes the geometric transit probability for the secondaries.

A more realistic question is ``what fraction of multiples have $\gamma_R$ 
smaller enough so that you'd still observe a radius close to the true one?''. 
We discuss this in Sec.~\ref{sec:close_enough}.

\subsection{Astronomer B counts host stars correctly}

Same as astronomer A above, except the denominator becomes $N_s + 2N_d$.

\subsection{Astronomer C counts host stars correctly and figures out diluted
radii}

In Sec.~\ref{sec:model_1}, Astronomer C only had to do high resolution 
imaging to measure $\gamma_R$ for each system. Since for $\gamma_R = 1$, the 
dilution is single-valued (Fig.~\ref{fig:rad_vs_gammaR}), this immediately 
told her the true planet radius (since the primary/secondary distinction was 
meaningless).

In Model \#2's universe, the dilution is double-valued for a given $\gamma_R$.
To correctly work out the diluted radii, Astronomer C now needs more than high 
resolution imaging -- she needs to resolve the stars during transit to 
discover which star the planet orbits.
She also needs to derive masses for the primaries and secondaries (to make the 
geometric transit probability correction).

Only after doing this does she find that all detected planets from this survey 
have radii $R_p$. She computes an occurrence rate
\begin{equation}
	\Gamma_{{\rm C},R_p} = 
	\frac
			{N_{\rm det,s}/f_{s,g} +
			  N_{\rm det,d1}/f_{d1,g} +
			  N_{\rm 	det,d2}/f_{d2,g}}
			{N_s + 2N_d}
\end{equation}

the closest yet to the true rate (Sec.~\ref{sec:true_rate}).

\subsection{Astronomer D counts host stars correctly, figures out diluted
radii, and accounts for completeness}

Same story as for Model \#1.
Now, they do injection recovery and derive correct estimates for their 
completeness functions about single stars $f_{s,c}$, the primaries of double 
star systems, $f_{d1,c}$, and the secondaries of double star systems 
$f_{d2,c}$.
They compute the respective single and binary 
occurrence rates
\begin{equation}
\Gamma_{t,s} = \frac{N_{\rm det,s}}{N_s f_{s,g} f_{s,c}},
\end{equation}
\begin{equation}
\Gamma_{t,d1} = \frac{N_{\rm det,d1}}{N_d f_{d1,g} f_{d1,c}},
\end{equation}
\begin{equation}
\Gamma_{t,d2} = \frac{N_{\rm det,d2}}{N_d f_{d2,g} f_{d2,c}}.
\end{equation}

With these in hand, they derive the overall occurrence rate
\begin{align}
\Gamma_{{\rm D}, R_p} 
&= \frac{
	N_{\rm det,s}/f_s + 
	N_{\rm det,d1}/f_{d1} +
	N_{\rm det,d2}/f_{d2}
	}
	{N_s + 2 N_d} \\
&= \frac{\Gamma_{t,s} N_s + (\Gamma_{t,d1} + \Gamma_{t,d2})N_d}
				{N_s + 2 N_d},
\end{align}
where $f_s \equiv f_{s,g}f_{s,c}, f_{d1} \equiv f_{d1,g}f_{d1,c}, 
f_{d2} \equiv f_{d2,g}f_{d2,c}$ for brevity.
Astronomer D's occurrence rate is the true occurrence rate (cf. 
Eq.~\ref{eq:true_occ_2}).

\subsection{Numerical verification}
\texttt{todo, if needed}

\subsection{Representative numbers for a few cases}

We have not given analytic expressions for the number of double star systems 
$N_d$ because even with simplified $M(L)$ relations, they are unwieldy.
However, we can ask similar numerical questions as from Sec.~\ref{sec:model_1}, 
and compare the answers.

\subsubsection{If we ignore binarity, for what fraction of detections do we
misclassify the radii?}

Ignoring binarity, we will detect $N_{\rm det,s}$ planets around single stars, 
and $N_{\rm det,d}$ planets around double stars. The latter set will be assumed 
to have radii $R_p \sqrt{\mathcal{D}}$, where $\mathcal{D}$ is the vector of 
dilutions appropriate for each binary star system.
The fraction of detections with misclassified radii can then be written
\begin{equation}
\frac{N_{\rm det,d}}{N_{\rm det,s} + N_{\rm det,d}} = \frac{1}{1+\alpha},
\end{equation}
for
\begin{equation}
\alpha \equiv 
\frac{ N_d (\Gamma_{t,d1} f_{d1} + \Gamma_{t,d2} f_{d2}) }{N_s \Gamma_{t,s} 
f_s},
\end{equation}
where $f_s \equiv f_{s,g}f_{s,c}, f_{d1} \equiv f_{d1,g}f_{d1,c}, 
f_{d2} \equiv f_{d2,g}f_{d2,c}$ for brevity.

% For the nominal G2V dwarf case of ${\rm BF}=0.45$, and the same distributions 
% used to generate 
% Figs.~\ref{fig:gammaR_distribn_vol_limited}~--~\ref{fig:gammaR_distribn_mag_limited},
% we found $N_d/N_s=0.59$.
% We are assuming $\Gamma_{t,d} / \Gamma_{t,s} = 1$.
% If we maintained the (incorrect) assumption that $f_{d,{\rm geom}} / f_{s,{\rm 
% 		geom}} = 1$, this would give a misclassification rate of 45\%. 
% 
% This latter assumption is wrong because although we are letting the occurrence 
% rate be the same across binary and single systems, our binary systems have 
% secondaries with varying masses. Thus they have varying radii.
% Taking [Demircan \& Kahraman 1991]'s empirical fit to eclipsing binary data,
% \begin{equation}
% R = 1.06 M^{0.945}
% \label{eq:mass_radius}
% \end{equation}
% for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 1991's 
% stated bounds).
% For a zero eccentricity system, $f_{\rm geom}=R_\star/a$.
% Since $f_{d,{\rm geom}} = 0.5\times (f_{d1,{\rm geom}} + f_{d2,{\rm geom}})$, 
% \textit{i.e.} it is the average of the transit probabilities for both the 
% primary and secondary (d1 and d2), we can write (dropping the ``geom'' 
% subscript)
% \begin{align}
% \frac{f_d}{f_s} &= \frac{1}{2} \left(1 + \left(\frac{M_1}{M_{d2}}\right)^{1/3} 
% \frac{R_{d2}}{R_1}\right) \\
% &= \frac{1}{2} \left( 1 + q^{-1/3 + 0.945}\right).
% \end{align}
% 
% Computing it all out in a numerical simulation, the only thing that changes 
% analytically is that
% \begin{equation}
% N_d f_d \rightarrow \sum_{i=1}^{N_d} f_{d,i},
% \end{equation}
% so when considering the ratio $f_d/f_s$, we are in fact interested in
% \begin{equation}
% \frac{\langle f_d \rangle}{f_s} = \frac{\frac{1}{N_d} \sum_{i=1}^{N_d} 
% 	f_{d,i}}{f_s},
% \end{equation}
% where $f_{d,i}$ is the vector of geometric transit probabilities across all 
% systems.
% For the same numerical realization with $N_d/N_s = 0.59$, and using the mass 
% radius relation given by Eq.~\ref{eq:mass_radius}, I get $\langle f_d 
% \rangle / f_s = 0.84$ (smaller secondaries lead to a smaller average transit 
% probability in double star systems), for a radius misclassification rate of 
% $0.51$.

\texttt{The different secondary completeness matters here if proceeding 
analytically. Would be easier to do numerically.
Regardless, this is not the most interesting question right now.}

\subsubsection{If we ignore binarity, how wrong is our occurrence rate for
planets of radius $R_p$?}

Write $\Gamma_{{\rm A,\ planets\ of\ }R_p} = \Gamma_{{\rm A,}R_p}$, 
and similarly for ${\rm D}$.
Just as in Model \#1, the answer to ``what is the relative difference between 
the occurrence rates derived by Astronomers D and A for planets of radius 
$R_p$?''
is simpler when we assume that Astronomer A has also derived a 
completeness, which we assume is the same as for Astronomer D in the single 
star case.
So Astronomer A now misclassifies planetary radii, and miscounts the total 
number of stars, but knows his completeness for single stars.
Astronomer D corrects all these errors.
Then
\begin{equation}
	\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d}.
\end{equation}

The relative difference between the two occurrence rates, $\Delta \equiv 
\left|(\Gamma_{{\rm A},R_p} - \Gamma_{{\rm D},R_p})/\Gamma_{{\rm A},R_p} 
\right|$, is
\begin{align}
	\Delta &=
	\left| 1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}}\right| \\
	&=
	\left| 1 - \left(\frac{\Gamma_{t,s}N_s + (\Gamma_{t,d1} + 
	\Gamma_{t,d2})N_d}{N_s + 2N_d}  
	\cdot 
	\frac{N_s + N_d}{\Gamma_{t,s} N_s }\right)\right| \\
	&= 
	\left|
	1 - \frac{(1 + \beta (\Gamma_{t,d1}+\Gamma_{t,d2})/\Gamma_{t,s})(1 + 
	\beta)}{(1+2\beta)}
	\right| \\
	&=
	\left| 1 - \left(
	\left[ 1 + \chi \beta \right]
	\cdot 
	\frac{1 + \beta}{1 + 2\beta }
	\right) \right|
\end{align}
for
\begin{equation}
\beta \equiv N_d/N_s, \quad \chi \equiv 
(\Gamma_{t,d1}+\Gamma_{t,d2})/\Gamma_{t,s}.
\end{equation}

In the limit where the $R_p$ planet has the same occurrence rate about every 
host type regardless of its mass, $\chi=2$, and this simplifies to $\Delta = 
\beta$, for a relative error of $59\%$.
This is less than for Model \#1 because of the smaller volume of searchable 
binary stars in Model \#2.

If we consider the opposite limit of the $R_p$ planet \textit{only} 
existing around single stars and the primaries of double star systems with 
equal occurrence, $\chi = 1$, and $\Delta = \beta^2 / (1+2\beta)$.
For $\beta = 0.59$, this gives $\Delta = 0.16$.

\subsubsection{What if we ignore binarity, but count derived planet radii that 
are ``close enough''?}
\label{sec:close_enough}

Let $x$ be the acceptable margin of error on the observed radius, so that we 
are interested in counting the number of planets detected with $R_p < R_p^{\rm 
obs} < (1\pm x) R_p$.
For Model \#2, since all planets have radius $R_p$, we only care about 
the smaller radius case, \textit{i.e.}, we want to count $R_p < R_p^{\rm 
	obs} < (1 - x) R_p$, for instance with $x=0.1$.

All of these diluted detections will be from binary systems. The number of 
planets detected in double star systems with observed radius $R_p^{\rm obs}$ 
is
\begin{align}
N_{\rm det,d}(R_p^{\rm obs}) &= N_{\rm det, d1}(R_p^{\rm obs}) + 
															  N_{\rm det, d2}(R_p^{\rm obs}) \\
&= \Gamma_{t,d1} f_{d1}(R_p^{\rm obs}) N_{d1}(R_p^{\rm obs})\  + \nonumber \\
&\quad\quad \quad\Gamma_{t,d2} f_{d2}(R_p^{\rm obs}) N_{d2}(R_p^{\rm obs}).
\end{align}
As always, the fractional terms $f_{di}$ for $i\in{1,2}$ encompass geometric 
and completeness corrections.
We're writing both them and the number of stars for which a given observed 
radius would be seen as functions of $R_p^{\rm obs}$.

To save on notational horribleness, write $R_p' \equiv \{(1-x) R_p < R_p^{\rm 
	obs} < R_p\}$, in other words let $R_p'$ denote the set of planets 
	that are observed in the desired radius range.

Then writing $N_{d1}(R_p')$ is relatively straight-forward:
\begin{equation}
N_{d1}(R_p') = \int_{0}^{{\rm min}(1,\gamma_{R,u})} N_d(\gamma_R) 
\,{\rm prob}(\gamma_R) \,{\rm d}(\gamma_R),
\end{equation}
where the upper limit is set by equating the limiting radius $(1-x)R_p$ to the 
observed diluted radius $(1+\gamma_R)^{-1/2}R_p$, giving
\begin{equation}
\gamma_{R,u} \equiv \frac{x (2-x)}{(1-x)^2}.
\end{equation}
If $x > (1 - 2^{-1/2})$, bad things happen (cf. Fig.~\ref{fig:rad_vs_gammaR}), 
hence the need for the minimum.

The analogous equation for secondaries is
\begin{equation}
N_{d2}(R_p') = 
  \Bigg\{\begin{array}{lr}
  \int_{\gamma_{R,l}}^{1} N_d(\gamma_R)
  \,{\rm prob}(\gamma_R) \,{\rm d}(\gamma_R),\\ 
  \quad\quad\quad\quad\quad\quad\quad\quad x > 1 - 
  2^{-1/2}\\
  0, \quad\quad\quad\quad\quad\quad\quad x \leq 1 - 2^{-1/2}
  \end{array}
\label{eq:N_Rp_secondary}
\end{equation}
where
\begin{equation}
\gamma_{R,l} \equiv \frac{(1-x)^2}{x (2-x)}.
\end{equation}
Eq.~\ref{eq:N_Rp_secondary} accounts for the fact that you cannot measure 
$R_p$ to better than $\approx 29\%$ if the planet orbits the secondary of an 
unidentified binary.
This was originally noted in Fig.~\ref{fig:rad_vs_gammaR}.

If we were forward-modelling (\textit{i.e.} we actually wanted to compute 
$N_{\rm det,d}(R_p')$) we would need to evaluate the completeness 
term and this ``fraction of slightly diluted stars'' terms together, and 
integrate jointly to find $N_{\rm det,d1}(R_p')$ and $N_{\rm det,d2}(R_p')$.

\texttt{If we ask the right question, this will not be necessary.}

Astronomer A misclassifies planetary radii, miscounts the total number of 
stars, but knows his completeness and geometric correction for single stars as 
a function of radius $f_s(R_p')$.
His occurrence rate for planets of radius \textit{near} $R_p$, \textit{i.e.} 
the set of planets with $\{ (1-x)R_p < R_p^{\rm obs} < R_p \}$, is
\begin{equation}
\Gamma_{{\rm A}, R_p'} = \frac{ (N_{\rm det,s} + N_{\rm det,d}(R_p')) 
/ f_s(R_p')   }{ N_s + N_d}.
\end{equation}

For him,
\begin{equation}
f_s(R_p') = f_{s,g}(R_p') f_{s,c}(R_p').
\end{equation}

Astronomer D, who had everything right, doesn't change anything.

The new difference between the two rates is
\begin{align}
\Delta' &\equiv \left|
\frac{\Gamma_{{\rm A},R_p'} - \Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p'} }
\right| \\
&= \left|
1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p'} } 
\right| \\
&=
\left| 1 -
\left( \frac{1+\chi \beta}{1 + \xi}  \cdot 
\frac{1 + \beta}{1 + 2\beta} \right)
\right|,
\end{align}
for
\begin{equation}
\beta \equiv N_d/N_s, \quad \chi \equiv (\Gamma_{t,d1} + 
\Gamma_{t,d2})/\Gamma_{t,s},
\end{equation}
and
\begin{equation}
\xi \equiv \frac{N_{\rm det,d}(R_p')}{f_s(R_p')} \cdot \frac{1}{N_s 
\Gamma_{t,s}}.
\end{equation}
If we let $x < 1 - 2^{-1/2}$, for instance $x=0.1$, then $N_{d2}(R_p') = 0$. 
The expression for $\xi$ simplifies to
\begin{equation}
\chi = \frac{N_{d1}(R_p')}{N_s} \cdot \frac{\Gamma_{t,d1}}{\Gamma_{t,s}} \cdot 
\frac{f_{d1}(R_p')}{f_s(R_p')},\quad ({\rm if\ }x < 1-2^{-1/2}).
\label{eq:chi_simplfied}
\end{equation}

If we assume $\Gamma_{t,d1} = \Gamma_{t,s}$, as we were doing 
anyway, Eq.~\ref{eq:chi_simplfied} is bounded above by
$\beta$. This is because $N_{d1}(R_p') \leq N_{d1} = N_d$, and $f_{d1}(R_p') / 
f_s(R_p') \leq 1$, since over a fixed radius interval dilution will always 
yield a lower completeness fraction for double star systems than singles.
Recall for the fixed $\gamma_R$ population we showed that $f_{d1}/f_s = 
(1+\gamma_R)^{-3}$.

Regardless, the cheap way of making the estimate without deriving the actual 
completeness fractions, or even running numerics, is just to plot the error as 
a function of $\xi$. We do this in Fig.~\ref{fig:deltaprime_vs_xi}, which 
shows that the
relative error varies between $\approx 5\% - 35\%$, depending on what 
fraction of secondaries you assume host planets.

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=.8]{figures/deltaprime_vs_xi.pdf}
	\end{center}
	\caption{\textit{Y axis}: relative error between the Astronomer who ignores 
	binarity 
	but counts planetary radii that are ``close enough'' (A') and the Astronomer 
	who knows everything (D).
	\textit{X axis}: $\chi$ as defined in Eq.~\ref{eq:chi_simplfied}. Taking a 
	guess at the true value, if $N_{d1}(R_p') = N_d/2$, and the occurrence rates 
	between `d1' and `s' are the same, and we put $\gamma_R = 0.1$ into the 
	completeness ratio, then $\xi = (\beta/2) \times (1+0.1)^{-3} = 0.22$. The 
	relative error then varies between $\approx 5\% - 35\%$, depending on what 
	fraction of secondaries you assume host planets.
		}
	\label{fig:deltaprime_vs_xi}
\end{figure}

%FIXME

% \begin{comment}
% \subsection{How many stars are in the sample?}
% \subsection{How many planets are in the sample?}
% \subsection{What is the true occurrence rate?}
% \subsection{How many planets are detected?}
% \subsubsection{Analytic completeness}
% \subsubsection{Deriving ${\rm prob}(x_i)$}
% \subsubsection{Number of detected planets}
% \subsection{Astronomer A ignores binarity}
% \subsection{Astronomer B counts host stars correctly}
% \subsection{Astronomer C counts host stars correctly and figures out diluted
% radii}
% \subsection{Astronomer D counts host stars correctly, figures out diluted
% radii, and accounts for completeness}
% \subsection{Numerical verification}
% \subsection{Representative numbers for a few cases}
% \subsubsection{If we ignore binarity, for what fraction of detections do we
% misclassify the radii?}
% \subsubsection{If we ignore binarity, how wrong is our occurrence rate for
% planets of radius $R_p$?}
% \subsubsection{If we ignore binarity, but allow derived planet radii to be
% ``close enough'', how do we do?}
% 
% \end{comment}

\newpage

%\begin{thebibliography}{}

%\end{thebibliography}

\end{document}

### 2017-08-30 memos/workdir/toy_analytic_surveys.tex
initial commit

--- 2017-08-30 memos/toy_analytic_surveys_170804.tex
@@
 
+\def \kepler {{\it Kepler}}
+
 % -------------------------------------------------------------------------
@@
 
-\title{ Analytic Transit Surveys }
+\title{ Toy Transit Surveys }
 
@@
 
-We describe a few different toy transit surveys.
-
-First, we derive the signal to noise distribution of planets detected in a 
-survey where all the stars and planets are the same (except some stars are twin 
-binary systems). We give expressions for the number of detected planets, and 
-for possible errors one can make when deriving occurrence rates.
+What errors does ignoring binarity introduce on the occurrence rates derived 
+from transit surveys?
+
+First, we show that in a universe where all stars have fixed properties
+(except that some are twin binaries), the results of a search for planets of a 
+given radius and orbital period can be expressed analytically.
 
 Then, we allow the properties of the secondary in binary systems to vary. A 
-distribution of light ratios introduces additional biases, which we discuss.
-We reassess the errors related to occurrence rates.
+distribution of light ratios introduces a heavy detection bias towards 
+equal-mass binaries in magnitude limited samples.
+We show that dependent upon planet occurrence's functional dependence on host 
+star mass (and binarity), the errors introduced in derived occurrence rates vs. 
+real occurrence rates can range from a few percent to at most $\approx 60\%$ 
+relative error.
+
+Finally, we perform a synthetic transit survey analogous to {\it Kepler}. We 
+show that if we assume every star in this {\it Kepler} survey analog is single, 
+we underestimate occurrence rates for a given planet radius and period by 
+$\approx 39\%$.
+This is immediately applicable to the question of Earth-like planets around 
+Sun-like stars.
 
@@
 \keywords{planets and satellites:\ detection}
+
+\section{Introduction}
+
+Ah, transiting planets! We learn so much by studying them. But how much can we 
+actually learn, and how much is messed up by binarity?
+
+Binary companions to single stars introduce a surprising number of biases into 
+transit surveys.
+For a start, any magnitude limited survey over-represents binaries compared to 
+a volume-limited survey -- the familiar Malmquist bias, though in a slightly 
+different context.
+While transit surveys are ideally signal to noise limited, if the noise 
+properties of the survey are dominated by photon counting noise, a signal to 
+noise limit becomes equivalent to a magnitude limit.
+Thus stars surveyed in transit surveys must also over-represent binaries 
+compared to those in a volume-limited sample.
+
+A more straight-forward effect of binarity on transit surveys is that any star 
+which is unknowingly a double introduces extra stars about which planets might 
+be detectable.
+However, this apparent increase in the number of searchable stars is mitigated 
+by dilution: the observed transit depths in such systems are reduced by 
+contaminating flux from the binary companion. This dilution is especially 
+drastic if the planet orbits the secondary, in which case dilution impacts 
+detectability the most.
+
+A related effect to dilution is misinterpretation of transit depths. If the 
+planet to star radius ratio is directly taken as the square-root of 
+the observed transit depth, derived planetary radii in binary systems will 
+systematically too small.
+Ciardi et al [2015] explored this issue in depth. They showed that for the case 
+of \kepler, assuming host stars to always be single will on 
+average underestimate radii by a multiplicative factor of $\approx 1.5$.
+
+Finally, binarity may affect transit surveys through actual astrophysics.
+For the relatively rare case [Duquennoy \& Mayor 1991] of close 
+binaries, dynamical stability imposes a limited range of periods in which 
+planets may orbit [e.g., Holman \& Weigert 
+1999, Li et al 2017], the effects of which have been observed [Kraus et al 
+2016].
+Also, the formation rates of planets, and thus their intrinsic occurrence 
+rates, may be influenced by properties of circumprimary and/or circumbinary 
+disks.
+Although the observations are still ongoing, early results seem to indicate 
+that once binary companions are $\gtrsim 100\,{\rm AU}$ apart, binarity does 
+not decrease planet occurrence rates [CITE], and may even increase them for 
+giant planets [Ngo et al 2017].
 
@@
 
-Ah, transiting planets! We learn so much by studying them. But how much can we 
-actually learn, and how much is messed up by binarity?
-
-Imagine the following transit survey:
+Binarity introduces several convoluted effects on transit 
+surveys. To disentangle them, imagine the following idealized survey:
+
 \begin{enumerate}
@@
 The wavelength-specific photon flux density $F_\lambda$ [${\rm 
-ph\,cm^{-2}\,s^{-1}\,\AA^{-1}}$] could be from Pickles' library, or could be 
-a blackbody function. The transmission function is, up to factor of order 
+ph\,cm^{-2}\,s^{-1}\,\AA^{-1}}$] could be from Pickles' library [1998, PASP 
+110, 863], or could be a blackbody function. The transmission function is, up 
+to factor of order 
 unity, a step function between two wavelengths $\lambda_{\rm min}$ and 
@@
 	\end{center}
-	\caption{Empirical fit to main sequence dwarf mass luminosity data from 
-		Torres et al. [2009] and Benedict et al. [2016]. The ``low fit'' is a 
-		least 
-		squares fit to data from $0.1-0.5M_\odot$, and the ``high fit'' is to 
-		data 
-		above that, and below the Kraft break ($1.2M_\odot$).
+	\caption{Empirical fit to main sequence dwarf mass luminosity data compiled 
+	from Torres et al. [2009] and Benedict et al. [2016]. The ``low fit'' is a 
+	least squares fit to data from $0.1-0.5M_\odot$, and the ``high fit'' is to 
+	data above that, and below the Kraft break ($1.2M_\odot$).
 	The $L(M)$ relation taken for subsequent numerics is the maximum 
@@
 planets of radius $R_p$?}
+\label{subsubsec:model2_A}
 
@@
 For Model \#2, since all planets have radius $R_p$, we only care about 
-the smaller radius case, \textit{i.e.}, we want to count $R_p < R_p^{\rm 
-	obs} < (1 - x) R_p$, for instance with $x=0.1$.
+the smaller radius case, \textit{i.e.}, we want to count $R_p > R_p^{\rm 
+	obs} > (1 - x) R_p$, for instance with $x=0.1$.
 
@@
 
-Then writing $N_{d1}(R_p')$ is relatively straight-forward:
+Then writing $N_{d1}(R_p')$, the number of double systems for which a $(R_p, 
+P)$ planet would be observed in $R_p'$ if it orbiting the primary, is 
+relatively straight-forward:
 \begin{equation}
@@
 
-The analogous equation for secondaries is
+The analogous equation for $N_{d1}(R_p')$, the number of double systems for 
+which a $(R_p, P)$ planet would be observed in $R_p'$ if it orbiting the 
+secondary, is
 \begin{equation}
@@
 
-%FIXME
+\section{Model \# 3: A Synthetic {\it Kepler} Analog}
+
+If we assume every KIC star to be single, what order of magnitude of error do 
+we make in occurrence rates estimated for planets of different sizes (and 
+periods)?
+
+We can attempt an answer with Monte Carlo simulations of something like the 
+\kepler\ field. I say ``something like'' because the binarity properties of the 
+\kepler\ field (or more specifically the stars selected in the KIC) are not 
+known.
+\todo[inline]{but it might be worth doing something directly on the KIC's 
+selected target stars, like what Ciardi did! The following is a good example, 
+but at most it can ``suggest'' that a similar level of error has been made from 
+the \kepler\ survey.}
+
+The plan is as follows:
+\begin{itemize}
+	\item Verify Galaxia can produce something resembling the KIC.
+	\item Introduce binaries (Galaxia does not, by default, have any).
+	\item Use Galaxia [Sharma et al 2010] to construct an analog of the KIC.
+	\item Introduce a planet population.
+	\item Numerically assess what occurrence rates would be derived, and 
+	compare to previous sections.
+\end{itemize}
+
+\subsection{Verifying Galaxia can produce a KIC analog }
+
+Galaxia [Sharma et al 2010] is a stellar population synthesis code for creating 
+surveys of the Milky Way. It samples an assumed analytic distribution function. 
+In other words, it assumes a number density of stars as a function of position, 
+velocity, age, metallicty, and mass, from which it then samples to construct a 
+``realistic'' stellar population.
+The assumed model includes:
+\begin{itemize}
+	\item A metallicity distribution (log-normal),
+	\item A velocity distribution (triaxial Gaussian),
+	\item The stellar density and galactic components specified by Robin et al 
+	[2003] (the Besan\c{c}on model) -- see Sharma et al [2010] Table 1. This 
+	includes the IMF for a young and old thin disk, a thick disk, a spheroidal 
+	component, a bulge component, as well as an ISM and dark halo. It also 
+	includes the star-formation rate.
+\end{itemize}
+
+While this model of course lacks some detail, Sharma et al [2010] show that 
+agreement with the volume-limited Hipparcos sample is probably good enough for 
+our purpose of estimating the number of ``solar-like'' (by definition 
+hereafter, $0.7 M_\odot < M_\star < 1.3 M_\odot$) stars in something like the 
+\kepler\ field.
+
+We then run the Galaxia in its ``circular survey'' mode, towards the direction 
+of the \kepler\ field\footnote{galactic longitude 76.532562, galactic latitude 
+13.289502.}. We apply the Schlegel, Finkbeiner et al [1999] extinction map when 
+converting absolute to apparent magnitudes.
+
+To verify the results are in reasonable agreement with the stars that actually 
+exist in that direction, we compare the output with the KIC\footnote{From 
+\url{http://archive.stsci.edu/kepler/kic.html}, we 
+downloaded the 13.1 million row "|"-delimited gzipped ASCII file containing the 
+complete Kepler Input Catalog (version 10).} [Brown et al 2011].
+Following Sharma et al [2016, ApJ 822:15], for each catalog we selected all 
+stars within $7.5^\circ$ of the center of the \kepler\ field, and with Sloan $r 
+< 14$.
+The KIC is complete to magnitudes fainter than $r=14$, so the comparison should 
+not be affected by its completeness.
+
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=.8]{figures/rmag_distribn.pdf}
+	\end{center}
+	\caption{Probably density distributions of $r$ magnitude for $r<14$ stars 
+	within $7.5^\circ$ of the center of the \kepler\ field, for Galaxia and the 
+	KIC.	}
+	\label{fig:rmag_distribn}
+\end{figure}
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=.8]{figures/rstar_more_distribn.pdf}
+	\end{center}
+	\caption{Same as Fig.~\ref{fig:rmag_distribn}, for $R_\star$.}
+	\label{fig:rstar_more_distribn}
+\end{figure}
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=.8]{figures/teff_distribn.pdf}
+	\end{center}
+	\caption{Same as Fig.~\ref{fig:rmag_distribn}, for $T_{\rm eff}$.}
+	\label{fig:teff_distribn}
+\end{figure}
+
+A few distributions are shown in 
+Figs.~\ref{fig:rmag_distribn}-~\ref{fig:teff_distribn}.
+From Fig.~\ref{fig:rmag_distribn}, we can see that the stellar counts as a 
+function of Sloan $r$ magnitude are in close agreement with the KIC.
+From Fig.~\ref{fig:rstar_more_distribn}, the KIC reports relatively more dwarf 
+stars, and fewer giants than Galaxia does.
+This is likely related to KIC's inability to distinguish between dwarfs and 
+giants with only limited photometry -- although Brown et al [2011]'s idea of 
+using a specific ``D51'' photometric band as a $\log g$ diagnostic was reported 
+to have helped with this problem.
+In Fig.~\ref{fig:teff_distribn}, we see something resembling the K dwarf desert 
+again (5000K $\approx$ K3V dwarf).
+Although there are a few slightly worrying discrepancies (notably in 
+Fig.~\ref{fig:rstar_more_distribn}), we get the idea that Galaxia at least 
+roughly resembles ``reality'' (inasmuch as the reported KIC parameters 
+represent reality!).
+
+\subsection{Introduce binaries}
+\label{subsec:introduce_binaries}
+
+By default, Galaxia does not include binaries.
+However, from Fig.~\ref{fig:rmag_distribn}, we've seen that its number counts 
+as a function of apparent magnitude are good.
+So we assign binarity to the Galaxia stars as follows.
+\begin{itemize}
+	\item Assume a binary fraction ${\rm BF} = 0.45$ [Raghavan et al 2010].
+	\item Ignore higher order multiples.
+	\item If a star is drawn to be a binary:
+	\subitem Its reported magnitude becomes a system magnitude.
+	\subitem Draw the binary's mass ratio $q$ from the probability distribution 
+	function for a magnitude limited survey, i.e. from 
+	Fig.~\ref{fig:q_distribn_mag_limited}. This is sensible so long as the pdf 
+	is independent of the primary's mass. We assume this to be the case over 
+	$0.7-1.3M_\odot$\footnote{It was not immediately self-evident to me that 
+	this distribution was the correct choice, since for fixed mass ratio as the 
+	primary mass changes so does the detectable volume. However, as long as 
+	${\rm prob}(q)$ is independent of the primary's mass, the average 
+	distribution over all the possible detectable volumes will remain the 
+	same.}.
+	\subitem For $q \geq 5/7$, it turns out the above is sufficient information 
+	to analytically specify the individual stellar masses, and thus 
+	luminosities with the relation shown in Fig.~\ref{fig:mass_luminosity}.
+	\subitem For $q < 5/7$, the individual stellar masses are not uniquely 
+	specified\footnote{See the derivation of 2017/08/25. This is because of the 
+	mass luminosity relation we specified, and the fact that we have $q$, and 
+	$L_d$, but want individual masses, and individual luminosities.}. Thus for 
+	this latter fraction of the population, we draw the 
+	primary mass from the pdf of single star primary masses.
+	\subitem With the individual masses known, use Eq.~\ref{eq:mass_radius} to 
+	assign a radius to each star in the binary system.
+\end{itemize}
+
+Benefits of the above procedure include that it does not change Galaxia's 
+star counts as a function of magnitude.
+It produces the correct distribution of the mass ratio for the binaries in a 
+magnitude limited sample, and thus a reasonable light ratio distribution.
+
+The drawbacks are that the individual component masses this procedure produces 
+are not drawn from the same distribution throughout.
+This might bias any subsequent synthetic transit survey.
+From Fig.~\ref{fig:galaxia_scatter_masses}, it's likely not excessively 
+important (it's more important to get the mass ratio and light ratio 
+distributions correct).
+
+In addition, the procedure produces a different mass-radius relation 
+for stars in binary systems, vs. stars in single star systems.
+This might also bias any subsequent synthetic transit survey.
+The mass-radius relation for single stars comes from the Padova isochrones 
+[Marigo et al 2008, Marigo \& Girardi 2007, Girardi et al. 2000, Bertelli et al 
+1994] used in Galaxia.
+Although Galaxia does not directly report radii, it reports luminosities and 
+effective temperatures, from which we compute the radii.
+The alternative choice would be to use the reported Galaxia masses with our 
+Eq.~\ref{eq:mass_radius} to produce radius estimates.
+This would produce consistency in the mass-radius relation, but would break $L 
+= 4\pi R^2 \sigma T_{\rm eff}$ for single star systems.
+The different mass-radius relations are shown in 
+Fig.~\ref{fig:galaxia_mass_radius}.
+Generally, it seems that the at a given mass, the stars in binary systems are 
+biased to have slightly greater radii. (Ignoring the evolved stars).
+However this is a small effect -- perhaps of order 10\% in the radius.
+This will affect the transit depths in a synthetic transit survey, making it 
+more difficult to detect planets around binaries.
+
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=.8]{figures/galaxia_gammaR_distribn_mag_limited.pdf}
+	\end{center}
+	\caption{Light ratio distribution for the $r<17$ Galaxia sample discussed 
+	in Sec.~\ref{subsec:introduce_binaries}.}
+	\label{fig:galaxia_gammar_distribn}
+\end{figure}
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=.8]{figures/galaxia_q_distribn_mag_limited.pdf}
+	\end{center}
+	\caption{Mass ratio distribution for the $r<17$ Galaxia sample discussed 
+		in Sec.~\ref{subsec:introduce_binaries}.}
+	\label{fig:galaxia_q_distribn}
+\end{figure}
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=.8]{figures/galaxia_gammaR_scatter_masses.pdf}
+	\end{center}
+	\caption{Scatter plot of primary and secondary mass (solar units).}
+	\label{fig:galaxia_scatter_masses}
+\end{figure}
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=.8]{figures/galaxia_gammaR_scatter_luminosities.pdf}
+	\end{center}
+	\caption{Scatter plot of primary and secondary luminosity (solar units).}
+	\label{fig:galaxia_scatter_luminosities}
+\end{figure}
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=.8]{figures/galaxia_compare_mass_radius.pdf}
+	\end{center}
+	\caption{Scatter plot of mass vs radius for singles and binaries. For 
+	binaries, different points are shown for the primary and secondary of a 
+	given system. Only points with $0.7<M_\star/M_\odot<1.3$ are selected.}
+	\label{fig:galaxia_mass_radius}
+\end{figure}
+
+\subsection{Using Galaxia to construct a KIC analog}
+
+With glowing confidence, we then proceed to construct a KIC analog as follows.
+We take all Galaxia stars within 7.5 degrees of the center of the \kepler\ 
+field, and take stars with $r<17$.
+\todo[inline]{a different magnitude cut, e.g. $r<16$, might matter?}
+This is now beyond where the KIC is complete, so direct comparisons of 
+distributions with the KIC will be confused.
+We then apply an analog of the prioritization scheme used by the \kepler 
+mission to select target stars, described by Batalha et al [2010]'s Table 1.
+
+For each star, we compute the minimum detectable planet radius at three 
+semimajor axes: (1) the inner radius of the so-called ``habitable'' zone,
+\begin{equation}
+a_{\rm HZ} = 0.95\, {\rm AU} \left(\frac{L}{L_\odot}\right)^{1/2},
+\end{equation}
+(2) $0.5 a_{\rm HZ}$, and (3) $5R_\star$. We make the same assumptions as 
+Batalha et al [2010] did about the transit duration. We assume a noise model of 
+pure shot noise, rather than using any CDPP estimates.
+We then apply the Batalha et al [2010] Table 1 prioritization, except 
+that we use SDSS $r$ magnitudes instead of Kepler $Kp$ magnitudes.
+The rankings are shown in Fig.~\ref{fig:batalha_table_1}.
+We get similar (within a factor of two) counts as they did towards the end of 
+the table, though at the beginning we seem to have fewer stars\footnote{I 
+suspect this is because of the slightly different noise calculation -- and 
+perhaps Batalha using $N_{\rm sample}$ rather than just counting photons 
+matters}.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=0.3]{figures/batalha_2010_table_1.png}
+	\end{center}
+	\caption{Prioritization applied by Batalha et al [2010], from whom I 
+	copy-pasted this Table. I did the same, but using $r$ magnitudes instead of 
+	$Kp$.}
+	\label{fig:batalha_table_1}
+\end{figure}
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=0.8]{figures/KIC_vs_GIC_teff_distribn.pdf}
+	\end{center}
+	\caption{Effective temperature distribution of the stars in our synthetic 
+		Kepler input catalog analog -- the Galaxia input catalog.
+		Note that the effective temperatures of binary star systems in the GIC 
+		are 
+		intentionally wrongly estimated, as if they were single star systems.
+		The listed $T_{\rm eff}$ to spectral type correspondence is from 
+		Mamajek's tables.
+	}
+	\label{fig:KIC_vs_GIC_teff_distribn}
+\end{figure}
+
+Note that, since we assigned binarity to Galaxia stars, applying the Batalha et 
+al [2010] prioritization procedure requires estimating \textit{incorrect} 
+single star parameters for what are really binary star systems!
+We do this by keeping the total system luminosity, inverting 
+Fig.~\ref{fig:mass_luminosity} to find an incorrect mass, and then applying 
+Eq.~\ref{eq:mass_radius} to get an incorrect radius.
+We can then calculate the minimum detectable planet radius as
+\begin{equation}
+R_{p,{\rm min}} = \left(\frac{7.1 \sigma_{\rm tot}}{r}\right)^{1/2} R_\star,
+\label{eq:min_det_radius}
+\end{equation}
+for $r=1$ the dilution that we ignore, and
+\begin{equation}
+\sigma_{\rm tot} \equiv {\rm N} = (F_\gamma^{\rm N} A N_{\rm tra} T_{\rm 
+dur})^{-1/2}
+\end{equation}
+
+In applying the prioritization, we omit the latter two priority classes of 
+Fig.~\ref{fig:batalha_table_1}, to obtain a catalog of 152682 stellar systems, 
+which we hereafter refer to as the \textit{Galaxia Input Catalog (GIC)}.
+We compare the effective temperature distribution of the GIC with 
+that of the KIC, reported by Batalha et al [2010] Table 2, in 
+Fig.~\ref{fig:KIC_vs_GIC_teff_distribn}. Evidently, the GIC is 
+dominated by sun-like stars, in a very similar manner to the exoplanet target 
+stars of the KIC.
+
+\subsection{Introduce a planet population}
+
+As in Sec.~\ref{sec:model_2}, we allow for three different occurrence rates: 
+$\Gamma_{t,s}$, the fraction of stars in single systems with a planet of radius 
+$R_p$ and orbital period $P$, and also $\Gamma_{t,d1}$ and $\Gamma_{t,d2}$ for 
+the fraction per sun-like primary and secondary (respectively) of double star 
+systems with a planet of $(R_p, P)$.
+
+We then randomly select which stars get a planet, and using the true stellar 
+masses and radii compute the corresponding impact parameters for the planets as
+\begin{equation}
+b = \frac{a}{R_\star} \cos i, \quad \cos i \sim \mathcal{U}(0,1).
+\end{equation}
+Any planet with $|b| < 1$ is transiting, and has its transit duration evaluated 
+as
+\begin{equation}
+T_{\rm dur} = 13\,{\rm hr} \left(\frac{P}{{\rm yr}}\right)^{1/3} 
+\left(\frac{\rho_\star}{\rho_\odot}\right)^{-1/3} \sqrt{1 - b^2},
+\end{equation}
+which assumes circular orbits.
+
+\subsection{Run the survey}
+We run the following survey:
+$A = 0.708\,{\rm m}^2$ (\kepler\ effective area), $T_{\rm obs} = 20\,{\rm 
+years}$ (to detect more Earth-like planets than \kepler\ did), and $x_{\rm min} 
+= 7.1$, the same signal to noise threshold as \kepler.
+
+For simplicity, we assign a $r$ band zero point of $10^6\,{\rm ph/s/cm}^2$ to 
+correspond to an $r=0$. We use this to compute the observed 
+photon number fluxes for every system, regardless of whether it has a planet.
+We then compute the signal to noise distribution of transit events, and the 
+number of detections, according to Eq.~\ref{eq:snr_ivory_tower}.
+
+\subsection{Numerically assess what occurrence rates would be derived, and 
+	compare to previous models}
+
+\subsubsection{If we ignore binarity, how wrong is our occurrence rate for 
+	planets of radius $R_p$?}
+
+This is similar to what we did in Sec.~\ref{subsubsec:model2_A}, with the 
+exception that we need to specify the calculation more precisely.
+
+Specifically, our ``Astronomer A'' in this case will estimate an occurrence 
+rate for planets of radius $R_p$ and orbital period $P$ as
+\begin{equation}
+\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}}{Z},
+\label{eq:rate_A}
+\end{equation}
+for
+\begin{equation}
+Z \approx (N_s + N_d) \frac{1}{J} \sum_{j=1}^{J} Q_j,
+\end{equation}
+for $Q_j$ the total detection efficiency, indexed over single and double 
+systems, given as the product of the geometric transit probability and the 
+system-specific completeness:
+\begin{equation}
+Q_j \equiv f_g^{(j)} f_c^{(j)}.
+\end{equation}
+
+Keep in mind that for Astronomer A, we need to compute the 
+\textit{incorrect} transit probability for double star systems (since 
+Astronomer A is assuming that they are single stars).
+
+To evaluate the system-level completeness, we assume that Astronomer A performs 
+something equivalent to perfect injection-recovery.
+For a stellar system with known properties (stellar and planetary), and a 
+perfectly known noise model, the probability of detection is a binary function: 
+either ${\rm S/N} > {\rm (S/N)_{min}}$, or it is not\footnote{There are 
+probabilities involved if we allow for any uncertainty in the parameters, but 
+this is too complicated for our current goal.}.
+So we compute $f_c^{(j)}$ as Astronomer A would: an array (over star systems) 
+of zeros wherever the $(R_p,P)$ planet could not be detected, and ones where it 
+could.
+We then evaluate the total detection efficiency $Q_j$ for each system, and 
+compute the occurrence rate as in Eq.~\ref{eq:rate_A}.
+The resulting occurrence rates are only slightly smaller than those in 
+Sec.~\ref{subsubsec:close_enough} discussed below and shown in 
+Fig.~\ref{fig:error_vs_occrate}.
+
+\subsubsection{What if we ignore binarity, but count planets whose derived 
+radii are ``close enough'' to $R_p$?}
+\label{subsubsec:close_enough}
+
+As in Sec.~\ref{sec:model_2}, we now count planets with observed radii from 
+$(1-x)R_\oplus < R_p^{\rm obs} < R_\oplus$. We'll use $x=0.1$, but it can be 
+whatever we want.
+The occurrence rate in this case will be
+\begin{equation}
+\Gamma_{A,R_p'} = \frac{N_{\rm det,s} + N_{\rm det,d}(R_p')}{Z'},
+\end{equation}
+for
+\begin{equation}
+Z' \approx (N_s + N_d) \frac{1}{J} \sum_{j=1}^{J} Q_j',
+\end{equation}
+where the geometric transit probability is the same as before, but the 
+system-specific completeness is now a function of the desired radius interval:
+\begin{equation}
+Q_j' = f_g^{(j)} f_c^{(j)}(R_p').
+\end{equation}
+There are three possible cases for the completeness, specified by the minimum 
+detectable planet radius $R_{p,{\rm min}}$ (computed according to 
+Eq.~\ref{eq:min_det_radius}):
+\begin{align}
+f_c^{(j)}(R_p') &= 
+\Bigg\{\begin{array}{lr}
+1 & {\rm\ if\ }R_{p,{\rm min}}>R_p,\\
+0, & {\rm\ if\ }R_{p,{\rm min}}<(1-x)R_p,\\
+\frac{R_p - R_{p,{\rm min}}}{R_p - (1-x)R_p}, &{\rm otherwise}.
+\end{array}
+\label{eq:completeness_close_enough}
+\end{align}
+The latter term above is the case in which the minimum detectable planet 
+radius happens to be between $(1-x)R_p$ and $R_p$. In that case the 
+completeness becomes the fraction of planets in that interval that would be 
+detected.
+Assuming a uniform prior for the planet radius distribution $R_p \sim 
+\mathcal{U}((1-x)R_p, R_p)$, this becomes the expression given in 
+Eq.~\ref{eq:completeness_close_enough}.
+
+These occurrence rates can then be compared to the ``true occurrence rate'' 
+derived by the all-knowing Astronomer D:
+\begin{equation}
+\Gamma_{{\rm D}, R_p} = \frac{\Gamma_{t,s} N_s + (\Gamma_{t,d1} + 
+						\Gamma_{t,d2})N_d}{N_s + 2 N_d}.
+\end{equation}
+A note in passing: the more information-rich parameters for Astronomer D to 
+report would be $(\Gamma_{t,s}, \Gamma_{t,d1}, \Gamma_{t,d2})$, rather than 
+$\Gamma_{{\rm D}, R_p}$. But I digress.
+
+Similar to Fig.~\ref{fig:deltaprime_vs_xi}, Fig.~\ref{fig:error_vs_occrate} 
+shows that the magnitude of the relative error 
+made by the $\Gamma_{{\rm A},R_p'}$ and $\Gamma_{{\rm A},R_p}$ estimates 
+depends strongly 
+on the fraction of secondaries that are assumed to host the $(R_p, P)$ planets.
+The simplest assumption is that the occurrence rate is the same for all solar 
+type stars: any ``Sun-like'' ($0.7<M_\star/M_\odot<1.3$) star hosts planets at 
+the ``true rate'' $\Gamma_t$, and any other mass of star does not.
+Given how we have defined the GIC, this means that all single stars and 
+primaries of double star systems are sun-like. The fraction of Sun-like 
+secondaries in this case is 37.4\% of all secondaries.
+Reading the appropriate value off Fig.~\ref{fig:error_vs_occrate}, this 
+corresponds to $\Gamma_{\rm D} = 1.39 \Gamma_{\rm A'}$.
+
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=0.8]{figures/error_vs_occrate.pdf}
+	\end{center}
+	\caption{ The ``most reasonable'' $\Gamma_{t,d2}$ value is obtained by 
+	assuming that the occurrence rate of $(R_p, P)$ planets is the same for all 
+	solar type stars, and zero otherwise.
+	This applies most immediately to the question of Earth-like planets around 
+	Sun-like stars.
+	}
+	\label{fig:error_vs_occrate}
+\end{figure}
 

### 2017-08-30 src/binary_distribution.py

@@ -1,0 +1,627 @@
+'''
+what is the light ratio distriubtion for binaries in a magnitude limited
+sample, vs a volume limited sample?
+'''
+from __future__ import division, print_function
+
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+from astropy import units as u, constants as c
+from math import pi as π
+
+def get_L(M):
+
+    m_lo = 1.8818719873988132
+    c_lo = -0.9799647314108376
+    m_hi = 5.1540712426599882
+    c_hi = 0.0127626185389781
+    M_merge = 0.4972991257826812
+    L_merge = 0.0281260412126928
+
+    L = np.ones_like(M)
+
+    # Method 1: totally fine.
+    #lo_mask = M < M_merge
+    #L[lo_mask] = 10**(np.log10(M[lo_mask])*m_lo + c_lo)
+
+    #hi_mask = M >= M_merge
+    #L[hi_mask] = 10**(np.log10(M[hi_mask])*m_hi + c_hi)
+
+    # Method 2: cuter.
+    L_lo = 10**(np.log10(M)*m_lo + c_lo)
+    L_hi = 10**(np.log10(M)*m_hi + c_hi)
+
+    L = np.maximum(L_lo, L_hi)
+
+    return L
+
+
+def _make_distribution_plots(vl, doubles):
+
+    savedir = '../results/analytic_model_plots/'
+    # Plot distribution of doubles. First do volume limited case. 
+    plt.close('all')
+    f,ax = plt.subplots(figsize=(4,4))
+    hist, bin_edges = np.histogram(vl['q'], bins=np.append(np.linspace(0,1,101),42),
+            normed=True)
+    ax.step(bin_edges[:-1], hist, 'k-', where='post',)
+    ax.set(xlabel='$q = M_2/M_1$', ylabel='prob')
+    ax.set_title(txt_vl, fontsize='small')
+    f.tight_layout()
+    f.savefig(savedir+'q_distribn_vol_limited.pdf', dpi=250, bbox_inches='tight')
+
+    plt.close('all')
+    f,ax = plt.subplots(figsize=(4,4))
+    hist, bin_edges = np.histogram(vl['γ_R'],
+            bins=np.append(np.linspace(0,1,501),42), normed=True)
+    ax.step(bin_edges[:-1], hist, 'k-', where='post')
+    ax.set(xlabel='$\gamma_R = L_2/L_1$',
+           ylabel='prob')
+    ax.set_title(txt_vl, fontsize='small')
[…]

### 2017-08-30 src/compare_mass_radius_relations.py

@@ -1,0 +1,46 @@
+'''
+Mon 28 Aug 2017 08:32:38 AM EDT
+
+Current mass-radius relation for stars of binaries is different for singles.
+Show how different.
+'''
+
+import numpy as np, pandas as pd
+import matplotlib.pyplot as plt
+import pickle
+
+gic = pickle.load(
+        open('/home/luke/local/selected_GIC_systems_kepler_analog.p','rb'))
+
+singles = gic['singles']
+doubles = gic['binaries']
+
+R_s = singles['Rstar']
+M_s = singles['mact']
+
+R_d1 = doubles['R_1']
+M_d1 = doubles['M_1']
+
+R_d2 = doubles['R_2'][ (doubles['M_2'] > 0.7) & (doubles['M_2'] < 1.3) ]
+M_d2 = doubles['M_2'][ (doubles['M_2'] > 0.7) & (doubles['M_2'] < 1.3) ]
+
+R_d = np.concatenate((R_d1, R_d2))
+M_d = np.concatenate((M_d1, M_d2))
+
+plt.close('all')
+f,ax = plt.subplots(figsize=(4,4))
+
+ax.scatter(R_d, M_d, s=1, c='black', alpha=1, lw=0, rasterized=True,
+        label='doubles', zorder=0)
+ax.scatter(R_s, M_s, s=1, c='blue', alpha=0.7, lw=0, rasterized=True,
+        label='singles', zorder=-2)
+
+ax.set(xlabel='radius [$R_\odot$]', ylabel='mass [$M_\odot$]')
+
+ax.legend(loc='best', fontsize='x-small')
+
+f.tight_layout()
+
+f.savefig('galaxia_compare_mass_radius.pdf',
+        dpi=250, bbox_inches='tight')
+

### 2017-08-30 src/convert_galaxia_output.py

@@ -1,0 +1,37 @@
+import pickle
+
+global DATADIR
+DATADIR = '/home/luke/local/GalaxiaData/kepler_synthesis/'
+
+def convert_galaxia_output():
+    '''
+    Convert from ebf to pickle (latter is more generally usable, e.g. not in
+    python 2.X)
+    '''
+    try:
+        import ebf
+    except ImportError:
+        print('ImportError: ebf requires python 2.X, and manual installation.')
+
+    # Keys described http://galaxia.sourceforge.net/Galaxia3pub.html#mozTocId124318
+    # ['rad', 'exbv_solar', 'teff', 'sdss_g', 'mag2', 'mag1', 'mag0', 'sdss_r',
+    # 'sdss_u', 'satid', 'vx', 'vy', 'vz', 'sdss_z', 'mtip', 'log', 'pz', 'px',
+    # 'py', 'feh', 'exbv_schlegel', 'sdss_i', 'lum', 'exbv_schlegel_inf', 'mact',
+    # 'glon', 'popid', 'glat', 'alpha', 'center', 'partid', 'age', 'grav', 'smass',
+    # 'fieldid']
+
+    datapath = DATADIR+'keplersynthesis.ebf'
+
+    data = ebf.read(datapath)
+
+    outd = {}
+
+    for key in data.keys():
+
+        outd[key] = data[key]
+
+    pickle.dump(outd, open(DATADIR+'keplersynthesis.p', 'wb'))
+
+if __name__ == '__main__':
+
+    convert_galaxia_output()

### 2017-08-30 src/deltaprime_vs_xi.py

@@ -1,0 +1,46 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+Nd_by_Ns = 0.59
+
+ξ = np.arange(0, Nd_by_Ns, 1e-3)
+
+f,ax = plt.subplots(nrows=1, ncols=1, figsize=(4,4))
+
+χ = 2 # greek chi
+β = 0.59
+Δprime = abs(
+        1 - ( (1 + χ*β)/(1+ξ) )*( (1+β)/(1+2*β)  )
+        )
+
+ax.plot(ξ, Δprime, 'k-', label=r'$\chi = 2, \beta = 0.59$')
+
+χ = 1 # greek chi
+β = 0.59
+Δprime = abs(
+        1 - ( (1 + χ*β)/(1+ξ) )*( (1+β)/(1+2*β)  )
+        )
+
+ax.plot(ξ, Δprime, 'b-', label=r'$\chi = 1, \beta = 0.59$')
+
+
+χ = 1.5 # greek chi
+β = 0.59
+Δprime = abs(
+        1 - ( (1 + χ*β)/(1+ξ) )*( (1+β)/(1+2*β)  )
+        )
+
+ax.plot(ξ, Δprime, 'g-', label=r'$\chi = 1.5, \beta = 0.59$')
+
+
+ax.legend(loc='best', fontsize='small', frameon=False)
+
+ax.set_xlim([0,β])
+ax.set_ylim([0,0.6])
+
+ax.set_ylabel("$\Delta'$ (relative error btwn A' and D)")
+ax.set_xlabel(r"$\xi$ (max is upper bound of $\beta$)")
+
+f.tight_layout()
+
+f.savefig('deltaprime_vs_xi.pdf', dpi=300, bbox_inches='tight')

### 2017-08-30 src/kepler_numerics.py

@@ -1,0 +1,1414 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+'''
+>  Can we ignore binarity when mapping KIC stars onto synthetic local stars?
+
+Or better yet
+
+> If we assume every KIC star is single, what OOM of error do we make in
+> occurrence rates estimated for planets of different sizes (& periods)?
+
+Answer the latter question with Monte Carlo simulations of the Kepler field.
+
+with TRILEGAL/Galaxia:
+    * construct the Kepler field (w/ known binarity properties, how?)
+    * introduce various planet populations
+    * numerically assess the error.
+    * Compare w/ analytic results.
+'''
+
+from __future__ import print_function, division
+import sys, os
+import pickle
+import numpy as np, pandas as pd
+
+from astropy import units as u, constants as c
+from astropy.coordinates import SkyCoord
+
+global DATADIR
+DATADIR = '/home/luke/local/GalaxiaData/kepler_synthesis/'
+
+def T_dur(P, ρ_star, b):
+    '''
+    Compute the transit duration given orbital period, host star density, and
+    impact parameter.
+    E.g., Sullivan et al 2015, Eq 10. (Assumes zero eccentricity).
+    '''
+
+    from math import pi as π
+    ρ_sun = 3*u.Msun/(4*π*u.Rsun**3)
+    return 13*u.hr * (P/(1*u.year))**(1/3) * \
+                     (ρ_star/ρ_sun)**(-1/3) * \
+                     np.sqrt(1-b*b)
+
+
+def are_synthetic_stars_on_silicon():
+    '''
+    Galaxia model was run with 200 sq deg FoV, centered at the true Kepler
+    field center. Select synthetic Kepler field stars by comparing coordinates
+    to viable KIC stars.
+    '''
+    import K2fov
+    # CITE: http://adsabs.harvard.edu/abs/2016ascl.soft01009M, Mullaly,
+    # Barclay, Barentsen 2017.
+
+    gal = pickle.load(
+            open(DATADIR+'keplersynthesis.p', 'rb'), encoding='latin1')
+
+    c = SkyCoord(frame='galactic', l=gal['glon']*u.deg, b=gal['glat']*u.deg)
+
+    df = pd.DataFrame(np.transpose([np.array(c.fk5.ra), np.array(c.fk5.dec)]),
+                      columns=['ra','dec'])
+
+    df['mags'] = np.ones_like(gal['glon'])
+
[…]

### 2017-08-30 src/mass_luminosity.py

@@ -1,0 +1,146 @@
+'''
+we want a good mass-luminosity relation for dwarfs with masses less than 2
+Msun.
+
+this script takes reported masses and luminosities from a few sources, and then
+derives a broken power law fit.
+'''
+from astropy.table import Table
+import astropy.units as u
+import pandas as pd, numpy as np, matplotlib.pyplot as plt
+
+##############
+# Wrangling. #
+##############
+# Collect M_v and masses from Benedict+ (2016).
+tab10 = pd.read_csv('../data/Benedict_et_al_2016_table10_proc.txt',
+                    delimiter=',')
+tab11 = pd.read_csv('../data/Benedict_et_al_2016_table11_proc.txt',
+                    delimiter=',')
+
+_0 = pd.DataFrame(np.array(tab10[['Mv_pri', 'M_pri']]), columns=['Mv', 'M'])
+_1 = pd.DataFrame(np.array(tab10[['Mv_sec', 'M_sec']]), columns=['Mv', 'M'])
+b16 = pd.concat([_0, _1])
+
+# Load in M_v to logL reported by E. Mamajek in
+# http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt,
+# downloaded 2017.08.02
+mmj = pd.read_csv('../data/Mamajek_Mv_logL.txt', delimiter=',')
+
+from scipy.interpolate import interp1d
+f_mmj = interp1d(mmj['Mv'], mmj['logL'])
+
+b16['logL'] = f_mmj(b16['Mv'])
+
+# Load in Torres+ 2009: highest quality binary data in existence.
+tab = Table.read('../data/Torres_Andersen_Gimenez_2009_table1.vot',
+                 format='votable')
+df = tab.to_pandas()
+# Select things that at least resemble dwarfs. Some of these will be "peculiar"
+# or have extra spectroscopic letters attached to them. That's OK.
+df = df[ [('V' in spt.decode('ascii')) for spt in df['SpType'] ] ]
+
+############
+# Fitting. #
+############
+from scipy.optimize import curve_fit
+
+def line_func(x, m, c):
[…]

### 2017-08-30 src/method_2.py

@@ -1,0 +1,617 @@
+'''
+numerical verification of "method 2" as described in
+memos/toy_analytic_surveys_170804.pdf
+'''
+from __future__ import division, print_function
+
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+from astropy import units as u, constants as c
+from math import pi as π
+
+def get_L(M):
+
+    m_lo = 1.8818719873988132
+    c_lo = -0.9799647314108376
+    m_hi = 5.1540712426599882
+    c_hi = 0.0127626185389781
+    M_merge = 0.4972991257826812
+    L_merge = 0.0281260412126928
+
+    L = np.ones_like(M)
+
+    # Method 1: totally fine.
+    #lo_mask = M < M_merge
+    #L[lo_mask] = 10**(np.log10(M[lo_mask])*m_lo + c_lo)
+
+    #hi_mask = M >= M_merge
+    #L[hi_mask] = 10**(np.log10(M[hi_mask])*m_hi + c_hi)
+
+    # Method 2: cuter.
+    L_lo = 10**(np.log10(M)*m_lo + c_lo)
+    L_hi = 10**(np.log10(M)*m_hi + c_hi)
+
+    L = np.maximum(L_lo, L_hi)
+
+    return L
+
+
+def _make_distribution_plots(vl, doubles):
+
+    # Plot distribution of doubles. First do volume limited case. 
+    plt.close('all')
+    f,ax = plt.subplots(figsize=(4,4))
+    hist, bin_edges = np.histogram(vl['q'], bins=np.append(np.linspace(0,1,51),42),
+            normed=True)
+    ax.step(bin_edges[:-1], hist, where='post')
+    ax.set(xlabel='q', ylabel='prob')
+    ax.set_title(txt_vl, fontsize='small')
+    f.tight_layout()
+    f.savefig('q_distribn_vol_limited.pdf', dpi=250, bbox_inches='tight')
+
+    plt.close('all')
+    f,ax = plt.subplots(figsize=(4,4))
+    hist, bin_edges = np.histogram(vl['γ_R'],
+            bins=np.append(np.linspace(0,1,100),42), normed=True)
+    ax.step(bin_edges[:-1], hist, where='post')
+    ax.set(xlabel='L2/L1 ($\gamma_R$)',
+           ylabel='prob')
+    ax.set_title(txt_vl, fontsize='small')
+    f.tight_layout()
[…]

### 2017-08-30 src/observed_radii_vs_gammaR.py

@@ -1,0 +1,41 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+γ_R = np.arange(0, 1, 1e-3)
+
+dil_pri = (1 + γ_R)**(-1)
+dil_sec = γ_R * (1 + γ_R)**(-1)
+
+Rpobs_by_Rptrue_pri = np.sqrt(dil_pri)
+Rpobs_by_Rptrue_sec = np.sqrt(dil_sec)
+
+f,axs = plt.subplots(nrows=2, ncols=1, figsize=(4,6), sharex=True)
+
+# Top plot: dilution vs gamma_R
+ax = axs[0]
+ax.plot(γ_R, dil_pri, 'k-', label='planet orbits primary')
+ax.plot(γ_R, dil_sec, 'b-', label='planet orbits secondary')
+ax.hlines(0.5, 0, 1, colors='k', linestyles='dashed', alpha=0.5)
+
+ax.legend(loc='best', fontsize='xx-small', frameon=False)
+
+ax.set_xlim([0,1])
+ax.set_ylim([0,1])
+
+ax.set_ylabel('$\mathcal{D}$')
+
+# Bottom plot: radius ratio vs gamma_R
+ax = axs[1]
+ax.plot(γ_R, Rpobs_by_Rptrue_pri, 'k-')
+ax.plot(γ_R, Rpobs_by_Rptrue_sec, 'b-')
+ax.hlines(1/2**(1/2), 0, 1, colors='k', linestyles='dashed', alpha=0.5)
+
+ax.set_xlim([0,1])
+ax.set_ylim([0,1])
+
+ax.set_xlabel('$\gamma_R = L_2/L_1$')
+ax.set_ylabel('$R_p^\mathrm{obs} / R_p^\mathrm{true} = \sqrt{\mathcal{D}}$')
+
+f.tight_layout()
+
+f.savefig('observed_radii_vs_gammaR.pdf', dpi=300, bbox_inches='tight')

### 2017-08-30 src/plot_Batalha_2010_table_2.py

@@ -1,0 +1,68 @@
+import numpy as np, matplotlib.pyplot as plt, pandas as pd
+from astropy.table import Table
+
+# Presume bins are in 1000K, and the listed Teffs are the centers of the bins
+df = pd.read_csv('Batalha_2010_Table_2.txt')
+
+mmj = pd.read_csv('Mamajek_Teff_sptype.txt', delim_whitespace=True)
+mmj = mmj[(mmj['Teff']<11000)&(mmj['Teff']>3000)]
+
+#############################
+# PLOT 1: Teff distribution #
+#############################
+plt.close('all')
+f, ax = plt.subplots()
+
+ax.step(df['teff'], df['val'], where='mid')
+for teff, yval, SpT in list(zip(mmj['Teff'][::2],
+                                np.ones_like(mmj['Teff'][::2])*8e4,
+                                mmj['SpT'][::2])):
+
+    t = ax.text(teff, yval, SpT, alpha=0.3, fontsize='xx-small', rotation=270)
+
+ax.set(
+    xlabel='teff',
+    ylabel='number',
+    yscale='log',
+    ylim=[1e2,1e5],
+    xlim=[3e3,11e3]
+    )
+ax.set_title(
+    'table 2 Batalha+ 2010. KIC Teff distribution for planet target stars. '
+    'SpTypes Pecault & Mamajek 2013',
+    fontsize='xx-small'
+    )
+
+f.savefig('Batalha_2010_Table_2_distribution.pdf', dpi=300)
+
+###############################
+# PLOT 2: Teff cumulative sum #
+###############################
+plt.close('all')
+f, ax = plt.subplots()
+
+ax.step(df['teff'],
+        np.cumsum(df['val'])/max(np.cumsum(df['val'])),
+        where='mid')
+
+for teff, yval, SpT in list(zip(mmj['Teff'][::2],
+                                np.ones_like(mmj['Teff'][::2])*0.95,
+                                mmj['SpT'][::2])):
+
+    t = ax.text(teff, yval, SpT, alpha=0.3, fontsize='xx-small', rotation=270)
+
+ax.set(
+    xlabel='teff',
+    ylabel='percent of {:d} stars'.format(int(max(np.cumsum(df['val'])))),
+    yscale='linear',
+    ylim=[0,1],
+    xlim=[3e3,11e3]
+    )
+
+ax.set_title(
[…]

### 2017-08-30 src/plot_KIC_v_GIC_teff_distribn.py

@@ -1,0 +1,71 @@
+'''
+compare your "Galaxia input catalog" teff distribn w/ KIC's from Batalha et al
+2010.
+'''
+import numpy as np, matplotlib.pyplot as plt, pandas as pd
+import pickle
+from astropy.table import Table
+from astropy import units as u, constants as c
+from math import pi as π
+
+# Presume bins are in 1000K, and the listed Teffs are the centers of the bins
+df = pd.read_csv('../Batalha_2010_Table_2.txt')
+
+mmj = pd.read_csv('../Mamajek_Teff_sptype.txt', delim_whitespace=True)
+mmj = mmj[(mmj['Teff']<11000)&(mmj['Teff']>3000)]
+
+gic = pickle.load(
+        open('/home/luke/local/selected_GIC_systems_kepler_analog.p','rb'))
+
+singles = gic['singles']
+doubles = gic['binaries']
+
+Teff_s = np.array(10**singles['teff'])
+
+# Get Teff distribution for doubles (as if they were singles)
+R_d = np.array(doubles['R_as_if_single'])*u.Rsun
+M_d = np.array(doubles['M_as_if_single'])
+
+import binary_distribution as bd
+L_d = bd.get_L(M_d)
+Teff_d = np.array((( L_d*u.Lsun / \
+            (4 * π * R_d**2 * c.sigma_sb))**(1/4)).to(u.K))
+
+gic_teff = np.concatenate((Teff_s, Teff_d))
+
+gic_hist, gic_bin_edges = np.histogram( gic_teff,
+        bins=np.append(np.arange(3e3,11e3,1e3),12e3),normed=False)
+
+#############################
+# PLOT 1: Teff distribution #
+#############################
+plt.close('all')
+f, ax = plt.subplots(figsize=(4,4))
+
+ax.step(df['teff'], df['val'], where='mid',
+        label='KIC (B+2010 Table 2)', lw=1)
+
+ax.step(gic_bin_edges[:-1], gic_hist, where='post',
+        label='GIC', lw=1, zorder=1)
+
+for teff, yval, SpT in list(zip(mmj['Teff'][::2],
+                                np.ones_like(mmj['Teff'][::2])*1.8e5,
+                                mmj['SpT'][::2])):
+
+    t = ax.text(teff, yval, SpT, alpha=0.3, fontsize=5, rotation=270,
[…]

### 2017-08-30 src/simplest_analytic_model.py

@@ -1,0 +1,366 @@
+'''
+verify the numerical expression for the number of detected planets in the
+"simplest analytic model" case (tex'd up 17/07/26) is correct.
+'''
+from __future__ import division, print_function
+
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+from astropy import units as u, constants as c
+from math import pi as π
+
+
+def draw_star_positions(d_max, N_stars, sys_type=None):
+    '''
+    Args:
+
+        d_max: maximum distance out to which stars (of either single or binary
+               class) are selected by mag-limit
+
+        N_stars (int): the number of total stars to draw
+
+        sys_type: 'single' or 'binary' (star systems)
+
+    Returns:
+
+        a pandas dataframe with the positions, and their distance from the
+        origin.
+
+    '''
+
+    assert sys_type == 'single' or sys_type == 'binary'
+    assert type(N_stars) == int
+
+    x_l,y_l,z_l,r_l = [], [], [], []
+
+    while len(r_l) < N_stars:
+        if len(r_l) % 10 == 0:
+            print('{:d}/{:d}'.format(len(r_l), N_stars))
+
+        x = d_max * np.random.rand()
+        y = d_max * np.random.rand()
+        z = d_max * np.random.rand()
+
+        if x**2 + y**2 + z**2 <= d_max**2:
+            x_l.append(x.value)
+            y_l.append(y.value)
+            z_l.append(z.value)
+            r_l.append( ((x**2 + y**2 + z**2)**(1/2)).value )
+
+    x_a, y_a, z_a = np.array(x_l)*u.pc, np.array(y_l)*u.pc, np.array(z_l)*u.pc
+    r_a = np.array(r_l)*u.pc
+
+    return pd.DataFrame({
+        'x': x_a,
+        'y': y_a,
+        'z': z_a,
+        'r': r_a,
+        'sys_type': [sys_type for _ in range(len(r_a))]
+        })
+
+
+if __name__ == '__main__':
+
+    np.random.seed(42)
+
+    ##############
+    # INSTRUMENT #
+    ##############
+    A = 50*u.cm**2
+
+    λ_min = 500*u.nm
[…]

### 2017-08-30 src/simplest_analytic_model.py
simplest_analytic_model: vectorize position draw

@@ -8,15 +8,14 @@ import numpy as np, pandas as pd, matplotlib.pyplot as plt
 from astropy import units as u, constants as c
 from math import pi as π
 
-
-def draw_star_positions(d_max, N_stars, sys_type=None):
+def draw_star_positions(d_max, N_sys, sys_type=None):
     '''
     Args:
 
         d_max: maximum distance out to which stars (of either single or binary
                class) are selected by mag-limit
 
-        N_stars (int): the number of total stars to draw
+        N_sys (int): the number of total stellar systems to draw
 
         sys_type: 'single' or 'binary' (star systems)
 
@@ -28,33 +27,29 @@ def draw_star_positions(d_max, N_stars, sys_type=None):
     '''
 
     assert sys_type == 'single' or sys_type == 'binary'
-    assert type(N_stars) == int
+    assert type(N_sys) == int
 
-    x_l,y_l,z_l,r_l = [], [], [], []
+    # Vectorized draw:
+    # vol sphere / vol cube = π/6 ~= half. Draw 10* the number of systems you
+    # really need. Then cut.
+    _x = d_max * np.random.rand( (10*N_sys) )
+    _y = d_max * np.random.rand( (10*N_sys) )
+    _z = d_max * np.random.rand( (10*N_sys) )
+    _r = np.sqrt(_x**2 + _y**2 + _z**2)
 
-    while len(r_l) < N_stars:
-        if len(r_l) % 10 == 0:
-            print('{:d}/{:d}'.format(len(r_l), N_stars))
+    assert len(_x[_r < d_max]) > N_sys
 
-        x = d_max * np.random.rand()
-        y = d_max * np.random.rand()
-        z = d_max * np.random.rand()
-
-        if x**2 + y**2 + z**2 <= d_max**2:
-            x_l.append(x.value)
-            y_l.append(y.value)
-            z_l.append(z.value)
-            r_l.append( ((x**2 + y**2 + z**2)**(1/2)).value )
-
-    x_a, y_a, z_a = np.array(x_l)*u.pc, np.array(y_l)*u.pc, np.array(z_l)*u.pc
-    r_a = np.array(r_l)*u.pc
+    x = _x[_r < d_max][:N_sys]*u.pc
+    y = _y[_r < d_max][:N_sys]*u.pc
+    z = _z[_r < d_max][:N_sys]*u.pc
+    r = _r[_r < d_max][:N_sys]*u.pc
 
     return pd.DataFrame({
[…]

### 2017-08-30 src/simplest_analytic_model.py
simplest_analytic_model: add print statements

@@ -272,6 +272,33 @@ if __name__ == '__main__':
     print('N detected numerically: {:d}\n N detected analytically: {:d}\n'.
           format(N_det_n, N_det_a))
 
+    # numerical completeness correction fraction for singles
+    f_s_c_n = len( x_s_n[x_s_n > x_min] ) / len(x_s_n)
+    Γ_A_Rp_n = ( N_det_s_n / f_s_c_n ) / (N_s + N_d)
+    print('Γ_A_Rp: {:.3f} (numerical)'.format(Γ_A_Rp_n))
+
+    # analytic completeness correction fraction for singles
+    f_s_c_a = len( x_s_a[x_s_a > x_min] ) / len(x_s_a)
+    Γ_A_Rp_a = ( N_det_s_a / f_s_c_a ) / (N_s + N_d)
+    print('Γ_A_Rp: {:.3f} (analytic)'.format(Γ_A_Rp_a.cgs))
+
+    Γ_D_Rp = Γ_ts
+    print('Γ_D_Rp: {:.3f}'.format(Γ_D_Rp))
+
+    X_Γ_n = Γ_D_Rp / Γ_A_Rp_n
+    β = BF * (1 + γ_R)**(3/2)
+    X_Γ_math = 1 + β
+    print('\nX_Γ_n: {:.3f}'.format(X_Γ_n))
+    print('X_Γ_a: {:.3f} (analytic from math)'.format(X_Γ_math))
+    print('X_Γ_a: {:.3f} (analytic from implementation)'.format(
+            (Γ_D_Rp / Γ_A_Rp_a).cgs))
+
+    Γ_A_math = Γ_D_Rp / X_Γ_math
+    print('\nAnalytic from math, Γ_A_Rp: {:.3f}'.format(Γ_A_math))
+
+    print('\nRelative percentage error: {:.3f}%'.format(
+           100*(Γ_A_math - Γ_D_Rp)/Γ_D_Rp ))
+
     ####################################################
     # COMPARE ANALYTIC AND NUMERICAL SNR DISTRIBUTIONS #
     ####################################################

### 2017-08-30 src/method_2.py
model 2: clean up comments

@@ -103,22 +103,6 @@ def draw_star_positions(d_max, N_sys, sys_type=None):
     assert sys_type == 'single' or sys_type == 'binary'
     assert type(N_sys) == int
 
-    ## Inefficient, non-vectorized draw:
-    #x_l,y_l,z_l,r_l = [], [], [], []
-    #while len(r_l) < N_sys:
-    #    if len(r_l) % 10 == 0:
-    #        print('{:d}/{:d}'.format(len(r_l), N_sys))
-
-    #    x = d_max * np.random.rand()
-    #    y = d_max * np.random.rand()
-    #    z = d_max * np.random.rand()
-
-    #    if x**2 + y**2 + z**2 <= d_max**2:
-    #        x_l.append(x.value)
-    #        y_l.append(y.value)
-    #        z_l.append(z.value)
-    #        r_l.append( ((x**2 + y**2 + z**2)**(1/2)).value )
-
     # Vectorized draw:
     # vol sphere / vol cube = π/6 ~= half. Draw 10* the number of systems you
     # really need. Then cut.

### 2017-08-30 memos/workdir/toy_analytic_surveys.tex
update memo with clearer correction factor.
also remove obnoxious TeX extras from repo

@@
 Finally, binarity may affect transit surveys through actual astrophysics.
-For the relatively rare case [Duquennoy \& Mayor 1991] of close 
-binaries, dynamical stability imposes a limited range of periods in which 
+For the relatively rare case [Duquennoy \& Mayor 1991, Raghavan et al 2010] of 
+close binaries, dynamical stability imposes a limited range of periods in which 
 planets may orbit [e.g., Holman \& Weigert 
@@
 \begin{equation}
-\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d},
-\end{equation}
-and the relative difference between the two occurrence rates is
-
+\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d}.
+\end{equation}
+
+To express ``how wrong'' our occurrence rate is, we define a correction factor 
+$X_\Gamma$ such that the true occurrence rate is the product of the correction 
+factor with the measured occurrence rate: $\Gamma_D \equiv \Gamma_A X_\Gamma$.
+
+The correction factor is then
 \begin{align}
-\left|\frac{\Gamma_{{\rm A},R_p} - \Gamma_{{\rm D},R_p}}
-		   {\Gamma_{{\rm A},R_p}} \right| &=
-\left| 1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}}\right| \\
+X_\Gamma &=
+\frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}} \\
 &=
-\left| 1 - \left(\frac{\Gamma_{t,s}N_s + 2\Gamma_{t,d}N_d}{N_s + 2N_d}  \cdot 
-\frac{N_s + N_d}{\Gamma_{t,s} N_s }\right)\right| \\
+\frac{\Gamma_{t,s}N_s + 2\Gamma_{t,d}N_d}{N_s + 2N_d}  \cdot 
+\frac{N_s + N_d}{\Gamma_{t,s} N_s } \\
 &= 
-\left|
-1 - \frac{(1 + 2 \beta \Gamma_{t,d}/\Gamma_{t,s})(1 + \beta)}{(1+2\beta)}
-\right|,
+\frac{(1 + \beta)(1 + 2 \beta \Gamma_{t,d}/\Gamma_{t,s})}{(1+2\beta)},
 \end{align}
@@
 For the nominal G2V dwarf case of ${\rm BF}=0.45$ with twin binaries ($\gamma_R 
-= 1$) and $\Gamma_{t,d} = \Gamma_{t,s}$ this gives a relative error of $127\%$.
+= 1$) and $\Gamma_{t,d} = \Gamma_{t,s}$ this gives $X_\Gamma = 2.27$.
 For instance, in the numerical simulation corresponding to 
-Fig.~\ref{fig:snr_dist_analytic_v_numeric}, Astronomer A finds $\Gamma_{\rm 
-A,R_p} = 0.22$, while Astronomer 
-D derives the true (input) occurrence rate of $\Gamma_{\rm D,R_p} = 0.5$.
+Fig.~\ref{fig:snr_dist_analytic_v_numeric}, Astronomer A finds 
+$\Gamma_{{\rm A},R_p} = 0.22$, while Astronomer 
+D derives the true (input) occurrence rate of $\Gamma_{{\rm D},R_p} = 0.5$.
+This gives a relative percentage error of $\delta\Gamma_{{\rm D},R_p} \equiv 
+(\Gamma_{\rm A,R_p} - \Gamma_{\rm D,R_p}) / \Gamma_{\rm D,R_p} = -56\%$.
 
@@
 So Astronomer A now misclassifies planetary radii, and miscounts the total 
-number of stars, but knows his completeness for single stars.
-Astronomer D corrects all these errors.
+number of stars, but somehow knows his completeness for single 
+stars. This is possible, because Astronomer A can correctly derive a 
+completeness estimate and geometric transit probability for single stars only. 
+For Astronomer $A'$ below, including the planets in binary star systems in the 
+numerator actually introduces confusion that precludes nice math.
 Then
@@
 
-The relative difference between the two occurrence rates, $\Delta \equiv 
-\left|(\Gamma_{{\rm A},R_p} - \Gamma_{{\rm D},R_p})/\Gamma_{{\rm A},R_p} 
-\right|$, is
+The correction factor $X_\Gamma$ is
 \begin{align}
-	\Delta &=
-	\left| 1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}}\right| \\
+X_\Gamma &=
+\frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}} \\
 &=
-	\left| 1 - \left(\frac{\Gamma_{t,s}N_s + (\Gamma_{t,d1} + 
+\frac{\Gamma_{t,s}N_s + (\Gamma_{t,d1} + 
 	\Gamma_{t,d2})N_d}{N_s + 2N_d}  
 \cdot 
-	\frac{N_s + N_d}{\Gamma_{t,s} N_s }\right)\right| \\
+\frac{N_s + N_d}{\Gamma_{t,s} N_s } \\
 &= 
-	\left|
-	1 - \frac{(1 + \beta (\Gamma_{t,d1}+\Gamma_{t,d2})/\Gamma_{t,s})(1 + 
-	\beta)}{(1+2\beta)}
-	\right| \\
+\frac{(1 + \beta (\Gamma_{t,d1}+\Gamma_{t,d2})/\Gamma_{t,s})(1 + 
+	\beta)}{(1+2\beta)} 	\\
 &=
-	\left| 1 - \left(
-	\left[ 1 + \chi \beta \right]
+(1 + \chi \beta )
 \cdot 
 \frac{1 + \beta}{1 + 2\beta }
-	\right) \right|
 \end{align} 
@@
 In the limit where the $R_p$ planet has the same occurrence rate about every 
-host type regardless of its mass, $\chi=2$, and this simplifies to $\Delta = 
-\beta$, for a relative error of $59\%$.
-This is less than for Model \#1 because of the smaller volume of searchable 
-binary stars in Model \#2.
+host type regardless of its mass, $\chi=2$, and this simplifies to $X_\Gamma = 
+1+\beta$.
+Although this is same expression as what we found for Model \#1, the value of 
+$\beta$ will be smaller in this model because of the 
+smaller volume of searchable binary stars.
 

### 2017-08-30 src/binary_distribution.py
add clarifying print statement

@@ -381,6 +381,10 @@ if __name__ == '__main__':
     print( ((1/N_d) * np.sum(doubles['f_d']) / f_s).cgs )
     print('\n')
 
+    print('''\nYou never ran full numerics for this.
+    But the above is actually all you need for the Rp rate.
+    For the Rp-prime rate, you might want numerics''')
+
 
 
 

### 2017-08-30 memos/workdir/toy_analytic_surveys.tex
update memo text

@@
 For Astronomer $A'$ below, including the planets in binary star systems in the 
-numerator actually introduces confusion that precludes nice math.
+numerator introduces mistakes in the estimated completeness, and so the picture 
+is more complicated.
 Then
@@
 1+\beta$.
-Although this is same expression as what we found for Model \#1, the value of 
+This is same expression as what we found for Model \#1, but now the value of 
 $\beta$ will be smaller in this model because of the 
 smaller volume of searchable binary stars.
+From numerics, we find $\beta = N_d/N_s = 0.59$, so the occurrence rate 
+correction factor is $X_\Gamma(\chi=2) = 1.59$.
 
@@
 existing around single stars and the primaries of double star systems with 
-equal occurrence, $\chi = 1$, and $\Delta = \beta^2 / (1+2\beta)$.
-For $\beta = 0.59$, this gives $\Delta = 0.16$.
+equal occurrence, $\chi = 1$, and $X_\Gamma = (1 +\beta)^2 / (1+2\beta)$.
+For $\beta = 0.59$, this gives $X_\Gamma(\chi=1) = 1.16$.
+The true correction factor is somewhere between these two limits.
 
@@
 
-All of these diluted detections will be from binary systems. The number of 
+The diluted detections only occur in binary systems. The number of 
 planets detected in double star systems with observed radius $R_p^{\rm obs}$ 
@@
 To save on notational horribleness, write $R_p' \equiv \{(1-x) R_p < R_p^{\rm 
-	obs} < R_p\}$, in other words let $R_p'$ denote the set of planets 
-	that are observed in the desired radius range.
+obs} < R_p\}$, in other words let $R_p'$ denote the set of planets that are 
+observed in the desired radius range.
 
@@
 \,{\rm prob}(\gamma_R) \,{\rm d}(\gamma_R),
+\label{eq:N_d1}
 \end{equation}
@@
 hence the need for the minimum.
-
-The analogous equation for $N_{d1}(R_p')$, the number of double systems for 
-which a $(R_p, P)$ planet would be observed in $R_p'$ if it orbiting the 
-secondary, is
+Note that Eq.~\ref{eq:N_d1} is not an expression for the number of detected 
+planets about primaries -- that quantity is expressed as $\Gamma_{t,d1} 
+f_{d1}(R_p^{\rm obs}) N_{d1}(R_p^{\rm obs})$.
+
+The analogous equation to Eq.~\ref{eq:N_d1} for $N_{d2}(R_p')$, the number of 
+double systems for which a $(R_p, P)$ planet would be observed in $R_p'$ if it 
+orbited the secondary, is
 \begin{equation}
@@
   2^{-1/2}\\
-  0, \quad\quad\quad\quad\quad\quad\quad x \leq 1 - 2^{-1/2}
+  0, \quad\quad\quad\quad\quad\quad\quad x \leq 1 - 2^{-1/2} % lol
   \end{array}
@@
 
-If we were forward-modelling (\textit{i.e.} we actually wanted to compute 
-$N_{\rm det,d}(R_p')$) we would need to evaluate the completeness 
-term and this ``fraction of slightly diluted stars'' terms together, and 
-integrate jointly to find $N_{\rm det,d1}(R_p')$ and $N_{\rm det,d2}(R_p')$.
-
-\texttt{If we ask the right question, this will not be necessary.}
+If we were forward-modelling, \textit{i.e.} if we actually wanted to compute 
+$N_{\rm det,d}(R_p')$, we would also need to evaluate the total detection 
+efficiency term $f_{d1}(R_p^{\rm obs})$, which is a product of the geometric 
+transit probability and the fraction of signals of a given depth that are 
+detectable.
+We would need to integrate this product over the domain of allowable radii 
+$R_p'$ to find the number of detected planets.
+
+However, the procedure of estimating the geometric transit probability, as well 
+as the completeness, is complicated by the errors that are introduced by 
+including binary star systems that are thought to be single in the numerator.
+These stars will have incorrectly estimated masses and radii, which influences 
+both $f_c(R_p')$ and $f_g(R_p')$.
+
+\texttt{If we ask the right question, we will not need to actually compute 
+either term.}
 
 Astronomer A misclassifies planetary radii, miscounts the total number of 
-stars, but knows his completeness and geometric correction for single stars as 
-a function of radius $f_s(R_p')$.
+stars, and (incorrectly) estimates his completeness and geometric correction 
+for ``single'' stars as a function of radius $f(R_p')$ (omitting a ``s'' 
+subscript because the stars for which this incorrect occurrence rate are 
+derived are not all single).
 His occurrence rate for planets of radius \textit{near} $R_p$, \textit{i.e.} 
@@
 \begin{equation}
-f_s(R_p') = f_{s,g}(R_p') f_{s,c}(R_p').
+f(R_p') = f_{g}(R_p') f_{c}(R_p').
 \end{equation}
@@
 Astronomer D, who had everything right, doesn't change anything.
+
+%TODO: LEFT OFF HERE!
 

### 2017-08-30 memos/workdir/toy_analytic_surveys.tex
model 2: fix presentation of correction factor

@@
 \Gamma_{{\rm A}, R_p'} = \frac{ (N_{\rm det,s} + N_{\rm det,d}(R_p')) 
-/ f_s(R_p')   }{ N_s + N_d}.
+/ f(R_p')   }{ N_s + N_d}.
 \end{equation}
@@
 \begin{equation}
-f(R_p') = f_{g}(R_p') f_{c}(R_p').
-\end{equation}
-
-Astronomer D, who had everything right, doesn't change anything.
-
-%TODO: LEFT OFF HERE!
-
-The new difference between the two rates is
+f(R_p') = f_g(R_p') f_c(R_p').
+\end{equation}
+
+Astronomer D, who had everything right, changes nothing.
+
+The new correction factor is
 \begin{align}
-\Delta' &\equiv \left|
-\frac{\Gamma_{{\rm A},R_p'} - \Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p'} }
-\right| \\
-&= \left|
-1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p'} } 
-\right| \\
+X_\Gamma &= \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p'} } \\
 &=
-\left| 1 -
-\left( \frac{1+\chi \beta}{1 + \xi}  \cdot 
-\frac{1 + \beta}{1 + 2\beta} \right)
-\right|,
+\frac{1+\chi \beta}{1 + \xi}  \cdot 
+\frac{1 + \beta}{1 + 2\beta},
 \end{align}
@@
 \begin{equation}
-\xi \equiv \frac{N_{\rm det,d}(R_p')}{f_s(R_p')} \cdot \frac{1}{N_s 
+\xi \equiv \frac{N_{\rm det,d}(R_p')}{f(R_p')} \cdot \frac{1}{N_s 
 \Gamma_{t,s}}.
@@
 If we let $x < 1 - 2^{-1/2}$, for instance $x=0.1$, then $N_{d2}(R_p') = 0$. 
-The expression for $\xi$ simplifies to
-\begin{equation}
-\chi = \frac{N_{d1}(R_p')}{N_s} \cdot \frac{\Gamma_{t,d1}}{\Gamma_{t,s}} \cdot 
-\frac{f_{d1}(R_p')}{f_s(R_p')},\quad ({\rm if\ }x < 1-2^{-1/2}).
-\label{eq:chi_simplfied}
-\end{equation}
-
+The expression for $\xi$ in this case becomes
+\begin{align}
+\xi &= \frac{N_{d1}(R_p')}{N_s} \cdot \frac{\Gamma_{t,d1}}{\Gamma_{t,s}} \cdot 
+\frac{f_{d1}(R_p')}{f(R_p')} \\
+&= \frac{N_{d1}(R_p')}{N_s} \cdot \frac{\Gamma_{t,d1}}{\Gamma_{t,s}} \cdot 
+\frac{f_{g,d1}(R_p')}{f_g(R_p')}\cdot \frac{f_{c,d1}(R_p')}{f_c(R_p')},
+\label{eq:chi_simplified}
+\end{align}
+provided that $x < 1-2^{-1/2}$.
+
+We have now parametrized our ignorance.
+The values that the latter two ratios of Eq.~\ref{eq:chi_simplified} take 
+depend on what we assume Astronomer A' does to derive the stellar parameters 
+for the systems that he thinks are singles but are in fact doubles.
+For instance, assume he somehow observed the bolometric flux from every point 
+on the sky, and used it with a distance to obtain single-star luminosities, and 
+thus radii and masses.
+His binary systems (thought to be singles) would be bluer than either component 
+star is in reality.
+On the main sequence, $\rho_\star$ decreases with increasing stellar radius, 
+so $\rho_\star$ would be underestimated for all binaries.
+Since $f_g \propto \rho_\star^{1/3}$, this means that $f_{g,d1}(R_p')/f_g(R_p') 
+> 1$.
+However, by the same token we can place an \textit{upper} bound on this ratio. 
+In the worst case scenario a given star's luminosity is twice the individual 
+star's luminosity. If $L\propto M^3 \propto R^3$, then the radius is 
+over-estimated by at most a factor of $2^{1/3}$. Thus the density is 
+underestimated by at most a factor of $2^{1/2}$, and $f_g(R_p')$ is 
+underestimated by at most a factor of $2^{1/6}=1.12$.
+
+This begins to take us somewhere. The left-most term of 
+Eq.~\ref{eq:chi_simplified} is bounded above by $\beta$. This is because 
+$N_{d1}(R_p') \leq N_{d1} = N_d$.
 If we assume $\Gamma_{t,d1} = \Gamma_{t,s}$, as we were doing 
-anyway, Eq.~\ref{eq:chi_simplfied} is bounded above by
-$\beta$. This is because $N_{d1}(R_p') \leq N_{d1} = N_d$, and $f_{d1}(R_p') / 
-f_s(R_p') \leq 1$, since over a fixed radius interval dilution will always 
-yield a lower completeness fraction for double star systems than singles.
-Recall for the fixed $\gamma_R$ population we showed that $f_{d1}/f_s = 
-(1+\gamma_R)^{-3}$.
-
-Regardless, the cheap way of making the estimate without deriving the actual 
+anyway, then the product of the three left-most fractions must be less than 
+$1.12\beta$.
+The only thing standing in our way of an analytic upper bound to $\xi$ is the 
+completeness fraction $f_{c,d1}(R_p')/f_c(R_p')$.
+Note that over a fixed radius interval dilution will always 
+yield a lower completeness fraction for double star systems than singles, so 
+$f_{c,d1}(R_p') / f_{c,s}(R_p') \leq 1$.
+
+For the completeness term, an argument that I don't wholly believe
+is that
+\begin{equation}
+\frac{f_{c,d1}(R_p')}{f_c(R_p')} = \frac{f_{c,d1}(R_p')}{f_{c,s}(R_p')} \cdot
+	\frac{f_{c,s}(R_p')}{f_{c}(R_p')} \lessapprox 1 \cdot (1+\langle \gamma_R 
+	\rangle)^3,
+\end{equation}
+where the first part of the inequality is obvious, but the second is by 
+assuming it's $\lessapprox f_{s,c}/f_{d,c}$, which might be wrong because of 
+the radius dependence.
+
+A way of making the estimate without deriving the actual 
 completeness fractions, or even running numerics, is just to plot the error as 
-a function of $\xi$. We do this in Fig.~\ref{fig:deltaprime_vs_xi}, which 
-shows that the
-relative error varies between $\approx 5\% - 35\%$, depending on what 
+a function of $\xi$. We do this in Fig.~\ref{fig:XGamma_vs_xi}, which 
+shows that the correction factor varies from 0.6 to 1.6, depending on what 
 fraction of secondaries you assume host planets.
+\todo[inline]{I could run numerics for this to make the result more 
+interesting.}
 
@@
 	\begin{center}
-		\includegraphics[scale=.8]{figures/deltaprime_vs_xi.pdf}
+		\includegraphics[scale=.8]{figures/XGamma_vs_xi.pdf}
 	\end{center}
-	\caption{\textit{Y axis}: relative error between the Astronomer who ignores 
-	binarity 
-	but counts planetary radii that are ``close enough'' (A') and the Astronomer 
-	who knows everything (D).
-	\textit{X axis}: $\chi$ as defined in Eq.~\ref{eq:chi_simplfied}. Taking a 
+	\caption{\textit{Y axis}: correction factor for the astronomer who 
+	ignores binarity but counts planetary radii that are ``close enough'' (A') 
+	and the astronomer who knows everything (D).
+	\textit{X axis}: $\chi$ as defined in Eq.~\ref{eq:chi_simplified}. Taking a 
 	guess at the true value, if $N_{d1}(R_p') = N_d/2$, and the occurrence rates 
@@
 		}
-	\label{fig:deltaprime_vs_xi}
+	\label{fig:XGamma_vs_xi}
 \end{figure}
@@
 known.
-\todo[inline]{but it might be worth doing something directly on the KIC's 
+\todo[inline]{but it might be better to do something directly on the KIC's 
 selected target stars, like what Ciardi did! The following is a good example, 
@@
 
-Similar to Fig.~\ref{fig:deltaprime_vs_xi}, Fig.~\ref{fig:error_vs_occrate} 
+Similar to Fig.~\ref{fig:XGamma_vs_xi}, Fig.~\ref{fig:error_vs_occrate} 
 shows that the magnitude of the relative error 

### 2017-08-30 src/XGamma_vs_xi.py

@@ -1,0 +1,40 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+Nd_by_Ns = 0.59
+
+ξ = np.arange(0, 1.5*1.12*Nd_by_Ns, 1e-3)
+
+f,ax = plt.subplots(nrows=1, ncols=1, figsize=(4,4))
+
+χ = 2 # greek chi
+β = 0.59
+
+X_Γ =  ((1 + χ*β)/(1+ξ)) * (1+β)/(1+2*β)
+
+ax.plot(ξ, X_Γ, 'k-', label=r'$\chi = 2, \beta = 0.59$')
+
+χ = 1 # greek chi
+β = 0.59
+X_Γ =  ((1 + χ*β)/(1+ξ)) * (1+β)/(1+2*β)
+
+ax.plot(ξ, X_Γ, 'b-', label=r'$\chi = 1, \beta = 0.59$')
+
+χ = 1.5 # greek chi
+β = 0.59
+X_Γ =  ((1 + χ*β)/(1+ξ)) * (1+β)/(1+2*β)
+
+ax.plot(ξ, X_Γ, 'g-', label=r'$\chi = 1.5, \beta = 0.59$')
+
+
+ax.legend(loc='best', fontsize='small', frameon=False)
+
+#ax.set_xlim([0,β])
+#ax.set_ylim([0,0.6])
+
+ax.set_ylabel("$X_\Gamma = \Gamma_{\mathrm{D},R_p}/\Gamma_{\mathrm{A},R_p\'}$")
+ax.set_xlabel(r"$\xi$ (maximum is $\approx (1+\langle \gamma \rangle)^3\cdot 1.12\cdot \beta \approx 1$)")
+
+f.tight_layout()
+
+f.savefig('XGamma_vs_xi.pdf', dpi=300, bbox_inches='tight')

### 2017-08-30 src/deltaprime_vs_xi.py

@@ -1,46 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-Nd_by_Ns = 0.59
-
-ξ = np.arange(0, Nd_by_Ns, 1e-3)
-
-f,ax = plt.subplots(nrows=1, ncols=1, figsize=(4,4))
-
-χ = 2 # greek chi
-β = 0.59
-Δprime = abs(
-        1 - ( (1 + χ*β)/(1+ξ) )*( (1+β)/(1+2*β)  )
-        )
-
-ax.plot(ξ, Δprime, 'k-', label=r'$\chi = 2, \beta = 0.59$')
-
-χ = 1 # greek chi
-β = 0.59
-Δprime = abs(
-        1 - ( (1 + χ*β)/(1+ξ) )*( (1+β)/(1+2*β)  )
-        )
-
-ax.plot(ξ, Δprime, 'b-', label=r'$\chi = 1, \beta = 0.59$')
-
-
-χ = 1.5 # greek chi
-β = 0.59
-Δprime = abs(
-        1 - ( (1 + χ*β)/(1+ξ) )*( (1+β)/(1+2*β)  )
-        )
-
-ax.plot(ξ, Δprime, 'g-', label=r'$\chi = 1.5, \beta = 0.59$')
-
-
-ax.legend(loc='best', fontsize='small', frameon=False)
-
-ax.set_xlim([0,β])
-ax.set_ylim([0,0.6])
-
-ax.set_ylabel("$\Delta'$ (relative error btwn A' and D)")
-ax.set_xlabel(r"$\xi$ (max is upper bound of $\beta$)")
-
-f.tight_layout()
-
-f.savefig('deltaprime_vs_xi.pdf', dpi=300, bbox_inches='tight')

### 2017-08-30 src/kepler_numerics.py
kepler numerics: fix correction factor

@@ -1202,10 +1202,9 @@ def evaluate_occurrence_rate_errors(true_occ_rates):
     chi = (Γ_t_d1 + Γ_t_d2)/Γ_t_s
     #print('chi: {:.3f}'.format(chi))
 
-    Δprime = 1 - Γ_D_Rp / Γ_A_Rpprime
-    #print('Δprime: {:.3f}'.format(Δprime))
+    X_Γ = Γ_D_Rp / Γ_A_Rpprime
 
-    return Δprime
+    return X_Γ
 
 
 def do_kic_complete_survey():
@@ -1276,11 +1275,11 @@ def do_kepler_analog():
         prioritize_target_stars()
 
     Γ_t_d2_arr = np.arange(0, 0.80, 0.05)
-    avg_Δprimes, std_Δprimes = [], []
+    avg_X_Γs, std_X_Γs = [], []
 
     for Γ_t_d2 in Γ_t_d2_arr:
 
-        Δprimes = []
+        X_Γs = []
 
         for i in range(N_trials):
 
@@ -1292,18 +1291,18 @@ def do_kepler_analog():
 
             add_planets_and_do_survey(true_occ_rates, is_kepler_analog=True)
 
-            Δprime = evaluate_occurrence_rate_errors(true_occ_rates)
+            X_Γ = evaluate_occurrence_rate_errors(true_occ_rates)
 
-            Δprimes.append(Δprime)
+            X_Γs.append(X_Γ)
 
-        avg_Δprimes.append(np.mean(Δprimes))
-        std_Δprimes.append(np.std(Δprimes))
+        avg_X_Γs.append(np.mean(X_Γs))
+        std_X_Γs.append(np.std(X_Γs))
 
-    out = np.array(avg_Δprimes)
-    out_err = np.array(std_Δprimes)
+    out = np.array(avg_X_Γs)
+    out_err = np.array(std_X_Γs)
 
-    out = pd.DataFrame({'Δprimes': out,
-                        'Δprimes_err': out_err,
+    out = pd.DataFrame({'X_Γ': out,
+                        'X_Γ_err': out_err,
                         'Γ_t_d2': Γ_t_d2_arr})
 
     out.to_csv('../results/error_v_occrate_data.csv', index=False)
@@ -1314,25 +1313,27 @@ def plot_error_v_occrate(Γ_t_d2_best_guess):
     import matplotlib.pyplot as plt
[…]

### 2017-08-30 memos/workdir/toy_analytic_surveys.tex
finished updating memo with correction factor

@@
 we underestimate occurrence rates for a given planet radius and period by 
-$\approx 39\%$.
-This is immediately applicable to the question of Earth-like planets around 
-Sun-like stars.
+a factor of $\approx 1.4$, or a percentage error relative to the true 
+occurrence rate $\Gamma$ of $\delta\Gamma=-28\%$.
+This is relevant for measuring accurate values of the rate at which Earth-like 
+planets exist around Sun-like stars.
 
@@
 \frac{f_{c,d1}(R_p')}{f_c(R_p')} = \frac{f_{c,d1}(R_p')}{f_{c,s}(R_p')} \cdot
-	\frac{f_{c,s}(R_p')}{f_{c}(R_p')} \lessapprox 1 \cdot (1+\langle \gamma_R 
+	\frac{f_{c,s}(R_p')}{f_{c}(R_p')} \lesssim 1 \cdot (1+\langle \gamma_R 
 	\rangle)^3,
@@
 where the first part of the inequality is obvious, but the second is by 
-assuming it's $\lessapprox f_{s,c}/f_{d,c}$, which might be wrong because of 
+assuming it's $\lesssim f_{s,c}/f_{d,c}$, which might be wrong because of 
 the radius dependence.
@@
 	and the astronomer who knows everything (D).
-	\textit{X axis}: $\chi$ as defined in Eq.~\ref{eq:chi_simplified}. Taking a 
-	guess at the true value, if $N_{d1}(R_p') = N_d/2$, and the occurrence rates 
-	between `d1' and `s' are the same, and we put $\gamma_R = 0.1$ into the 
-	completeness ratio, then $\xi = (\beta/2) \times (1+0.1)^{-3} = 0.22$. The 
-	relative error then varies between $\approx 5\% - 35\%$, depending on what 
-	fraction of secondaries you assume host planets.
+	\textit{X axis}: $\chi$ as defined in Eq.~\ref{eq:chi_simplified}.
 	}
@@
 Similar to Fig.~\ref{fig:XGamma_vs_xi}, Fig.~\ref{fig:error_vs_occrate} 
-shows that the magnitude of the relative error 
-made by the $\Gamma_{{\rm A},R_p'}$ and $\Gamma_{{\rm A},R_p}$ estimates 
-depends strongly 
+shows that the magnitude of the correction needed by the $\Gamma_{{\rm 
+A},R_p'}$ estimate depends strongly 
 on the fraction of secondaries that are assumed to host the $(R_p, P)$ planets.
@@
 corresponds to $\Gamma_{\rm D} = 1.39 \Gamma_{\rm A'}$.
+This gives a relative percentage error of $\delta\Gamma_{{\rm D},R_p} \equiv 
+(\Gamma_{\rm A,R_p'} - \Gamma_{\rm D,R_p}) / \Gamma_{\rm D,R_p} = -28\%$.
 

### 2017-08-30 memos/workdir/toy_analytic_surveys.tex
memo: update abstract

@@
 
-Then, we allow the properties of the secondary in binary systems to vary. A 
-distribution of light ratios introduces a heavy detection bias towards 
+We then allow the properties of the secondary in binary systems to vary. A 
+distribution of light ratios introduces a detection bias towards 
 equal-mass binaries in magnitude limited samples.
 We show that dependent upon planet occurrence's functional dependence on host 
-star mass (and binarity), the errors introduced in derived occurrence rates vs. 
-real occurrence rates can range from a few percent to at most $\approx 60\%$ 
-relative error.
+star mass, as well as the specifics of the assumed survey completeness, the 
+occurrence rate correction factor $X_\Gamma \equiv \Gamma_{\rm 
+true}/\Gamma_{\rm obs}$ ranges from $0.6-1.6$.
 

### 2017-08-31 src/binary_distribution.py
add analytic mag-limited prob(q) derivn and plots

@@ -65,7 +65,18 @@ def _make_distribution_plots(vl, doubles):
     f,ax = plt.subplots(figsize=(4,4))
     hist, bin_edges = np.histogram(doubles['q'],
             bins=np.append(np.linspace(0,1,101),42), normed=True)
-    ax.step(bin_edges[:-1], hist, 'k-', where='post')
+    ax.step(bin_edges[:-1], hist, where='post',
+            label='numerical, empirical $L(M)$')
+    # Analytic distribution from 17/08/31.2 result
+    _q = np.arange(0,1+1e-3,1e-3)
+    I_1 = 0.4645286925158471
+    I_2 = 1.323588493214896
+    norm = 9/(I_1*I_2)
+    pdf_q_analytic = norm*I_1/9*(1+_q**3)**(3/2)
+    pdf_q_analytic[_q<0.1] = 0
+    ax.plot(_q, pdf_q_analytic, label='analytic, $L=M^3$')
+    ax.legend(loc='upper left', fontsize='small')
+
     ax.set(xlabel='$q = M_2/M_1$', ylabel='prob')
     ax.set_title(txt_ml, fontsize='small')
     f.tight_layout()

### 2017-08-31 src/joint_prob_r_q.py

@@ -1,0 +1,133 @@
+'''
+compute and plot the joint 2d probability distribution function of a binary
+star's position and mass ratio in a magnitude-limited sample.
+'''
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.integrate import trapz
+
+r = np.linspace(0,0.5,3e3)
+q = np.linspace(0,1,3e3)[::-1]
+
+L_1_N = 1
+F_lim_N = 1
+
+d_max_of_q = ( L_1_N / (4*np.pi*F_lim_N) * (1+q**3) )**(1/2)
+
+d_max_s = (L_1_N / (4*np.pi*F_lim_N))**(1/2)
+# integral of (1+q^3)^(-9/2)dq from q=0.1 to 1 evaluates to
+I_1 = 0.4645286925158471
+# integral of (1+q^3)^(3/2)dq from q=0.1 to 1 evaluates to
+I_2 = 1.323588493214896
+# normalization constant brings integrals to 1
+norm = 9/(I_1*I_2)
+
+prob = norm * I_1 * r**2 / (3*d_max_s**3)
+prob = prob[:,None] * np.ones(len(q))[None,:]
+
+# Below line: incorrect calculation, based on math that I don't understand my
+# error from, but that I work around.
+## prob = np.outer(1/3 * r**2, 1/(d_max_of_q**3))
+
+prob_pre = np.copy(prob)
+
+print('before mask, distribn integrates to {:.3f}'.format(
+    trapz(trapz(prob, r, axis=0), q[::-1])))
+print('before mask, distribn integrates to {:.3f}'.format(
+    trapz(trapz(prob, q[::-1], axis=1), r)))
+
+# masks must be bool. if you try int, odd things happen.
+mask_q = (q<0.1) * np.ones(len(r))[:,None].astype(bool)
+prob[ mask_q ] = 0
+
+d_max_qr = np.transpose(d_max_of_q[:,None] * np.ones(len(r))[None,:])
+r_qr = r[:,None] * np.ones(len(q))[None,:]
+
+mask_r = (r_qr > d_max_qr).astype(bool)
+
+prob[ mask_r ] = 0
+
+print('after mask, distribn integrates to {:.3f}'.format(
+    trapz(trapz(prob, r, axis=0), q[::-1])))
+print('after mask, distribn integrates to {:.3f}'.format(
[…]

### 2017-08-31 memos/workdir/toy_analytic_surveys.tex
memo: fix typo
memo: add analytic mag-limited prob(q)

@@
 	because they are detectable out to a greater distance\footnote{This is a 
-	scarcy systematic w.r.t. the claimed intrinsic excess of twin binaries.}. 
-	This is shown in 
-	Fig.~\ref{fig:q_distribn_mag_limited}.
+	scary systematic w.r.t. the claimed intrinsic excess of twin binaries.}. 
+	This is shown in Fig.~\ref{fig:q_distribn_mag_limited}, and explained 
+	analytically below.
 	\item In a magnitude-limited sample of binary star systems in which the 
@@
 $10^{-3}$, where the probability distribution peaks.
+
+Similarly, the probability distribution function for the mass ratio of binaries 
+in a magnitude limited sample can be found by marginalizing over the joint 
+distribution for a binary star's position and mass ratio.
+Explicitly, using `ml' as a subscript for `magnitude-limited', and assuming 
+$L\propto M^3$,
+\begin{align}
+{\rm prob}_{\rm ml}(q) &= 
+	\int_{0}^{d_{\rm max}(q)} {\rm prob}_{\rm ml}(r,q) \,{\rm d}r \\
+&= \mathcal{Z}\cdot (1+q^3)^{3/2}\quad {\rm if\ }0.1<q\leq 1,
+\label{eq:pdf_q_mag_limited}
+\end{align}
+and otherwise 0, for $\mathcal{Z}$ the appropriate normalization constant.
+Eq.~\ref{eq:pdf_q_mag_limited} is the ``analytic'' form overplotted in 
+Fig.~\ref{fig:q_distribn_mag_limited}. Its independence from both the primary 
+star mass as well as from the magnitude cutoff will be used in 
+Sec.~\ref{sec:model_3}.
+The form given in Eq.~\ref{eq:pdf_q_mag_limited} is found by writing 
+${\rm prob}_{\rm ml}(r,q) = {\rm prob}(q|r){\rm prob}_{\rm ml}(r)$, using 
+Eq.~\ref{eq:mass_ratio} for ${\rm prob}(q|r)$, and marginalizing over the mass 
+ratio
+\begin{align}
+{\rm prob}_{\rm ml}(r) &= \int_{0}^{1} {\rm prob}_{\rm ml}(r|q)\,{\rm 
+prob}(q)\,{\rm d}q
+\\
+&= \int_{0.1}^{1} \frac{1}{3} \frac{r^2}{d_{\rm max}^3(q)}\,{\rm d}q
+\quad {\rm if\ }r<d_{\rm max}(q),
+\end{align}
+and otherwise zero. The resulting joint probability distribution is plotted in 
+Fig.~\ref{fig:joint_prob_r_q}.
 
@@
 	\caption{The distribution of the mass ratio for a magnitude limited 
-		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
-		total 
-		number density from Bovy 2017; $M(L)$ relation from 
-		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
+		sample of binary stars. For numerics, ${\rm BF} = 0.45$ [Duchene and 
+		Kraus 2013], total number density from Bovy 2017, empirical $M(L)$ 
+		from Fig.~\ref{fig:mass_luminosity}, and mass ratios are drawn from 
 		Eq.~\ref{eq:mass_ratio} -- a \textit{uniform} distribution in a 
-		volume-limited sample!}
+		volume-limited sample!
+		The analytic line (Eq.~\ref{eq:pdf_q_mag_limited}) is shown to indicate 
+		that the entire effect, albeit with a simpler $M(L)$, can be understood 
+		without resorting to numerics. 
+		}
 	\label{fig:q_distribn_mag_limited}
@@
 	\label{fig:gammaR_distribn_mag_limited}
+\end{figure}
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=.8]{figures/joint_prob_r_q.pdf}
+	\end{center}
+	\caption{The joint probability distribution of a binary star's position and 
+	mass ratio in a magnitude-limited sample. This plot assumes a different 
+	$M(L)$ relation than previous plots: $L=M^3$. Note that the marginalized 
+	mass ratio distribution is uniform at any given distance.}
+	\label{fig:joint_prob_r_q}
 \end{figure}
@@
 \section{Model \# 3: A Synthetic {\it Kepler} Analog}
+\label{sec:model_3}
 

### 2017-09-22 memos/workdir/toy_analytic_surveys.tex
general cleanup. logg subgiant check. read papers.
renamed things

parsed Furlan et al (2017)'s summary of all the high resolution imaging
following paper.

added the "logg subgiant bias check": Ji Wang claimed in his ~2015 HJ
occ rate paper that like ~50% of the Kepler stars (or something of
equally large number) were subgiants. This was based on an erroneous
extrapolation from Bastien et al's result which showed that for BRIGHT
Kepler stars (where the subgiant contamination is greatest because of a
Malmquist bias). For most of the sample, from 13<Kp<16, it's not such a
big concern.

also wro

@@
 \shortauthors{LGB, KM, JNW}
-\shorttitle{Toy transit surveys}
+\shorttitle{Binary Biases}
 
@@
 
-\title{ Toy Transit Surveys }
+\title{ Binary Biases \& Toy Transit Surveys }
 
@@
 a factor of $\approx 1.4$, or a percentage error relative to the true 
-occurrence rate $\Gamma$ of $\delta\Gamma=-28\%$.
-This is relevant for measuring accurate values of the rate at which Earth-like 
-planets exist around Sun-like stars.
+occurrence rate $\Gamma_{\rm true}$ of $\delta\Gamma_{\rm 
+true}=-28\%$. This is relevant for measuring accurate values of the rate at 
+which Earth-like planets exist around Sun-like stars.
 
@@
 \section{Introduction}
-
-Ah, transiting planets! We learn so much by studying them. But how much can we 
-actually learn, and how much is messed up by binarity?
 
@@
 detectability the most.
-
-A related effect to dilution is misinterpretation of transit depths. If the 
+Note that dilution can come from visually associated companions as well as 
+gravitationally bound companions.
+
+A consequence of dilution is misinterpretation of transit depths. If the 
 planet to star radius ratio is directly taken as the square-root of 
@@
 systematically too small.
-Ciardi et al [2015] explored this issue in depth. They showed that for the case 
-of \kepler, assuming host stars to always be single will on 
-average underestimate radii by a multiplicative factor of $\approx 1.5$.
-
-Finally, binarity may affect transit surveys through actual astrophysics.
-For the relatively rare case [Duquennoy \& Mayor 1991, Raghavan et al 2010] of 
-close binaries, dynamical stability imposes a limited range of periods in which 
-planets may orbit [e.g., Holman \& Weigert 
-1999, Li et al 2017], the effects of which have been observed [Kraus et al 
+Ciardi et al [2015] explored this issue in depth. They claimed that for the 
+case of \kepler, if ones assumes host stars are always single, one will
+underestimate radii by an average multiplicative factor of $\approx 1.5$ 
+(presumably, compared to the case of even weights for whether the 
+primary/secondary gets a given planet).
+
+Finally, binarity may affect transit surveys through actual astrophysical 
+processes.
+Duquennoy \& Mayor [1991], and alter Raghavan et al [2010] showed that over 
+half of solar-like are separated by less than 100\,AU.
+One immediate consequence on circumprimary and circumsecondary architectures is 
+that dynamical stability limits begin to carve away in realms of planet space 
+where planets might otherwise be detectable [e.g., Holman \& Weigert 1999, Li 
+et al 2017].
+The formation efficiency of various types of planetary systems may also 
+be different in binary systems compared to single-star systems.
+For instance, the properties of the circumstellar and circumbinary disks 
+influence the dynamics of planetesimals [Rafikov, Silsbee];
+
+ the effects of which have been observed [Kraus et al 
 2016].
@@
 but at most it can ``suggest'' that a similar level of error has been made from 
-the \kepler\ survey.}
+the \kepler\ survey.
+
+But it has MAJOR benefits compared to the GIC method.
+It eliminates the need to simulate the biases of the KIC (since they're already 
+baked in). The process of ``adding'' binaries to the KIC is the closest thing 
+to REALITY, since in reality the KIC is missing binary companion information.
+It BAKES IN all the real biases that come with mis-estimating stellar 
+radii/masses when they are actually binaries (because it's already been done).
+It also has literature precedent in Ciardi's work, so you don't need to cook up 
+a crazy new procedure -- can just take his with the flaws baked in.
+
+So this method is best if we care about what errors were made in Kepler/the KIC.
+It's not as nice if we want something ``general''.
+}
 

### 2017-09-22 src/binary_distribution.py

@@ -97,6 +97,30 @@ def _make_distribution_plots(vl, doubles):
     plt.close('all')
     f,ax = plt.subplots(figsize=(4,4))
     hist, bin_edges = np.histogram(doubles['γ_R'],
+            bins=np.append(np.logspace(-4,1,11),42), normed=True)
+    ax.step(bin_edges[:-1], hist, 'k-', where='post')
+    ymax = 1.25
+    ax.vlines(1/40, 0, ymax,
+          label='Furlan+17 fig 21,'
+          r'$\theta<1$arcsec: '
+          '{:.1f}% of doubles'.format(
+              len(doubles['γ_R'][doubles['γ_R']<(1/40)])\
+              /len(doubles['γ_R'])*100)+\
+          ' have $\Delta m > 4$'
+          )
+    ax.set(xlabel='$\gamma_R = L_2/L_1$',
+           ylabel='prob',
+           xlim=[5e-4,1.05],
+           xscale='log')
+    ax.legend(loc='upper left', fontsize=5)
+    ax.set_title(txt_ml, fontsize='small')
+    f.tight_layout()
+    f.savefig(savedir+'gammaR_distribn_mag_limited_logx.pdf', dpi=250, bbox_inches='tight')
+
+
+    plt.close('all')
+    f,ax = plt.subplots(figsize=(4,4))
+    hist, bin_edges = np.histogram(doubles['γ_R'],
             bins=np.append(np.linspace(0,1,501),42), normed=True)
     ax.plot(np.append(0, bin_edges[:-1]),
             np.append(0, np.cumsum(hist)/np.max(np.cumsum(hist))),

### 2017-09-22 src/logg_subgiant_bias_check.py

@@ -1,0 +1,185 @@
+'''
+Bastien et al (2014, ApJ 788:L9) showed that for a Kepmag<13 sample, the NASA
+Exoplanet Archive derived stellar parameters were biased to have smaller logg's
+than those measured with their flicker method (cf Bastien+ 2013).
+
+This matters for the derived planetary radius distribution -- in Bastien et al
+(2014), it systematically shifted the derived planetary properties for this
+sub-population by a factor of ~1.3x
+
+However, it also probably matters for the HJ rate of this subpopulation. This
+is because the HJ occ rate is claimed to be lower for subgiants than MS stars.
+This claim probably first came from Hatzes et al. (2003), but was substantiated
+by the California group adding "retired A stars" to their target list
+("retired" because they were thought to be formerly F stars by mass, but now A
+stars and evolved -- though there has been lots of confusion about this). This
+work was associated with publications by Johnson et al. (2007, 2008, 2010c,
+2010b, 2011b, 2011a).  Robinson et al.  (2007), Fischer et al.  (2007), and
+Bowler et al. (2010) also contributed to this observational question.
+
+Regardless, Schlaufman & Winn (2013) looked at this, came up with an
+independent kinematic argument for why HJ occ rates are probably lower around
+subgiants, and then argued that tidal evolution is probably the best
+explanation (though no physical models convincingly produce the rates).
+
+Regardless, ASSUMING we believe HJ occ rates are lower about subgiants,
+subgiant contamination of the Kepler sample then matters. (For both the radius
+distribution, and the occurrence rates of any close-in planets).
+
+Wang et al (2015 ApJ 799:229) noticed this in the context of the
+barely-statistically significant difference between the HJ occ rates measured
+by transit surveys (mostly Kepler, Howard et al 2012, & Fressin et al 2013),
+Γ_HJ,Kepler ~= 0.5%, and RV surveys (mostly CKS, Wright et al 2012, but also
+the unpublished Mayor et al 2011), Γ_HJ,Kepler ~= 1%.
+
+Wang et al (2015) did the following. Start with 1e5 planets. Give them a power
[…]

### 2017-09-22 src/model_2.py

@@ -1,6 +1,8 @@
 '''
-numerical verification of "method 2" as described in
-memos/toy_analytic_surveys_170804.pdf
+numerical verification of "model 2" as described in
+memos/toy_analytic_surveys_170804.pdf.
+
+fixed Rp. fixed primary L1. varying light ratio (L2/L1).
 '''
 from __future__ import division, print_function
 

### 2017-09-22 src/model_3.py

@@ -1,0 +1,1419 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+'''
+>  Can we ignore binarity when mapping KIC stars onto synthetic local stars?
+
+Or better yet
+
+> If we assume every KIC star is single, what OOM of error do we make in
+> occurrence rates estimated for planets of different sizes (& periods)?
+
+Answer the latter question with Monte Carlo simulations of the Kepler field.
+
+with TRILEGAL/Galaxia:
+    * construct the Kepler field (w/ known binarity properties, how?)
+    * introduce various planet populations
+    * numerically assess the error.
+    * Compare w/ analytic results.
+'''
+
+from __future__ import print_function, division
+import sys, os
+import pickle
+import numpy as np, pandas as pd
+
+from astropy import units as u, constants as c
+from astropy.coordinates import SkyCoord
+
+global DATADIR
+DATADIR = '/home/luke/local/GalaxiaData/kepler_synthesis/'
+
+def T_dur(P, ρ_star, b):
+    '''
+    Compute the transit duration given orbital period, host star density, and
+    impact parameter.
+    E.g., Sullivan et al 2015, Eq 10. (Assumes zero eccentricity).
+    '''
+
+    from math import pi as π
+    ρ_sun = 3*u.Msun/(4*π*u.Rsun**3)
+    return 13*u.hr * (P/(1*u.year))**(1/3) * \
+                     (ρ_star/ρ_sun)**(-1/3) * \
+                     np.sqrt(1-b*b)
+
+
+def are_synthetic_stars_on_silicon():
+    '''
+    Galaxia model was run with 200 sq deg FoV, centered at the true Kepler
+    field center. Select synthetic Kepler field stars by comparing coordinates
+    to viable KIC stars.
+    '''
+    import K2fov
+    # CITE: http://adsabs.harvard.edu/abs/2016ascl.soft01009M, Mullaly,
+    # Barclay, Barentsen 2017.
+
+    gal = pickle.load(
+            open(DATADIR+'keplersynthesis.p', 'rb'), encoding='latin1')
+
+    c = SkyCoord(frame='galactic', l=gal['glon']*u.deg, b=gal['glat']*u.deg)
+
+    df = pd.DataFrame(np.transpose([np.array(c.fk5.ra), np.array(c.fk5.dec)]),
+                      columns=['ra','dec'])
+
+    df['mags'] = np.ones_like(gal['glon'])
+
[…]

### 2017-09-22 src/model_4.py

@@ -1,20 +1,11 @@
 #!/usr/bin/env python
 # -*- coding: utf-8 -*-
 '''
->  Can we ignore binarity when mapping KIC stars onto synthetic local stars?
+varying Rp, stellar luminosities, light ratios -- varying everything!
 
-Or better yet
+How dos this affect the inferred radius distributions?
 
-> If we assume every KIC star is single, what OOM of error do we make in
-> occurrence rates estimated for planets of different sizes (& periods)?
-
-Answer the latter question with Monte Carlo simulations of the Kepler field.
-
-with TRILEGAL/Galaxia:
-    * construct the Kepler field (w/ known binarity properties, how?)
-    * introduce various planet populations
-    * numerically assess the error.
-    * Compare w/ analytic results.
+Answer the question with Monte Carlo simulations (the GIC).
 '''
 
 from __future__ import print_function, division
@@ -40,48 +31,6 @@ def T_dur(P, ρ_star, b):
     return 13*u.hr * (P/(1*u.year))**(1/3) * \
                      (ρ_star/ρ_sun)**(-1/3) * \
                      np.sqrt(1-b*b)
-
-
-def are_synthetic_stars_on_silicon():
-    '''
-    Galaxia model was run with 200 sq deg FoV, centered at the true Kepler
-    field center. Select synthetic Kepler field stars by comparing coordinates
-    to viable KIC stars.
-    '''
-    import K2fov
-    # CITE: http://adsabs.harvard.edu/abs/2016ascl.soft01009M, Mullaly,
-    # Barclay, Barentsen 2017.
-
-    gal = pickle.load(
-            open(DATADIR+'keplersynthesis.p', 'rb'), encoding='latin1')
-
-    c = SkyCoord(frame='galactic', l=gal['glon']*u.deg, b=gal['glat']*u.deg)
-
-    df = pd.DataFrame(np.transpose([np.array(c.fk5.ra), np.array(c.fk5.dec)]),
-                      columns=['ra','dec'])
-
-    df['mags'] = np.ones_like(gal['glon'])
-
-    df.to_csv(DATADIR+'raw_galaxia_coords.csv', index=False, header=False)
-
-    # skimming source of K2fov, they *do* tabulate the original Kepler field,
-    # as "preliminary" (for a single season). Access with "campaign number"
-    # 1000.
[…]

### 2017-09-22 src/vary_Rp_study_pdfs.py

@@ -1,0 +1,502 @@
+'''
+A straight numerical approach would be:
+
+Take all Galaxia stars within 7.5 degrees of the center of Kepler's
+field, and only take stars with r<17.
+
+This is beyond where the KIC is complete, so comparisons between the two
+will be confused.  Then apply analogs of the Batalha+ 2010 Table 1
+prioritization.
+
+This gives the "GIC", and it's Teff distribution is OK. Its binary
+radius distribution is currently a bit wonky, which is not nice.
+
+If you give the GIC a planet population from various obvious options
+(uniform in Rp, uniform in logRp, normal in logRp) you can study what the
+binarity does to the inferred radius distribution.
+
+ALTERNATIVELY
+
+you can do it all semianalytically, following notes of 17/09/03.2-3
+'''
+
+import numpy as np
+from scipy.integrate import trapz
+import matplotlib.pyplot as plt
+
+def prob_gammaR(γ_R):
+
+    Z = 1 # constants seem to work
+    prob_γ_R = np.zeros_like(γ_R)
+
+    inds = (γ_R<1) & (γ_R>0.1**3)
+    prob_γ_R[inds] = Z * (1 + γ_R[inds])**(3/2) * (1/3) * γ_R[inds]**(-2/3)
+
+    return prob_γ_R
+
+
+def prob_Rpt_uniform(R_pt, R_pu, R_pl):
+    '''
+    uniform distribution
+    R_pt: array for to return the probability over
+    R_pu, R_pl: floats of upper and lower bounds
+    '''
+
+    c_Rp = 1/(R_pu - R_pl)
+
+    prob_Rpt = np.zeros_like(R_pt)
+
+    inds = (R_pt > R_pl) & (R_pt < R_pu)
+    prob_Rpt[inds] = c_Rp
+
+    return prob_Rpt
+
+
+def prob_Rpt_loguniform(R_pt, R_pu, R_pl):
+    '''
+    loguniform distribution (not normalized correctly; done in post)
+    R_pt: array for to return the probability over
+    R_pu, R_pl: floats of upper and lower bounds
+    '''
+
+    const = 1
+
[…]

### 2017-09-25 src/binary_distribution.py
modify M(L) to M~L^3.5

@@ -4,6 +4,8 @@ sample, vs a volume limited sample?
 '''
 from __future__ import division, print_function
 
+import matplotlib as mpl
+mpl.use('Agg')
 import numpy as np, pandas as pd, matplotlib.pyplot as plt
 from astropy import units as u, constants as c
 from math import pi as π
@@ -67,14 +69,14 @@ def _make_distribution_plots(vl, doubles):
             bins=np.append(np.linspace(0,1,101),42), normed=True)
     ax.step(bin_edges[:-1], hist, where='post',
             label='numerical, empirical $L(M)$')
-    # Analytic distribution from 17/08/31.2 result
+    # Analytic distribution from 17/09/24 result
     _q = np.arange(0,1+1e-3,1e-3)
-    I_1 = 0.4645286925158471
-    I_2 = 1.323588493214896
+    I_1 = 0.5060577377392849
+    I_2 = 1.275894325140383
     norm = 9/(I_1*I_2)
-    pdf_q_analytic = norm*I_1/9*(1+_q**3)**(3/2)
+    pdf_q_analytic = norm*I_1/9*(1+_q**3.5)**(3/2)
     pdf_q_analytic[_q<0.1] = 0
-    ax.plot(_q, pdf_q_analytic, label='analytic, $L=M^3$')
+    ax.plot(_q, pdf_q_analytic, label='analytic, $L=M^{3.5}$')
     ax.legend(loc='upper left', fontsize='small')
 
     ax.set(xlabel='$q = M_2/M_1$', ylabel='prob')

### 2017-09-25 src/joint_prob_r_q.py

@@ -3,6 +3,8 @@ compute and plot the joint 2d probability distribution function of a binary
 star's position and mass ratio in a magnitude-limited sample.
 '''
 import numpy as np
+import matplotlib as mpl
+mpl.use('Agg')
 import matplotlib.pyplot as plt
 from scipy.integrate import trapz
 
@@ -12,13 +14,13 @@ q = np.linspace(0,1,3e3)[::-1]
 L_1_N = 1
 F_lim_N = 1
 
-d_max_of_q = ( L_1_N / (4*np.pi*F_lim_N) * (1+q**3) )**(1/2)
+d_max_of_q = ( L_1_N / (4*np.pi*F_lim_N) * (1+q**3.5) )**(1/2)
 
 d_max_s = (L_1_N / (4*np.pi*F_lim_N))**(1/2)
-# integral of (1+q^3)^(-9/2)dq from q=0.1 to 1 evaluates to
-I_1 = 0.4645286925158471
+# integral of (1+q^3.5)^(-9/2)dq from q=0.1 to 1 evaluates to
+I_1 = 0.5060577377392849
 # integral of (1+q^3)^(3/2)dq from q=0.1 to 1 evaluates to
-I_2 = 1.323588493214896
+I_2 = 1.275894325140383
 # normalization constant brings integrals to 1
 norm = 9/(I_1*I_2)
 
@@ -93,13 +95,13 @@ im = ax.imshow(prob.T,
         aspect=0.5,
         cmap='Blues')
 
-cbar = plt.colorbar(im, fraction=0.046, pad=0.04)
-cbar.set_label(r'$\mathrm{prob}(r,q) \propto r^2$ if '+\
-        '$r<d_{\mathrm{max}}(q)$ and $q>0.1$', rotation=270, labelpad=12,
-        fontsize='small')
-
+ax.set_xlim([0,0.4])
 ax.set_xlabel('distance from origin, $r$ [arbitrary units]')
 ax.set_ylabel('$q = M_2/M_1$')
+
+cbar = plt.colorbar(im, fraction=0.052, pad=0.04)
+cbar.set_label(r'$\mathrm{prob}(r,q),\ \mathrm{magnitude\ limited}$', rotation=270,
+        labelpad=12, fontsize='small')
 
 f.tight_layout()
[…]

### 2017-09-25 paper/ms.tex
paper: add working draft

--- 2017-09-22 memos/workdir/toy_analytic_surveys.tex
@@
 \usepackage{todonotes}
+\usepackage{comment}
 
 \shortauthors{LGB, KM, JNW}
-\shorttitle{Binary Biases}
+\shorttitle{Binary Biases \& Toy Transit Surveys}
 
@@
 
-\title{ Binary Biases \& Toy Transit Surveys }
+\title{ The Effects of Stellar Binarity on Transit Survey Occurrence 
+Rates }
 
@@
 % \journalinfo{Draft version}
-\slugcomment{Memo for internal use}
+\slugcomment{Draft version}
 
@@
 
+\begin{comment}
 First, we show that in a universe where all stars have fixed properties
@@
 which Earth-like planets exist around Sun-like stars.
+\end{comment}
 
@@
 
-Binary companions to single stars introduce a surprising number of biases into 
-transit surveys.
-For a start, any magnitude limited survey over-represents binaries compared to 
-a volume-limited survey -- the familiar Malmquist bias, though in a slightly 
-different context.
-While transit surveys are ideally signal to noise limited, if the noise 
-properties of the survey are dominated by photon counting noise, a signal to 
-noise limit becomes equivalent to a magnitude limit.
-Thus stars surveyed in transit surveys must also over-represent binaries 
-compared to those in a volume-limited sample.
-
-A more straight-forward effect of binarity on transit surveys is that any star 
-which is unknowingly a double introduces extra stars about which planets might 
-be detectable.
-However, this apparent increase in the number of searchable stars is mitigated 
-by dilution: the observed transit depths in such systems are reduced by 
-contaminating flux from the binary companion. This dilution is especially 
-drastic if the planet orbits the secondary, in which case dilution impacts 
-detectability the most.
-Note that dilution can come from visually associated companions as well as 
-gravitationally bound companions.
-
-A consequence of dilution is misinterpretation of transit depths. If the 
-planet to star radius ratio is directly taken as the square-root of 
-the observed transit depth, derived planetary radii in binary systems will 
-systematically too small.
-Ciardi et al [2015] explored this issue in depth. They claimed that for the 
-case of \kepler, if ones assumes host stars are always single, one will
-underestimate radii by an average multiplicative factor of $\approx 1.5$ 
-(presumably, compared to the case of even weights for whether the 
-primary/secondary gets a given planet).
-
-Finally, binarity may affect transit surveys through actual astrophysical 
-processes.
-Duquennoy \& Mayor [1991], and alter Raghavan et al [2010] showed that over 
-half of solar-like are separated by less than 100\,AU.
-One immediate consequence on circumprimary and circumsecondary architectures is 
-that dynamical stability limits begin to carve away in realms of planet space 
-where planets might otherwise be detectable [e.g., Holman \& Weigert 1999, Li 
-et al 2017].
-The formation efficiency of various types of planetary systems may also 
-be different in binary systems compared to single-star systems.
-For instance, the properties of the circumstellar and circumbinary disks 
-influence the dynamics of planetesimals [Rafikov, Silsbee];
-
- the effects of which have been observed [Kraus et al 
-2016].
-Also, the formation rates of planets, and thus their intrinsic occurrence 
-rates, may be influenced by properties of circumprimary and/or circumbinary 
-disks.
-Although the observations are still ongoing, early results seem to indicate 
-that once binary companions are $\gtrsim 100\,{\rm AU}$ apart, binarity does 
-not decrease planet occurrence rates [CITE], and may even increase them for 
-giant planets [Ngo et al 2017].
+An astronomer who does not believe in binaries wants to measure the
+occurrence of planets of a certain type around stars of a certain type.
+In other words, they wish to find the mean number of planets within specific 
+planetary $(R_p, P)$ bounds orbiting the stars in a given volume 
+of stellar $(M_\star, R_\star, L_\star)$ phase-space.
+They perform a signal-to-noise limited transit survey and detect $N_{\rm det}$ 
+transit signals that appear to be from planets of the desired type.  They 
+calculate the number of stars $N_\star$ that appear to be searchable for the 
+desired type of planet, and also of the desired stellar type.  These 
+``searchable stars'' are the points on the sky for which they think planets 
+are observed with 100\% detection efficiency. Accounting for the geometric 
+transit probability $f_{\rm g}$, they compute an apparent occurrence rate 
+$\Gamma_{\rm a}$:
+
+\begin{equation}
+\Gamma_{\rm a} = \frac{N_{\rm det}}{N_\star} \times \frac{1}{f_g}.
+\end{equation}
+
+There are many potential pitfalls.  Some genuine transit signals can be missed
+by the detection pipeline.  Some apparent transit signals are spurious, from
+noise fluctuations, failures of `detrending', or instrumental effects.  Stars
+and planets can be misclassified due to statistical and systematic errors in
+the measurements of their properties.  Poor angular resolution causes false
+positives due to blends with background eclipsing binaries. {\it Et cetera}.
+
+Here we focus on problems that arise from the fact that stars of the desired
+type often exist in binaries. 
+We assume for simplicity that all binaries in the transit survey are 
+spatially unresolved.
+
+One immediate complication is that~--~due to dynamical stability or some 
+aspect of planet formation~--~ the occurrence rate of planets of the desired 
+type may differ between binary and single-star systems.
+If ``occurrence rate'' is defined purely as the mean number of planets within 
+set radius and period bounds per star in a given interval $\Delta M_\star, 
+\Delta R_\star, \Delta L_\star$, it must include an implicit marginalization 
+over stellar multiplicity.
+Otherwise, one must discuss ``occurrence rates in single star systems'', 
+``occurrence rates about primaries of double systems'', and
+``occurrence rates about secondaries of double star systems''.
+
+Outside of bonafide astrophysical differences, there are observational biases.
+A given apparently-searchable star may truly be a single star of the desired 
+type. If not, and ignoring higher order multiples, one of the following must 
+be true:
+\begin{enumerate}	
+\item The apparently-searchable star is a binary system, with two searchable 
+stars of the desired type. $N_\star$ is under-counted by one for every such 
+system.
+%
+\item The apparently-searchable star is a binary system, with one searchable 
+star of the desired type. 
+The searchable star of the desired type could be the primary or the secondary.
+$N_\star$ is correctly counted for every such system, though there may be 
+systematic errors in estimates of stellar parameters.
+%
+\item  The apparently-searchable star is a binary system, in which neither of 
+the components is a searchable star, but at least one is of the desired type. 
+$N_\star$ is over-counted by one for every such system.
+\end{enumerate}
+
+One might also imagine an apparently-searchable star which in no stellar 
+component is of the desired type.
+For single stars, this may simply be from erroneous stellar parameters.
+In binaries, one might imagine the combined light of two stars somehow making 
+them resemble a single star of the desired type.  We will 
+not consider errors of this type.  We will assume that all the 
+apparently-searchable stars are either single stars of
+the desired type, or binaries in which either the primary or else both 
+components are of the desired type.
+
+The above enumerates possible errors when counting $N_\star$.
+When tallying $N_{\rm det}$, the detected signals from planets of the desired 
+type, neglecting binarity introduces the following error cases:
+\begin{enumerate}
+\item 
+A detected signal is from a star of undesired type (by the above assumptions, 
+the secondary of a binary system), from a planet of undesired type, and is 
+incorrectly counted as a planet of desired type.
+%
+\item
+A detected signal is from a star of desired type (the primary or secondary of 
+a binary systems), from a planet of undesired type, and is incorrectly counted 
+as a planet of desired type.
+This and the case above occur when the host star parameters are miscalculated 
+({\it e.g.,} the host star is the faint secondary, but is assumed to be 
+primary), or when the constant diluting light from the binary companion is not 
+corrected.
+$N_{\rm det}$ should be lowered by 1 for every such case.
+%
+\item
+A detected signal is from a star of desired type, from a planet of desired 
+type, but incorrectly counted as planet of undesired type.
+$N_{\rm det}$ should be raised by 1 for every such case.
+\item
+An {\it undetected} signal from a planet of the desired type, around a star of 
+the desired type, was not detected because of dilution from the companion 
+(because the SNR floor of the survey is set for single stars). 
+This happens during case \#3 in $N_\star$-counting errors (the 
+apparently-searchable star is a binary system, in which neither component is 
+in fact searchable). If $N_\star$ is corrected for this, {\it i.e.,} the 
+system is correctly counted as ``not searchable'', no correction to $N_{\rm 
+det}$ is needed.
+\end{enumerate}
+
+Errors (1)-(3) in calculating $N_{\rm det}$ above broadly fall under ``radius 
+misclassification'', while error (4) falls under ``completeness 
+miscalculation''.
+Note that in the above list, a planet cannot be both ``of desired type'' and 
+``orbiting a star of undesired type''. We ignore this type of error for 
+analytic convenience, but \texttt{reintroduce it later during numerics}.
+
+Our approach is step-by-step starting from a very simple scenario.
+\begin{itemize}
+\item We consider a universe consisting of only the desired types of stars and
+planets.  We give analytic expressions for eta(apparent)/eta(true) taking
+into account all applicable errors.
+%
+\item We allow for a distribution of the mass ratio $dn/dq \propto q^\gamma$, 
+and a concurrent power-law distribution in the luminosity ratio, $dn/dL 
+\propto L^\beta$. 
+%We assume the secondary is not of the desired stellar type.
+%NOTE: LB notes that yes, while analytically simple, it doesn't make sense. 
+%For the twin binaries case, what possible reason could there be for the 
+%secondary to not be of the desired stellar type? Similar for the high mass 
+%ratio binaries -- the stars are ~identical.
+We give analytic expressions for eta(apparent)/eta(true) taking into 
+account errors of type 2.
+%NOTE: empirical L(M) maybe OK for numerics -- not for analytics.
+%
+\item We allow for a nonzero tolerance $\Delta R_p$ in the planet radius, and 
+tolerance $\Delta L_\star$ in stellar luminosity, to qualify as "desirable".  
+We recalculate eta(apparent)/eta(true).
+(LB: I think this has to become numerical to actually evaluate the occurrence
+rate correction factor, $X_\Gamma$.)
+%
+\item We assume the planets have a power-law radius distribution, $d\eta/dR_p
+\propto r^\alpha$.  We calculate the apparent $d\eta/dR_p$ taking into account 
+errors of type 1-3.  We do so for stars of the desired type within some 
+tolerance dL.
+%
+\item Everything up to now is more-or-less analytic, perhaps supported by 
+numerical checking or numerical integration. At the end we summarize the 
+radius-dependence of the occurrence rate shifting.
+"Accounting only for dilution \& its numerator effect, big radius planets have
+underestimated inferred rates, the smallest radius planets have overestimated
+inferred rates, and the in-between radius planets are in-between"
+
+\end{itemize}
+
+We then dig in, w/ numerics, to what happens with different assumed
+completeness vs Rp functions.
+We note that the the claimed HJ occ rate discrepency btwn RV \&
+transit surveys gets much weaker when accounting for binarity.
+This is b/c the more believable occ rate papers do not heavily rely on their
+completeness calibration -- they know they're complete.
+This is a much more understandable regime than the wonky completeness estimate 
+dependent zone of eta-Earth for Kepler.
+
+Finally, for numerics, we simulate a sample of 20,000 
+apparently-searchable-stars (or however many "stars" appear to be searchable
+for Earthlike planets in the Kepler field).  Basically this means a magnitude
+limited survey of 20,000 ``stars'' that are either single or are binaries with
+the primary of the desired type.  The sample has a realistic distribution of
+binaries (occurrence and luminosity ratio).  We calculate 
+eta(apparent)/eta(true).  We can also play with errors of type 4 in this
+numerical model (i.e. we let the relative occurrence rates between primaries 
+and secondaries vary).  (It may not worth doing so for the analytic 
+work-- though it's pretty simple. The way to play with this is just present 
+both limits, and then also present a few weight factors (Furlan+17 do a good 
+job).
+
+We discuss ``how do we interpret Fulton+ 2017?''
+That paper did (or claims to have done) a good job at weeding out numerator
+errors. If true, means $\Gamma(R_{\rm p})$ has the right shape, and just the 
+wrong normalization. Important if we ever manage to convince ourselves 
+``brightest transit hosts'' could work.
+
+In passing, at the end, we mention how this changes attempts at eta-Earth 
+measurements.
+
+\begin{comment}
+UNSOLVED PROBLEMS W/ THE NUMERICAL MODEL-
+* currently have not implemented self-consistent (L,M,R) relation between
+primaries and secondaries.
+Wang et al (2015) avoided that problem by doing a similar empirical thing to
+what I currently have implemented. But I do not think that was a good paper.
+Ciardi et al (2015) have an approach that was Kepler-specific -- they used
+the same Dartmouth isochrones used for KIC stellar parameters.
+But they did not conserve $dn/dm_x$, for $m_x$ apparent mag in band x.
+I think the latter choice might be better for Kepler-specific discussion.
+But it's less-nice if we want this to be general enough for e.g., occ HJ
+rates w/ TESS.
+
+My main gripe w/ taking the 150k KIC stars \& adding binaries is that less
+is known about those parameters and biases. Using e.g., Galaxia has the major
+benefit of you understand what's been cooked in much better (it's minimal).
+
+I think if we want to talk about Kepler, it's the best possible approach
+though.
+
+If we want to talk about Fulton+17 we can go further, and apply the CKS cuts.
+
+Some questions based on the current draft:
+
+- Why should we ever consider $F_{lim} < F_{min}$?
+
+LB: in practice, surveys look at many stars for which they could not detect
+Rp=Re planets, but could detect Rp=10Re planets.  One possible error is
+miscounting which stars are which (i.e. getting completeness corrections
+wrong). So this was more for plausbility than b/c it's immediately needed in
+the analytics. I can cut it.
+
+- Why do we need to consider the distribution of S/N, as opposed to simply the
+limiting distance for detection?
+
+LB: for "model 1" (fixed stars, fixed planets, fixed light ratio binaries)
+you're correct, it's not immediately necessary.
+
+For this model, I did it as a way to think through the completeness effects
+systematically -- i.e. how for the twin-binary case, you get a completeness
+fraction of 1/8, though your limiting distance is sqrt(2) times bigger. (Even
+when your completeness for the singles is 100%).
+
+If we go the path of completing "model 2"'s analytics (vary light ratio, fixed
+primary luminosity, fixed planet), I think this formulation would become
+necessary in order to actually evaluate the completeness fractions.
+
+(I never completed these analytics b/c of the L(M) mess + confusion about
+assumed stellar masses+radii. Power law L(M) helps this... I think there needs
+to be an extra assumption about the stellar mass+radius change though that we
+have not spelled out)
+
+- I don't think the bandpass, effective area, photon count rate, etc.  are
+needed, either. All that matters is the limiting distance.
+
+Are needed for what? 
+In Model 1, for analytic expressions of the occurrence rate, I agree.
+For analytic expressions of the number of detected planets, I disagree.
+
+For Model 2, I'm not sure because I got bogged down. I suspect that for
+occurrence rate errors, you are correct.
+
+The analytic expression I gave for the occurrence rate depends on this "xi"
+factor, which needs to incorporate an assumption about how Astronomer A
+misestimates the completeness (relatedly, how he misestimates Rstar).  I never
+specified that assumption.
+
+- In Eq. 4, is this a traditional definition of BF?  I would have defined it as
+nd/(ns+nd) or perhaps 2*nd/(ns+ 2*nd).  Let's make sure we are using
+traditional terminology, and also interpreting the literature correctly, i.e.,
+is the right number really 0.45?
+
+Oops. nd/(ns+nd) = BF = MF (if assuming all multiple systems are binaries).
+This minor change propagates.
+
+Touching up on correct terminology, from Duchene \& Kraus 2013, Sec 2.1:
+
+"Frequency of multiple systems" (MF)
+vs
+"Companion frequency" (CF; average \# of companions per target, can exceed 
+100\%)
+
+my understanding of the terms:
+
+survey 1: {*, **, *, ***, *, *}
+
+6 systems.
+2 are multiple.
+MF = 2/6 = 1/3
+CF = (0+1+0+2+0+0)/6 = 3/6 = 1/2
+
+survey 2: {**, **, ***, ****, *}
+
+5 systems
+4 are multiple
+MF = 4/5
+CF = (1+1+2+3+0)/5 = 6/5
+
+DK2013 Sec 3.1.1 give (citing Raghavan+ 2010)
+MF^{MS}_{1-1.3Msun} = 50 \pm 4%
+MF^{MS}_{0.7-1Msun} = 41 \pm 3%
+
+They also give
+CF^{MS}_{1-1.3Msun} = 75 \pm 5%
+CF^{MS}_{0.7-1Msun} = 56 \pm 4%
+
+I understand the first pair of numbers to mean:
+
+In a volume-limited survey (of the stellar neighborhood), if you select systems
+for which the primary is from 1-1.3Msun, 50% of them will be multiple.
+
+I take the second pair of numbers to mean:
+
+In a volume-limited survey (of the stellar neighborhood), if you select systems
+for which the primary is from 1-1.3Msun, each system will on average have 0.75
+companions.
+
+So YES, need to correct this throughout.
+\end{comment}
 
@@
 
-Binarity introduces several convoluted effects on transit 
-surveys. To disentangle them, imagine the following idealized survey:
-
-\begin{enumerate}
-\item You are going to observe the entire sky for a duration $T_{\rm obs}$, 
-with a detector of area $A$, and known bandpass. Your detector is photon-noise 
-limited.
-%
-\item You are interested only in detecting planets of radius $R_p$, and orbital 
+Stellar binarity introduces convoluted effects on transit surveys.
+To disentangle them, imagine the following idealized survey.
+
+You observe the entire sky for some duration with a detector of known 
+area and bandpass. Your detector is photon-noise limited.
+You are interested only in detecting planets of radius $R_p$, and orbital 
 period $P$. For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
-%
-\item You are only interested in detecting them around stars of radius $R_1$, 
+You are only interested in detecting them around stars of radius $R_1$, 
 and luminosity $L_1$. For instance, G2V dwarfs.
-%
-\item You can only detect your `nominal planet' when you observe signals with
-${\rm S/N} > {\rm (S/N)_{min}}$.
+You can only detect your desired type of planet when you observe signals 
+with ${\rm S/N} > {\rm (S/N)_{min}}$.
 For a photon-noise limited survey, this minimum signal to noise ratio is 
-equivalent to a minimum flux, $F_{\rm min}$.
-%
-\item However, you opt to observe all the ``points'' on-sky with apparent 
-magnitude $m < m_{\rm lim}$, or equivalently with energy flux in your bandpass 
-greater than some limit $F > F_{\rm lim}$.
-(You will not be able to detect planets around the faintest of 
-these, and will need to perform a completeness correction to later derive 
-occurrence rates.)
-Generally $F_{\rm lim} < F_{\rm min}$, \textit{e.g.}, whenever the original 
-magnitude cut defines a ``source catalog'' from which ``searchable stars'' are 
-selected, as in the KIC [Batalha et al, 2010, Brown et al 2011] and the TIC 
-[Stassun et al 2017].
-\end{enumerate}
-
-You get funding, and perform the S/N-limited survey. All planets with 
-${\rm S/N} > {\rm (S/N)_{min}}$ are detected. The rest are not.
+equivalent to a minimum flux required for detection, $F_{\rm min}$.
+
+You get funding, and perform the S/N-limited survey. 
+You observe all the ``points'' on-sky with apparent 
+magnitude $m < m_{\rm min}$, or equivalently with energy flux in your bandpass 
+greater than some limit $F > F_{\rm min}$.
+All planets with ${\rm S/N} > {\rm (S/N)_{min}}$ are detected. The rest are 
+not.
+
 You now wish to derive an occurrence rate for planets of radius $R_p$ and 
@@
 	\item The true population of ``points'' (stellar systems, all unresolved) 
-	comprises both single and double star systems. Single star systems have 
+	comprises only single and double star systems. Single star systems have 
 	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
@@
 	the secondary to the primary. 
-	In this section, $\gamma_R$ is a constant across the population 
-	of star systems.
-	The ratio of the two number densities in a 
-	volume-limited sample is the binary fraction\footnote{The binary fraction 
-		is equivalent to the multiplicity fraction if there are no triple, 
-		quadruple, $\ldots$ systems.}.
+	In this section, $\gamma_R=1$ across the population 
+	of star systems~--~all binaries are assumed to be twin systems.
+	The ratio of the number density of binary systems and the number density of 
+	all systems in a volume-limited sample is the binary fraction\footnote{The 
+	binary fraction is equivalent to the multiplicity fraction if there are 
+	no triple, quadruple, $\ldots$ systems.}.
 	\item The true population of planets around these stars is as follows:
-	\subitem A fraction $\Gamma_{t,s}$ of stars in single star systems 
-	have a planet of radius $R_p$, with orbital period $P$.
-	\subitem A fraction $\Gamma_{t,d}$ of each star in a double star 
-	system has a planet of radius $R_p$, with orbital period $P$. For instance, 
-	if $\Gamma_{t,s} = \Gamma_{t,d} = 0.1$, on average each double 
-	system contributes 0.2 planets, and each single system 0.1 planets.
+	\begin{itemize}
+	\item A fraction $\Gamma_{t,s}$ of stars in single star systems 
+	have planets of radius $R_p$, with orbital period $P$.
+	\item A fraction $\Gamma_{t,d}$ of primary stars in double star 
+	systems have planets of radius $R_p$, with orbital period $P$. 
+	A fraction $w_{d2}\Gamma_{t,d}$ of secondary stars in a double star 
+	systems have the same type of 
+	planet, where the weight factor $w_{d2}$ specifies the secondary star 
+	occurrence rate as a fraction	of the primary star occurrence rate.
 	Any astrophysical difference in planet formation between singles and 
-	binaries is captured by these two terms.
-	Note that we are taking an limit by assuming the primary and secondary of 
-	a binary system host planets at equal rates. If the secondary is 
-	non-identical to the primary, it's clearly wrong.	
+	binaries is captured by $\Gamma_{t,s}, \Gamma_{t,d},$ and $w_{d2}$.
+	If one assumes the primary and secondary of a binary system host planets at 
+	equal rates, this is the $w_{d2}=1$ limit.
+	Similarly, if one assumes  secondaries 
+	cannot host planets, this is the $w_{d2}=0$ limit.
 	\end{itemize}
-
-For simplicity, one might wish to imagine the observer correctly 
-pre-selects all of the ``searchable stars'', in the style of [Pepper et al, 
-2003].
-In other words, they would begin by knowing ${\rm (S/N)_{min}}$, and then only 
-observe stars for which they would have adequate flux to achieve it -- setting 
-$F_{\rm lim} = F_{\rm min}$.
-This simplification would naively allow the observer to ignore 
+\end{itemize}
+
+The assumption of [Pepper et al, 2003] in a very similar context was that the 
+observer correctly pre-selects all of the ``searchable stars''.
+This simplification would allow the observer to ignore 
 completeness effects when deriving occurrence rates, since their completeness 
@@
 
-We do not begin this way because in a stellar population of both singles and 
-binaries, the above approach ignores necessary incompleteness for 
-binary systems.
+The above approach ignores incompleteness for binary systems.
 If we set the limiting flux to give 100\% completeness for single stars, we 
@@
 will push transit signals below the detection threshold.
-This is because a flux limit maps onto three different maximum 
+This is because a flux limit maps onto three distinct maximum 
 detection distances for a population of single stars and twin binaries:
 \begin{enumerate}
-\item $d_{{\rm max},s}$: the maximum distance out to which single stars are
-  selected. This can be chosen to coincide with the maximum distance out to
-  which planets are detectable around single stars.
+\item $d_{{\rm max},s}$: the maximum distance out to which the desired type 
+	of planet is detectable around single stars.
+	These are the selected ``searchable, single stars''.
 %
-\item $d_{{\rm max},d}^\star$: the maximum distance out to which double stars
-  are selected. This is different from $d_{{\rm max},d}^{\rm p}$, the maximum
+\item $d_{{\rm max},d}$: the maximum distance out to which double stars
+  are selected (down to the same minimum flux as that defining $d_{{\rm 
+  max},s}$). 
+  This is different from $d_{{\rm max},d}^{\rm p}$, the maximum
   distance out to which planets are detectable around double stars. For a 
@@
     d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
-    (1 + \gamma_R)^{-1} d_{{\rm max},d}^\star.
-  \end{equation}
-  For $\gamma_R = 1$, this means only 1 in 8 selected binary stars can yield
-  detectable planets.
+    (1 + \gamma_R)^{-1} d_{{\rm max},d}.
+  \end{equation}
+  For $\gamma_R = 1$, this means only 1 in 8 apparently-searchable stars which 
+  are in fact binary systems can yield detectable planets.
+
 \end{enumerate}
-We develop tools to address this and related completeness effects from the 
-outset.
-
-Consider the following questions:
-\begin{enumerate}
-\item How many single and double star systems, respectively, are in the sample?
-Correspondingly, how many stars are in the sample?
-
-\item How many planets are in the sample? (Orbiting single stars, and orbiting 
-double stars respectively).
-
-\item What is the true occurrence rate?
-
-\item How many planets are detected?
-
-\item What occurrence rate does astronomer A, who has never heard of binary 
-star systems, or completeness, derive for planets of radius $R_p$ and period 
-$P$?
-
-\item What occurrence rate does astronomer B, who accounts for the ``2 for 1'' 
-effect of binarity (\textit{i.e.} that the sample actually has more stars than 
-astronomer A thought) derive? (Astronomer B neglects completeness).
-
-\item What about astronomer C, who accounts for ``2 for 1'' \textit{and} 
-misclassification due to diluted radii? In other words, astronomer C did a 
-combination of high resolution imaging and RV followup on every candidate, and
-correctly classifies the planetary radii in every case.
-
-\item What about astronomer D, who additionally notes the importance of 
-completeness?
-\end{enumerate}
-
-\subsection{How many stars are in the sample?}
-
-Let $N_s$ be the number of single star systems, and $N_d$ the number of double 
-star systems. Then the total number of stars in the sample is
-\begin{equation}
-N_{\rm stars} = N_s + 2N_d.
-\end{equation}
-
-In a magnitude-limited sample in which stars are uniformly distributed in 
-volume, the number of stars will be the number density times the volume.
-If the volume is taken to be a sphere over which the number density is uniform,
+
+We note in passing that we are considering a SNR-limited survey, which in this 
+case is the same as a magnitude-limited one.
+The observer might prefer to perform a completeness-limited survey, which in 
+this case would be the same as a volume-limited one with limiting distance
+$d_{{\rm max},d}^{\rm p}$.
+This change would likely improve interpretability.
+
+With the stage set, we ask:
+
+\subsection{What is the true occurrence rate?}
+\label{sec:true_rate}
+
+The ``true occurrence rate'' $\Gamma_t$ is the average number of planets of 
+the desired type per star of the desired type:
+\begin{align}
+\Gamma_t &= \frac{N_{\rm planets}}{N_{\rm stars}} \\
+\Gamma_t &= \frac{\Gamma_{t,s} N_s + (1+w_{d2}) \Gamma_{t,d} N_d}{N_s + 2N_d}.
+\label{eq:true_occ}
+\end{align}
+where $N_s$ is the number of single star systems, and $N_d$ the number of 
+double star systems. 
+The factor of 2 in the total number of stars accounts for the fact that there 
+are twice as many stars in double star systems.
+
+To develop more analytic expressions in terms of observables, assume that the 
+stars are homogeneously distributed\footnote{
+If we wished to write a stellar number density profile that accounted for the 
+vertical structure of the Milky Way, we might choose a profile either $\propto 
+\exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from the 
+galactic midplane and $H$ a scale-height. Both density profiles would lead 
+closed form analytic solutions.	}.
+Counting the number of single and double systems in their respective selected 
+volumes,
 \begin{equation}
@@
 \begin{equation}
-\frac{n_{d}}{n_{s}} = {\rm binary\ fraction} \equiv {\rm BF}
-\end{equation}
-by definition. The absolute normalization of the number density is a measured 
-quantity, as is the binary fraction. For G2V dwarfs, the latter is $\approx 
-0.45$ [Duchene \& Kraus, 2013]. The former is given by [Bovy 2017].
+\frac{n_{d}}{n_s+n_d} = {\rm binary\ fraction} \equiv {\rm BF}
+\end{equation}
+The absolute normalization of the number density $n_s + n_d$ is an observed
+quantity [Bovy 2017], as is the binary fraction.
+For ``sun-like'' dwarfs with $0.7<M_\star/M_\odot<1.3$,
+${\rm BF}=0.44\pm0.02$ [Duchene \& Kraus, 2013, Raghavan et al 2010]. 
+Note that these latter authors really quote a {\it multiplicity} fraction of
+0.44 for sun-like dwarfs~--~we count the $\approx 25\%$ of multiple 
+systems that are higher order multiples as doubles.
 
@@
 \end{equation}
-where the limiting flux in the bandpass $F_{{\rm min}}$ can also be 
-stated in terms of the limiting magnitude $m_{{\rm min}}$,
-\begin{equation}
-m_{{\rm min}} = m_{0} - \frac{5}{2} \log_{10} \left(\frac{F_{{\rm 
-min}}}{F_{0}}\right),
-\end{equation}
-for $m_{0}$ a zero-point magnitude and $F_{0}$ its corresponding flux (as 
-always, everything is implicitly written in a defined bandpass).
-
-In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$, and as a 
-consequence the maximum distance to which binary stars will be selected is 
-greater than that of single stars, simply as a consequence of imposing a 
-magnitude cut.
-The ratio of double to single systems is
+where the limiting flux in the bandpass $F_{{\rm min}}$ could equivalently be 
+stated in terms of a limiting magnitude $m_{{\rm min}}$.
+In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$. Simply as a 
+consequence of imposing a magnitude cut, the maximum distance to which binary 
+stars will be selected is greater than that of single stars.
+The ratio of double to single systems in the observed SNR-limited sample is 
+then
 \begin{align}
 \frac{N_d}{N_s} &= 
-	\frac{n_d}{n_s} \left(\frac{d_{{\rm max}, d} }{d_{{\rm max}, s}}\right)^3 \\
-&= {\rm BF} \times (1+\gamma_R)^{3/2}.
+	\frac{n_d}{n_s}
+	\left( \frac{d_{{\rm max}, d} }{d_{{\rm max}, s}} \right)^3 \\
+&= \frac{{\rm BF}}{1 - {\rm BF}} \times (1+\gamma_R)^{3/2}.
 \end{align}
-In the nominal case of twin binaries ($\gamma_R = 1$), with a binary fraction 
+In the case of twin binaries ($\gamma_R = 1$), with a binary fraction 
 ${\rm BF} = 0.5$, there are 
-$\sqrt{2}$ more binary systems than single systems in the sample.
-Correspondingly, there are $2\sqrt{2}$ more stars in binary systems than stars 
-in single systems.
-
-As a comment on Eq.~\ref{eq:number_systems}, if 
-we wished to write a stellar number density profile that accounted for the 
-vertical structure of the Milky Way, we might choose a profile either 
-$\propto \exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from 
-the galactic midplane and $H$ a scale-height. Both density profiles would lead 
-closed form analytic solutions.
-
-\subsection{How many planets are in the sample?}
-
-The number of planets in the sample is
-\begin{align}
-N_{\rm planets} =& N_{\rm planets\ in\ single\ star\ systems}  +  \\
-				  &\quad\quad N_{\rm planets\ in\ double\ star\ systems} 
-				  \nonumber \\
-			   =& \Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d.
-\end{align}
-
-The factor of 2 accounts for the fact that there are twice as many stars in 
-double star systems.
-
-\subsection{What is the true occurrence rate?}
-\label{sec:true_rate}
-
-The ``true occurrence rate'' is the average number of planets per star. Thus
-
-\begin{align}
-\Gamma_t &= \frac{N_{\rm planets}}{N_{\rm stars}} \\
-\Gamma_t &= \frac{\Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d}{N_s + 2N_d}.
-\label{eq:true_occ}
-\end{align}
-
-\subsection{How many planets are detected?}
-The total number of planet detections is the sum of the number of planets 
-detected in single star systems $N_{{\rm det},s}$ and the number of planets 
-detected in double star systems $N_{{\rm det},d}$.
-These can be expressed individually. 
-The former is
-\begin{equation}
-N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,{\rm geom}} f_{s,{\rm S/N} > {\rm 
-(S/N)_{min}}},
-\label{eq:N_det_s}
-\end{equation}
-where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
-star systems of the sample, $f_{s,{\rm geom}}\equiv f_{s,g}$ is the geometric 
-transit probability, and $f_{s,({\rm S/N} > {\rm (S/N)_{min}})} \equiv f_{s,c}$ 
-is the fraction of these transiting planets that are observed with signal to 
-noise greater than the minimum detection threshold (the completeness). 
-Analogously,
-\begin{equation}
-N_{{\rm det},d} = 2 N_d \Gamma_{t,d} f_{d,g} f_{d,c},
-\label{eq:N_det_d}
-\end{equation}
-where now $2 N_d \Gamma_{t,d}$ is the number of planets in the double star 
-systems of the sample, the geometric transit probability is the same (if we 
-have twin binaries -- not if we consider a distribution of secondaries) and the 
-completeness term must account for any differences in the signal to noise 
-distribution that come from stellar binarity.
-
-\subsubsection{Analytic completeness}
-Since the geometric transit probability is known, the only terms we have yet to 
-compute are the completeness terms, $f_{i,c}$ for $i\in{\rm \{single,double\} 
-}$. We proceed as follows.
-
-The signal ${\rm S}$ for a box-car train transiting planet is
-\begin{align}
-{\rm S} &= \delta \mathcal{D} \\
-&= \left(\frac{R_p}{R_\star}\right)^2 \mathcal{D},
-\end{align}
-for $R_p$ the planet's radius, $R_\star$ that of its host star, and 
-$\mathcal{D}$ the dilution parameter defined as
+$2^{3/2}\approx2.8$ times more binary systems than single systems in the 
+SNR-limited sample.
+Since the completeness fraction for binary systems is $1/8$, there are 
+$2^{-3/2}\approx0.35$ times fewer binary systems in the sample that can yield 
+planets as single star systems in the system that can yield planets.
+
+\subsection{What is the measured occurrence rate?}
+% % % % % % % % % % % %
+
+The observer who has never heard of binary star systems wants to find the 
+occurrence rate of planets with radius $R_p$ and period $P$, orbiting stars of 
+radius $R_1$ and luminosity $L_1$.
+They detected $N_{\rm det}$ transit signals that appear to be from planets of 
+this desired type.
+The occurrence rate they then find, {\it i.e.} the number of planets divided 
+by number of ``stars'' (really, stellar systems) is
+\begin{equation}
+\Gamma_{\rm a } = \frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_g}.
+\end{equation}
+
+Since only the transit depth is observed, the planetary radii apparent to this 
+astronomer, $R_{\rm p, a}$, are different from the true planetary radii 
+$R_{\rm p, t}$. 
+This is because the assumed stellar radii $R_{\star,{\rm a}}$ may differ from 
+the true stellar radii $R_{\star,{\rm t}}$, and the flux will be diluted:
+\begin{equation}
+R_{p,{\rm a}} = R_{p,{\rm t}} 
+\frac{R_{\star,{\rm a}}}
+		 {R_{\star,{\rm t}}}
+		 \,\mathcal{D}^{1/2},
+\end{equation}
+where the dilution parameter $\mathcal{D}$ is defined for binary systems as
 \begin{align}
@@
 \Bigg\{\begin{array}{lr}
-  L_1 / L_d, & {\rm\ if\ binary\ and\ target\ primary}\\
-  \gamma_R L_1 / L_d, & {\rm\ if\ binary\ and\ target\ 
-  secondary}\\
-  1, & {\rm\ if\ single},
+L_1 / L_d, & \text{if\ planet\ orbits\ primary}\\
+\gamma_R L_1 / L_d, & \text{\ if\ planet\ orbits\ secondary}
 \end{array} \nonumber\\
@@
 \Bigg\{\begin{array}{lr}
-    (1+\gamma_R)^{-1}, & {\rm\ if\ binary\ and\ target\ primary}\\
-    (1 + \gamma_R^{-1})^{-1}, & {\rm\ if\ binary\ and\ target\ 
-    	secondary}\\
-    1, & {\rm\ if\ single},
+(1+\gamma_R)^{-1}, & \text{if\ planet\ orbits\ primary}\\
+(1 + \gamma_R^{-1})^{-1}, & \text{if\ planet\ orbits\ secondary}
 \end{array} 
@@
 
-Assuming the only source of noise is Poissonian counting noise, the noise ${\rm 
-N}$ can be written
-\begin{equation}
-{\rm N} = \frac{1}{\sqrt{N_\gamma}},
-\end{equation}
-for $N_\gamma$ the number of photons received by the detector. This noise model 
-is a useful simplification -- see [Howell 2006, pg 75] for the full 
-CCD equation.
-The number of received photons can be written
-\begin{equation}
-N_\gamma = F^{\rm N}_\gamma A N_{\rm tra} T_{\rm dur},
-\end{equation}
-for $F^{\rm N}_\gamma$ the photon number flux from the system [${\rm 
-ph\,cm^{-2}\,s^{-1}}$], 
-$A$ the detector area, $T_{\rm dur}$ the transit duration, and $N_{\rm tra} $ 
-the number of transits observed, which is multiplied in assuming the transits 
-are ``phase-folded''.
-
-Thus the signal to noise ratio can be written
-\begin{equation}
-{\rm S/N} = \delta \mathcal{D} \sqrt{F^{\rm N}_\gamma A N_{\rm tra} T_{\rm 
-dur}}.
-\label{eq:snr_ivory_tower}
-\end{equation}
-
-This means we can write the minimum number flux of photons required for a 
-detection at threshold as
-\begin{equation}
-F_{\rm lim}^N = \left[ \left({\rm \frac{S}{N}}\right)_{\rm min} \frac{1}{\delta 
-	\mathcal{D}} \right]^2 \frac{1}{A N_{\rm tra} T_{\rm dur}}.
-\end{equation}
-To convert this to $F_{\rm lim}$, multiply by the average photon energy in the 
-bandpass.
-
-In passing, given the parameters that define a survey and planet type, 
-Eq.~\ref{eq:snr_ivory_tower} would need to be re-expressed with 
-$N_{\rm tra}$ roughly the ratio of the observing baseline to the planet period, 
-and $T_{\rm dur}$ a function of $R_\star, P, a$, and impact parameter $b$, and 
-then perhaps averaged over $b$. We leave them as-is for subsequent 
-development.
-
-The interesting term in Eq.~\ref{eq:snr_ivory_tower} that changes between stars 
-of the same binarity class in our idealized sample is the square root of 
-$F^{\rm N}_\gamma$.
-This is the term that leads to a distribution of signal to noises for different 
-stars.
-The completenesses $f_{i,c}$ can be directly 
-expressed in terms of those probability density functions:
-\begin{equation}
-f_{i,c} = 
-	\int_{{\rm (S/N)_{min}}}^{\infty} 
-		{\rm d}\left(\frac{{\rm S}}{{\rm N}}\right)_i \ 
-		{\rm prob}\left(\frac{{\rm S}}{{\rm N}}\right)_i.
-\label{eq:completeness_long}
-\end{equation}
-We keep the subscript $i$ because the signal to noise distributions are 
-different for the cases of single star systems ($i=s$) and double star systems 
-($i=d$).
-Notably:
-\begin{itemize}
-	\item The dilution differs (Eq.~\ref{eq:dilution}).
-	\item The photon number flux from the system differs.
-\end{itemize}
-
-To simplify notation, we let $x_i \equiv {\rm (S/N)}_i$, and rewrite 
-Eq.~\ref{eq:completeness_long} as
-\begin{equation}
-f_{i,c} = 
-\int_{{\rm x_{\rm min}}}^{\infty} 
-{\rm d}x_i \ 
-{\rm prob}(x_i).
-\label{eq:completeness_short}
-\end{equation}
-
-\subsubsection{Deriving ${\rm prob}(x_i)$}
-We want expressions for the probability density function of the observed signal 
-to noise ratio, ${\rm prob}(x_i)$, for both the single and binary 
-system case.
-
-First, note that a star placed uniformly in the volume of the search space will 
-have a probability density function for its distance $r$ from the origin of
-\begin{equation}
-{\rm prob} (r) = \frac{3 r^2}{d_{\rm max}^3},
-\end{equation}
-where the appropriate maximum distances should be substituted per 
-Eq.~\ref{eq:d_max}.
-Noting the transformation rule for probability density functions, we can 
-evaluate the probability of a star having a observed number flux $F^{\rm 
-N}_{i,\gamma}$ in the bandpass,
-\begin{align}
-{\rm prob} (F^{\rm N}_{i,\gamma}) 
-&= {\rm prob}(r(F^{\rm N}_{i,\gamma}))
-	\left| \frac{{\rm d} r}{{\rm d} F^{\rm N}_{i,\gamma}} \right| \\
-&= \frac{3}{2 d_{\rm max}^3} c_i^{3/2} (F^{\rm N}_{i,\gamma})^{-5/2},
-\label{eq:pdf_observed_flux}
-\end{align}
-where in the latter equality we have written a ``number luminosity'' $c_i$ 
-(units of inverse time) defined for $i\in{\rm \{single,double\} }$ as
-\begin{equation}
-c_i = 
-\Bigg\{\begin{array}{lr}
-R_1^2 F^{\rm N}_{s1,\gamma}, & {\rm\ if\ single}\\
-R_1^2 F^{\rm N}_{s1,\gamma} + R_2^2 F^{\rm N}_{s2,\gamma} & {\rm\ if\ double}.
-\end{array}
-\label{eq:c_i}
-\end{equation}
-In Eq.~\ref{eq:c_i}, $F^{\rm N}_{s1,\gamma}$ and $F^{\rm N}_{s2,\gamma}$ are 
-the photon number fluxes at the surfaces of the stars. To derive 
-Eq.~\ref{eq:pdf_observed_flux}, we simply scaled these by the distance:
-\begin{equation}
-F^{\rm N}_{i,\gamma} = \frac{c_i}{r^2}.
-\label{eq:flux_at_detector}
-\end{equation}
-
-The surface photon number fluxes $F^{\rm N}_{si,\gamma}$ in Eq.~\ref{eq:c_i} 
-are usually evaluated numerically, by convolving the wavelength-specific photon 
-flux density of a star with the dimensionless spectral response function of the 
-instrument. In other words,
-\begin{equation}
-F^{\rm N}_{s,\gamma} = \int F_\lambda T_\lambda \, {\rm d}\lambda.
-\label{eq:pfd}
-\end{equation}
-The wavelength-specific photon flux density $F_\lambda$ [${\rm 
-ph\,cm^{-2}\,s^{-1}\,\AA^{-1}}$] could be from Pickles' library [1998, PASP 
-110, 863], or could be a blackbody function. The transmission function is, up 
-to factor of order 
-unity, a step function between two wavelengths $\lambda_{\rm min}$ and 
-$\lambda_{\rm max}$.
-If we assume a blackbody source, Eq.~\ref{eq:pfd} becomes
-\begin{equation}
-F^{\rm N}_{s,\gamma} = 8\pi c \left(\frac{k T}{h c}\right)^3 
-\int_{hc/(\lambda_{\rm max} kT)}^{hc/(\lambda_{\rm min} kT)}
-\frac{u^2}{e^u - 1} \,{\rm d} u,
-\end{equation}
-which can be evaluated numerically\footnote{It may help in the numerics to note 
-that infinite series representations of this type of dimensionless integral 
-exist and converge quickly. For instance, one can show that
-\begin{equation}
-\int_{0}^{a} \frac{u^3}{e^u -1} \,{\rm d}u = \sum_{n=1}^{\infty} \frac{6}{n^4} 
-- \frac{e^{-an}}{n^4} (6 + 6an + 3(an)^2 + (an)^3).
-\nonumber
-\end{equation}
-A similar expression exists for the similar integral in the text. 
-[Michels 1968] explains an analogous derivation.
-}.
-
-The importance of the functional form of $F^{\rm N}_{s,\gamma}$ is that it is 
-to first order only a function of the blackbody temperature and the bandpass 
-wavelength bounds.
-Thus in the most general case $c_i(R_1, R_2, T_1, T_2, \lambda_{\rm min}, 
-\lambda_{\rm max})$ and nothing else.
-The only random variable involved in the flux being received at the detector is 
-$r$, so we can indeed write the flux received at the detector as in 
-Eq.~\ref{eq:flux_at_detector}.
-
-We can finally write out the probability density functions for the signal to 
-noise ratios in the single and double-star cases by using the transformation 
-rule for pdfs, and applying Eq.~\ref{eq:snr_ivory_tower}.
-For single stars,
-\begin{equation}
-{\rm prob}(x_s) = \frac{3}{d_{\rm max, s}^3} c_s^{3/2} \delta^3 \left( A T_{\rm 
-dur} N_{\rm tra}\right)^{3/2} x_s^{-4}.
-\label{eq:prob_xs}
-\end{equation}
-Analogously for double stars,
-\begin{equation}
-{\rm prob}(x_d) = \frac{3}{d_{\rm max, d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 
-\left( A T_{\rm dur} N_{\rm tra}\right)^{3/2} x_d^{-4}.
-\label{eq:prob_xd}
-\end{equation}
-
-\subsubsection{Number of detected planets}
-
-Performing the integrals of Eq.~\ref{eq:completeness_short}, we get:
-
-\begin{align}
-N_{{\rm det},s} &= N_s \Gamma_{t,s} f_{s,g} f_{s,c} \\
-&= N_s \Gamma_{t,s} \frac{R_\star}{a} \frac{1}{d_{\rm max,s}^3} c_s^{3/2} 
-\delta^3 \left(A 
-T_{\rm dur} N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3},
-\end{align}
-and
-\begin{align}
-N_{{\rm det},d} &= 2 N_d \Gamma_{t,d} f_{d,g} f_{d,c} \\
-&= 2 N_d \Gamma_{t,d} \frac{R_\star}{a}
-\frac{1}{d_{\rm max,d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 \left(A T_{\rm dur} 
-N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3}.
-\end{align}
-
-Formally, the $f_{i,c}$ terms should be written as ${\rm min}(1, \ldots)$, 
-where $(\ldots)$ is the given expression. This ensures that the fraction of 
-planets above the signal to noise threshold is less than 1.
-Note that we assumed twin binaries in the geometric transit probabilities.
-Interestingly, the ratio of the two completeness fractions is
-\begin{equation}
-\frac{f_{d,c}}{f_{s,c}} = \left(\frac{d_{\rm max,s}}{d_{\rm max,d}}\right)^3 
-\left(\frac{c_d}{c_s}\right)^{3/2} \mathcal{D}^3 = (1+\gamma_R)^{-3},
-\end{equation}
-which, for the $\gamma_R = 1$ case, gives 
-$1/8$. (We got the same number from simple scaling laws at the outset, but the 
-above machinery will be useful when generalizing in Sec.~\ref{sec:model_2}.)
-
-The number of detected planets $N_{\rm det}$ is the sum of the two appropriate 
-equations above, and can be written
-\begin{align}
-N_{\rm det} = 
-&\left( \frac{\delta}{x_{\rm min}} \right)^3 (A T_{\rm dur} N_{\rm tra})^{3/2}
-\frac{R_\star}{a}
-\nonumber\\
-&\quad \times
-\left[ N_s \Gamma_{t,s} \frac{c_s^{3/2}}{d_{\rm max,s}^3}  +
-	   2 N_d \Gamma_{t,d} \frac{c_d^{3/2}}{d_{\rm max,d}^3}  \right].
-\label{eq:N_det}
-\end{align}
-All terms can be input to a computer, and checked against a Monte Carlo 
-simulation if desired.
-
-\subsection{Astronomer A ignores binarity}
-Astronomer A has never heard of binary star systems. 
-Nor has he heard of completeness corrections.
-He knows about geometric transit probabilities.
-What occurrence rate does he derive for planets of radius $R_p$ and period $P$?
-
-The {\it total} occurrence rate (number of planets divided by number of 
-``stars'') for Astronomer A would be $N_{\rm det}/(N_s + N_d)$, up to the 
-geometric correction. 
-However, even though Astronomer A does not know about binaries, the radii he 
-derives for any planets in binary systems are too small, by a factor 
-$\sqrt{\mathcal{D}}$ (for twin binaries, they are all $R_p/\sqrt{2}$).
-Astronomer A wants an occurrence rate for planets of radius $R_p$ and period 
-$P$.
-The answer is thus
-\begin{equation}
-\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/f_{s,g}}{N_s + N_d}.
-\end{equation}
-
-This astronomer will also think there is a second population of planets, with 
-radius $R_p \sqrt{\mathcal{D}}$, and will thus rush to Nature claiming to also
-have derived a second occurrence rate,
-\begin{equation}
-\Gamma_{{\rm A}, R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm det,d}/f_{s,g}}{N_s 
-+ N_d}.
-\end{equation}
-Note in the second case that this astronomer thinks these are single stars, and 
-generally will miscompute the geometric transit probability.
-
-\subsection{Astronomer B counts host stars correctly}
-Astronomer B can somehow account correctly for the ``2 for 1'' 
-effect of binarity, \textit{i.e.} that the sample actually has more stars than 
-astronomer A thought.
-
-By the same token as above,
-\begin{equation}
-\Gamma_{{\rm B},R_p} = \frac{N_{\rm det,s}/f_{s,g}}{N_s + 2N_d},
-\end{equation}
-and
-\begin{equation}
-\Gamma_{{\rm B},R_p \sqrt{\mathcal{D}}} = 
-		\frac{N_{\rm det,d}/f_{d,g}}{N_s + 2N_d}.
-\end{equation}
-
-\subsection{Astronomer C counts host stars correctly and figures out diluted 
-radii}
-Astronomer C did high resolution imaging followup on every candidate, and 
-correctly classifies the planetary radii.
-Thus, she also knows which planets are in binary systems, and which are in 
-single star systems.
-
-She knows that the purported population of planets with radii $R_p 
-\sqrt{\mathcal{D}}$ does not exist. All detected planets from this survey have 
-radii $R_p$. She computes an occurrence rate
-\begin{equation}
-\Gamma_{{\rm C},R_p} = 
-\frac{N_{\rm det,s}/f_{s,g} + N_{\rm det,d}/f_{d,g}}{N_s + 2N_d}
-\end{equation}
-
-the closest yet to the true rate (Sec.~\ref{sec:true_rate}).
-
-\subsection{Astronomer D counts host stars correctly, figures out diluted 
-radii, 
-and accounts for completeness}
-
-Astronomer D, knows which detections were around 
-binaries and the associated radius correction.
-They do injection recovery, and derive correct estimates for their 
-completeness functions about single stars $f_{s,c}$, and 
-about double star systems, $f_{d,c}$.
-With this knowledge in hand, they compute the respective single and binary 
-occurrence rates
-\begin{equation}
-\Gamma_{t,s} = \frac{N_{\rm det,s}}{N_s f_{s,g} f_{s,c}},
-\end{equation}
-\begin{equation}
-\Gamma_{t,d} = \frac{N_{\rm det,d}}{2 N_d f_{d,g} f_{d,c}}.
-\end{equation}
-
-With these in hand, they derive the overall occurrence rate
-\begin{align}
-\Gamma_{{\rm D}, R_p} 
-&= \frac{N_{\rm det,s}/(f_{s,g}f_{s,c}) + N_{\rm det,d}/(f_{d,g}f_{d,c})  }{N_s 
-+ 2 N_d} \\
-&= \frac{\Gamma_{t,s} N_s + 2\Gamma_{t,d} N_d}{N_s + 2 N_d}.
-\end{align}
-
-Although we admit it's been a bit of a slog, it happens that Astronomer D's 
-occurrence rate is the true occurrence rate (cf. Eq.~\ref{eq:true_occ}).
-
-All it takes is $\approx$ a full semester at Keck, a system-by-system analysis, 
-and perfect understanding of the completeness of the detection efficiency for 
-single and double star systems.
-
-\subsection{Numerical verification}
-
-To check the preceding analytic development, we implemented a Monte Carlo 
-simulation of this idealized transit survey.
-To run the survey, we defined the instrument specifications (detector area and 
-transmission function), the stellar population (binary fraction, total number 
-density of a given stellar class, fixed stellar 
-properties), the planet population (fixed planet radius, period, and occurrence 
-rate about single and binary stars), and finally the survey parameters 
-(observing baseline, minimum SNR for ``detection'').
-We then randomly drew star positions, randomly assigned planets to stars in 
-single and binary systems, and computed the resulting signal to noise 
-(Eq.~\ref{eq:snr_ivory_tower}) with which 
-the transits would be observed.
-As in the preceding analytics, we assumed ``twin'' binaries (same stellar 
-radii, same effective 
-temperature, and dilution does not depend on which stellar binary is the 
-``target'').
-
-The results are shown in Fig.~\ref{fig:snr_dist_analytic_v_numeric}, and 
-indicate that the analytic probability distribution functions 
-Eqs.~\ref{eq:prob_xs},~\ref{eq:prob_xd}, and the number of 
-detections (Eq.~\ref{eq:N_det}) are correct.
-
-A point evident in Fig.~\ref{fig:snr_dist_analytic_v_numeric} is that, for 
-fixed planet parameters, and fixed stellar parameters ($R_\star, L_\star$, and 
-distance $r$) the SNR distribution for planets in binaries is poorer than that 
-of planets in single star systems.
-We can see analytically that this simply due to dilution:
-\begin{equation}
-\frac{{\rm prob}(x_d)}{{\rm prob}(x_s)} =(1 + \gamma_R)^{-1} = \mathcal{D}.
-\end{equation}
-Deriving this simple form requires noting that the ratios of the 
-bandpass-specific number luminosities (Eq.~\ref{eq:c_i}) is equal to the ratio 
-of the bandpass-specific energy luminosities (otherwise a term with $c_s/c_d$ 
-must be included).
-
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=0.4]{figures/snr_distribution.pdf}
-	\end{center}
-	\caption{Comparison of analytic and numeric probability 
-	density functions of the SNR in an idealized transit survey.
-	The analytic lines are Eqs.~\ref{eq:prob_xs} and~\ref{eq:prob_xd} for the 
-	planet populations orbiting single and binary stars. The underlying stepped 
-	histogram is output from Monte Carlo simulations. Poisson noise leads to 
-	a small deviation at the faint and bright limits, but the numerics and 
-	analytics otherwise agree.
-		 }
-	\label{fig:snr_dist_analytic_v_numeric}
-\end{figure}
-
-\subsection{Representative numbers for a few cases}
-
-\subsubsection{Twin binaries: if we ignore binarity, for what fraction of 
-detections do we misclassify the radii?}
-
-Ignoring binarity, we will detect $N_{\rm det,s}$ planets around single stars, 
-and $N_{\rm det,d}$ planets around double stars. The latter set will be assumed 
-to have radii $R_p \sqrt{\mathcal{D}}$.
-The fraction of detections with misclassified radii can then be written
-\begin{equation}
-\frac{N_{\rm det,d}}{N_{\rm det,s} + N_{\rm det,d}} = \frac{1}{1+\alpha},
-\end{equation}
-for
-\begin{equation}
-\alpha \equiv \frac{1}{2({\rm BF})} (1+\gamma_R)^{3/2} 
-\frac{\Gamma_{t,s}}{\Gamma_{t,d}}.
-\end{equation}
-
-For the nominal G2V dwarf case of ${\rm BF}=0.45$, twin binaries with equal 
-occurrence rates this produces a misclassification rate of $24\%$, in agreement 
-with Fig.~\ref{fig:snr_dist_analytic_v_numeric}.
-
-\subsubsection{Twin binaries: if we ignore binarity, how wrong is our
-occurrence rate for planets of radius $R_p$?}
-
-This is almost simply asking ``what is the relative difference between the 
-occurrence rates derived by Astronomers D and A for planets of radius $R_p$?''
-However, in the more realistic case, Astronomer A also has derived a 
-completeness, which we assume is the same as for Astronomer D in the single 
-star case.
-So Astronomer A now misclassifies planetary radii, and miscounts the total 
-number of stars, but knows his completeness for single stars.
-Astronomer D corrects all these errors.
-Then
-\begin{equation}
-\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d}.
-\end{equation}
-
-To express ``how wrong'' our occurrence rate is, we define a correction factor 
-$X_\Gamma$ such that the true occurrence rate is the product of the correction 
-factor with the measured occurrence rate: $\Gamma_D \equiv \Gamma_A X_\Gamma$.
-
-The correction factor is then
+In our twin binary thought-experiment, dilution means that all planets 
+detected in binary systems will have apparent radii of $R_p/\sqrt{2}$.
+Since for twin binaries the stellar radii are the same, the transit 
+probability will be correctly calculated.
+
+The total number of planet detections, irrespective of whether they are ``of 
+the desired type'', is the sum of the number of planets 
+detected in single star systems $N_{{\rm det},s}$ and the number of planets 
+detected in double star systems $N_{{\rm det},d}$.
+Expressing each individually,
+\begin{equation}
+N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,g} f_{s,c},
+\label{eq:N_det_s}
+\end{equation}
+where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
+star systems of the sample, $f_{s,g}$ is the geometric transit probability, 
+and $f_{s,c}$ is the fraction of these transiting planets that are observed 
+with signal to noise greater than the minimum detection threshold (the 
+completeness). 
+For our SNR-limited, twin-binary survey, $f_{s,g} = R_1/a\equiv f_g$, and 
+$f_{s,c}=1$.
+
+Analogously,
+\begin{equation}
+N_{{\rm det},d} = (1+w_{d2}) N_d \Gamma_{t,d} f_{d,g} f_{d,c},
+\label{eq:N_det_d}
+\end{equation}
+where now $(1+w_{d2}) N_d \Gamma_{t,d}$ is the number of planets in the double 
+star systems of the sample, the geometric transit probability $f_{d,g}=R_1/a$ 
+for our twin-binary survey, and $f_{d,c}=1/8$ by the geometric argument 
+presented in the initial discussion of limiting distances.
+
+The occurrence rate this astronomer derives for planets of assumed 
+radius $R_p$ orbiting stars thought to be of the desired type is then
+\begin{equation}
+\Gamma_{\rm a} = \frac{\Gamma_{t,s}N_s}{N_s + N_d}.
+\end{equation}
+
+\subsection{Errors on derived occurrence rates for twin-binary thought 
+experiment}
+The ratio of the true to calculated occurrence rates for planets of radius 
+$R_p$, $X_\Gamma$, is
 \begin{align}
 X_\Gamma &=
-\frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}} \\
+\frac{\Gamma_t}{\Gamma_{\rm a}} \\
 &=
-\frac{\Gamma_{t,s}N_s + 2\Gamma_{t,d}N_d}{N_s + 2N_d}  \cdot 
-\frac{N_s + N_d}{\Gamma_{t,s} N_s } \\
+\frac{\Gamma_{t,s}N_s + (1+w_{d2})\Gamma_{t,d}N_d}{\Gamma_{t,s} N_s } \cdot 
+\frac{N_s + N_d}{N_s + 2N_d} \\
 &= 
-\frac{(1 + \beta)(1 + 2 \beta \Gamma_{t,d}/\Gamma_{t,s})}{(1+2\beta)},
+\left( 
+1 + \frac{(1+w_{d2}) \Gamma_{t,d}}{\Gamma_{t,s}} \beta
+\right) \cdot 
+\frac{1+\beta}{1+2\beta},
 \end{align}
@@
 \begin{equation}
-\beta \equiv N_d/N_s = {\rm BF} \times (1+\gamma_R)^{3/2}.
-\end{equation}
-
-For the nominal G2V dwarf case of ${\rm BF}=0.45$ with twin binaries ($\gamma_R 
-= 1$) and $\Gamma_{t,d} = \Gamma_{t,s}$ this gives $X_\Gamma = 2.27$.
-For instance, in the numerical simulation corresponding to 
-Fig.~\ref{fig:snr_dist_analytic_v_numeric}, Astronomer A finds 
-$\Gamma_{{\rm A},R_p} = 0.22$, while Astronomer 
-D derives the true (input) occurrence rate of $\Gamma_{{\rm D},R_p} = 0.5$.
-This gives a relative percentage error of $\delta\Gamma_{{\rm D},R_p} \equiv 
-(\Gamma_{\rm A,R_p} - \Gamma_{\rm D,R_p}) / \Gamma_{\rm D,R_p} = -56\%$.
-
-\section{Model \#2: fixed stars, fixed planets, varying light-ratio binaries}
+\beta \equiv N_d/N_s = \frac{\rm BF}{1- {\rm BF}} \times (1+\gamma_R)^{3/2}.
+\end{equation}
+
+For the nominal G2V dwarf case of ${\rm BF}=0.44$, if we assume that all stars 
+have equal occurrence rates, $\Gamma_{t,s} = \Gamma_{t,d}$ and 
+$w_{d2}=1$, then we find $X_\Gamma = 1+\beta = 1+ 2^{3/2}\approx4$~--~ the 
+observed occurrence rate is underestimated by a factor of 4.
+
+For the planets around double stars with observed radii $R_p 
+\sqrt{\mathcal{D}}$, there is also the error of their misclassification.
+The fraction of detections with misclassified radii can be written
+\begin{equation}
+\frac{N_{\rm det,d}}{N_{\rm det,s} + N_{\rm det,d}} = \frac{1}{1+\zeta},
+\end{equation}
+for
+\begin{equation}
+\zeta \equiv \frac{N_s \Gamma_{t,s} f_{s,g} f_{s,c}}{
+	 (1+w_{d2})N_d \Gamma_{t,d} f_{d,g} f_{d,c }}.
+\end{equation}
+
+For our nominal G2V dwarf case, the twin binaries with equal 
+occurrence rates gives $\zeta = 4/\beta = \sqrt{2}$, so 41\% of detections 
+have misclassified radii, all of $R_p/\sqrt{2}$.
+
+Of course more generally, the observer might define their 
+desired ``type of planet'' to be anything within some margin of $R_p$.
+Considering this for the idealized twin-binary case, if the margin includes 
+the planets in binary systems, the number of detections ``of the desired 
+type'' would change.
+The correction factor $X_\Gamma$ would then become
+\begin{equation}
+X_\Gamma = 
+\frac{\Gamma_{t,s} + (1+w_{d2})\Gamma_{t,d}\beta}{\Gamma_{t,s} + 
+(1+w_{d2})\Gamma_{t,d} f_{d,c} } \cdot 
+\frac{1+\beta}{1+2\beta},
+\end{equation}
+which simplifies to $(1+2f_{d,c})^{-1} (1+\beta) = 4(1+\beta)/5\approx 3$ for 
+our twin-binary thought experiment.
+This makes sense: increasing the radius interval of ``desired planets'' 
+decreases the error in the numerator, but the error of mis-counted stars in 
+the denominator remains identical.
+
+\section{Model \#2: fixed primaries, fixed planets, varying light-ratio 
+secondaries}
 \label{sec:model_2}
 
-Same story as Model \#1, except now $\gamma_R$ varies across the population 
+The model of Sec.~\ref{sec:model_1} is of course idealized, but it helps
+for developing intuition.
+We now let the light ratio $\gamma_R = L_2/L_1$ vary across the population 
 of star systems.
-It does so because the underlying mass ratio varies. 
-Since we are interested in solar type binaries, we take the distribution of the 
-mass ratio $q=M_2/M_1$ by approximating [Rhagavan et al 2010, Fig 16]:
-\begin{equation}
-{\rm prob}(q) =
+It does so because the underlying mass ratio $q=M_2/M_1$ varies.
+We keep the primary mass fixed as $M_1$, which is also the mass of all single 
+stars.
+
+One parametrization for the distribution of 
+binary mass ratios in a volume-limited sample is $f(q)\propto q^\gamma$.
+[Duchene and Kraus 2013], fitting all the multiple systems of [Raghavan et al 
+(2010)]'s Fig 16, quote $\gamma = 0.28\pm0.05$ for $0.7<M_\star/M_\odot<1.3$.
+Examining only the binary systems of Rhagavan et al 2010, Fig 16, the 
+distribution is roughly uniform,
+\begin{equation}
+f(q) =
 \Bigg\{\begin{array}{lr}
@@
 \end{equation}
-for $c_q = 1/9$ to normalize the distribution to 1.
-The mass-luminosity relation can, for analytic convenience, be approximated as 
-$L = M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
-While we use this in subsequent analytic development, for numerics we fit a 
-lines to mass-luminosity data collected by Torres et al. [2009] for 
-dwarfs above M, and Benedict et al. [2016] for dwarfs below M.
+provided we ignore the peak at high mass ratios for analytic convenience. $c_q 
+= 1/9$ is the normalization.
+We use this latter mass ratio distribution.
+
+For the mass-luminosity relation, for analytic convenience we assume 
+$L \propto M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
+As a simple check, we fit lines to mass-luminosity data collected by 
+Torres et al. [2009] for dwarfs earlier than spectral class M, and Benedict 
+et al. [2016] for dwarfs later than spectral class M.
 To convert Benedict et al. [2016]'s reported $M_V$ values to absolute 
@@
 	downloaded 2017.08.02}.
-In log-log space, we let the intersection point of the lines float, and make 
+In log-log space, we let the intersection point of the lines float, and made 
 various cuts on the data as indicated in Fig~\ref{fig:mass_luminosity}. The 
@@
 	L at merge: 0.0281260412126928.
-}.
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.9]{figures/mass_luminosity.pdf}
-	\end{center}
-	\caption{Empirical fit to main sequence dwarf mass luminosity data compiled 
-	from Torres et al. [2009] and Benedict et al. [2016]. The ``low fit'' is a 
-	least squares fit to data from $0.1-0.5M_\odot$, and the ``high fit'' is to 
-	data above that, and below the Kraft break ($1.2M_\odot$).
-	The $L(M)$ relation taken for subsequent numerics is the maximum 
-	of the two fits.
-	}
-	\label{fig:mass_luminosity}
-\end{figure}
-
-The other changes between Model \#2 and Model \#1 are:
-\begin{enumerate}
-\item We allow a third occurrence rate, splitting $\Gamma_{t,d} \rightarrow 
-(\Gamma_{t,d1}, \Gamma_{t,d2})$ where the latter terms represent the occurrence 
-rate around the primary of a double star system, and the occurrence rate around 
-the secondary of a double star system. For the case of \textit{e.g.}, $q=0.1$, 
-this seems more relevant.
-We can then take the limit that says ``any planet around the secondary is as 
-likely as around the primary'', or go to the opposite extreme of ``there are 
-only planets around the primaries, not the secondaries''.
-%
-\item Our binary systems have secondaries with varying masses. Thus they have 
-varying radii. Taking [Demircan \& Kahraman 1991]'s empirical fit to eclipsing 
+}. Though the data show a break at $M_\star \approx 0.5M_\odot$, the 
+$L=M^{3.5}$ relation is a reasonable fit for $M\lesssim 2M_\odot$.
+
+The final addition to our model is that we must assume a stellar mass-radius 
+relation.
+This affects the transit probabilities and signal sizes about the 
+secondaries of binary star systems.
+[Demircan \& Kahraman 1991]'s empirical fit to eclipsing 
 binary data,
@@
 \end{equation}
-for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 1991's 
-stated bounds), and for each in solar units. This only affects the transit 
-probabilities about the secondaries of binary star systems.
-\end{enumerate}
-
-We then ask the same questions as for Model \#1.
-
-\subsection{How many stars are in the sample?}
-
-Let $N_s$ be the number of single star systems, and $N_d$ the number of double 
-star systems. $N_s$ is the same as in Sec.~\ref{sec:model_1}.
-
-Now
-\begin{equation}
-N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3(L_1^N, \gamma_R, F_{\rm 
-	lim}^N)
+applies for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 
+1991's stated bounds), where each quantity is given in solar units. 
+Since again, our main purpose is to develop intuition about observational 
+biases, we approximate this as $R\propto M$.
+
+In passing, we note that while $w_{d2}=1$ was a reasonable assumption in 
+Sec.~\ref{sec:model_1} (since secondaries were identical to primaries), in 
+binary systems we now have secondaries with different stellar properties from 
+primaries.
+Recall that our occurrence rate is defined for planets of a given type 
+orbiting stars of a given type.
+Assume our universe is one in which the primary mass of all stars is 
+$M_1=M_\odot$.
+If this is the case, we might define our ``stars of interest'' to be 
+those with $0.7<M_\star/M_\odot<1.3$.
+Then, in a volume-limited sample, 1 in 3 secondaries would be ``of interest'', 
+and could contribute to the true occurrence rate.
+The general point is that $w_{d2}$ will be less than 1, unless planets of the 
+desired type are more common around secondaries of lower mass (as may actually 
+be true for small planets around M dwarfs).
+Keeping this in mind, we ask the same questions as for Model \#1.
+
+\subsection{What is the true occurrence rate?}
+
+The true occurrence rate is the average number of planets of the desired type 
+per star of the desired type.
+The number of planets of the desired type is
+\begin{equation}
+N_{\rm planets} = \Gamma_{t,s}N_s + \Gamma_{t,d}N_d + 
+\Gamma_{t,d}w_{d2}f_{{\rm desired}}N_d,
+\end{equation}
+where as before $N_s$ and $N_d$ are the number of single and double systems, 
+and the fraction $f_{{\rm desired}}$ is defined so that $f_{{\rm desired}} 
+N_d$ is the number of secondaries that are of the desired type, and as before 
+$\Gamma_{t,d}w_{d2}$ is the average number of planets each of these desired 
+stars hosts.
+
+Counting the number of desired stars, $N_s$ is the same as in 
+Sec.~\ref{sec:model_1}. However
+\begin{equation}
+N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3(L_1, \gamma_R, F_{\rm 
+	min})
 \end{equation}
@@
 \begin{equation}
-d_{\rm max,d}(L_1^N, \gamma_R, F_{\rm lim}^N) =
-d_{\rm max,s}(L_1^N, F_{\rm lim}^N)\times (1+\gamma_R)^{1/2}.
-\end{equation}
-Since $q$ is a random variable, $\gamma_R$ is a random variable, and $d_{\rm 
-	max,d}$ is a random variable. Since $N_d$ is a function of $d_{\rm max,d}$, 
+d_{\rm max,d}(L_1, \gamma_R, F_{\rm min}) =
+d_{\rm max,s}(L_1, F_{\rm min})\times (1+\gamma_R)^{1/2},
+\end{equation}
+where $d_{\rm max,s}$ is given by Eq.~\ref{eq:d_max}.
+
+For a given binary system, $q$, $\gamma_R$, and $d_{\rm 
+	max,d}(\gamma_R)$ are all random variables.
+Since $N_d$ is a function of $d_{\rm max,d}$, 
 the number of double star systems becomes a random variable.
-The expected (mean) number of double star systems in the sample is
-\begin{equation}
-\langle N_d \rangle = \int_{0}^{\infty} N_d\, {\rm prob}(N_d) \,{\rm d}N_d.
-\end{equation}
-By applying the chain rule for probability density functions, the distribution 
-${\rm prob}(N_d)$ can be written 
+Applying the chain rule for probability density functions, the distribution 
+function for the number of binary systems,
+${\rm prob}(N_d)$, can be written 
 \begin{equation}
@@
 \end{equation}
-Doing some algebra, and assuming $\gamma_R = q^3$, this can be shown to be
+
+\paragraph{Necessary aside on biases in mass ratio distributions}
+
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=.8]{figures/joint_prob_r_q.pdf}
+	\end{center}
+	\caption{The joint probability distribution of a binary star's position and 
+		mass ratio in a magnitude-limited sample. This plot assumes a different 
+		$M(L)$ relation than previous plots: $L=M^3$. Note that the marginalized 
+		mass ratio distribution is uniform at any given distance.}
+	\label{fig:joint_prob_r_q}
+\end{figure}
+
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=.8]{figures/q_distribn_mag_limited.pdf}
+	\end{center}
+	\caption{The distribution of the mass ratio for a magnitude limited 
+		sample of binary stars. The empirical mass-luminosity relation is shown
+		in Fig.~\ref{fig:mass_luminosity}.
+		The underlying mass ratios are drawn from Eq.~\ref{eq:mass_ratio} -- a 
+		\textit{uniform} distribution in a volume-limited sample!
+		The entire bias can be understood analytically 
+		(Eq.~\ref{eq:pdf_q_mag_limited}) to reasonable precision. 
+	}
+	\label{fig:q_distribn_mag_limited}
+\end{figure}
+
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=.8]{figures/gammaR_distribn_mag_limited.pdf}
+	\end{center}
+	\caption{The distribution of the luminosity ratio for a magnitude limited 
+		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
+		total 
+		number density from Bovy 2017; $M(L)$ relation from 
+		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
+		Eq.~\ref{eq:mass_ratio}.}
+	\label{fig:gammaR_distribn_mag_limited}
+\end{figure}
+
+Before blindly plugging in Eq.~\ref{eq:mass_ratio}'s assumed mass ratio 
+distribution to find the expected number of double systems, we must
+emphasize that Eq.~\ref{eq:mass_ratio}'s distribution applies only for 
+volume-limited samples.
+If one is interested in a SNR-limited transit survey (or really any
+magnitude-limited survey), the Malmquist-like bias associated with binarity, 
+and affecting the distribution of binary mass ratios, must be included.
+
+The distribution function for the mass ratio of binaries in a magnitude 
+limited sample can be found by marginalizing over the joint 
+distribution for a binary star's position $r$ and mass ratio $q$.
+Explicitly, using `ml' as a subscript for `magnitude-limited',
+\begin{equation}
+{\rm prob}_{\rm ml}(r,q) = {\rm prob}(q|r){\rm prob}_{\rm ml}(r),
+\end{equation} 
+where ${\rm prob}(q|r)$ is given by Eq.~\ref{eq:mass_ratio}, and 
 \begin{align}
-{\rm prob}(N_d) = \frac{2}{81} &N_d^{-1/3} \left(\frac{3}{4\pi 
-	n_d}\right)^{2/3} d_{\rm max,s}^{-2/3} \ \times \nonumber\\
-&\left[ \left(\frac{3 N_d}{4\pi n_d}\right)^{2/3} - d_{\rm 
-	max,s}^2 \right]^{-2/3}
+{\rm prob}_{\rm ml}(r) &= \int_{0}^{1} {\rm prob}_{\rm ml}(r|q)\,{\rm 
+	prob}(q)\,{\rm d}q
+\\
+&= \int_{0.1}^{1} \frac{1}{3} \frac{r^2}{d_{\rm max,d}^3(q)}\,{\rm d}q
+\quad {\rm if\ }r<d_{\rm max}(q),
+\end{align}
+and otherwise zero. The resulting joint probability distribution is plotted in 
+Fig.~\ref{fig:joint_prob_r_q}.
+
+Marginalizing the joint distribution over distance, and assuming $L\propto 
+M^{3.5}$,
+\begin{align}
+{\rm prob}_{\rm ml}(q) &= 
+\int_{0}^{d_{\rm max}(q)} {\rm prob}_{\rm ml}(r,q) \,{\rm d}r \\
+&= \mathcal{Z}\cdot (1+q^{3.5})^{3/2}\quad {\rm if\ }0.1<q\leq 1,
+\label{eq:pdf_q_mag_limited}
+\end{align}
+and otherwise 0, for $\mathcal{Z}$ the appropriate normalization constant.
+This marginalized distribution~--~the distribution function of mass ratios for 
+binaries in a magnitude-limited sample~--~is plotted in 
+Fig.~\ref{fig:q_distribn_mag_limited}.
+
+More correctly then, we must write
+\begin{equation}
+{\rm prob}(N_d) = {\rm prob_{ml}}(q(\gamma_R)) 
+\left| \frac{{\rm d}q}{{\rm d}\gamma_R}  \right|
+\left| \frac{{\rm d}\gamma_R}{{\rm d}d_{\rm max,d}}  \right|				
+\left| \frac{{\rm d}d_{\rm max,d}}{{\rm d}N_d}  \right|.
+\end{equation}
+
+Doing some algebra, and assuming $\gamma_R = q^\alpha$, this can be shown to be
+\begin{align}
+{\rm prob}(N_d) = \frac{2\mathcal{Z}}{3\alpha}& N_d^{1/3} \left(\frac{3}{4\pi 
+	n_d}\right)^{4/3} d_{\rm max,s}^{-4} \ \nonumber\\
+&\times \left[ 
+	\left(\frac{3 N_d}{4\pi n_d}\right)^{2/3}  d_{\rm max,s}^{-2} - 1
+ \right]^{\frac{1-\alpha}{\alpha} }
 \end{align}
@@
 \begin{equation}
-{\rm lower\ bound} = \frac{4 \pi n_d}{3} (\sqrt{0.1^3 + 1} d_{\rm max,s})^3
-\end{equation}
-\begin{equation}
-{\rm upper\ bound} = \frac{4 \pi n_d}{3} (\sqrt{2}d_{\rm max,s})^3,
+N_d^{\rm lower} = \frac{4 \pi n_d}{3} (\sqrt{0.1^3 + 1} d_{\rm max,s})^3,
+\end{equation}
+\begin{equation}
+N_d^{\rm upper} = \frac{4 \pi n_d}{3} (\sqrt{2}d_{\rm max,s})^3,
 \end{equation}
 and outside the stated interval ${\rm prob}(N_d) = 0$.
+
+An interesting intermediate result is the magnitude-limited luminosity ratio 
+distribution,
+\begin{equation}
+{\rm prob_{ml}}(\gamma_R) = \frac{\mathcal{Z}}{\alpha} (1+\gamma_R) 
+\gamma_R^{\frac{1-\alpha}{\alpha}},
+\end{equation}
+for $(0.1)^{\alpha}<\gamma_R<1$, and otherwise zero.
+For $\alpha=3.5$, ${\rm prob_{ml}}(\gamma_R) \propto 
+(1+\gamma_R)\gamma_R^{-5/7}$, which tells us the distribution will be highly 
+biased towards low-light ratio binaries.
+The distribution is plotted in Fig.~\ref{fig:gammaR_distribn_mag_limited}.
+
+The reason we derived all this is that we wanted to know the number of 
+searchable binary stars.
+The expected value for the number of double star systems in the sample is
+\begin{equation}
+\langle N_d \rangle = \int_{0}^{\infty} N_d\, {\rm prob}(N_d) \,{\rm d}N_d.
+\end{equation}
+With some algebra, and letting $\alpha=3.5$, this can be expressed as
+\begin{equation}
+\langle N_d \rangle = \frac{2 \mathcal{Z}}{7 R^{1/2}}
+\int_{(0.1^\alpha + 1)^{1/3}}^{\sqrt{2}} u^{3/2} (u-1)^{5/7} \,{\rm d}u ,
+\label{eq:integral_Nd}
+\end{equation}
+for
+\begin{equation}
+R \equiv \left(\frac{3}{4\pi n_d}\right)^{2/3} d_{\rm max,s}^{-2}.
+\end{equation}
+The integral in Eq.~\ref{eq:integral_Nd} has a closed-form solution in terms 
+of Gauss hypergeometric functions, $_2F_1$. Once evaluated, the integral takes 
+a numerical value of $0.1829$.
+The normalization of
+
+%TODO GIVE THE EXPLICIT FORM, SEMI-ANALYTICALLY.
 
@@
 $10^{-3}$, where the probability distribution peaks.
-
-Similarly, the probability distribution function for the mass ratio of binaries 
-in a magnitude limited sample can be found by marginalizing over the joint 
-distribution for a binary star's position and mass ratio.
-Explicitly, using `ml' as a subscript for `magnitude-limited', and assuming 
-$L\propto M^3$,
-\begin{align}
-{\rm prob}_{\rm ml}(q) &= 
-	\int_{0}^{d_{\rm max}(q)} {\rm prob}_{\rm ml}(r,q) \,{\rm d}r \\
-&= \mathcal{Z}\cdot (1+q^3)^{3/2}\quad {\rm if\ }0.1<q\leq 1,
-\label{eq:pdf_q_mag_limited}
-\end{align}
-and otherwise 0, for $\mathcal{Z}$ the appropriate normalization constant.
-Eq.~\ref{eq:pdf_q_mag_limited} is the ``analytic'' form overplotted in 
-Fig.~\ref{fig:q_distribn_mag_limited}. Its independence from both the primary 
-star mass as well as from the magnitude cutoff will be used in 
-Sec.~\ref{sec:model_3}.
-The form given in Eq.~\ref{eq:pdf_q_mag_limited} is found by writing 
-${\rm prob}_{\rm ml}(r,q) = {\rm prob}(q|r){\rm prob}_{\rm ml}(r)$, using 
-Eq.~\ref{eq:mass_ratio} for ${\rm prob}(q|r)$, and marginalizing over the mass 
-ratio
-\begin{align}
-{\rm prob}_{\rm ml}(r) &= \int_{0}^{1} {\rm prob}_{\rm ml}(r|q)\,{\rm 
-prob}(q)\,{\rm d}q
-\\
-&= \int_{0.1}^{1} \frac{1}{3} \frac{r^2}{d_{\rm max}^3(q)}\,{\rm d}q
-\quad {\rm if\ }r<d_{\rm max}(q),
-\end{align}
-and otherwise zero. The resulting joint probability distribution is plotted in 
-Fig.~\ref{fig:joint_prob_r_q}.
-
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/gammaR_distribn_vol_limited.pdf}
-	\end{center}
-	\caption{The distribution of the luminosity ratio for a volume limited 
-		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
-		total 
-		number density from Bovy 2017; $M(L)$ relation from 
-		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
-		Eq.~\ref{eq:mass_ratio}.}
-	\label{fig:gammaR_distribn_vol_limited}
-\end{figure}
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/q_distribn_mag_limited.pdf}
-	\end{center}
-	\caption{The distribution of the mass ratio for a magnitude limited 
-		sample of binary stars. For numerics, ${\rm BF} = 0.45$ [Duchene and 
-		Kraus 2013], total number density from Bovy 2017, empirical $M(L)$ 
-		from Fig.~\ref{fig:mass_luminosity}, and mass ratios are drawn from 
-		Eq.~\ref{eq:mass_ratio} -- a \textit{uniform} distribution in a 
-		volume-limited sample!
-		The analytic line (Eq.~\ref{eq:pdf_q_mag_limited}) is shown to indicate 
-		that the entire effect, albeit with a simpler $M(L)$, can be understood 
-		without resorting to numerics. 
-		}
-	\label{fig:q_distribn_mag_limited}
-\end{figure}
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/gammaR_distribn_mag_limited.pdf}
-	\end{center}
-	\caption{The distribution of the luminosity ratio for a magnitude limited 
-		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
-		total 
-		number density from Bovy 2017; $M(L)$ relation from 
-		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
-		Eq.~\ref{eq:mass_ratio}.}
-	\label{fig:gammaR_distribn_mag_limited}
-\end{figure}
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/joint_prob_r_q.pdf}
-	\end{center}
-	\caption{The joint probability distribution of a binary star's position and 
-	mass ratio in a magnitude-limited sample. This plot assumes a different 
-	$M(L)$ relation than previous plots: $L=M^3$. Note that the marginalized 
-	mass ratio distribution is uniform at any given distance.}
-	\label{fig:joint_prob_r_q}
-\end{figure}
 
@@
 
+\begin{figure}[!t]
+	\begin{center}
+		\includegraphics[scale=0.9]{figures/mass_luminosity.pdf}
+	\end{center}
+	\caption{Empirical fit to main sequence dwarf mass luminosity data compiled 
+		from Torres et al. [2009] and Benedict et al. [2016]. The ``low fit'' is a 
+		least squares fit to data from $0.1-0.5M_\odot$, and the ``high fit'' is 
+		to 
+		data above that, and below the Kraft break ($1.2M_\odot$).
+		The $L(M)$ relation taken for numerics is the maximum 
+		of the two fits. For analytics, we assume $L\propto M^{3.5}$.
+	}
+	\label{fig:mass_luminosity}
+\end{figure}
+
 If we assume every KIC star to be single, what order of magnitude of error do 

### 2017-09-25 paper/ms.tex
gammaR distribn: update plot

@@
   population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
-  L^{-1/2} d$,
+  L^{1/2} d^{-1}$,
   \begin{equation}
@@
 \begin{equation}
-N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3(L_1, \gamma_R, F_{\rm 
-	min})
+N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3(\gamma_R)
 \end{equation}
@@
 \begin{equation}
-d_{\rm max,d}(L_1, \gamma_R, F_{\rm min}) =
-d_{\rm max,s}(L_1, F_{\rm min})\times (1+\gamma_R)^{1/2},
+d_{\rm max,d}(\gamma_R) =
+d_{\rm max,s}\times (1+\gamma_R)^{1/2},
 \end{equation}
@@
 	\caption{The distribution of the mass ratio for a magnitude limited 
-		sample of binary stars. The empirical mass-luminosity relation is shown
-		in Fig.~\ref{fig:mass_luminosity}.
-		The underlying mass ratios are drawn from Eq.~\ref{eq:mass_ratio} -- a 
-		\textit{uniform} distribution in a volume-limited sample!
+		sample of binary stars. The underlying mass ratios are drawn from 
+		Eq.~\ref{eq:mass_ratio} -- a \textit{uniform} distribution in a 
+		volume-limited sample.
 		The entire bias can be understood analytically 
-		(Eq.~\ref{eq:pdf_q_mag_limited}) to reasonable precision. 
+		(Eq.~\ref{eq:pdf_q_mag_limited}) to good precision. 
+        The numerical comparison uses the empirical mass-luminosity relation 
+        shown in Fig.~\ref{fig:mass_luminosity}.
 	}
@@
 	\caption{The distribution of the luminosity ratio for a magnitude limited 
-		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
-		total 
-		number density from Bovy 2017; $M(L)$ relation from 
-		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
-		Eq.~\ref{eq:mass_ratio}.}
+		sample of binary stars. Despite the preference in the sample towards 
+		large $q$ binaries (Fig.~\protect\ref{fig:q_distribn_mag_limited}), 
+		the bias is towards small $\gamma_R$ binaries because of the steepness 
+		of the mass-luminosity relation.
+        The underlying mass ratios are uniform in volume.
+        The numerical comparison uses the empirical mass-luminosity relation 
+        shown in Fig.~\ref{fig:mass_luminosity}.       
+    }
 	\label{fig:gammaR_distribn_mag_limited}
@@
 If one is interested in a SNR-limited transit survey (or really any
-magnitude-limited survey), the Malmquist-like bias associated with binarity, 
-and affecting the distribution of binary mass ratios, must be included.
+magnitude-limited survey), the Malmquist-like bias associated with binarity
+must be included.
 

### 2017-09-25 src/binary_distribution.py

@@ -62,7 +62,7 @@ def _make_distribution_plots(vl, doubles):
     f.tight_layout()
     f.savefig(savedir+'gammaR_distribn_vol_limited.pdf', dpi=250, bbox_inches='tight')
 
-    # Now do magnitude limited case.
+    # Now do magnitude limited case. Maglimited mass ratio.
     plt.close('all')
     f,ax = plt.subplots(figsize=(4,4))
     hist, bin_edges = np.histogram(doubles['q'],
@@ -84,14 +84,25 @@ def _make_distribution_plots(vl, doubles):
     f.tight_layout()
     f.savefig(savedir+'q_distribn_mag_limited.pdf', dpi=250, bbox_inches='tight')
 
+    # Maglimited light ratio
     plt.close('all')
     f,ax = plt.subplots(figsize=(4,4))
     hist, bin_edges = np.histogram(doubles['γ_R'],
-            bins=np.append(np.linspace(0,1.01,501),42), normed=True)
-    ax.step(bin_edges[:-1], hist, 'k-', where='post')
+            bins=np.append(np.linspace(0,1.01,101),42), normed=True)
+    ax.step(bin_edges[:-1], hist, where='post',
+            label='numerical, empirical $L(M)$')
+    # Analytic distribution from 17/09/24 result
+    _gammaR = np.arange(0,1+1e-4,1e-4)
+    curly_Z = 1/I_2
+    pdf_gammaR_analytic = (curly_Z / 3.5) * (1+_gammaR)*_gammaR**(-5/7)
+    pdf_gammaR_analytic[_gammaR<(0.1**(3.5))] = 0
+    ax.plot(_gammaR, pdf_gammaR_analytic, label='analytic, $L=M^{3.5}$')
+    ax.legend(loc='upper right', fontsize='small')
     ax.set(xlabel='$\gamma_R = L_2/L_1$',
            ylabel='prob',
-           xlim=[0,1.05])
+           xlim=[0,1.05],
+           ylim=[0.1,90],
+           yscale='log')
     ax.set_title(txt_ml, fontsize='small')
     f.tight_layout()
     f.savefig(savedir+'gammaR_distribn_mag_limited.pdf', dpi=250, bbox_inches='tight')

### 2017-09-25 src/binary_distribution.py
bugfix: light ratio distribn

@@ -92,10 +92,12 @@ def _make_distribution_plots(vl, doubles):
     ax.step(bin_edges[:-1], hist, where='post',
             label='numerical, empirical $L(M)$')
     # Analytic distribution from 17/09/24 result
-    _gammaR = np.arange(0,1+1e-4,1e-4)
-    curly_Z = 1/I_2
-    pdf_gammaR_analytic = (curly_Z / 3.5) * (1+_gammaR)*_gammaR**(-5/7)
+    from scipy.integrate import trapz
+    _gammaR = np.arange(0,1+2e-5,2e-5)
+    pdf_gammaR_analytic = (1+_gammaR)**(3/2)*_gammaR**(-5/7)
     pdf_gammaR_analytic[_gammaR<(0.1**(3.5))] = 0
+    norm = trapz(pdf_gammaR_analytic, _gammaR)
+    pdf_gammaR_analytic = pdf_gammaR_analytic/norm
     ax.plot(_gammaR, pdf_gammaR_analytic, label='analytic, $L=M^{3.5}$')
     ax.legend(loc='upper right', fontsize='small')
     ax.set(xlabel='$\gamma_R = L_2/L_1$',

### 2017-09-25 paper/ms.tex
include corrected plot; update text

@@
 	\begin{center}
-		\includegraphics[scale=.8]{figures/q_distribn_mag_limited.pdf}
+		\includegraphics[width=0.5\textwidth]{figures/q_distribn_mag_limited.pdf}
 	\end{center}
@@
 	\begin{center}
-		\includegraphics[scale=.8]{figures/gammaR_distribn_mag_limited.pdf}
+		\includegraphics[width=.5\textwidth]{figures/gammaR_distribn_mag_limited.pdf}
 	\end{center}
@@
 \int_{0}^{d_{\rm max}(q)} {\rm prob}_{\rm ml}(r,q) \,{\rm d}r \\
-&= \mathcal{Z}\cdot (1+q^{3.5})^{3/2}\quad {\rm if\ }0.1<q\leq 1,
+&= \mathcal{Z}\cdot (1+q^{7/2})^{3/2}\quad {\rm if\ }0.1<q\leq 1,
 \label{eq:pdf_q_mag_limited}
@@
 \begin{align}
-{\rm prob}(N_d) = \frac{2\mathcal{Z}}{3\alpha}& N_d^{1/3} \left(\frac{3}{4\pi 
-	n_d}\right)^{4/3} d_{\rm max,s}^{-4} \ \nonumber\\
+{\rm prob}(N_d) \propto N_d^{2/3} &\left(\frac{3}{4\pi 
+	n_d}\right)^{5/3} d_{\rm max,s}^{-4} \ \nonumber\\
 &\times \left[ 
@@
 \begin{equation}
-{\rm prob_{ml}}(\gamma_R) = \frac{\mathcal{Z}}{\alpha} (1+\gamma_R) 
+{\rm prob_{ml}}(\gamma_R) \propto (1+\gamma_R)^{3/2} 
 \gamma_R^{\frac{1-\alpha}{\alpha}},
@@
 For $\alpha=3.5$, ${\rm prob_{ml}}(\gamma_R) \propto 
-(1+\gamma_R)\gamma_R^{-5/7}$, which tells us the distribution will be highly 
-biased towards low-light ratio binaries.
-The distribution is plotted in Fig.~\ref{fig:gammaR_distribn_mag_limited}.
+(1+\gamma_R)^{3/2}\gamma_R^{-5/7}$, which tells us the distribution will be 
+highly biased towards low-light ratio binaries, as shown in 
+Fig.~\ref{fig:gammaR_distribn_mag_limited}.
 
@@
 \end{equation}
-With some algebra, and letting $\alpha=3.5$, this can be expressed as
-\begin{equation}
-\langle N_d \rangle = \frac{2 \mathcal{Z}}{7 R^{1/2}}
-\int_{(0.1^\alpha + 1)^{1/3}}^{\sqrt{2}} u^{3/2} (u-1)^{5/7} \,{\rm d}u ,
+Letting $\alpha=3.5$, this can be expressed as
+\begin{equation}
+\langle N_d \rangle \propto \mathcal{R}^{-3/2}
+\int_{0.1^{7/2} + 1}^{2} u^{3/2} (u-1)^{5/7} \,{\rm d}u ,
 \label{eq:integral_Nd}
 \end{equation}
-for
-\begin{equation}
-R \equiv \left(\frac{3}{4\pi n_d}\right)^{2/3} d_{\rm max,s}^{-2}.
+where non-dimensional constants are not shown, and $\mathcal{R}$ is defined as
+\begin{equation}
+\mathcal{R} \equiv \left(\frac{3}{4\pi n_d}\right)^{2/3} d_{\rm max,s}^{-2}.
 \end{equation}
 The integral in Eq.~\ref{eq:integral_Nd} has a closed-form solution in terms 
-of Gauss hypergeometric functions, $_2F_1$. Once evaluated, the integral takes 
-a numerical value of $0.1829$.
-The normalization of
-
-%TODO GIVE THE EXPLICIT FORM, SEMI-ANALYTICALLY.
-
-While a closed analytic expression for $\langle N_d \rangle$ does exist, it is 
-messy, and it does not yield much intuition. Instead, summarizing the important 
-points and supporting them numerically:
-\begin{itemize}
-	\item In a volume-limited sample of binary star systems in which the 
-	primary mass is fixed, and the mass ratio is drawn from a bounded uniform 
-	distribution, the distribution of $\gamma_R$ will be biased towards low 
-	values ($\gamma_R \approx 0.1$). This is shown in 
-	Fig.~\ref{fig:gammaR_distribn_vol_limited}.
-	\item In a magnitude-limited sample of binary star systems in which the 
-	primary mass is fixed, and the mass ratio of the \textit{population} is 
-	drawn from a bounded uniform distribution, the observed distribution of 
-	mass ratios will be biased towards high values. You will see more twins,
-	because they are detectable out to a greater distance\footnote{This is a 
-	scary systematic w.r.t. the claimed intrinsic excess of twin binaries.}. 
-	This is shown in Fig.~\ref{fig:q_distribn_mag_limited}, and explained 
-	analytically below.
-	\item In a magnitude-limited sample of binary star systems in which the 
-	primary mass is fixed, the distribution of $\gamma_R$ will be biased 
-	towards low values ($\approx 0.1$), but less so than in a 
-	volume-limited sample. This is shown in 
-	Fig.~\ref{fig:gammaR_distribn_mag_limited}.
-\end{itemize}
-
-In passing, the bias in ${\rm prob}(\gamma_R)$ towards low luminosity ratios 
-can be seen analytically. If we assume $\gamma_R = q^\alpha$, then
-\begin{equation}
-{\rm prob}(\gamma_R) = \frac{1}{9 \alpha} \ \gamma_R^{\frac{1-\alpha}{\alpha}}
-\quad {\rm for\ } (0.1)^\alpha < \gamma_R < 1,
-\end{equation}
-and otherwise zero. For instance if $\alpha = 3$, ${\rm prob}(\gamma_R) 
-\propto \gamma_R^{-2/3}$ and the domain extends from $1$ to 
-$10^{-3}$, where the probability distribution peaks.
-
-\paragraph{How many stars are in the sample?}
-
-We return to the original question: how many stars?
-\begin{equation}
-\langle N_{\rm stars} \rangle = N_s + 2 \langle N_d \rangle,
-\end{equation}
-as before.
-Since we are not giving analytic expressions for either of the $\langle \ldots 
-\rangle$ terms, we just note that in the case of a population with a single 
-$\gamma_R$ value, $N_d/N_s = {\rm BF} \times (1 + \gamma_R)^{3/2}$. A lower
-bound for this ratio will always be the binary fraction.
-For a single run of a Monte Carlo code, where ${\rm mean}(\gamma_R) = 0.28$, 
-${\rm median}(\gamma_R) = 0.13$, and the distribution was that given in 
-Fig.~\ref{fig:gammaR_distribn_mag_limited}, the observed ratio was $N_d/N_s = 
-0.59$.
-This corresponds to a single-valued population with $\gamma_R \sim 0.20$, a
-number between the mean and median of the true distribution.
-It means $\sim 1.2$ stars in binary star systems for every star in a single 
-star system from this sample.
-
-We also note in passing that $N_d/N_s = 0.59$ is a smaller ratio than for the 
-case of a fixed $\gamma_R$ population with $\gamma_R = 1$, which gave $N_d/N_s 
-= 1.27$. The difference is that in the latter population, the volume of 
-searchable stars is greater. Based on the fixed-$\gamma_R$ population scaling 
-law $N_d / N_s \propto (1+\gamma_R)^{3/2}$, we 
-would expect the ratio of single stars to be $\sim (2/1.1)^{3/2} = 2.4$ 
-times smaller, which is roughly (though not exactly) what we observe.
+of the Gauss hypergeometric function, $_2F_1$. Once evaluated, and after 
+normalizing ${\rm prob}(N_d)$, to 4 significant digits it gives
+\begin{equation}
+\langle N_d \rangle = 1.601 \left(\frac{4\pi n_d}{3}\right) d_{\rm max,s}^3.
+\end{equation}
+This gives the useful expression
+\begin{align}
+\frac{\langle N_d \rangle}{N_s} &= 1.601\, \frac{n_d}{n_s} \\
+&= 1.601 \left(\frac{{\rm BF}}{1-{\rm BF}}\right).
+\end{align}
+Taking the sun-like star binary fraction of ${\rm BF} = 0.44$, this gives
+$\langle N_d \rangle / N_s = 1.258$.
+Compared to the binary-twin thought experiment of Sec.~\ref{sec:model_1}, in 
+which the SNR-limited sample had $\approx 2.8$ times more binary systems than 
+single systems, there are now $\approx 1.3$ times as many.
+This is sensible, because the mean light ratio has decreased; it was 
+previously 1, it is now $\langle \gamma_R \rangle \approx 0.27$.
+
+\paragraph{What is the true occurrence rate?}
+
+Returning to the original question, the total number of stars in the sample is
+\begin{equation}
+\langle N_{\rm stars,tot} \rangle = N_s + 2 \langle N_d \rangle.
+\end{equation}
+For stars in binary systems, only a fraction of them are of the desired type.
+Defining the ``desired type'' to be $0.7<M_\star/M_\odot<1.3$, this can be 
+evaluated by integrating Eq.~\ref{eq:pdf_q_mag_limited} from $0.7M_\odot$ to 
+$M_\odot$.
+The result is that $47.5\%$ of secondaries are of the desired type.
+Thus the expected number of stars of the desired type in the sample is
+\begin{equation}
+\langle N_{\rm stars} \rangle = N_s + 1.475 \langle N_d \rangle.
+\end{equation}
+
+The number of planets of the desired type is then
+\begin{equation}
+\langle N_{\rm planets} \rangle = 
+\Gamma_{t,s}N_s + \Gamma_{t,d}\langle N_d\rangle + 
+0.475\Gamma_{t,d}w_{d2} \langle N_d\rangle,
+\end{equation}
+and the true occurrence rate is
+\begin{align}
+\Gamma_t &= \frac{\langle N_{\rm planets} \rangle }{\langle N_{\rm stars} 
+\rangle} \\
+\Gamma_t &= 
+\frac{
+	\Gamma_{t,s}N_s + (1+0.475w_{d2})\Gamma_{t,d}\langle N_d\rangle
+	}{
+	N_s + 1.475 \langle N_d \rangle
+	}.
+\end{align}
+
+% % % % % % % % % % % % % % % % % % % % % % % % % % % %
+\subsection{What is the measured occurrence rate?}
+
+The observer who has never heard of binary star systems wants to find the 
+occurrence rate of planets with radius $R_p$ and period $P$, orbiting stars 
+with mass $0.7<M_\star/M_\odot < 1.3$ (with corresponding stellar radii and 
+luminosities).
+They detected $N_{\rm det}$ transit signals that appear to be from planets of 
+this desired type.
+The occurrence rate they then find, {\it i.e.} the number of planets divided 
+by number of ``stars'' (really, stellar systems) is
+\begin{equation}
+\Gamma_{\rm a } = \frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_g}.
+\end{equation}
+
+Unlike in Sec.~\ref{sec:model_1}, where the observer saw signals of only two 
+apparent radii (undiluted and diluted), they will now find signals with a 
+spectrum of apparent radii.
+The dilutions $\mathcal{D}$ in binary systems, given by Eq.~\ref{eq:dilution}, 
+vary depending on which star the planet orbits, and depending on the system's 
+luminosity ratio $\gamma_R$.
+Moreover, the apparent planetary radii will be affected by the assumptions the 
+observer makes about the apparently-single star's radius, $R_{\star,{\rm a}}$.
+We posit that the observer assumes $R_{\star,{\rm a}} = R_1$; in other words 
+they assume that the primary star hosts any signal they observe.
+The apparent planetary radii, $R_{\rm p, a}$, are then
+\begin{equation}
+R_{p,{\rm a}} = R_p 
+\frac{R_1}
+{R_{\star,{\rm t}}}
+\,\mathcal{D}^{1/2}.
+\end{equation}
+
+The total number of planet detections of the desired type is 
+\begin{equation}
+\langle N_{\rm det}\rangle = 
+N_{{\rm det},s} 
++ 
+\langle N_{{\rm det},d}\rangle.
+\end{equation}
+As in Sec.~\ref{sec:model_1},
+\begin{equation}
+N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,g} f_{s,c},
+\label{eq:N_det_s_model2}
+\end{equation}
+where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
+star systems of the sample, $f_{s,g}$ is the geometric transit probability, 
+and $f_{s,c}$ is the fraction of these transiting planets that are observed 
+with signal to noise greater than the minimum detection threshold (the 
+completeness). 
+For our SNR-limited, varying-$\gamma_R$ survey, $f_{s,g} = R_1/a\equiv f_g$, 
+and $f_{s,c}=1$.
+
+For planets detected in binary systems, only a fraction will be ``of the 
+desired type'', since only a fraction of the stars are of the desired type.
+This is the same factor $f_{\rm desired}$ that we have shown is 0.475. Thus
+\begin{equation}
+\langle N_{{\rm det},d} \rangle
+= 
+(1+0.475w_{d2}) \langle N_d\rangle \Gamma_{t,d} f_{d,g} f_{d,c}.
+\label{eq:N_det_d_model2}
+\end{equation}
+Now $(1+0.475w_{d2}) \langle N_d\rangle \Gamma_{t,d}$ is the number of 
+planets of the desired type in the double star systems of the sample.
+The geometric transit probability $f_{d,g}$ and completeness $f_{d,c}$ require 
+more careful evaluation.
+First, separate them into their sub-components about primaries ``d1'' and 
+secondaries ``d2'':
+\begin{align}
+\langle N_{{\rm det},d} \rangle
+= 
+\langle N_d\rangle \Gamma_{t,d} ( f_{d1,g} f_{d1,c} 
++
+0.475w_{d2}  f_{d2,g} f_{d2,c}
+)
+\end{align}
+
+The geometric transit probability $f_{d,g}=R_1/a$ 
+for our twin-binary survey, and $f_{d,c}=1/8$ by the geometric argument 
+presented in the initial discussion of limiting distances.
+
+The occurrence rate this astronomer derives for planets of assumed 
+radius $R_p$ orbiting stars thought to be of the desired type is then
+\begin{equation}
+\Gamma_{\rm a} = \frac{\Gamma_{t,s}N_s}{N_s + N_d}.
+\end{equation}
+
+\subsection{Errors on derived occurrence rates for Model \#2}
 
@@
 \begin{itemize}
-	\item Assume a binary fraction ${\rm BF} = 0.45$ [Raghavan et al 2010].
+	\item Assume a binary fraction ${\rm BF} = 0.44$ [Raghavan et al 2010].
 	\item Ignore higher order multiples.

### 2017-09-25 src/completeness_semianalytic.py

@@ -1,0 +1,25 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+alpha = np.linspace(1,4,num=2000)
+f,ax=plt.subplots()
+ax.plot(alpha, (1-alpha)/alpha)
+
+f.savefig('alpha_power_allowed_vals.pdf',
+        bbox_inches='tight')
+
+
+alpha = 3.5
+f_d1c_upper = (1+0.1**alpha)**(-3)
+f_d1c_lower = 2**(-3)
+
+f_d1c = np.linspace(f_d1c_lower, f_d1c_upper, num=int(1e4))
+
+import IPython; IPython.embed()
+integrand = f_d1c**5 * (f_d1c**3 -1)**(-5/7)
+
+f,ax=plt.subplots()
+ax.plot(f_d1c, integrand)
+
+f.savefig('temp.pdf',
+        bbox_inches='tight')

### 2017-10-01 paper/ms.tex
paper: minor edits

@@
 {\rm prob}(N_d) \propto N_d^{2/3} &\left(\frac{3}{4\pi 
-	n_d}\right)^{5/3} d_{\rm max,s}^{-4} \ \nonumber\\
+	n_d}\right)^{5/3} d_{\rm max,s}^{-5} \ \nonumber\\
 &\times \left[ 
@@
 \langle N_d \rangle \propto \mathcal{R}^{-3/2}
-\int_{0.1^{7/2} + 1}^{2} u^{3/2} (u-1)^{5/7} \,{\rm d}u ,
+\int_{0.1^{7/2} + 1}^{2} u^3 (u-1)^{-5/7} \,{\rm d}u ,
 \label{eq:integral_Nd}
@@
 The integral in Eq.~\ref{eq:integral_Nd} has a closed-form solution in terms 
-of the Gauss hypergeometric function, $_2F_1$. Once evaluated, and after 
+of polynomials. Once evaluated, and after 
 normalizing ${\rm prob}(N_d)$, to 4 significant digits it gives
 \begin{equation}
-\langle N_d \rangle = 1.601 \left(\frac{4\pi n_d}{3}\right) d_{\rm max,s}^3.
+\langle N_d \rangle = 1.590 \left(\frac{4\pi n_d}{3}\right) d_{\rm max,s}^3.
 \end{equation}
@@
 \begin{align}
-\frac{\langle N_d \rangle}{N_s} &= 1.601\, \frac{n_d}{n_s} \\
-&= 1.601 \left(\frac{{\rm BF}}{1-{\rm BF}}\right).
+\frac{\langle N_d \rangle}{N_s} &= 1.590\, \frac{n_d}{n_s} \\
+&= 1.590 \left(\frac{{\rm BF}}{1-{\rm BF}}\right).
 \end{align}
 Taking the sun-like star binary fraction of ${\rm BF} = 0.44$, this gives
-$\langle N_d \rangle / N_s = 1.258$.
+$\langle N_d \rangle / N_s = 1.25$.
 Compared to the binary-twin thought experiment of Sec.~\ref{sec:model_1}, in 
 which the SNR-limited sample had $\approx 2.8$ times more binary systems than 
-single systems, there are now $\approx 1.3$ times as many.
+single systems, there are now $\approx 1.25$ times as many.
 This is sensible, because the mean light ratio has decreased; it was 

### 2017-10-01 src/completeness_semianalytic.py

@@ -1,25 +1,28 @@
+from scipy.integrate import trapz
 import matplotlib.pyplot as plt
 import numpy as np
-
-alpha = np.linspace(1,4,num=2000)
-f,ax=plt.subplots()
-ax.plot(alpha, (1-alpha)/alpha)
-
-f.savefig('alpha_power_allowed_vals.pdf',
-        bbox_inches='tight')
-
 
 alpha = 3.5
 f_d1c_upper = (1+0.1**alpha)**(-3)
 f_d1c_lower = 2**(-3)
 
-f_d1c = np.linspace(f_d1c_lower, f_d1c_upper, num=int(1e4))
+f_d1c = np.linspace(f_d1c_lower, f_d1c_upper, num=int(5e4))
 
-import IPython; IPython.embed()
-integrand = f_d1c**5 * (f_d1c**3 -1)**(-5/7)
+integrand = f_d1c**5 * (np.abs(f_d1c**3 -1))**(-5/7)
+
+norm = trapz(integrand, f_d1c)
+print('normalization is: {:.8f}'.format(norm))
+
+integrand /= norm
 
 f,ax=plt.subplots()
 ax.plot(f_d1c, integrand)
 
-f.savefig('temp.pdf',
+ax.set(
+xlabel='$f_{d1,c}$',
+ylabel='prob',
+yscale='log'
+)
+
+f.savefig('prob_fd1c.pdf',
         bbox_inches='tight')

### 2017-10-03 paper/ms.tex
wordsmithed; did needlessly hard integrals

@@
 component is of the desired type.
-For single stars, this may simply be from erroneous stellar parameters.
-In binaries, one might imagine the combined light of two stars somehow making 
-them resemble a single star of the desired type.  We will 
-not consider errors of this type.  We will assume that all the 
+%For single stars, this may simply be from erroneous stellar parameters.
+%In binaries, one might imagine the combined light of two stars somehow making 
+%them resemble a single star of the desired type.  
+We will not consider errors of this type.  We will assume that all the 
 apparently-searchable stars are either single stars of
@@
   For $\gamma_R = 1$, this means only 1 in 8 apparently-searchable stars which 
-  are in fact binary systems can yield detectable planets.
+  are in fact binary systems can yield detectable planets (the ratio of 
+  volume in which planets can be detected to the actual selected volume).
+  For a more realistic sample, $\gamma_R \approx 0.1$, and there is an 
+  interesting asymmetry.
+  One finds the completeness for planets which orbit the primary star, 
+  $f_{d1,c}$ to be
+  \begin{equation}
+  f_{d1,c} = (1+\gamma_R)^{-3} \approx 3/4.
+  \end{equation}
+  However, for planets which orbit the secondary star, the completeness is
+  \begin{equation}
+    f_{d2,c} = (1+\gamma_R^{-1})^{-3} \approx 3/4000.
+  \end{equation}
 
@@
 
-The model of Sec.~\ref{sec:model_1} is of course idealized, but it helps
-for developing intuition.
-We now let the light ratio $\gamma_R = L_2/L_1$ vary across the population 
+The model of Sec.~\ref{sec:model_1} is idealized, but it helps
+develop intuition.
+Now, we let the light ratio $\gamma_R = L_2/L_1$ vary across the population 
 of star systems.
@@
 \begin{equation}
-f(q) =
+{\rm prob}(q) =
 \Bigg\{\begin{array}{lr}
@@
 = 1/9$ is the normalization.
-We use this latter mass ratio distribution.
+We choose this latter mass ratio distribution.
 
@@
 $L \propto M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
-As a simple check, we fit lines to mass-luminosity data collected by 
-Torres et al. [2009] for dwarfs earlier than spectral class M, and Benedict 
-et al. [2016] for dwarfs later than spectral class M.
+As a simple check, we fit lines to mass-luminosity data\footnote{
+    Data collected by Torres et al. [2009] for dwarfs earlier than 
+    spectral class M, and Benedict et al. [2016] for dwarfs later than 
+    spectral class M.
     To convert Benedict et al. [2016]'s reported $M_V$ values to absolute 
     luminosities, we interpolated over E. Mamajek's 
-table\footnote{
+    table
 	\texttt{pas.rochester.edu/\textasciitilde 
@@
 	downloaded 2017.08.02}.
-In log-log space, we let the intersection point of the lines float, and made 
-various cuts on the data as indicated in Fig~\ref{fig:mass_luminosity}. The 
-fit parameters are available in a footnote\footnote{
+In log-log space, we let the intersection point of the lines float. The 
+resulting fit parameters are available in a footnote\footnote{
 	m lo: 1.8818719873988132~--~
@@
 secondaries of binary star systems.
+Since again, our main purpose is to develop intuition about observational 
+biases, we approximate this as 
+\begin{equation}
+R\propto M.
+\end{equation}
+\begin{comment}
 [Demircan \& Kahraman 1991]'s empirical fit to eclipsing 
@@
 1991's stated bounds), where each quantity is given in solar units. 
-Since again, our main purpose is to develop intuition about observational 
-biases, we approximate this as $R\propto M$.
+\end{comment}
 
@@
 \begin{equation}
-N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3(\gamma_R)
+N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3
 \end{equation}
@@
 \begin{equation}
-d_{\rm max,d}(\gamma_R) =
+d_{\rm max,d} =
 d_{\rm max,s}\times (1+\gamma_R)^{1/2},
@@
 
-For a given binary system, $q$, $\gamma_R$, and $d_{\rm 
+For a given binary system, $q$, $\gamma_R$, and thus $d_{\rm 
 	max,d}(\gamma_R)$ are all random variables.
 Since $N_d$ is a function of $d_{\rm max,d}$, 
-the number of double star systems becomes a random variable.
+the number of double star systems becomes a random variable, drawn from the 
+ensemble of all possible transit surveys.
 Applying the chain rule for probability density functions, the distribution 
@@
 \left| \frac{{\rm d}d_{\rm max,d}}{{\rm d}N_d}  \right|.
+\label{eq:prob_Nd_raw}
 \end{equation}
@@
 	\caption{The joint probability distribution of a binary star's position and 
-		mass ratio in a magnitude-limited sample. This plot assumes a different 
-		$M(L)$ relation than previous plots: $L=M^3$. Note that the marginalized 
+		mass ratio in a magnitude-limited sample. This plot assumes $L\propto 
+		M^{3.5}$. Note that the marginalized 
 		mass ratio distribution is uniform at any given distance.}
@@
 \end{equation} 
-where ${\rm prob}(q|r)$ is given by Eq.~\ref{eq:mass_ratio}, and 
-\begin{align}
-{\rm prob}_{\rm ml}(r) &= \int_{0}^{1} {\rm prob}_{\rm ml}(r|q)\,{\rm 
-	prob}(q)\,{\rm d}q
-\\
-&= \int_{0.1}^{1} \frac{1}{3} \frac{r^2}{d_{\rm max,d}^3(q)}\,{\rm d}q
-\quad {\rm if\ }r<d_{\rm max}(q),
-\end{align}
+where ${\rm prob}(q|r)$ is given by Eq.~\ref{eq:mass_ratio}.
+One can show
+\begin{equation}
+{\rm prob}_{\rm ml}(r,q) \propto r^2 \quad{\rm if}\ r<d_{\rm max,d}(q)\ {\rm 
+and}\ 0.1<q<1, 
+\end{equation}
 and otherwise zero. The resulting joint probability distribution is plotted in 
 Fig.~\ref{fig:joint_prob_r_q}.
-
 Marginalizing the joint distribution over distance, and assuming $L\propto 
-M^{3.5}$,
+M^\alpha$, one finds
 \begin{align}
-{\rm prob}_{\rm ml}(q) &= 
-\int_{0}^{d_{\rm max}(q)} {\rm prob}_{\rm ml}(r,q) \,{\rm d}r \\
-&= \mathcal{Z}\cdot (1+q^{7/2})^{3/2}\quad {\rm if\ }0.1<q\leq 1,
+{\rm prob}_{\rm ml}(q) \propto (1+q^\alpha)^{3/2}\quad {\rm if\ 
+}0.1<q\leq 1,
 \label{eq:pdf_q_mag_limited}
 \end{align}
-and otherwise 0, for $\mathcal{Z}$ the appropriate normalization constant.
+and otherwise 0.
 This marginalized distribution~--~the distribution function of mass ratios for 
@@
 Fig.~\ref{fig:q_distribn_mag_limited}.
-
-More correctly then, we must write
-\begin{equation}
-{\rm prob}(N_d) = {\rm prob_{ml}}(q(\gamma_R)) 
-\left| \frac{{\rm d}q}{{\rm d}\gamma_R}  \right|
-\left| \frac{{\rm d}\gamma_R}{{\rm d}d_{\rm max,d}}  \right|				
-\left| \frac{{\rm d}d_{\rm max,d}}{{\rm d}N_d}  \right|.
-\end{equation}
-
-Doing some algebra, and assuming $\gamma_R = q^\alpha$, this can be shown to be
+Reworded, the probability of drawing a binary of mass ratio 
+$q$ in a magnitude-limited sample scales as $d_{\rm max,d}^3(q)$.
+
+Applying Eq.~\ref{eq:prob_Nd_raw} with the correct magnitude-limited 
+distributions gives
+\begin{equation}
+{\rm prob}(N_d) \propto N_d^{2/3} 
+\left( \left(\frac{N_d n_s}{N_s n_d}\right)^{2/3} - 1
+\right)^{\frac{1-\alpha}{\alpha}}
+\end{equation}
+\begin{comment}
 \begin{align}
@@
 \end{align}
+\end{comment}
 over the interval
 \begin{equation}
-N_d^{\rm lower} = \frac{4 \pi n_d}{3} (\sqrt{0.1^3 + 1} d_{\rm max,s})^3,
+N_d^{\rm lower} = \frac{4 \pi n_d}{3} (\sqrt{(0.1)^\alpha + 1} d_{\rm 
+max,s})^3,
 \end{equation}
@@
 searchable binary stars.
-The expected value for the number of double star systems in the sample is
+For any given realization of the survey, this number is a random draw.
+Computing the \textit{expected value} for the number of double star systems in 
+the sample,
 \begin{equation}
@@
 \end{equation}
-where non-dimensional constants are not shown, and $\mathcal{R}$ is defined as
-\begin{equation}
-\mathcal{R} \equiv \left(\frac{3}{4\pi n_d}\right)^{2/3} d_{\rm max,s}^{-2}.
+where non-dimensional constants are not shown,
+$u = \mathcal{R} N_d^{2/3}$,
+and $\mathcal{R}$ is defined as
+\begin{equation}
+\mathcal{R} \equiv \left(\frac{3}{4\pi n_d}\right)^{2/3} d_{\rm max,s}^{-2}
+= \left(\frac{n_s}{N_s n_d}\right)^{2/3}.
 \end{equation}
@@
 \begin{align}
-\frac{\langle N_d \rangle}{N_s} &= 1.590\, \frac{n_d}{n_s} \\
-&= 1.590 \left(\frac{{\rm BF}}{1-{\rm BF}}\right).
+\frac{\langle N_d \rangle}{N_s} &= 1.590\, \frac{n_d}{n_s}
 \end{align}
@@
 
-Returning to the original question, the total number of stars in the sample is
-\begin{equation}
-\langle N_{\rm stars,tot} \rangle = N_s + 2 \langle N_d \rangle.
-\end{equation}
-For stars in binary systems, only a fraction of them are of the desired type.
-Defining the ``desired type'' to be $0.7<M_\star/M_\odot<1.3$, this can be 
-evaluated by integrating Eq.~\ref{eq:pdf_q_mag_limited} from $0.7M_\odot$ to 
-$M_\odot$.
-The result is that $47.5\%$ of secondaries are of the desired type.
-Thus the expected number of stars of the desired type in the sample is
-\begin{equation}
-\langle N_{\rm stars} \rangle = N_s + 1.475 \langle N_d \rangle.
-\end{equation}
-
-The number of planets of the desired type is then
-\begin{equation}
-\langle N_{\rm planets} \rangle = 
-\Gamma_{t,s}N_s + \Gamma_{t,d}\langle N_d\rangle + 
-0.475\Gamma_{t,d}w_{d2} \langle N_d\rangle,
-\end{equation}
-and the true occurrence rate is
+Returning to the original question, the total number of stars for any given 
+realization of the survey is
+\begin{equation}
+N_{\rm stars,tot} = N_s + 2 N_d.
+\end{equation}
+Since $N_d$ is a random number, so is the total number of stars.
+By assumption all primaries are of the desired type, but for secondaries only 
+a fraction $f_{\rm desired}$ of them are of the desired type:
+\begin{equation}
+N_{\rm stars} = N_s + (1+f_{\rm desired}) N_d.
+\end{equation}
+Explicitly,
+\begin{equation}
+f_{\rm desired} = \frac{\int_{M_{\rm min}}^{M_{\rm max}} {\rm prob}_{\rm 
+ml}(q)\,{\rm d}q }
+{\int_{0.1M_\odot}^{M_\odot} {\rm prob}_{\rm 
+        ml}(q)\,{\rm d}q}.
+\end{equation}
+For instance, if we define the ``desired type'' to be 
+$0.7<M_\star/M_\odot<1.3$, we find $f_{\rm desired}\approx0.475$.
+
+For a given survey draw, the number of planets of the desired type is 
+then
+\begin{equation}
+N_{\rm planets} = 
+\Gamma_{t,s}N_s + \Gamma_{t,d} N_d + 
+f_{\rm des}\Gamma_{t,d}w_{d2} N_d,
+\end{equation}
+and the drawn ``true occurrence rate'' is
 \begin{align}
-\Gamma_t &= \frac{\langle N_{\rm planets} \rangle }{\langle N_{\rm stars} 
-\rangle} \\
+\Gamma_t &= \frac{ N_{\rm planets}  }{ N_{\rm stars} } \\
 \Gamma_t &= 
 \frac{
-	\Gamma_{t,s}N_s + (1+0.475w_{d2})\Gamma_{t,d}\langle N_d\rangle
+	\Gamma_{t,s}N_s + (1+f_{\rm desired}w_{d2})\Gamma_{t,d} N_d
 	}{
-	N_s + 1.475 \langle N_d \rangle
+	N_s + (1+f_{\rm desired}) N_d
 	}.
@@
 this desired type.
-The occurrence rate they then find, {\it i.e.} the number of planets divided 
-by number of ``stars'' (really, stellar systems) is
+The occurrence rate they then find, {\it i.e.} the apparent number of planets 
+divided by the apparent number of ``stars'' (really, stellar systems) is
+\begin{comment}
 \begin{equation}
 \Gamma_{\rm a } = \frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_g}.
+\end{equation}
+\end{comment}
+\begin{equation}
+\Gamma_{\rm a } =
+\frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_{s,g}}.
 \end{equation}
@@
 
-The total number of planet detections of the desired type is 
-\begin{equation}
-\langle N_{\rm det}\rangle = 
+The drawn number of planet detections of the desired type is 
+\begin{equation}
+N_{\rm det} = 
 N_{{\rm det},s} 
 + 
-\langle N_{{\rm det},d}\rangle.
+N_{{\rm det},d}.
 \end{equation}
@@
 desired type'', since only a fraction of the stars are of the desired type.
-This is the same factor $f_{\rm desired}$ that we have shown is 0.475. Thus
-\begin{equation}
-\langle N_{{\rm det},d} \rangle
+This is the same factor $f_{\rm desired}$ as previously discussed. Accounting 
+for this and the independent geometric and completeness probabilities,
+\begin{equation}
+N_{{\rm det},d}
 = 
-(1+0.475w_{d2}) \langle N_d\rangle \Gamma_{t,d} f_{d,g} f_{d,c}.
+N_d \Gamma_{t,d} (f_{d1,g}f_{d1,c} + f_{\rm desired}w_{d2}f_{d2,g}f_{d2,c}).
 \label{eq:N_det_d_model2}
 \end{equation}
-Now $(1+0.475w_{d2}) \langle N_d\rangle \Gamma_{t,d}$ is the number of 
-planets of the desired type in the double star systems of the sample.
-The geometric transit probability $f_{d,g}$ and completeness $f_{d,c}$ require 
-more careful evaluation.
-First, separate them into their sub-components about primaries ``d1'' and 
-secondaries ``d2'':
+
+For any given system,
 \begin{align}
-\langle N_{{\rm det},d} \rangle
-= 
-\langle N_d\rangle \Gamma_{t,d} ( f_{d1,g} f_{d1,c} 
-+
-0.475w_{d2}  f_{d2,g} f_{d2,c}
-)
+f_{d1,c} &= (1+\gamma_R)^{-3} \\
+f_{d1,g} &= f_{s,g} \\
+f_{d2,c} &= (1+\gamma_R^{-1})^{-3} \\
+f_{d2,g} &= f_{s,g}q^{2/3}
 \end{align}
 
-The geometric transit probability $f_{d,g}=R_1/a$ 
-for our twin-binary survey, and $f_{d,c}=1/8$ by the geometric argument 
-presented in the initial discussion of limiting distances.
-
-The occurrence rate this astronomer derives for planets of assumed 
-radius $R_p$ orbiting stars thought to be of the desired type is then
-\begin{equation}
-\Gamma_{\rm a} = \frac{\Gamma_{t,s}N_s}{N_s + N_d}.
-\end{equation}
-
-\subsection{Errors on derived occurrence rates for Model \#2}
-
+\subsection{Errors on derived occurrence rates}
+
+Recall $X_\Gamma = \Gamma_t / \Gamma_a$. Let $f_{\rm desired} = f_{\rm d}$ for 
+brevity.
+Combining the above results, for any draw from the ensemble of transit 
+surveys the necessary correction to the derived occurrence rate is
+\begin{align}
+X_\Gamma =
+&\frac{\Gamma_{t,s}N_s + (1+f_{\rm d}w_{d2})\Gamma_{t,d}N_d }{N_s + (1+f_{\rm 
+d})N_d}
+\nonumber \\
+&\times
+\frac{(N_s + N_d)f_{s,g}}{N_s \Gamma_{t,s} f_{s,g} + N_d \Gamma_{t,d}
+      (f_{d1,g}f_{d1,c} + f_{\rm d}w_{d2}f_{d2,g}f_{d2,c})  }.
+\end{align}
+Letting $\beta = N_d/N_s$, 
+\begin{align}
+X_\Gamma =
+&\frac{1+\beta}{1 + \beta(1+f_{\rm d})} 
+\nonumber \\
+&\times
+\frac{ 1+ \beta (\Gamma_{t,d}/\Gamma_{t,s}) (1+f_{\rm d}w_{d2}) }
+{1 +\beta(\Gamma_{t,d}/\Gamma_{t,s})(f_{d1,c} + f_{\rm 
+d}w_{d2}f_{d2,g}f_{d2,c}/f_{s,g})}.
+\end{align}
+
+One typically assumes that $\Gamma_{t,s}=\Gamma_{t,d}$, and that the 
+fraction of secondaries of the desired type with planets, $w_{d2}$ can be set 
+to 1.
+If so,
+\begin{align}
+X_\Gamma =
+&\frac{1+\beta}
+{1 +\beta(f_{d1,c} + f_{\rm d}f_{d2,g}f_{d2,c}/f_{s,g})}.
+\label{eq:model2_xgamma_simple}
+\end{align}
+
+If $\gamma_R \ll 1$, $f_{d1,c}\approx 1$, $f_{d2,c}\ll f_{d1,c}$, and the 
+number of detections from secondaries of binary systems is negligible.
+In this limit, binarity does not affect the derived occurrence rate:
+\begin{equation}
+X_\Gamma = 1 \quad ({\rm limit\ of}\ \gamma_R \ll 1).
+\end{equation}
+
+To assess whether it is reasonable to take this limit, we want the mean
+value of $X_\Gamma$, or the median, or else some other summary statistic.
+Recall that all the terms in Eq.~\ref{eq:model2_xgamma_simple} are
+random variables, and note that they are inter-dependent.
+This latter fact impedes analytic progress.
+
+Running a Monte Carlo simulation in which we fix the planet and primary star 
+properties, draw the binary mass ratio from Eq.~\ref{eq:pdf_q_mag_limited}, 
+and take 
+the luminosity as $L\propto M^{3.5}$, we find 
+a mean of $\langle X_\Gamma \rangle ={\bf VALUE HERE}$,
+and a median of ${\rm median}(X_\Gamma) ={\bf VALUE HERE}$.
+The distribution is shown in Figure BLAH.
+
+\begin{comment}
 \subsection{How many planets are in the sample?}
@@
 
+\end{comment}
+
 % \begin{comment}

### 2017-10-03 src/binary_distribution.py

@@ -320,7 +320,7 @@ if __name__ == '__main__':
     # STELLAR POPULATION #
     ######################
     # Binary fraction. BF = n_d / n_s.
-    BF = 0.45
+    BF = 0.44
     # Bovy (2017) gives n_tot = n_s + n_d. We want the number density of single
     # star systems, n_s, and the number density of double star systems, n_d.
     #n_tot = 4.76e-4 / (u.pc**3)

### 2017-10-03 src/completeness_semianalytic.py

@@ -1,28 +1,215 @@
+import pickle
 from scipy.integrate import trapz
 import matplotlib.pyplot as plt
 import numpy as np
 
-alpha = 3.5
-f_d1c_upper = (1+0.1**alpha)**(-3)
-f_d1c_lower = 2**(-3)
+def primary_star_completeness():
+    '''
+    Find prob(N_d * f_d1,c)
+    and < N_d * f_d1,c >.
 
-f_d1c = np.linspace(f_d1c_lower, f_d1c_upper, num=int(5e4))
+    These are the relevant terms for the observed occ rate.
+    '''
+    alpha = 3.5
+    f_d1c_upper = (1+0.1**alpha)**(-3)
+    f_d1c_lower = 2**(-3)
 
-integrand = f_d1c**5 * (np.abs(f_d1c**3 -1))**(-5/7)
+    f_d1c = np.linspace(f_d1c_lower, f_d1c_upper, num=int(1e4))
 
-norm = trapz(integrand, f_d1c)
-print('normalization is: {:.8f}'.format(norm))
+    # N_s is a survey dependent number. The following verifies that it doesn't
+    # matter. (We just need it to implement semi-analytics)
+    for N_s in np.logspace(3,6,7):
 
-integrand /= norm
+        N_s = int(N_s)
+        R = N_s**(-2/3)
+        BF = 0.44 # star-type dependent binary fraction
+        n_d_by_n_s = BF / (1-BF)
+        N_d_lower = (0.1**(3.5) + 1)**(1/2) * N_s * n_d_by_n_s
+        N_d_upper = (2)**(1/2) * N_s * n_d_by_n_s
 
-f,ax=plt.subplots()
-ax.plot(f_d1c, integrand)
 
-ax.set(
-xlabel='$f_{d1,c}$',
-ylabel='prob',
-yscale='log'
-)
+        # Numerically integrate 17/09/25.6 expression
+        curlyF_min = f_d1c_lower * N_d_lower
+        curlyF_max = f_d1c_upper * N_d_upper
 
-f.savefig('prob_fd1c.pdf',
-        bbox_inches='tight')
+        F_grid = np.linspace(curlyF_min, curlyF_max, num=int(5e3))
+
+        prob_product = []
+
+        # calculate prob_maglimited(N_d * f_{d1,c})
+        for ind, F in enumerate(F_grid):
+
+            if ind%500 == 0:
[…]

### 2017-10-03 src/completeness_test.py

@@ -1,0 +1,23 @@
+import numpy as np
+from scipy.integrate import trapz
+
+x_0, x_1 = 1, 3
+y_0, y_1 = 1, 2
+x = np.linspace(x_0, x_1, num=int(1e3))
+y = np.linspace(y_0, y_1, num=int(2e3))
+
+# 1k x 2k array
+f = x[:,None] * y[None,:]
+
+int1 = trapz(f, y, axis=1)
+
+g = x**2
+
+int2 = trapz(g*int1, x)
+
+print('numeric')
+print(int2)
+
+print('analytic')
+val = 1/8 * (y_1**2 - y_0**2) * (x_1**4 - x_0**4)
+print(val)

### 2017-10-03 src/binary_distribution.py
wrote model 2 for X_Γ monte carlo

@@ -1,6 +1,7 @@
 '''
-what is the light ratio distriubtion for binaries in a magnitude limited
-sample, vs a volume limited sample?
+fixed Rp. fixed primary L1. varying light ratio (L2/L1).
+
+Find the stellar system properties (ONLY!)
 '''
 from __future__ import division, print_function
 
@@ -434,248 +435,3 @@ if __name__ == '__main__':
     print('''\nYou never ran full numerics for this.
     But the above is actually all you need for the Rp rate.
     For the Rp-prime rate, you might want numerics''')
-
-
-
-
-
-#
-#
-#    # Stellar luminosities, radii, Teffs, and masses.
-#    L_2 = L_d - L_1
-#    R_2 = 1*u.Rsun
-#    T_eff2 = (1/c.sigma_sb * L_2 / (4*π*R_2**2))**(1/4)
-#    M_2 = 1*u.Msun
-#
-#    # Dilution parameter
-#    dil = L_1 / L_d
-#
-#    # Surface photon number fluxes
-#    u1_lower = c.h*c.c/(λ_max*c.k_B*T_eff1)
-#    u1_upper = c.h*c.c/(λ_min*c.k_B*T_eff1)
-#    u1 = np.linspace(u1_lower, u1_upper, num=5e4)
-#    integral = np.trapz( u1*u1/(np.exp(u1) - 1),  u1)
-#    F_s1γ_N = 8*π*c.c * (c.k_B * T_eff1 / (c.h * c.c))**3 * integral
-#
-#    u2_lower = c.h*c.c/(λ_max*c.k_B*T_eff2)
-#    u2_upper = c.h*c.c/(λ_min*c.k_B*T_eff2)
-#    u2 = np.linspace(u2_lower, u2_upper, num=5e4)
-#    integral = np.trapz( u2*u2/(np.exp(u2) - 1),  u2)
-#    F_s2γ_N = 8*π*c.c * (c.k_B * T_eff2 / (c.h * c.c))**3 * integral
-#
-#    c_s = R_1**2 * F_s1γ_N
-#    c_d = R_1**2 * F_s1γ_N  +  R_2**2 * F_s2γ_N
-#
-#    #####################
-#    # PLANET POPULATION #
-#    #####################
-#
-#    R_p = 1*c.R_earth
-#    P = 100*u.day
-#
-#    assert M_1 == M_2, 'currently a is written independent of which star '+\
-#                       'is being transited'
-#    a = (P**2 * c.G * M_1 / (4*π*π) )**(1/3)
[…]

### 2017-10-03 src/model_2.py

@@ -1,87 +1,32 @@
 '''
-numerical verification of "model 2" as described in
-memos/toy_analytic_surveys_170804.pdf.
+what is the light ratio distriubtion for binaries in a magnitude limited
+sample, vs a volume limited sample?
 
 fixed Rp. fixed primary L1. varying light ratio (L2/L1).
+
+This code is the Monte Carlo simulation needed to produce summary statistics
+for X_Γ.
 '''
 from __future__ import division, print_function
 
-import numpy as np, pandas as pd, matplotlib.pyplot as plt
+import numpy as np, pandas as pd
 from astropy import units as u, constants as c
 from math import pi as π
+import os
 
+#####################
+# UTILITY FUNCTIONS #
+#####################
 def get_L(M):
+    '''
+    assume input mass, M, a float or vector, is in solar units.
+    '''
 
-    m_lo = 1.8818719873988132
-    c_lo = -0.9799647314108376
-    m_hi = 5.1540712426599882
-    c_hi = 0.0127626185389781
-    M_merge = 0.4972991257826812
-    L_merge = 0.0281260412126928
+    α = 3.5
 
-    L = np.ones_like(M)
-
-    # Method 1: totally fine.
-    #lo_mask = M < M_merge
-    #L[lo_mask] = 10**(np.log10(M[lo_mask])*m_lo + c_lo)
-
-    #hi_mask = M >= M_merge
-    #L[hi_mask] = 10**(np.log10(M[hi_mask])*m_hi + c_hi)
-
-    # Method 2: cuter.
-    L_lo = 10**(np.log10(M)*m_lo + c_lo)
-    L_hi = 10**(np.log10(M)*m_hi + c_hi)
-
-    L = np.maximum(L_lo, L_hi)
+    L = M**α
 
     return L
-
-
-def _make_distribution_plots(vl, doubles):
-
-    # Plot distribution of doubles. First do volume limited case. 
-    plt.close('all')
-    f,ax = plt.subplots(figsize=(4,4))
-    hist, bin_edges = np.histogram(vl['q'], bins=np.append(np.linspace(0,1,51),42),
-            normed=True)
-    ax.step(bin_edges[:-1], hist, where='post')
-    ax.set(xlabel='q', ylabel='prob')
-    ax.set_title(txt_vl, fontsize='small')
-    f.tight_layout()
-    f.savefig('q_distribn_vol_limited.pdf', dpi=250, bbox_inches='tight')
-
-    plt.close('all')
[…]

### 2017-10-04 paper/ms.tex
add model_2_analysis scripts

@@
 
-To assess whether it is reasonable to take this limit, we want the mean
+For the stellar population described in this model, with a mean light ratio 
+$\langle \gamma_R \rangle \approx 0.27$, this limit is not strictly applicable.
+To assess the importance of the error, we want the mean
 value of $X_\Gamma$, or the median, or else some other summary statistic.
@@
 random variables, and note that they are inter-dependent.
-This latter fact impedes analytic progress.
+This latter fact impedes us from analytically expressing these 
+descriptive statistics.
 
@@
 the luminosity as $L\propto M^{3.5}$, we find 
-a mean of $\langle X_\Gamma \rangle ={\bf VALUE HERE}$,
-and a median of ${\rm median}(X_\Gamma) ={\bf VALUE HERE}$.
-The distribution is shown in Figure BLAH.
+a mean of 
+\begin{equation}
+\langle X_\Gamma \rangle =1.291\quad ({\rm Monte\ Carlo\ result}),
+\end{equation}
+and a median that is essentially the same.
+This value is comparable to what one would guess using the analytic results 
+for $\langle N_d \rangle / N_s$ to substitute $\beta$, and $\gamma_R$ to 
+compute $f_{d1,c}$.
 

### 2017-10-04 src/model_2.py

@@ -108,7 +108,7 @@ if __name__ == '__main__':
     #############################
     # !!! BEGIN MONTE CARLO !!! #
     #############################
-    for seed in np.arange(int(4e2),int(5e2),1):
+    for seed in np.arange(int(2e3),int(3e3),1):
 
         ################################################################
         # BEGIN OBNOXIOUS PARAMETERS NEEDED FOR MONTE CARLO SIMULATION #

### 2017-10-04 src/model_2_analysis.py

@@ -1,0 +1,46 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+
+df = pd.read_csv('model2_results.csv')
+
+f,ax = plt.subplots()
+ax.scatter(df['seed'],df['X_Γ'])
+f.savefig('tests/seed_num_vs_X_Gamma.pdf')
+
+plt.close('all')
+f,ax = plt.subplots()
+ax.scatter(df['m_lim'],df['X_Γ'])
+f.savefig('tests/m_lim_vs_X_Gamma.pdf')
+
+plt.close('all')
+f,ax = plt.subplots()
+ax.scatter(df['N_s'],df['X_Γ'])
+f.savefig('tests/N_s_vs_X_Gamma.pdf')
+
+plt.close('all')
+f,ax = plt.subplots()
+ax.scatter(df['N_d'],df['X_Γ'])
+f.savefig('tests/N_d_vs_X_Gamma.pdf')
+
+plt.close('all')
+f,ax = plt.subplots()
+ax.scatter(df['β'],df['X_Γ'])
+f.savefig('tests/beta_vs_X_Gamma.pdf')
+
+plt.close('all')
+import seaborn as sns
+f,ax = plt.subplots()
+sns.distplot(df['X_Γ'], bins=15, kde=True, norm_hist=True, ax=ax)
+pd.set_option('precision',3)
+txt = repr(df['X_Γ'].describe())
+ax.text(
+        0.96, 0.96, txt,
+        horizontalalignment='right',
+        verticalalignment='top',
+        transform=ax.transAxes, fontsize='small'
+        )
+ax.set_ylabel('prob')
+f.savefig('tests/dist_X_Gamma.pdf')
+
+pd.set_option('precision',5)
+print(df.describe())

### 2017-10-04 paper/ms.tex
model 2 mostly done

@@
 ``searchable stars'' are the points on the sky for which they think planets 
-are observed with 100\% detection efficiency. Accounting for the geometric 
+are observed with 100\% detection efficiency. Correcting for the geometric 
 transit probability $f_{\rm g}$, they compute an apparent occurrence rate 
@@
 
-One immediate complication is that~--~due to dynamical stability or some 
-aspect of planet formation~--~ the occurrence rate of planets of the desired 
-type may differ between binary and single-star systems.
+An immediate complication is that~--~due to dynamical stability or some 
+aspect of planet formation~--~ the occurrence rate of planets may differ 
+between binary and single-star systems.
 If ``occurrence rate'' is defined purely as the mean number of planets within 
@@
 
-Errors (1)-(3) in calculating $N_{\rm det}$ above broadly fall under ``radius 
-misclassification'', while error (4) falls under ``completeness 
-miscalculation''.
-Note that in the above list, a planet cannot be both ``of desired type'' and 
-``orbiting a star of undesired type''. We ignore this type of error for 
-analytic convenience, but \texttt{reintroduce it later during numerics}.
+Errors (1)-(3) in calculating $N_{\rm det}$ can be broadly described as 
+``radius misclassification'', while error (4) is an error in the assumed 
+survey completeness.
+%Note that in the above list, a planet cannot be both ``of desired type'' and 
+%``orbiting a star of undesired type''. We ignore this type of error for 
+%analytic convenience, but \texttt{reintroduce it later during numerics}.
 
 Our approach is step-by-step starting from a very simple scenario.
-\begin{itemize}
-\item We consider a universe consisting of only the desired types of stars and
-planets.  We give analytic expressions for eta(apparent)/eta(true) taking
-into account all applicable errors.
-%
-\item We allow for a distribution of the mass ratio $dn/dq \propto q^\gamma$, 
+In Sec.~\ref{sec:model_1} we first consider a universe consisting of only the 
+desired types of stars and planets.  
+We give an analytic expression for $\Gamma_{\rm a}/\Gamma_{\rm t}$, the ratio 
+of the true to apparent occurrence rate, taking into account all applicable 
+errors.
+In Sec.~\ref{sec:model_2} we allow for a distribution of the mass ratio $dn/dq 
+\propto q^\gamma$, 
 and a concurrent power-law distribution in the luminosity ratio, $dn/dL 
 \propto L^\beta$. 
+We give expressions for $\Gamma_{\rm a}/\Gamma_{\rm t}$, and evaluate them 
+numerically for the case of planets orbiting Sun-like dwarf stars.
 %We assume the secondary is not of the desired stellar type.
@@
 %ratio binaries -- the stars are ~identical.
-We give analytic expressions for eta(apparent)/eta(true) taking into 
-account errors of type 2.
-%NOTE: empirical L(M) maybe OK for numerics -- not for analytics.
-%
-\item We allow for a nonzero tolerance $\Delta R_p$ in the planet radius, and 
-tolerance $\Delta L_\star$ in stellar luminosity, to qualify as "desirable".  
-We recalculate eta(apparent)/eta(true).
-(LB: I think this has to become numerical to actually evaluate the occurrence
-rate correction factor, $X_\Gamma$.)
-%
-\item We assume the planets have a power-law radius distribution, $d\eta/dR_p
-\propto r^\alpha$.  We calculate the apparent $d\eta/dR_p$ taking into account 
-errors of type 1-3.  We do so for stars of the desired type within some 
-tolerance dL.
-%
-\item Everything up to now is more-or-less analytic, perhaps supported by 
-numerical checking or numerical integration. At the end we summarize the 
-radius-dependence of the occurrence rate shifting.
-"Accounting only for dilution \& its numerator effect, big radius planets have
+In Sec.~\ref{sec:model_3}, 
+we assume the planets have a power-law radius distribution, ${\rm 
+d}N_p/{\rm d}R_p\propto R_p^\alpha$.  We calculate the 
+effects of binarity on the apparent $\Gamma_{\rm a}(R_p)$ taking into account 
+all relevant errors.
+Accounting only for dilution \& its numerator effect, big radius planets have
 underestimated inferred rates, the smallest radius planets have overestimated
-inferred rates, and the in-between radius planets are in-between"
-
-\end{itemize}
-
+inferred rates, and the in-between radius planets are in-between.
+One consequence is that the claimed HJ occurrence rate discrepancy between 
+radial velocity \& transit surveys gets much weaker when accounting for 
+binarity.
+In Sec.~\ref{sec:discussion} we summarize these models and discuss their 
+utility in interpreting transit surveys.
+
+\begin{comment}
 We then dig in, w/ numerics, to what happens with different assumed
 completeness vs Rp functions.
-We note that the the claimed HJ occ rate discrepency btwn RV \&
-transit surveys gets much weaker when accounting for binarity.
+
 This is b/c the more believable occ rate papers do not heavily rely on their
@@
 measurements.
+\end{comment}
 
@@
 
-Stellar binarity introduces convoluted effects on transit surveys.
-To disentangle them, imagine the following idealized survey.
-
-You observe the entire sky for some duration with a detector of known 
-area and bandpass. Your detector is photon-noise limited.
-You are interested only in detecting planets of radius $R_p$, and orbital 
-period $P$. For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
-You are only interested in detecting them around stars of radius $R_1$, 
-and luminosity $L_1$. For instance, G2V dwarfs.
-You can only detect your desired type of planet when you observe signals 
-with ${\rm S/N} > {\rm (S/N)_{min}}$.
+Consider the following idealized survey.
+An astronomer observes the entire sky for some duration with a detector of 
+known area and bandpass.
+The detector is photon-noise limited.
+They are interested only in detecting planets of radius $R_p$, and orbital 
+period $P$. %For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
+They are only interested in detecting them around stars of radius $R_1$, 
+and luminosity $L_1$. %For instance, G2V dwarfs.
+They detect the desired type of planet when they observe signals 
+above some signal-to-noise floor.
 For a photon-noise limited survey, this minimum signal to noise ratio is 
 equivalent to a minimum flux required for detection, $F_{\rm min}$.
-
-You get funding, and perform the S/N-limited survey. 
-You observe all the ``points'' on-sky with apparent 
-magnitude $m < m_{\rm min}$, or equivalently with energy flux in your bandpass 
-greater than some limit $F > F_{\rm min}$.
-All planets with ${\rm S/N} > {\rm (S/N)_{min}}$ are detected. The rest are 
-not.
-
-You now wish to derive an occurrence rate for planets of radius $R_p$ and 
+Therefore they observe all of the ``points'' on-sky with apparent 
+magnitude $m < m_{\rm min}$, or equivalently with energy flux in their 
+bandpass greater than the limiting flux $F > F_{\rm min}$.
+After detecting all planets with ${\rm S/N} > {\rm (S/N)_{min}}$, they now 
+wish to derive an occurrence rate for planets of radius $R_p$ and 
 orbital period $P$.
-Assume your universe is a universe in which:
+
+Assume this astronomer lives in a universe in which the following assumptions 
+hold.
 \begin{itemize}
@@
 	In this section, $\gamma_R=1$ across the population 
-	of star systems~--~all binaries are assumed to be twin systems.
+	of star systems~--~all binaries are assumed to be twins.
 	The ratio of the number density of binary systems and the number density of 
@@
 The assumption of [Pepper et al, 2003] in a very similar context was that the 
-observer correctly pre-selects all of the ``searchable stars''.
+observer can correctly pre-select all of the ``searchable stars''.
 This simplification would allow the observer to ignore 
@@
 
-The above approach ignores incompleteness for binary systems.
-If we set the limiting flux to give 100\% completeness for single stars, we 
-will unwittingly select binaries near that flux limit, for which dilution 
-will push transit signals below the detection threshold.
-This is because a flux limit maps onto three distinct maximum 
-detection distances for a population of single stars and twin binaries:
-\begin{enumerate}
-\item $d_{{\rm max},s}$: the maximum distance out to which the desired type 
-	of planet is detectable around single stars.
-	These are the selected ``searchable, single stars''.
-%
-\item $d_{{\rm max},d}$: the maximum distance out to which double stars
-  are selected (down to the same minimum flux as that defining $d_{{\rm 
-  max},s}$). 
-  This is different from $d_{{\rm max},d}^{\rm p}$, the maximum
-  distance out to which planets are detectable around double stars. For a 
-  population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
+However, any signal-to-noise limited survey of a population that includes 
+binaries cannot ignore incompleteness.
+%If we set the limiting flux to give 100\% completeness for single stars, we 
+%will unwittingly select binaries near that flux limit, for which dilution 
+%will push transit signals below the detection threshold.
+This is because any given flux limit maps onto three distinct ``maximum 
+distances'' for a population of single stars and twin binaries.
+First, there is $d_{{\rm max},s}$: the maximum distance out to which the 
+desired type of planet is detectable around single stars. Typically this is 
+set to be the distance out to which stars in a SNR-limited survey are 
+selected, since these are the ``searchable stars''.
+Using the same minimum flux as that defining $d_{{\rm max},s}$, double stars 
+are then selected out to a larger limiting distance, $d_{{\rm 
+max},d}$~--~effectively a Malmquist bias.
+
+However, the distance out to which double stars are {\it selected} is 
+different from the distance $d_{{\rm max},d}^{\rm p}$ out to which planets can 
+be {\it detected} (around the primary of a double star system).
+For a population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
 L^{1/2} d^{-1}$,
@@
 
-\end{enumerate}
-
+\begin{comment}
 We note in passing that we are considering a SNR-limited survey, which in this 
@@
 $d_{{\rm max},d}^{\rm p}$.
-This change would likely improve interpretability.
-
-With the stage set, we ask:
+This change might circumvent various biases.
+\end{comment}
+With our model defined, we ask:
 
@@
 Note that these latter authors really quote a {\it multiplicity} fraction of
-0.44 for sun-like dwarfs~--~we count the $\approx 25\%$ of multiple 
-systems that are higher order multiples as doubles.
+0.44 for sun-like dwarfs~--~for simplicity we count the $\approx 25\%$ of 
+multiple systems that are higher order multiples as doubles.
 
@@
 stated in terms of a limiting magnitude $m_{{\rm min}}$.
-In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$. Simply as a 
+In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$. As a 
 consequence of imposing a magnitude cut, the maximum distance to which binary 
@@
 &= \frac{{\rm BF}}{1 - {\rm BF}} \times (1+\gamma_R)^{3/2}.
+\label{eq:beta_model1}
 \end{align}
@@
 this desired type.
-The occurrence rate they then find, {\it i.e.} the number of planets divided 
-by number of ``stars'' (really, stellar systems) is
+The occurrence rate they then find is
 \begin{equation}
@@
 $L \propto M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
-As a simple check, we fit lines to mass-luminosity data\footnote{
+As a simple check, we fit lines to measured masses and luminosities of dwarf 
+stars\footnote{
     Data collected by Torres et al. [2009] for dwarfs earlier than 
@@
 }. Though the data show a break at $M_\star \approx 0.5M_\odot$, the 
-$L=M^{3.5}$ relation is a reasonable fit for $M\lesssim 2M_\odot$.
+$L\propto M^{3.5}$ relation is a reasonable fit for $M\lesssim 2M_\odot$.
 
@@
 Since again, our main purpose is to develop intuition about observational 
-biases, we approximate this as 
+biases, we coarsely assume
 \begin{equation}
@@
 \end{comment}
-
-In passing, we note that while $w_{d2}=1$ was a reasonable assumption in 
-Sec.~\ref{sec:model_1} (since secondaries were identical to primaries), in 
-binary systems we now have secondaries with different stellar properties from 
-primaries.
-Recall that our occurrence rate is defined for planets of a given type 
-orbiting stars of a given type.
-Assume our universe is one in which the primary mass of all stars is 
-$M_1=M_\odot$.
-If this is the case, we might define our ``stars of interest'' to be 
-those with $0.7<M_\star/M_\odot<1.3$.
-Then, in a volume-limited sample, 1 in 3 secondaries would be ``of interest'', 
-and could contribute to the true occurrence rate.
-The general point is that $w_{d2}$ will be less than 1, unless planets of the 
-desired type are more common around secondaries of lower mass (as may actually 
-be true for small planets around M dwarfs).
-Keeping this in mind, we ask the same questions as for Model \#1.
+With our model specified, we ask:
 
@@
 stars hosts.
-
-Counting the number of desired stars, $N_s$ is the same as in 
-Sec.~\ref{sec:model_1}. However
-\begin{equation}
-N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3
-\end{equation}
-for
+Explicitly,
+\begin{equation}
+f_{\rm desired} = \frac{\int_{M_{\rm min}}^{M_{\rm max}} {\rm prob}_{\rm 
+        ml}(q)\,{\rm d}q }
+{\int_{0.1M_1}^{M_1} {\rm prob}_{\rm 
+        ml}(q)\,{\rm d}q}.
+\end{equation}
+For instance, assuming a primary mass of $M_1 = M_\odot$, and defining the 
+``desired type'' of stars being to be $0.7<M_\star/M_\odot<1.3$, we find 
+$f_{\rm desired}\approx0.475$.
+
+Counting the number of desired stars that are single, $N_s$, is the same as in 
+Sec.~\ref{sec:model_1}. However now the maximum selected distance for a binary 
+system is now a function of the system's light ratio,
 \begin{equation}
@@
 where $d_{\rm max,s}$ is given by Eq.~\ref{eq:d_max}.
+This affects our ability to count the number of desired stars that are double,
+since
+\begin{equation}
+N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3.
+\end{equation}
 
@@
 ensemble of all possible transit surveys.
-Applying the chain rule for probability density functions, the distribution 
+Transforming the appropriate probability density functions, the distribution 
 function for the number of binary systems,
-${\rm prob}(N_d)$, can be written 
+${\rm prob}(N_d)$, can be expressed as 
 \begin{equation}
@@
     \begin{center}
-		\includegraphics[scale=.8]{figures/joint_prob_r_q.pdf}
+        \includegraphics[width=.5\textwidth]{figures/joint_prob_r_q.pdf}
     \end{center}
-	\caption{The joint probability distribution of a binary star's position and 
+    \caption{The joint probability distribution of a binary star's position 
+        and 
         mass ratio in a magnitude-limited sample. This plot assumes $L\propto 
@@
         (Eq.~\ref{eq:pdf_q_mag_limited}) to good precision. 
-        The numerical comparison uses the empirical mass-luminosity relation 
-        shown in Fig.~\ref{fig:mass_luminosity}.
+        The numerical comparison uses the ``realistic'' empirical 
+        mass-luminosity relation shown in Fig.~\ref{fig:mass_luminosity}.
     }
@@
 Before blindly plugging in Eq.~\ref{eq:mass_ratio}'s assumed mass ratio 
-distribution to find the expected number of double systems, we must
-emphasize that Eq.~\ref{eq:mass_ratio}'s distribution applies only for 
+distribution to find the expected number of double systems, it is critical to 
+note that Eq.~\ref{eq:mass_ratio}'s distribution applies only for 
 volume-limited samples.
@@
 distribution for a binary star's position $r$ and mass ratio $q$.
-Explicitly, using `ml' as a subscript for `magnitude-limited',
+Using `ml' as a subscript for `magnitude-limited',
 \begin{equation}
@@
 
-The reason we derived all this is that we wanted to know the number of 
-searchable binary stars.
+Our original reason for deriving these expression was that we wanted to know 
+the number of searchable binary stars.
 For any given realization of the survey, this number is a random draw.
-Computing the \textit{expected value} for the number of double star systems in 
-the sample,
+We can compute the \textit{expected value} for the number of double star 
+systems in the sample,
 \begin{equation}
@@
 \end{equation}
-Letting $\alpha=3.5$, this can be expressed as
+Letting $\alpha=3.5$, this becomes
 \begin{equation}
@@
 \end{equation}
-where non-dimensional constants are not shown,
-$u = \mathcal{R} N_d^{2/3}$,
+where $u = \mathcal{R} N_d^{2/3}$,
+non-dimensional constants are not shown,
 and $\mathcal{R}$ is defined as
@@
 \frac{\langle N_d \rangle}{N_s} &= 1.590\, \frac{n_d}{n_s}
+\label{eq:beta_model2}
 \end{align}
@@
 $\langle N_d \rangle / N_s = 1.25$.
-Compared to the binary-twin thought experiment of Sec.~\ref{sec:model_1}, in 
-which the SNR-limited sample had $\approx 2.8$ times more binary systems than 
-single systems, there are now $\approx 1.25$ times as many.
+Compared to Eq.~\ref{eq:beta_model1} from the binary-twin thought experiment, 
+while the SNR-limited sample previously had $\approx 2.8$ times more binary 
+systems than single systems, there are now $\approx 1.25$ times as many.
 This is sensible, because the mean light ratio has decreased; it was 
@@
 \end{equation}
-Explicitly,
-\begin{equation}
-f_{\rm desired} = \frac{\int_{M_{\rm min}}^{M_{\rm max}} {\rm prob}_{\rm 
-ml}(q)\,{\rm d}q }
-{\int_{0.1M_\odot}^{M_\odot} {\rm prob}_{\rm 
-        ml}(q)\,{\rm d}q}.
-\end{equation}
-For instance, if we define the ``desired type'' to be 
-$0.7<M_\star/M_\odot<1.3$, we find $f_{\rm desired}\approx0.475$.
 
@@
 occurrence rate of planets with radius $R_p$ and period $P$, orbiting stars 
-with mass $0.7<M_\star/M_\odot < 1.3$ (with corresponding stellar radii and 
-luminosities).
+with mass $M_1$.
 They detected $N_{\rm det}$ transit signals that appear to be from planets of 
@@
 
-The drawn number of planet detections of the desired type is 
+The drawn total number of planet detections is 
 \begin{equation}
@@
 \end{equation}
-As in Sec.~\ref{sec:model_1},
+Formally, if we are only interested in planets of radius $R_p$ then no planets 
+around binaries are ``of the desired type'' because of dilution.
+For the following development, ignore this caveat~--~consider an 
+observer who is interested in planets of any apparent radius.
+
+Then as in Sec.~\ref{sec:model_1},
 \begin{equation}
@@
 
-For planets detected in binary systems, only a fraction will be ``of the 
-desired type'', since only a fraction of the stars are of the desired type.
-This is the same factor $f_{\rm desired}$ as previously discussed. Accounting 
-for this and the independent geometric and completeness probabilities,
+For planets detected in binary systems,
 \begin{equation}
@@
 \end{equation}
-
-For any given system,
+The factor $f_{\rm desired}$, as previously discussed, accounts for the 
+fraction of secondaries not of the desired {\it stellar} type. The above 
+counts the total number of detected planets around stars of the desired type, 
+though all these planets have diluted radii.
+The detection and geometric transit probabilities for a given system are
 \begin{align}
@@
 f_{d2,c} &= (1+\gamma_R^{-1})^{-3} \\
-f_{d2,g} &= f_{s,g}q^{2/3}
+f_{d2,g} &= f_{s,g}q^{2/3}.
 \end{align}
@@
 For the stellar population described in this model, with a mean light ratio 
-$\langle \gamma_R \rangle \approx 0.27$, this limit is not strictly applicable.
+$\langle \gamma_R \rangle \approx 0.27$, this limit is not applicable.
 To assess the importance of the error, we want the mean
@@
 
-Running a Monte Carlo simulation in which we fix the planet and primary star 
-properties, draw the binary mass ratio from Eq.~\ref{eq:pdf_q_mag_limited}, 
-and take 
-the luminosity as $L\propto M^{3.5}$, we find 
+Instead, we run a Monte Carlo simulation in which we fix the planet and 
+primary star properties, draw the binary mass ratio from 
+Eq.~\ref{eq:pdf_q_mag_limited}, 
+and take the luminosity as $L\propto M^{3.5}$. We find 
 a mean of 
 \begin{equation}
-\langle X_\Gamma \rangle =1.291\quad ({\rm Monte\ Carlo\ result}),
-\end{equation}
-and a median that is essentially the same.
+\langle X_\Gamma \rangle =1.289\quad ({\rm lower\ limit\ from\ Monte\ Carlo}),
+\label{eq:X_Gamma_model2_lower}
+\end{equation}
+and a median that is the same to within one percent.
 This value is comparable to what one would guess using the analytic results 
@@
 compute $f_{d1,c}$.
+
+%WARNING
+%WARNING
+%the reported value of <N_d>/N_s is higher from analytics than vs what 
+%simulations give. means a mistake in one of them.
+
+\paragraph{Varying what is meant by ``planets of the desired type'' }
+The above occurrence rate correction factor is a lower limit for this model, 
+because if we were to be any stricter about the range of allowed apparent 
+radii, the apparent occurrence rate would only decrease.
+
+Taking the opposite limit of not counting any detections that deviate 
+from $R_p$ as ``planets of the desired type'',
+we find
+\begin{equation}
+X_{\Gamma} = \frac{1 + \beta }{1 + \beta f_{d1,c}}.
+\end{equation}
+From the same Monte Carlo simulation, this gives an upper limit on the 
+correction term
+\begin{equation}
+\langle X_\Gamma \rangle = 2.114\quad ({\rm upper\ limit\ from\ Monte\ 
+Carlo}).
+\label{eq:X_Gamma_model2_upper}
+\end{equation}
+
+The general point is that starting from a single planet radius, dilution 
+produces a spectrum of apparent planetary radii. 
+This skews occurrence rate measurements, by an amount between 
+Eqs.~\ref{eq:X_Gamma_model2_lower} and~\ref{eq:X_Gamma_model2_upper},
+depending on the radius bounds set by the analyst.
+We investigate this further in our final model below.
+
+\section{Model \#3: allow a distribution of planetary radii}
+\label{sec:model_3}
+
+Keep the same stars as in Sec.~\ref{sec:model_2}.
+Now let the planets have a power-law radius distribution, ${\rm d}N_p/{\rm 
+d}R_p\propto R_p^\delta$. 
+We want to know the effects of binarity on the apparent distribution of 
+planetary radii.
+Specifically, the average number of planets in a given size range per star of 
+the desired type in this universe is
+\begin{equation}
+\frac{N_{\rm planets}(R_p^{\rm min}<R_p <R_p^{\rm max})}{N_{\rm stars}} = 
+\Gamma_0 \int_{R_p^{\rm min}}^{R_p^{\rm max}} R_p^\delta \,{\rm d}R_p,
+\end{equation}
+for $\Gamma_0$ one of $\{\Gamma_{t,s}, \Gamma_{t,d}, w_{d2}\Gamma_{t,d}\}$
+corresponding to single star systems, the primaries of double star systems, 
+and the secondaries of double star systems.
+
+\section{Discussion}
+\label{sec:discussion}
+
+\appendix
+\begin{figure}[!t]
+    \begin{center}
+        \includegraphics[scale=0.9]{figures/mass_luminosity.pdf}
+    \end{center}
+    \caption{Empirical fit to main sequence dwarf mass luminosity data 
+    compiled 
+        from Torres et al. [2009] and Benedict et al. [2016]. The ``low fit'' 
+        is a 
+        least squares fit to data from $0.1-0.5M_\odot$, and the ``high fit'' 
+        is 
+        to 
+        data above that, and below the Kraft break ($1.2M_\odot$).
+        The $L(M)$ relation taken for numerics is the maximum 
+        of the two fits. For analytics, we assume $L\propto M^{3.5}$.
+    }
+    \label{fig:mass_luminosity}
+\end{figure}
 
@@
 \section{Model \# 3: A Synthetic {\it Kepler} Analog}
-\label{sec:model_3}
+\label{sec:kepler_analog}
 

### 2017-10-04 src/model_2.py

@@ -108,7 +108,7 @@ if __name__ == '__main__':
     #############################
     # !!! BEGIN MONTE CARLO !!! #
     #############################
-    for seed in np.arange(int(2e3),int(3e3),1):
+    for index, seed in enumerate(np.arange(int(8e3),int(9e3),1)):
 
         ################################################################
         # BEGIN OBNOXIOUS PARAMETERS NEEDED FOR MONTE CARLO SIMULATION #
@@ -197,6 +197,30 @@ if __name__ == '__main__':
 
         # impose magnitude cut.
         doubles = vl[vl['F'] > F_lim.cgs.value]
+
+        if index==0:
+            import matplotlib.pyplot as plt
+            import seaborn as sns
+            from scipy.integrate import trapz
+
+            f,ax=plt.subplots()
+            sns.distplot(doubles['q'], ax=ax, kde=False, norm_hist=True)
+            q = np.arange(0.1,1+1e-3,1e-3)
+            prob_ml_q = (1+q**(3.5))**(3/2)
+            prob_ml_q /= trapz(prob_ml_q, q)
+            ax.plot(q, prob_ml_q)
+            f.savefig('tests/doubles_q_distribution.pdf')
+
+            plt.close('all')
+            f,ax=plt.subplots()
+            sns.distplot(doubles['γ_R'], ax=ax, kde=False, norm_hist=True)
+            γ_R = np.arange(0.1**3.5,1+1e-3,1e-3)
+            prob_ml_γR = (1+γ_R)**(3/2) * γ_R**(-5/7)
+            prob_ml_γR /= trapz(prob_ml_γR, γ_R)
+            ax.plot(γ_R, prob_ml_γR)
+            ax.set_yscale('log')
+
+            f.savefig('tests/doubles_gammaR_distribution.pdf')
 
         N_d = int(len(doubles))
 
@@ -302,7 +326,12 @@ if __name__ == '__main__':
         Γ_t = N_planets/N_stars
         Γ_a = (N_det_s + N_det_d)/(N_s + N_d) * (1/f_sg[0])
 
+        # the above is in the limit of counting all planets in binaries in the
+        # occ rate. the below does not count any of them.
+        Γ_a_upperlimit = N_det_s/(N_s+N_d)*(1/f_sg[0])
+
         X_Γ = Γ_t/Γ_a
+        X_Γ_upperlimit = Γ_t/Γ_a_upperlimit
 
         ######################
[…]

### 2017-10-04 src/model_2_analysis.py

@@ -42,5 +42,22 @@ ax.text(
 ax.set_ylabel('prob')
 f.savefig('tests/dist_X_Gamma.pdf')
 
+
+plt.close('all')
+import seaborn as sns
+f,ax = plt.subplots()
+sns.distplot(df['X_Γ_upperlimit'], bins=15, kde=True, norm_hist=True, ax=ax)
+pd.set_option('precision',3)
+txt = repr(df['X_Γ_upperlimit'].describe())
+ax.text(
+        0.96, 0.96, txt,
+        horizontalalignment='right',
+        verticalalignment='top',
+        transform=ax.transAxes, fontsize='small'
+        )
+ax.set_ylabel('prob')
+f.savefig('tests/dist_X_Gamma_upperlimit.pdf')
+
+
 pd.set_option('precision',5)
 print(df.describe())

### 2017-10-05 paper/ms.tex
wordsmith
corrected f_d2c bug

@@
 be {\it detected} (around the primary of a double star system).
-For a population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
+For the binary twin population, since ${\rm S/N} \propto \mathcal{D} 
 L^{1/2} d^{-1}$,
@@
 volume in which planets can be detected to the actual selected volume).
-For a more realistic sample, $\gamma_R \approx 0.1$, and there is an 
-interesting asymmetry.
-One finds the completeness for planets which orbit the primary star, 
-$f_{d1,c}$ to be
-\begin{equation}
-f_{d1,c} = (1+\gamma_R)^{-3} \approx 3/4.
-\end{equation}
-However, for planets which orbit the secondary star, the completeness is
-\begin{equation}
-f_{d2,c} = (1+\gamma_R^{-1})^{-3} \approx 3/4000.
-\end{equation}
 
@@
 f_{d1,g} &= f_{s,g} \\
-f_{d2,c} &= (1+\gamma_R^{-1})^{-3} \\
-f_{d2,g} &= f_{s,g}q^{2/3}.
+f_{d2,c} &= (1+\gamma_R^{-1})^{-3} \gamma_R^{-6/\alpha} 
+\label{eq:f_d2c}\\
+f_{d2,g} &= f_{s,g}q^{2/3},
 \end{align}
+where typically $\alpha=3.5$ (and the second term in Eq.~\ref{eq:f_d2c} comes 
+from the stellar radius correction).
+For a typical binary drawn from the local stellar neighborhood, $\gamma_R 
+\approx 0.2$ (see Fig.~\ref{fig:gammaR_distribn_mag_limited}), and there is an 
+interesting asymmetry.
+One finds the completeness for planets which orbit the primary star, 
+$f_{d1,c}$ to be
+\begin{equation}
+f_{d1,c} = (1+\gamma_R)^{-3} \sim 70\%
+\end{equation}
+However, for planets which orbit the secondary star, the completeness is
+\begin{equation}
+f_{d2,c} = (1+\gamma_R^{-1})^{-3} \gamma_R^{-6/\alpha} \sim 7\%.
+\end{equation}
+Controlling for all other variables, in the typical binary system planets 
+around secondaries are more difficult to detect than those around primaries.
 
@@
 Keep the same stars as in Sec.~\ref{sec:model_2}.
-Now let the planets have a power-law radius distribution, ${\rm d}N_p/{\rm 
+Now let the planets have a power-law radius distribution, ${\rm 
+d}(N_p/N_\star)/{\rm 
 d}R_p\propto R_p^\delta$. 
@@
 \begin{equation}
-\frac{N_{\rm planets}(R_p^{\rm min}<R_p <R_p^{\rm max})}{N_{\rm stars}} = 
-\Gamma_0 \int_{R_p^{\rm min}}^{R_p^{\rm max}} R_p^\delta \,{\rm d}R_p,
+\frac{N_{\rm p}(R_p^{\rm \ell}<R_p <R_p^{\rm u})}{N_{\rm \star}} = 
+\Gamma_0 \int_{R_p^{\rm \ell}}^{R_p^{\rm u}} R_p^\delta \,{\rm d}R_p,
 \end{equation}

### 2017-10-05 src/model_2.py

@@ -108,7 +108,7 @@ if __name__ == '__main__':
     #############################
     # !!! BEGIN MONTE CARLO !!! #
     #############################
-    for index, seed in enumerate(np.arange(int(8e3),int(9e3),1)):
+    for index, seed in enumerate(np.arange(int(1e3),int(2e3),1)):
 
         ################################################################
         # BEGIN OBNOXIOUS PARAMETERS NEEDED FOR MONTE CARLO SIMULATION #
@@ -206,7 +206,8 @@ if __name__ == '__main__':
             f,ax=plt.subplots()
             sns.distplot(doubles['q'], ax=ax, kde=False, norm_hist=True)
             q = np.arange(0.1,1+1e-3,1e-3)
-            prob_ml_q = (1+q**(3.5))**(3/2)
+            alpha = 3.5
+            prob_ml_q = (1+q**alpha)**(3/2)
             prob_ml_q /= trapz(prob_ml_q, q)
             ax.plot(q, prob_ml_q)
             f.savefig('tests/doubles_q_distribution.pdf')
@@ -230,9 +231,9 @@ if __name__ == '__main__':
         # Fraction of stars in single star systems with planet of (R_p, P). NB this
         # is the same as the average number of planets per star in the single
         # planet system limit. (...)
-        Γ_ts = 0.2
+        Γ_ts = 0.3
         # Fraction per primary of double star systems with planet of (R_p, P).
-        Γ_td = 0.2
+        Γ_td = 0.3
         # Weight of secondaries of desired type with planets of desired type
         w_d2 = 1
 
@@ -265,7 +266,7 @@ if __name__ == '__main__':
         #################################
         # COMPUTE TRANSIT PROBABILITIES #
         #################################
-        P = 0.1*u.year
+        P = 30*u.day
         a_1 = (P**2 * c.G * M_1 / (4*π*π))**(1/3)
         ones = np.ones_like(np.array(singles['r']))
         f_sg = (R_1/a_1).cgs.value * ones
@@ -299,7 +300,9 @@ if __name__ == '__main__':
                 & s_has_transiting_planet
 
         f_d1c = (1+np.array(doubles['γ_R']))**(-3)
-        f_d2c = (1+np.array(doubles['γ_R'])**(-1))**(-3)
+        alpha = 3.5
+        f_d2c = (1+np.array(doubles['γ_R'])**(-1))**(-3) \
+                *np.array(doubles['γ_R'])**(-6/alpha)
 
[…]

### 2017-10-05 src/vary_Rp_study_pdfs.py

@@ -318,7 +318,7 @@ def _make_Rpobs_plot(R_pt, prob_Rpts, R_pos, prob_Rpos_d1, prob_Rpos_d2,
         ax.tick_params(which='both', direction='in', zorder=0)
 
     f.tight_layout(h_pad=0)
-    f.savefig('prob_Rpobs_{savstr}.pdf'.format(savstr=sstr),
+    f.savefig('prob_Rpobs/prob_Rpobs_{savstr}.pdf'.format(savstr=sstr),
                dpi=250, bbox_inches='tight')
 
 

### 2017-10-05 src/vary_Rp_study_pdfs.py
apparent planet radius distribution improvements
* correct bug in secondary radius for the distribution function when
planet orbits the secondary

* add case of narrow uniform plot. (todo is reality check it)

@@ -1,38 +1,27 @@
 '''
-A straight numerical approach would be:
-
-Take all Galaxia stars within 7.5 degrees of the center of Kepler's
-field, and only take stars with r<17.
-
-This is beyond where the KIC is complete, so comparisons between the two
-will be confused.  Then apply analogs of the Batalha+ 2010 Table 1
-prioritization.
-
-This gives the "GIC", and it's Teff distribution is OK. Its binary
-radius distribution is currently a bit wonky, which is not nice.
-
-If you give the GIC a planet population from various obvious options
-(uniform in Rp, uniform in logRp, normal in logRp) you can study what the
-binarity does to the inferred radius distribution.
-
-ALTERNATIVELY
+How does binarity affect the apparent radius distribution? (vs an input true
+radius distribution)
 
 you can do it all semianalytically, following notes of 17/09/03.2-3
+
+and, more ACCURATELY (AKA with math bug fixes) the notes of 17/10/04.3-4
 '''
 
 import numpy as np
 from scipy.integrate import trapz
 import matplotlib.pyplot as plt
 
+global α
+α = 3.5
+
 def prob_gammaR(γ_R):
 
-    Z = 1 # constants seem to work
     prob_γ_R = np.zeros_like(γ_R)
 
-    inds = (γ_R<1) & (γ_R>0.1**3)
-    prob_γ_R[inds] = Z * (1 + γ_R[inds])**(3/2) * (1/3) * γ_R[inds]**(-2/3)
+    inds = (γ_R<1) & (γ_R>0.1**α)
+    prob_γ_R[inds] = (1 + γ_R[inds])**(3/2) * γ_R[inds]**((1-α)/α)
 
-    return prob_γ_R
+    return prob_γ_R/trapz(prob_γ_R, γ_R)
 
 
 def prob_Rpt_uniform(R_pt, R_pu, R_pl):
@@ -138,7 +127,7 @@ def logistic_fn(x, L, k, x_0):
     val = L / (1 + np.exp(-k * (x-x_0)))
     return val
 
[…]

### 2017-10-05 paper/ms.tex
update MC numbers given bugfix

@@
 \begin{equation}
-f_{d1,c} = (1+\gamma_R)^{-3} \sim 70\%
+f_{d1,c} \sim 70\% \quad(\alpha=3.5, \gamma_R=0.2).
 \end{equation}
@@
 \begin{equation}
-f_{d2,c} = (1+\gamma_R^{-1})^{-3} \gamma_R^{-6/\alpha} \sim 7\%.
-\end{equation}
-Controlling for all other variables, in the typical binary system planets 
-around secondaries are more difficult to detect than those around primaries.
+f_{d2,c} \sim 7\% \quad(\alpha=3.5, \gamma_R=0.2).
+\end{equation}
+This is the obvious statement that controlling for all other variables, 
+in the typical binary system planets around secondaries are more difficult to 
+detect than those around primaries.
 
@@
 Eq.~\ref{eq:pdf_q_mag_limited}, 
-and take the luminosity as $L\propto M^{3.5}$. We find 
-a mean of 
-\begin{equation}
-\langle X_\Gamma \rangle =1.289\quad ({\rm lower\ limit\ from\ Monte\ Carlo}),
+and take the luminosity as $L\propto M^{3.5}$. We find a mean occurrence 
+correction of 
+\begin{equation}
+\langle X_\Gamma \rangle =1.268\quad ({\rm lower\ limit\ from\ Monte\ Carlo}),
 \label{eq:X_Gamma_model2_lower}

### 2017-10-05 src/model_2_analysis.py

@@ -1,3 +1,8 @@
+'''
+run after running model_2.py
+
+(processes those results)
+'''
 import pandas as pd
 import matplotlib.pyplot as plt
 

### 2017-10-05 paper/ms.tex
bugfix T_dur factor in f_d2c.
hopefully last such bugfix. I *think* I have all the terms included...

@@
 f_{d1,g} &= f_{s,g} \\
-f_{d2,c} &= (1+\gamma_R^{-1})^{-3} \gamma_R^{-6/\alpha} 
+f_{d2,c} &= (1+\gamma_R^{-1})^{-3} \gamma_R^{-5/\alpha} 
 \label{eq:f_d2c}\\
@@
 \end{align}
-where typically $\alpha=3.5$ (and the second term in Eq.~\ref{eq:f_d2c} comes 
-from the stellar radius correction).
+where typically $\alpha=3.5$.
+To derive Eq.~\ref{eq:f_d2c}, one must note that for fixed period, planet 
+radius, and signal to noise, 
+$$d_{\rm det}\propto \mathcal{D} R_2^{-2} L_d^{1/2} T_{\rm dur}^{1/2}.$$
 For a typical binary drawn from the local stellar neighborhood, $\gamma_R 

### 2017-10-05 src/model_2.py

@@ -113,18 +113,20 @@ if __name__ == '__main__':
         ################################################################
         # BEGIN OBNOXIOUS PARAMETERS NEEDED FOR MONTE CARLO SIMULATION #
         ################################################################
-        ##############
-        # INSTRUMENT #
-        ##############
-        A = 50*u.cm**2      # effective observing area
-        λ_min = 500*u.nm    #
-        λ_max = 1000*u.nm
 
-        #################
-        # SURVEY PARAMS #
-        #################
-        T_obs = 4*u.year
-        x_min = 25 # minimum SNR for detection.
+        # Those comments below are not needed (now). They may be in the future.
+        ###############
+        ## INSTRUMENT #
+        ###############
+        #A = 50*u.cm**2      # effective observing area
+        #λ_min = 500*u.nm    #
+        #λ_max = 1000*u.nm
+
+        ##################
+        ## SURVEY PARAMS #
+        ##################
+        #T_obs = 4*u.year
+        #x_min = 25 # minimum SNR for detection.
 
         # Zombeck 2007, p.103: Vega, V=0.03 has a wavelength-specific photon flux of
         # 1e3 ph/s/cm^2/angstrom. So for our instrument's bandpass, we
@@ -302,7 +304,7 @@ if __name__ == '__main__':
         f_d1c = (1+np.array(doubles['γ_R']))**(-3)
         alpha = 3.5
         f_d2c = (1+np.array(doubles['γ_R'])**(-1))**(-3) \
-                *np.array(doubles['γ_R'])**(-6/alpha)
+                *np.array(doubles['γ_R'])**(-5/alpha)
 
         doubles['primary_has_detected_planet'] = \
                 (np.random.rand(N_d) < f_d1c) \

### 2017-10-05 paper/ms.tex
wordsmithing

@@
 In other words, they wish to find the mean number of planets within specific 
-planetary $(R_p, P)$ bounds orbiting the stars in a given volume 
-of stellar $(M_\star, R_\star, L_\star)$ phase-space.
+planetary bounds orbiting the stars in a given volume 
+of stellar phase-space $(\Delta R_p, \Delta P, \Delta M_\star,\ldots$).
 They perform a signal-to-noise limited transit survey and detect $N_{\rm det}$ 
 transit signals that appear to be from planets of the desired type.  They 
-calculate the number of stars $N_\star$ that appear to be searchable for the 
-desired type of planet, and also of the desired stellar type.  These 
-``searchable stars'' are the points on the sky for which they think planets 
-are observed with 100\% detection efficiency. Correcting for the geometric 
+calculate the number of stars $N_\star$ that seem to be of the desired type, 
+and searchable for the planet.  These ``searchable stars'' are the points on 
+the sky for which they think planets are observed with 100\% detection 
+efficiency. Correcting for the geometric 
 transit probability $f_{\rm g}$, they compute an apparent occurrence rate 
@@
 type often exist in binaries. 
-We assume for simplicity that all binaries in the transit survey are 
+We assume for simplicity that all binaries in the survey are 
 spatially unresolved.
@@
 If ``occurrence rate'' is defined purely as the mean number of planets within 
-set radius and period bounds per star in a given interval $\Delta M_\star, 
-\Delta R_\star, \Delta L_\star$, it must include an implicit marginalization 
-over stellar multiplicity.
-Otherwise, one must discuss ``occurrence rates in single star systems'', 
+set radius and period bounds per star in a given interval of say $\Delta 
+M_\star$, it must implicitly marginalize over stellar multiplicity.
+Otherwise, one must address ``occurrence rates in single star systems'', 
 ``occurrence rates about primaries of double systems'', and
@@
 
-Outside of bonafide astrophysical differences, there are observational biases.
+Outside of astrophysical differences, there are observational biases.
 A given apparently-searchable star may truly be a single star of the desired 
@@
 
-The above enumerates possible errors when counting $N_\star$.
-When tallying $N_{\rm det}$, the detected signals from planets of the desired 
+When counting $N_{\rm det}$, the detected signals from planets of the desired 
 type, neglecting binarity introduces the following error cases:
@@
 A detected signal is from a star of desired type (the primary or secondary of 
-a binary systems), from a planet of undesired type, and is incorrectly counted 
+a binary), from a planet of undesired type, and is incorrectly counted 
 as a planet of desired type.
@@
 errors.
-In Sec.~\ref{sec:model_2} we allow for a distribution of the mass ratio $dn/dq 
-\propto q^\gamma$, 
+In Sec.~\ref{sec:model_2} {\bf we allow for a distribution of the mass ratio 
+$dn/dq \propto q^\gamma$} {\bf (not done yet)}, 
 and a concurrent power-law distribution in the luminosity ratio, $dn/dL 
@@
 binarity.
-In Sec.~\ref{sec:discussion} we summarize these models and discuss their 
-utility in interpreting transit surveys.
+In Sec.~\ref{sec:discussion} we summarize the models and discuss how they can 
+be used to interpret transit surveys.
 
@@
 	of star systems~--~all binaries are assumed to be twins.
-	The ratio of the number density of binary systems and the number density of 
-	all systems in a volume-limited sample is the binary fraction\footnote{The 
+	The ratio of the number density of binary systems to all systems in a 
+	volume-limited sample is the binary fraction\footnote{The 
 	binary fraction is equivalent to the multiplicity fraction if there are 
 	no triple, quadruple, $\ldots$ systems.}.
-	\item The true population of planets around these stars is as follows:
-	\begin{itemize}
-	\item A fraction $\Gamma_{t,s}$ of stars in single star systems 
+	\item The true population of planets around these stars is as follows.
+    In single star systems, a fraction $\Gamma_{t,s}$ of stars 
 	have planets of radius $R_p$, with orbital period $P$.
-	\item A fraction $\Gamma_{t,d}$ of primary stars in double star 
-	systems have planets of radius $R_p$, with orbital period $P$. 
-	A fraction $w_{d2}\Gamma_{t,d}$ of secondary stars in a double star 
+	In double star systems, a fraction $\Gamma_{t,d}$ of primary stars have 
+	planets of radius $R_p$, with orbital period $P$. 
+	A fraction $w_{d2}\Gamma_{t,d}$ of secondary stars in double star 
 	systems have the same type of 
@@
 	binaries is captured by $\Gamma_{t,s}, \Gamma_{t,d},$ and $w_{d2}$.
-	If one assumes the primary and secondary of a binary system host planets at 
-	equal rates, this is the $w_{d2}=1$ limit.
+	If one assumes the primary and secondary of a binary system host planets 
+	at equal rates, this is the $w_{d2}=1$ limit.
 	Similarly, if one assumes  secondaries 
 	cannot host planets, this is the $w_{d2}=0$ limit.
-	\end{itemize}
 \end{itemize}
@@
 be {\it detected} (around the primary of a double star system).
-For the binary twin population, since ${\rm S/N} \propto \mathcal{D} 
+For the binary twin population, all planet and stellar radii are identical, so 
+${\rm S/N} \propto \mathcal{D}, and
 L^{1/2} d^{-1}$,
@@
 ${\rm BF}=0.44\pm0.02$ [Duchene \& Kraus, 2013, Raghavan et al 2010]. 
-Note that these latter authors really quote a {\it multiplicity} fraction of
+These latter authors really quote a {\it multiplicity} fraction of
 0.44 for sun-like dwarfs~--~for simplicity we count the $\approx 25\%$ of 
@@
 stars will be selected is greater than that of single stars.
-The ratio of double to single systems in the observed SNR-limited sample is 
+The ratio of double to single systems in the observed sample is 
 then
@@
 \Bigg\{\begin{array}{lr}
-c_q & 0.1 < q \leq 1  \\
+{\rm constant} & 0.1 < q \leq 1  \\
 0 & {\rm otherwise},
@@
 \end{equation}
-provided we ignore the peak at high mass ratios for analytic convenience. $c_q 
-= 1/9$ is the normalization.
+though we ignore the peak at high mass ratios for analytic convenience. 
 We choose this latter mass ratio distribution.
 
-For the mass-luminosity relation, for analytic convenience we assume 
+For the mass-luminosity relation, %for analytic convenience 
+we assume 
 $L \propto M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
-As a simple check, we fit lines to measured masses and luminosities of dwarf 
+To check, we fit lines to measured masses and luminosities of dwarf 
 stars\footnote{
@@
 
-The final addition to our model is that we must assume a stellar mass-radius 
-relation.
+The final necessary assumption is a stellar mass-radius relation.
 This affects the transit probabilities and signal sizes about the 
 secondaries of binary star systems.
-Since again, our main purpose is to develop intuition about observational 
+Since our main purpose is to develop intuition about observational 
 biases, we coarsely assume

### 2017-10-05 paper/ms.tex
update model 3

@@
 \begin{align}
-f_{d1,c} &= (1+\gamma_R)^{-3} \\
+f_{d1,c} &= (1+\gamma_R)^{-3} 
+\label{eq:f_d1c}
+\\
 f_{d1,g} &= f_{s,g} \\
@@
 f_{d2,g} &= f_{s,g}q^{2/3},
+\label{eq:f_d2g}
 \end{align}
@@
 \begin{equation}
-\langle X_\Gamma \rangle =1.268\quad ({\rm lower\ limit\ from\ Monte\ Carlo}),
+\langle X_\Gamma \rangle =1.273\quad ({\rm lower\ limit\ from\ Monte\ Carlo}),
 \label{eq:X_Gamma_model2_lower}
@@
 
-Keep the same stars as in Sec.~\ref{sec:model_2}.
-Now let the planets have a power-law radius distribution, ${\rm 
+For this model, we keep the same stars as in Sec.~\ref{sec:model_2}, but the 
+planets have a power-law radius distribution, ${\rm 
 d}(N_p/N_\star)/{\rm 
@@
 the desired type in this universe is
-\begin{equation}
-\frac{N_{\rm p}(R_p^{\rm \ell}<R_p <R_p^{\rm u})}{N_{\rm \star}} = 
+\begin{align}
+\frac{N_{\rm p}(R_p^{\rm \ell}<R_p <R_p^{\rm u})}{N_{\rm \star}} 
+&= 
+\int_{R_p^{\rm \ell}}^{R_p^{\rm u}} \Gamma(R_p) \,{\rm d}R_p
+\\
+&=
 \Gamma_0 \int_{R_p^{\rm \ell}}^{R_p^{\rm u}} R_p^\delta \,{\rm d}R_p,
-\end{equation}
+\end{align}
 for $\Gamma_0$ one of $\{\Gamma_{t,s}, \Gamma_{t,d}, w_{d2}\Gamma_{t,d}\}$
@@
 and the secondaries of double star systems.
+More strictly, $\Gamma$ should now be referred to as an occurrence rate 
+density (it always has been, but in previous models the planet-radius 
+dependence was a delta function).
+
+The observer now calculates occurrence rates slightly differently.
+The original survey was of stars with $m<m_{\rm min}$, and the
+planet detections are limited by the achievable signal-to-noise from this 
+sample.
+For a given planet radius, the observer must now count $N_\star (R_p)$ stars 
+that they think were searchable.
+In reality, some of these stars will be binaries:
+\begin{equation}
+N_\star(R_p) = N_s(R_p) + N_d(R_p),
+\end{equation}
+where the number of selected stars in single and double star systems is 
+analogous to those of Sec.~\ref{sec:model_1}, except they scale as the 
+sixth power of planet radius.
+The completeness should now be thought of as
+\begin{equation}
+{\rm completeness} = \frac{{\rm actual\ number\ of\ searchable\ stars}}{{\rm 
+assumed\ number\ of\ searchable\ stars}}.
+\end{equation}
+If the completeness is calculated in this way, 
+Equations~\ref{eq:f_d1c} through~\ref{eq:f_d2g} remain the same.
 
@@
 
+\newpage
 \appendix

### 2017-10-05 src/model_3.py

@@ -1,11 +1,10 @@
 '''
-what is the light ratio distriubtion for binaries in a magnitude limited
-sample, vs a volume limited sample?
-
-fixed Rp. fixed primary L1. varying light ratio (L2/L1).
+fixed primary L1. varying light ratio (L2/L1). varying Rp.
 
 This code is the Monte Carlo simulation needed to produce summary statistics
 for X_Γ.
+
+(To the extent possible, it is the same as model_2.py)
 '''
 from __future__ import division, print_function
 
@@ -13,6 +12,10 @@ import numpy as np, pandas as pd
 from astropy import units as u, constants as c
 from math import pi as π
 import os
+from scipy.interpolate import interp1d
+
+global α
+α = 3.5 # coefficient in the L~M^α power law scaling.
 
 #####################
 # UTILITY FUNCTIONS #
@@ -22,11 +25,72 @@ def get_L(M):
     assume input mass, M, a float or vector, is in solar units.
     '''
 
-    α = 3.5
+    return M**α
 
-    L = M**α
 
-    return L
+def draw_planet_radii(df, singleordouble, δ=None, R_pl=None, R_pu=None):
+    '''
+    Assign planets radii from power law radius probability distribution
+    function:
+    $$
+        ρ_Rpt (R_pt) ~ R_pt^δ
+    $$
+
+    df: DataFrame of singles or doubles
+
+    singleordouble (str): "single" or "double"
+
+    δ (required float): power of the distribution. For instance, Howard et al
+    (2012) find δ=-2.92 +/- 0.11.
+
+    R_pl, R_pu (optional floats): lower and upper bound for truncation. If
+    None, assumed the distribution is not truncated
+    '''
+
+    assert δ
+
+    power = δ
+    ΔR = 0.01
+    R_max = 50 # earth Radii, for grid.
+    R_min = 0
+
+    if isinstance(R_pu,float):
+        assert R_pu < R_max
+    if not isinstance(R_pu,float):
+        R_pu = R_max
+
+    R_pt_grid = np.arange(R_min, R_max+ΔR, ΔR)
+    prob_Rpt = np.zeros_like(R_pt_grid)
+
+    inds = (R_pt_grid > R_pl) & (R_pt_grid < R_pu)
+    prob_Rpt[inds] = (R_pt_grid[inds])**power
+    prob_Rpt /= trapz(prob_Rpt, R_pt_grid)
+
[…]

### 2017-10-08 src/model_2_semianalytic.py
minor changes

@@ -1,0 +1,44 @@
+'''
+integrate expression for <X_Γ>
+17/10/06.1
+'''
+
+import numpy as np
+from scipy.integrate import trapz
+
+q = np.arange(0.1,1+1e-4,1e-4)
+α = 3.5 # M~L^α
+
+BF = 0.44
+curlyB = BF/(1-BF)
+
+f_d = 0.475
+
+# lower limit:
+integrand = (1+q**α)**(3/2) * (1 + curlyB*(1+q**α)**(3/2)) / \
+            ( 1 + curlyB*(1+q**α)**(3/2)* \
+              ((1+q**α)**(-3) + f_d * q**(2/3) * (1+q**(-α))**(-3)*q**(-5)  )
+            )
+
+numerator = trapz(integrand, q)
+
+denominator = trapz((1+q**α)**(3/2), q)
+
+expected_X_Γ = numerator/denominator
+
+print('lower limit <X_Γ>: ', expected_X_Γ)
+
+# upper limit:
+
+integrand = (1+q**α)**(3/2) * (1 + curlyB*(1+q**α)**(3/2)) \
+            / \
+            ( 1 + curlyB*(1+q**α)**(-3/2)
+            )
+
+numerator = trapz(integrand, q)
+
+denominator = trapz((1+q**α)**(3/2), q)
+
+expected_X_Γ = numerator/denominator
+
+print('upper limit <X_Γ>: ', expected_X_Γ)

### 2017-10-08 src/vary_Rp_study_pdfs.py

@@ -140,6 +140,8 @@ def _make_Rpobs_plot(R_pt, prob_Rpts, R_pas, prob_Rpos_d1, prob_Rpos_d2,
 
     axs[0].set_ylabel(r'$\rho_{R_p^{\mathrm{true}}} (R_p^{\mathrm{true}}) $')
     axs[0].legend(loc='upper right', fontsize='small')
+    if sstr == 'uniform_narrow':
+        axs[0].set_ylim([0,1])
 
     norm_d1 = trapz(prob_Rpos_d1, R_pas)
     prob_Rpos_d1 /= norm_d1
@@ -152,6 +154,8 @@ def _make_Rpobs_plot(R_pt, prob_Rpts, R_pas, prob_Rpos_d1, prob_Rpos_d2,
 
     axs[1].set_ylabel(r'$\rho_{R_p^{\mathrm{obs}}} (R_p^{\mathrm{obs}}) $')
     axs[1].legend(loc='upper right', fontsize='small')
+    if sstr == 'uniform_narrow':
+        axs[1].set_ylim([0,1])
 
     if not howardpower:
         f_det = logistic_fn(R_pas, 0.95, 1, 10)
@@ -476,15 +480,15 @@ def case_5(γ_R, R_pt, R_pas):
 
 if __name__ == '__main__':
 
-    γ_R = np.logspace(-4, 0, num=int(4e3))
-    R_pt = np.linspace(0, 30, int(3e3)) # Re units
-    R_pas = np.linspace(0, 30, int(3e3)) # Re units
+    γ_R = np.logspace(-4, 0, num=int(1e4))
+    R_pt = np.linspace(0, 30, int(1e4)) # Re units
+    R_pas = np.linspace(0, 30, int(1e4)) # Re units
 
     case_1(γ_R, R_pt, R_pas, R_pu=15, R_pl=14.99, sstr='uniform_narrow')
-    case_1(γ_R, R_pt, R_pas)
-    case_2(γ_R, R_pt, R_pas)
-    case_3(γ_R, R_pt, R_pas)
-    case_4(γ_R, R_pt, R_pas)
-    case_5(γ_R, R_pt, R_pas)
+    #case_1(γ_R, R_pt, R_pas)
+    #case_2(γ_R, R_pt, R_pas)
+    #case_3(γ_R, R_pt, R_pas)
+    #case_4(γ_R, R_pt, R_pas)
+    #case_5(γ_R, R_pt, R_pas)
 
     #TODO: bin it. 

### 2017-10-11 memos/model_1.tex
added new memo and associated numerics

\section*{Generalities}

Define the occurrence rate density, $\Gamma$,
as the expected number of planets per star per natural logarithmic bin 
of planetary and stellar phase space:
\begin{equation}
\Gamma(\vec{x}) \equiv \frac{{\rm d}^n(N_p/N_\star)}{\prod_{i=1}^{n} {\rm d} 
\ln x_i}
= \frac{d^n\Lambda}{\prod_{i=1}^{n} {\rm d} \ln x_i  }.
\end{equation}
$\vec{x}$ is an $n$-dimensional list of the continuous physical properties 
that might affect the occurrence rate density. For example,
$\vec{x}=(r,P,R)$ where $r$ is the planet radius, $P$ is its orbital period, 
and $R$ is the host star radius.
In the latter equality, $\Lambda$ is the occurrence rate~--~a quantity found 
by integrating the rate density over a specified volume of phase space ({\it 
e.g.}, Foreman-Mackey et al, 2014).

According to the previous definition, the rate density implicitly marginalizes 
over stellar multiplicity.
For simplicity, this work only considers single and binary star systems.
Then for a selected population of stars and planets the rate density can be 
written
\begin{equation}
\Gamma(\vec{x}) = \sum_{i=0}^{2} w_i \Lambda_i p_i(\vec{x}) \equiv \sum_i 
\Gamma_i(\vec{x}),
\end{equation}
where $i=0$ corresponds to single star systems, $i=1$ primaries of a binaries, 
and $i=2$ secondaries of binaries.
$\Lambda_i$ is the occurrence rate for the $i^{\rm th}$ system 
type, $p_i(\vec{x})$ is the corresponding joint probability density function, 
and the weights are given by 
\begin{equation}
w_i = N_i/N_{\rm tot},
\end{equation}
for $N_{\rm tot} = \sum_i N_i$ the total number of selected stars, and 
$N_0,N_1,N_2$ the number of selected single stars, primaries, and 
secondaries respectively.

A transit survey will have a rate density of detected planets $\hat{\Gamma}$, 
which will be the (dot) product of the rate density and the detection 
efficiency $Q(\vec{x})$:
\begin{equation}
\hat{\Gamma}(\vec{x}) = \sum_i Q_i(\vec{x}) \Gamma_i(\vec{x}) 
\equiv \sum_i \hat{\Gamma}_i(\vec{x}),
\label{eq:detected_rate_density}
\end{equation}
where again the index $i$ is over each type of system (singles, primaries, 
and secondaries).
The detection efficiency includes the geometric transit probability, as well 
as any incompleteness effects.

\section{Model \#1: fixed stars, fixed planets, twin binaries}
\label{sec:model_1}

Consider a universe in which all planets are identical, and all stars are 
either single or twin stars with otherwise identical physical properties.
Then $\vec{x} = (r,R,a)$, and 
\begin{equation}
p_i(\vec{x})= \delta(r-r_p)\delta(R-R_\star)\delta(a-a_p)
\equiv \delta^3(r_p,R_\star,a_p),
\label{eq:model_1_prob}
\end{equation}
for $r_p$ and $R_\star$ some fixed planet and stellar radii, $a_p$ a fixed 
semi-major axis, and $\delta$ the Dirac delta function, whose latter compact 
form will be used for brevity.

The occurrence rate density for this model is
\begin{equation}
\Gamma(r,R,a) = \sum_i w_i \Lambda_i \delta^3(r_p,R_\star,a_p),
\label{eq:model1_occ_rate_density}
\end{equation}
and the occurrence rate over any interval that includes $r_p$, $R_\star$, and 
$a_p$ is
\begin{equation}
\Lambda = \sum_i w_i \Lambda_i = \frac{ \sum_i N_i \Lambda_i }{N_{\rm tot}}.
\end{equation}
The rate is zero over intervals that do not.

To express the rate density of detected planets, $\hat{\Gamma} = \sum 
Q_i\Gamma_i$, we need the detection efficiencies for each system type, which 
are products of the geometric and selection probabilities:
\begin{align}
Q_i(\vec{x}) &= Q_{g,i}(\vec{x}) Q_{c,i}(\vec{x}),\quad {\rm where}\ 
\vec{x}=(r,R,a).
\label{eq:general_detection_efficiency}
\end{align}
Similar to Pepper et al. (2003), but in a new context, we take $Q_c$ as the 
ratio of the number of stars that were searchable to the number of stars that 
were selected.
Assuming a homogeneous distribution of stars, this gives
\begin{equation}
Q_{c,i}(\vec{x}) = \left(
\frac{d_{{\rm det},i}(\vec{x})}{d_{\rm sel}(\vec{x})}
\right)^3,
\end{equation}
for $d_{\rm sel}$ the maximum distance to which surveyed stars are selected, 
and $d_{{\rm det},i}$ the maximum distance to which planets can actually be 
detected about the $i^{\rm th}$ system type.
Note that $d_{\rm sel} \geq d_{{\rm det},i}$.
In a signal-to-noise limited transit survey in which the observer does not 
know which stars are binaries, 
\begin{equation}
d_{\rm sel} \propto (r/R)^2 (L_{\rm sys} T_{\rm dur} A N_{\rm tra})^{1/2},
\end{equation}
for $L_{\rm sys}=L_1(1+\gamma_R)$ the system luminosity, $T_{\rm dur}$ the 
transit duration, $A$ the detector area, and $N_{\rm tra}$ 
the number of observed transits.
However,
\begin{equation}
d_{{\rm det},i} \propto \mathcal{D}_i(r/R)^2 (L_{\rm sys} T_{\rm dur} A N_{\rm 
tra})^{1/2},
\label{eq:d_det_i}
\end{equation}
for the dilution $\mathcal{D}_i$ given by
\begin{align}
\mathcal{D}_i
&=
\left.
\begin{cases}
1 & \text{for } i=0,\ {\rm single} \\
L_1 / L_d = (1+\gamma_R)^{-1}, & \text{for } i=1,\ {\rm primary} \\
\gamma_R L_1 / L_d = (1 + \gamma_R^{-1})^{-1}, & \text{for } i=2,\ {\rm 
secondary},
\end{cases}
\right.
\label{eq:dilution}
\end{align}
where the light ratio $\gamma_R$ of a given binary is defined as the ratio of 
the luminosity of the secondary to the primary.

The maximum detectable distance to single stars is assumed to be known, and so 
$d_{{\rm sel},0} = d_{{\rm det},0}$.
For binary systems there is a necessary incompleteness, and combining 
Eqs.~\ref{eq:general_detection_efficiency} through~\ref{eq:dilution} in the 
case of twin binaries yields
\begin{align}
Q_i(\vec{x})
&=
\left.
\begin{cases}
R/a, & \text{for } i=0\ {\rm single} \\
(R/a) (1+\gamma_R)^{-3} = (R/8a), & \text{for } i=1\ {\rm primary} \\
(R/a) (1+\gamma_R^{-1})^{-3}\gamma_R^{-5/\alpha} = (R/8a), & \text{for } i=2\ 
{\rm 
secondary}.
\end{cases}
\right.
\label{eq:model1_detection_efficiency}
\end{align}

The factor of $\gamma_R^{-5/\alpha}$ in the latter expression of 
Eq.~\ref{eq:model1_occ_rate_density} comes from the mass-luminosity-radius 
relation of stars in the model: we assume $R\propto M \propto L^{1/\alpha}$.
For $\gamma_R=1$ this term is unimportant, but it will later become relevant.

Summarizing, since we have written the rate density for each system type
(Eq.~\ref{eq:model1_occ_rate_density}) and the detection efficiency for each 
system type (Eq.~\ref{eq:model1_detection_efficiency}), we have fully 
specified the rate density of detected planets, $\hat{\Gamma}=\sum Q_i 
\Gamma_i$, in addition to the true rate density.

\subsection{What does an observer ignoring binarity infer? }

Consider an observer who ignores binarity.
They assume a detection efficiency $\tilde{Q}=Q_0$,
measure a detected planet rate density $\tilde{\Gamma}$, 
and infer an apparent rate density $\Gamma_a$.
Analogous to Eq.~\ref{eq:detected_rate_density},
\begin{equation}
\tilde{\Gamma} = \Gamma_a \tilde{Q}.
\end{equation}
Accounting for dilution, one can show
\begin{equation}
\Gamma_a = 
w_a \Lambda_0 \delta^3(r_p, R_\star, a_p) +
w_b (\Lambda_1 Q_{c,1} + \Lambda_2 Q_{c,2}) 
				\delta^3(r_p/\sqrt{2}, R_\star, a_p),
\end{equation}
for $w_a = N_0/(N_0+N_1)$, and $w_b = N_1/(N_0+N_1)$.
This observer miscounts the number of total searched stars, does not correct 
for completeness, and misclassifies the planetary radii because of dilution.

\subsection{Correction to inferred rate density and inferred rate}

Define the rate density correction factor, $X_\Gamma$, as the ratio of the 
apparent to true rate densities:
\begin{equation}
X_\Gamma \equiv \frac{\Gamma_a}{\Gamma}.
\end{equation}
This factor can be a function of whatever parameters $\Gamma_a$ and $\Gamma$ 
depend on; in this study, the planet radius is most relevant.
For the twin-binaries model,
\begin{equation}
X_\Gamma(r)
=
\frac{w_a \Lambda_0\delta^3(r_p) + 
	w_b(\Lambda_1 Q_{c,1} + \Lambda_2 Q_{c,2}) \delta^3(r_p/\sqrt{2})  }
	{(w_0\Lambda_0 + w_1\Lambda_1 + w_2\Lambda_2)\delta^3(r_p)}
	\label{eq:model1_correction}
	\end{equation}
where $\delta^3(r_p)$ is shorthand for $\delta^3(r-r_p,R-R_\star,a-a_p)$.

If we take the rates $\Lambda_i$ to be equal, applying the definitions of 
the weights gives a rate density correction factor at $r=r_p$ of
$X_\Gamma(r_p) = (1+\mu)^{-1}$, where 
\begin{align}
\mu \equiv \frac{N_1}{N_0} &=
\frac{n_b}{n_s} \left(\frac{d_{\rm sel,b}}{d_{\rm sel,s}}\right)^3 = 
\frac{{\rm BF}}{1-{\rm BF}} (1+\gamma_R)^{3/2},
\label{eq:mu_definition}
\end{align}
for $n_b$ and $n_s$ the number density of binaries and singles in a 
volume limited sample.
Using Raghavan et al. (2010)'s $0.7-1.3M_\odot$ multiplicity fraction as our 
binary fraction, we set ${\rm BF}=0.44$.
The resulting correction to the rate density is $X_\Gamma(r_p) \approx 0.31$. 
The correction at $r_p/\sqrt{2}$ is infinite.
%beta = 2.2223355980148636

If we assume that $\Lambda_0 = \Lambda_1$, but that $\Lambda_2=0$, we find 
$X_\Gamma(r_p) = (1+2\mu)/(1+\mu)^2$.
Taking the same binary fraction, this evaluates to $X_\Gamma(r_p)\approx 0.53$.

In passing, note that a correction to the inferred rate, $X_\Lambda$, can be 
defined analogously:
\begin{equation}
X_\Lambda \equiv \frac{\Lambda_a}{\Lambda}.
\end{equation}
For the twin binary model, the correction to the rate is the same as that to 
the rate density.

### 2017-10-11 memos/model_2.tex
added new memo and associated numerics

\section{Model \#2: fixed planets and primaries, varying secondaries}
\label{sec:model_2}

The main use of our binary-twin model was to help develop intuition.
Gradually introducing complexity, we now let the light ratio $\gamma_R = 
L_2/L_1$ vary across the binary population.
It does so because the underlying mass ratio $q=M_2/M_1$ varies.
We keep the primary mass fixed as $M_1$, which is also the mass of all single 
stars.

We parametrize the distribution of binary mass ratios in a volume-limited 
sample as a power law: $p(q)\propto q^\beta$.
For binaries with solar-type primaries\footnote{
Duchene and Kraus (2013), fitting all the multiple systems of Raghavan et al. 
(2010)'s Fig 16, find $\beta = 0.28\pm0.05$ for $0.7<M_\star/M_\odot<1.3$.
Examining only the binary systems of Rhagavan et al 2010, Fig 16, the 
distribution seems roughly uniform, $\beta \approx 0$, except for a claimed 
excess of twin binaries with $q\approx 1$, and a deficiency of $q<0.1$ 
stellar companions.
}, $\beta$ is probably between 0 and 0.3.
Since we assume stars are a one-parameter family, $R \propto M \propto 
L^{1/\alpha}$, a drawn value of $q$ determines everything about a secondary.

The rate density in this model, $\Gamma(\vec{x})$, is the sum over system 
types of $w_i \Lambda_i p_i(\vec{x})$:
\begin{equation}
\Gamma(\vec{x})
=
\delta^4(r_p,R_\star,a_p,P_p)(w_0 \Lambda_0 + w_1 \Lambda_1)
+ w_2 \Lambda_2 \delta^3(r_p, P_p, a_p) p_2(q),
\label{eq:model2_rate_density}
\end{equation}
where the semimajor axis of the planet must be such that its period is $P_p$, 
and $p_2(q)$ is expressed in terms of the mass ratio instead of the secondary 
star's radius for convenience ($q$ and $R_2$ are interchangeable).
The probability that a secondary hosts a planet, as a function of the mass 
ratio, is
\begin{align}
p_2(q) &= p({\rm has\ planet}\,|\,{\rm secondary\ with\ }q) \times
          p({\rm secondary\ with\ }q)
          \\
p_2(q) &\propto q^{\gamma + \beta} (1+q^\alpha)^{3/2}.
\label{eq:model_2_p_2}
\end{align}
We take first term, $p({\rm has\ planet}\,|\,{\rm secondary\ with\ }q)$, as a 
power law of $q$ with exponent $\gamma$.
For the second term, since the selected sample at a given $(r,P,a)$ is
magnitude-limited, $p({\rm secondary\ with\ }q)$ 
is the product of the volume limited probability and a 
Malmquist-like bias $(1+q^\alpha)^{3/2}$.

The occurrence rate corresponding to Eq.~\ref{eq:model2_rate_density}'s rate 
density is
\begin{equation}
\Lambda = \sum_i
\left[
\left(
    w_i \Lambda_i \int_{\Omega_{\rm desired}} p_i(\vec{x}) \,{\rm d}\Omega
\right)
\cdot
\left(
    w_i \Lambda_i \int_{\Omega_{\rm total}} p_i(\vec{x}) \,{\rm d}\Omega
\right)^{-1}
\right]
\end{equation}
for a desired volume of phase space $\Omega_{\rm desired}$.
Specifying the desired mass ratios of interest as $q_{\rm min} < q < q_{\rm 
max}$, this simplifies to
\begin{equation}
\Lambda = 
\frac{N_0 \Lambda_0 + N_1 \Lambda_1 + N_2 \Lambda_2 f_2}
{N_{\rm tot}},
\end{equation}
for
\begin{equation}
f_2 \equiv
\left(
\int_{q_{\rm min}}^{q_{\rm max}} p_2(q) \,{\rm d}q
\right)
\cdot
\left(
\int_{0}^{1} p_2(q) \,{\rm d}q
\right)^{-1}.
\end{equation}

The detected rate density, $\hat{\Gamma} = \sum_i Q_i \Gamma_i$, will be 
specified by the detection efficiencies for each type of system.
These are nearly identical to Eq.~\ref{eq:model1_detection_efficiency}:
\begin{align}
Q_0 &= Q_{g,0}Q_{c,0} = Q_{g,0} \label{eq:model2_Q_0}\\
Q_1 &= Q_{g,1}Q_{c,1} = Q_{g,0} (1+q^\alpha)^{-3} \\
Q_2 &= Q_{g,2}Q_{c,2} = Q_{g,0} q^{2/3} (1+q^{-\alpha})^{-3} q^{-5}, 
\label{eq:model2_Q_2}
\end{align}
for $Q_{g,0}=R/a$, the transit probability in single star systems.
The detection efficiency for secondaries (Eq.~\ref{eq:model2_Q_2}) includes 
the transit probability from the smaller stellar radius, and combines 
dilution, the transit duration, and stellar radius for the completeness
probability.

\subsection{What does an observer ignoring binarity infer?}

As a reminder, the apparent rate density is found by correcting the detected 
apparent rate density for the transit probability:
$\Gamma_a = \tilde{\Gamma} Q_{g,0}^{-1}$.
The observer's errors are as follows:
\begin{enumerate}
    \item The true planetary radii $r$ are interpreted as apparent radii $r_a$.
    \item The true stellar radii $R$ are all thought to be $R_\star$. In 
    ``reality'', this only holds for single stars and primaries.
    \item The selected and searchable stars are miscounted.
\end{enumerate}

Writing the apparent rate density as a function of $\vec{x}=(r_a,P,a,R)$, 
where $R$ is written interchangeably with the binary mass ratio $q$,
\begin{equation}
\Gamma_a(\vec{x}) =
Q_0 w_a \Lambda_0 \delta^4(r_p)
+
Q_1 w_b \Lambda_1 \delta^3(P_p) p_1(r_a)
+
Q_2 w_b \Lambda_2 \delta^2(P_p,a_p) p_2(q) p_2(r_a),
\label{eq:model2_Gamma_a}
\end{equation}
where the detection efficiencies are given in 
Eqs.~\ref{eq:model2_Q_0}-\ref{eq:model2_Q_2}, $p_2(q)$ is 
given by Eq.~\ref{eq:model_2_p_2}, and as in the first model, 
$w_a=N_0/(N_0+N_1)$, $w_b=N_1/(N_0+N_1)$.
%WARNING: is it actually correct to separate the q and r_a dependence?
%THEY are dependent on each-other...

The apparent radii depend on the system type:
\begin{align}
r_a
&=
\left.
\begin{cases}
r_p (1+q^\alpha)^{-1/2} & \text{for } i=1,\ {\rm primary} \\
r_p (1+q^{-\alpha})^{-1/2} q^{-1}, & \text{for } i=2,\ {\rm secondary}.
\end{cases}
\right.
\label{eq:model2_ra}
\end{align}
The factor of $q^{-1}$ for the secondary case accounts for the observer 
assuming that the host star is the primary.

Since the apparent radii directly depend on the mass ratio, we wish to write
\begin{equation}
p_i (r_a)
=
p_i(q(r_a))
\left|
    \frac{{\rm d}q}{{\rm d}r_a}
\right|,
\quad i\in\{1,2\},
\label{eq:model2_p_i}
\end{equation}
where $q(r_a)$ must be found by inverting Eq.~\ref{eq:model2_ra}.
For $i=1$, this gives
\begin{equation}
p_1(r_a) \propto \frac{r_p^5}{r_a^6}
\left( 
    \left(\frac{r_p}{r_a}\right)^2 - 1
\right)^{\frac{\beta - \alpha + 1}{\alpha}},
\end{equation}
but for $i=2$, $r_a(q)$ is not invertible.
Instead, we find $p_2(r_a)$ numerically\footnote{We 
sample from $p_2(q)$ (Eq.~\ref{eq:model_2_p_2}), and use 
Eq.~\ref{eq:model2_p_i} to compute the resulting apparent radii}, 
and show it in Fig.~\ref{fig:model2_prob_r_a}.

\subsection{Correction to inferred rate density}

Recall that the rate density correction factor, $X_\Gamma$, is the ratio of 
the apparent to true rate densities.
Applying Eqs.~\ref{eq:model2_rate_density} and~\ref{eq:model2_Gamma_a},
\begin{align}
X_\Gamma(r,q)
=
\frac{
    w_a \Lambda_0 \delta(r_p,R_\star)
    +
    Q_{c,1} w_b \Lambda_1 \delta(R_\star) p_1(r_a)
    +
    Q_{c,2} q^{2/3} w_b \Lambda_2 p_2(q) p_2(r_a)
}
{
w_0 \Lambda_0 \delta^2(r_p,R_\star) +
w_1 \Lambda_1 \delta^2(r_p,R_\star) +
w_2 \Lambda_2 p_2(q) \delta(r_p)
},
\label{eq:model2_X_Gamma}
\end{align}
where the stellar radius-dependent $\delta$-functions are not important.
To marginalize over $q$, we must write
\begin{equation}
X_\Gamma(r) = \frac{\Gamma_a(r)}{\Gamma(r)}
= \frac{\sum_i \int_0^1 \Gamma_{a,i}(r,q) \,{\rm d}q}
{\sum_i \int_0^1 \Gamma_{i}(r,q) \,{\rm d}q}.
\label{eq:model2_X_Gamma_of_r}
\end{equation}
Using the same definition of $\mu = N_1/N_0$ from Eq.~\ref{eq:mu_definition}, 
it is important when performing each integral to note that $\mu$ is a function 
of $q$:
\begin{equation}
\mu = \mathcal{B} (1+q^\alpha)^{3/2},
\quad \mathcal{B} \equiv \frac{{\rm BF}}{1 - {\rm BF}} .
\end{equation}
Thus the weights in the apparent and actual rate densities, $(w_a,w_b)$ and 
$(w_0,w_1,w_2)$, are all functions of $q$.

The crucial point of Eq.~\ref{eq:model2_X_Gamma_of_r} is that since both the 
real and apparent rate densities are separable, the expressions will simplify.

Our ``nominal model'' is a stellar population similar to Sun-like stars in the 
local neighborhood:
${\rm BF}=0.44$, $\alpha=3.5$, $\beta=0$.
We also assume that the occurrence of planets is independent of stellar mass 
($\gamma=0$).
Under these nominal assumptions, the planetary rate density is
\begin{equation}
\Gamma(r) \approx \delta(r-r_p) \left( \Lambda_0 + \Lambda_1 + 
\Lambda_2 \right) / 3,
\label{eq:model2_Gamma_r}
\end{equation}
where the coefficients of $1/3$ are accurate to within one percent of the true 
coefficients.
Ignoring binarity, the observer finds an apparent rate density
\begin{equation}
\Gamma_a(r) = c_0 \Lambda_0 \delta(r-r_p)
             +c_1 \Lambda_1 p_1(r_a)
             +c_2 \Lambda_2 p_2(r_a),
\label{eq:model2_Gamma_a_r}
\end{equation}
for $c_0\approx 0.49$, $c_1\approx 0.32$, $c_2\approx 0.03$, and the two 
latter apparent-radius dependences are shown in Fig.~\ref{fig:model2_prob_r_a}.

While the detailed radial dependence of the error between 
Eqs.~\ref{eq:model2_Gamma_r} 
and~\ref{eq:model2_Gamma_a_r} is interesting, the general point is
that dilution produces a spectrum of apparent planetary radii, which
skews occurrence rate and occurrence rate density measurements.

Evaluating the correction term at $r=r_p$, since $\lim_{q\rightarrow0} 
p_i(r_a)=0$ for $i\in\{1,2\}$,
\begin{equation}
X_\Gamma(r=r_p) \approx \frac{3c_0 \Lambda_0}{\Lambda_0+\Lambda_1+\Lambda_2}.
\end{equation}
If all the rates are equal, $X_\Gamma(r=r_p)\approx0.49$.
If there are no planets around the secondaries, $X_\Gamma(r=r_p)\approx0.74$.

\section{Model \#3: Fixed primaries, varying planets and secondaries}

\begin{figure}[!b]
    \begin{center}
        \includegraphics[width=.9\textwidth]{figures/prob_r_a.pdf}
    \end{center}
    \caption{Model \#2 (Sec.~\ref{sec:model_2}): 
    probability of observing an apparent radius $r_a$ given a 
    planet orbiting the primary, $p_1(r_a)$, or the secondary, $p_2(r_a)$ of a 
    binary system.
    The true planet radius is fixed~--~a delta function centered on ``1''.
    This plot takes $\alpha=3.5, \beta=0$, for $\alpha$ the exponent in 
    the mass-luminosity relation $L \propto M^\alpha$, and $\beta$ the 
    exponent in the distribution of mass ratios in a volume-limited sample.
    Each distribution is separately normalized.
    }
    \label{fig:model2_prob_r_a}
\end{figure}

### 2017-10-11 src/model_2_prob_ra.py

@@ -1,0 +1,154 @@
+'''
+expressions for the probability density of observing a planet with apparent
+radius $r_a$, in model 2
+
+also makes plot for paper
+'''
+# -*- coding: utf-8 -*-
+from __future__ import division, print_function
+import matplotlib as mpl
+mpl.use('pgf')
+pgf_with_custom_preamble = {
+    'pgf.texsystem': 'pdflatex', # xelatex is default; i don't have it
+    'font.family': 'serif', # use serif/main font for text elements
+    'text.usetex': True,    # use inline math for ticks
+    'pgf.rcfonts': False,   # don't setup fonts from rc parameters
+    'pgf.preamble': [
+        '\\usepackage{amsmath}',
+        '\\usepackage{amssymb}'
+        ]
+    }
+mpl.rcParams.update(pgf_with_custom_preamble)
+
+import matplotlib.pyplot as plt
+import numpy as np, pandas as pd
+from scipy.integrate import trapz
+from scipy.integrate import quad
+from scipy.interpolate import interp1d, InterpolatedUnivariateSpline, \
+    LSQUnivariateSpline, UnivariateSpline
+
+global α, β
+β = 0 # p(q)~q^β vol limited
+α = 3.5 # M~L^α
+r_p = 1 #Re, Rj - units don't matter
+
+f,ax = plt.subplots(figsize=(4,4))
+
+#PRIMARIES
+#boxed eqn 17/10/10.1 bottom. Define a lambda function for p1_ra so that you
+#can use quad to integrate over the inf.
+p1_ra = lambda r_a: (r_p/r_a)**3 * r_p**2/(r_a**3) * \
+                    ( (r_p/r_a)**2 -1 )**((β-α+1)/α)
+
+r_a = np.linspace(r_p*2**(-1/2), r_p, num=int(1e5))
+
+p1_ra_vals = p1_ra(r_a)/quad(p1_ra, r_p/np.sqrt(2), r_p)[0]
+
+ax.plot(r_a/r_p, p1_ra_vals, label='$p_1(r_a)$')
+
+#SECONDARIES
+#case 2, eqn 17/10/10.2 middle.
+#probability density if we ignore the secondary stellar radius correction
+p2_ra_case2 = lambda r_a: \
[…]

### 2017-10-11 src/model_2_semianalytic.py

@@ -12,17 +12,25 @@ q = np.arange(0.1,1+1e-4,1e-4)
 BF = 0.44
 curlyB = BF/(1-BF)
 
+β = curlyB * (1+q**α)**(3/2)
+
 f_d = 0.475
 
+
 # lower limit:
-integrand = (1+q**α)**(3/2) * (1 + curlyB*(1+q**α)**(3/2)) / \
-            ( 1 + curlyB*(1+q**α)**(3/2)* \
-              ((1+q**α)**(-3) + f_d * q**(2/3) * (1+q**(-α))**(-3)*q**(-5)  )
-            )
+f_d1c = (1+q**α)**(-3)
+f_d2c = (1+q**(-α))**(-3) * q**(-5)
+
+p_of_q = (1+q**α)**(3/2)
+
+
+integrand = (1 + β*(1+f_d))/(1+2*β) * (1+β) / \
+            (1 + β*(f_d1c + f_d*f_d2c*q**(2/3))) * \
+            p_of_q
 
 numerator = trapz(integrand, q)
 
-denominator = trapz((1+q**α)**(3/2), q)
+denominator = trapz(p_of_q, q)
 
 expected_X_Γ = numerator/denominator
 
@@ -30,15 +38,15 @@ print('lower limit <X_Γ>: ', expected_X_Γ)
 
 # upper limit:
 
-integrand = (1+q**α)**(3/2) * (1 + curlyB*(1+q**α)**(3/2)) \
-            / \
-            ( 1 + curlyB*(1+q**α)**(-3/2)
-            )
-
-numerator = trapz(integrand, q)
-
-denominator = trapz((1+q**α)**(3/2), q)
-
-expected_X_Γ = numerator/denominator
-
-print('upper limit <X_Γ>: ', expected_X_Γ)
+#integrand = (1+q**α)**(3/2) * (1 + curlyB*(1+q**α)**(3/2)) \
+#            / \
+#            ( 1 + curlyB*(1+q**α)**(-3/2)
+#            )
+#
+#numerator = trapz(integrand, q)
+#
+#denominator = trapz((1+q**α)**(3/2), q)
+#
+#expected_X_Γ = numerator/denominator
+#
+#print('upper limit <X_Γ>: ', expected_X_Γ)

### 2017-10-14 src/model_1.py
add model 1 monte carlo

@@ -1,0 +1,241 @@
+'''
+MODEL #1: FIXED STARS, FIXED PLANETS, TWIN BINARIES
+
+The point of this exercise is to develop a numerical procedure to simulate
+these toy surveys.
+'''
+
+from __future__ import division, print_function
+
+import numpy as np, pandas as pd
+from astropy import units as u, constants as c
+from math import pi as π
+import os
+
+global α, β
+α = 3.5
+
+#####################
+# UTILITY FUNCTIONS #
+#####################
+def get_L(M):
+    '''
+    assume input mass, M, a float or vector, is in solar units.
+    '''
+
+    return M**α
+
+
+if __name__ == '__main__':
+
+    ######################
+    # STELLAR POPULATION #
+    ######################
+    #
+    # Construct a pandas DataFrame in which each row is a star that has been
+    # "selected". Each selected star has properties:
+    #   * star_type (str): single, primary, or secondary
+    #   * is_searchable (bool): whether the star is searchable.
+    #   * q (float): binary mass ratio, if applicable
+
+    quickrun = False
+    slowrun = not quickrun
+
+    # arbitrary number of selected single stars
+    N_0 = int(5e2) if quickrun else int(1e5)
+
+    BF = 0.44 # binary fraction. BF = n_d / (n_s+n_d). Raghavan+ 2010 solar.
+    B = BF / (1-BF) # prefactor in definition of μ
+
+    # For a given light ratio (or q), everything is defined.
+    # In model #1, q=1.
+    q = 1
+    μ = B * (1 + q**α)**(3/2)
+
+    N_1 = int(N_0 * μ) # number of selected primaries
+    N_2 = int(N_1)     # number of selected secondaries
+
+    # End with a population of stars: singles, primaries, secondaries.
+    df = pd.concat((
+        pd.DataFrame({'star_type':[type_i for _ in range(N_i)]}) for N_i,type_i in
+        [(N_0,'single'),(N_1,'primary'),(N_2,'secondary')]
+        ), ignore_index=True )
+
+    df['q'] = q
+
+    # Make sure only primaries and secondaries are assigned a value of `q`
+    df.loc[df['star_type'] == 'single', 'q'] = np.nan
+
+    # Select stars that are searchable. This means they win the completeness
[…]

### 2017-10-14 src/model_2_apparent_detected_occ_rate_density.py
add clearer numeris

@@ -1,0 +1,163 @@
+'''
+expressions for the probability density of observing a planet with apparent
+radius $r_a$, in model 2
+
+also makes plot for paper
+'''
+# -*- coding: utf-8 -*-
+from __future__ import division, print_function
+import matplotlib as mpl
+mpl.use('pgf')
+pgf_with_custom_preamble = {
+    'pgf.texsystem': 'pdflatex', # xelatex is default; i don't have it
+    'font.family': 'serif', # use serif/main font for text elements
+    'text.usetex': True,    # use inline math for ticks
+    'pgf.rcfonts': False,   # don't setup fonts from rc parameters
+    'pgf.preamble': [
+        '\\usepackage{amsmath}',
+        '\\usepackage{amssymb}'
+        ]
+    }
+mpl.rcParams.update(pgf_with_custom_preamble)
+
+import matplotlib.pyplot as plt
+import numpy as np, pandas as pd
+from scipy.integrate import trapz
+from scipy.integrate import quad
+from scipy.interpolate import interp1d, InterpolatedUnivariateSpline, \
+    LSQUnivariateSpline, UnivariateSpline
+
+global α, β
+β = 0 # p(q)~q^β vol limited
+α = 3.5 # M~L^α
+γ = 0
+r_p = 1 #Re, Rj - units don't matter
+bf = 0.44
+B = bf / (1-bf)
+
+f,ax = plt.subplots(figsize=(4,4))
+
+#PRIMARIES
+#boxed eqn 17/10/13.2. Define a lambda function for p1_ra so that you
+#can use quad to integrate over the inf.
+
+p1_ra = lambda r_a: \
+        (1 + (( (r_p/r_a)**2 - 1)**(1/α))**α)**(-3) * \
+        (( (r_p/r_a)**2 - 1)**(1/α))**(γ+β) * \
+        (1+(( (r_p/r_a)**2 - 1)**(1/α))**α)**(3/2) * \
+        (B*(1+(( (r_p/r_a)**2 - 1)**(1/α))**α)**(3/2))*( 1 - (B*(1+(( (r_p/r_a)**2 - 1)**(1/α))**α)**(3/2)))**(-1)
+
+r_a = np.linspace(r_p*2**(-1/2), r_p, num=int(1e5))
+
+p1_ra_vals = p1_ra(r_a)/quad(p1_ra, r_p/np.sqrt(2), r_p)[0]
+
[…]

### 2017-10-14 src/model_2_prob_ra.py

@@ -151,4 +151,4 @@ ax.set_ylim([1e-1,1e2])
 ax.tick_params(which='both', direction='in', zorder=0)
 
 f.tight_layout(h_pad=0)
-f.savefig('prob_r_a.pdf', dpi=300, bbox_inches='tight')
+f.savefig('../results/prob_r_a.pdf', dpi=300, bbox_inches='tight')

### 2017-10-14 src/numerical_models.py

@@ -1,0 +1,340 @@
+'''
+This program simulates the following toy surveys:
+
+MODEL #1: Fixed stars, fixed planets, twin binaries.
+
+MODEL #2: Fixed planets and primaries, varying secondaries.
+
+For Model #1, there are analytic predictions for the correction at various
+points -- these can be verified.
+
+For Model #2, there are fewer, but there are still good tests we can apply
+along the way to ensure the result is believable.
+'''
+
+from __future__ import division, print_function
+
+import numpy as np, pandas as pd, matplotlib.pyplot as plt
+from astropy import units as u, constants as c
+from scipy.integrate import trapz
+from scipy.interpolate import interp1d
+from math import pi as π
+import os
+
+global α, β, γ
+
+if __name__ == '__main__':
+    np.random.seed(41)
+
+    ##########################################
+    # Inputs.
+    quickrun = False
+    slowrun = not quickrun
+    model_number = 3
+
+    BF = 0.44 # binary fraction. BF = n_d / (n_s+n_d). Raghavan+ 2010 solar.
+
+    α = 3.5     # exponent in L~M^α
+    β = 0       # exponent in p_vol_limited(q) ~ q^β
+    γ = 0       # exponent in p(has planet|secondary w/ q) ~ q^γ
+    δ = -2.92   # exponent in p(r) ~ r^δ, from Howard+ 2012
+
+    # planet occ rates (only relative values matter).
+    Λ_0 = 0.5 # fraction of selected singles with planet
+    Λ_1 = 0.5 # fraction of selected primaries with planet
+    Λ_2 = 0.5 # fraction of selected secondaries with planet
+    ##########################################
+
+    if γ != 0:
+        raise NotImplementedError
+
+    ######################
+    # STELLAR POPULATION #
+    ######################
+    # Construct a pandas DataFrame in which each row is a star that has been
+    # "selected". Each selected star has properties:
+    #   * star_type (str): single, primary, or secondary
+    #   * is_searchable (bool): whether the star is searchable (determined by
+    #       completeness for star-specific q).
+    #   * q (float): binary mass ratio, if applicable
+    ######################
+
+    # arbitrary number of selected single stars
[…]

### 2017-10-14 src/plot_rate_densities_vs_radius.py

@@ -1,0 +1,86 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+
+def make_plot(model_number, logx=None, logy=None, withtext=None):
+
+    # Make summary plot
+    fname = '../data/results_model_'+repr(model_number)+'.out'
+    df = pd.read_csv(fname)
+
+    f, ax = plt.subplots(figsize=(4,4))
+
+    ax.step(df['bin_left'], df['true_Γ'], where='pre', label='true')
+
+    ax.step(df['bin_left'], df['inferred_Γ'], where='pre', label='inferred')
+
+    ax.legend(loc='best',fontsize='small')
+
+    ax.set_xlabel('planet radius, $r$ [$R_\oplus$]')
+
+    ax.set_ylabel('planets per star per bin, $\Gamma(r)$ [$R_\oplus^{-1}$]')
+
+    if logx:
+        ax.set_xscale('log')
+    if logy:
+        ax.set_yscale('log')
+
+    if logx or logy:
+        outname = '../results/rate_density_vs_radius_logs_model_'+\
+                 repr(model_number)
+    else:
+        outname = '../results/rate_density_vs_radius_model_'+\
+                 repr(model_number)
+
+    if model_number == 3:
+        # Assess HJ rate difference.
+        from scipy.integrate import trapz
+
+        #Howard 2012 boundary #1 and boundary #2:
+        for lower_bound in [8,5.6]:
+            inds = df['bin_left'] > lower_bound
+            Λ_HJ_true = trapz(df[inds]['true_Γ'], df[inds]['bin_left'])
+            Λ_HJ_inferred = trapz(df[inds]['inferred_Γ'], df[inds]['bin_left'])
+
+            Λ_true = trapz(df['true_Γ'], df['bin_left'])
+            Λ_inferred = trapz(df['inferred_Γ'], df['bin_left'])
+
+            txt = \
+            '''
+            with $r>${:.1f}$R_\oplus$,
+            Λ_HJ_true:      {:.4f} planets per star
+            Λ_HJ_inferred:  {:.4f} planets per star
+            true/inferred:  {:.2f}.
+
+            Integrated over all $r$,
+            Λ_true:         {:.3f} planets per star
+            Λ_inferred:     {:.3f} planets per star
+            true/inferred:  {:.2f}.
+            '''.format(
+            lower_bound,
+            Λ_HJ_true,
+            Λ_HJ_inferred,
+            Λ_HJ_true/Λ_HJ_inferred,
[…]

### 2017-10-16 src/numerical_models.py
add README

@@ -1,92 +1,6 @@
+# -*- coding: utf-8 -*-
 '''
-####################
-DESCRIPTION
-
-This program simulates the following toy surveys:
-
-MODEL #1: Fixed stars, fixed planets, twin binaries.
-
-MODEL #2: Fixed planets and primaries, varying secondaries.
-
-MODEL #3: Fixed primaries, varying planets and secondaries.
-
-The simulation works as follows.
-
-First, the user specifies their inputs: the binary fraction, the model class
-(#1,2,3), various exponents (α,β,γ,δ), and the true occurrence rates, Λ_i.
-
-The population of selected stars is constructed as follows.  First, we note
-that each selected star has a `star_type` (single, primary, secondary), a
-binary mass ratio, if applicable, and the property of whether it is
-"searchable".
-
-The absolute number of stars is arbitrary; we take 10^6 single stars
-throughout. The number of binaries is calculated according to analytic
-formulae. The binary mass ratios are, when applicable, samples from the
-appropriate magnitude-limited distribution (given α and β).
-
-Whether a star is "searchable" depends entirely on its "completeness" fraction.
-By "completeness", we mean the ratio of the actual number of searchable stars
-to the assumed number of searchable stars (for a given planet size, period,
-etc.).  Assuming homogeneuously distributed stars, this is equivalent to the
-ratio of the searchable to selected volumes.  In our model, this volume ratio
-is a function of only the binary mass ratio.
-
-The procedure for assigning planets is then as follows:
-    * each selected star of type i gets a planet at rate Λ_i
-    * the radii of planets are assigned independently of any host system
-    property, as sampled from p_r(r) ~ r^δ for Model #3 or a δ function for
-    Models #1 and #2.
-    * a planet is detected when a) it transits, and and b) its host star is
-    searchable.
-
-The probability of transiting single stars in our model is assumed to be known,
-and so it can be mostly corrected by the observer attempting to infer an
-occurrence rate. The only bias is for secondaries, which can be smaller. This
-effect is including when computing the transit probability.
-
-For detected planets, apparent radii are computed according to analytic
-formulae that account for both dilution and the misclassification of stellar
-radii. We assume that the observer thinks they observe the primary.
-
-The rates are then computed in bins of true planet radius and apparent planet
-radius.
-
-In a given radius bin, the true rate is found by counting the number of planets
-that exist around selected stars of all types (singles, primaries,
-secondaries), and dividing by the total number of these stars.
-
[…]

### 2017-10-16 src/numerical_models.py
add errcases output parsing

@@ -21,9 +21,9 @@ if __name__ == '__main__':
     # Inputs.
     quickrun = False
     slowrun = not quickrun
-    model_number = 3
+    model_number = 2
 
-    BF = 0.44 # binary fraction. BF = n_d / (n_s+n_d). Raghavan+ 2010 solar.
+    BF = 0.44   # binary fraction. BF = n_d / (n_s+n_d). Raghavan+ 2010 solar.
 
     α = 3.5     # exponent in L~M^α
     β = 0       # exponent in p_vol_limited(q) ~ q^β
@@ -245,9 +245,9 @@ if __name__ == '__main__':
             (df['has_det_planet']==False)]['r_a'])))
 
 
-    ##############################################
-    # TRUE RATE DENSITY VS APPARENT RATE DENSITY #
-    ##############################################
+    ##############################################################
+    # TRUE RATE DENSITY VS APPARENT RATE DENSITY (NOMINAL MODEL) #
+    ##############################################################
 
     if model_number == 1 or model_number == 2:
         Δr = 0.01
@@ -316,9 +316,9 @@ if __name__ == '__main__':
             decimal=3
             )
 
-    #######################
-    # Final sanity checks #
-    #######################
+    ###############################
+    # Nominal model sanity checks #
+    ###############################
     if model_number == 1 or model_number== 2:
         Γ_inferred = N_p_at_r_p_inferred / N_tot_apparent
         Γ_true = N_p_at_r_p_true / N_tot
@@ -356,3 +356,116 @@ if __name__ == '__main__':
                 X_Γ_rp_analytic,
                 decimal=2
                 )
+
+    ###############
+    # ERROR CASES #
+    ###############
+    for error_case_number in [1,2,3]:
+        print('Beginning error case {:d}'.format(error_case_number))
+
+        true_dict = {}
+        inferred_dict = {}
+
+        for type_i in types:
+
+            # True rate densities are invariant of error case.
+            r = df[(df['star_type']==type_i) & (df['has_planet'])]['r']
+            N_p, edges = np.histogram(r, bins=radius_bins)
+
+            true_dict[type_i] = {}
+            true_dict[type_i]['N_p'] = N_p
+            true_dict[type_i]['edges'] = edges
+
+            # Inferred rate densities. Error case #3 gets incorrect, other two
+            # are correct.
+            if error_case_number == 1 or error_case_number == 2:
+                r_a = df[(df['star_type']==type_i) & (df['has_det_planet'])]['r']
+            elif error_case_number == 3:
[…]

### 2017-10-16 src/plot_errcases_rate_densities_vs_radius.py

@@ -2,7 +2,7 @@
 first, for models 1,2, 3:
 >>> python numerical_models.py
 then:
->>> python plot_rate_densities_vs_radius.py
+>>> python plot_absolute_rate_densities_vs_radius.py
 
 makes plots for paper
 '''
@@ -30,10 +30,10 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
         stdout=False):
 
     # Make summary plot
+    f, ax = plt.subplots(figsize=(4,4))
+
     fname = '../data/results_model_'+repr(model_number)+'.out'
     df = pd.read_csv(fname)
-
-    f, ax = plt.subplots(figsize=(4,4))
 
     if model_number == 3:
         xvals = np.append(0, df['bin_left'])
@@ -46,7 +46,31 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
 
     ax.step(xvals, ytrue, where='post', label='true')
 
-    ax.step(xvals, yinferred, where='post', label='inferred')
+    ax.step(xvals, yinferred, where='post', label='all errors')
+
+    for error_case_number in [1,2,3]:
+
+        fname = '../data/results_model_{:d}_error_case_{:d}.out'.format(
+                model_number, error_case_number)
+        df = pd.read_csv(fname)
+
+        if model_number == 3:
+            xvals = np.append(0, df['bin_left'])
+            ytrue = np.append(0, df['true_Γ'])
+            yinferred = np.append(0, df['inferred_Γ'])
+        elif model_number == 1 or model_number == 2:
+            xvals = np.append(df['bin_left'],1)
+            ytrue = np.append(df['true_Γ'],0)
+            yinferred = np.append(df['inferred_Γ'],0)
+
+        if error_case_number == 1:
+            label = 'only incorrect $N_\star$'
+        elif error_case_number == 2:
+            label = 'only incorrect $Q$'
+        elif error_case_number == 3:
+            label = 'only incorrect $r_a$'
+
+        ax.step(xvals, yinferred, where='post', label=label)
 
     ax.legend(loc='best',fontsize='medium')
 
@@ -60,16 +84,18 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
         ax.set_yscale('log')
 
     if logx or logy:
-        outname = '../results/rate_density_vs_radius_logs_model_'+\
+        outname = '../results/errcases_rate_density_vs_radius_logs_model_'+\
                  repr(model_number)
     else:
[…]

### 2017-10-17 src/numerical_models.py
two bugfixes

@@ -15,13 +15,13 @@ import os
 global α, β, γ
 
 if __name__ == '__main__':
-    np.random.seed(1337)
+    np.random.seed(42)
 
     ##########################################
     # Inputs.
     quickrun = False
     slowrun = not quickrun
-    model_number = 2
+    model_number = 3
 
     BF = 0.44   # binary fraction. BF = n_d / (n_s+n_d). Raghavan+ 2010 solar.
 
@@ -52,8 +52,8 @@ if __name__ == '__main__':
     #   * q (float): binary mass ratio, if applicable
     ######################
 
-    # arbitrary number of selected single stars
-    N_0 = int(5e2) if quickrun else int(1e6)
+    # arbitrary number of selected single stars. 1e6 is pretty fast.
+    N_0 = int(5e2) if quickrun else int(5e6)
 
     B = BF / (1-BF) # prefactor in definition of μ
 
@@ -85,7 +85,7 @@ if __name__ == '__main__':
         df['q'] = np.nan
 
         # Inverse transform sampling to get samples of q
-        q_grid = np.arange(0, 1+1e-6, 1e-6)
+        q_grid = np.arange(1e-6, 1+1e-6, 1e-6)
         prob_ml_q = q_grid**β * (1+q_grid**α)**(3/2)
         prob_ml_q /= trapz(prob_ml_q, q_grid)
         cdf_ml_q = np.append(0, np.cumsum(prob_ml_q)/np.max(np.cumsum(prob_ml_q)))
@@ -383,7 +383,6 @@ if __name__ == '__main__':
             elif error_case_number == 3:
                 r_a = df[(df['star_type']==type_i) & (df['has_det_planet'])]['r_a']
 
-
             inferred_dict[type_i] = {}
 
             if error_case_number == 1 or error_case_number == 3:
@@ -404,11 +403,10 @@ if __name__ == '__main__':
 
                     inds = (r_a > lower_edge) & (r_a <= upper_edge)
 
-                    if len(inds) == 0:
+                    if not np.any(inds):
                         # If there are no detected planets in this bin,
                         # detection efficiency is irrelevant for inferred rate.
-                        Q = 1
-                        this_N_p_det = 0
+                        this_weighted_N_p_det = 0
 
                     else:
                         # This bin has detected planets. Each planet contibutes
@@ -417,7 +415,8 @@ if __name__ == '__main__':
                         q = df_detd[inds]['q']
 
                         if type_i == 'single':
-                            Q_g = Q_g0
[…]

### 2017-10-17 src/plot_errcases_rate_densities_vs_radius.py
update errcases plots

@@ -54,6 +54,17 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
                 model_number, error_case_number)
         df = pd.read_csv(fname)
 
+        if error_case_number == 3:
+            try:
+                np.testing.assert_almost_equal(
+                        np.sum(df['inferred_Γ']),
+                        np.sum(df['true_Γ']),
+                        decimal=1
+                        )
+            except:
+                print(model_number, error_case_number)
+                print('assert failed')
+
         if model_number == 3:
             xvals = np.append(0, df['bin_left'])
             ytrue = np.append(0, df['true_Γ'])
@@ -63,12 +74,15 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
             ytrue = np.append(df['true_Γ'],0)
             yinferred = np.append(df['inferred_Γ'],0)
 
+            #if model_number==1:
+            #    import IPython; IPython.embed()
+
         if error_case_number == 1:
-            label = 'only incorrect $N_\star$'
+            label = 'only wrong $N_\star$'
         elif error_case_number == 2:
-            label = 'only incorrect $Q$'
+            label = 'only wrong $Q$'
         elif error_case_number == 3:
-            label = 'only incorrect $r_a$'
+            label = 'only wrong $r_a$'
 
         ax.step(xvals, yinferred, where='post', label=label)
 
@@ -94,49 +108,58 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
         ax.set_xlim([0.5,1.02])
 
     if model_number == 3:
-        fname = '../data/results_model_'+repr(model_number)+'.out'
-        df = pd.read_csv(fname)
-        # Assess HJ rate difference.
+
         from scipy.integrate import trapz
 
+        # Assess HJ rate difference.
         #Howard 2012 boundary #1 and boundary #2:
         for lower_bound in [5.6,8]:
-            inds = df['bin_left'] > lower_bound
-            #Λ_HJ_true = trapz(df[inds]['true_Γ'], df[inds]['bin_left'])
-            #Λ_HJ_inferred = trapz(df[inds]['inferred_Γ'], df[inds]['bin_left'])
-            Λ_HJ_true = np.sum(df[inds]['true_Γ'])
-            Λ_HJ_inferred = np.sum(df[inds]['inferred_Γ'])
[…]

### 2017-10-19 memos/generalities.tex
clarify generalies text

--- 2017-10-11 memos/model_1.tex
@@
 \begin{equation}
-\Gamma(\vec{x}) \equiv \frac{{\rm d}^n(N_p/N_\star)}{\prod_{i=1}^{n} {\rm d} 
-\ln x_i}
-= \frac{d^n\Lambda}{\prod_{i=1}^{n} {\rm d} \ln x_i  }.
+\Gamma(\vec{x}) = \frac{d^n\Lambda}{\prod_{i=1}^{n} {\rm d} \ln x_i  }.
+\label{eq:rate_density_defn}
 \end{equation}
 $\vec{x}$ is an $n$-dimensional list of the continuous physical properties 
-that might affect the occurrence rate density. For example,
+that might affect the occurrence rate density. 
+For example,
 $\vec{x}=(r,P,R)$ where $r$ is the planet radius, $P$ is its orbital period, 
 and $R$ is the host star radius.
-In the latter equality, $\Lambda$ is the occurrence rate~--~a quantity found 
-by integrating the rate density over a specified volume of phase space ({\it 
-e.g.}, Foreman-Mackey et al, 2014).
+The occurrence rate $\Lambda$ is found by integrating the rate density over a 
+specified volume of phase space.
 
-According to the previous definition, the rate density implicitly marginalizes 
-over stellar multiplicity.
+The previous definition implicitly marginalizes the rate density over stellar 
+multiplicity.
 For simplicity, this work only considers single and binary star systems.
@@
 \begin{equation}
-\Gamma(\vec{x}) = \sum_{i=0}^{2} w_i \Lambda_i p_i(\vec{x}) \equiv \sum_i 
-\Gamma_i(\vec{x}),
+\Gamma(\vec{x})
+= \sum_{i=0}^{2} w_i \Gamma_i(\vec{x})
+= \sum_i w_i \Lambda_i p_i(\vec{x})
+\label{eq:rate_density_marginalized}
 \end{equation}
-where $i=0$ corresponds to single star systems, $i=1$ primaries of a binaries, 
+where $i=0$ corresponds to single star systems, $i=1$ primaries of binaries, 
 and $i=2$ secondaries of binaries.
-$\Lambda_i$ is the occurrence rate for the $i^{\rm th}$ system 
-type, $p_i(\vec{x})$ is the corresponding joint probability density function, 
+$\Lambda_i$ is the occurrence rate integrated over all possible phase space  
+for the $i^{\rm th}$ system type, $p_i(\vec{x})$ is the joint 
+probability density function so that $\Gamma_i(\vec{x}) = \Lambda_i 
+p_i(\vec{x})$, 
 and the weights are given by 
@@
 $N_0,N_1,N_2$ the number of selected single stars, primaries, and 
-secondaries respectively.
+secondaries respectively\footnote{
+Eq.~\ref{eq:rate_density_marginalized} follows by writing the $i^{\rm 
+th}$ system type's rate density as some normalization multiplied by a 
+probability density:
+$\Gamma_i(\vec{x}) = \mathcal{Z}_i p_i(\vec{x})$.
+For Eq.~\ref{eq:rate_density_defn} to hold, we must have $\mathcal{Z}_i = 
+\Lambda_i$.
+}.
+The relationship between the rate $\Lambda$ over a desired volume of 
+phase space $\Omega_{\rm desired}$ and $\Lambda_i$ is
+\begin{equation}
+\Lambda = \sum_i
+\left[
+\left(
+w_i \Lambda_i \int_{\Omega_{\rm desired}} p_i(\vec{x}) \,{\rm d}\Omega
+\right)
+\cdot
+\left(
+w_i \Lambda_i \int_{\Omega_{\rm total}} p_i(\vec{x}) \,{\rm d}\Omega
+\right)^{-1}
+\right],
+\label{eq:occ_rate}
+\end{equation}
+where the inverse term is unity if $p_i(\vec{x})$ is appropriately normalized.
 
@@
 as any incompleteness effects.
-
-\section{Model \#1: fixed stars, fixed planets, twin binaries}
-\label{sec:model_1}
-
-Consider a universe in which all planets are identical, and all stars are 
-either single or twin stars with otherwise identical physical properties.
-Then $\vec{x} = (r,R,a)$, and 
-\begin{equation}
-p_i(\vec{x})= \delta(r-r_p)\delta(R-R_\star)\delta(a-a_p)
-\equiv \delta^3(r_p,R_\star,a_p),
-\label{eq:model_1_prob}
-\end{equation}
-for $r_p$ and $R_\star$ some fixed planet and stellar radii, $a_p$ a fixed 
-semi-major axis, and $\delta$ the Dirac delta function, whose latter compact 
-form will be used for brevity.
-
-The occurrence rate density for this model is
-\begin{equation}
-\Gamma(r,R,a) = \sum_i w_i \Lambda_i \delta^3(r_p,R_\star,a_p),
-\label{eq:model1_occ_rate_density}
-\end{equation}
-and the occurrence rate over any interval that includes $r_p$, $R_\star$, and 
-$a_p$ is
-\begin{equation}
-\Lambda = \sum_i w_i \Lambda_i = \frac{ \sum_i N_i \Lambda_i }{N_{\rm tot}}.
-\end{equation}
-The rate is zero over intervals that do not.
-
-To express the rate density of detected planets, $\hat{\Gamma} = \sum 
-Q_i\Gamma_i$, we need the detection efficiencies for each system type, which 
-are products of the geometric and selection probabilities:
-\begin{align}
-Q_i(\vec{x}) &= Q_{g,i}(\vec{x}) Q_{c,i}(\vec{x}),\quad {\rm where}\ 
-\vec{x}=(r,R,a).
-\label{eq:general_detection_efficiency}
-\end{align}
-Similar to Pepper et al. (2003), but in a new context, we take $Q_c$ as the 
-ratio of the number of stars that were searchable to the number of stars that 
-were selected.
-Assuming a homogeneous distribution of stars, this gives
-\begin{equation}
-Q_{c,i}(\vec{x}) = \left(
-\frac{d_{{\rm det},i}(\vec{x})}{d_{\rm sel}(\vec{x})}
-\right)^3,
-\end{equation}
-for $d_{\rm sel}$ the maximum distance to which surveyed stars are selected, 
-and $d_{{\rm det},i}$ the maximum distance to which planets can actually be 
-detected about the $i^{\rm th}$ system type.
-Note that $d_{\rm sel} \geq d_{{\rm det},i}$.
-In a signal-to-noise limited transit survey in which the observer does not 
-know which stars are binaries, 
-\begin{equation}
-d_{\rm sel} \propto (r/R)^2 (L_{\rm sys} T_{\rm dur} A N_{\rm tra})^{1/2},
-\end{equation}
-for $L_{\rm sys}=L_1(1+\gamma_R)$ the system luminosity, $T_{\rm dur}$ the 
-transit duration, $A$ the detector area, and $N_{\rm tra}$ 
-the number of observed transits.
-However,
-\begin{equation}
-d_{{\rm det},i} \propto \mathcal{D}_i(r/R)^2 (L_{\rm sys} T_{\rm dur} A N_{\rm 
-tra})^{1/2},
-\label{eq:d_det_i}
-\end{equation}
-for the dilution $\mathcal{D}_i$ given by
-\begin{align}
-\mathcal{D}_i
-&=
-\left.
-\begin{cases}
-1 & \text{for } i=0,\ {\rm single} \\
-L_1 / L_d = (1+\gamma_R)^{-1}, & \text{for } i=1,\ {\rm primary} \\
-\gamma_R L_1 / L_d = (1 + \gamma_R^{-1})^{-1}, & \text{for } i=2,\ {\rm 
-secondary},
-\end{cases}
-\right.
-\label{eq:dilution}
-\end{align}
-where the light ratio $\gamma_R$ of a given binary is defined as the ratio of 
-the luminosity of the secondary to the primary.
-
-The maximum detectable distance to single stars is assumed to be known, and so 
-$d_{{\rm sel},0} = d_{{\rm det},0}$.
-For binary systems there is a necessary incompleteness, and combining 
-Eqs.~\ref{eq:general_detection_efficiency} through~\ref{eq:dilution} in the 
-case of twin binaries yields
-\begin{align}
-Q_i(\vec{x})
-&=
-\left.
-\begin{cases}
-R/a, & \text{for } i=0\ {\rm single} \\
-(R/a) (1+\gamma_R)^{-3} = (R/8a), & \text{for } i=1\ {\rm primary} \\
-(R/a) (1+\gamma_R^{-1})^{-3}\gamma_R^{-5/\alpha} = (R/8a), & \text{for } i=2\ 
-{\rm 
-secondary}.
-\end{cases}
-\right.
-\label{eq:model1_detection_efficiency}
-\end{align}
-
-The factor of $\gamma_R^{-5/\alpha}$ in the latter expression of 
-Eq.~\ref{eq:model1_occ_rate_density} comes from the mass-luminosity-radius 
-relation of stars in the model: we assume $R\propto M \propto L^{1/\alpha}$.
-For $\gamma_R=1$ this term is unimportant, but it will later become relevant.
-
-Summarizing, since we have written the rate density for each system type
-(Eq.~\ref{eq:model1_occ_rate_density}) and the detection efficiency for each 
-system type (Eq.~\ref{eq:model1_detection_efficiency}), we have fully 
-specified the rate density of detected planets, $\hat{\Gamma}=\sum Q_i 
-\Gamma_i$, in addition to the true rate density.
-
-\subsection{What does an observer ignoring binarity infer? }
-
-Consider an observer who ignores binarity.
-They assume a detection efficiency $\tilde{Q}=Q_0$,
-measure a detected planet rate density $\tilde{\Gamma}$, 
-and infer an apparent rate density $\Gamma_a$.
-Analogous to Eq.~\ref{eq:detected_rate_density},
-\begin{equation}
-\tilde{\Gamma} = \Gamma_a \tilde{Q}.
-\end{equation}
-Accounting for dilution, one can show
-\begin{equation}
-\Gamma_a = 
-w_a \Lambda_0 \delta^3(r_p, R_\star, a_p) +
-w_b (\Lambda_1 Q_{c,1} + \Lambda_2 Q_{c,2}) 
-				\delta^3(r_p/\sqrt{2}, R_\star, a_p),
-\end{equation}
-for $w_a = N_0/(N_0+N_1)$, and $w_b = N_1/(N_0+N_1)$.
-This observer miscounts the number of total searched stars, does not correct 
-for completeness, and misclassifies the planetary radii because of dilution.
-
-\subsection{Correction to inferred rate density and inferred rate}
-
-Define the rate density correction factor, $X_\Gamma$, as the ratio of the 
-apparent to true rate densities:
-\begin{equation}
-X_\Gamma \equiv \frac{\Gamma_a}{\Gamma}.
-\end{equation}
-This factor can be a function of whatever parameters $\Gamma_a$ and $\Gamma$ 
-depend on; in this study, the planet radius is most relevant.
-For the twin-binaries model,
-\begin{equation}
-X_\Gamma(r)
-=
-\frac{w_a \Lambda_0\delta^3(r_p) + 
-	w_b(\Lambda_1 Q_{c,1} + \Lambda_2 Q_{c,2}) \delta^3(r_p/\sqrt{2})  }
-	{(w_0\Lambda_0 + w_1\Lambda_1 + w_2\Lambda_2)\delta^3(r_p)}
-	\label{eq:model1_correction}
-	\end{equation}
-where $\delta^3(r_p)$ is shorthand for $\delta^3(r-r_p,R-R_\star,a-a_p)$.
-
-If we take the rates $\Lambda_i$ to be equal, applying the definitions of 
-the weights gives a rate density correction factor at $r=r_p$ of
-$X_\Gamma(r_p) = (1+\mu)^{-1}$, where 
-\begin{align}
-\mu \equiv \frac{N_1}{N_0} &=
-\frac{n_b}{n_s} \left(\frac{d_{\rm sel,b}}{d_{\rm sel,s}}\right)^3 = 
-\frac{{\rm BF}}{1-{\rm BF}} (1+\gamma_R)^{3/2},
-\label{eq:mu_definition}
-\end{align}
-for $n_b$ and $n_s$ the number density of binaries and singles in a 
-volume limited sample.
-Using Raghavan et al. (2010)'s $0.7-1.3M_\odot$ multiplicity fraction as our 
-binary fraction, we set ${\rm BF}=0.44$.
-The resulting correction to the rate density is $X_\Gamma(r_p) \approx 0.31$. 
-The correction at $r_p/\sqrt{2}$ is infinite.
-%beta = 2.2223355980148636
-
-If we assume that $\Lambda_0 = \Lambda_1$, but that $\Lambda_2=0$, we find 
-$X_\Gamma(r_p) = (1+2\mu)/(1+\mu)^2$.
-Taking the same binary fraction, this evaluates to $X_\Gamma(r_p)\approx 0.53$.
-
-In passing, note that a correction to the inferred rate, $X_\Lambda$, can be 
-defined analogously:
-\begin{equation}
-X_\Lambda \equiv \frac{\Lambda_a}{\Lambda}.
-\end{equation}
-For the twin binary model, the correction to the rate is the same as that to 
-the rate density.

### 2017-10-19 memos/model_1.tex
clarify generalies text

@@
-\section*{Generalities}
-
-Define the occurrence rate density, $\Gamma$,
-as the expected number of planets per star per natural logarithmic bin 
-of planetary and stellar phase space:
-\begin{equation}
-\Gamma(\vec{x}) \equiv \frac{{\rm d}^n(N_p/N_\star)}{\prod_{i=1}^{n} {\rm d} 
-\ln x_i}
-= \frac{d^n\Lambda}{\prod_{i=1}^{n} {\rm d} \ln x_i  }.
-\end{equation}
-$\vec{x}$ is an $n$-dimensional list of the continuous physical properties 
-that might affect the occurrence rate density. For example,
-$\vec{x}=(r,P,R)$ where $r$ is the planet radius, $P$ is its orbital period, 
-and $R$ is the host star radius.
-In the latter equality, $\Lambda$ is the occurrence rate~--~a quantity found 
-by integrating the rate density over a specified volume of phase space ({\it 
-e.g.}, Foreman-Mackey et al, 2014).
-
-According to the previous definition, the rate density implicitly marginalizes 
-over stellar multiplicity.
-For simplicity, this work only considers single and binary star systems.
-Then for a selected population of stars and planets the rate density can be 
-written
-\begin{equation}
-\Gamma(\vec{x}) = \sum_{i=0}^{2} w_i \Lambda_i p_i(\vec{x}) \equiv \sum_i 
-\Gamma_i(\vec{x}),
-\end{equation}
-where $i=0$ corresponds to single star systems, $i=1$ primaries of a binaries, 
-and $i=2$ secondaries of binaries.
-$\Lambda_i$ is the occurrence rate for the $i^{\rm th}$ system 
-type, $p_i(\vec{x})$ is the corresponding joint probability density function, 
-and the weights are given by 
-\begin{equation}
-w_i = N_i/N_{\rm tot},
-\end{equation}
-for $N_{\rm tot} = \sum_i N_i$ the total number of selected stars, and 
-$N_0,N_1,N_2$ the number of selected single stars, primaries, and 
-secondaries respectively.
-
-A transit survey will have a rate density of detected planets $\hat{\Gamma}$, 
-which will be the (dot) product of the rate density and the detection 
-efficiency $Q(\vec{x})$:
-\begin{equation}
-\hat{\Gamma}(\vec{x}) = \sum_i Q_i(\vec{x}) \Gamma_i(\vec{x}) 
-\equiv \sum_i \hat{\Gamma}_i(\vec{x}),
-\label{eq:detected_rate_density}
-\end{equation}
-where again the index $i$ is over each type of system (singles, primaries, 
-and secondaries).
-The detection efficiency includes the geometric transit probability, as well 
-as any incompleteness effects.
-
 \section{Model \#1: fixed stars, fixed planets, twin binaries}
@@
 1 & \text{for } i=0,\ {\rm single} \\
-L_1 / L_d = (1+\gamma_R)^{-1}, & \text{for } i=1,\ {\rm primary} \\
-\gamma_R L_1 / L_d = (1 + \gamma_R^{-1})^{-1}, & \text{for } i=2,\ {\rm 
-secondary},
+L_1 / L_{\rm sys} = (1+\gamma_R)^{-1}, & \text{for } i=1,\ {\rm primary} \\
+\gamma_R L_1 / L_{\rm sys} = (1 + \gamma_R^{-1})^{-1}, & 
+    \text{for } i=2,\ {\rm secondary},
 \end{cases}

### 2017-10-19 results/make_merged.sh
improve rate density and errcase  plots

@@ -1,0 +1,20 @@
+#!/usr/bin/env bash
+pdftk \
+  errcases_rate_density_vs_radius_model_1.pdf \
+  errcases_rate_density_vs_radius_model_1_brokenx.pdf \
+  errcases_rate_density_vs_radius_model_2.pdf \
+  errcases_rate_density_vs_radius_logs_model_2.pdf \
+  errcases_rate_density_vs_radius_model_3.pdf \
+  errcases_rate_density_vs_radius_logs_model_3.pdf \
+  errcases_rate_density_vs_radius_model_3_withtext.pdf \
+  output errcases_merged.pdf
+
+pdftk \
+  rate_density_vs_radius_model_1.pdf \
+  rate_density_vs_radius_model_1_brokenx.pdf \
+  rate_density_vs_radius_model_2.pdf \
+  rate_density_vs_radius_logs_model_2.pdf \
+  rate_density_vs_radius_model_3.pdf \
+  rate_density_vs_radius_logs_model_3.pdf \
+  rate_density_vs_radius_model_3_withtext.pdf \
+  output rates_merged.pdf

### 2017-10-19 src/plot_errcases_rate_densities_vs_radius.py

@@ -26,11 +26,24 @@ import matplotlib.pyplot as plt
 import pandas as pd
 import numpy as np
 
+from brokenaxes import brokenaxes
+
 def make_plot(model_number, logx=None, logy=None, withtext=None,
-        stdout=False):
+        stdout=False, brokenx=False):
 
     # Make summary plot
-    f, ax = plt.subplots(figsize=(4,4))
+    if not brokenx:
+        f, ax = plt.subplots(figsize=(4,4))
+    else:
+        #cf https://github.com/bendichter/brokenaxes/blob/master/brokenaxes.py
+        f = plt.figure(figsize=(4,4))
+        bigax = brokenaxes(
+                xlims=((0.695, .715), (.985, 1.005)),
+                d=0.02,
+                tilt=85,
+                hspace=.05,
+                despine=True)
+        ax=bigax
 
     fname = '../data/results_model_'+repr(model_number)+'.out'
     df = pd.read_csv(fname)
@@ -40,9 +53,9 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
         ytrue = np.append(0, df['true_Γ'])
         yinferred = np.append(0, df['inferred_Γ'])
     elif model_number == 1 or model_number == 2:
-        xvals = np.append(df['bin_left'],1)
-        ytrue = np.append(df['true_Γ'],0)
-        yinferred = np.append(df['inferred_Γ'],0)
+        xvals = np.append(df['bin_left'],[1,1.1])
+        ytrue = np.append(df['true_Γ'],[0,0])
+        yinferred = np.append(df['inferred_Γ'],[0,0])
 
     ax.step(xvals, ytrue, where='post', label='true')
 
@@ -70,12 +83,9 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
             ytrue = np.append(0, df['true_Γ'])
             yinferred = np.append(0, df['inferred_Γ'])
         elif model_number == 1 or model_number == 2:
-            xvals = np.append(df['bin_left'],1)
-            ytrue = np.append(df['true_Γ'],0)
-            yinferred = np.append(df['inferred_Γ'],0)
-
-            #if model_number==1:
-            #    import IPython; IPython.embed()
+            xvals = np.append(df['bin_left'],[1,1.1])
+            ytrue = np.append(df['true_Γ'],[0,0])
+            yinferred = np.append(df['inferred_Γ'],[0,0])
 
         if error_case_number == 1:
[…]

### 2017-10-19 src/plot_rate_densities_vs_radius.py

@@ -26,29 +26,45 @@ import matplotlib.pyplot as plt
 import pandas as pd
 import numpy as np
 
+from brokenaxes import brokenaxes
+
 def make_plot(model_number, logx=None, logy=None, withtext=None,
-        stdout=False):
+        stdout=False, brokenx=None):
 
     # Make summary plot
     fname = '../data/results_model_'+repr(model_number)+'.out'
     df = pd.read_csv(fname)
 
-    f, ax = plt.subplots(figsize=(4,4))
+    if not brokenx:
+        f, ax = plt.subplots(figsize=(4,4))
+    else:
+        #cf https://github.com/bendichter/brokenaxes/blob/master/brokenaxes.py
+        f = plt.figure(figsize=(4,4))
+        bigax = brokenaxes(
+                xlims=((0.695, .715), (.985, 1.005)),
+                d=0.02,
+                tilt=85,
+                hspace=.05,
+                despine=True)
+        ax=bigax
 
     if model_number == 3:
         xvals = np.append(0, df['bin_left'])
         ytrue = np.append(0, df['true_Γ'])
         yinferred = np.append(0, df['inferred_Γ'])
     elif model_number == 1 or model_number == 2:
-        xvals = np.append(df['bin_left'],1)
-        ytrue = np.append(df['true_Γ'],0)
-        yinferred = np.append(df['inferred_Γ'],0)
+        xvals = np.append(df['bin_left'],[1,1.1])
+        ytrue = np.append(df['true_Γ'],[0,0])
+        yinferred = np.append(df['inferred_Γ'],[0,0])
 
     ax.step(xvals, ytrue, where='post', label='true')
 
     ax.step(xvals, yinferred, where='post', label='inferred')
 
-    ax.legend(loc='best',fontsize='medium')
+    if brokenx:
+        ax.legend(loc='upper left',fontsize='medium')
+    else:
+        ax.legend(loc='best',fontsize='medium')
 
     ax.set_xlabel('planet radius, $r$ [$R_\oplus$]', fontsize='large')
 
@@ -65,9 +81,14 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
     else:
         outname = '../results/rate_density_vs_radius_model_'+\
                  repr(model_number)
+    if brokenx:
+        outname += '_brokenx'
 
-    if model_number == 1 or model_number == 2 and not (logx or logy):
-        ax.set_xlim([0.5,1.02])
+    if not brokenx:
[…]

### 2017-10-19 memos/toy_analytic_surveys_171010.tex
moving things around

--- 2017-10-05 paper/ms.tex
@@
 For the binary twin population, all planet and stellar radii are identical, so 
-${\rm S/N} \propto \mathcal{D}, and
-L^{1/2} d^{-1}$,
+${\rm S/N} \propto \mathcal{D} L^{1/2} d^{-1}$, and
 \begin{equation}
@@
 \begin{equation}
-f_{\rm desired} = \frac{\int_{M_{\rm min}}^{M_{\rm max}} {\rm prob}_{\rm 
+f_{\rm desired} = \frac{\int_{q_{\rm min}}^{q_{\rm max}} {\rm prob}_{\rm 
         ml}(q)\,{\rm d}q }
-{\int_{0.1M_1}^{M_1} {\rm prob}_{\rm 
+{\int_{0.1}^{1} {\rm prob}_{\rm 
         ml}(q)\,{\rm d}q}.

### 2017-10-19 paper/main.tex
moving things around

--- 2017-08-30 memos/toy_analytic_surveys_170804.tex
@@
-%\documentclass[12pt,preprint]{aastex}
-%\documentclass[preprint2,12pt]{aastex}
-\documentclass{emulateapj}
-%\usepackage{apjfonts}
+\documentclass[12pt,modern]{aastex61}
 \usepackage{graphicx}
@@
 
-% ------------------------------------------------------------------------
-% New commands
-%
-\def\ltsima{$\; \buildrel < \over \sim \;$}
-\def\lsim{\lower.5ex\hbox{\ltsima}}
-\def\gtsima{$\; \buildrel > \over \sim \;$}
-\def\gsim{\lower.5ex\hbox{\gtsima}}
-\def\tess{{\it TESS} }
-\def \teff {T_{\rm eff}}
-\def \phir {\Phi_{\rm R}}
-\def \fov {24$^{\circ}$}
-\def \pixsz {21.1''}
-\def \aeff {69.1 cm$^2$ }    
-\def \epd {105 mm}
-
-% -------------------------------------------------------------------------
-%
-
-\bibliographystyle{apj}
-
-\title{ Analytic Transit Surveys }
+\title{ Binarity's effects on occurrence rates measured by transit surveys}
 
@@
 
-% \journalinfo{Draft version}
-\slugcomment{Memo for internal use}
-
-%\altaffiltext{1}{Princeton University}
-
-\begin{abstract}
-
-We describe a few different toy transit surveys.
-
-First, we derive the signal to noise distribution of planets detected in a 
-survey where all the stars and planets are the same (except some stars are twin 
-binary systems). We give expressions for the number of detected planets, and 
-for possible errors one can make when deriving occurrence rates.
-
-Then, we allow the properties of the secondary in binary systems to vary. A 
-distribution of light ratios introduces additional biases, which we discuss.
-We reassess the errors related to occurrence rates.
-
-\end{abstract}
-
-\keywords{planets and satellites:\ detection}
-
-\section{Model \#1: fixed stars, fixed planets, fixed light-ratio binaries}
-\label{sec:model_1}
-
-Ah, transiting planets! We learn so much by studying them. But how much can we 
-actually learn, and how much is messed up by binarity?
-
-Imagine the following transit survey:
-\begin{enumerate}
-\item You are going to observe the entire sky for a duration $T_{\rm obs}$, 
-with a detector of area $A$, and known bandpass. Your detector is photon-noise 
-limited.
-%
-\item You are interested only in detecting planets of radius $R_p$, and orbital 
-period $P$. For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
-%
-\item You are only interested in detecting them around stars of radius $R_1$, 
-and luminosity $L_1$. For instance, G2V dwarfs.
-%
-\item You can only detect your `nominal planet' when you observe signals with
-${\rm S/N} > {\rm (S/N)_{min}}$.
-For a photon-noise limited survey, this minimum signal to noise ratio is 
-equivalent to a minimum flux, $F_{\rm min}$.
-%
-\item However, you opt to observe all the ``points'' on-sky with apparent 
-magnitude $m < m_{\rm lim}$, or equivalently with energy flux in your bandpass 
-greater than some limit $F > F_{\rm lim}$.
-(You will not be able to detect planets around the faintest of 
-these, and will need to perform a completeness correction to later derive 
-occurrence rates.)
-Generally $F_{\rm lim} < F_{\rm min}$, \textit{e.g.}, whenever the original 
-magnitude cut defines a ``source catalog'' from which ``searchable stars'' are 
-selected, as in the KIC [Batalha et al, 2010, Brown et al 2011] and the TIC 
-[Stassun et al 2017].
-\end{enumerate}
-
-You get funding, and perform the S/N-limited survey. All planets with 
-${\rm S/N} > {\rm (S/N)_{min}}$ are detected. The rest are not.
-You now wish to derive an occurrence rate for planets of radius $R_p$ and 
-orbital period $P$.
-Assume your universe is a universe in which:
-\begin{itemize}
-	\item The true population of ``points'' (stellar systems, all unresolved) 
-	comprises both single and double star systems. Single star systems have 
-	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
-	temperature $T_{\rm eff,1}$.
-	Double star systems have luminosity in the observed bandpass $L_d = 
-	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
-	the secondary to the primary. 
-	In this section, $\gamma_R$ is a constant across the population 
-	of star systems.
-	The ratio of the two number densities in a 
-	volume-limited sample is the binary fraction\footnote{The binary fraction 
-		is equivalent to the multiplicity fraction if there are no triple, 
-		quadruple, $\ldots$ systems.}.
-	\item The true population of planets around these stars is as follows:
-	\subitem A fraction $\Gamma_{t,s}$ of stars in single star systems 
-	have a planet of radius $R_p$, with orbital period $P$.
-	\subitem A fraction $\Gamma_{t,d}$ of each star in a double star 
-	system has a planet of radius $R_p$, with orbital period $P$. For instance, 
-	if $\Gamma_{t,s} = \Gamma_{t,d} = 0.1$, on average each double 
-	system contributes 0.2 planets, and each single system 0.1 planets.
-	Any astrophysical difference in planet formation between singles and 
-	binaries is captured by these two terms.
-	Note that we are taking an limit by assuming the primary and secondary of 
-	a binary system host planets at equal rates. If the secondary is 
-	non-identical to the primary, it's clearly wrong.	
-\end{itemize}
-
-For simplicity, one might wish to imagine the observer correctly 
-pre-selects all of the ``searchable stars'', in the style of [Pepper et al, 
-2003].
-In other words, they would begin by knowing ${\rm (S/N)_{min}}$, and then only 
-observe stars for which they would have adequate flux to achieve it -- setting 
-$F_{\rm lim} = F_{\rm min}$.
-This simplification would naively allow the observer to ignore 
-completeness effects when deriving occurrence rates, since their completeness 
-is 100\%.
-
-We do not begin this way because in a stellar population of both singles and 
-binaries, the above approach ignores necessary incompleteness for 
-binary systems.
-If we set the limiting flux to give 100\% completeness for single stars, we 
-will unwittingly select binaries near that flux limit, for which dilution 
-will push transit signals below the detection threshold.
-This is because a flux limit maps onto three different maximum 
-detection distances for a population of single stars and twin binaries:
-\begin{enumerate}
-\item $d_{{\rm max},s}$: the maximum distance out to which single stars are
-  selected. This can be chosen to coincide with the maximum distance out to
-  which planets are detectable around single stars.
-%
-\item $d_{{\rm max},d}^\star$: the maximum distance out to which double stars
-  are selected. This is different from $d_{{\rm max},d}^{\rm p}$, the maximum
-  distance out to which planets are detectable around double stars. For a 
-  population with fixed  $\gamma_R$, since ${\rm S/N} \propto \mathcal{D} 
-  L^{-1/2} d$,
-  \begin{equation}
-    d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
-    (1 + \gamma_R)^{-1} d_{{\rm max},d}^\star.
-  \end{equation}
-  For $\gamma_R = 1$, this means only 1 in 8 selected binary stars can yield
-  detectable planets.
-\end{enumerate}
-We develop tools to address this and related completeness effects from the 
-outset.
-
-Consider the following questions:
-\begin{enumerate}
-\item How many single and double star systems, respectively, are in the sample?
-Correspondingly, how many stars are in the sample?
-
-\item How many planets are in the sample? (Orbiting single stars, and orbiting 
-double stars respectively).
-
-\item What is the true occurrence rate?
-
-\item How many planets are detected?
-
-\item What occurrence rate does astronomer A, who has never heard of binary 
-star systems, or completeness, derive for planets of radius $R_p$ and period 
-$P$?
-
-\item What occurrence rate does astronomer B, who accounts for the ``2 for 1'' 
-effect of binarity (\textit{i.e.} that the sample actually has more stars than 
-astronomer A thought) derive? (Astronomer B neglects completeness).
-
-\item What about astronomer C, who accounts for ``2 for 1'' \textit{and} 
-misclassification due to diluted radii? In other words, astronomer C did a 
-combination of high resolution imaging and RV followup on every candidate, and
-correctly classifies the planetary radii in every case.
-
-\item What about astronomer D, who additionally notes the importance of 
-completeness?
-\end{enumerate}
-
-\subsection{How many stars are in the sample?}
-
-Let $N_s$ be the number of single star systems, and $N_d$ the number of double 
-star systems. Then the total number of stars in the sample is
-\begin{equation}
-N_{\rm stars} = N_s + 2N_d.
-\end{equation}
-
-In a magnitude-limited sample in which stars are uniformly distributed in 
-volume, the number of stars will be the number density times the volume.
-If the volume is taken to be a sphere over which the number density is uniform,
-\begin{equation}
-N_i = n_{i} \frac{4\pi}{3} d_{{\rm max},i}^3,
-\label{eq:number_systems}
-\end{equation}
-for $i\in{\rm \{single,double\} } \equiv { \{s,d\} }$, and
-\begin{equation}
-\frac{n_{d}}{n_{s}} = {\rm binary\ fraction} \equiv {\rm BF}
-\end{equation}
-by definition. The absolute normalization of the number density is a measured 
-quantity, as is the binary fraction. For G2V dwarfs, the latter is $\approx 
-0.45$ [Duchene \& Kraus, 2013]. The former is given by [Bovy 2017].
-
-$d_{{\rm max},i}$ in Eq.~\ref{eq:number_systems} is the maximum distance 
-corresponding to the given magnitude limit:
-\begin{equation}
-d_{{\rm max},i} = \left(\frac{L_{i}}{4 \pi F_{{\rm min}}}\right)^{1/2},
-\label{eq:d_max}
-\end{equation}
-where the limiting flux in the bandpass $F_{{\rm min}}$ can also be 
-stated in terms of the limiting magnitude $m_{{\rm min}}$,
-\begin{equation}
-m_{{\rm min}} = m_{0} - \frac{5}{2} \log_{10} \left(\frac{F_{{\rm 
-min}}}{F_{0}}\right),
-\end{equation}
-for $m_{0}$ a zero-point magnitude and $F_{0}$ its corresponding flux (as 
-always, everything is implicitly written in a defined bandpass).
-
-In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$, and as a 
-consequence the maximum distance to which binary stars will be selected is 
-greater than that of single stars, simply as a consequence of imposing a 
-magnitude cut.
-The ratio of double to single systems is
-\begin{align}
-\frac{N_d}{N_s} &= 
-	\frac{n_d}{n_s} \left(\frac{d_{{\rm max}, d} }{d_{{\rm max}, s}}\right)^3 \\
-&= {\rm BF} \times (1+\gamma_R)^{3/2}.
-\end{align}
-In the nominal case of twin binaries ($\gamma_R = 1$), with a binary fraction 
-${\rm BF} = 0.5$, there are 
-$\sqrt{2}$ more binary systems than single systems in the sample.
-Correspondingly, there are $2\sqrt{2}$ more stars in binary systems than stars 
-in single systems.
-
-As a comment on Eq.~\ref{eq:number_systems}, if 
-we wished to write a stellar number density profile that accounted for the 
-vertical structure of the Milky Way, we might choose a profile either 
-$\propto \exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from 
-the galactic midplane and $H$ a scale-height. Both density profiles would lead 
-closed form analytic solutions.
-
-\subsection{How many planets are in the sample?}
-
-The number of planets in the sample is
-\begin{align}
-N_{\rm planets} =& N_{\rm planets\ in\ single\ star\ systems}  +  \\
-				  &\quad\quad N_{\rm planets\ in\ double\ star\ systems} 
-				  \nonumber \\
-			   =& \Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d.
-\end{align}
-
-The factor of 2 accounts for the fact that there are twice as many stars in 
-double star systems.
-
-\subsection{What is the true occurrence rate?}
-\label{sec:true_rate}
-
-The ``true occurrence rate'' is the average number of planets per star. Thus
-
-\begin{align}
-\Gamma_t &= \frac{N_{\rm planets}}{N_{\rm stars}} \\
-\Gamma_t &= \frac{\Gamma_{t,s} N_s + 2 \Gamma_{t,d} N_d}{N_s + 2N_d}.
-\label{eq:true_occ}
-\end{align}
-
-\subsection{How many planets are detected?}
-The total number of planet detections is the sum of the number of planets 
-detected in single star systems $N_{{\rm det},s}$ and the number of planets 
-detected in double star systems $N_{{\rm det},d}$.
-These can be expressed individually. 
-The former is
-\begin{equation}
-N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,{\rm geom}} f_{s,{\rm S/N} > {\rm 
-(S/N)_{min}}},
-\label{eq:N_det_s}
-\end{equation}
-where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
-star systems of the sample, $f_{s,{\rm geom}}\equiv f_{s,g}$ is the geometric 
-transit probability, and $f_{s,({\rm S/N} > {\rm (S/N)_{min}})} \equiv f_{s,c}$ 
-is the fraction of these transiting planets that are observed with signal to 
-noise greater than the minimum detection threshold (the completeness). 
-Analogously,
-\begin{equation}
-N_{{\rm det},d} = 2 N_d \Gamma_{t,d} f_{d,g} f_{d,c},
-\label{eq:N_det_d}
-\end{equation}
-where now $2 N_d \Gamma_{t,d}$ is the number of planets in the double star 
-systems of the sample, the geometric transit probability is the same (if we 
-have twin binaries -- not if we consider a distribution of secondaries) and the 
-completeness term must account for any differences in the signal to noise 
-distribution that come from stellar binarity.
-
-\subsubsection{Analytic completeness}
-Since the geometric transit probability is known, the only terms we have yet to 
-compute are the completeness terms, $f_{i,c}$ for $i\in{\rm \{single,double\} 
-}$. We proceed as follows.
-
-The signal ${\rm S}$ for a box-car train transiting planet is
-\begin{align}
-{\rm S} &= \delta \mathcal{D} \\
-&= \left(\frac{R_p}{R_\star}\right)^2 \mathcal{D},
-\end{align}
-for $R_p$ the planet's radius, $R_\star$ that of its host star, and 
-$\mathcal{D}$ the dilution parameter defined as
-\begin{align}
-  \mathcal{D} &= 
-  \Bigg\{\begin{array}{lr}
-  L_1 / L_d, & {\rm\ if\ binary\ and\ target\ primary}\\
-  \gamma_R L_1 / L_d, & {\rm\ if\ binary\ and\ target\ 
-  secondary}\\
-  1, & {\rm\ if\ single},
-  \end{array} \nonumber\\
-  &=
-    \Bigg\{\begin{array}{lr}
-    (1+\gamma_R)^{-1}, & {\rm\ if\ binary\ and\ target\ primary}\\
-    (1 + \gamma_R^{-1})^{-1}, & {\rm\ if\ binary\ and\ target\ 
-    	secondary}\\
-    1, & {\rm\ if\ single},
-    \end{array} 
-  \label{eq:dilution}
-\end{align}
-where $L_1, L_d,$ and $\gamma_R$ were defined in the opening monograph.
-
-Assuming the only source of noise is Poissonian counting noise, the noise ${\rm 
-N}$ can be written
-\begin{equation}
-{\rm N} = \frac{1}{\sqrt{N_\gamma}},
-\end{equation}
-for $N_\gamma$ the number of photons received by the detector. This noise model 
-is a useful simplification -- see [Howell 2006, pg 75] for the full 
-CCD equation.
-The number of received photons can be written
-\begin{equation}
-N_\gamma = F^{\rm N}_\gamma A N_{\rm tra} T_{\rm dur},
-\end{equation}
-for $F^{\rm N}_\gamma$ the photon number flux from the system [${\rm 
-ph\,cm^{-2}\,s^{-1}}$], 
-$A$ the detector area, $T_{\rm dur}$ the transit duration, and $N_{\rm tra} $ 
-the number of transits observed, which is multiplied in assuming the transits 
-are ``phase-folded''.
-
-Thus the signal to noise ratio can be written
-\begin{equation}
-{\rm S/N} = \delta \mathcal{D} \sqrt{F^{\rm N}_\gamma A N_{\rm tra} T_{\rm 
-dur}}.
-\label{eq:snr_ivory_tower}
-\end{equation}
-
-This means we can write the minimum number flux of photons required for a 
-detection at threshold as
-\begin{equation}
-F_{\rm lim}^N = \left[ \left({\rm \frac{S}{N}}\right)_{\rm min} \frac{1}{\delta 
-	\mathcal{D}} \right]^2 \frac{1}{A N_{\rm tra} T_{\rm dur}}.
-\end{equation}
-To convert this to $F_{\rm lim}$, multiply by the average photon energy in the 
-bandpass.
-
-In passing, given the parameters that define a survey and planet type, 
-Eq.~\ref{eq:snr_ivory_tower} would need to be re-expressed with 
-$N_{\rm tra}$ roughly the ratio of the observing baseline to the planet period, 
-and $T_{\rm dur}$ a function of $R_\star, P, a$, and impact parameter $b$, and 
-then perhaps averaged over $b$. We leave them as-is for subsequent 
-development.
-
-The interesting term in Eq.~\ref{eq:snr_ivory_tower} that changes between stars 
-of the same binarity class in our idealized sample is the square root of 
-$F^{\rm N}_\gamma$.
-This is the term that leads to a distribution of signal to noises for different 
-stars.
-The completenesses $f_{i,c}$ can be directly 
-expressed in terms of those probability density functions:
-\begin{equation}
-f_{i,c} = 
-	\int_{{\rm (S/N)_{min}}}^{\infty} 
-		{\rm d}\left(\frac{{\rm S}}{{\rm N}}\right)_i \ 
-		{\rm prob}\left(\frac{{\rm S}}{{\rm N}}\right)_i.
-\label{eq:completeness_long}
-\end{equation}
-We keep the subscript $i$ because the signal to noise distributions are 
-different for the cases of single star systems ($i=s$) and double star systems 
-($i=d$).
-Notably:
-\begin{itemize}
-	\item The dilution differs (Eq.~\ref{eq:dilution}).
-	\item The photon number flux from the system differs.
-\end{itemize}
-
-To simplify notation, we let $x_i \equiv {\rm (S/N)}_i$, and rewrite 
-Eq.~\ref{eq:completeness_long} as
-\begin{equation}
-f_{i,c} = 
-\int_{{\rm x_{\rm min}}}^{\infty} 
-{\rm d}x_i \ 
-{\rm prob}(x_i).
-\label{eq:completeness_short}
-\end{equation}
-
-\subsubsection{Deriving ${\rm prob}(x_i)$}
-We want expressions for the probability density function of the observed signal 
-to noise ratio, ${\rm prob}(x_i)$, for both the single and binary 
-system case.
-
-First, note that a star placed uniformly in the volume of the search space will 
-have a probability density function for its distance $r$ from the origin of
-\begin{equation}
-{\rm prob} (r) = \frac{3 r^2}{d_{\rm max}^3},
-\end{equation}
-where the appropriate maximum distances should be substituted per 
-Eq.~\ref{eq:d_max}.
-Noting the transformation rule for probability density functions, we can 
-evaluate the probability of a star having a observed number flux $F^{\rm 
-N}_{i,\gamma}$ in the bandpass,
-\begin{align}
-{\rm prob} (F^{\rm N}_{i,\gamma}) 
-&= {\rm prob}(r(F^{\rm N}_{i,\gamma}))
-	\left| \frac{{\rm d} r}{{\rm d} F^{\rm N}_{i,\gamma}} \right| \\
-&= \frac{3}{2 d_{\rm max}^3} c_i^{3/2} (F^{\rm N}_{i,\gamma})^{-5/2},
-\label{eq:pdf_observed_flux}
-\end{align}
-where in the latter equality we have written a ``number luminosity'' $c_i$ 
-(units of inverse time) defined for $i\in{\rm \{single,double\} }$ as
-\begin{equation}
-c_i = 
-\Bigg\{\begin{array}{lr}
-R_1^2 F^{\rm N}_{s1,\gamma}, & {\rm\ if\ single}\\
-R_1^2 F^{\rm N}_{s1,\gamma} + R_2^2 F^{\rm N}_{s2,\gamma} & {\rm\ if\ double}.
-\end{array}
-\label{eq:c_i}
-\end{equation}
-In Eq.~\ref{eq:c_i}, $F^{\rm N}_{s1,\gamma}$ and $F^{\rm N}_{s2,\gamma}$ are 
-the photon number fluxes at the surfaces of the stars. To derive 
-Eq.~\ref{eq:pdf_observed_flux}, we simply scaled these by the distance:
-\begin{equation}
-F^{\rm N}_{i,\gamma} = \frac{c_i}{r^2}.
-\label{eq:flux_at_detector}
-\end{equation}
-
-The surface photon number fluxes $F^{\rm N}_{si,\gamma}$ in Eq.~\ref{eq:c_i} 
-are usually evaluated numerically, by convolving the wavelength-specific photon 
-flux density of a star with the dimensionless spectral response function of the 
-instrument. In other words,
-\begin{equation}
-F^{\rm N}_{s,\gamma} = \int F_\lambda T_\lambda \, {\rm d}\lambda.
-\label{eq:pfd}
-\end{equation}
-The wavelength-specific photon flux density $F_\lambda$ [${\rm 
-ph\,cm^{-2}\,s^{-1}\,\AA^{-1}}$] could be from Pickles' library, or could be 
-a blackbody function. The transmission function is, up to factor of order 
-unity, a step function between two wavelengths $\lambda_{\rm min}$ and 
-$\lambda_{\rm max}$.
-If we assume a blackbody source, Eq.~\ref{eq:pfd} becomes
-\begin{equation}
-F^{\rm N}_{s,\gamma} = 8\pi c \left(\frac{k T}{h c}\right)^3 
-\int_{hc/(\lambda_{\rm max} kT)}^{hc/(\lambda_{\rm min} kT)}
-\frac{u^2}{e^u - 1} \,{\rm d} u,
-\end{equation}
-which can be evaluated numerically\footnote{It may help in the numerics to note 
-that infinite series representations of this type of dimensionless integral 
-exist and converge quickly. For instance, one can show that
-\begin{equation}
-\int_{0}^{a} \frac{u^3}{e^u -1} \,{\rm d}u = \sum_{n=1}^{\infty} \frac{6}{n^4} 
-- \frac{e^{-an}}{n^4} (6 + 6an + 3(an)^2 + (an)^3).
-\nonumber
-\end{equation}
-A similar expression exists for the similar integral in the text. 
-[Michels 1968] explains an analogous derivation.
-}.
-
-The importance of the functional form of $F^{\rm N}_{s,\gamma}$ is that it is 
-to first order only a function of the blackbody temperature and the bandpass 
-wavelength bounds.
-Thus in the most general case $c_i(R_1, R_2, T_1, T_2, \lambda_{\rm min}, 
-\lambda_{\rm max})$ and nothing else.
-The only random variable involved in the flux being received at the detector is 
-$r$, so we can indeed write the flux received at the detector as in 
-Eq.~\ref{eq:flux_at_detector}.
-
-We can finally write out the probability density functions for the signal to 
-noise ratios in the single and double-star cases by using the transformation 
-rule for pdfs, and applying Eq.~\ref{eq:snr_ivory_tower}.
-For single stars,
-\begin{equation}
-{\rm prob}(x_s) = \frac{3}{d_{\rm max, s}^3} c_s^{3/2} \delta^3 \left( A T_{\rm 
-dur} N_{\rm tra}\right)^{3/2} x_s^{-4}.
-\label{eq:prob_xs}
-\end{equation}
-Analogously for double stars,
-\begin{equation}
-{\rm prob}(x_d) = \frac{3}{d_{\rm max, d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 
-\left( A T_{\rm dur} N_{\rm tra}\right)^{3/2} x_d^{-4}.
-\label{eq:prob_xd}
-\end{equation}
-
-\subsubsection{Number of detected planets}
-
-Performing the integrals of Eq.~\ref{eq:completeness_short}, we get:
-
-\begin{align}
-N_{{\rm det},s} &= N_s \Gamma_{t,s} f_{s,g} f_{s,c} \\
-&= N_s \Gamma_{t,s} \frac{R_\star}{a} \frac{1}{d_{\rm max,s}^3} c_s^{3/2} 
-\delta^3 \left(A 
-T_{\rm dur} N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3},
-\end{align}
-and
-\begin{align}
-N_{{\rm det},d} &= 2 N_d \Gamma_{t,d} f_{d,g} f_{d,c} \\
-&= 2 N_d \Gamma_{t,d} \frac{R_\star}{a}
-\frac{1}{d_{\rm max,d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 \left(A T_{\rm dur} 
-N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3}.
-\end{align}
-
-Formally, the $f_{i,c}$ terms should be written as ${\rm min}(1, \ldots)$, 
-where $(\ldots)$ is the given expression. This ensures that the fraction of 
-planets above the signal to noise threshold is less than 1.
-Note that we assumed twin binaries in the geometric transit probabilities.
-Interestingly, the ratio of the two completeness fractions is
-\begin{equation}
-\frac{f_{d,c}}{f_{s,c}} = \left(\frac{d_{\rm max,s}}{d_{\rm max,d}}\right)^3 
-\left(\frac{c_d}{c_s}\right)^{3/2} \mathcal{D}^3 = (1+\gamma_R)^{-3},
-\end{equation}
-which, for the $\gamma_R = 1$ case, gives 
-$1/8$. (We got the same number from simple scaling laws at the outset, but the 
-above machinery will be useful when generalizing in Sec.~\ref{sec:model_2}.)
-
-The number of detected planets $N_{\rm det}$ is the sum of the two appropriate 
-equations above, and can be written
-\begin{align}
-N_{\rm det} = 
-&\left( \frac{\delta}{x_{\rm min}} \right)^3 (A T_{\rm dur} N_{\rm tra})^{3/2}
-\frac{R_\star}{a}
-\nonumber\\
-&\quad \times
-\left[ N_s \Gamma_{t,s} \frac{c_s^{3/2}}{d_{\rm max,s}^3}  +
-	   2 N_d \Gamma_{t,d} \frac{c_d^{3/2}}{d_{\rm max,d}^3}  \right].
-\label{eq:N_det}
-\end{align}
-All terms can be input to a computer, and checked against a Monte Carlo 
-simulation if desired.
-
-\subsection{Astronomer A ignores binarity}
-Astronomer A has never heard of binary star systems. 
-Nor has he heard of completeness corrections.
-He knows about geometric transit probabilities.
-What occurrence rate does he derive for planets of radius $R_p$ and period $P$?
-
-The {\it total} occurrence rate (number of planets divided by number of 
-``stars'') for Astronomer A would be $N_{\rm det}/(N_s + N_d)$, up to the 
-geometric correction. 
-However, even though Astronomer A does not know about binaries, the radii he 
-derives for any planets in binary systems are too small, by a factor 
-$\sqrt{\mathcal{D}}$ (for twin binaries, they are all $R_p/\sqrt{2}$).
-Astronomer A wants an occurrence rate for planets of radius $R_p$ and period 
-$P$.
-The answer is thus
-\begin{equation}
-\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/f_{s,g}}{N_s + N_d}.
-\end{equation}
-
-This astronomer will also think there is a second population of planets, with 
-radius $R_p \sqrt{\mathcal{D}}$, and will thus rush to Nature claiming to also
-have derived a second occurrence rate,
-\begin{equation}
-\Gamma_{{\rm A}, R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm det,d}/f_{s,g}}{N_s 
-+ N_d}.
-\end{equation}
-Note in the second case that this astronomer thinks these are single stars, and 
-generally will miscompute the geometric transit probability.
-
-\subsection{Astronomer B counts host stars correctly}
-Astronomer B can somehow account correctly for the ``2 for 1'' 
-effect of binarity, \textit{i.e.} that the sample actually has more stars than 
-astronomer A thought.
-
-By the same token as above,
-\begin{equation}
-\Gamma_{{\rm B},R_p} = \frac{N_{\rm det,s}/f_{s,g}}{N_s + 2N_d},
-\end{equation}
-and
-\begin{equation}
-\Gamma_{{\rm B},R_p \sqrt{\mathcal{D}}} = 
-		\frac{N_{\rm det,d}/f_{d,g}}{N_s + 2N_d}.
-\end{equation}
-
-\subsection{Astronomer C counts host stars correctly and figures out diluted 
-radii}
-Astronomer C did high resolution imaging followup on every candidate, and 
-correctly classifies the planetary radii.
-Thus, she also knows which planets are in binary systems, and which are in 
-single star systems.
-
-She knows that the purported population of planets with radii $R_p 
-\sqrt{\mathcal{D}}$ does not exist. All detected planets from this survey have 
-radii $R_p$. She computes an occurrence rate
-\begin{equation}
-\Gamma_{{\rm C},R_p} = 
-\frac{N_{\rm det,s}/f_{s,g} + N_{\rm det,d}/f_{d,g}}{N_s + 2N_d}
-\end{equation}
-
-the closest yet to the true rate (Sec.~\ref{sec:true_rate}).
-
-\subsection{Astronomer D counts host stars correctly, figures out diluted 
-radii, 
-and accounts for completeness}
-
-Astronomer D, knows which detections were around 
-binaries and the associated radius correction.
-They do injection recovery, and derive correct estimates for their 
-completeness functions about single stars $f_{s,c}$, and 
-about double star systems, $f_{d,c}$.
-With this knowledge in hand, they compute the respective single and binary 
-occurrence rates
-\begin{equation}
-\Gamma_{t,s} = \frac{N_{\rm det,s}}{N_s f_{s,g} f_{s,c}},
-\end{equation}
-\begin{equation}
-\Gamma_{t,d} = \frac{N_{\rm det,d}}{2 N_d f_{d,g} f_{d,c}}.
-\end{equation}
-
-With these in hand, they derive the overall occurrence rate
-\begin{align}
-\Gamma_{{\rm D}, R_p} 
-&= \frac{N_{\rm det,s}/(f_{s,g}f_{s,c}) + N_{\rm det,d}/(f_{d,g}f_{d,c})  }{N_s 
-+ 2 N_d} \\
-&= \frac{\Gamma_{t,s} N_s + 2\Gamma_{t,d} N_d}{N_s + 2 N_d}.
-\end{align}
-
-Although we admit it's been a bit of a slog, it happens that Astronomer D's 
-occurrence rate is the true occurrence rate (cf. Eq.~\ref{eq:true_occ}).
-
-All it takes is $\approx$ a full semester at Keck, a system-by-system analysis, 
-and perfect understanding of the completeness of the detection efficiency for 
-single and double star systems.
-
-\subsection{Numerical verification}
-
-To check the preceding analytic development, we implemented a Monte Carlo 
-simulation of this idealized transit survey.
-To run the survey, we defined the instrument specifications (detector area and 
-transmission function), the stellar population (binary fraction, total number 
-density of a given stellar class, fixed stellar 
-properties), the planet population (fixed planet radius, period, and occurrence 
-rate about single and binary stars), and finally the survey parameters 
-(observing baseline, minimum SNR for ``detection'').
-We then randomly drew star positions, randomly assigned planets to stars in 
-single and binary systems, and computed the resulting signal to noise 
-(Eq.~\ref{eq:snr_ivory_tower}) with which 
-the transits would be observed.
-As in the preceding analytics, we assumed ``twin'' binaries (same stellar 
-radii, same effective 
-temperature, and dilution does not depend on which stellar binary is the 
-``target'').
-
-The results are shown in Fig.~\ref{fig:snr_dist_analytic_v_numeric}, and 
-indicate that the analytic probability distribution functions 
-Eqs.~\ref{eq:prob_xs},~\ref{eq:prob_xd}, and the number of 
-detections (Eq.~\ref{eq:N_det}) are correct.
-
-A point evident in Fig.~\ref{fig:snr_dist_analytic_v_numeric} is that, for 
-fixed planet parameters, and fixed stellar parameters ($R_\star, L_\star$, and 
-distance $r$) the SNR distribution for planets in binaries is poorer than that 
-of planets in single star systems.
-We can see analytically that this simply due to dilution:
-\begin{equation}
-\frac{{\rm prob}(x_d)}{{\rm prob}(x_s)} =(1 + \gamma_R)^{-1} = \mathcal{D}.
-\end{equation}
-Deriving this simple form requires noting that the ratios of the 
-bandpass-specific number luminosities (Eq.~\ref{eq:c_i}) is equal to the ratio 
-of the bandpass-specific energy luminosities (otherwise a term with $c_s/c_d$ 
-must be included).
-
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=0.4]{figures/snr_distribution.pdf}
-	\end{center}
-	\caption{Comparison of analytic and numeric probability 
-	density functions of the SNR in an idealized transit survey.
-	The analytic lines are Eqs.~\ref{eq:prob_xs} and~\ref{eq:prob_xd} for the 
-	planet populations orbiting single and binary stars. The underlying stepped 
-	histogram is output from Monte Carlo simulations. Poisson noise leads to 
-	a small deviation at the faint and bright limits, but the numerics and 
-	analytics otherwise agree.
-		 }
-	\label{fig:snr_dist_analytic_v_numeric}
-\end{figure}
-
-\subsection{Representative numbers for a few cases}
-
-\subsubsection{Twin binaries: if we ignore binarity, for what fraction of 
-detections do we misclassify the radii?}
-
-Ignoring binarity, we will detect $N_{\rm det,s}$ planets around single stars, 
-and $N_{\rm det,d}$ planets around double stars. The latter set will be assumed 
-to have radii $R_p \sqrt{\mathcal{D}}$.
-The fraction of detections with misclassified radii can then be written
-\begin{equation}
-\frac{N_{\rm det,d}}{N_{\rm det,s} + N_{\rm det,d}} = \frac{1}{1+\alpha},
-\end{equation}
-for
-\begin{equation}
-\alpha \equiv \frac{1}{2({\rm BF})} (1+\gamma_R)^{3/2} 
-\frac{\Gamma_{t,s}}{\Gamma_{t,d}}.
-\end{equation}
-
-For the nominal G2V dwarf case of ${\rm BF}=0.45$, twin binaries with equal 
-occurrence rates this produces a misclassification rate of $24\%$, in agreement 
-with Fig.~\ref{fig:snr_dist_analytic_v_numeric}.
-
-\subsubsection{Twin binaries: if we ignore binarity, how wrong is our
-occurrence rate for planets of radius $R_p$?}
-
-This is almost simply asking ``what is the relative difference between the 
-occurrence rates derived by Astronomers D and A for planets of radius $R_p$?''
-However, in the more realistic case, Astronomer A also has derived a 
-completeness, which we assume is the same as for Astronomer D in the single 
-star case.
-So Astronomer A now misclassifies planetary radii, and miscounts the total 
-number of stars, but knows his completeness for single stars.
-Astronomer D corrects all these errors.
-Then
-\begin{equation}
-\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d},
-\end{equation}
-and the relative difference between the two occurrence rates is
-
-\begin{align}
-\left|\frac{\Gamma_{{\rm A},R_p} - \Gamma_{{\rm D},R_p}}
-		   {\Gamma_{{\rm A},R_p}} \right| &=
-\left| 1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}}\right| \\
-&=
-\left| 1 - \left(\frac{\Gamma_{t,s}N_s + 2\Gamma_{t,d}N_d}{N_s + 2N_d}  \cdot 
-\frac{N_s + N_d}{\Gamma_{t,s} N_s }\right)\right| \\
-&= 
-\left|
-1 - \frac{(1 + 2 \beta \Gamma_{t,d}/\Gamma_{t,s})(1 + \beta)}{(1+2\beta)}
-\right|,
-\end{align}
-for
-\begin{equation}
-\beta \equiv N_d/N_s = {\rm BF} \times (1+\gamma_R)^{3/2}.
-\end{equation}
-
-For the nominal G2V dwarf case of ${\rm BF}=0.45$ with twin binaries ($\gamma_R 
-= 1$) and $\Gamma_{t,d} = \Gamma_{t,s}$ this gives a relative error of $127\%$.
-For instance, in the numerical simulation corresponding to 
-Fig.~\ref{fig:snr_dist_analytic_v_numeric}, Astronomer A finds $\Gamma_{\rm 
-A,R_p} = 0.22$, while Astronomer 
-D derives the true (input) occurrence rate of $\Gamma_{\rm D,R_p} = 0.5$.
-
-\section{Model \#2: fixed stars, fixed planets, varying light-ratio binaries}
-\label{sec:model_2}
-
-Same story as Model \#1, except now $\gamma_R$ varies across the population 
-of star systems.
-It does so because the underlying mass ratio varies. 
-Since we are interested in solar type binaries, we take the distribution of the 
-mass ratio $q=M_2/M_1$ by approximating [Rhagavan et al 2010, Fig 16]:
-\begin{equation}
-{\rm prob}(q) =
-\Bigg\{\begin{array}{lr}
-c_q & 0.1 < q \leq 1  \\
-0 & {\rm otherwise},
-\end{array}
-\label{eq:mass_ratio}
-\end{equation}
-for $c_q = 1/9$ to normalize the distribution to 1.
-The mass-luminosity relation can, for analytic convenience, be approximated as 
-$L = M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
-While we use this in subsequent analytic development, for numerics we fit a 
-lines to mass-luminosity data collected by Torres et al. [2009] for 
-dwarfs above M, and Benedict et al. [2016] for dwarfs below M.
-To convert Benedict et al. [2016]'s reported $M_V$ values to absolute 
-luminosities, we interpolated over E. Mamajek's 
-table\footnote{
-	\texttt{pas.rochester.edu/\textasciitilde 
-		emamajek/EEM\_dwarf\_UBVIJHK\_colors\_Teff.txt},
-	downloaded 2017.08.02}.
-In log-log space, we let the intersection point of the lines float, and make 
-various cuts on the data as indicated in Fig~\ref{fig:mass_luminosity}. The 
-fit parameters are available in a footnote\footnote{
-	m lo: 1.8818719873988132~--~
-	c lo: -0.9799647314108376~--~
-	m hi: 5.1540712426599882~--~
-	c hi: 0.0127626185389781~--~
-	M at merge: 0.4972991257826812~--~
-	L at merge: 0.0281260412126928.
-}.
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.9]{figures/mass_luminosity.pdf}
-	\end{center}
-	\caption{Empirical fit to main sequence dwarf mass luminosity data from 
-		Torres et al. [2009] and Benedict et al. [2016]. The ``low fit'' is a 
-		least 
-		squares fit to data from $0.1-0.5M_\odot$, and the ``high fit'' is to 
-		data 
-		above that, and below the Kraft break ($1.2M_\odot$).
-		The $L(M)$ relation taken for subsequent numerics is the maximum 
-		of the two fits.
-	}
-	\label{fig:mass_luminosity}
-\end{figure}
-
-The other changes between Model \#2 and Model \#1 are:
-\begin{enumerate}
-\item We allow a third occurrence rate, splitting $\Gamma_{t,d} \rightarrow 
-(\Gamma_{t,d1}, \Gamma_{t,d2})$ where the latter terms represent the occurrence 
-rate around the primary of a double star system, and the occurrence rate around 
-the secondary of a double star system. For the case of \textit{e.g.}, $q=0.1$, 
-this seems more relevant.
-We can then take the limit that says ``any planet around the secondary is as 
-likely as around the primary'', or go to the opposite extreme of ``there are 
-only planets around the primaries, not the secondaries''.
-%
-\item Our binary systems have secondaries with varying masses. Thus they have 
-varying radii. Taking [Demircan \& Kahraman 1991]'s empirical fit to eclipsing 
-binary data,
-\begin{equation}
-R = 1.06 M^{0.945}
-\label{eq:mass_radius}
-\end{equation}
-for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 1991's 
-stated bounds), and for each in solar units. This only affects the transit 
-probabilities about the secondaries of binary star systems.
-\end{enumerate}
-
-We then ask the same questions as for Model \#1.
-
-\subsection{How many stars are in the sample?}
-
-Let $N_s$ be the number of single star systems, and $N_d$ the number of double 
-star systems. $N_s$ is the same as in Sec.~\ref{sec:model_1}.
-
-Now
-\begin{equation}
-N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3(L_1^N, \gamma_R, F_{\rm 
-	lim}^N)
-\end{equation}
-for
-\begin{equation}
-d_{\rm max,d}(L_1^N, \gamma_R, F_{\rm lim}^N) =
-d_{\rm max,s}(L_1^N, F_{\rm lim}^N)\times (1+\gamma_R)^{1/2}.
-\end{equation}
-Since $q$ is a random variable, $\gamma_R$ is a random variable, and $d_{\rm 
-	max,d}$ is a random variable. Since $N_d$ is a function of $d_{\rm max,d}$, 
-the number of double star systems becomes a random variable.
-The expected (mean) number of double star systems in the sample is
-\begin{equation}
-\langle N_d \rangle = \int_{0}^{\infty} N_d\, {\rm prob}(N_d) \,{\rm d}N_d.
-\end{equation}
-By applying the chain rule for probability density functions, the distribution 
-${\rm prob}(N_d)$ can be written 
-\begin{equation}
-{\rm prob}(N_d) = {\rm prob}(q(\gamma_R)) 
-\left| \frac{{\rm d}q}{{\rm d}\gamma_R}  \right|
-\left| \frac{{\rm d}\gamma_R}{{\rm d}d_{\rm max,d}}  \right|				
-\left| \frac{{\rm d}d_{\rm max,d}}{{\rm d}N_d}  \right|.
-\end{equation}
-Doing some algebra, and assuming $\gamma_R = q^3$, this can be shown to be
-\begin{align}
-{\rm prob}(N_d) = \frac{2}{81} &N_d^{-1/3} \left(\frac{3}{4\pi 
-	n_d}\right)^{2/3} d_{\rm max,s}^{-2/3} \ \times \nonumber\\
-&\left[ \left(\frac{3 N_d}{4\pi n_d}\right)^{2/3} - d_{\rm 
-	max,s}^2 \right]^{-2/3}
-\end{align}
-over the interval
-\begin{equation}
-{\rm lower\ bound} = \frac{4 \pi n_d}{3} (\sqrt{0.1^3 + 1} d_{\rm max,s})^3
-\end{equation}
-\begin{equation}
-{\rm upper\ bound} = \frac{4 \pi n_d}{3} (\sqrt{2}d_{\rm max,s})^3,
-\end{equation}
-and outside the stated interval ${\rm prob}(N_d) = 0$.
-
-While a closed analytic expression for $\langle N_d \rangle$ does exist, it is 
-messy, and it does not yield much intuition. Instead, summarizing the important 
-points and supporting them numerically:
-\begin{itemize}
-	\item In a volume-limited sample of binary star systems in which the 
-	primary mass is fixed, and the mass ratio is drawn from a bounded uniform 
-	distribution, the distribution of $\gamma_R$ will be biased towards low 
-	values ($\gamma_R \approx 0.1$). This is shown in 
-	Fig.~\ref{fig:gammaR_distribn_vol_limited}.
-	\item In a magnitude-limited sample of binary star systems in which the 
-	primary mass is fixed, and the mass ratio of the \textit{population} is 
-	drawn from a bounded uniform distribution, the observed distribution of 
-	mass ratios will be biased towards high values. You will see more twins,
-	because they are detectable out to a greater distance\footnote{This is a 
-	scarcy systematic w.r.t. the claimed intrinsic excess of twin binaries.}. 
-	This is shown in 
-	Fig.~\ref{fig:q_distribn_mag_limited}.
-	\item In a magnitude-limited sample of binary star systems in which the 
-	primary mass is fixed, the distribution of $\gamma_R$ will be biased 
-	towards low values ($\approx 0.1$), but less so than in a 
-	volume-limited sample. This is shown in 
-	Fig.~\ref{fig:gammaR_distribn_mag_limited}.
-\end{itemize}
-
-In passing, the bias in ${\rm prob}(\gamma_R)$ towards low luminosity ratios 
-can be seen analytically. If we assume $\gamma_R = q^\alpha$, then
-\begin{equation}
-{\rm prob}(\gamma_R) = \frac{1}{9 \alpha} \ \gamma_R^{\frac{1-\alpha}{\alpha}}
-\quad {\rm for\ } (0.1)^\alpha < \gamma_R < 1,
-\end{equation}
-and otherwise zero. For instance if $\alpha = 3$, ${\rm prob}(\gamma_R) 
-\propto \gamma_R^{-2/3}$ and the domain extends from $1$ to 
-$10^{-3}$, where the probability distribution peaks.
-
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/gammaR_distribn_vol_limited.pdf}
-	\end{center}
-	\caption{The distribution of the luminosity ratio for a volume limited 
-		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
-		total 
-		number density from Bovy 2017; $M(L)$ relation from 
-		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
-		Eq.~\ref{eq:mass_ratio}.}
-	\label{fig:gammaR_distribn_vol_limited}
-\end{figure}
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/q_distribn_mag_limited.pdf}
-	\end{center}
-	\caption{The distribution of the mass ratio for a magnitude limited 
-		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
-		total 
-		number density from Bovy 2017; $M(L)$ relation from 
-		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
-		Eq.~\ref{eq:mass_ratio} -- a \textit{uniform} distribution in a 
-		volume-limited sample!}
-	\label{fig:q_distribn_mag_limited}
-\end{figure}
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/gammaR_distribn_mag_limited.pdf}
-	\end{center}
-	\caption{The distribution of the luminosity ratio for a magnitude limited 
-		sample of binary stars. ${\rm BF} = 0.45$ [Duchene and Kraus 2013]; 
-		total 
-		number density from Bovy 2017; $M(L)$ relation from 
-		Fig.~\ref{fig:mass_luminosity}; mass ratios are drawn from 
-		Eq.~\ref{eq:mass_ratio}.}
-	\label{fig:gammaR_distribn_mag_limited}
-\end{figure}
-
-\paragraph{How many stars are in the sample?}
-
-We return to the original question: how many stars?
-\begin{equation}
-\langle N_{\rm stars} \rangle = N_s + 2 \langle N_d \rangle,
-\end{equation}
-as before.
-Since we are not giving analytic expressions for either of the $\langle \ldots 
-\rangle$ terms, we just note that in the case of a population with a single 
-$\gamma_R$ value, $N_d/N_s = {\rm BF} \times (1 + \gamma_R)^{3/2}$. A lower
-bound for this ratio will always be the binary fraction.
-For a single run of a Monte Carlo code, where ${\rm mean}(\gamma_R) = 0.28$, 
-${\rm median}(\gamma_R) = 0.13$, and the distribution was that given in 
-Fig.~\ref{fig:gammaR_distribn_mag_limited}, the observed ratio was $N_d/N_s = 
-0.59$.
-This corresponds to a single-valued population with $\gamma_R \sim 0.20$, a
-number between the mean and median of the true distribution.
-It means $\sim 1.2$ stars in binary star systems for every star in a single 
-star system from this sample.
-
-We also note in passing that $N_d/N_s = 0.59$ is a smaller ratio than for the 
-case of a fixed $\gamma_R$ population with $\gamma_R = 1$, which gave $N_d/N_s 
-= 1.27$. The difference is that in the latter population, the volume of 
-searchable stars is greater. Based on the fixed-$\gamma_R$ population scaling 
-law $N_d / N_s \propto (1+\gamma_R)^{3/2}$, we 
-would expect the ratio of single stars to be $\sim (2/1.1)^{3/2} = 2.4$ 
-times smaller, which is roughly (though not exactly) what we observe.
-
-\subsection{How many planets are in the sample?}
-
-The mean number of planets in the sample is
-\begin{align}
-\langle N_{\rm planets} \rangle
-=& N_{\rm planets\ in\ single\ star\ systems}  +  \\
-&\quad\quad N_{\rm planets\ in\ double\ star\ systems} 
-\nonumber \\
-=& \Gamma_{t,s} N_s + (\Gamma_{t,d1} + \Gamma_{t,d2}) \langle N_d \rangle.
-\end{align}
-
-What was formerly a factor of 2 has now been split into the true occurrence 
-rates about the primaries and secondaries of double star systems.
-
-\subsection{What is the true occurrence rate?}
-
-The ``true occurrence rate'' is the average number of planets per star. Thus
-
-\begin{align}
-\Gamma_t &= \frac{\langle N_{\rm planets}\rangle}{\langle N_{\rm stars} 
-\rangle} \\
-\Gamma_t &= \frac{\Gamma_{t,s} N_s + (\Gamma_{t,d1} + \Gamma_{t,d2}) \langle 
-N_d\rangle}{N_s +	2\langle N_d\rangle}.
-\label{eq:true_occ_2}
-\end{align}
-
-\subsection{How many planets are detected?}
-
-The total number of planet detections is the sum of the number of planets 
-detected in single star systems $N_{{\rm det},s}$ and the number of planets 
-detected in double star systems $N_{{\rm det},d}$.
-The latter of these is a random variable, and can be further split into the 
-primary and secondary contributions $N_{{\rm det},d1}$, $N_{{\rm det},d2}$.
-
-The number of planets detected in single star systems is
-\begin{equation}
-N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,g} f_{s,c},
-\end{equation}
-where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
-star systems of the sample, $f_{s,g}$ is the geometric probability of the 
-planets transiting, and $f_{s,c}$ is the fraction of transiting planets around 
-single star systems that are detected (the completeness).
-
-The mean number of planets detected in double star systems is
-\begin{equation}
-\langle N_{{\rm det},d}\rangle = \langle N_d\rangle  
-(\Gamma_{t,d1} f_{d1,g} f_{d1,c} + 
-\Gamma_{t,d2} f_{d2,g} f_{d2,c} ).
-\label{eq:N_det_d_2}
-\end{equation}
-Now $\langle N_d\rangle  \Gamma_{t,d1}$ is the mean number of planets orbiting 
-the primaries of double star systems, ditto the corresponding expression for 
-the secondaries.
-For this population, $f_{d1,g} = f_{s,g}$, but $f_{d1,g} \neq f_{d2,g}$.
-The completeness fractions differ between the primary and secondary because 
-the dilution as a function of the light ratio, $\mathcal{D}(\gamma_R)$, has 
-different behavior for the two cases.
-We show this in Fig.~\ref{fig:rad_vs_gammaR}.
-Specifically, a planet orbiting the secondary will have worse dilution than a 
-planet orbiting the primary, and consequently lower completeness.
-
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.9]{figures/observed_radii_vs_gammaR.pdf}
-	\end{center}
-	\caption{Dilution $\mathcal{D}$ and observed to true radius ratio 
-	$\sqrt{\mathcal{D}}$ plotted against binary light ratio $\gamma_R$. This 
-	illustrates	Eq.~\ref{eq:dilution}. The lower panel's horizontal line is at 
-	$1/\sqrt{2}$.
-	The implication is that if a detected planet orbits the secondary and you do 
-	not know it, you cannot measure $R_p$ to better than $\approx 29\%$ 
-	accuracy.
-	Conversely, if it orbits the primary and you do not know it, you will get 
-	the radius wrong by at most $\approx 29\%$.
-	}
-	\label{fig:rad_vs_gammaR}
-\end{figure}
-
-We could probably derive analytic expressions for all the terms between 
-brackets in Eq.~\ref{eq:N_det_d_2}, completeness included.
-I'm not yet convinced we need to -- we should get cancellations for
-most of the questions we want to ask.
-
-\subsubsection{Analytic completeness}
-\texttt{todo, if necessary}
-\subsubsection{Deriving ${\rm prob}(x_i)$}
-\texttt{todo, if necessary}
-\subsubsection{Number of detected planets}
-\texttt{todo, if necessary}
-
-\subsection{Astronomer A ignores binarity}
-
-Just as in Sec.~\ref{sec:model_1}, Astronomer A
-\begin{itemize}
-\item has never heard about binary star systems.
-\item has never heard about completeness corrections.
-\item knows about geometric transit probabilities.
-\end{itemize}
-
-What occurrence rate does he derive for planets of radius $R_p$ and period $P$?
-
-The answer is the same as in Sec.~\ref{sec:model_1},
-\begin{equation}
-\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/f_{s,g}}{N_s + N_d}.
-\end{equation}
-
-However, now rather than planets detected in binary systems being perceived as 
-a second population of planets with fixed radius $R_p \sqrt{\mathcal{D}}$, 
-there will be an apparent spectrum of diluted radii, since $\mathcal{D}$ 
-varies by system.
-Writing the number of detections in double systems as a function of observed 
-radius, $N_{\rm det,d}(R_p^{\rm obs})$, Astronomer A would derive an 
-occurrence rate for this separate population of
-\begin{equation}
-\Gamma_{\rm A}(R_p^{\rm obs}) = \frac{N_{\rm det,d}(R_p^{\rm 
-obs})/f_{s,g}}{N_s + N_d}.
-\end{equation}
-Note that this astronomer thinks these planets orbit single stars, 
-and so miscomputes the geometric transit probability for the secondaries.
-
-A more realistic question is ``what fraction of multiples have $\gamma_R$ 
-smaller enough so that you'd still observe a radius close to the true one?''. 
-We discuss this in Sec.~\ref{sec:close_enough}.
-
-\subsection{Astronomer B counts host stars correctly}
-
-Same as astronomer A above, except the denominator becomes $N_s + 2N_d$.
-
-\subsection{Astronomer C counts host stars correctly and figures out diluted
-radii}
-
-In Sec.~\ref{sec:model_1}, Astronomer C only had to do high resolution 
-imaging to measure $\gamma_R$ for each system. Since for $\gamma_R = 1$, the 
-dilution is single-valued (Fig.~\ref{fig:rad_vs_gammaR}), this immediately 
-told her the true planet radius (since the primary/secondary distinction was 
-meaningless).
-
-In Model \#2's universe, the dilution is double-valued for a given $\gamma_R$.
-To correctly work out the diluted radii, Astronomer C now needs more than high 
-resolution imaging -- she needs to resolve the stars during transit to 
-discover which star the planet orbits.
-She also needs to derive masses for the primaries and secondaries (to make the 
-geometric transit probability correction).
-
-Only after doing this does she find that all detected planets from this survey 
-have radii $R_p$. She computes an occurrence rate
-\begin{equation}
-	\Gamma_{{\rm C},R_p} = 
-	\frac
-			{N_{\rm det,s}/f_{s,g} +
-			  N_{\rm det,d1}/f_{d1,g} +
-			  N_{\rm 	det,d2}/f_{d2,g}}
-			{N_s + 2N_d}
-\end{equation}
-
-the closest yet to the true rate (Sec.~\ref{sec:true_rate}).
-
-\subsection{Astronomer D counts host stars correctly, figures out diluted
-radii, and accounts for completeness}
-
-Same story as for Model \#1.
-Now, they do injection recovery and derive correct estimates for their 
-completeness functions about single stars $f_{s,c}$, the primaries of double 
-star systems, $f_{d1,c}$, and the secondaries of double star systems 
-$f_{d2,c}$.
-They compute the respective single and binary 
-occurrence rates
-\begin{equation}
-\Gamma_{t,s} = \frac{N_{\rm det,s}}{N_s f_{s,g} f_{s,c}},
-\end{equation}
-\begin{equation}
-\Gamma_{t,d1} = \frac{N_{\rm det,d1}}{N_d f_{d1,g} f_{d1,c}},
-\end{equation}
-\begin{equation}
-\Gamma_{t,d2} = \frac{N_{\rm det,d2}}{N_d f_{d2,g} f_{d2,c}}.
-\end{equation}
-
-With these in hand, they derive the overall occurrence rate
-\begin{align}
-\Gamma_{{\rm D}, R_p} 
-&= \frac{
-	N_{\rm det,s}/f_s + 
-	N_{\rm det,d1}/f_{d1} +
-	N_{\rm det,d2}/f_{d2}
-	}
-	{N_s + 2 N_d} \\
-&= \frac{\Gamma_{t,s} N_s + (\Gamma_{t,d1} + \Gamma_{t,d2})N_d}
-				{N_s + 2 N_d},
-\end{align}
-where $f_s \equiv f_{s,g}f_{s,c}, f_{d1} \equiv f_{d1,g}f_{d1,c}, 
-f_{d2} \equiv f_{d2,g}f_{d2,c}$ for brevity.
-Astronomer D's occurrence rate is the true occurrence rate (cf. 
-Eq.~\ref{eq:true_occ_2}).
-
-\subsection{Numerical verification}
-\texttt{todo, if needed}
-
-\subsection{Representative numbers for a few cases}
-
-We have not given analytic expressions for the number of double star systems 
-$N_d$ because even with simplified $M(L)$ relations, they are unwieldy.
-However, we can ask similar numerical questions as from Sec.~\ref{sec:model_1}, 
-and compare the answers.
-
-\subsubsection{If we ignore binarity, for what fraction of detections do we
-misclassify the radii?}
-
-Ignoring binarity, we will detect $N_{\rm det,s}$ planets around single stars, 
-and $N_{\rm det,d}$ planets around double stars. The latter set will be assumed 
-to have radii $R_p \sqrt{\mathcal{D}}$, where $\mathcal{D}$ is the vector of 
-dilutions appropriate for each binary star system.
-The fraction of detections with misclassified radii can then be written
-\begin{equation}
-\frac{N_{\rm det,d}}{N_{\rm det,s} + N_{\rm det,d}} = \frac{1}{1+\alpha},
-\end{equation}
-for
-\begin{equation}
-\alpha \equiv 
-\frac{ N_d (\Gamma_{t,d1} f_{d1} + \Gamma_{t,d2} f_{d2}) }{N_s \Gamma_{t,s} 
-f_s},
-\end{equation}
-where $f_s \equiv f_{s,g}f_{s,c}, f_{d1} \equiv f_{d1,g}f_{d1,c}, 
-f_{d2} \equiv f_{d2,g}f_{d2,c}$ for brevity.
-
-% For the nominal G2V dwarf case of ${\rm BF}=0.45$, and the same distributions 
-% used to generate 
-% Figs.~\ref{fig:gammaR_distribn_vol_limited}~--~\ref{fig:gammaR_distribn_mag_limited},
-% we found $N_d/N_s=0.59$.
-% We are assuming $\Gamma_{t,d} / \Gamma_{t,s} = 1$.
-% If we maintained the (incorrect) assumption that $f_{d,{\rm geom}} / f_{s,{\rm 
-% 		geom}} = 1$, this would give a misclassification rate of 45\%. 
-% 
-% This latter assumption is wrong because although we are letting the occurrence 
-% rate be the same across binary and single systems, our binary systems have 
-% secondaries with varying masses. Thus they have varying radii.
-% Taking [Demircan \& Kahraman 1991]'s empirical fit to eclipsing binary data,
-% \begin{equation}
-% R = 1.06 M^{0.945}
-% \label{eq:mass_radius}
-% \end{equation}
-% for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 1991's 
-% stated bounds).
-% For a zero eccentricity system, $f_{\rm geom}=R_\star/a$.
-% Since $f_{d,{\rm geom}} = 0.5\times (f_{d1,{\rm geom}} + f_{d2,{\rm geom}})$, 
-% \textit{i.e.} it is the average of the transit probabilities for both the 
-% primary and secondary (d1 and d2), we can write (dropping the ``geom'' 
-% subscript)
-% \begin{align}
-% \frac{f_d}{f_s} &= \frac{1}{2} \left(1 + \left(\frac{M_1}{M_{d2}}\right)^{1/3} 
-% \frac{R_{d2}}{R_1}\right) \\
-% &= \frac{1}{2} \left( 1 + q^{-1/3 + 0.945}\right).
-% \end{align}
-% 
-% Computing it all out in a numerical simulation, the only thing that changes 
-% analytically is that
-% \begin{equation}
-% N_d f_d \rightarrow \sum_{i=1}^{N_d} f_{d,i},
-% \end{equation}
-% so when considering the ratio $f_d/f_s$, we are in fact interested in
-% \begin{equation}
-% \frac{\langle f_d \rangle}{f_s} = \frac{\frac{1}{N_d} \sum_{i=1}^{N_d} 
-% 	f_{d,i}}{f_s},
-% \end{equation}
-% where $f_{d,i}$ is the vector of geometric transit probabilities across all 
-% systems.
-% For the same numerical realization with $N_d/N_s = 0.59$, and using the mass 
-% radius relation given by Eq.~\ref{eq:mass_radius}, I get $\langle f_d 
-% \rangle / f_s = 0.84$ (smaller secondaries lead to a smaller average transit 
-% probability in double star systems), for a radius misclassification rate of 
-% $0.51$.
-
-\texttt{The different secondary completeness matters here if proceeding 
-analytically. Would be easier to do numerically.
-Regardless, this is not the most interesting question right now.}
-
-\subsubsection{If we ignore binarity, how wrong is our occurrence rate for
-planets of radius $R_p$?}
-
-Write $\Gamma_{{\rm A,\ planets\ of\ }R_p} = \Gamma_{{\rm A,}R_p}$, 
-and similarly for ${\rm D}$.
-Just as in Model \#1, the answer to ``what is the relative difference between 
-the occurrence rates derived by Astronomers D and A for planets of radius 
-$R_p$?''
-is simpler when we assume that Astronomer A has also derived a 
-completeness, which we assume is the same as for Astronomer D in the single 
-star case.
-So Astronomer A now misclassifies planetary radii, and miscounts the total 
-number of stars, but knows his completeness for single stars.
-Astronomer D corrects all these errors.
-Then
-\begin{equation}
-	\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d}.
-\end{equation}
-
-The relative difference between the two occurrence rates, $\Delta \equiv 
-\left|(\Gamma_{{\rm A},R_p} - \Gamma_{{\rm D},R_p})/\Gamma_{{\rm A},R_p} 
-\right|$, is
-\begin{align}
-	\Delta &=
-	\left| 1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}}\right| \\
-	&=
-	\left| 1 - \left(\frac{\Gamma_{t,s}N_s + (\Gamma_{t,d1} + 
-	\Gamma_{t,d2})N_d}{N_s + 2N_d}  
-	\cdot 
-	\frac{N_s + N_d}{\Gamma_{t,s} N_s }\right)\right| \\
-	&= 
-	\left|
-	1 - \frac{(1 + \beta (\Gamma_{t,d1}+\Gamma_{t,d2})/\Gamma_{t,s})(1 + 
-	\beta)}{(1+2\beta)}
-	\right| \\
-	&=
-	\left| 1 - \left(
-	\left[ 1 + \chi \beta \right]
-	\cdot 
-	\frac{1 + \beta}{1 + 2\beta }
-	\right) \right|
-\end{align}
-for
-\begin{equation}
-\beta \equiv N_d/N_s, \quad \chi \equiv 
-(\Gamma_{t,d1}+\Gamma_{t,d2})/\Gamma_{t,s}.
-\end{equation}
-
-In the limit where the $R_p$ planet has the same occurrence rate about every 
-host type regardless of its mass, $\chi=2$, and this simplifies to $\Delta = 
-\beta$, for a relative error of $59\%$.
-This is less than for Model \#1 because of the smaller volume of searchable 
-binary stars in Model \#2.
-
-If we consider the opposite limit of the $R_p$ planet \textit{only} 
-existing around single stars and the primaries of double star systems with 
-equal occurrence, $\chi = 1$, and $\Delta = \beta^2 / (1+2\beta)$.
-For $\beta = 0.59$, this gives $\Delta = 0.16$.
-
-\subsubsection{What if we ignore binarity, but count derived planet radii that 
-are ``close enough''?}
-\label{sec:close_enough}
-
-Let $x$ be the acceptable margin of error on the observed radius, so that we 
-are interested in counting the number of planets detected with $R_p < R_p^{\rm 
-obs} < (1\pm x) R_p$.
-For Model \#2, since all planets have radius $R_p$, we only care about 
-the smaller radius case, \textit{i.e.}, we want to count $R_p < R_p^{\rm 
-	obs} < (1 - x) R_p$, for instance with $x=0.1$.
-
-All of these diluted detections will be from binary systems. The number of 
-planets detected in double star systems with observed radius $R_p^{\rm obs}$ 
-is
-\begin{align}
-N_{\rm det,d}(R_p^{\rm obs}) &= N_{\rm det, d1}(R_p^{\rm obs}) + 
-															  N_{\rm det, d2}(R_p^{\rm obs}) \\
-&= \Gamma_{t,d1} f_{d1}(R_p^{\rm obs}) N_{d1}(R_p^{\rm obs})\  + \nonumber \\
-&\quad\quad \quad\Gamma_{t,d2} f_{d2}(R_p^{\rm obs}) N_{d2}(R_p^{\rm obs}).
-\end{align}
-As always, the fractional terms $f_{di}$ for $i\in{1,2}$ encompass geometric 
-and completeness corrections.
-We're writing both them and the number of stars for which a given observed 
-radius would be seen as functions of $R_p^{\rm obs}$.
-
-To save on notational horribleness, write $R_p' \equiv \{(1-x) R_p < R_p^{\rm 
-	obs} < R_p\}$, in other words let $R_p'$ denote the set of planets 
-	that are observed in the desired radius range.
-
-Then writing $N_{d1}(R_p')$ is relatively straight-forward:
-\begin{equation}
-N_{d1}(R_p') = \int_{0}^{{\rm min}(1,\gamma_{R,u})} N_d(\gamma_R) 
-\,{\rm prob}(\gamma_R) \,{\rm d}(\gamma_R),
-\end{equation}
-where the upper limit is set by equating the limiting radius $(1-x)R_p$ to the 
-observed diluted radius $(1+\gamma_R)^{-1/2}R_p$, giving
-\begin{equation}
-\gamma_{R,u} \equiv \frac{x (2-x)}{(1-x)^2}.
-\end{equation}
-If $x > (1 - 2^{-1/2})$, bad things happen (cf. Fig.~\ref{fig:rad_vs_gammaR}), 
-hence the need for the minimum.
-
-The analogous equation for secondaries is
-\begin{equation}
-N_{d2}(R_p') = 
-  \Bigg\{\begin{array}{lr}
-  \int_{\gamma_{R,l}}^{1} N_d(\gamma_R)
-  \,{\rm prob}(\gamma_R) \,{\rm d}(\gamma_R),\\ 
-  \quad\quad\quad\quad\quad\quad\quad\quad x > 1 - 
-  2^{-1/2}\\
-  0, \quad\quad\quad\quad\quad\quad\quad x \leq 1 - 2^{-1/2}
-  \end{array}
-\label{eq:N_Rp_secondary}
-\end{equation}
-where
-\begin{equation}
-\gamma_{R,l} \equiv \frac{(1-x)^2}{x (2-x)}.
-\end{equation}
-Eq.~\ref{eq:N_Rp_secondary} accounts for the fact that you cannot measure 
-$R_p$ to better than $\approx 29\%$ if the planet orbits the secondary of an 
-unidentified binary.
-This was originally noted in Fig.~\ref{fig:rad_vs_gammaR}.
-
-If we were forward-modelling (\textit{i.e.} we actually wanted to compute 
-$N_{\rm det,d}(R_p')$) we would need to evaluate the completeness 
-term and this ``fraction of slightly diluted stars'' terms together, and 
-integrate jointly to find $N_{\rm det,d1}(R_p')$ and $N_{\rm det,d2}(R_p')$.
-
-\texttt{If we ask the right question, this will not be necessary.}
-
-Astronomer A misclassifies planetary radii, miscounts the total number of 
-stars, but knows his completeness and geometric correction for single stars as 
-a function of radius $f_s(R_p')$.
-His occurrence rate for planets of radius \textit{near} $R_p$, \textit{i.e.} 
-the set of planets with $\{ (1-x)R_p < R_p^{\rm obs} < R_p \}$, is
-\begin{equation}
-\Gamma_{{\rm A}, R_p'} = \frac{ (N_{\rm det,s} + N_{\rm det,d}(R_p')) 
-/ f_s(R_p')   }{ N_s + N_d}.
-\end{equation}
-
-For him,
-\begin{equation}
-f_s(R_p') = f_{s,g}(R_p') f_{s,c}(R_p').
-\end{equation}
-
-Astronomer D, who had everything right, doesn't change anything.
-
-The new difference between the two rates is
-\begin{align}
-\Delta' &\equiv \left|
-\frac{\Gamma_{{\rm A},R_p'} - \Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p'} }
-\right| \\
-&= \left|
-1 - \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p'} } 
-\right| \\
-&=
-\left| 1 -
-\left( \frac{1+\chi \beta}{1 + \xi}  \cdot 
-\frac{1 + \beta}{1 + 2\beta} \right)
-\right|,
-\end{align}
-for
-\begin{equation}
-\beta \equiv N_d/N_s, \quad \chi \equiv (\Gamma_{t,d1} + 
-\Gamma_{t,d2})/\Gamma_{t,s},
-\end{equation}
-and
-\begin{equation}
-\xi \equiv \frac{N_{\rm det,d}(R_p')}{f_s(R_p')} \cdot \frac{1}{N_s 
-\Gamma_{t,s}}.
-\end{equation}
-If we let $x < 1 - 2^{-1/2}$, for instance $x=0.1$, then $N_{d2}(R_p') = 0$. 
-The expression for $\xi$ simplifies to
-\begin{equation}
-\chi = \frac{N_{d1}(R_p')}{N_s} \cdot \frac{\Gamma_{t,d1}}{\Gamma_{t,s}} \cdot 
-\frac{f_{d1}(R_p')}{f_s(R_p')},\quad ({\rm if\ }x < 1-2^{-1/2}).
-\label{eq:chi_simplfied}
-\end{equation}
-
-If we assume $\Gamma_{t,d1} = \Gamma_{t,s}$, as we were doing 
-anyway, Eq.~\ref{eq:chi_simplfied} is bounded above by
-$\beta$. This is because $N_{d1}(R_p') \leq N_{d1} = N_d$, and $f_{d1}(R_p') / 
-f_s(R_p') \leq 1$, since over a fixed radius interval dilution will always 
-yield a lower completeness fraction for double star systems than singles.
-Recall for the fixed $\gamma_R$ population we showed that $f_{d1}/f_s = 
-(1+\gamma_R)^{-3}$.
-
-Regardless, the cheap way of making the estimate without deriving the actual 
-completeness fractions, or even running numerics, is just to plot the error as 
-a function of $\xi$. We do this in Fig.~\ref{fig:deltaprime_vs_xi}, which 
-shows that the
-relative error varies between $\approx 5\% - 35\%$, depending on what 
-fraction of secondaries you assume host planets.
-
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/deltaprime_vs_xi.pdf}
-	\end{center}
-	\caption{\textit{Y axis}: relative error between the Astronomer who ignores 
-	binarity 
-	but counts planetary radii that are ``close enough'' (A') and the Astronomer 
-	who knows everything (D).
-	\textit{X axis}: $\chi$ as defined in Eq.~\ref{eq:chi_simplfied}. Taking a 
-	guess at the true value, if $N_{d1}(R_p') = N_d/2$, and the occurrence rates 
-	between `d1' and `s' are the same, and we put $\gamma_R = 0.1$ into the 
-	completeness ratio, then $\xi = (\beta/2) \times (1+0.1)^{-3} = 0.22$. The 
-	relative error then varies between $\approx 5\% - 35\%$, depending on what 
-	fraction of secondaries you assume host planets.
-		}
-	\label{fig:deltaprime_vs_xi}
-\end{figure}
-
-%FIXME
-
-% \begin{comment}
-% \subsection{How many stars are in the sample?}
-% \subsection{How many planets are in the sample?}
-% \subsection{What is the true occurrence rate?}
-% \subsection{How many planets are detected?}
-% \subsubsection{Analytic completeness}
-% \subsubsection{Deriving ${\rm prob}(x_i)$}
-% \subsubsection{Number of detected planets}
-% \subsection{Astronomer A ignores binarity}
-% \subsection{Astronomer B counts host stars correctly}
-% \subsection{Astronomer C counts host stars correctly and figures out diluted
-% radii}
-% \subsection{Astronomer D counts host stars correctly, figures out diluted
-% radii, and accounts for completeness}
-% \subsection{Numerical verification}
-% \subsection{Representative numbers for a few cases}
-% \subsubsection{If we ignore binarity, for what fraction of detections do we
-% misclassify the radii?}
-% \subsubsection{If we ignore binarity, how wrong is our occurrence rate for
-% planets of radius $R_p$?}
-% \subsubsection{If we ignore binarity, but allow derived planet radii to be
-% ``close enough'', how do we do?}
-% 
-% \end{comment}
-
-\newpage
-
-%\begin{thebibliography}{}
-
-%\end{thebibliography}
+\input{generalities}
+\input{model_1}
+\input{model_2}
 

### 2017-10-19 paper/model_2.tex
moving things around

--- 2017-10-11 memos/model_2.tex
@@
 The occurrence rate corresponding to Eq.~\ref{eq:model2_rate_density}'s rate 
-density is
-\begin{equation}
-\Lambda = \sum_i
-\left[
-\left(
-    w_i \Lambda_i \int_{\Omega_{\rm desired}} p_i(\vec{x}) \,{\rm d}\Omega
-\right)
-\cdot
-\left(
-    w_i \Lambda_i \int_{\Omega_{\rm total}} p_i(\vec{x}) \,{\rm d}\Omega
-\right)^{-1}
-\right]
-\end{equation}
-for a desired volume of phase space $\Omega_{\rm desired}$.
+density for a desired volume of phase space $\Omega_{\rm desired}$ is given by 
+Eq.~\ref{eq:occ_rate}.
 Specifying the desired mass ratios of interest as $q_{\rm min} < q < q_{\rm 

### 2017-10-20 paper/analytics_v_numerics.tex
update paper

\subsection{Numerical Approach}
\label{sec:numerical_methods}

To progressively gain intuition, we study the following idealized transit 
surveys:
\begin{itemize}
    \item Model \#1: fixed stars, fixed planets, twin binaries;
    \item Model \#2: fixed planets and primaries, varying secondaries;
    \item Model \#3: fixed primaries, varying planets and secondaries.
\end{itemize}

Assuming a signal-to-noise limited survey, we would like to find the true
occurrence rate density, and also that inferred by an observer ignoring 
binarity.
Though analytic or semi-analytic solutions exist for each model, beyond Model 
\#2 the equations become burdensome. To preserve simplicity, we develop a 
Monte Carlo approach, which works as follows.

First, the user specifies the model class (\#1, 2, or 3) and various free 
parameters describing the stellar and planetary population.
Most importantly, these parameters include the binary fraction, and the true 
planet occurrence rates around single stars, primaries, and secondaries.

We then generate the population of selected stars.
Each selected star has a type (single, primary, secondary), 
a binary mass ratio (if it is not single), and the property of whether it 
is ``searchable''.               
The absolute number of stars is arbitrary.
The relative number of binaries to singles is calculated according to analytic
formulae. The binary mass ratios are sampled from the
appropriate magnitude-limited distributions.

Whether a star is ``searchable'' depends entirely on its ``completeness''
fraction. By ``completeness'', we mean the ratio of the actual number of 
searchable stars to the assumed number of searchable stars (for a given planet 
size, period, etc.).
Assuming homogeneously distributed stars, we will show 
(Sec.~\ref{sec:model_1}) that this is equivalent to the
ratio of the searchable to selected volumes.
For single stars, we assume these volumes are identical~--~exactly the case 
discussed by Pepper et al. (2003).
For binaries, this volume ratio is a function of only the binary mass ratio.

To assign planets, each selected star receives a planet at the initially 
specified rate, according to its type.
The radii of planets are assigned independently of any host system
property, as sampled from $p_r(r) \sim r^\delta$ for Model \#3 or a 
$\delta$--function 
for Models \#1 and \#2.
A planet is detected when a) it transits, and b) its host star is
searchable.

The probability of transiting single stars in our model is assumed to be known,
and so it is mostly corrected by the observer attempting to infer an
occurrence rate. The only bias is for secondaries, which can be smaller than 
primaries in Models \#2 and \#3.
This effect is included when computing the transit probability.

For detected planets, apparent radii are computed according to analytic
formulae that account for both dilution and the misclassification of stellar
radii. We assume that the observer thinks all transits are around primaries.

The rates are then computed in bins of true planet radius and apparent planet
radius.
In a given radius bin, the true rate is found by counting the number of planets
that exist around selected stars of all types (singles, primaries,
secondaries), and dividing by the total number of these stars.
The apparent rates are found by counting the number of detected planets that
were found in an apparent radius bin, dividing by the geometric transit
probability for single stars, and dividing by the apparent total number of
stars.

The simplest realization of this scheme is described analytically in 
Sec.~\ref{sec:model_1}, but we first clarify our terminology.

### 2017-10-20 paper/figures.tex
update paper

\begin{figure}
    \begin{center}
        \includegraphics[width=\textwidth]{figures/errcases_rate_density_vs_radius_model_1_brokenx.pdf}
    \end{center}
    \caption{
    Inferred planet occurrence rates as a function of planet radius in Model 
    \#1.
    This model has fixed stars, fixed planets, twin binaries;
    Note the break in the $x$ axis.
    If the true planet radius is $r_p$, all planets 
    detected in binaries will have apparent radii $r_a = r_p/\sqrt{2}$.
    Only 1 in 8 selected binaries is actually searchable (see 
    Sec.~\ref{sec:model_1}).
    }
    \label{fig:errcases_model_1}
\end{figure}

\begin{figure}[!tb]
    \centering
    \includegraphics[width=.7\textwidth]{figures/errcases_rate_density_vs_radius_model_2.pdf}
    %\includegraphics[width=1.0\textwidth]{figures/errcases_rate_density_vs_radius_logs_model_2.pdf}
    \caption{
    Inferred planet occurrence rates as a function of planet radius in Model 
    \#2.
    This model has fixed planets and primaries, but varying secondaries.
    }
    \label{fig:errcases_model_2_linear}
\end{figure}

\begin{figure}[!tb]
    \centering
    %\includegraphics[width=1.0\textwidth]{figures/errcases_rate_density_vs_radius_model_2.pdf}
    \includegraphics[width=.7\textwidth]{figures/errcases_rate_density_vs_radius_logs_model_2.pdf}
    \caption{
    Same as Fig.~\ref{fig:errcases_model_2_linear}, but with logarithmic 
    $y$-axis, and different $x$ scale.
    }
    \label{fig:errcases_model_2_log}
\end{figure}

\begin{figure}[!tb]
    \centering
    \includegraphics[width=\textwidth]{figures/errcases_rate_density_vs_radius_logs_model_3.pdf}
    \caption{
    Inferred planet occurrence rates as a function of planet radius in Model 
    \#3.
    This model has fixed primaries and single stars, but varying 
    secondaries.
    The true planet radius distribution is a power law with exponent $-2.92$ 
    above $2R_\oplus$, below which it is uniform (e.g., Howard et al., 
    2012).
    }
    \label{fig:errcases_model_3_log}
\end{figure}

%\begin{figure}[!b]
%    \begin{center}
%        \includegraphics[width=.9\textwidth]{figures/prob_r_a.pdf}
%    \end{center}
%    \caption{Model \#2 (Sec.~\ref{sec:model_2}): 
%    probability of observing an apparent radius $r_a$ given a 
%    planet orbiting the primary, $p_1(r_a)$, or the secondary, $p_2(r_a)$ of 
%    a 
%    binary system.
%    The true planet radius is fixed~--~a delta function centered on ``1''.
%    This plot takes $\alpha=3.5, \beta=0$, for $\alpha$ the exponent in 
%    the mass-luminosity relation $L \propto M^\alpha$, and $\beta$ the 
%    exponent in the distribution of mass ratios in a volume-limited sample.
%    Each distribution is separately normalized.
%    }
%    \label{fig:model2_prob_r_a}
%\end{figure}

### 2017-10-20 paper/generalities.tex
update paper

--- 2017-10-19 memos/generalities.tex
@@
-\section*{Generalities}
+\subsection{Analytic Preliminaries}
 
@@
 A transit survey will have a rate density of detected planets $\hat{\Gamma}$, 
-which will be the (dot) product of the rate density and the detection 
-efficiency $Q(\vec{x})$:
+which for each system type will be the product of the rate density and the 
+detection efficiency $Q_i(\vec{x})$:
 \begin{equation}
@@
 The detection efficiency includes the geometric transit probability, as well 
-as any incompleteness effects.
+as any incompleteness effects. Foreman-Mackey et al. (2014) discuss how this 
+is calculated in practice.

### 2017-10-20 paper/main.tex
update paper

@@
 \documentclass[12pt,modern]{aastex61}
-\usepackage{graphicx}
+\usepackage{graphics,graphicx}
+\usepackage{hyperref}
 \usepackage{amssymb}
 \usepackage{amsmath}
-\usepackage{todonotes}
+\usepackage{comment}
 
@@
 
-\title{ Binarity's effects on occurrence rates measured by transit surveys}
+\title{ The effects of binarity on planet occurrence rates measured by transit 
+surveys}
 
 \author{
-LGB, KM, JNW
+L. G. Bouma, K. Masuda, J. N. Winn
 }
 
-\input{generalities}
-\input{model_1}
-\input{model_2}
+\begin{abstract}
+
+The aim of this work is to clarify what biases, if any, stellar binarity 
+introduces to occurrence rates inferred from transit surveys.
+Ignoring binarity leads to diluted planetary radii, overestimated 
+detection efficiencies, and an undercounted number of selected stars.
+These effects skew occurrence rate measurements in different directions, and 
+we develop simplified analytic and numerical models to clarify which are the 
+most important.
+For models in which all planets have the same radii, we find that 
+ignoring binarity leads to an inferred occurrence rate at the true radius two 
+to three times smaller than the actual rate (an upper limit to the possible 
+error).
+%However, the direction and size of the error depend on the assumed 
+%planet population's radius distribution.
+More realistically, using Howard et al. (2012)'s radius distribution and a 
+simple stellar model, we find that ignoring binarity leads to a $\sim10-30\%$ 
+underestimate of the number of planets per star over most planetary radii.
+In particular, the 30\% underestimate applies to hot Jupiter occurrence rates, 
+which suggests that the discrepancy between rates measured by {\it Kepler} and 
+RV surveys might be attributable to the {\it Kepler} sample containing more 
+binaries.
+
+\end{abstract}
+
+\input{introduction}
+\input{methods}
+\input{results}
+\input{discussion}
+
+\input{figures.tex}
 

### 2017-10-20 paper/model_1.tex
update paper

--- 2017-10-19 memos/model_1.tex
@@
-\section{Model \#1: fixed stars, fixed planets, twin binaries}
+\subsection{Model \#1: fixed stars, fixed planets, twin binaries}
 \label{sec:model_1}
@@
 either single or twin stars with otherwise identical physical properties.
-Then $\vec{x} = (r,R,a)$, and 
-\begin{equation}
-p_i(\vec{x})= \delta(r-r_p)\delta(R-R_\star)\delta(a-a_p)
-\equiv \delta^3(r_p,R_\star,a_p),
-\label{eq:model_1_prob}
-\end{equation}
-for $r_p$ and $R_\star$ some fixed planet and stellar radii, $a_p$ a fixed 
-semi-major axis, and $\delta$ the Dirac delta function, whose latter compact 
-form will be used for brevity.
-
-The occurrence rate density for this model is
+The occurrence rate density at a planet radius $r$, stellar radius $R$, and 
+semimajor axis $a$ for this model is
 \begin{equation}
@@
 \end{equation}
-and the occurrence rate over any interval that includes $r_p$, $R_\star$, and 
+for $r_p$ and $R_\star$ some fixed planet and stellar radii, $a_p$ a fixed 
+semi-major axis, and $\delta$ the Dirac delta function, whose compact
+form will be used for brevity.
+The occurrence rate over any interval that includes $r_p$, $R_\star$, and 
 $a_p$ is
@@
 For binary systems there is a necessary incompleteness, and combining 
-Eqs.~\ref{eq:general_detection_efficiency} through~\ref{eq:dilution} in the 
-case of twin binaries yields
+Eqs.~\ref{eq:general_detection_efficiency} through~\ref{eq:dilution} yields
 \begin{align}
-Q_i(\vec{x})
-&=
-\left.
-\begin{cases}
-R/a, & \text{for } i=0\ {\rm single} \\
-(R/a) (1+\gamma_R)^{-3} = (R/8a), & \text{for } i=1\ {\rm primary} \\
-(R/a) (1+\gamma_R^{-1})^{-3}\gamma_R^{-5/\alpha} = (R/8a), & \text{for } i=2\ 
-{\rm 
-secondary}.
-\end{cases}
-\right.
-\label{eq:model1_detection_efficiency}
+Q_0 &= Q_{g,0}Q_{c,0} = Q_{g,0} \label{eq:detection_efficiency_0}\\
+Q_1 &= Q_{g,1}Q_{c,1} = Q_{g,0} (1+q^\alpha)^{-3} \\
+Q_2 &= Q_{g,2}Q_{c,2} = Q_{g,0} q^{2/3} (1+q^{-\alpha})^{-3} q^{-5}, 
+\label{eq:detection_efficiency_2}
 \end{align}
+for $Q_{g,0}=R/a$, the transit probability in single star systems.
+The factors of $q^{2/3}$ and $q^{-5}$ in
+Eq.~\ref{eq:detection_efficiency_2} come from the assumed stellar 
+mass-luminosity-radius relation: $R\propto M \propto L^{1/\alpha}$.
+For $q=1$ both terms evaluate to unity, but they will later become relevant.
 
-The factor of $\gamma_R^{-5/\alpha}$ in the latter expression of 
-Eq.~\ref{eq:model1_occ_rate_density} comes from the mass-luminosity-radius 
-relation of stars in the model: we assume $R\propto M \propto L^{1/\alpha}$.
-For $\gamma_R=1$ this term is unimportant, but it will later become relevant.
+Summarizing, we have written the rate density for each system type
+(Eq.~\ref{eq:model1_occ_rate_density}) and the detection efficiency for each 
+system type 
+(Eq.~\ref{eq:detection_efficiency_0}-\ref{eq:detection_efficiency_2}),
+and so have fully specified the rate density of detected planets, 
+in addition to the true rate density.
 
-Summarizing, since we have written the rate density for each system type
-(Eq.~\ref{eq:model1_occ_rate_density}) and the detection efficiency for each 
-system type (Eq.~\ref{eq:model1_detection_efficiency}), we have fully 
-specified the rate density of detected planets, $\hat{\Gamma}=\sum Q_i 
-\Gamma_i$, in addition to the true rate density.
-
-\subsection{What does an observer ignoring binarity infer? }
-
-Consider an observer who ignores binarity.
-They assume a detection efficiency $\tilde{Q}=Q_0$,
-measure a detected planet rate density $\tilde{\Gamma}$, 
-and infer an apparent rate density $\Gamma_a$.
+\paragraph{What does an observer ignoring binarity infer?} 
+An observer who ignores binarity assumes a detection efficiency 
+$\tilde{Q}=Q_0$,
+measures a detected planet rate density $\tilde{\Gamma}$, 
+and infers an apparent rate density $\Gamma_a$.
 Analogous to Eq.~\ref{eq:detected_rate_density},
@@
 This observer miscounts the number of total searched stars, does not correct 
-for completeness, and misclassifies the planetary radii because of dilution.
+for incompleteness, and misclassifies the planetary radii because of dilution.
 
-\subsection{Correction to inferred rate density and inferred rate}
+\paragraph{Correction to inferred rate density and inferred rate}
 
-Define the rate density correction factor, $X_\Gamma$, as the ratio of the 
+Define a rate density correction factor, $X_\Gamma$, as the ratio of the 
 apparent to true rate densities:
@@
 Using Raghavan et al. (2010)'s $0.7-1.3M_\odot$ multiplicity fraction as our 
-binary fraction, we set ${\rm BF}=0.44$.
+binary fraction\footnote{
+The binary fraction is the fraction of systems in a volume-limited sample that 
+are binary. It is equivalent to the multiplicity fraction if there are no 
+triple, quadruple, or higher order multiples.
+}, we set ${\rm BF}=0.44$.
 The resulting correction to the rate density is $X_\Gamma(r_p) \approx 0.31$. 
 The correction at $r_p/\sqrt{2}$ is infinite.
+The numerical realization of this model agrees with these analytic values, and 
+its output is shown in Fig~\ref{fig:errcases_model_1}.
 %beta = 2.2223355980148636
-
-If we assume that $\Lambda_0 = \Lambda_1$, but that $\Lambda_2=0$, we find 
+If instead we assume that $\Lambda_0 = \Lambda_1$, but that $\Lambda_2=0$, we 
+find 
 $X_\Gamma(r_p) = (1+2\mu)/(1+\mu)^2$.
 Taking the same binary fraction, this evaluates to $X_\Gamma(r_p)\approx 0.53$.
+Since the correction to the rate is equal to that of rate density, at 
+$r=r_p$, the occurrence rate is underestimated by a factor of roughly 2 to 3.
 
-In passing, note that a correction to the inferred rate, $X_\Lambda$, can be 
-defined analogously:
-\begin{equation}
-X_\Lambda \equiv \frac{\Lambda_a}{\Lambda}.
-\end{equation}
-For the twin binary model, the correction to the rate is the same as that to 
-the rate density.
+%Note that a correction to the inferred rate, $X_\Lambda$, can be 
+%defined analogously:
+%\begin{equation}
+%X_\Lambda \equiv \frac{\Lambda_a}{\Lambda}.
+%\end{equation}
+%For this twin binary model, the correction to the rate is the same as that to 
+%the rate density.

### 2017-10-20 paper/model_2.tex
update paper

@@
-\section{Model \#2: fixed planets and primaries, varying secondaries}
+\subsection{Model \#2: fixed planets and primaries, varying secondaries}
 \label{sec:model_2}
 
-The main use of our binary-twin model was to help develop intuition.
-Gradually introducing complexity, we now let the light ratio $\gamma_R = 
-L_2/L_1$ vary across the binary population.
+The main use of our binary-twin model is to help develop intuition.
+We now let the light ratio $\gamma_R = L_2/L_1$ vary across the binary 
+population.
 It does so because the underlying mass ratio $q=M_2/M_1$ varies.
@@
 distribution seems roughly uniform, $\beta \approx 0$, except for a claimed 
-excess of twin binaries with $q\approx 1$, and a deficiency of $q<0.1$ 
+excess of twin binaries with $q\approx 1$, and an obvious lack of $q<0.1$ 
 stellar companions.
@@
 Malmquist-like bias $(1+q^\alpha)^{3/2}$.
+In passing, we note that various authors (ourselves included) have incorrectly 
+used volume-limited binary distributions in Monte Carlo simulations of 
+transit surveys (Sullivan et al 2015, Hartman and Bakos XX, Guenther et al 
+2017, Bouma et al 2017).
 
@@
 specified by the detection efficiencies for each type of system.
-These are nearly identical to Eq.~\ref{eq:model1_detection_efficiency}:
-\begin{align}
-Q_0 &= Q_{g,0}Q_{c,0} = Q_{g,0} \label{eq:model2_Q_0}\\
-Q_1 &= Q_{g,1}Q_{c,1} = Q_{g,0} (1+q^\alpha)^{-3} \\
-Q_2 &= Q_{g,2}Q_{c,2} = Q_{g,0} q^{2/3} (1+q^{-\alpha})^{-3} q^{-5}, 
-\label{eq:model2_Q_2}
-\end{align}
-for $Q_{g,0}=R/a$, the transit probability in single star systems.
-The detection efficiency for secondaries (Eq.~\ref{eq:model2_Q_2}) includes 
-the transit probability from the smaller stellar radius, and combines 
+These are identical to 
+Eqs.~\ref{eq:detection_efficiency_0}-\ref{eq:detection_efficiency_2}.
+The detection efficiency for secondaries (Eq.~\ref{eq:detection_efficiency_2}) 
+includes the transit probability from the smaller stellar radius, and combines 
 dilution, the transit duration, and stellar radius for the completeness
@@
 
-\subsection{What does an observer ignoring binarity infer?}
-
+\paragraph{What does an observer ignoring binarity infer?}
 As a reminder, the apparent rate density is found by correcting the detected 
@@
     \item The true planetary radii $r$ are interpreted as apparent radii $r_a$.
-    \item The true stellar radii $R$ are all thought to be $R_\star$. In 
-    ``reality'', this only holds for single stars and primaries.
-    \item The selected and searchable stars are miscounted.
-\end{enumerate}
-
-Writing the apparent rate density as a function of $\vec{x}=(r_a,P,a,R)$, 
-where $R$ is written interchangeably with the binary mass ratio $q$,
-\begin{equation}
-\Gamma_a(\vec{x}) =
-Q_0 w_a \Lambda_0 \delta^4(r_p)
-+
-Q_1 w_b \Lambda_1 \delta^3(P_p) p_1(r_a)
-+
-Q_2 w_b \Lambda_2 \delta^2(P_p,a_p) p_2(q) p_2(r_a),
-\label{eq:model2_Gamma_a}
-\end{equation}
-where the detection efficiencies are given in 
-Eqs.~\ref{eq:model2_Q_0}-\ref{eq:model2_Q_2}, $p_2(q)$ is 
-given by Eq.~\ref{eq:model_2_p_2}, and as in the first model, 
-$w_a=N_0/(N_0+N_1)$, $w_b=N_1/(N_0+N_1)$.
-%WARNING: is it actually correct to separate the q and r_a dependence?
-%THEY are dependent on each-other...
-
     The apparent radii depend on the system type:
@@
     assuming that the host star is the primary.
+    \item The true stellar radii $R$ are all thought to be $R_\star$. In 
+    ``reality'', this only holds for single stars and primaries.
+    \item The selected and searchable stars are miscounted.
+\end{enumerate}
 
-Since the apparent radii directly depend on the mass ratio, we wish to write
+To write the apparent rate density as a function of the apparent 
+radius $r_a$, we marginalize out the planet period, semimajor axis, and 
+stellar radius (or equivalently the mass ratio, for binaries):
 \begin{equation}
-p_i (r_a)
-=
-p_i(q(r_a))
-\left|
-    \frac{{\rm d}q}{{\rm d}r_a}
-\right|,
-\quad i\in\{1,2\},
-\label{eq:model2_p_i}
+\Gamma_a(r_a) =
+w_a \Lambda_0 \delta(r_p)
++
+w_b \Lambda_1 I_1(r_a)
++
+w_b \Lambda_2 I_2(r_a),
+\label{eq:model2_Gamma_a}
 \end{equation}
-where $q(r_a)$ must be found by inverting Eq.~\ref{eq:model2_ra}.
-For $i=1$, this gives
+where the detection efficiencies are given in 
+Eqs.~\ref{eq:detection_efficiency_0}-\ref{eq:detection_efficiency_2}, and as 
+in the first model, $w_a=N_0/(N_0+N_1)$, $w_b=N_1/(N_0+N_1)$. The ratio of 
+primaries to singles, $\mu$, is now given by a variant of 
+Eq.~\ref{eq:mu_definition}:
 \begin{equation}
-p_1(r_a) \propto \frac{r_p^5}{r_a^6}
-\left( 
-    \left(\frac{r_p}{r_a}\right)^2 - 1
-\right)^{\frac{\beta - \alpha + 1}{\alpha}},
+\mu \equiv \frac{N_1}{N_0} = \frac{\rm BF}{1 - {\rm BF}} \left(2^{3/2} - 
+\int_{1}^{\sqrt{2}} u^2 (u^2 -1)^{1/\alpha} \,{\rm d}u\right),
+\label{eq:mu_model_2}
 \end{equation}
-but for $i=2$, $r_a(q)$ is not invertible.
-Instead, we find $p_2(r_a)$ numerically\footnote{We 
-sample from $p_2(q)$ (Eq.~\ref{eq:model_2_p_2}), and use 
-Eq.~\ref{eq:model2_p_i} to compute the resulting apparent radii}, 
-and show it in Fig.~\ref{fig:model2_prob_r_a}.
+where the latter dimensionless integral is easily found numerically.
+The $I_1(r_a)$ and $I_2(r_a)$ terms marginalize over the joint distribution of 
+apparent radius and mass ratio:
+\begin{align}
+I_i(r_a) &= 
+\int_0^1 p({\rm has\ detected\ planet}, r_a, q | {\rm star\ is\ type\ }i)
+    \,{\rm d}q,
+\quad
+{\rm for}\ i\in\{1, 2\}, \\
+&=
+\int_0^1 
+    p({\rm has\ detected\ planet} | r_a, q, {\rm star\ is\ type\ }i) 
+    \nonumber\\
+    &\quad\quad\quad\quad\times p(r_a | q, {\rm star\ is\ type\ }i)
+    \,p(q | {\rm star\ is\ type\ }i)
+\,{\rm d}q.
+\end{align}
+The first term is the detection efficiency; the second is a $\delta$-function 
+of the apparent radius; the latter is the mass ratio distribution given by 
+Eq.~\ref{eq:model_2_p_2}.
+The analytic solution for $i=1$ is
+\begin{equation}
+I_1(r_a) = \frac{1}{\mathcal{N}_1} \left(\frac{r_p}{r_a}\right)^{-3}
+\left( \left(\frac{r_p}{r_a}\right)^2 -1  
+    \right)^{\frac{\gamma+\beta}{\alpha}},
+\quad {\rm for}\ r_p/\sqrt{2} < r_a < r_p,
+\end{equation}
+where $\mathcal{N}_1$ is the normalization term of the binary mass ratio 
+distribution (Eq.~\ref{eq:model_2_p_2}): $\mathcal{N}_1 = \int_0^1 
+q^{\gamma+\beta} (1+q^\alpha)^{3/2} {\rm d}q$.
 
-\subsection{Correction to inferred rate density}
+For $i=2$, there is no analytic solution, because evaluating the integral 
+requires imposing the constraint that $r_a  = r_p 
+(1+q^{-\alpha})^{-1/2}q^{-1}$. This equation can be re-written
+\begin{equation}
+\left(\frac{r_p}{r_a}\right)^2 = q^2 + q^{-\alpha + 2},
+\end{equation}
+which has no analytic solution except for special values of $\alpha$, the 
+mass-luminosity exponent.
+For $\alpha=3.5$, our nominal case, semianalytic solutions can be found.
 
+However, since our main interest is in understanding the qualitative behavior 
+of the solutions, we focus instead on a few limiting cases.
+
+\paragraph{Correction to inferred rate density}
 Recall that the rate density correction factor, $X_\Gamma$, is the ratio of 
 the apparent to true rate densities.
-Applying Eqs.~\ref{eq:model2_rate_density} and~\ref{eq:model2_Gamma_a},
-\begin{align}
-X_\Gamma(r,q)
-=
-\frac{
-    w_a \Lambda_0 \delta(r_p,R_\star)
-    +
-    Q_{c,1} w_b \Lambda_1 \delta(R_\star) p_1(r_a)
-    +
-    Q_{c,2} q^{2/3} w_b \Lambda_2 p_2(q) p_2(r_a)
-}
-{
-w_0 \Lambda_0 \delta^2(r_p,R_\star) +
-w_1 \Lambda_1 \delta^2(r_p,R_\star) +
-w_2 \Lambda_2 p_2(q) \delta(r_p)
-},
-\label{eq:model2_X_Gamma}
-\end{align}
-where the stellar radius-dependent $\delta$-functions are not important.
-To marginalize over $q$, we must write
+We consider a ``nominal model'' in which the stellar population is similar to 
+Sun-like stars in the local neighborhood:
+${\rm BF}=0.44$, $\alpha=3.5$, $\beta=0$.
+Our default assumption is also that the occurrence of planets is independent 
+of stellar mass ($\gamma=0$), so secondaries have the same occurrence rate as 
+primaries and single stars.
+Under these assumptions, the planetary rate density is
 \begin{equation}
-X_\Gamma(r) = \frac{\Gamma_a(r)}{\Gamma(r)}
-= \frac{\sum_i \int_0^1 \Gamma_{a,i}(r,q) \,{\rm d}q}
-{\sum_i \int_0^1 \Gamma_{i}(r,q) \,{\rm d}q}.
-\label{eq:model2_X_Gamma_of_r}
-\end{equation}
-Using the same definition of $\mu = N_1/N_0$ from Eq.~\ref{eq:mu_definition}, 
-it is important when performing each integral to note that $\mu$ is a function 
-of $q$:
-\begin{equation}
-\mu = \mathcal{B} (1+q^\alpha)^{3/2},
-\quad \mathcal{B} \equiv \frac{{\rm BF}}{1 - {\rm BF}} .
-\end{equation}
-Thus the weights in the apparent and actual rate densities, $(w_a,w_b)$ and 
-$(w_0,w_1,w_2)$, are all functions of $q$.
-
-The crucial point of Eq.~\ref{eq:model2_X_Gamma_of_r} is that since both the 
-real and apparent rate densities are separable, the expressions will simplify.
-
-Our ``nominal model'' is a stellar population similar to Sun-like stars in the 
-local neighborhood:
-${\rm BF}=0.44$, $\alpha=3.5$, $\beta=0$.
-We also assume that the occurrence of planets is independent of stellar mass 
-($\gamma=0$).
-Under these nominal assumptions, the planetary rate density is
-\begin{equation}
-\Gamma(r) \approx \delta(r-r_p) \left( \Lambda_0 + \Lambda_1 + 
+\Gamma(r) \approx \delta(r_p) \left( \Lambda_0 + \Lambda_1 + 
 \Lambda_2 \right) / 3,
@@
 \begin{equation}
-\Gamma_a(r) = c_0 \Lambda_0 \delta(r-r_p)
-             +c_1 \Lambda_1 p_1(r_a)
-             +c_2 \Lambda_2 p_2(r_a),
+\Gamma_a(r) = c_0 \Lambda_0 \delta(r_p)
+             +c_1 \Lambda_1 I_1(r_a)
+             +c_2 \Lambda_2 I_2(r_a),
 \label{eq:model2_Gamma_a_r}
 \end{equation}
-for $c_0\approx 0.49$, $c_1\approx 0.32$, $c_2\approx 0.03$, and the two 
-latter apparent-radius dependences are shown in Fig.~\ref{fig:model2_prob_r_a}.
-
-While the detailed radial dependence of the error between 
-Eqs.~\ref{eq:model2_Gamma_r} 
-and~\ref{eq:model2_Gamma_a_r} is interesting, the general point is
-that dilution produces a spectrum of apparent planetary radii, which
-skews occurrence rate and occurrence rate density measurements.
-
+for $c_0\approx 0.49$, $c_1\approx 0.32$, $c_2\approx 0.03$.
+%171011_integrals.nb
 Evaluating the correction term at $r=r_p$, since $\lim_{q\rightarrow0} 
@@
 
-\section{Model \#3: Fixed primaries, varying planets and secondaries}
+We also run this model in our numerical simulation.
+The results are shown in Figs~\ref{fig:errcases_model_2_linear} 
+and~\ref{fig:errcases_model_2_log}.
+They produce the same correction factors as predicted analytically.
+They also show a peak in the inferred rate at $r_p/\sqrt{2}$ from 
+secondaries.
+This effect requires the observer to incorrectly estimate the host star's 
+radius; if they somehow knew the host star's radius, but did not correct for 
+binary dilution, the peak would instead be at an apparent radius of 0.
 
-\begin{figure}[!b]
-    \begin{center}
-        \includegraphics[width=.9\textwidth]{figures/prob_r_a.pdf}
-    \end{center}
-    \caption{Model \#2 (Sec.~\ref{sec:model_2}): 
-    probability of observing an apparent radius $r_a$ given a 
-    planet orbiting the primary, $p_1(r_a)$, or the secondary, $p_2(r_a)$ of a 
-    binary system.
-    The true planet radius is fixed~--~a delta function centered on ``1''.
-    This plot takes $\alpha=3.5, \beta=0$, for $\alpha$ the exponent in 
-    the mass-luminosity relation $L \propto M^\alpha$, and $\beta$ the 
-    exponent in the distribution of mass ratios in a volume-limited sample.
-    Each distribution is separately normalized.
-    }
-    \label{fig:model2_prob_r_a}
-\end{figure}
+To summarize, dilution produces a spectrum of apparent planetary radii. In 
+this model, this produces overestimated rates everywhere except where there 
+are actually planets, where the rate is underestimated by a factor of 2.

### 2017-10-20 paper/model_3.tex
update paper

\subsection{Model \#3: Fixed primaries, varying planets and secondaries}
\label{sec:model_3}

See Fig.~\ref{fig:errcases_model_3_log}.

### 2017-10-21 paper/model_2.tex
paper: begin model #3 section

@@
 The first term is the detection efficiency; the second is a $\delta$-function 
-of the apparent radius; the latter is the mass ratio distribution given by 
+of the apparent radius; the last is the mass ratio distribution given by 
 Eq.~\ref{eq:model_2_p_2}.

### 2017-10-21 paper/model_3.tex
paper: begin model #3 section

@@
 
-See Fig.~\ref{fig:errcases_model_3_log}.
+In this model, as in the previous one, all single and primary stars have 
+identical properties.
+Only the secondaries have masses, radii, and luminosities that vary between 
+systems: $M\propto R \propto L^{1/\alpha}$.
+The radii of planets are assigned independently of any host system
+property, and are sampled from an intrinsic radius distribution, which we take 
+as
+\begin{align}
+p_r(r)
+&\propto
+\left.
+\begin{cases}
+r^\delta & \text{for } r\geq 2R_\oplus \\
+{\rm constant} & \text{for } r\leq2R_\oplus.
+\end{cases}
+\right.
+\label{eq:model3_radius_distribution}
+\end{align}
+Following Howard et al. (2012)'s measurement, we take $\delta = -2.92$.
+Our ``nominal model'' remains the same: the binary fraction is 0.44, 
+$\alpha=\beta=\gamma=0$.
+We take $\Lambda_i$, the occurrence rate integrated over all possible phase 
+space for the $i^{\rm th}$ system type, to be equal for singles, primaries, 
+and secondaries.
+
+For this model, we forgo analytic development and focus only on numerics 
+(Sec.~\ref{sec:numerical_methods}).
+The nominal results are shown in Fig.~\ref{fig:errcases_model_3_log}.
+For the assumed planetary and stellar distributions, the inferred rate is 
+underestimated over all radii.
+Taking Fig.~\ref{fig:errcases_model_3_log} and counting the number of planets 
+per star with $r>8R_\oplus$, we can compare the true and inferred hot Jupiter 
+occurrence rates.
+Under the previously described assumptions, the true rate is 9.1 hot Jupiters 
+per thousand stars.
+The inferred rate is 6.9 per thousand stars.
+This means that the inferred rate underestimates the true rate by a 
+multiplicative factor of $\sim 1.3$.
+This number is suggestively close to the claimed discrepancy between 
+hot Jupiter occurrence rates measured by radial velocity and transit surveys 
+(Wright et al. 2012).
+
+However, this result only applies under the assumption that $\Lambda_0 = 
+\Lambda_1 = \Lambda_2$.
+Given that hot Jupiters are thought to be less common around less-massive 
+stars, it would be more sensible to consider $\Lambda_2<\Lambda_0$, while 
+letting single stars and primaries host planets at the same rate.
+We show the HJ multiplicative correction factor as a function of 
+$\Lambda_2/\Lambda_0$ in Fig BLAH.

### 2017-10-21 src/numerical_models.py
add Lambda2 variation in code

@@ -304,8 +304,9 @@ if __name__ == '__main__':
              'inferred_Γ': inferred_dict['Γ']
             }
             )
-    fname = '../data/results_model_'+repr(model_number)+'.out'
-    outdf.to_csv(fname, index=False)
+    fname = '../data/results_model_{:d}_Lambda2_{:.1f}'.format(
+            model_number, Λ_2)
+    outdf.to_csv(fname+'.out', index=False)
     print('wrote output to {:s}'.format(fname))
 
     # The sum of the true distribution's histogrammed rate densities is the true
@@ -464,7 +465,7 @@ if __name__ == '__main__':
                  'inferred_Γ': inferred_dict['Γ']
                 }
                 )
-        fname = '../data/results_model_{:d}_error_case_{:d}.out'.format(
-                model_number, error_case_number)
+        fname = '../data/results_model_{:d}_error_case_{:d}_Lambda2_{:.1f}.out'.format(
+                model_number, error_case_number, Λ_2)
         outdf.to_csv(fname, index=False)
         print('\twrote output to {:s}'.format(fname))

### 2017-10-21 src/plot_errcases_rate_densities_vs_radius.py

@@ -29,7 +29,7 @@ import numpy as np
 from brokenaxes import brokenaxes
 
 def make_plot(model_number, logx=None, logy=None, withtext=None,
-        stdout=False, brokenx=False):
+        stdout=False, brokenx=False, Λ_2=None):
 
     # Make summary plot
     if not brokenx:
@@ -45,7 +45,12 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
                 despine=True)
         ax=bigax
 
-    fname = '../data/results_model_'+repr(model_number)+'.out'
+    if isinstance(Λ_2,float) or isinstance(Λ_2,int):
+        fname = '../data/results_model_{:d}_Lambda2_{:.1f}.out'.format(
+                model_number, Λ_2)
+    else:
+        fname = '../data/results_model_'+repr(model_number)+'.out'
+
     df = pd.read_csv(fname)
 
     if model_number == 3:
@@ -63,8 +68,12 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
 
     for error_case_number in [1,2,3]:
 
-        fname = '../data/results_model_{:d}_error_case_{:d}.out'.format(
-                model_number, error_case_number)
+        if isinstance(Λ_2,float) or isinstance(Λ_2,int):
+            fname = '../data/results_model_{:d}_error_case_{:d}_Lambda2_{:.1f}.out'.format(
+                    model_number, error_case_number, Λ_2)
+        else:
+            fname = '../data/results_model_{:d}_error_case_{:d}.out'.format(
+                    model_number, error_case_number)
         df = pd.read_csv(fname)
 
         if error_case_number == 3:
@@ -132,7 +141,12 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
         # Assess HJ rate difference.
         #Howard 2012 boundary #1 and boundary #2:
         for lower_bound in [5.6,8]:
-            fname = '../data/results_model_'+repr(model_number)+'.out'
+            if isinstance(Λ_2,float) or isinstance(Λ_2,int):
+                fname = '../data/results_model_{:d}_Lambda2_{:.1f}.out'.format(
+                        model_number, Λ_2)
+            else:
+                fname = '../data/results_model_{:d}.out'.format(
+                        model_number)
             df = pd.read_csv(fname)
 
             inds = df['bin_left'] > lower_bound
[…]

### 2017-10-21 src/plot_rate_densities_vs_radius.py

@@ -29,10 +29,15 @@ import numpy as np
 from brokenaxes import brokenaxes
 
 def make_plot(model_number, logx=None, logy=None, withtext=None,
-        stdout=False, brokenx=None):
+        stdout=False, brokenx=None, Λ_2=None):
 
     # Make summary plot
-    fname = '../data/results_model_'+repr(model_number)+'.out'
+    if isinstance(Λ_2,float) or isinstance(Λ_2,int):
+        fname = '../data/results_model_{:d}_Lambda2_{:.1f}.out'.format(
+                model_number, Λ_2)
+    else:
+        fname = '../data/results_model_'+repr(model_number)+'.out'
+
     df = pd.read_csv(fname)
 
     if not brokenx:
@@ -136,6 +141,9 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
                     transform=ax.transAxes, fontsize='x-small')
             outname += '_withtext'
 
+        if isinstance(Λ_2,float) or isinstance(Λ_2,int):
+            outname += '_Lambda2_{:.1f}'.format(Λ_2)
+
     f.savefig(outname+'.pdf', bbox_inches='tight')
 
 
@@ -150,3 +158,7 @@ if __name__ == '__main__':
     make_plot(2, logy=True)
     make_plot(3, logy=True)
     make_plot(3, withtext=True, stdout=True)
+
+    make_plot(3, Λ_2=0)
+    make_plot(3, logy=True, Λ_2=0)
+    make_plot(3, withtext=True, stdout=True, Λ_2=0)

### 2017-10-21 src/numerical_models.py
new interface to numerical models; new wrapper too
also added `plot_HJ_correction_vs_LambdaTwo.py`, a thing to plot the
normalization-dependence of the HJ correction.

@@ -1,6 +1,11 @@
 # -*- coding: utf-8 -*-
 '''
-see README.md
+See README.md for full description.
+
+Example usage:
+>>> python numerical_models.py --quickrun --modelnumber 3 --LambdaTwo 0.5
+
+Alternatively, write a wrapper like `run_numerical_models.py`.
 '''
 
 from __future__ import division, print_function
@@ -10,35 +15,58 @@ from astropy import units as u, constants as c
 from scipy.integrate import trapz
 from scipy.interpolate import interp1d
 from math import pi as π
-import os
+import os, argparse, decimal
 
-global α, β, γ
+def numerical_transit_survey(
+        quickrun,
+        model_number,
+        Λ_2,
+        BF=0.44,
+        α=3.5,
+        β=0,
+        γ=0,
+        δ=-2.92,
+        Λ_0=0.5,
+        Λ_1=0.5
+        ):
+    '''
+    Inputs:
+        quickrun = False
+        model_number = 3
 
-if __name__ == '__main__':
-    np.random.seed(42)
+        BF = 0.44   # binary fraction. BF = n_d / (n_s+n_d). Raghavan+ 2010 solar.
 
-    ##########################################
-    # Inputs.
-    quickrun = False
+        α = 3.5     # exponent in L~M^α
+        β = 0       # exponent in p_vol_limited(q) ~ q^β
+        γ = 0       # exponent in p(has planet|secondary w/ q) ~ q^γ
+        δ = -2.92   # exponent in p(r) ~ r^δ, from Howard+ 2012
+
+        Λ_0 = 0.5 # fraction of selected singles with planet
+        Λ_1 = 0.5 # fraction of selected primaries with planet
+        Λ_2 = 0.5 # fraction of selected secondaries with planet.
+
+    Outputs:
+        Four data files (in ../data/):
+
+        'results_model_{:d}_Lambda2_{:.2f}'.format( model_number, Λ_2)
+
+        'results_model_{:d}_error_case_{:d}_Lambda2_{:.2f}.out'.format(
+            model_number, error_case_number, Λ_2)
+
+        The output data files contain binned occurrence rates as a function of
+        true and apparent radius.
+
+    '''
+
     slowrun = not quickrun
-    model_number = 3
-
-    BF = 0.44   # binary fraction. BF = n_d / (n_s+n_d). Raghavan+ 2010 solar.
-
-    α = 3.5     # exponent in L~M^α
-    β = 0       # exponent in p_vol_limited(q) ~ q^β
[…]

### 2017-10-21 src/plot_HJ_correction_vs_LambdaTwo.py

@@ -1,0 +1,87 @@
+'''
+Given that hot Jupiters are thought to be less common around less-massive
+stars, it would be more sensible to consider Λ_2 < Λ_0 , while letting single
+stars and primaries host planets at the same rate.
+
+This plots the HJ multiplicative correction factor as a function of Λ_2 /Λ_0.
+'''
+# -*- coding: utf-8 -*-
+from __future__ import division, print_function
+import matplotlib as mpl
+mpl.use('pgf')
+pgf_with_custom_preamble = {
+    'pgf.texsystem': 'pdflatex', # xelatex is default; i don't have it
+    'font.family': 'serif', # use serif/main font for text elements
+    'text.usetex': True,    # use inline math for ticks
+    'pgf.rcfonts': False,   # don't setup fonts from rc parameters
+    'pgf.preamble': [
+        '\\usepackage{amsmath}',
+        '\\usepackage{amssymb}'
+        ]
+    }
+mpl.rcParams.update(pgf_with_custom_preamble)
+
+import matplotlib.pyplot as plt
+import numpy as np, pandas as pd
+import os
+
+if __name__ == '__main__':
+
+    model_number = 3
+
+    Λ_2_arr = np.arange(0, 0.5+0.05, 0.05)
+
+    Λ_HJ_true_arr, Λ_HJ_inferred_arr = [], []
+
+    # compute the HJ correction factors for each Λ_2
+    for Λ_2 in Λ_2_arr:
+
+        savdir = '../data/numerics/'
+        fname = 'results_model_{:d}_Lambda2_{:.2f}.out'.format(
+                model_number, Λ_2)
+        df = pd.read_csv(savdir+fname)
+
+        lower_bound = 8
+        inds = df['bin_left'] > lower_bound
+        Λ_HJ_true = np.sum(df[inds]['true_Γ'])
+        Λ_HJ_inferred = np.sum(df[inds]['inferred_Γ'])
+
+        Λ_HJ_true_arr.append(float(Λ_HJ_true))
+        Λ_HJ_inferred_arr.append(float(Λ_HJ_inferred))
+
+    Λ_HJ_true_arr = np.array(Λ_HJ_true_arr)
+    Λ_HJ_inferred_arr = np.array(Λ_HJ_inferred_arr)
+
+    X_correction = Λ_HJ_inferred_arr/Λ_HJ_true_arr
+
+    Λ_0 = 0.5
+
[…]

### 2017-10-21 src/plot_errcases_rate_densities_vs_radius.py

@@ -46,7 +46,7 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
         ax=bigax
 
     if isinstance(Λ_2,float) or isinstance(Λ_2,int):
-        fname = '../data/results_model_{:d}_Lambda2_{:.1f}.out'.format(
+        fname = '../data/numerics/results_model_{:d}_Lambda2_{:.1f}.out'.format(
                 model_number, Λ_2)
     else:
         fname = '../data/results_model_'+repr(model_number)+'.out'

### 2017-10-21 src/plot_rate_densities_vs_radius.py

@@ -33,7 +33,7 @@ def make_plot(model_number, logx=None, logy=None, withtext=None,
 
     # Make summary plot
     if isinstance(Λ_2,float) or isinstance(Λ_2,int):
-        fname = '../data/results_model_{:d}_Lambda2_{:.1f}.out'.format(
+        fname = '../data/numerics/results_model_{:d}_Lambda2_{:.1f}.out'.format(
                 model_number, Λ_2)
     else:
         fname = '../data/results_model_'+repr(model_number)+'.out'

### 2017-10-21 src/run_numerical_models.py

@@ -1,0 +1,24 @@
+'''
+Wrapper to `numerical_models.py`.
+
+Useful for running Model #3 with varying Λ_2 in parallel.
+'''
+
+import numpy as np
+from multiprocessing import Pool
+
+from numerical_models import numerical_transit_survey
+
+if __name__ == '__main__':
+
+    model_number = 3
+    quickrun = False
+    Λ_2_arr = np.arange(0., 0.5+0.05, 0.05)
+
+    paramlist = [(quickrun, model_number, Λ_2) for Λ_2 in Λ_2_arr]
+
+    with Pool(6) as p:
+        p.starmap(
+            numerical_transit_survey,
+            paramlist
+            )

### 2017-10-21 paper/figures.tex
paper: begin model #3 section
paper: add HJ correction vs free param

@@
+\newpage
 \begin{figure}
@@
 
+\begin{figure}[!tb]
+    \centering
+    \includegraphics[width=.6\textwidth]{figures/HJ_correction_inputrate_vs_HJratevalues.pdf}
+    \caption{
+    $\Lambda_i$ is the occurrence rate integrated over all possible phase 
+    space for the $i^{\rm th}$ system type. $\Lambda_2/\Lambda_0=1$ 
+    corresponds to an equal number of planets per secondary as per single star;
+    $\Lambda_2/\Lambda_0=0$ corresponds to secondaries not having any planets.
+    While the true HJ occurrence rate (marginalized over stellar multiplicity, 
+    Eq.~\ref{eq:occ_rate}) is highly dependent on $\Lambda_2$, 
+    the inferred rate hardly depends on whether secondaries have HJs.
+    }
+    \label{fig:HJ_correction_inputrate_vs_HJratevalues}
+\end{figure}
+
+\begin{figure}[!tb]
+    \centering
+    \includegraphics[width=.6\textwidth]{figures/HJ_correction_inputrate_vs_HJrate_correction_factor.pdf}
+    \caption{
+    Ratio of inferred to true HJ rates from 
+    Fig.~\ref{fig:HJ_correction_inputrate_vs_HJratevalues}, for Model \#3.
+    A plausible value for $\Lambda_2/\Lambda_0$ is $\approx 0.5$, 
+    corresponding to an underestimate by a factor of 1.1 (see text).
+    }
+    \label{fig:HJ_correction_inputrate_vs_HJrate_correction_factor}
+\end{figure}
+
 %\begin{figure}[!b]

### 2017-10-21 paper/main.tex
paper: add HJ correction vs free param

@@
 \shortauthors{LGB, KM, JNW}
-\shorttitle{Toy transit surveys}
+\shorttitle{Binarity and Occurrence Rates}
 
@@
 underestimate of the number of planets per star over most planetary radii.
-In particular, the 30\% underestimate applies to hot Jupiter occurrence rates, 
-which suggests that the discrepancy between rates measured by {\it Kepler} and 
-RV surveys might be attributable to the {\it Kepler} sample containing more 
-binaries.
+In particular, this underestimate applies to hot Jupiter occurrence rates, 
+which suggests that at least part of the discrepancy between rates measured by 
+{\it Kepler} and RV surveys might come from the {\it Kepler} sample 
+containing more binaries.
 

### 2017-10-21 paper/model_3.tex
paper: add HJ correction vs free param

@@
 \begin{cases}
-r^\delta & \text{for } r\geq 2R_\oplus \\
-{\rm constant} & \text{for } r\leq2R_\oplus.
+r^\delta & \text{for } r\geq 2r_\oplus \\
+{\rm constant} & \text{for } r\leq2r_\oplus.
 \end{cases}
@@
 Taking Fig.~\ref{fig:errcases_model_3_log} and counting the number of planets 
-per star with $r>8R_\oplus$, we can compare the true and inferred hot Jupiter 
+per star with $r>8r_\oplus$, we can compare the true and inferred hot Jupiter 
 occurrence rates.
@@
 \Lambda_1 = \Lambda_2$.
-Given that hot Jupiters are thought to be less common around less-massive 
-stars, it would be more sensible to consider $\Lambda_2<\Lambda_0$, while 
-letting single stars and primaries host planets at the same rate.
-We show the HJ multiplicative correction factor as a function of 
-$\Lambda_2/\Lambda_0$ in Fig BLAH.
+If hot Jupiters are less common around less-massive stars, it would be more 
+sensible to consider $\Lambda_2<\Lambda_0$, while letting single stars and 
+primaries host planets at the same rate.
+Therefore in Fig.~\ref{fig:HJ_correction_inputrate_vs_HJratevalues}, we let 
+$\Lambda_2$ vary, and show the resulting inferred and true hot Jupiter 
+($r>8r_\oplus$) rates.
+We see that the inferred rate is nearly independent of $\Lambda_2$~--~this is 
+because most ($<1/10$) secondaries are not searchable, and so their 
+completeness fraction is much smaller than that of primaries or single stars.
+This means far fewer detected hot Jupiters orbit secondaries, and so they 
+hardly affect the inferred rate.
+In contrast, the true rate is highly dependent on $\Lambda_2$, as one would 
+expect from Eq.~\ref{eq:occ_rate}..
+
+This means that the ``correction factor'' from an inferred rate to a true 
+rate, in this case for hot Jupiters but also more generally, depends strongly 
+on the intrinsic occurrence of the planet type of interest.
+This is illustrated visually in 
+Fig.~\ref{fig:HJ_correction_inputrate_vs_HJrate_correction_factor}, which 
+shows that for $\Lambda_2 = \Lambda_0$, the inferred rate is underestimated by 
+a factor of $\sim1.3$.
+However if $\Lambda_2 = 0$, the inferred rate is actually an overestimate, 
+by a factor of $\sim1.1$.
+Though the true ratio $\Lambda_2/\Lambda_0$ is unknown for hot Jupiters, a 
+reasonable estimate might be to imagine that only stars with mass 
+$>0.7M_\odot$ might be capable of hosting them.
+For our nominal mass ratio distribution, this is $48\%$ of secondaries, so a 
+reasonable interpretation of 
+Fig.~\ref{fig:HJ_correction_inputrate_vs_HJrate_correction_factor} would be 
+that an observer ignoring binarity might underestimate HJ occurrence rates by 
+a factor of $\sim1.1$.
+Assuming that RV surveys contain far fewer binary systems than transit 
+surveys, this suggests that the effect likely contributes towards the HJ rate 
+discrepancy.

### 2017-10-21 src/plot_HJ_correction_vs_LambdaTwo.py

@@ -52,7 +52,7 @@ if __name__ == '__main__':
     Λ_HJ_true_arr = np.array(Λ_HJ_true_arr)
     Λ_HJ_inferred_arr = np.array(Λ_HJ_inferred_arr)
 
-    X_correction = Λ_HJ_inferred_arr/Λ_HJ_true_arr
+    X_correction = Λ_HJ_true_arr/Λ_HJ_inferred_arr
 
     Λ_0 = 0.5
 
@@ -81,7 +81,7 @@ if __name__ == '__main__':
     ax.plot(Λ_2_arr/Λ_0, X_correction, marker='o')
 
     ax.set_xlabel('$\Lambda_2/\Lambda_0$', fontsize='large')
-    ax.set_ylabel('inferred HJ rate / true HJ rate', fontsize='large')
+    ax.set_ylabel('true HJ rate / inferred HJ rate', fontsize='large')
 
     outname = '../results/HJ_correction_inputrate_vs_HJrate_correction_factor'
     f.savefig(outname+'.pdf', bbox_inches='tight')

### 2017-10-24 src/numerical_models.py
add type-specific rate output
also updated HJ correction plot for clarity.

@@ -328,7 +328,10 @@ def numerical_transit_survey(
     outdf = pd.DataFrame(
             {'bin_left': edges[:-1],
              'true_Γ': true_dict['Γ'],
-             'inferred_Γ': inferred_dict['Γ']
+             'inferred_Γ': inferred_dict['Γ'],
+             'true_single_Γ': true_dict['single']['N_p']/N_0,
+             'true_primary_Γ': true_dict['primary']['N_p']/N_1,
+             'true_secondary_Γ': true_dict['secondary']['N_p']/N_2
             }
             )
 
@@ -493,7 +496,10 @@ def numerical_transit_survey(
         outdf = pd.DataFrame(
                 {'bin_left': edges[:-1],
                  'true_Γ': true_dict['Γ'],
-                 'inferred_Γ': inferred_dict['Γ']
+                 'inferred_Γ': inferred_dict['Γ'],
+                 'true_single_Γ': true_dict['single']['N_p']/N_0,
+                 'true_primary_Γ': true_dict['primary']['N_p']/N_1,
+                 'true_secondary_Γ': true_dict['secondary']['N_p']/N_2
                 }
                 )
 

### 2017-10-24 src/plot_HJ_correction_vs_LambdaTwo.py

@@ -31,7 +31,7 @@ if __name__ == '__main__':
 
     Λ_2_arr = np.arange(0, 0.5+0.05, 0.05)
 
-    Λ_HJ_true_arr, Λ_HJ_inferred_arr = [], []
+    Λ_HJ_true_arr, Λ_HJ_inferred_arr, Λ_HJ_true_primary_arr = [], [], []
 
     # compute the HJ correction factors for each Λ_2
     for Λ_2 in Λ_2_arr:
@@ -45,12 +45,15 @@ if __name__ == '__main__':
         inds = df['bin_left'] > lower_bound
         Λ_HJ_true = np.sum(df[inds]['true_Γ'])
         Λ_HJ_inferred = np.sum(df[inds]['inferred_Γ'])
+        Λ_HJ_true_primary = np.sum(df[inds]['true_primary_Γ'])
 
         Λ_HJ_true_arr.append(float(Λ_HJ_true))
         Λ_HJ_inferred_arr.append(float(Λ_HJ_inferred))
+        Λ_HJ_true_primary_arr.append(float(Λ_HJ_true_primary))
 
     Λ_HJ_true_arr = np.array(Λ_HJ_true_arr)
     Λ_HJ_inferred_arr = np.array(Λ_HJ_inferred_arr)
+    Λ_HJ_true_primary_arr = np.array(Λ_HJ_true_primary_arr)
 
     X_correction = Λ_HJ_true_arr/Λ_HJ_inferred_arr
 
@@ -61,10 +64,13 @@ if __name__ == '__main__':
 
     f,ax = plt.subplots(figsize=(4,4))
 
-    ax.plot(Λ_2_arr/Λ_0, Λ_HJ_true_arr*1e3, marker='o', label='true')
+    ax.plot(Λ_2_arr/Λ_0, Λ_HJ_true_arr*1e3, marker='o',
+        label='true (whole population)')
     ax.plot(Λ_2_arr/Λ_0, Λ_HJ_inferred_arr*1e3, marker='o', label='inferred')
+    ax.plot(Λ_2_arr/Λ_0, Λ_HJ_true_primary_arr*1e3, marker='o',
+        label='true (only primaries)')
 
-    ax.legend(loc='upper left',fontsize='medium')
+    ax.legend(loc='best',fontsize='medium')
 
     ax.set_xlabel('$\Lambda_2/\Lambda_0$', fontsize='large')
     ax.set_ylabel('HJ rate (planets per thousand stars)', fontsize='large')

### 2017-10-27 paper/discussion.tex
wordsmithed text

\section{Discussion}

\paragraph{Does a detected planet orbit the primary or secondary?}
Ciardi et al. (2015) studied the effects of stellar multiplicity on the 
planet radii derived from transit surveys.
They modeled the problem for {\it Kepler}\ objects of interest by matching a 
population of binary and tertiary companions to KOI stars, 
under the assumption that the KIC-listed stars were the primaries.
They then computed planet radius correction factors assuming that {\it 
Kepler}-detected planets orbited the primary or companion stars
with equal probability (their Sec. 5).
Under these assumptions, they found that any given planet's radius is on 
average underestimated by a multiplicative factor of 1.5.

Our models show that the uniform prior on planet assignment overstates 
binarity's population-level effects.
A planet orbiting the secondary does lead to extreme corrections, but these 
cases are rare outliers, because the searchable volume for secondaries is so 
much smaller than that for primaries.
Phrased in terms of the completeness, in our Model \#3 only $\sim 6\%$ of 
selected secondaries are searchable, compared to $\sim 60\%$ of selected 
primaries.
This means that when high-resolution imaging discovers a binary companion in 
system that hosts a detected transiting planet, the planet is much more likely 
to orbit the primary.
It also means that the average radius correction factor across the entire {\it 
Kepler} sample is smaller than that advertised by Ciardi et al.

\paragraph{On the utility for future occurrence rate measurements}
Though they will be difficult to distinguish from false positives, {\it TESS}\ 
is expected to discover over $10^4$ giant planets (Sullivan et al. 2015).
One possible use of this overwhelmingly large sample will to measure an
occurrence rate of short-period giant planets.
Our work indicates that if this measurement is to be more precise than $\sim 
30\%$, binarity cannot be neglected.

\paragraph{What about $\eta_\oplus$?}

\paragraph{What about the radius gap?}

\paragraph{What about {\it Kepler}?}

### 2017-10-27 paper/figures.tex
wordsmithed text

--- 2017-10-20 paper/figures.tex
@@
     \#1.
-    This model has fixed stars, fixed planets, twin binaries;
-    Note the break in the $x$ axis.
+    This model has fixed stars, fixed planets, and twin binaries.
     If the true planet radius is $r_p$, all planets 
@@
     Sec.~\ref{sec:model_1}).
+    To help illustrate the individual effects of errors, we 
+    separated them:
+    ``only wrong $N_\star$'' means the only error is an incorrectly assumed 
+    number of selected stars;
+    ``only wrong $Q$'' means the only error is an incorrectly assumed 
+    detection efficiency (including both miscalculated transit probabilities 
+    and 
+    fraction of selected stars that are searchable);
+    ``only wrong $r_a$'' means the only error is misinterpretation of 
+    planetary radii, due to both transit depth dilution and also wrongly 
+    assumed host star radii.
     }
@@
 
+\begin{figure}[!tb]
+    \centering
+    \includegraphics[width=.9\textwidth]{figures/HJ_correction_inputrate_vs_HJratevalues.pdf}
+    \caption{
+    $\Lambda_i$ is the occurrence rate integrated over all possible phase 
+    space for the $i^{\rm th}$ system type. $\Lambda_2/\Lambda_0=1$ 
+    corresponds to an equal number of planets per secondary as per single star;
+    $\Lambda_2/\Lambda_0=0$ corresponds to secondaries not having any planets.
+    In our Model \#3, though the true HJ occurrence rate 
+    (Eq.~\ref{eq:occ_rate}) is highly dependent on $\Lambda_2$, 
+    the inferred rate hardly depends on whether secondaries have HJs.
+    This means that the ``correction factor'' between the inferred rate and 
+    the true rate around single stars is underestimated by a factor of 
+    $\sim1.3$, independent of the HJ rate around secondaries.
+    }
+    \label{fig:HJ_correction_inputrate_vs_HJratevalues}
+\end{figure}
+
 %\begin{figure}[!b]

### 2017-10-27 paper/model_2.tex
wordsmithed text

@@
 magnitude-limited, $p({\rm secondary\ with\ }q)$ 
-is the product of the volume limited probability and a 
-Malmquist-like bias $(1+q^\alpha)^{3/2}$.
-In passing, we note that various authors (ourselves included) have incorrectly 
+is the product of the volume limited probability, $q^\beta$, and a 
+Malmquist-like bias, $(1+q^\alpha)^{3/2}$.
+We note that various authors (ourselves included) have incorrectly 
 used volume-limited binary distributions in Monte Carlo simulations of 

### 2017-10-27 paper/model_3.tex
wordsmithed text

--- 2017-10-21 paper/model_3.tex
@@
 \Lambda_1 = \Lambda_2$.
-Given that hot Jupiters are thought to be less common around less-massive 
-stars, it would be more sensible to consider $\Lambda_2<\Lambda_0$, while 
-letting single stars and primaries host planets at the same rate.
-We show the HJ multiplicative correction factor as a function of 
-$\Lambda_2/\Lambda_0$ in Fig BLAH.
+If hot Jupiters are less common around lower mass stars, it would be more 
+sensible to consider $\Lambda_2<\Lambda_0$, while letting single stars and 
+primaries host planets at the same rate.
+Therefore in Fig.~\ref{fig:HJ_correction_inputrate_vs_HJratevalues}, we let 
+$\Lambda_2$ vary, and show the resulting inferred and true hot Jupiter 
+($r>8r_\oplus$) rates.
+We see that the inferred rate is nearly independent of $\Lambda_2$~--~this is 
+because most ($<1/10$) secondaries are not searchable, and so their 
+completeness fraction is much smaller than that of primaries or single stars.
+This means far fewer detected hot Jupiters orbit secondaries, and so they 
+hardly affect the inferred rate.
+While the true rate across the entire population is highly dependent on 
+$\Lambda_2$, as one would expect from Eq.~\ref{eq:occ_rate}, the rate around 
+singles and primaries (green line) is independent of that around secondaries.
+
+This means that, in this case, the inferred rate should be revised upward by a 
+factor of $\sim1.3$, independent of the HJ rate around secondaries.
+Assuming that RV surveys are measuring the true rate around single stars (or 
+primaries), this suggests that the effect appreciably contributes to the 
+HJ rate discrepancy.

### 2017-10-28 src/plot_HJ_correction_vs_LambdaTwo.py
add eta earth assessment

@@ -3,7 +3,7 @@ Given that hot Jupiters are thought to be less common around less-massive
 stars, it would be more sensible to consider Λ_2 < Λ_0 , while letting single
 stars and primaries host planets at the same rate.
 
-This plots the HJ multiplicative correction factor as a function of Λ_2 /Λ_0.
+This plots the rates as a function of Λ_2 /Λ_0.
 '''
 # -*- coding: utf-8 -*-
 from __future__ import division, print_function
@@ -23,7 +23,6 @@ mpl.rcParams.update(pgf_with_custom_preamble)
 
 import matplotlib.pyplot as plt
 import numpy as np, pandas as pd
-import os
 
 if __name__ == '__main__':
 
@@ -59,8 +58,9 @@ if __name__ == '__main__':
 
     Λ_0 = 0.5
 
-    ##########################################
-    # First plot: the inferred rates vs the true values.
+    ######################################################
+    # First plot: the inferred rates vs the true values. #
+    ######################################################
 
     f,ax = plt.subplots(figsize=(4,4))
 
@@ -78,8 +78,9 @@ if __name__ == '__main__':
     outname = '../results/HJ_correction_inputrate_vs_HJratevalues'
     f.savefig(outname+'.pdf', bbox_inches='tight')
 
-    ##########################################
-    # Then plot: the inferred rates vs the ratio.
+    ###############################################
+    # Then plot: the inferred rates vs the ratio. #
+    ###############################################
 
     plt.close('all')
     f,ax = plt.subplots(figsize=(4,4))

### 2017-10-28 src/plot_print_earth_rate_analog.py

@@ -1,0 +1,97 @@
+'''
+What is the rate correction for earth analogs in Model #3?
+
+Print the result.
+
+Make the plot of rate as a function of Λ_2/Λ_0 as well.
+'''
+# -*- coding: utf-8 -*-
+from __future__ import division, print_function
+import matplotlib as mpl
+mpl.use('pgf')
+pgf_with_custom_preamble = {
+    'pgf.texsystem': 'pdflatex', # xelatex is default; i don't have it
+    'font.family': 'serif', # use serif/main font for text elements
+    'text.usetex': True,    # use inline math for ticks
+    'pgf.rcfonts': False,   # don't setup fonts from rc parameters
+    'pgf.preamble': [
+        '\\usepackage{amsmath}',
+        '\\usepackage{amssymb}'
+        ]
+    }
+mpl.rcParams.update(pgf_with_custom_preamble)
+
+import matplotlib.pyplot as plt
+
+import numpy as np, pandas as pd
+
+if __name__ == '__main__':
+
+    model_number = 3
+
+    Λ_0 = 0.5
+    Λ_2_arr = np.arange(0, 0.5+0.05, 0.05)
+
+    Λ_earth_true_arr, Λ_earth_inferred_arr, Λ_earth_true_primary_arr = [], [], []
+
+    # compute the rate for each Λ_2
+    for Λ_2 in Λ_2_arr:
+
+        savdir = '../data/numerics/'
+        fname = 'results_model_{:d}_Lambda2_{:.2f}.out'.format(
+                model_number, Λ_2)
+        df = pd.read_csv(savdir+fname)
+
+        for ix, bl in enumerate([0.5, 1.0]):
+
+            inds = (df['bin_left'] == bl)
+
+            Λ_earth_true = df[inds]['true_Γ']
+            Λ_earth_inferred = df[inds]['inferred_Γ']
+            Λ_earth_true_primary = df[inds]['true_primary_Γ']
+
+            Λ_earth_true_arr.append(float(Λ_earth_true))
+            Λ_earth_inferred_arr.append(float(Λ_earth_inferred))
+            Λ_earth_true_primary_arr.append(float(Λ_earth_true_primary))
+
+    Λ_earth_true_arr = np.array(Λ_earth_true_arr)
+    Λ_earth_inferred_arr = np.array(Λ_earth_inferred_arr)
[…]

### 2017-10-28 paper/analytics_v_numerics.tex
paper: add HJ correction vs free param
paper: η_earth plot and more discussion

@@
 \label{sec:numerical_methods}
-
-To progressively gain intuition, we study the following idealized transit 
-surveys:
-\begin{itemize}
-    \item Model \#1: fixed stars, fixed planets, twin binaries;
-    \item Model \#2: fixed planets and primaries, varying secondaries;
-    \item Model \#3: fixed primaries, varying planets and secondaries.
-\end{itemize}
 
@@
 Though analytic or semi-analytic solutions exist for each model, beyond Model 
-\#2 the equations become burdensome. To preserve simplicity, we develop a 
+\#2 the equations become burdensome. To maintain simplicity, we develop a 
 Monte Carlo approach, which works as follows.
@@
 Most importantly, these parameters include the binary fraction, and the true 
-planet occurrence rates around single stars, primaries, and secondaries.
+planet occurrence rates around single stars, primaries, and secondaries (the 
+$\Lambda_i$'s, to be defined in Sec.~\ref{sec:analytic_preliminaries}).
 
@@
 size, period, etc.).
-Assuming homogeneously distributed stars, we will show 
-(Sec.~\ref{sec:model_1}) that this is equivalent to the
+Assuming homogeneously distributed stars, this is equivalent to the
 ratio of the searchable to selected volumes.
@@
 formulae that account for both dilution and the misclassification of stellar
-radii. We assume that the observer thinks all transits are around primaries.
+radii. We assume that the observer think all transits are around single 
+stars.
 
@@
 stars.
-
 The simplest realization of this scheme is described analytically in 

### 2017-10-28 paper/conclusion.tex
paper: η_earth plot and more discussion

\section{Conclusion}
\label{sec:conclusion}

This work presented three simple models for the effects of binarity on 
occurrence rates measured by transit surveys.

### 2017-10-28 paper/discussion.tex
paper: η_earth plot and more discussion

@@
 \section{Discussion}
+\label{sec:discussion}
 
@@
 
-Our models show that the uniform prior on planet assignment overstates 
+Our models show that assuming a detected planet has equal probability of 
+orbiting the primary or secondary leads to an overstatement of
 binarity's population-level effects.
@@
 This means that when high-resolution imaging discovers a binary companion in 
-system that hosts a detected transiting planet, the planet is much more likely 
-to orbit the primary.
-It also means that the average radius correction factor across the entire {\it 
-Kepler} sample is smaller than that advertised by Ciardi et al.
+system that hosts a detected transiting planet, the planet is overwhelmingly 
+more likely to orbit the primary.
+The average radius correction factor across the entire {\it Kepler} sample is 
+smaller than that Ciardi et al's factor of 1.5.
+This statement is independent of the fact that planets are often confirmed to 
+orbit the primary by inferring the stellar density from the transit duration.
 
@@
 
-\paragraph{What about $\eta_\oplus$?}
+\paragraph{What about {\it Kepler}?}
+Barclay \& Collaborator (in preparation) have performed the exercise 
+of taking stars selected by the {\it Kepler}\ team, pairing them with a 
+population of secondaries, injecting a realistic distribution of planet radii, 
+and then comparing the inferred occurrence rates with the true ones.
+In their model, they find that binarity leads to an inferred rate of 
+Earth-sized planets $\approx 10\%$ less than the true rate.
+In our Model \#3, if all $\Lambda_i$'s are equal (a plausible assumption in 
+the lack of evidence to the contrary), the underestimate is by 16\%.
 
 \paragraph{What about the radius gap?}
-
-\paragraph{What about {\it Kepler}?}

### 2017-10-28 paper/figures.tex
paper: η_earth plot and more discussion

@@
     \includegraphics[width=.7\textwidth]{figures/errcases_rate_density_vs_radius_model_2.pdf}
-    %\includegraphics[width=1.0\textwidth]{figures/errcases_rate_density_vs_radius_logs_model_2.pdf}
     \caption{
@@
     \centering
-    %\includegraphics[width=1.0\textwidth]{figures/errcases_rate_density_vs_radius_model_2.pdf}
     \includegraphics[width=.7\textwidth]{figures/errcases_rate_density_vs_radius_logs_model_2.pdf}
@@
     $\sim1.3$, independent of the HJ rate around secondaries.
+    The ``HJ rate'' is the summed rate from     
+    Fig.~\ref{fig:errcases_model_3_log} above $8r_\oplus$.
     }
     \label{fig:HJ_correction_inputrate_vs_HJratevalues}
+\end{figure}
+
+\begin{figure}[!tb]
+    \centering
+    \includegraphics[width=.9\textwidth]{figures/earth_inputrate_vs_etaearthratevalues.pdf}
+    \caption{
+        Same as Fig.~\ref{fig:HJ_correction_inputrate_vs_HJratevalues}, but 
+        for Earth-sized planets.
+        The absolute values given on the $y$-axis found by summing the rate 
+        from Fig.~\ref{fig:errcases_model_3_log} for planetary radii from 
+        $0.5$ to $1.5r_\oplus$ (this is a toy model, and they do not reflect 
+        an actual determination of $\eta_\oplus$).
+        The relative values show that the inferred rate for Earths is roughly 
+        independent of the occurrence rate (integrated over all radii) around 
+        secondaries.
+        However, it is systematically lower than the true rate around single 
+        and primary stars, by $\approx 20\%$.
+    }
+    \label{fig:earth_inputrate_vs_etaearthratevalues}
 \end{figure}

### 2017-10-28 paper/generalities.tex
paper: η_earth plot and more discussion

@@
 \subsection{Analytic Preliminaries}
+\label{sec:analytic_preliminaries}
 
@@
 The occurrence rate $\Lambda$ is found by integrating the rate density over a 
-specified volume of phase space.
+specified volume of phase space (e.g., Youdin 2011).
 
@@
 $\Lambda_i$ is the occurrence rate integrated over all possible phase space  
-for the $i^{\rm th}$ system type, $p_i(\vec{x})$ is the joint 
+for the $i^{\rm th}$ system type, and $p_i(\vec{x})$ is the joint 
 probability density function so that $\Gamma_i(\vec{x}) = \Lambda_i 
-p_i(\vec{x})$, 
-and the weights are given by 
+p_i(\vec{x})$.
+The weights are given by 
 \begin{equation}
@@
 The detection efficiency includes the geometric transit probability, as well 
-as any incompleteness effects. Foreman-Mackey et al. (2014) discuss how this 
+as any incompleteness effects.
+Foreman-Mackey et al. (2014) discuss how this 
 is calculated in practice.

### 2017-10-28 paper/introduction.tex
paper: η_earth plot and more discussion

\section{Introduction}

{\bf TODO} -- how should we open this? Should I include less on the HJ rate 
thing below?

\paragraph{Binarity and Occurrence Rates}

\paragraph{The Hot Jupiter Rate Discrepancy}

Radial velocity surveys of nearby bright Sun-like stars have reported hot 
Jupiter ($a<0.1\,{\rm AU}, P\lesssim 10\,{\rm day}$) occurrence rates of
$12\pm 1$, $15\pm 6$, $8.9 \pm 3.6$, and $12.0 \pm 3.8$ per thousand stars
(Marcy et al 2005, Cumming et al 2008, Mayor et al 2011, Wright et al 2012 
respectively).

In transit surveys, OGLE-III reported an absolute rate of $3.1^{+ 
4.3}_{-1.8}$ hot Jupiters with $P<5\,{\rm days}$ per thousand stars (Gould et 
al 2006).
The SuperLupus survey quoted $10^{+27}_{-8}$ per thousand stars, allowing 
planets with periods less than 10 days (Bayliss \& Sackett 2011).
Through the {\it Kepler}\ survey, Howard et al. (2012) later reported a rate 
of 
$4 \pm 1$ per thousand Sun-like stars.
This rate was for $P<10\,{\rm day}, 8-32r_\oplus$ planets, and 
was specific to their ``solar subset''\footnote{The ``solar subset'' was 
defined as $4100<T_{\rm eff}/{\rm K}<6100$, {\it Kepler}\ 
magnitude $<15$, $4.0 < \log g < 4.9$, and only took planets with measured 
signal to noise $>10$.
}.
Expanding their sample to fainter magnitudes, they quoted a rate of $5 \pm 
1$.
Expanding down to $r_p>5.6r_\oplus$, to avoid excluding hot Jupiters reported 
by RV surveys, they reported $7.6 \pm 1.3$.
Recent work by Petigura et al. using the updated parameters of the 
California-{\it Kepler}\ Survey has found a rate of $X.X \pm 
Y.Y$ (2017, in preparation)

The trend is that hot Jupiter occurrence rates measured by transit 
surveys seem to be marginally lower than those found by radial velocity 
surveys.
While the actual discrepancy is of sub-$3\sigma$ significance,
one reason to expect a difference in the rates is that the populations have 
different metallicities.
As originally argued by Gould et al. (2006), the RV sample is biased towards 
metal-rich stars, which have been measured by RV surveys to preferentially 
host more giant planets (Santos et al 2004, Fischer and Valenti 2005).
The {\it Kepler}\ sample specifically has been measured to be more metal poor 
than the local neighborhood, with a mean metallicity of $[{\rm M/H}]_{\rm 
mean}\approx -0.05$ (Dong et al., 2014; Guo et al., 2017).
Studying the problem in detail, Guo et al. recently argued that the 
metallicity difference could account for a $\approx 0.1\%$ difference in the 
measured rates between the CKS and {\it Kepler}\ samples~--~not a $\approx 
0.5\%$ difference.
These authors concluded that ``other factors, such as binary contamination and 
imperfect stellar properties'' must also be at play (Guo et al., 2017).

Aside from surveying stars of different metallicities, radial velocity and 
transit surveys differ in how they treat binarity.
Radial velocity surveys typically reject both visual and spectroscopic binaries
({\it e.g.}, Wright et al. 2012).
Transit surveys typically observe binaries, but the question of whether they 
were searchable to begin with is usually left for later interpretation.
In spectroscopic follow-up of transiting candidates, the prevalence of 
astrophysical false-positives may also lead to a bias against confirmation of 
transiting planets in binary systems.

Ignoring these complications, in this work we focus on whether
binarity might intrinsically bias occurrence rate measurements, simply 
through its effects on the number of searchable stars and the apparent radii 
of detected planets.

To progressively gain intuition, we study the following idealized transit 
surveys:
\begin{itemize}
    \item Model \#1: fixed stars, fixed planets, twin binaries;
    \item Model \#2: fixed planets and primaries, varying secondaries;
    \item Model \#3: fixed primaries, varying planets and secondaries.
\end{itemize}
In Sec.~\ref{sec:numerical_methods}, we introduce our numerical approach 
to the problem, and in Sec.~\ref{sec:analytic_preliminaries} we clarify our 
terminology.
We present the analytic and numerical results in 
Secs.~\ref{sec:model_1}-\ref{sec:model_3}, where each subsection corresponds 
to each model above.
We interpret these calculations throughout, and in 
Sec.~\ref{sec:discussion} discuss their relevance to various questions in 
the interpretation of transit survey occurrence rates.

### 2017-10-28 paper/model_2.tex
paper: η_earth plot and more discussion

@@
     assuming that the host star is the primary.
-    \item The true stellar radii $R$ are all thought to be $R_\star$. In 
-    ``reality'', this only holds for single stars and primaries.
+    \item The completeness fraction is miscalculated for any stars in binary 
+    systems.
     \item The selected and searchable stars are miscounted.
@@
 
-However, since our main interest is in understanding the qualitative behavior 
-of the solutions, we focus instead on a few limiting cases.
+Since our main interest is in understanding the qualitative behavior 
+of the solutions, we focus on a few analytic limiting cases, and then 
+proceed numerically.
 
@@
 We also run this model in our numerical simulation.
-The results are shown in Figs~\ref{fig:errcases_model_2_linear} 
+The results are shown in Figs.~\ref{fig:errcases_model_2_linear} 
 and~\ref{fig:errcases_model_2_log}.

### 2017-10-28 paper/model_3.tex
paper: η_earth plot and more discussion

--- 2017-10-21 paper/model_3.tex
@@
 underestimated over all radii.
+
+\paragraph{Hot Jupiter Occurrence Rates}
 Taking Fig.~\ref{fig:errcases_model_3_log} and counting the number of planets 
-per star with $r>8R_\oplus$, we can compare the true and inferred hot Jupiter 
+per star with $r>8r_\oplus$, we can compare the true and inferred hot Jupiter 
 occurrence rates.
-Under the previously described assumptions, the true rate is 9.1 hot Jupiters 
+Under the above assumptions, the true rate is 9.1 hot Jupiters 
 per thousand stars.
@@
 This means that the inferred rate underestimates the true rate by a 
-multiplicative factor of $\sim 1.3$.
-This number is suggestively close to the claimed discrepancy between 
-hot Jupiter occurrence rates measured by radial velocity and transit surveys 
-(Wright et al. 2012).
+multiplicative factor of $\sim\!1.3$.
 
@@
 \Lambda_1 = \Lambda_2$.
-Given that hot Jupiters are thought to be less common around less-massive 
-stars, it would be more sensible to consider $\Lambda_2<\Lambda_0$, while 
-letting single stars and primaries host planets at the same rate.
-We show the HJ multiplicative correction factor as a function of 
-$\Lambda_2/\Lambda_0$ in Fig BLAH.
+If hot Jupiters are less common around lower mass stars, it would be more 
+sensible to consider $\Lambda_2<\Lambda_0$, while letting single stars and 
+primaries host planets at the same rate.
+Therefore in Fig.~\ref{fig:HJ_correction_inputrate_vs_HJratevalues} we let 
+$\Lambda_2$ vary, and show the resulting inferred and true hot Jupiter 
+($r>8r_\oplus$) rates.
+The result is that the inferred rate is nearly independent of 
+$\Lambda_2$~--~this is because most ($<1/10$) secondaries are not searchable, 
+and so their completeness fraction is much smaller than that of primaries or 
+single stars.
+This means far fewer detected hot Jupiters orbit secondaries, and so they 
+hardly affect the inferred rate.
+While the ``true rate'' across the entire population is highly dependent on 
+$\Lambda_2$, as one would expect from Eq.~\ref{eq:occ_rate}, the rate around 
+singles and primaries (green line) is independent of that around secondaries.
+Assuming that RV surveys are measuring the true rate around single stars (or 
+primaries), this suggests that binarity might contributes to the 
+HJ rate discrepancy at the $\sim 0.2\%$ level, independent of the HJ rate 
+around secondaries.
+
+\paragraph{The Rate of Earth Analogs}
+The rate (density) of Earth-like planets orbiting Sun-like stars has 
+been independently measured by Youdin, Petigura, Dong \& Zhu, 
+Foreman-Mackey et al., and Burke et al., (2011, 2013, 2013, 2014, and 2015 
+respectively).
+These efforts have found that the one-year terrestrial planet occurrence rate 
+varies between $\approx 0.03$ and $\approx 1$ per Sun-like star, depending on 
+assumptions that are made when retrieving the rate (Burke et al. 2015's 
+Fig.~17).
+
+Our model does not explicitly include the rate density's period-dependence, 
+because stellar binarity does not appreciably bias the period-dependence of 
+occurrence rates measured by transit surveys\footnote{
+    Note that stellar binarity does bias the {\it intrinsic} planet 
+    occurrence as a function of planetary and binary periods. This is expected 
+    from dynamical stability limits in $\geq$3 body systems, and has been 
+    observationally confirmed (theory by Holman \& Wiegert 1999, and others 
+    including Gongjie; confirmation from Wang et al 2014a, 2014b, Kraus et al 
+    2014). 
+    However our statement is that inferred rates as a function of planet     
+    period should be negligibly affected by this, given the geometric bias 
+    against long-period transit detections, and the fact that the period 
+    distribution of solar-type binaries peaks at $\approx 100\,{\rm years}$ 
+    (Raghavan et al 2010, Fig.~13).
+}.
+Instead, it allows us to evaluate the difference in the apparent and true rate 
+as a function of radius 
+(Fig.~\ref{fig:errcases_model_3_log}).
+At Earth's radius, the result is that the inferred rate is $0.84\times$ the 
+true rate around single stars, assuming that the $\Lambda_i$'s are equal.
+Similar to the above case of the hot Jupiters, if we vary the true $\Lambda_2$ 
+while keeping $\Lambda_0 = \Lambda_1$, the ratio of the inferred to true rate 
+around single stars hardly 
+changes; $(\Lambda_{\rm inferred}/\Lambda_1)_{r=r_\oplus} \approx 0.135$, to 
+within a few percent, independent of $\Lambda_2$ (see 
+Fig.~\ref{fig:earth_inputrate_vs_etaearthratevalues}).
+The ratio of the inferred to the true rate, $(\Lambda_{\rm 
+inferred}/\Lambda)_{r=r_\oplus}$ varies substantially, but by at most $50\%$ 
+in the (unrealistic) limiting case that secondaries do not host planets.

### 2017-10-28 paper/conclusion.tex
paper: add intro/disc/conclusion/HJ table

@@
 occurrence rates measured by transit surveys.
+The most realistic of these models (Model \#3) suggests that binarity leads to 
+underestimates of occurrence rates at the $\lesssim 30\%$ relative 
+error level.
+The effects suggest that hot Jupiter rates measured by transit surveys are 
+biased to infer $\approx 2$ fewer hot Jupiters per thousand single stars.
+Our third model also indicates that binarity's effects on the measured 
+occurrence rates of Earth-sized planets are far smaller than current 
+systematic uncertainties.
+Though our models are simplistic, their agreement with Barclay \& 
+Collaborator's recent detailed simulations indicate that they capture the 
+important ingredients.

### 2017-10-28 paper/discussion.tex
paper: add intro/disc/conclusion/HJ table

@@
 This means that when high-resolution imaging discovers a binary companion in 
-system that hosts a detected transiting planet, the planet is overwhelmingly 
+system that hosts a detected transiting planet, the planet is much
 more likely to orbit the primary.
-The average radius correction factor across the entire {\it Kepler} sample is 
-smaller than that Ciardi et al's factor of 1.5.
 This statement is independent of the fact that planets are often confirmed to 
@@
 the lack of evidence to the contrary), the underestimate is by 16\%.
-
-\paragraph{What about the radius gap?}

### 2017-10-28 paper/figures.tex
paper: add intro/disc/conclusion/HJ table

@@
     This means that the ``correction factor'' between the inferred rate and 
-    the true rate around single stars is underestimated by a factor of 
-    $\sim1.3$, independent of the HJ rate around secondaries.
+    the true rate around single stars is underestimated by a multiplicative 
+    factor of $\approx1.3$, independent of the HJ rate around secondaries.
     The ``HJ rate'' is the summed rate from     

### 2017-10-28 paper/introduction.tex
paper: add intro/disc/conclusion/HJ table

@@
 
-{\bf TODO} -- how should we open this? Should I include less on the HJ rate 
-thing below?
+An astronomer who does not believe in stellar multiplicity wants to measure the
+mean number of planets of a certain type per star of a certain type.
+They perform a signal-to-noise limited transit survey and detect $N_{\rm det}$ 
+transit signals that appear to be from planets of the desired type.  They 
+calculate the apparent number of searchable stars, $N_\star$.  ``Searchable 
+stars'' are the stars for which 
+planets are observed with 100\% detection efficiency. 
+Correcting for the geometric 
+transit probability $f_{\rm g}$, they compute an apparent occurrence rate 
+$\Lambda_{\rm a}$:
 
-\paragraph{Binarity and Occurrence Rates}
+\begin{equation}
+\Lambda_{\rm a} = \frac{N_{\rm det}}{N_\star} \times \frac{1}{f_g}.
+\end{equation}
+
+There are many potential pitfalls.  Some genuine transit signals can be missed
+by the detection pipeline.  Some apparent transit signals are spurious, from
+noise fluctuations, failures of `detrending', or instrumental effects.  Stars
+and planets can be misclassified due to statistical and systematic errors in
+the measurements of their properties.  Poor angular resolution causes false
+positives due to blends with background eclipsing binaries. {\it Et cetera}.
+
+Here we focus on problems that arise from the fact that many stars exist in 
+multiple star systems.
+For simplicity, we only consider binaries, and we assume that they are all 
+spatially unresolved.
+
+An immediate complication is that, due to dynamical stability or some 
+aspect of planet formation, the occurrence rate of planets may differ 
+between binary and single-star systems.
+If ``occurrence rate'' is defined purely as the mean number of planets within 
+set radius and period bounds per star in a given mass interval, it must 
+implicitly marginalize over stellar multiplicity.
+This means marginalizing over ``occurrence rates in single star systems'', 
+``occurrence rates about primaries'', and
+``occurrence rates about secondaries'' (see {\it e.g.,} Wang et al., 
+2015).
+
+Outside of astrophysical differences, there are observational biases.
+A given apparently-searchable star may truly be a single star of the desired 
+type. If it is binary, it may contain two, one, or no searchable stars of the 
+desired type, which affects $N_\star$.
+There may be systematic errors in stellar parameter estimates of such 
+systems, which in this work we neglect.
+One might also imagine an apparently-searchable star which in no stellar 
+component is of the desired type. We will not consider errors of this type.  
+We will assume that all the apparently-searchable stars are either single 
+stars of the desired type, or binaries in which either the primary or else 
+both components are of the desired type.
+
+When counting the detected signals from planets of the desired 
+type, $N_{\rm det}$, ignoring binarity leads to errors in planet radius 
+classification, and the assumed survey completeness.
+Detected planets in binary systems will have underestimated radii because of 
+the diluting flux from the companions, and possibly because they are assumed 
+to orbit the wrong star ({\it e.g.}, Furlan et al. 2017).
+The number of selected stars should change dependent on the planet radius 
+and period for which the rate measurement is desired, and on the survey's 
+achieved photometric precision.
+This mean our naive astronomer assumes the fraction of their selected stars 
+which are searchable is 1 (this is their ``assumed completeness'').
+Binarity confuses this, because the ability to search for planets 
+in binary systems depends on which stellar component hosts the planet, and on 
+how much dilution affects the measurement.
+
+\paragraph{How Has Binarity Been Considered in Occurrence Rate Measurements?}
+Many authors have reported planet occurrence rates from transit 
+survey data\footnote{
+    An online list of occurrence rate papers is maintained at 
+    \url{https://exoplanetarchive.ipac.caltech.edu/docs/occurrence_rate_papers.html}
+}.
+With regard to binarity, the typical approach has been to ignore the 
+issue\footnote{Deacon et al., 2015 are an exception to the rule.}.
+For instance, in a laudable study by Burke et al. (2015), they
+\begin{quote}
+``treat the pipeline completeness as having no uncertainty 
+due to the incomplete understanding of the stellar parameters, eccentricity, 
+and stellar binarity.''
+\end{quote}
+Though they do ``explore sensitivity in the derived planet occurrence rates 
+to alternative assumptions for the stellar parameters and non-zero 
+eccentricities'', they do not for binarity.
+
+Of course, on a system-by-system level stellar multiplicity affects the 
+interpretation of planet candidates. High resolution imaging 
+campaigns have consequently been undertaken to measure the multiplicity of 
+almost all {\it Kepler}\ Objects of Interest 
+(Howell et al. 2011; Adams et al. 2012, 2013; Horch et al. 2012, 2014; 
+Lillo-Box et al. 2012, 2014; Dressing et al. 2014; Law et al. 2014; Cartier et 
+al. 2015; Everett et al. 2015; Gilliland et al. 2015; Wang et al. 2015a, 
+2015b; Baranec et al. 2016).
+The results of these programs have been collected by Furlan et al. 
+(2017), and they represent an important advance in understanding the KOI 
+sample's multiplicity statistics.
+In particular, they can be immediately applied to rectify binarity's influence 
+on the mass-radius diagram (Furlan \& Howell 2017).
+
+For {\it Kepler}-derived occurrence rates, the high resolution imaging 
+campaigns have not yet come full circle to observe a comparison sample of 
+non-KOI host stars.
+The most recent rate studies have thus used Furlan et al. (2017)'s 
+catalog to test the effects of removing KOI hosts with known companions, which 
+is an important step towards reducing contamination in the ``numerator'' of 
+the occurrence rate (Fulton et al. 2017).
+However, without an understanding of the multiplicity statistics of the 
+non-KOI hosting stars that are assumed to be searchable, the 
+true completeness, the true number of searchable stars, and thus the true 
+occurrence rates will remain uncertain.
 
 \paragraph{The Hot Jupiter Rate Discrepancy}
-
-Radial velocity surveys of nearby bright Sun-like stars have reported hot 
-Jupiter ($a<0.1\,{\rm AU}, P\lesssim 10\,{\rm day}$) occurrence rates of
-$12\pm 1$, $15\pm 6$, $8.9 \pm 3.6$, and $12.0 \pm 3.8$ per thousand stars
-(Marcy et al 2005, Cumming et al 2008, Mayor et al 2011, Wright et al 2012 
-respectively).
-
-In transit surveys, OGLE-III reported an absolute rate of $3.1^{+ 
-4.3}_{-1.8}$ hot Jupiters with $P<5\,{\rm days}$ per thousand stars (Gould et 
-al 2006).
-The SuperLupus survey quoted $10^{+27}_{-8}$ per thousand stars, allowing 
-planets with periods less than 10 days (Bayliss \& Sackett 2011).
-Through the {\it Kepler}\ survey, Howard et al. (2012) later reported a rate 
-of 
-$4 \pm 1$ per thousand Sun-like stars.
-This rate was for $P<10\,{\rm day}, 8-32r_\oplus$ planets, and 
-was specific to their ``solar subset''\footnote{The ``solar subset'' was 
-defined as $4100<T_{\rm eff}/{\rm K}<6100$, {\it Kepler}\ 
-magnitude $<15$, $4.0 < \log g < 4.9$, and only took planets with measured 
-signal to noise $>10$.
-}.
-Expanding their sample to fainter magnitudes, they quoted a rate of $5 \pm 
-1$.
-Expanding down to $r_p>5.6r_\oplus$, to avoid excluding hot Jupiters reported 
-by RV surveys, they reported $7.6 \pm 1.3$.
-Recent work by Petigura et al. using the updated parameters of the 
-California-{\it Kepler}\ Survey has found a rate of $X.X \pm 
-Y.Y$ (2017, in preparation)
-
-The trend is that hot Jupiter occurrence rates measured by transit 
-surveys seem to be marginally lower than those found by radial velocity 
-surveys.
-While the actual discrepancy is of sub-$3\sigma$ significance,
-one reason to expect a difference in the rates is that the populations have 
-different metallicities.
+There is at least one context in which measurement of absolute 
+occurrence rates may already be showing the signatures of binarity.
+Hot Jupiter occurrence rates measured by transit surveys ($\approx 0.5\%$) are 
+marginally lower than those found by radial velocity surveys ($\approx 1\%$; 
+see Table~\ref{tab:hj_rates}).
+Though the discrepancy is of weak statistical significance ($<3\sigma$),
+one plausible explanation for the difference is that the populations have 
+distinct metallicities.
 As originally argued by Gould et al. (2006), the RV sample is biased towards 
@@
 0.5\%$ difference.
-These authors concluded that ``other factors, such as binary contamination and 
-imperfect stellar properties'' must also be at play (Guo et al., 2017).
+Guo et al. concluded that ``other factors, such as binary contamination and 
+imperfect stellar properties'' must also be at play.
 
-Aside from surveying stars of different metallicities, radial velocity and 
+Aside from surveying stars of varying metallicities, radial velocity and 
 transit surveys differ in how they treat binarity.
@@
 were searchable to begin with is usually left for later interpretation.
-In spectroscopic follow-up of transiting candidates, the prevalence of 
+In spectroscopic follow-up of candidate transiting planets, the prevalence of 
 astrophysical false-positives may also lead to a bias against confirmation of 
@@
 Ignoring these complications, in this work we focus on whether
-binarity might intrinsically bias occurrence rate measurements, simply 
+binarity might intrinsically bias transit survey occurrence rates, simply 
 through its effects on the number of searchable stars and the apparent radii 

### 2017-10-28 paper/main.tex
paper: η_earth plot and more discussion
paper: add intro/disc/conclusion/HJ table

--- 2017-10-19 paper/main.tex
@@
 \documentclass[12pt,modern]{aastex61}
-\usepackage{graphicx}
+\usepackage{graphics,graphicx}
+\usepackage{hyperref}
 \usepackage{amssymb}
 \usepackage{amsmath}
-\usepackage{todonotes}
+\usepackage{comment}
 
-\shortauthors{LGB, KM, JNW}
-\shorttitle{Toy transit surveys}
+\shortauthors{LGB, JNW}
+\shorttitle{Binarity and Occurrence Rates}
 
@@
 
-\title{ Binarity's effects on occurrence rates measured by transit surveys}
+\title{ The effects of binarity on planet occurrence rates measured by transit 
+surveys}
 
 \author{
-LGB, KM, JNW
+L. G. Bouma, J. N. Winn
 }
 
-\input{generalities}
-\input{model_1}
-\input{model_2}
+\begin{abstract}
+
+This work aims to clarify the biases that stellar binarity introduces
+to occurrence rates inferred from transit surveys.
+In general, stellar multiplicity leads to diluted planetary 
+radii, overestimated detection efficiencies, and an undercounted number of 
+selected stars (though possibly an overcounted number of searched stars).
+These effects skew occurrence rate measurements in different directions, and 
+we develop simple models that allow us to understand the crucial effects.
+For models in which all stellar systems are either single or twin binaries, 
+and all planets are identical, we find that ignoring binarity leads to a
+two-to-threefold underestimate of the occurrence at the true radius.
+We believe this represents an upper bound on binarity's effects.
+Using more realistic models for the stellar population and planetary 
+radii, we find that ignoring binarity leads to a $10$\,-\,$30\%$ 
+underestimate of the number of planets per star, depending on planet radii.
+For the occurrence of Earth-sized planets, in our most realistic model the 
+underestimate is by $\approx 15\%$~--~at present far smaller than other 
+systematic uncertainties.
+For hot Jupiters, we find that the inferred occurrence rate is 
+$\approx 1.3\times$ smaller than the true rate around single stars.
+We suggest that this latter effect contributes to the discrepant hot Jupiter 
+rates measured by {\it Kepler}\ and the California group's respective surveys.
+\end{abstract}
+
+\input{introduction}
+\input{methods}
+\input{results}
+\input{discussion}
+\input{conclusion}
+
+\acknowledgements
+It was a pleasure to share discussions with Kento Masuda, who pointed us in 
+this direction, and helped clarify that something like this was worth studying.
+
+\input{figures.tex}
+\input{hj_rates.tex}
 

### 2017-10-28 paper/model_3.tex
paper: add intro/disc/conclusion/HJ table

@@
 changes; $(\Lambda_{\rm inferred}/\Lambda_1)_{r=r_\oplus} \approx 0.135$, to 
-within a few percent, independent of $\Lambda_2$ (see 
+within a few percent, independent of $\Lambda_2$ (shown in 
 Fig.~\ref{fig:earth_inputrate_vs_etaearthratevalues}).

### 2017-10-28 memos/introduction_v0.tex
memos: add draft introductions

--- 2017-10-05 paper/ms.tex
@@
-%\documentclass[12pt,preprint]{aastex}
-%\documentclass[preprint2,12pt]{aastex}
-\documentclass{emulateapj}
-%\usepackage{apjfonts}
-\usepackage{graphicx}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{todonotes}
-\usepackage{comment}
-
-\shortauthors{LGB, KM, JNW}
-\shorttitle{Binary Biases \& Toy Transit Surveys}
-
-\begin{document}
-
-% ------------------------------------------------------------------------
-% New commands
-%
-\def\ltsima{$\; \buildrel < \over \sim \;$}
-\def\lsim{\lower.5ex\hbox{\ltsima}}
-\def\gtsima{$\; \buildrel > \over \sim \;$}
-\def\gsim{\lower.5ex\hbox{\gtsima}}
-\def\tess{{\it TESS} }
-\def \teff {T_{\rm eff}}
-\def \phir {\Phi_{\rm R}}
-\def \fov {24$^{\circ}$}
-\def \pixsz {21.1''}
-\def \aeff {69.1 cm$^2$ }    
-\def \epd {105 mm}
-
-\def \kepler {{\it Kepler}}
-
-% -------------------------------------------------------------------------
-%
-
-\bibliographystyle{apj}
-
-\title{ The Effects of Stellar Binarity on Transit Survey Occurrence 
-Rates }
-
-\author{
-  LGB, KM, JNW
-}
-
-% \journalinfo{Draft version}
-\slugcomment{Draft version}
-
-%\altaffiltext{1}{Princeton University}
-
-\begin{abstract}
-
-What errors does ignoring binarity introduce on the occurrence rates derived 
-from transit surveys?
-
-\begin{comment}
-First, we show that in a universe where all stars have fixed properties
-(except that some are twin binaries), the results of a search for planets of a 
-given radius and orbital period can be expressed analytically.
-
-We then allow the properties of the secondary in binary systems to vary. A 
-distribution of light ratios introduces a detection bias towards 
-equal-mass binaries in magnitude limited samples.
-We show that dependent upon planet occurrence's functional dependence on host 
-star mass, as well as the specifics of the assumed survey completeness, the 
-occurrence rate correction factor $X_\Gamma \equiv \Gamma_{\rm 
-true}/\Gamma_{\rm obs}$ ranges from $0.6-1.6$.
-
-Finally, we perform a synthetic transit survey analogous to {\it Kepler}. We 
-show that if we assume every star in this {\it Kepler} survey analog is single, 
-we underestimate occurrence rates for a given planet radius and period by 
-a factor of $\approx 1.4$, or a percentage error relative to the true 
-occurrence rate $\Gamma_{\rm true}$ of $\delta\Gamma_{\rm 
-true}=-28\%$. This is relevant for measuring accurate values of the rate at 
-which Earth-like planets exist around Sun-like stars.
-\end{comment}
-
-\end{abstract}
-
-\keywords{planets and satellites:\ detection}
-
 \section{Introduction}
 
-An astronomer who does not believe in binaries wants to measure the
-occurrence of planets of a certain type around stars of a certain type.
-In other words, they wish to find the mean number of planets within specific 
-planetary bounds orbiting the stars in a given volume 
-of stellar phase-space $(\Delta R_p, \Delta P, \Delta M_\star,\ldots$).
+An astronomer who does not believe in stellar multiplicity wants to measure the
+mean number of planets of a certain type per star of a certain type.
 They perform a signal-to-noise limited transit survey and detect $N_{\rm det}$ 
@@
 calculate the number of stars $N_\star$ that seem to be of the desired type, 
-and searchable for the planet.  These ``searchable stars'' are the points on 
-the sky for which they think planets are observed with 100\% detection 
-efficiency. Correcting for the geometric 
+and searchable for the planet.  ``Searchable stars'' are the stars for which 
+planets are observed with 100\% detection efficiency. 
+Correcting for the geometric 
 transit probability $f_{\rm g}$, they compute an apparent occurrence rate 
-$\Gamma_{\rm a}$:
+$\Lambda_{\rm a}$:
 
 \begin{equation}
-\Gamma_{\rm a} = \frac{N_{\rm det}}{N_\star} \times \frac{1}{f_g}.
+\Lambda_{\rm a} = \frac{N_{\rm det}}{N_\star} \times \frac{1}{f_g}.
 \end{equation}
@@
 Here we focus on problems that arise from the fact that stars of the desired
-type often exist in binaries. 
-We assume for simplicity that all binaries in the survey are 
+type often exist in multiple star systems.
+For simplicity, we only consider binaries, and we assume that they are all 
 spatially unresolved.
@@
 An immediate complication is that~--~due to dynamical stability or some 
-aspect of planet formation~--~ the occurrence rate of planets may differ 
+aspect of planet formation~--~the occurrence rate of planets may differ 
 between binary and single-star systems.
@@
 M_\star$, it must implicitly marginalize over stellar multiplicity.
-Otherwise, one must address ``occurrence rates in single star systems'', 
-``occurrence rates about primaries of double systems'', and
-``occurrence rates about secondaries of double star systems''.
+Otherwise, one must discuss ``occurrence rates in single star systems'', 
+``occurrence rates about primaries'', and
+``occurrence rates about secondaries'' (this has been noted by Wang et al., 
+2015)
 
@@
 A given apparently-searchable star may truly be a single star of the desired 
-type. If not, and ignoring higher order multiples, one of the following must 
-be true:
+type. If not one of the following must be true:
 \begin{enumerate}	
-\item The apparently-searchable star is a binary system, with two searchable 
-stars of the desired type. $N_\star$ is under-counted by one for every such 
+    \item The apparently-searchable star is a binary system, with two 
+    searchable 
+    stars of the desired type. $N_\star$ is under-counted by one for every 
+    such 
     system.
     %
-\item The apparently-searchable star is a binary system, with one searchable 
-star of the desired type. 
-The searchable star of the desired type could be the primary or the secondary.
-$N_\star$ is correctly counted for every such system, though there may be 
-systematic errors in estimates of stellar parameters.
+    \item The apparently-searchable star is a binary system, with one 
+    searchable star of the desired type. 
+    $N_\star$ is correctly counted for every such system.
+    There may be systematic errors in stellar parameter estimates of such 
+    systems, but in this work we neglect these.
     %
-\item  The apparently-searchable star is a binary system, in which neither of 
-the components is a searchable star, but at least one is of the desired type. 
+    \item  The apparently-searchable star is a binary system, in which neither 
+    of the components is a searchable star, but at least one is of the desired 
+    type. 
     $N_\star$ is over-counted by one for every such system.
@@
 component is of the desired type.
-%For single stars, this may simply be from erroneous stellar parameters.
-%In binaries, one might imagine the combined light of two stars somehow making 
-%them resemble a single star of the desired type.  
 We will not consider errors of this type.  We will assume that all the 
@@
     \item 
-A detected signal is from a star of undesired type (by the above assumptions, 
+    A detected signal is from a star of undesired type (by the above 
+    assumptions, 
     the secondary of a binary system), from a planet of undesired type, and is 
@@
     \item
-A detected signal is from a star of desired type (the primary or secondary of 
+    A detected signal is from a star of desired type (the primary or secondary 
+    of 
     a binary), from a planet of undesired type, and is incorrectly counted 
     as a planet of desired type.
-This and the case above occur when the host star parameters are miscalculated 
+    This and the case above occur when the host star parameters are 
+    miscalculated 
     ({\it e.g.,} the host star is the faint secondary, but is assumed to be 
-primary), or when the constant diluting light from the binary companion is not 
-corrected.
+    primary), or when the constant diluting light from the binary companion is 
+    not corrected.
     $N_{\rm det}$ should be lowered by 1 for every such case.
@@
     \item
-An {\it undetected} signal from a planet of the desired type, around a star of 
+    An {\it undetected} signal from a planet of the desired type, around a 
+    star of 
     the desired type, was not detected because of dilution from the companion 
@@
     This happens during case \#3 in $N_\star$-counting errors (the 
-apparently-searchable star is a binary system, in which neither component is 
+    apparently-searchable star is a binary system, in which neither component 
+    is 
     in fact searchable). If $N_\star$ is corrected for this, {\it i.e.,} the 
-system is correctly counted as ``not searchable'', no correction to $N_{\rm 
+    system is correctly counted as ``not searchable'', no correction to 
+    $N_{\rm 
         det}$ is needed.
@@
 survey completeness.
-%Note that in the above list, a planet cannot be both ``of desired type'' and 
-%``orbiting a star of undesired type''. We ignore this type of error for 
-%analytic convenience, but \texttt{reintroduce it later during numerics}.
 
-Our approach is step-by-step starting from a very simple scenario.
-In Sec.~\ref{sec:model_1} we first consider a universe consisting of only the 
-desired types of stars and planets.  
-We give an analytic expression for $\Gamma_{\rm a}/\Gamma_{\rm t}$, the ratio 
-of the true to apparent occurrence rate, taking into account all applicable 
-errors.
-In Sec.~\ref{sec:model_2} {\bf we allow for a distribution of the mass ratio 
-$dn/dq \propto q^\gamma$} {\bf (not done yet)}, 
-and a concurrent power-law distribution in the luminosity ratio, $dn/dL 
-\propto L^\beta$. 
-We give expressions for $\Gamma_{\rm a}/\Gamma_{\rm t}$, and evaluate them 
-numerically for the case of planets orbiting Sun-like dwarf stars.
-%We assume the secondary is not of the desired stellar type.
-%NOTE: LB notes that yes, while analytically simple, it doesn't make sense. 
-%For the twin binaries case, what possible reason could there be for the 
-%secondary to not be of the desired stellar type? Similar for the high mass 
-%ratio binaries -- the stars are ~identical.
-In Sec.~\ref{sec:model_3}, 
-we assume the planets have a power-law radius distribution, ${\rm 
-d}N_p/{\rm d}R_p\propto R_p^\alpha$.  We calculate the 
-effects of binarity on the apparent $\Gamma_{\rm a}(R_p)$ taking into account 
-all relevant errors.
-Accounting only for dilution \& its numerator effect, big radius planets have
-underestimated inferred rates, the smallest radius planets have overestimated
-inferred rates, and the in-between radius planets are in-between.
-One consequence is that the claimed HJ occurrence rate discrepancy between 
-radial velocity \& transit surveys gets much weaker when accounting for 
-binarity.
-In Sec.~\ref{sec:discussion} we summarize the models and discuss how they can 
-be used to interpret transit surveys.
+\paragraph{The Hot Jupiter Rate Discrepancy}
 
-\begin{comment}
-We then dig in, w/ numerics, to what happens with different assumed
-completeness vs Rp functions.
+Radial velocity surveys of nearby bright Sun-like stars have reported hot 
+Jupiter ($a<0.1\,{\rm AU}, P\lesssim 10\,{\rm day}$) occurrence rates of
+$12\pm 1$, $15\pm 6$, $8.9 \pm 3.6$, and $12.0 \pm 3.8$ per thousand stars
+(Marcy et al 2005, Cumming et al 2008, Mayor et al 2011, Wright et al 2012 
+respectively).
 
-This is b/c the more believable occ rate papers do not heavily rely on their
-completeness calibration -- they know they're complete.
-This is a much more understandable regime than the wonky completeness estimate 
-dependent zone of eta-Earth for Kepler.
+In transit surveys, OGLE-III reported an absolute rate of $3.1^{+ 
+4.3}_{-1.8}$ hot Jupiters with $P<5\,{\rm days}$ per thousand stars (Gould et 
+al 2006).
+The SuperLupus survey quoted $10^{+27}_{-8}$ per thousand stars, allowing 
+planets with periods less than 10 days (Bayliss \& Sackett 2011).
+Through the {\it Kepler}\ survey, Howard et al. (2012) later reported a rate 
+of 
+$4 \pm 1$ per thousand Sun-like stars.
+This rate was for $P<10\,{\rm day}, 8-32r_\oplus$ planets, and 
+was specific to their ``solar subset''\footnote{The ``solar subset'' was 
+defined as $4100<T_{\rm eff}/{\rm K}<6100$, {\it Kepler}\ 
+magnitude $<15$, $4.0 < \log g < 4.9$, and only took planets with measured 
+signal to noise $>10$.
+}.
+Expanding their sample to fainter magnitudes, they quoted a rate of $5 \pm 
+1$.
+Expanding down to $r_p>5.6r_\oplus$, to avoid excluding hot Jupiters reported 
+by RV surveys, they reported $7.6 \pm 1.3$.
+Recent work by Petigura et al. using the updated parameters of the 
+California-{\it Kepler}\ Survey has found a rate of $X.X \pm 
+Y.Y$ (2017, in preparation)
 
-Finally, for numerics, we simulate a sample of 20,000 
-apparently-searchable-stars (or however many "stars" appear to be searchable
-for Earthlike planets in the Kepler field).  Basically this means a magnitude
-limited survey of 20,000 ``stars'' that are either single or are binaries with
-the primary of the desired type.  The sample has a realistic distribution of
-binaries (occurrence and luminosity ratio).  We calculate 
-eta(apparent)/eta(true).  We can also play with errors of type 4 in this
-numerical model (i.e. we let the relative occurrence rates between primaries 
-and secondaries vary).  (It may not worth doing so for the analytic 
-work-- though it's pretty simple. The way to play with this is just present 
-both limits, and then also present a few weight factors (Furlan+17 do a good 
-job).
+The trend is that hot Jupiter occurrence rates measured by transit 
+surveys seem to be marginally lower than those found by radial velocity 
+surveys.
+While the actual discrepancy is of sub-$3\sigma$ significance,
+one reason to expect a difference in the rates is that the populations have 
+different metallicities.
+As originally argued by Gould et al. (2006), the RV sample is biased towards 
+metal-rich stars, which have been measured by RV surveys to preferentially 
+host more giant planets (Santos et al 2004, Fischer and Valenti 2005).
+The {\it Kepler}\ sample specifically has been measured to be more metal poor 
+than the local neighborhood, with a mean metallicity of $[{\rm M/H}]_{\rm 
+mean}\approx -0.05$ (Dong et al., 2014; Guo et al., 2017).
+Studying the problem in detail, Guo et al. recently argued that the 
+metallicity difference could account for a $\approx 0.1\%$ difference in the 
+measured rates between the CKS and {\it Kepler}\ samples~--~not a $\approx 
+0.5\%$ difference.
+These authors concluded that ``other factors, such as binary contamination and 
+imperfect stellar properties'' must also be at play (Guo et al., 2017).
 
-We discuss ``how do we interpret Fulton+ 2017?''
-That paper did (or claims to have done) a good job at weeding out numerator
-errors. If true, means $\Gamma(R_{\rm p})$ has the right shape, and just the 
-wrong normalization. Important if we ever manage to convince ourselves 
-``brightest transit hosts'' could work.
+Aside from surveying stars of different metallicities, radial velocity and 
+transit surveys differ in how they treat binarity.
+Radial velocity surveys typically reject both visual and spectroscopic binaries
+({\it e.g.}, Wright et al. 2012).
+Transit surveys typically observe binaries, but the question of whether they 
+were searchable to begin with is usually left for later interpretation.
+In spectroscopic follow-up of transiting candidates, the prevalence of 
+astrophysical false-positives may also lead to a bias against confirmation of 
+transiting planets in binary systems.
 
-In passing, at the end, we mention how this changes attempts at eta-Earth 
-measurements.
-\end{comment}
+Ignoring these complications, in this work we focus on whether
+binarity might intrinsically bias occurrence rate measurements, simply 
+through its effects on the number of searchable stars and the apparent radii 
+of detected planets.
 
-\begin{comment}
-UNSOLVED PROBLEMS W/ THE NUMERICAL MODEL-
-* currently have not implemented self-consistent (L,M,R) relation between
-primaries and secondaries.
-Wang et al (2015) avoided that problem by doing a similar empirical thing to
-what I currently have implemented. But I do not think that was a good paper.
-Ciardi et al (2015) have an approach that was Kepler-specific -- they used
-the same Dartmouth isochrones used for KIC stellar parameters.
-But they did not conserve $dn/dm_x$, for $m_x$ apparent mag in band x.
-I think the latter choice might be better for Kepler-specific discussion.
-But it's less-nice if we want this to be general enough for e.g., occ HJ
-rates w/ TESS.
-
-My main gripe w/ taking the 150k KIC stars \& adding binaries is that less
-is known about those parameters and biases. Using e.g., Galaxia has the major
-benefit of you understand what's been cooked in much better (it's minimal).
-
-I think if we want to talk about Kepler, it's the best possible approach
-though.
-
-If we want to talk about Fulton+17 we can go further, and apply the CKS cuts.
-
-Some questions based on the current draft:
-
-- Why should we ever consider $F_{lim} < F_{min}$?
-
-LB: in practice, surveys look at many stars for which they could not detect
-Rp=Re planets, but could detect Rp=10Re planets.  One possible error is
-miscounting which stars are which (i.e. getting completeness corrections
-wrong). So this was more for plausbility than b/c it's immediately needed in
-the analytics. I can cut it.
-
-- Why do we need to consider the distribution of S/N, as opposed to simply the
-limiting distance for detection?
-
-LB: for "model 1" (fixed stars, fixed planets, fixed light ratio binaries)
-you're correct, it's not immediately necessary.
-
-For this model, I did it as a way to think through the completeness effects
-systematically -- i.e. how for the twin-binary case, you get a completeness
-fraction of 1/8, though your limiting distance is sqrt(2) times bigger. (Even
-when your completeness for the singles is 100%).
-
-If we go the path of completing "model 2"'s analytics (vary light ratio, fixed
-primary luminosity, fixed planet), I think this formulation would become
-necessary in order to actually evaluate the completeness fractions.
-
-(I never completed these analytics b/c of the L(M) mess + confusion about
-assumed stellar masses+radii. Power law L(M) helps this... I think there needs
-to be an extra assumption about the stellar mass+radius change though that we
-have not spelled out)
-
-- I don't think the bandpass, effective area, photon count rate, etc.  are
-needed, either. All that matters is the limiting distance.
-
-Are needed for what? 
-In Model 1, for analytic expressions of the occurrence rate, I agree.
-For analytic expressions of the number of detected planets, I disagree.
-
-For Model 2, I'm not sure because I got bogged down. I suspect that for
-occurrence rate errors, you are correct.
-
-The analytic expression I gave for the occurrence rate depends on this "xi"
-factor, which needs to incorporate an assumption about how Astronomer A
-misestimates the completeness (relatedly, how he misestimates Rstar).  I never
-specified that assumption.
-
-- In Eq. 4, is this a traditional definition of BF?  I would have defined it as
-nd/(ns+nd) or perhaps 2*nd/(ns+ 2*nd).  Let's make sure we are using
-traditional terminology, and also interpreting the literature correctly, i.e.,
-is the right number really 0.45?
-
-Oops. nd/(ns+nd) = BF = MF (if assuming all multiple systems are binaries).
-This minor change propagates.
-
-Touching up on correct terminology, from Duchene \& Kraus 2013, Sec 2.1:
-
-"Frequency of multiple systems" (MF)
-vs
-"Companion frequency" (CF; average \# of companions per target, can exceed 
-100\%)
-
-my understanding of the terms:
-
-survey 1: {*, **, *, ***, *, *}
-
-6 systems.
-2 are multiple.
-MF = 2/6 = 1/3
-CF = (0+1+0+2+0+0)/6 = 3/6 = 1/2
-
-survey 2: {**, **, ***, ****, *}
-
-5 systems
-4 are multiple
-MF = 4/5
-CF = (1+1+2+3+0)/5 = 6/5
-
-DK2013 Sec 3.1.1 give (citing Raghavan+ 2010)
-MF^{MS}_{1-1.3Msun} = 50 \pm 4%
-MF^{MS}_{0.7-1Msun} = 41 \pm 3%
-
-They also give
-CF^{MS}_{1-1.3Msun} = 75 \pm 5%
-CF^{MS}_{0.7-1Msun} = 56 \pm 4%
-
-I understand the first pair of numbers to mean:
-
-In a volume-limited survey (of the stellar neighborhood), if you select systems
-for which the primary is from 1-1.3Msun, 50% of them will be multiple.
-
-I take the second pair of numbers to mean:
-
-In a volume-limited survey (of the stellar neighborhood), if you select systems
-for which the primary is from 1-1.3Msun, each system will on average have 0.75
-companions.
-
-So YES, need to correct this throughout.
-\end{comment}
-
-\section{Model \#1: fixed stars, fixed planets, fixed light-ratio binaries}
-\label{sec:model_1}
-
-Consider the following idealized survey.
-An astronomer observes the entire sky for some duration with a detector of 
-known area and bandpass.
-The detector is photon-noise limited.
-They are interested only in detecting planets of radius $R_p$, and orbital 
-period $P$. %For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
-They are only interested in detecting them around stars of radius $R_1$, 
-and luminosity $L_1$. %For instance, G2V dwarfs.
-They detect the desired type of planet when they observe signals 
-above some signal-to-noise floor.
-For a photon-noise limited survey, this minimum signal to noise ratio is 
-equivalent to a minimum flux required for detection, $F_{\rm min}$.
-Therefore they observe all of the ``points'' on-sky with apparent 
-magnitude $m < m_{\rm min}$, or equivalently with energy flux in their 
-bandpass greater than the limiting flux $F > F_{\rm min}$.
-After detecting all planets with ${\rm S/N} > {\rm (S/N)_{min}}$, they now 
-wish to derive an occurrence rate for planets of radius $R_p$ and 
-orbital period $P$.
-
-Assume this astronomer lives in a universe in which the following assumptions 
-hold.
+To progressively gain intuition, we study the following idealized transit 
+surveys:
 \begin{itemize}
-	\item The true population of ``points'' (stellar systems, all unresolved) 
-	comprises only single and double star systems. Single star systems have 
-	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
-	temperature $T_{\rm eff,1}$.
-	Double star systems have luminosity in the observed bandpass $L_d = 
-	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
-	the secondary to the primary. 
-	In this section, $\gamma_R=1$ across the population 
-	of star systems~--~all binaries are assumed to be twins.
-	The ratio of the number density of binary systems to all systems in a 
-	volume-limited sample is the binary fraction\footnote{The 
-	binary fraction is equivalent to the multiplicity fraction if there are 
-	no triple, quadruple, $\ldots$ systems.}.
-	\item The true population of planets around these stars is as follows.
-    In single star systems, a fraction $\Gamma_{t,s}$ of stars 
-	have planets of radius $R_p$, with orbital period $P$.
-	In double star systems, a fraction $\Gamma_{t,d}$ of primary stars have 
-	planets of radius $R_p$, with orbital period $P$. 
-	A fraction $w_{d2}\Gamma_{t,d}$ of secondary stars in double star 
-	systems have the same type of 
-	planet, where the weight factor $w_{d2}$ specifies the secondary star 
-	occurrence rate as a fraction	of the primary star occurrence rate.
-	Any astrophysical difference in planet formation between singles and 
-	binaries is captured by $\Gamma_{t,s}, \Gamma_{t,d},$ and $w_{d2}$.
-	If one assumes the primary and secondary of a binary system host planets 
-	at equal rates, this is the $w_{d2}=1$ limit.
-	Similarly, if one assumes  secondaries 
-	cannot host planets, this is the $w_{d2}=0$ limit.
+    \item Model \#1: fixed stars, fixed planets, twin binaries;
+    \item Model \#2: fixed planets and primaries, varying secondaries;
+    \item Model \#3: fixed primaries, varying planets and secondaries.
 \end{itemize}
-
-The assumption of [Pepper et al, 2003] in a very similar context was that the 
-observer can correctly pre-select all of the ``searchable stars''.
-This simplification would allow the observer to ignore 
-completeness effects when deriving occurrence rates, since their completeness 
-is 100\%.
-
-However, any signal-to-noise limited survey of a population that includes 
-binaries cannot ignore incompleteness.
-%If we set the limiting flux to give 100\% completeness for single stars, we 
-%will unwittingly select binaries near that flux limit, for which dilution 
-%will push transit signals below the detection threshold.
-This is because any given flux limit maps onto three distinct ``maximum 
-distances'' for a population of single stars and twin binaries.
-First, there is $d_{{\rm max},s}$: the maximum distance out to which the 
-desired type of planet is detectable around single stars. Typically this is 
-set to be the distance out to which stars in a SNR-limited survey are 
-selected, since these are the ``searchable stars''.
-Using the same minimum flux as that defining $d_{{\rm max},s}$, double stars 
-are then selected out to a larger limiting distance, $d_{{\rm 
-max},d}$~--~effectively a Malmquist bias.
-
-However, the distance out to which double stars are {\it selected} is 
-different from the distance $d_{{\rm max},d}^{\rm p}$ out to which planets can 
-be {\it detected} (around the primary of a double star system).
-For the binary twin population, all planet and stellar radii are identical, so 
-${\rm S/N} \propto \mathcal{D}, and
-L^{1/2} d^{-1}$,
-\begin{equation}
-d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
-(1 + \gamma_R)^{-1} d_{{\rm max},d}.
-\end{equation}
-For $\gamma_R = 1$, this means only 1 in 8 apparently-searchable stars which 
-are in fact binary systems can yield detectable planets (the ratio of 
-volume in which planets can be detected to the actual selected volume).
-
-\begin{comment}
-We note in passing that we are considering a SNR-limited survey, which in this 
-case is the same as a magnitude-limited one.
-The observer might prefer to perform a completeness-limited survey, which in 
-this case would be the same as a volume-limited one with limiting distance
-$d_{{\rm max},d}^{\rm p}$.
-This change might circumvent various biases.
-\end{comment}
-With our model defined, we ask:
-
-\subsection{What is the true occurrence rate?}
-\label{sec:true_rate}
-
-The ``true occurrence rate'' $\Gamma_t$ is the average number of planets of 
-the desired type per star of the desired type:
-\begin{align}
-\Gamma_t &= \frac{N_{\rm planets}}{N_{\rm stars}} \\
-\Gamma_t &= \frac{\Gamma_{t,s} N_s + (1+w_{d2}) \Gamma_{t,d} N_d}{N_s + 2N_d}.
-\label{eq:true_occ}
-\end{align}
-where $N_s$ is the number of single star systems, and $N_d$ the number of 
-double star systems. 
-The factor of 2 in the total number of stars accounts for the fact that there 
-are twice as many stars in double star systems.
-
-To develop more analytic expressions in terms of observables, assume that the 
-stars are homogeneously distributed\footnote{
-If we wished to write a stellar number density profile that accounted for the 
-vertical structure of the Milky Way, we might choose a profile either $\propto 
-\exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from the 
-galactic midplane and $H$ a scale-height. Both density profiles would lead 
-closed form analytic solutions.	}.
-Counting the number of single and double systems in their respective selected 
-volumes,
-\begin{equation}
-N_i = n_{i} \frac{4\pi}{3} d_{{\rm max},i}^3,
-\label{eq:number_systems}
-\end{equation}
-for $i\in{\rm \{single,double\} } \equiv { \{s,d\} }$, and
-\begin{equation}
-\frac{n_{d}}{n_s+n_d} = {\rm binary\ fraction} \equiv {\rm BF}
-\end{equation}
-The absolute normalization of the number density $n_s + n_d$ is an observed
-quantity [Bovy 2017], as is the binary fraction.
-For ``sun-like'' dwarfs with $0.7<M_\star/M_\odot<1.3$,
-${\rm BF}=0.44\pm0.02$ [Duchene \& Kraus, 2013, Raghavan et al 2010]. 
-These latter authors really quote a {\it multiplicity} fraction of
-0.44 for sun-like dwarfs~--~for simplicity we count the $\approx 25\%$ of 
-multiple systems that are higher order multiples as doubles.
-
-$d_{{\rm max},i}$ in Eq.~\ref{eq:number_systems} is the maximum distance 
-corresponding to the given magnitude limit:
-\begin{equation}
-d_{{\rm max},i} = \left(\frac{L_{i}}{4 \pi F_{{\rm min}}}\right)^{1/2},
-\label{eq:d_max}
-\end{equation}
-where the limiting flux in the bandpass $F_{{\rm min}}$ could equivalently be 
-stated in terms of a limiting magnitude $m_{{\rm min}}$.
-In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$. As a 
-consequence of imposing a magnitude cut, the maximum distance to which binary 
-stars will be selected is greater than that of single stars.
-The ratio of double to single systems in the observed sample is 
-then
-\begin{align}
-\frac{N_d}{N_s} &= 
-	\frac{n_d}{n_s}
-	\left( \frac{d_{{\rm max}, d} }{d_{{\rm max}, s}} \right)^3 \\
-&= \frac{{\rm BF}}{1 - {\rm BF}} \times (1+\gamma_R)^{3/2}.
-\label{eq:beta_model1}
-\end{align}
-In the case of twin binaries ($\gamma_R = 1$), with a binary fraction 
-${\rm BF} = 0.5$, there are 
-$2^{3/2}\approx2.8$ times more binary systems than single systems in the 
-SNR-limited sample.
-Since the completeness fraction for binary systems is $1/8$, there are 
-$2^{-3/2}\approx0.35$ times fewer binary systems in the sample that can yield 
-planets as single star systems in the system that can yield planets.
-
-\subsection{What is the measured occurrence rate?}
-% % % % % % % % % % % %
-
-The observer who has never heard of binary star systems wants to find the 
-occurrence rate of planets with radius $R_p$ and period $P$, orbiting stars of 
-radius $R_1$ and luminosity $L_1$.
-They detected $N_{\rm det}$ transit signals that appear to be from planets of 
-this desired type.
-The occurrence rate they then find is
-\begin{equation}
-\Gamma_{\rm a } = \frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_g}.
-\end{equation}
-
-Since only the transit depth is observed, the planetary radii apparent to this 
-astronomer, $R_{\rm p, a}$, are different from the true planetary radii 
-$R_{\rm p, t}$. 
-This is because the assumed stellar radii $R_{\star,{\rm a}}$ may differ from 
-the true stellar radii $R_{\star,{\rm t}}$, and the flux will be diluted:
-\begin{equation}
-R_{p,{\rm a}} = R_{p,{\rm t}} 
-\frac{R_{\star,{\rm a}}}
-		 {R_{\star,{\rm t}}}
-		 \,\mathcal{D}^{1/2},
-\end{equation}
-where the dilution parameter $\mathcal{D}$ is defined for binary systems as
-\begin{align}
-\mathcal{D} &= 
-\Bigg\{\begin{array}{lr}
-L_1 / L_d, & \text{if\ planet\ orbits\ primary}\\
-\gamma_R L_1 / L_d, & \text{\ if\ planet\ orbits\ secondary}
-\end{array} \nonumber\\
-&=
-\Bigg\{\begin{array}{lr}
-(1+\gamma_R)^{-1}, & \text{if\ planet\ orbits\ primary}\\
-(1 + \gamma_R^{-1})^{-1}, & \text{if\ planet\ orbits\ secondary}
-\end{array} 
-\label{eq:dilution}
-\end{align}
-where $L_1, L_d,$ and $\gamma_R$ were defined in the opening monograph.
-
-In our twin binary thought-experiment, dilution means that all planets 
-detected in binary systems will have apparent radii of $R_p/\sqrt{2}$.
-Since for twin binaries the stellar radii are the same, the transit 
-probability will be correctly calculated.
-
-The total number of planet detections, irrespective of whether they are ``of 
-the desired type'', is the sum of the number of planets 
-detected in single star systems $N_{{\rm det},s}$ and the number of planets 
-detected in double star systems $N_{{\rm det},d}$.
-Expressing each individually,
-\begin{equation}
-N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,g} f_{s,c},
-\label{eq:N_det_s}
-\end{equation}
-where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
-star systems of the sample, $f_{s,g}$ is the geometric transit probability, 
-and $f_{s,c}$ is the fraction of these transiting planets that are observed 
-with signal to noise greater than the minimum detection threshold (the 
-completeness). 
-For our SNR-limited, twin-binary survey, $f_{s,g} = R_1/a\equiv f_g$, and 
-$f_{s,c}=1$.
-
-Analogously,
-\begin{equation}
-N_{{\rm det},d} = (1+w_{d2}) N_d \Gamma_{t,d} f_{d,g} f_{d,c},
-\label{eq:N_det_d}
-\end{equation}
-where now $(1+w_{d2}) N_d \Gamma_{t,d}$ is the number of planets in the double 
-star systems of the sample, the geometric transit probability $f_{d,g}=R_1/a$ 
-for our twin-binary survey, and $f_{d,c}=1/8$ by the geometric argument 
-presented in the initial discussion of limiting distances.
-
-The occurrence rate this astronomer derives for planets of assumed 
-radius $R_p$ orbiting stars thought to be of the desired type is then
-\begin{equation}
-\Gamma_{\rm a} = \frac{\Gamma_{t,s}N_s}{N_s + N_d}.
-\end{equation}
-
-\subsection{Errors on derived occurrence rates for twin-binary thought 
-experiment}
-The ratio of the true to calculated occurrence rates for planets of radius 
-$R_p$, $X_\Gamma$, is
-\begin{align}
-X_\Gamma &=
-\frac{\Gamma_t}{\Gamma_{\rm a}} \\
-&=
-\frac{\Gamma_{t,s}N_s + (1+w_{d2})\Gamma_{t,d}N_d}{\Gamma_{t,s} N_s } \cdot 
-\frac{N_s + N_d}{N_s + 2N_d} \\
-&= 
-\left( 
-1 + \frac{(1+w_{d2}) \Gamma_{t,d}}{\Gamma_{t,s}} \beta
-\right) \cdot 
-\frac{1+\beta}{1+2\beta},
-\end{align}
-for
-\begin{equation}
-\beta \equiv N_d/N_s = \frac{\rm BF}{1- {\rm BF}} \times (1+\gamma_R)^{3/2}.
-\end{equation}
-
-For the nominal G2V dwarf case of ${\rm BF}=0.44$, if we assume that all stars 
-have equal occurrence rates, $\Gamma_{t,s} = \Gamma_{t,d}$ and 
-$w_{d2}=1$, then we find $X_\Gamma = 1+\beta = 1+ 2^{3/2}\approx4$~--~ the 
-observed occurrence rate is underestimated by a factor of 4.
-
-For the planets around double stars with observed radii $R_p 
-\sqrt{\mathcal{D}}$, there is also the error of their misclassification.
-The fraction of detections with misclassified radii can be written
-\begin{equation}
-\frac{N_{\rm det,d}}{N_{\rm det,s} + N_{\rm det,d}} = \frac{1}{1+\zeta},
-\end{equation}
-for
-\begin{equation}
-\zeta \equiv \frac{N_s \Gamma_{t,s} f_{s,g} f_{s,c}}{
-	 (1+w_{d2})N_d \Gamma_{t,d} f_{d,g} f_{d,c }}.
-\end{equation}
-
-For our nominal G2V dwarf case, the twin binaries with equal 
-occurrence rates gives $\zeta = 4/\beta = \sqrt{2}$, so 41\% of detections 
-have misclassified radii, all of $R_p/\sqrt{2}$.
-
-Of course more generally, the observer might define their 
-desired ``type of planet'' to be anything within some margin of $R_p$.
-Considering this for the idealized twin-binary case, if the margin includes 
-the planets in binary systems, the number of detections ``of the desired 
-type'' would change.
-The correction factor $X_\Gamma$ would then become
-\begin{equation}
-X_\Gamma = 
-\frac{\Gamma_{t,s} + (1+w_{d2})\Gamma_{t,d}\beta}{\Gamma_{t,s} + 
-(1+w_{d2})\Gamma_{t,d} f_{d,c} } \cdot 
-\frac{1+\beta}{1+2\beta},
-\end{equation}
-which simplifies to $(1+2f_{d,c})^{-1} (1+\beta) = 4(1+\beta)/5\approx 3$ for 
-our twin-binary thought experiment.
-This makes sense: increasing the radius interval of ``desired planets'' 
-decreases the error in the numerator, but the error of mis-counted stars in 
-the denominator remains identical.
-
-\section{Model \#2: fixed primaries, fixed planets, varying light-ratio 
-secondaries}
-\label{sec:model_2}
-
-The model of Sec.~\ref{sec:model_1} is idealized, but it helps
-develop intuition.
-Now, we let the light ratio $\gamma_R = L_2/L_1$ vary across the population 
-of star systems.
-It does so because the underlying mass ratio $q=M_2/M_1$ varies.
-We keep the primary mass fixed as $M_1$, which is also the mass of all single 
-stars.
-
-One parametrization for the distribution of 
-binary mass ratios in a volume-limited sample is $f(q)\propto q^\gamma$.
-[Duchene and Kraus 2013], fitting all the multiple systems of [Raghavan et al 
-(2010)]'s Fig 16, quote $\gamma = 0.28\pm0.05$ for $0.7<M_\star/M_\odot<1.3$.
-Examining only the binary systems of Rhagavan et al 2010, Fig 16, the 
-distribution is roughly uniform,
-\begin{equation}
-{\rm prob}(q) =
-\Bigg\{\begin{array}{lr}
-{\rm constant} & 0.1 < q \leq 1  \\
-0 & {\rm otherwise},
-\end{array}
-\label{eq:mass_ratio}
-\end{equation}
-though we ignore the peak at high mass ratios for analytic convenience. 
-We choose this latter mass ratio distribution.
-
-For the mass-luminosity relation, %for analytic convenience 
-we assume 
-$L \propto M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
-To check, we fit lines to measured masses and luminosities of dwarf 
-stars\footnote{
-    Data collected by Torres et al. [2009] for dwarfs earlier than 
-    spectral class M, and Benedict et al. [2016] for dwarfs later than 
-    spectral class M.
-    To convert Benedict et al. [2016]'s reported $M_V$ values to absolute 
-    luminosities, we interpolated over E. Mamajek's 
-    table
-	\texttt{pas.rochester.edu/\textasciitilde 
-		emamajek/EEM\_dwarf\_UBVIJHK\_colors\_Teff.txt},
-	downloaded 2017.08.02}.
-In log-log space, we let the intersection point of the lines float. The 
-resulting fit parameters are available in a footnote\footnote{
-	m lo: 1.8818719873988132~--~
-	c lo: -0.9799647314108376~--~
-	m hi: 5.1540712426599882~--~
-	c hi: 0.0127626185389781~--~
-	M at merge: 0.4972991257826812~--~
-	L at merge: 0.0281260412126928.
-}. Though the data show a break at $M_\star \approx 0.5M_\odot$, the 
-$L\propto M^{3.5}$ relation is a reasonable fit for $M\lesssim 2M_\odot$.
-
-The final necessary assumption is a stellar mass-radius relation.
-This affects the transit probabilities and signal sizes about the 
-secondaries of binary star systems.
-Since our main purpose is to develop intuition about observational 
-biases, we coarsely assume
-\begin{equation}
-R\propto M.
-\end{equation}
-\begin{comment}
-[Demircan \& Kahraman 1991]'s empirical fit to eclipsing 
-binary data,
-\begin{equation}
-R = 1.06 M^{0.945}
-\label{eq:mass_radius}
-\end{equation}
-applies for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 
-1991's stated bounds), where each quantity is given in solar units. 
-\end{comment}
-With our model specified, we ask:
-
-\subsection{What is the true occurrence rate?}
-
-The true occurrence rate is the average number of planets of the desired type 
-per star of the desired type.
-The number of planets of the desired type is
-\begin{equation}
-N_{\rm planets} = \Gamma_{t,s}N_s + \Gamma_{t,d}N_d + 
-\Gamma_{t,d}w_{d2}f_{{\rm desired}}N_d,
-\end{equation}
-where as before $N_s$ and $N_d$ are the number of single and double systems, 
-and the fraction $f_{{\rm desired}}$ is defined so that $f_{{\rm desired}} 
-N_d$ is the number of secondaries that are of the desired type, and as before 
-$\Gamma_{t,d}w_{d2}$ is the average number of planets each of these desired 
-stars hosts.
-Explicitly,
-\begin{equation}
-f_{\rm desired} = \frac{\int_{M_{\rm min}}^{M_{\rm max}} {\rm prob}_{\rm 
-        ml}(q)\,{\rm d}q }
-{\int_{0.1M_1}^{M_1} {\rm prob}_{\rm 
-        ml}(q)\,{\rm d}q}.
-\end{equation}
-For instance, assuming a primary mass of $M_1 = M_\odot$, and defining the 
-``desired type'' of stars being to be $0.7<M_\star/M_\odot<1.3$, we find 
-$f_{\rm desired}\approx0.475$.
-
-Counting the number of desired stars that are single, $N_s$, is the same as in 
-Sec.~\ref{sec:model_1}. However now the maximum selected distance for a binary 
-system is now a function of the system's light ratio,
-\begin{equation}
-d_{\rm max,d} =
-d_{\rm max,s}\times (1+\gamma_R)^{1/2},
-\end{equation}
-where $d_{\rm max,s}$ is given by Eq.~\ref{eq:d_max}.
-This affects our ability to count the number of desired stars that are double,
-since
-\begin{equation}
-N_d(\gamma_R) = n_d \frac{4\pi}{3} d_{\rm max,d}^3.
-\end{equation}
-
-For a given binary system, $q$, $\gamma_R$, and thus $d_{\rm 
-	max,d}(\gamma_R)$ are all random variables.
-Since $N_d$ is a function of $d_{\rm max,d}$, 
-the number of double star systems becomes a random variable, drawn from the 
-ensemble of all possible transit surveys.
-Transforming the appropriate probability density functions, the distribution 
-function for the number of binary systems,
-${\rm prob}(N_d)$, can be expressed as 
-\begin{equation}
-{\rm prob}(N_d) = {\rm prob}(q(\gamma_R)) 
-\left| \frac{{\rm d}q}{{\rm d}\gamma_R}  \right|
-\left| \frac{{\rm d}\gamma_R}{{\rm d}d_{\rm max,d}}  \right|				
-\left| \frac{{\rm d}d_{\rm max,d}}{{\rm d}N_d}  \right|.
-\label{eq:prob_Nd_raw}
-\end{equation}
-
-\paragraph{Necessary aside on biases in mass ratio distributions}
-
-\begin{figure}[!t]
-    \begin{center}
-        \includegraphics[width=.5\textwidth]{figures/joint_prob_r_q.pdf}
-    \end{center}
-    \caption{The joint probability distribution of a binary star's position 
-        and 
-        mass ratio in a magnitude-limited sample. This plot assumes $L\propto 
-        M^{3.5}$. Note that the marginalized 
-        mass ratio distribution is uniform at any given distance.}
-    \label{fig:joint_prob_r_q}
-\end{figure}
-
-\begin{figure}[!t]
-    \begin{center}
-        \includegraphics[width=0.5\textwidth]{figures/q_distribn_mag_limited.pdf}
-    \end{center}
-    \caption{The distribution of the mass ratio for a magnitude limited 
-        sample of binary stars. The underlying mass ratios are drawn from 
-        Eq.~\ref{eq:mass_ratio} -- a \textit{uniform} distribution in a 
-        volume-limited sample.
-        The entire bias can be understood analytically 
-        (Eq.~\ref{eq:pdf_q_mag_limited}) to good precision. 
-        The numerical comparison uses the ``realistic'' empirical 
-        mass-luminosity relation shown in Fig.~\ref{fig:mass_luminosity}.
-    }
-    \label{fig:q_distribn_mag_limited}
-\end{figure}
-
-\begin{figure}[!t]
-    \begin{center}
-        \includegraphics[width=.5\textwidth]{figures/gammaR_distribn_mag_limited.pdf}
-    \end{center}
-    \caption{The distribution of the luminosity ratio for a magnitude limited 
-        sample of binary stars. Despite the preference in the sample towards 
-        large $q$ binaries (Fig.~\protect\ref{fig:q_distribn_mag_limited}), 
-        the bias is towards small $\gamma_R$ binaries because of the steepness 
-        of the mass-luminosity relation.
-        The underlying mass ratios are uniform in volume.
-        The numerical comparison uses the empirical mass-luminosity relation 
-        shown in Fig.~\ref{fig:mass_luminosity}.       
-    }
-    \label{fig:gammaR_distribn_mag_limited}
-\end{figure}
-
-Before blindly plugging in Eq.~\ref{eq:mass_ratio}'s assumed mass ratio 
-distribution to find the expected number of double systems, it is critical to 
-note that Eq.~\ref{eq:mass_ratio}'s distribution applies only for 
-volume-limited samples.
-If one is interested in a SNR-limited transit survey (or really any
-magnitude-limited survey), the Malmquist-like bias associated with binarity
-must be included.
-
-The distribution function for the mass ratio of binaries in a magnitude 
-limited sample can be found by marginalizing over the joint 
-distribution for a binary star's position $r$ and mass ratio $q$.
-Using `ml' as a subscript for `magnitude-limited',
-\begin{equation}
-{\rm prob}_{\rm ml}(r,q) = {\rm prob}(q|r){\rm prob}_{\rm ml}(r),
-\end{equation} 
-where ${\rm prob}(q|r)$ is given by Eq.~\ref{eq:mass_ratio}.
-One can show
-\begin{equation}
-{\rm prob}_{\rm ml}(r,q) \propto r^2 \quad{\rm if}\ r<d_{\rm max,d}(q)\ {\rm 
-and}\ 0.1<q<1, 
-\end{equation}
-and otherwise zero. The resulting joint probability distribution is plotted in 
-Fig.~\ref{fig:joint_prob_r_q}.
-Marginalizing the joint distribution over distance, and assuming $L\propto 
-M^\alpha$, one finds
-\begin{align}
-{\rm prob}_{\rm ml}(q) \propto (1+q^\alpha)^{3/2}\quad {\rm if\ 
-}0.1<q\leq 1,
-\label{eq:pdf_q_mag_limited}
-\end{align}
-and otherwise 0.
-This marginalized distribution~--~the distribution function of mass ratios for 
-binaries in a magnitude-limited sample~--~is plotted in 
-Fig.~\ref{fig:q_distribn_mag_limited}.
-Reworded, the probability of drawing a binary of mass ratio 
-$q$ in a magnitude-limited sample scales as $d_{\rm max,d}^3(q)$.
-
-Applying Eq.~\ref{eq:prob_Nd_raw} with the correct magnitude-limited 
-distributions gives
-\begin{equation}
-{\rm prob}(N_d) \propto N_d^{2/3} 
-\left( \left(\frac{N_d n_s}{N_s n_d}\right)^{2/3} - 1
-\right)^{\frac{1-\alpha}{\alpha}}
-\end{equation}
-\begin{comment}
-\begin{align}
-{\rm prob}(N_d) \propto N_d^{2/3} &\left(\frac{3}{4\pi 
-	n_d}\right)^{5/3} d_{\rm max,s}^{-5} \ \nonumber\\
-&\times \left[ 
-	\left(\frac{3 N_d}{4\pi n_d}\right)^{2/3}  d_{\rm max,s}^{-2} - 1
- \right]^{\frac{1-\alpha}{\alpha} }
-\end{align}
-\end{comment}
-over the interval
-\begin{equation}
-N_d^{\rm lower} = \frac{4 \pi n_d}{3} (\sqrt{(0.1)^\alpha + 1} d_{\rm 
-max,s})^3,
-\end{equation}
-\begin{equation}
-N_d^{\rm upper} = \frac{4 \pi n_d}{3} (\sqrt{2}d_{\rm max,s})^3,
-\end{equation}
-and outside the stated interval ${\rm prob}(N_d) = 0$.
-
-An interesting intermediate result is the magnitude-limited luminosity ratio 
-distribution,
-\begin{equation}
-{\rm prob_{ml}}(\gamma_R) \propto (1+\gamma_R)^{3/2} 
-\gamma_R^{\frac{1-\alpha}{\alpha}},
-\end{equation}
-for $(0.1)^{\alpha}<\gamma_R<1$, and otherwise zero.
-For $\alpha=3.5$, ${\rm prob_{ml}}(\gamma_R) \propto 
-(1+\gamma_R)^{3/2}\gamma_R^{-5/7}$, which tells us the distribution will be 
-highly biased towards low-light ratio binaries, as shown in 
-Fig.~\ref{fig:gammaR_distribn_mag_limited}.
-
-Our original reason for deriving these expression was that we wanted to know 
-the number of searchable binary stars.
-For any given realization of the survey, this number is a random draw.
-We can compute the \textit{expected value} for the number of double star 
-systems in the sample,
-\begin{equation}
-\langle N_d \rangle = \int_{0}^{\infty} N_d\, {\rm prob}(N_d) \,{\rm d}N_d.
-\end{equation}
-Letting $\alpha=3.5$, this becomes
-\begin{equation}
-\langle N_d \rangle \propto \mathcal{R}^{-3/2}
-\int_{0.1^{7/2} + 1}^{2} u^3 (u-1)^{-5/7} \,{\rm d}u ,
-\label{eq:integral_Nd}
-\end{equation}
-where $u = \mathcal{R} N_d^{2/3}$,
-non-dimensional constants are not shown,
-and $\mathcal{R}$ is defined as
-\begin{equation}
-\mathcal{R} \equiv \left(\frac{3}{4\pi n_d}\right)^{2/3} d_{\rm max,s}^{-2}
-= \left(\frac{n_s}{N_s n_d}\right)^{2/3}.
-\end{equation}
-The integral in Eq.~\ref{eq:integral_Nd} has a closed-form solution in terms 
-of polynomials. Once evaluated, and after 
-normalizing ${\rm prob}(N_d)$, to 4 significant digits it gives
-\begin{equation}
-\langle N_d \rangle = 1.590 \left(\frac{4\pi n_d}{3}\right) d_{\rm max,s}^3.
-\end{equation}
-This gives the useful expression
-\begin{align}
-\frac{\langle N_d \rangle}{N_s} &= 1.590\, \frac{n_d}{n_s}
-\label{eq:beta_model2}
-\end{align}
-Taking the sun-like star binary fraction of ${\rm BF} = 0.44$, this gives
-$\langle N_d \rangle / N_s = 1.25$.
-Compared to Eq.~\ref{eq:beta_model1} from the binary-twin thought experiment, 
-while the SNR-limited sample previously had $\approx 2.8$ times more binary 
-systems than single systems, there are now $\approx 1.25$ times as many.
-This is sensible, because the mean light ratio has decreased; it was 
-previously 1, it is now $\langle \gamma_R \rangle \approx 0.27$.
-
-\paragraph{What is the true occurrence rate?}
-
-Returning to the original question, the total number of stars for any given 
-realization of the survey is
-\begin{equation}
-N_{\rm stars,tot} = N_s + 2 N_d.
-\end{equation}
-Since $N_d$ is a random number, so is the total number of stars.
-By assumption all primaries are of the desired type, but for secondaries only 
-a fraction $f_{\rm desired}$ of them are of the desired type:
-\begin{equation}
-N_{\rm stars} = N_s + (1+f_{\rm desired}) N_d.
-\end{equation}
-
-For a given survey draw, the number of planets of the desired type is 
-then
-\begin{equation}
-N_{\rm planets} = 
-\Gamma_{t,s}N_s + \Gamma_{t,d} N_d + 
-f_{\rm des}\Gamma_{t,d}w_{d2} N_d,
-\end{equation}
-and the drawn ``true occurrence rate'' is
-\begin{align}
-\Gamma_t &= \frac{ N_{\rm planets}  }{ N_{\rm stars} } \\
-\Gamma_t &= 
-\frac{
-	\Gamma_{t,s}N_s + (1+f_{\rm desired}w_{d2})\Gamma_{t,d} N_d
-	}{
-	N_s + (1+f_{\rm desired}) N_d
-	}.
-\end{align}
-
-% % % % % % % % % % % % % % % % % % % % % % % % % % % %
-\subsection{What is the measured occurrence rate?}
-
-The observer who has never heard of binary star systems wants to find the 
-occurrence rate of planets with radius $R_p$ and period $P$, orbiting stars 
-with mass $M_1$.
-They detected $N_{\rm det}$ transit signals that appear to be from planets of 
-this desired type.
-The occurrence rate they then find, {\it i.e.} the apparent number of planets 
-divided by the apparent number of ``stars'' (really, stellar systems) is
-\begin{comment}
-\begin{equation}
-\Gamma_{\rm a } = \frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_g}.
-\end{equation}
-\end{comment}
-\begin{equation}
-\Gamma_{\rm a } =
-\frac{N_{\rm det}}{N_s + N_d} \times \frac{1}{f_{s,g}}.
-\end{equation}
-
-Unlike in Sec.~\ref{sec:model_1}, where the observer saw signals of only two 
-apparent radii (undiluted and diluted), they will now find signals with a 
-spectrum of apparent radii.
-The dilutions $\mathcal{D}$ in binary systems, given by Eq.~\ref{eq:dilution}, 
-vary depending on which star the planet orbits, and depending on the system's 
-luminosity ratio $\gamma_R$.
-Moreover, the apparent planetary radii will be affected by the assumptions the 
-observer makes about the apparently-single star's radius, $R_{\star,{\rm a}}$.
-We posit that the observer assumes $R_{\star,{\rm a}} = R_1$; in other words 
-they assume that the primary star hosts any signal they observe.
-The apparent planetary radii, $R_{\rm p, a}$, are then
-\begin{equation}
-R_{p,{\rm a}} = R_p 
-\frac{R_1}
-{R_{\star,{\rm t}}}
-\,\mathcal{D}^{1/2}.
-\end{equation}
-
-The drawn total number of planet detections is 
-\begin{equation}
-N_{\rm det} = 
-N_{{\rm det},s} 
-+ 
-N_{{\rm det},d}.
-\end{equation}
-Formally, if we are only interested in planets of radius $R_p$ then no planets 
-around binaries are ``of the desired type'' because of dilution.
-For the following development, ignore this caveat~--~consider an 
-observer who is interested in planets of any apparent radius.
-
-Then as in Sec.~\ref{sec:model_1},
-\begin{equation}
-N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,g} f_{s,c},
-\label{eq:N_det_s_model2}
-\end{equation}
-where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
-star systems of the sample, $f_{s,g}$ is the geometric transit probability, 
-and $f_{s,c}$ is the fraction of these transiting planets that are observed 
-with signal to noise greater than the minimum detection threshold (the 
-completeness). 
-For our SNR-limited, varying-$\gamma_R$ survey, $f_{s,g} = R_1/a\equiv f_g$, 
-and $f_{s,c}=1$.
-
-For planets detected in binary systems,
-\begin{equation}
-N_{{\rm det},d}
-= 
-N_d \Gamma_{t,d} (f_{d1,g}f_{d1,c} + f_{\rm desired}w_{d2}f_{d2,g}f_{d2,c}).
-\label{eq:N_det_d_model2}
-\end{equation}
-The factor $f_{\rm desired}$, as previously discussed, accounts for the 
-fraction of secondaries not of the desired {\it stellar} type. The above 
-counts the total number of detected planets around stars of the desired type, 
-though all these planets have diluted radii.
-The detection and geometric transit probabilities for a given system are
-\begin{align}
-f_{d1,c} &= (1+\gamma_R)^{-3} \\
-f_{d1,g} &= f_{s,g} \\
-f_{d2,c} &= (1+\gamma_R^{-1})^{-3} \gamma_R^{-5/\alpha} 
-\label{eq:f_d2c}\\
-f_{d2,g} &= f_{s,g}q^{2/3},
-\end{align}
-where typically $\alpha=3.5$.
-To derive Eq.~\ref{eq:f_d2c}, one must note that for fixed period, planet 
-radius, and signal to noise, 
-$$d_{\rm det}\propto \mathcal{D} R_2^{-2} L_d^{1/2} T_{\rm dur}^{1/2}.$$
-For a typical binary drawn from the local stellar neighborhood, $\gamma_R 
-\approx 0.2$ (see Fig.~\ref{fig:gammaR_distribn_mag_limited}), and there is an 
-interesting asymmetry.
-One finds the completeness for planets which orbit the primary star, 
-$f_{d1,c}$ to be
-\begin{equation}
-f_{d1,c} \sim 70\% \quad(\alpha=3.5, \gamma_R=0.2).
-\end{equation}
-However, for planets which orbit the secondary star, the completeness is
-\begin{equation}
-f_{d2,c} \sim 7\% \quad(\alpha=3.5, \gamma_R=0.2).
-\end{equation}
-This is the obvious statement that controlling for all other variables, 
-in the typical binary system planets around secondaries are more difficult to 
-detect than those around primaries.
-
-\subsection{Errors on derived occurrence rates}
-
-Recall $X_\Gamma = \Gamma_t / \Gamma_a$. Let $f_{\rm desired} = f_{\rm d}$ for 
-brevity.
-Combining the above results, for any draw from the ensemble of transit 
-surveys the necessary correction to the derived occurrence rate is
-\begin{align}
-X_\Gamma =
-&\frac{\Gamma_{t,s}N_s + (1+f_{\rm d}w_{d2})\Gamma_{t,d}N_d }{N_s + (1+f_{\rm 
-d})N_d}
-\nonumber \\
-&\times
-\frac{(N_s + N_d)f_{s,g}}{N_s \Gamma_{t,s} f_{s,g} + N_d \Gamma_{t,d}
-      (f_{d1,g}f_{d1,c} + f_{\rm d}w_{d2}f_{d2,g}f_{d2,c})  }.
-\end{align}
-Letting $\beta = N_d/N_s$, 
-\begin{align}
-X_\Gamma =
-&\frac{1+\beta}{1 + \beta(1+f_{\rm d})} 
-\nonumber \\
-&\times
-\frac{ 1+ \beta (\Gamma_{t,d}/\Gamma_{t,s}) (1+f_{\rm d}w_{d2}) }
-{1 +\beta(\Gamma_{t,d}/\Gamma_{t,s})(f_{d1,c} + f_{\rm 
-d}w_{d2}f_{d2,g}f_{d2,c}/f_{s,g})}.
-\end{align}
-
-One typically assumes that $\Gamma_{t,s}=\Gamma_{t,d}$, and that the 
-fraction of secondaries of the desired type with planets, $w_{d2}$ can be set 
-to 1.
-If so,
-\begin{align}
-X_\Gamma =
-&\frac{1+\beta}
-{1 +\beta(f_{d1,c} + f_{\rm d}f_{d2,g}f_{d2,c}/f_{s,g})}.
-\label{eq:model2_xgamma_simple}
-\end{align}
-
-If $\gamma_R \ll 1$, $f_{d1,c}\approx 1$, $f_{d2,c}\ll f_{d1,c}$, and the 
-number of detections from secondaries of binary systems is negligible.
-In this limit, binarity does not affect the derived occurrence rate:
-\begin{equation}
-X_\Gamma = 1 \quad ({\rm limit\ of}\ \gamma_R \ll 1).
-\end{equation}
-
-For the stellar population described in this model, with a mean light ratio 
-$\langle \gamma_R \rangle \approx 0.27$, this limit is not applicable.
-To assess the importance of the error, we want the mean
-value of $X_\Gamma$, or the median, or else some other summary statistic.
-Recall that all the terms in Eq.~\ref{eq:model2_xgamma_simple} are
-random variables, and note that they are inter-dependent.
-This latter fact impedes us from analytically expressing these 
-descriptive statistics.
-
-Instead, we run a Monte Carlo simulation in which we fix the planet and 
-primary star properties, draw the binary mass ratio from 
-Eq.~\ref{eq:pdf_q_mag_limited}, 
-and take the luminosity as $L\propto M^{3.5}$. We find a mean occurrence 
-correction of 
-\begin{equation}
-\langle X_\Gamma \rangle =1.268\quad ({\rm lower\ limit\ from\ Monte\ Carlo}),
-\label{eq:X_Gamma_model2_lower}
-\end{equation}
-and a median that is the same to within one percent.
-This value is comparable to what one would guess using the analytic results 
-for $\langle N_d \rangle / N_s$ to substitute $\beta$, and $\gamma_R$ to 
-compute $f_{d1,c}$.
-
-%WARNING
-%WARNING
-%the reported value of <N_d>/N_s is higher from analytics than vs what 
-%simulations give. means a mistake in one of them.
-
-\paragraph{Varying what is meant by ``planets of the desired type'' }
-The above occurrence rate correction factor is a lower limit for this model, 
-because if we were to be any stricter about the range of allowed apparent 
-radii, the apparent occurrence rate would only decrease.
-
-Taking the opposite limit of not counting any detections that deviate 
-from $R_p$ as ``planets of the desired type'',
-we find
-\begin{equation}
-X_{\Gamma} = \frac{1 + \beta }{1 + \beta f_{d1,c}}.
-\end{equation}
-From the same Monte Carlo simulation, this gives an upper limit on the 
-correction term
-\begin{equation}
-\langle X_\Gamma \rangle = 2.114\quad ({\rm upper\ limit\ from\ Monte\ 
-Carlo}).
-\label{eq:X_Gamma_model2_upper}
-\end{equation}
-
-The general point is that starting from a single planet radius, dilution 
-produces a spectrum of apparent planetary radii. 
-This skews occurrence rate measurements, by an amount between 
-Eqs.~\ref{eq:X_Gamma_model2_lower} and~\ref{eq:X_Gamma_model2_upper},
-depending on the radius bounds set by the analyst.
-We investigate this further in our final model below.
-
-\section{Model \#3: allow a distribution of planetary radii}
-\label{sec:model_3}
-
-Keep the same stars as in Sec.~\ref{sec:model_2}.
-Now let the planets have a power-law radius distribution, ${\rm 
-d}(N_p/N_\star)/{\rm 
-d}R_p\propto R_p^\delta$. 
-We want to know the effects of binarity on the apparent distribution of 
-planetary radii.
-Specifically, the average number of planets in a given size range per star of 
-the desired type in this universe is
-\begin{equation}
-\frac{N_{\rm p}(R_p^{\rm \ell}<R_p <R_p^{\rm u})}{N_{\rm \star}} = 
-\Gamma_0 \int_{R_p^{\rm \ell}}^{R_p^{\rm u}} R_p^\delta \,{\rm d}R_p,
-\end{equation}
-for $\Gamma_0$ one of $\{\Gamma_{t,s}, \Gamma_{t,d}, w_{d2}\Gamma_{t,d}\}$
-corresponding to single star systems, the primaries of double star systems, 
-and the secondaries of double star systems.
-
-\section{Discussion}
-\label{sec:discussion}
-
-\appendix
-\begin{figure}[!t]
-    \begin{center}
-        \includegraphics[scale=0.9]{figures/mass_luminosity.pdf}
-    \end{center}
-    \caption{Empirical fit to main sequence dwarf mass luminosity data 
-    compiled 
-        from Torres et al. [2009] and Benedict et al. [2016]. The ``low fit'' 
-        is a 
-        least squares fit to data from $0.1-0.5M_\odot$, and the ``high fit'' 
-        is 
-        to 
-        data above that, and below the Kraft break ($1.2M_\odot$).
-        The $L(M)$ relation taken for numerics is the maximum 
-        of the two fits. For analytics, we assume $L\propto M^{3.5}$.
-    }
-    \label{fig:mass_luminosity}
-\end{figure}
-
-\begin{comment}
-\subsection{How many planets are in the sample?}
-
-The mean number of planets in the sample is
-\begin{align}
-\langle N_{\rm planets} \rangle
-=& N_{\rm planets\ in\ single\ star\ systems}  +  \\
-&\quad\quad N_{\rm planets\ in\ double\ star\ systems} 
-\nonumber \\
-=& \Gamma_{t,s} N_s + (\Gamma_{t,d1} + \Gamma_{t,d2}) \langle N_d \rangle.
-\end{align}
-
-What was formerly a factor of 2 has now been split into the true occurrence 
-rates about the primaries and secondaries of double star systems.
-
-\subsection{What is the true occurrence rate?}
-
-The ``true occurrence rate'' is the average number of planets per star. Thus
-
-\begin{align}
-\Gamma_t &= \frac{\langle N_{\rm planets}\rangle}{\langle N_{\rm stars} 
-\rangle} \\
-\Gamma_t &= \frac{\Gamma_{t,s} N_s + (\Gamma_{t,d1} + \Gamma_{t,d2}) \langle 
-N_d\rangle}{N_s +	2\langle N_d\rangle}.
-\label{eq:true_occ_2}
-\end{align}
-
-\subsection{How many planets are detected?}
-
-The total number of planet detections is the sum of the number of planets 
-detected in single star systems $N_{{\rm det},s}$ and the number of planets 
-detected in double star systems $N_{{\rm det},d}$.
-The latter of these is a random variable, and can be further split into the 
-primary and secondary contributions $N_{{\rm det},d1}$, $N_{{\rm det},d2}$.
-
-The number of planets detected in single star systems is
-\begin{equation}
-N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,g} f_{s,c},
-\end{equation}
-where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
-star systems of the sample, $f_{s,g}$ is the geometric probability of the 
-planets transiting, and $f_{s,c}$ is the fraction of transiting planets around 
-single star systems that are detected (the completeness).
-
-The mean number of planets detected in double star systems is
-\begin{equation}
-\langle N_{{\rm det},d}\rangle = \langle N_d\rangle  
-(\Gamma_{t,d1} f_{d1,g} f_{d1,c} + 
-\Gamma_{t,d2} f_{d2,g} f_{d2,c} ).
-\label{eq:N_det_d_2}
-\end{equation}
-Now $\langle N_d\rangle  \Gamma_{t,d1}$ is the mean number of planets orbiting 
-the primaries of double star systems, ditto the corresponding expression for 
-the secondaries.
-For this population, $f_{d1,g} = f_{s,g}$, but $f_{d1,g} \neq f_{d2,g}$.
-The completeness fractions differ between the primary and secondary because 
-the dilution as a function of the light ratio, $\mathcal{D}(\gamma_R)$, has 
-different behavior for the two cases.
-We show this in Fig.~\ref{fig:rad_vs_gammaR}.
-Specifically, a planet orbiting the secondary will have worse dilution than a 
-planet orbiting the primary, and consequently lower completeness.
-
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.9]{figures/observed_radii_vs_gammaR.pdf}
-	\end{center}
-	\caption{Dilution $\mathcal{D}$ and observed to true radius ratio 
-	$\sqrt{\mathcal{D}}$ plotted against binary light ratio $\gamma_R$. This 
-	illustrates	Eq.~\ref{eq:dilution}. The lower panel's horizontal line is at 
-	$1/\sqrt{2}$.
-	The implication is that if a detected planet orbits the secondary and you do 
-	not know it, you cannot measure $R_p$ to better than $\approx 29\%$ 
-	accuracy.
-	Conversely, if it orbits the primary and you do not know it, you will get 
-	the radius wrong by at most $\approx 29\%$.
-	}
-	\label{fig:rad_vs_gammaR}
-\end{figure}
-
-We could probably derive analytic expressions for all the terms between 
-brackets in Eq.~\ref{eq:N_det_d_2}, completeness included.
-I'm not yet convinced we need to -- we should get cancellations for
-most of the questions we want to ask.
-
-\subsubsection{Analytic completeness}
-\texttt{todo, if necessary}
-\subsubsection{Deriving ${\rm prob}(x_i)$}
-\texttt{todo, if necessary}
-\subsubsection{Number of detected planets}
-\texttt{todo, if necessary}
-
-\subsection{Astronomer A ignores binarity}
-
-Just as in Sec.~\ref{sec:model_1}, Astronomer A
-\begin{itemize}
-\item has never heard about binary star systems.
-\item has never heard about completeness corrections.
-\item knows about geometric transit probabilities.
-\end{itemize}
-
-What occurrence rate does he derive for planets of radius $R_p$ and period $P$?
-
-The answer is the same as in Sec.~\ref{sec:model_1},
-\begin{equation}
-\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/f_{s,g}}{N_s + N_d}.
-\end{equation}
-
-However, now rather than planets detected in binary systems being perceived as 
-a second population of planets with fixed radius $R_p \sqrt{\mathcal{D}}$, 
-there will be an apparent spectrum of diluted radii, since $\mathcal{D}$ 
-varies by system.
-Writing the number of detections in double systems as a function of observed 
-radius, $N_{\rm det,d}(R_p^{\rm obs})$, Astronomer A would derive an 
-occurrence rate for this separate population of
-\begin{equation}
-\Gamma_{\rm A}(R_p^{\rm obs}) = \frac{N_{\rm det,d}(R_p^{\rm 
-obs})/f_{s,g}}{N_s + N_d}.
-\end{equation}
-Note that this astronomer thinks these planets orbit single stars, 
-and so miscomputes the geometric transit probability for the secondaries.
-
-A more realistic question is ``what fraction of multiples have $\gamma_R$ 
-smaller enough so that you'd still observe a radius close to the true one?''. 
-We discuss this in Sec.~\ref{sec:close_enough}.
-
-\subsection{Astronomer B counts host stars correctly}
-
-Same as astronomer A above, except the denominator becomes $N_s + 2N_d$.
-
-\subsection{Astronomer C counts host stars correctly and figures out diluted
-radii}
-
-In Sec.~\ref{sec:model_1}, Astronomer C only had to do high resolution 
-imaging to measure $\gamma_R$ for each system. Since for $\gamma_R = 1$, the 
-dilution is single-valued (Fig.~\ref{fig:rad_vs_gammaR}), this immediately 
-told her the true planet radius (since the primary/secondary distinction was 
-meaningless).
-
-In Model \#2's universe, the dilution is double-valued for a given $\gamma_R$.
-To correctly work out the diluted radii, Astronomer C now needs more than high 
-resolution imaging -- she needs to resolve the stars during transit to 
-discover which star the planet orbits.
-She also needs to derive masses for the primaries and secondaries (to make the 
-geometric transit probability correction).
-
-Only after doing this does she find that all detected planets from this survey 
-have radii $R_p$. She computes an occurrence rate
-\begin{equation}
-	\Gamma_{{\rm C},R_p} = 
-	\frac
-			{N_{\rm det,s}/f_{s,g} +
-			  N_{\rm det,d1}/f_{d1,g} +
-			  N_{\rm 	det,d2}/f_{d2,g}}
-			{N_s + 2N_d}
-\end{equation}
-
-the closest yet to the true rate (Sec.~\ref{sec:true_rate}).
-
-\subsection{Astronomer D counts host stars correctly, figures out diluted
-radii, and accounts for completeness}
-
-Same story as for Model \#1.
-Now, they do injection recovery and derive correct estimates for their 
-completeness functions about single stars $f_{s,c}$, the primaries of double 
-star systems, $f_{d1,c}$, and the secondaries of double star systems 
-$f_{d2,c}$.
-They compute the respective single and binary 
-occurrence rates
-\begin{equation}
-\Gamma_{t,s} = \frac{N_{\rm det,s}}{N_s f_{s,g} f_{s,c}},
-\end{equation}
-\begin{equation}
-\Gamma_{t,d1} = \frac{N_{\rm det,d1}}{N_d f_{d1,g} f_{d1,c}},
-\end{equation}
-\begin{equation}
-\Gamma_{t,d2} = \frac{N_{\rm det,d2}}{N_d f_{d2,g} f_{d2,c}}.
-\end{equation}
-
-With these in hand, they derive the overall occurrence rate
-\begin{align}
-\Gamma_{{\rm D}, R_p} 
-&= \frac{
-	N_{\rm det,s}/f_s + 
-	N_{\rm det,d1}/f_{d1} +
-	N_{\rm det,d2}/f_{d2}
-	}
-	{N_s + 2 N_d} \\
-&= \frac{\Gamma_{t,s} N_s + (\Gamma_{t,d1} + \Gamma_{t,d2})N_d}
-				{N_s + 2 N_d},
-\end{align}
-where $f_s \equiv f_{s,g}f_{s,c}, f_{d1} \equiv f_{d1,g}f_{d1,c}, 
-f_{d2} \equiv f_{d2,g}f_{d2,c}$ for brevity.
-Astronomer D's occurrence rate is the true occurrence rate (cf. 
-Eq.~\ref{eq:true_occ_2}).
-
-\subsection{Numerical verification}
-\texttt{todo, if needed}
-
-\subsection{Representative numbers for a few cases}
-
-We have not given analytic expressions for the number of double star systems 
-$N_d$ because even with simplified $M(L)$ relations, they are unwieldy.
-However, we can ask similar numerical questions as from Sec.~\ref{sec:model_1}, 
-and compare the answers.
-
-\subsubsection{If we ignore binarity, for what fraction of detections do we
-misclassify the radii?}
-
-Ignoring binarity, we will detect $N_{\rm det,s}$ planets around single stars, 
-and $N_{\rm det,d}$ planets around double stars. The latter set will be assumed 
-to have radii $R_p \sqrt{\mathcal{D}}$, where $\mathcal{D}$ is the vector of 
-dilutions appropriate for each binary star system.
-The fraction of detections with misclassified radii can then be written
-\begin{equation}
-\frac{N_{\rm det,d}}{N_{\rm det,s} + N_{\rm det,d}} = \frac{1}{1+\alpha},
-\end{equation}
-for
-\begin{equation}
-\alpha \equiv 
-\frac{ N_d (\Gamma_{t,d1} f_{d1} + \Gamma_{t,d2} f_{d2}) }{N_s \Gamma_{t,s} 
-f_s},
-\end{equation}
-where $f_s \equiv f_{s,g}f_{s,c}, f_{d1} \equiv f_{d1,g}f_{d1,c}, 
-f_{d2} \equiv f_{d2,g}f_{d2,c}$ for brevity.
-
-% For the nominal G2V dwarf case of ${\rm BF}=0.45$, and the same distributions 
-% used to generate 
-% Figs.~\ref{fig:gammaR_distribn_vol_limited}~--~\ref{fig:gammaR_distribn_mag_limited},
-% we found $N_d/N_s=0.59$.
-% We are assuming $\Gamma_{t,d} / \Gamma_{t,s} = 1$.
-% If we maintained the (incorrect) assumption that $f_{d,{\rm geom}} / f_{s,{\rm 
-% 		geom}} = 1$, this would give a misclassification rate of 45\%. 
-% 
-% This latter assumption is wrong because although we are letting the occurrence 
-% rate be the same across binary and single systems, our binary systems have 
-% secondaries with varying masses. Thus they have varying radii.
-% Taking [Demircan \& Kahraman 1991]'s empirical fit to eclipsing binary data,
-% \begin{equation}
-% R = 1.06 M^{0.945}
-% \label{eq:mass_radius}
-% \end{equation}
-% for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 1991's 
-% stated bounds).
-% For a zero eccentricity system, $f_{\rm geom}=R_\star/a$.
-% Since $f_{d,{\rm geom}} = 0.5\times (f_{d1,{\rm geom}} + f_{d2,{\rm geom}})$, 
-% \textit{i.e.} it is the average of the transit probabilities for both the 
-% primary and secondary (d1 and d2), we can write (dropping the ``geom'' 
-% subscript)
-% \begin{align}
-% \frac{f_d}{f_s} &= \frac{1}{2} \left(1 + \left(\frac{M_1}{M_{d2}}\right)^{1/3} 
-% \frac{R_{d2}}{R_1}\right) \\
-% &= \frac{1}{2} \left( 1 + q^{-1/3 + 0.945}\right).
-% \end{align}
-% 
-% Computing it all out in a numerical simulation, the only thing that changes 
-% analytically is that
-% \begin{equation}
-% N_d f_d \rightarrow \sum_{i=1}^{N_d} f_{d,i},
-% \end{equation}
-% so when considering the ratio $f_d/f_s$, we are in fact interested in
-% \begin{equation}
-% \frac{\langle f_d \rangle}{f_s} = \frac{\frac{1}{N_d} \sum_{i=1}^{N_d} 
-% 	f_{d,i}}{f_s},
-% \end{equation}
-% where $f_{d,i}$ is the vector of geometric transit probabilities across all 
-% systems.
-% For the same numerical realization with $N_d/N_s = 0.59$, and using the mass 
-% radius relation given by Eq.~\ref{eq:mass_radius}, I get $\langle f_d 
-% \rangle / f_s = 0.84$ (smaller secondaries lead to a smaller average transit 
-% probability in double star systems), for a radius misclassification rate of 
-% $0.51$.
-
-\texttt{The different secondary completeness matters here if proceeding 
-analytically. Would be easier to do numerically.
-Regardless, this is not the most interesting question right now.}
-
-\subsubsection{If we ignore binarity, how wrong is our occurrence rate for
-planets of radius $R_p$?}
-\label{subsubsec:model2_A}
-
-Write $\Gamma_{{\rm A,\ planets\ of\ }R_p} = \Gamma_{{\rm A,}R_p}$, 
-and similarly for ${\rm D}$.
-Just as in Model \#1, the answer to ``what is the relative difference between 
-the occurrence rates derived by Astronomers D and A for planets of radius 
-$R_p$?''
-is simpler when we assume that Astronomer A has also derived a 
-completeness, which we assume is the same as for Astronomer D in the single 
-star case.
-So Astronomer A now misclassifies planetary radii, and miscounts the total 
-number of stars, but somehow knows his completeness for single 
-stars. This is possible, because Astronomer A can correctly derive a 
-completeness estimate and geometric transit probability for single stars only. 
-For Astronomer $A'$ below, including the planets in binary star systems in the 
-numerator introduces mistakes in the estimated completeness, and so the picture 
-is more complicated.
-Then
-\begin{equation}
-	\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}/(f_{s,g} f_{s,c})}{N_s + N_d}.
-\end{equation}
-
-The correction factor $X_\Gamma$ is
-\begin{align}
-X_\Gamma &=
-\frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p}} \\
-&=
-\frac{\Gamma_{t,s}N_s + (\Gamma_{t,d1} + 
-	\Gamma_{t,d2})N_d}{N_s + 2N_d}  
-\cdot 
-\frac{N_s + N_d}{\Gamma_{t,s} N_s } \\
-&= 
-\frac{(1 + \beta (\Gamma_{t,d1}+\Gamma_{t,d2})/\Gamma_{t,s})(1 + 
-	\beta)}{(1+2\beta)} 	\\
-&=
-(1 + \chi \beta )
-\cdot 
-\frac{1 + \beta}{1 + 2\beta }
-\end{align} 
-for
-\begin{equation}
-\beta \equiv N_d/N_s, \quad \chi \equiv 
-(\Gamma_{t,d1}+\Gamma_{t,d2})/\Gamma_{t,s}.
-\end{equation}
-
-In the limit where the $R_p$ planet has the same occurrence rate about every 
-host type regardless of its mass, $\chi=2$, and this simplifies to $X_\Gamma = 
-1+\beta$.
-This is same expression as what we found for Model \#1, but now the value of 
-$\beta$ will be smaller in this model because of the 
-smaller volume of searchable binary stars.
-From numerics, we find $\beta = N_d/N_s = 0.59$, so the occurrence rate 
-correction factor is $X_\Gamma(\chi=2) = 1.59$.
-
-If we consider the opposite limit of the $R_p$ planet \textit{only} 
-existing around single stars and the primaries of double star systems with 
-equal occurrence, $\chi = 1$, and $X_\Gamma = (1 +\beta)^2 / (1+2\beta)$.
-For $\beta = 0.59$, this gives $X_\Gamma(\chi=1) = 1.16$.
-The true correction factor is somewhere between these two limits.
-
-\subsubsection{What if we ignore binarity, but count derived planet radii that 
-are ``close enough''?}
-\label{sec:close_enough}
-
-Let $x$ be the acceptable margin of error on the observed radius, so that we 
-are interested in counting the number of planets detected with $R_p < R_p^{\rm 
-obs} < (1\pm x) R_p$.
-For Model \#2, since all planets have radius $R_p$, we only care about 
-the smaller radius case, \textit{i.e.}, we want to count $R_p > R_p^{\rm 
-	obs} > (1 - x) R_p$, for instance with $x=0.1$.
-
-The diluted detections only occur in binary systems. The number of 
-planets detected in double star systems with observed radius $R_p^{\rm obs}$ 
-is
-\begin{align}
-N_{\rm det,d}(R_p^{\rm obs}) &= N_{\rm det, d1}(R_p^{\rm obs}) + 
-								N_{\rm det, d2}(R_p^{\rm obs}) \\
-&= \Gamma_{t,d1} f_{d1}(R_p^{\rm obs}) N_{d1}(R_p^{\rm obs})\  + \nonumber \\
-&\quad\quad \quad\Gamma_{t,d2} f_{d2}(R_p^{\rm obs}) N_{d2}(R_p^{\rm obs}).
-\end{align}
-As always, the fractional terms $f_{di}$ for $i\in{1,2}$ encompass geometric 
-and completeness corrections.
-We're writing both them and the number of stars for which a given observed 
-radius would be seen as functions of $R_p^{\rm obs}$.
-
-To save on notational horribleness, write $R_p' \equiv \{(1-x) R_p < R_p^{\rm 
-obs} < R_p\}$, in other words let $R_p'$ denote the set of planets that are 
-observed in the desired radius range.
-
-Then writing $N_{d1}(R_p')$, the number of double systems for which a $(R_p, 
-P)$ planet would be observed in $R_p'$ if it orbiting the primary, is 
-relatively straight-forward:
-\begin{equation}
-N_{d1}(R_p') = \int_{0}^{{\rm min}(1,\gamma_{R,u})} N_d(\gamma_R) 
-\,{\rm prob}(\gamma_R) \,{\rm d}(\gamma_R),
-\label{eq:N_d1}
-\end{equation}
-where the upper limit is set by equating the limiting radius $(1-x)R_p$ to the 
-observed diluted radius $(1+\gamma_R)^{-1/2}R_p$, giving
-\begin{equation}
-\gamma_{R,u} \equiv \frac{x (2-x)}{(1-x)^2}.
-\end{equation}
-If $x > (1 - 2^{-1/2})$, bad things happen (cf. Fig.~\ref{fig:rad_vs_gammaR}), 
-hence the need for the minimum.
-Note that Eq.~\ref{eq:N_d1} is not an expression for the number of detected 
-planets about primaries -- that quantity is expressed as $\Gamma_{t,d1} 
-f_{d1}(R_p^{\rm obs}) N_{d1}(R_p^{\rm obs})$.
-
-The analogous equation to Eq.~\ref{eq:N_d1} for $N_{d2}(R_p')$, the number of 
-double systems for which a $(R_p, P)$ planet would be observed in $R_p'$ if it 
-orbited the secondary, is
-\begin{equation}
-N_{d2}(R_p') = 
-  \Bigg\{\begin{array}{lr}
-  \int_{\gamma_{R,l}}^{1} N_d(\gamma_R)
-  \,{\rm prob}(\gamma_R) \,{\rm d}(\gamma_R),\\ 
-  \quad\quad\quad\quad\quad\quad\quad\quad x > 1 - 
-  2^{-1/2}\\
-  0, \quad\quad\quad\quad\quad\quad\quad x \leq 1 - 2^{-1/2} % lol
-  \end{array}
-\label{eq:N_Rp_secondary}
-\end{equation}
-where
-\begin{equation}
-\gamma_{R,l} \equiv \frac{(1-x)^2}{x (2-x)}.
-\end{equation}
-Eq.~\ref{eq:N_Rp_secondary} accounts for the fact that you cannot measure 
-$R_p$ to better than $\approx 29\%$ if the planet orbits the secondary of an 
-unidentified binary.
-This was originally noted in Fig.~\ref{fig:rad_vs_gammaR}.
-
-If we were forward-modelling, \textit{i.e.} if we actually wanted to compute 
-$N_{\rm det,d}(R_p')$, we would also need to evaluate the total detection 
-efficiency term $f_{d1}(R_p^{\rm obs})$, which is a product of the geometric 
-transit probability and the fraction of signals of a given depth that are 
-detectable.
-We would need to integrate this product over the domain of allowable radii 
-$R_p'$ to find the number of detected planets.
-
-However, the procedure of estimating the geometric transit probability, as well 
-as the completeness, is complicated by the errors that are introduced by 
-including binary star systems that are thought to be single in the numerator.
-These stars will have incorrectly estimated masses and radii, which influences 
-both $f_c(R_p')$ and $f_g(R_p')$.
-
-\texttt{If we ask the right question, we will not need to actually compute 
-either term.}
-
-Astronomer A misclassifies planetary radii, miscounts the total number of 
-stars, and (incorrectly) estimates his completeness and geometric correction 
-for ``single'' stars as a function of radius $f(R_p')$ (omitting a ``s'' 
-subscript because the stars for which this incorrect occurrence rate are 
-derived are not all single).
-His occurrence rate for planets of radius \textit{near} $R_p$, \textit{i.e.} 
-the set of planets with $\{ (1-x)R_p < R_p^{\rm obs} < R_p \}$, is
-\begin{equation}
-\Gamma_{{\rm A}, R_p'} = \frac{ (N_{\rm det,s} + N_{\rm det,d}(R_p')) 
-/ f(R_p')   }{ N_s + N_d}.
-\end{equation}
-
-For him,
-\begin{equation}
-f(R_p') = f_g(R_p') f_c(R_p').
-\end{equation}
-
-Astronomer D, who had everything right, changes nothing.
-
-The new correction factor is
-\begin{align}
-X_\Gamma &= \frac{\Gamma_{{\rm D},R_p}}{\Gamma_{{\rm A},R_p'} } \\
-&=
-\frac{1+\chi \beta}{1 + \xi}  \cdot 
-\frac{1 + \beta}{1 + 2\beta},
-\end{align}
-for
-\begin{equation}
-\beta \equiv N_d/N_s, \quad \chi \equiv (\Gamma_{t,d1} + 
-\Gamma_{t,d2})/\Gamma_{t,s},
-\end{equation}
-and
-\begin{equation}
-\xi \equiv \frac{N_{\rm det,d}(R_p')}{f(R_p')} \cdot \frac{1}{N_s 
-\Gamma_{t,s}}.
-\end{equation}
-If we let $x < 1 - 2^{-1/2}$, for instance $x=0.1$, then $N_{d2}(R_p') = 0$. 
-The expression for $\xi$ in this case becomes
-\begin{align}
-\xi &= \frac{N_{d1}(R_p')}{N_s} \cdot \frac{\Gamma_{t,d1}}{\Gamma_{t,s}} \cdot 
-\frac{f_{d1}(R_p')}{f(R_p')} \\
-&= \frac{N_{d1}(R_p')}{N_s} \cdot \frac{\Gamma_{t,d1}}{\Gamma_{t,s}} \cdot 
-\frac{f_{g,d1}(R_p')}{f_g(R_p')}\cdot \frac{f_{c,d1}(R_p')}{f_c(R_p')},
-\label{eq:chi_simplified}
-\end{align}
-provided that $x < 1-2^{-1/2}$.
-
-We have now parametrized our ignorance.
-The values that the latter two ratios of Eq.~\ref{eq:chi_simplified} take 
-depend on what we assume Astronomer A' does to derive the stellar parameters 
-for the systems that he thinks are singles but are in fact doubles.
-For instance, assume he somehow observed the bolometric flux from every point 
-on the sky, and used it with a distance to obtain single-star luminosities, and 
-thus radii and masses.
-His binary systems (thought to be singles) would be bluer than either component 
-star is in reality.
-On the main sequence, $\rho_\star$ decreases with increasing stellar radius, 
-so $\rho_\star$ would be underestimated for all binaries.
-Since $f_g \propto \rho_\star^{1/3}$, this means that $f_{g,d1}(R_p')/f_g(R_p') 
-> 1$.
-However, by the same token we can place an \textit{upper} bound on this ratio. 
-In the worst case scenario a given star's luminosity is twice the individual 
-star's luminosity. If $L\propto M^3 \propto R^3$, then the radius is 
-over-estimated by at most a factor of $2^{1/3}$. Thus the density is 
-underestimated by at most a factor of $2^{1/2}$, and $f_g(R_p')$ is 
-underestimated by at most a factor of $2^{1/6}=1.12$.
-
-This begins to take us somewhere. The left-most term of 
-Eq.~\ref{eq:chi_simplified} is bounded above by $\beta$. This is because 
-$N_{d1}(R_p') \leq N_{d1} = N_d$.
-If we assume $\Gamma_{t,d1} = \Gamma_{t,s}$, as we were doing 
-anyway, then the product of the three left-most fractions must be less than 
-$1.12\beta$.
-The only thing standing in our way of an analytic upper bound to $\xi$ is the 
-completeness fraction $f_{c,d1}(R_p')/f_c(R_p')$.
-Note that over a fixed radius interval dilution will always 
-yield a lower completeness fraction for double star systems than singles, so 
-$f_{c,d1}(R_p') / f_{c,s}(R_p') \leq 1$.
-
-For the completeness term, an argument that I don't wholly believe
-is that
-\begin{equation}
-\frac{f_{c,d1}(R_p')}{f_c(R_p')} = \frac{f_{c,d1}(R_p')}{f_{c,s}(R_p')} \cdot
-	\frac{f_{c,s}(R_p')}{f_{c}(R_p')} \lesssim 1 \cdot (1+\langle \gamma_R 
-	\rangle)^3,
-\end{equation}
-where the first part of the inequality is obvious, but the second is by 
-assuming it's $\lesssim f_{s,c}/f_{d,c}$, which might be wrong because of 
-the radius dependence.
-
-A way of making the estimate without deriving the actual 
-completeness fractions, or even running numerics, is just to plot the error as 
-a function of $\xi$. We do this in Fig.~\ref{fig:XGamma_vs_xi}, which 
-shows that the correction factor varies from 0.6 to 1.6, depending on what 
-fraction of secondaries you assume host planets.
-\todo[inline]{I could run numerics for this to make the result more 
-interesting.}
-
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/XGamma_vs_xi.pdf}
-	\end{center}
-	\caption{\textit{Y axis}: correction factor for the astronomer who 
-	ignores binarity but counts planetary radii that are ``close enough'' (A') 
-	and the astronomer who knows everything (D).
-	\textit{X axis}: $\chi$ as defined in Eq.~\ref{eq:chi_simplified}.
-	}
-	\label{fig:XGamma_vs_xi}
-\end{figure}
-
-\section{Model \# 3: A Synthetic {\it Kepler} Analog}
-\label{sec:kepler_analog}
-
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=0.9]{figures/mass_luminosity.pdf}
-	\end{center}
-	\caption{Empirical fit to main sequence dwarf mass luminosity data compiled 
-		from Torres et al. [2009] and Benedict et al. [2016]. The ``low fit'' is a 
-		least squares fit to data from $0.1-0.5M_\odot$, and the ``high fit'' is 
-		to 
-		data above that, and below the Kraft break ($1.2M_\odot$).
-		The $L(M)$ relation taken for numerics is the maximum 
-		of the two fits. For analytics, we assume $L\propto M^{3.5}$.
-	}
-	\label{fig:mass_luminosity}
-\end{figure}
-
-If we assume every KIC star to be single, what order of magnitude of error do 
-we make in occurrence rates estimated for planets of different sizes (and 
-periods)?
-
-We can attempt an answer with Monte Carlo simulations of something like the 
-\kepler\ field. I say ``something like'' because the binarity properties of the 
-\kepler\ field (or more specifically the stars selected in the KIC) are not 
-known.
-\todo[inline]{but it might be better to do something directly on the KIC's 
-selected target stars, like what Ciardi did! The following is a good example, 
-but at most it can ``suggest'' that a similar level of error has been made from 
-the \kepler\ survey.
-
-But it has MAJOR benefits compared to the GIC method.
-It eliminates the need to simulate the biases of the KIC (since they're already 
-baked in). The process of ``adding'' binaries to the KIC is the closest thing 
-to REALITY, since in reality the KIC is missing binary companion information.
-It BAKES IN all the real biases that come with mis-estimating stellar 
-radii/masses when they are actually binaries (because it's already been done).
-It also has literature precedent in Ciardi's work, so you don't need to cook up 
-a crazy new procedure -- can just take his with the flaws baked in.
-
-So this method is best if we care about what errors were made in Kepler/the KIC.
-It's not as nice if we want something ``general''.
-}
-
-The plan is as follows:
-\begin{itemize}
-	\item Verify Galaxia can produce something resembling the KIC.
-	\item Introduce binaries (Galaxia does not, by default, have any).
-	\item Use Galaxia [Sharma et al 2010] to construct an analog of the KIC.
-	\item Introduce a planet population.
-	\item Numerically assess what occurrence rates would be derived, and 
-	compare to previous sections.
-\end{itemize}
-
-\subsection{Verifying Galaxia can produce a KIC analog }
-
-Galaxia [Sharma et al 2010] is a stellar population synthesis code for creating 
-surveys of the Milky Way. It samples an assumed analytic distribution function. 
-In other words, it assumes a number density of stars as a function of position, 
-velocity, age, metallicty, and mass, from which it then samples to construct a 
-``realistic'' stellar population.
-The assumed model includes:
-\begin{itemize}
-	\item A metallicity distribution (log-normal),
-	\item A velocity distribution (triaxial Gaussian),
-	\item The stellar density and galactic components specified by Robin et al 
-	[2003] (the Besan\c{c}on model) -- see Sharma et al [2010] Table 1. This 
-	includes the IMF for a young and old thin disk, a thick disk, a spheroidal 
-	component, a bulge component, as well as an ISM and dark halo. It also 
-	includes the star-formation rate.
-\end{itemize}
-
-While this model of course lacks some detail, Sharma et al [2010] show that 
-agreement with the volume-limited Hipparcos sample is probably good enough for 
-our purpose of estimating the number of ``solar-like'' (by definition 
-hereafter, $0.7 M_\odot < M_\star < 1.3 M_\odot$) stars in something like the 
-\kepler\ field.
-
-We then run the Galaxia in its ``circular survey'' mode, towards the direction 
-of the \kepler\ field\footnote{galactic longitude 76.532562, galactic latitude 
-13.289502.}. We apply the Schlegel, Finkbeiner et al [1999] extinction map when 
-converting absolute to apparent magnitudes.
-
-To verify the results are in reasonable agreement with the stars that actually 
-exist in that direction, we compare the output with the KIC\footnote{From 
-\url{http://archive.stsci.edu/kepler/kic.html}, we 
-downloaded the 13.1 million row "|"-delimited gzipped ASCII file containing the 
-complete Kepler Input Catalog (version 10).} [Brown et al 2011].
-Following Sharma et al [2016, ApJ 822:15], for each catalog we selected all 
-stars within $7.5^\circ$ of the center of the \kepler\ field, and with Sloan $r 
-< 14$.
-The KIC is complete to magnitudes fainter than $r=14$, so the comparison should 
-not be affected by its completeness.
-
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/rmag_distribn.pdf}
-	\end{center}
-	\caption{Probably density distributions of $r$ magnitude for $r<14$ stars 
-	within $7.5^\circ$ of the center of the \kepler\ field, for Galaxia and the 
-	KIC.	}
-	\label{fig:rmag_distribn}
-\end{figure}
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/rstar_more_distribn.pdf}
-	\end{center}
-	\caption{Same as Fig.~\ref{fig:rmag_distribn}, for $R_\star$.}
-	\label{fig:rstar_more_distribn}
-\end{figure}
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/teff_distribn.pdf}
-	\end{center}
-	\caption{Same as Fig.~\ref{fig:rmag_distribn}, for $T_{\rm eff}$.}
-	\label{fig:teff_distribn}
-\end{figure}
-
-A few distributions are shown in 
-Figs.~\ref{fig:rmag_distribn}-~\ref{fig:teff_distribn}.
-From Fig.~\ref{fig:rmag_distribn}, we can see that the stellar counts as a 
-function of Sloan $r$ magnitude are in close agreement with the KIC.
-From Fig.~\ref{fig:rstar_more_distribn}, the KIC reports relatively more dwarf 
-stars, and fewer giants than Galaxia does.
-This is likely related to KIC's inability to distinguish between dwarfs and 
-giants with only limited photometry -- although Brown et al [2011]'s idea of 
-using a specific ``D51'' photometric band as a $\log g$ diagnostic was reported 
-to have helped with this problem.
-In Fig.~\ref{fig:teff_distribn}, we see something resembling the K dwarf desert 
-again (5000K $\approx$ K3V dwarf).
-Although there are a few slightly worrying discrepancies (notably in 
-Fig.~\ref{fig:rstar_more_distribn}), we get the idea that Galaxia at least 
-roughly resembles ``reality'' (inasmuch as the reported KIC parameters 
-represent reality!).
-
-\subsection{Introduce binaries}
-\label{subsec:introduce_binaries}
-
-By default, Galaxia does not include binaries.
-However, from Fig.~\ref{fig:rmag_distribn}, we've seen that its number counts 
-as a function of apparent magnitude are good.
-So we assign binarity to the Galaxia stars as follows.
-\begin{itemize}
-	\item Assume a binary fraction ${\rm BF} = 0.44$ [Raghavan et al 2010].
-	\item Ignore higher order multiples.
-	\item If a star is drawn to be a binary:
-	\subitem Its reported magnitude becomes a system magnitude.
-	\subitem Draw the binary's mass ratio $q$ from the probability distribution 
-	function for a magnitude limited survey, i.e. from 
-	Fig.~\ref{fig:q_distribn_mag_limited}. This is sensible so long as the pdf 
-	is independent of the primary's mass. We assume this to be the case over 
-	$0.7-1.3M_\odot$\footnote{It was not immediately self-evident to me that 
-	this distribution was the correct choice, since for fixed mass ratio as the 
-	primary mass changes so does the detectable volume. However, as long as 
-	${\rm prob}(q)$ is independent of the primary's mass, the average 
-	distribution over all the possible detectable volumes will remain the 
-	same.}.
-	\subitem For $q \geq 5/7$, it turns out the above is sufficient information 
-	to analytically specify the individual stellar masses, and thus 
-	luminosities with the relation shown in Fig.~\ref{fig:mass_luminosity}.
-	\subitem For $q < 5/7$, the individual stellar masses are not uniquely 
-	specified\footnote{See the derivation of 2017/08/25. This is because of the 
-	mass luminosity relation we specified, and the fact that we have $q$, and 
-	$L_d$, but want individual masses, and individual luminosities.}. Thus for 
-	this latter fraction of the population, we draw the 
-	primary mass from the pdf of single star primary masses.
-	\subitem With the individual masses known, use Eq.~\ref{eq:mass_radius} to 
-	assign a radius to each star in the binary system.
-\end{itemize}
-
-Benefits of the above procedure include that it does not change Galaxia's 
-star counts as a function of magnitude.
-It produces the correct distribution of the mass ratio for the binaries in a 
-magnitude limited sample, and thus a reasonable light ratio distribution.
-
-The drawbacks are that the individual component masses this procedure produces 
-are not drawn from the same distribution throughout.
-This might bias any subsequent synthetic transit survey.
-From Fig.~\ref{fig:galaxia_scatter_masses}, it's likely not excessively 
-important (it's more important to get the mass ratio and light ratio 
-distributions correct).
-
-In addition, the procedure produces a different mass-radius relation 
-for stars in binary systems, vs. stars in single star systems.
-This might also bias any subsequent synthetic transit survey.
-The mass-radius relation for single stars comes from the Padova isochrones 
-[Marigo et al 2008, Marigo \& Girardi 2007, Girardi et al. 2000, Bertelli et al 
-1994] used in Galaxia.
-Although Galaxia does not directly report radii, it reports luminosities and 
-effective temperatures, from which we compute the radii.
-The alternative choice would be to use the reported Galaxia masses with our 
-Eq.~\ref{eq:mass_radius} to produce radius estimates.
-This would produce consistency in the mass-radius relation, but would break $L 
-= 4\pi R^2 \sigma T_{\rm eff}$ for single star systems.
-The different mass-radius relations are shown in 
-Fig.~\ref{fig:galaxia_mass_radius}.
-Generally, it seems that the at a given mass, the stars in binary systems are 
-biased to have slightly greater radii. (Ignoring the evolved stars).
-However this is a small effect -- perhaps of order 10\% in the radius.
-This will affect the transit depths in a synthetic transit survey, making it 
-more difficult to detect planets around binaries.
-
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/galaxia_gammaR_distribn_mag_limited.pdf}
-	\end{center}
-	\caption{Light ratio distribution for the $r<17$ Galaxia sample discussed 
-	in Sec.~\ref{subsec:introduce_binaries}.}
-	\label{fig:galaxia_gammar_distribn}
-\end{figure}
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/galaxia_q_distribn_mag_limited.pdf}
-	\end{center}
-	\caption{Mass ratio distribution for the $r<17$ Galaxia sample discussed 
-		in Sec.~\ref{subsec:introduce_binaries}.}
-	\label{fig:galaxia_q_distribn}
-\end{figure}
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/galaxia_gammaR_scatter_masses.pdf}
-	\end{center}
-	\caption{Scatter plot of primary and secondary mass (solar units).}
-	\label{fig:galaxia_scatter_masses}
-\end{figure}
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/galaxia_gammaR_scatter_luminosities.pdf}
-	\end{center}
-	\caption{Scatter plot of primary and secondary luminosity (solar units).}
-	\label{fig:galaxia_scatter_luminosities}
-\end{figure}
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=.8]{figures/galaxia_compare_mass_radius.pdf}
-	\end{center}
-	\caption{Scatter plot of mass vs radius for singles and binaries. For 
-	binaries, different points are shown for the primary and secondary of a 
-	given system. Only points with $0.7<M_\star/M_\odot<1.3$ are selected.}
-	\label{fig:galaxia_mass_radius}
-\end{figure}
-
-\subsection{Using Galaxia to construct a KIC analog}
-
-With glowing confidence, we then proceed to construct a KIC analog as follows.
-We take all Galaxia stars within 7.5 degrees of the center of the \kepler\ 
-field, and take stars with $r<17$.
-\todo[inline]{a different magnitude cut, e.g. $r<16$, might matter?}
-This is now beyond where the KIC is complete, so direct comparisons of 
-distributions with the KIC will be confused.
-We then apply an analog of the prioritization scheme used by the \kepler 
-mission to select target stars, described by Batalha et al [2010]'s Table 1.
-
-For each star, we compute the minimum detectable planet radius at three 
-semimajor axes: (1) the inner radius of the so-called ``habitable'' zone,
-\begin{equation}
-a_{\rm HZ} = 0.95\, {\rm AU} \left(\frac{L}{L_\odot}\right)^{1/2},
-\end{equation}
-(2) $0.5 a_{\rm HZ}$, and (3) $5R_\star$. We make the same assumptions as 
-Batalha et al [2010] did about the transit duration. We assume a noise model of 
-pure shot noise, rather than using any CDPP estimates.
-We then apply the Batalha et al [2010] Table 1 prioritization, except 
-that we use SDSS $r$ magnitudes instead of Kepler $Kp$ magnitudes.
-The rankings are shown in Fig.~\ref{fig:batalha_table_1}.
-We get similar (within a factor of two) counts as they did towards the end of 
-the table, though at the beginning we seem to have fewer stars\footnote{I 
-suspect this is because of the slightly different noise calculation -- and 
-perhaps Batalha using $N_{\rm sample}$ rather than just counting photons 
-matters}.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.3]{figures/batalha_2010_table_1.png}
-	\end{center}
-	\caption{Prioritization applied by Batalha et al [2010], from whom I 
-	copy-pasted this Table. I did the same, but using $r$ magnitudes instead of 
-	$Kp$.}
-	\label{fig:batalha_table_1}
-\end{figure}
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=0.8]{figures/KIC_vs_GIC_teff_distribn.pdf}
-	\end{center}
-	\caption{Effective temperature distribution of the stars in our synthetic 
-		Kepler input catalog analog -- the Galaxia input catalog.
-		Note that the effective temperatures of binary star systems in the GIC 
-		are 
-		intentionally wrongly estimated, as if they were single star systems.
-		The listed $T_{\rm eff}$ to spectral type correspondence is from 
-		Mamajek's tables.
-	}
-	\label{fig:KIC_vs_GIC_teff_distribn}
-\end{figure}
-
-Note that, since we assigned binarity to Galaxia stars, applying the Batalha et 
-al [2010] prioritization procedure requires estimating \textit{incorrect} 
-single star parameters for what are really binary star systems!
-We do this by keeping the total system luminosity, inverting 
-Fig.~\ref{fig:mass_luminosity} to find an incorrect mass, and then applying 
-Eq.~\ref{eq:mass_radius} to get an incorrect radius.
-We can then calculate the minimum detectable planet radius as
-\begin{equation}
-R_{p,{\rm min}} = \left(\frac{7.1 \sigma_{\rm tot}}{r}\right)^{1/2} R_\star,
-\label{eq:min_det_radius}
-\end{equation}
-for $r=1$ the dilution that we ignore, and
-\begin{equation}
-\sigma_{\rm tot} \equiv {\rm N} = (F_\gamma^{\rm N} A N_{\rm tra} T_{\rm 
-dur})^{-1/2}
-\end{equation}
-
-In applying the prioritization, we omit the latter two priority classes of 
-Fig.~\ref{fig:batalha_table_1}, to obtain a catalog of 152682 stellar systems, 
-which we hereafter refer to as the \textit{Galaxia Input Catalog (GIC)}.
-We compare the effective temperature distribution of the GIC with 
-that of the KIC, reported by Batalha et al [2010] Table 2, in 
-Fig.~\ref{fig:KIC_vs_GIC_teff_distribn}. Evidently, the GIC is 
-dominated by sun-like stars, in a very similar manner to the exoplanet target 
-stars of the KIC.
-
-\subsection{Introduce a planet population}
-
-As in Sec.~\ref{sec:model_2}, we allow for three different occurrence rates: 
-$\Gamma_{t,s}$, the fraction of stars in single systems with a planet of radius 
-$R_p$ and orbital period $P$, and also $\Gamma_{t,d1}$ and $\Gamma_{t,d2}$ for 
-the fraction per sun-like primary and secondary (respectively) of double star 
-systems with a planet of $(R_p, P)$.
-
-We then randomly select which stars get a planet, and using the true stellar 
-masses and radii compute the corresponding impact parameters for the planets as
-\begin{equation}
-b = \frac{a}{R_\star} \cos i, \quad \cos i \sim \mathcal{U}(0,1).
-\end{equation}
-Any planet with $|b| < 1$ is transiting, and has its transit duration evaluated 
-as
-\begin{equation}
-T_{\rm dur} = 13\,{\rm hr} \left(\frac{P}{{\rm yr}}\right)^{1/3} 
-\left(\frac{\rho_\star}{\rho_\odot}\right)^{-1/3} \sqrt{1 - b^2},
-\end{equation}
-which assumes circular orbits.
-
-\subsection{Run the survey}
-We run the following survey:
-$A = 0.708\,{\rm m}^2$ (\kepler\ effective area), $T_{\rm obs} = 20\,{\rm 
-years}$ (to detect more Earth-like planets than \kepler\ did), and $x_{\rm min} 
-= 7.1$, the same signal to noise threshold as \kepler.
-
-For simplicity, we assign a $r$ band zero point of $10^6\,{\rm ph/s/cm}^2$ to 
-correspond to an $r=0$. We use this to compute the observed 
-photon number fluxes for every system, regardless of whether it has a planet.
-We then compute the signal to noise distribution of transit events, and the 
-number of detections, according to Eq.~\ref{eq:snr_ivory_tower}.
-
-\subsection{Numerically assess what occurrence rates would be derived, and 
-	compare to previous models}
-
-\subsubsection{If we ignore binarity, how wrong is our occurrence rate for 
-	planets of radius $R_p$?}
-
-This is similar to what we did in Sec.~\ref{subsubsec:model2_A}, with the 
-exception that we need to specify the calculation more precisely.
-
-Specifically, our ``Astronomer A'' in this case will estimate an occurrence 
-rate for planets of radius $R_p$ and orbital period $P$ as
-\begin{equation}
-\Gamma_{{\rm A}, R_p} = \frac{N_{\rm det,s}}{Z},
-\label{eq:rate_A}
-\end{equation}
-for
-\begin{equation}
-Z \approx (N_s + N_d) \frac{1}{J} \sum_{j=1}^{J} Q_j,
-\end{equation}
-for $Q_j$ the total detection efficiency, indexed over single and double 
-systems, given as the product of the geometric transit probability and the 
-system-specific completeness:
-\begin{equation}
-Q_j \equiv f_g^{(j)} f_c^{(j)}.
-\end{equation}
-
-Keep in mind that for Astronomer A, we need to compute the 
-\textit{incorrect} transit probability for double star systems (since 
-Astronomer A is assuming that they are single stars).
-
-To evaluate the system-level completeness, we assume that Astronomer A performs 
-something equivalent to perfect injection-recovery.
-For a stellar system with known properties (stellar and planetary), and a 
-perfectly known noise model, the probability of detection is a binary function: 
-either ${\rm S/N} > {\rm (S/N)_{min}}$, or it is not\footnote{There are 
-probabilities involved if we allow for any uncertainty in the parameters, but 
-this is too complicated for our current goal.}.
-So we compute $f_c^{(j)}$ as Astronomer A would: an array (over star systems) 
-of zeros wherever the $(R_p,P)$ planet could not be detected, and ones where it 
-could.
-We then evaluate the total detection efficiency $Q_j$ for each system, and 
-compute the occurrence rate as in Eq.~\ref{eq:rate_A}.
-The resulting occurrence rates are only slightly smaller than those in 
-Sec.~\ref{subsubsec:close_enough} discussed below and shown in 
-Fig.~\ref{fig:error_vs_occrate}.
-
-\subsubsection{What if we ignore binarity, but count planets whose derived 
-radii are ``close enough'' to $R_p$?}
-\label{subsubsec:close_enough}
-
-As in Sec.~\ref{sec:model_2}, we now count planets with observed radii from 
-$(1-x)R_\oplus < R_p^{\rm obs} < R_\oplus$. We'll use $x=0.1$, but it can be 
-whatever we want.
-The occurrence rate in this case will be
-\begin{equation}
-\Gamma_{A,R_p'} = \frac{N_{\rm det,s} + N_{\rm det,d}(R_p')}{Z'},
-\end{equation}
-for
-\begin{equation}
-Z' \approx (N_s + N_d) \frac{1}{J} \sum_{j=1}^{J} Q_j',
-\end{equation}
-where the geometric transit probability is the same as before, but the 
-system-specific completeness is now a function of the desired radius interval:
-\begin{equation}
-Q_j' = f_g^{(j)} f_c^{(j)}(R_p').
-\end{equation}
-There are three possible cases for the completeness, specified by the minimum 
-detectable planet radius $R_{p,{\rm min}}$ (computed according to 
-Eq.~\ref{eq:min_det_radius}):
-\begin{align}
-f_c^{(j)}(R_p') &= 
-\Bigg\{\begin{array}{lr}
-1 & {\rm\ if\ }R_{p,{\rm min}}>R_p,\\
-0, & {\rm\ if\ }R_{p,{\rm min}}<(1-x)R_p,\\
-\frac{R_p - R_{p,{\rm min}}}{R_p - (1-x)R_p}, &{\rm otherwise}.
-\end{array}
-\label{eq:completeness_close_enough}
-\end{align}
-The latter term above is the case in which the minimum detectable planet 
-radius happens to be between $(1-x)R_p$ and $R_p$. In that case the 
-completeness becomes the fraction of planets in that interval that would be 
-detected.
-Assuming a uniform prior for the planet radius distribution $R_p \sim 
-\mathcal{U}((1-x)R_p, R_p)$, this becomes the expression given in 
-Eq.~\ref{eq:completeness_close_enough}.
-
-These occurrence rates can then be compared to the ``true occurrence rate'' 
-derived by the all-knowing Astronomer D:
-\begin{equation}
-\Gamma_{{\rm D}, R_p} = \frac{\Gamma_{t,s} N_s + (\Gamma_{t,d1} + 
-						\Gamma_{t,d2})N_d}{N_s + 2 N_d}.
-\end{equation}
-A note in passing: the more information-rich parameters for Astronomer D to 
-report would be $(\Gamma_{t,s}, \Gamma_{t,d1}, \Gamma_{t,d2})$, rather than 
-$\Gamma_{{\rm D}, R_p}$. But I digress.
-
-Similar to Fig.~\ref{fig:XGamma_vs_xi}, Fig.~\ref{fig:error_vs_occrate} 
-shows that the magnitude of the correction needed by the $\Gamma_{{\rm 
-A},R_p'}$ estimate depends strongly 
-on the fraction of secondaries that are assumed to host the $(R_p, P)$ planets.
-The simplest assumption is that the occurrence rate is the same for all solar 
-type stars: any ``Sun-like'' ($0.7<M_\star/M_\odot<1.3$) star hosts planets at 
-the ``true rate'' $\Gamma_t$, and any other mass of star does not.
-Given how we have defined the GIC, this means that all single stars and 
-primaries of double star systems are sun-like. The fraction of Sun-like 
-secondaries in this case is 37.4\% of all secondaries.
-Reading the appropriate value off Fig.~\ref{fig:error_vs_occrate}, this 
-corresponds to $\Gamma_{\rm D} = 1.39 \Gamma_{\rm A'}$.
-This gives a relative percentage error of $\delta\Gamma_{{\rm D},R_p} \equiv 
-(\Gamma_{\rm A,R_p'} - \Gamma_{\rm D,R_p}) / \Gamma_{\rm D,R_p} = -28\%$.
-
-\begin{figure}[!t]
-	\begin{center}
-		\includegraphics[scale=0.8]{figures/error_vs_occrate.pdf}
-	\end{center}
-	\caption{ The ``most reasonable'' $\Gamma_{t,d2}$ value is obtained by 
-	assuming that the occurrence rate of $(R_p, P)$ planets is the same for all 
-	solar type stars, and zero otherwise.
-	This applies most immediately to the question of Earth-like planets around 
-	Sun-like stars.
-	}
-	\label{fig:error_vs_occrate}
-\end{figure}
-
-\end{comment}
-
-% \begin{comment}
-% \subsection{How many stars are in the sample?}
-% \subsection{How many planets are in the sample?}
-% \subsection{What is the true occurrence rate?}
-% \subsection{How many planets are detected?}
-% \subsubsection{Analytic completeness}
-% \subsubsection{Deriving ${\rm prob}(x_i)$}
-% \subsubsection{Number of detected planets}
-% \subsection{Astronomer A ignores binarity}
-% \subsection{Astronomer B counts host stars correctly}
-% \subsection{Astronomer C counts host stars correctly and figures out diluted
-% radii}
-% \subsection{Astronomer D counts host stars correctly, figures out diluted
-% radii, and accounts for completeness}
-% \subsection{Numerical verification}
-% \subsection{Representative numbers for a few cases}
-% \subsubsection{If we ignore binarity, for what fraction of detections do we
-% misclassify the radii?}
-% \subsubsection{If we ignore binarity, how wrong is our occurrence rate for
-% planets of radius $R_p$?}
-% \subsubsection{If we ignore binarity, but allow derived planet radii to be
-% ``close enough'', how do we do?}
-% 
-% \end{comment}
-
-\newpage
-
-%\begin{thebibliography}{}
-
-%\end{thebibliography}
-
-\end{document}
+In Sec.~\ref{sec:numerical_methods}, we introduce our numerical approach 
+to the problem, and in Sec.~\ref{sec:analytic_preliminaries} we clarify our 
+terminology.
+We present the analytic and numerical results in 
+Secs.~\ref{sec:model_1}-\ref{sec:model_3}, where each subsection corresponds 
+to each model above.
+We interpret these calculations throughout, and in 
+Sec.~\ref{sec:discussion} discuss their relevance to various questions in 
+the interpretation of transit survey occurrence rates.

### 2017-10-28 memos/introduction_v1.tex
memos: add draft introductions

--- 2017-10-28 paper/introduction.tex
@@
 
-An immediate complication is that, due to dynamical stability or some 
-aspect of planet formation, the occurrence rate of planets may differ 
+An immediate complication is that~--~due to dynamical stability or some 
+aspect of planet formation~--~the occurrence rate of planets may differ 
 between binary and single-star systems.
@@
 ``occurrence rates about primaries'', and
-``occurrence rates about secondaries'' (see {\it e.g.,} Wang et al., 
+``occurrence rates about secondaries'' (this is discussed by Wang et al., 
 2015).
@@
 There may be systematic errors in stellar parameter estimates of such 
-systems, which in this work we neglect.
+systems, but in this work we neglect all but the simplest.
 One might also imagine an apparently-searchable star which in no stellar 
@@
 classification, and the assumed survey completeness.
-Detected planets in binary systems will have underestimated radii because of 
-the diluting flux from the companions, and possibly because they are assumed 
-to orbit the wrong star ({\it e.g.}, Furlan et al. 2017).
+At a given apparent radius, detected planets in binary systems will have 
+underestimated radii because of the diluting flux from the companions, and 
+possibly because they are assumed to orbit the wrong star.
 The number of selected stars should change dependent on the planet radius 
@@
 achieved photometric precision.
-This mean our naive astronomer assumes the fraction of their selected stars 
-which are searchable is 1 (this is their ``assumed completeness'').
+This leads our naive astronomer to an assumed ``completeness'' (fraction of 
+selected stars which are searchable) of 1.
 Binarity confuses this, because the ability to search for planets 
-in binary systems depends on which stellar component hosts the planet, and on 
+in binary system depends on which stellar component hosts the planet, and on 
 how much dilution affects the measurement.
@@
 \paragraph{How Has Binarity Been Considered in Occurrence Rate Measurements?}
-Many authors have reported planet occurrence rates from transit 
-survey data\footnote{
+Many authors have investigated planet occurrence rates\footnote{
     An online list of occurrence rate papers is maintained at 
@@
 }.
-With regard to binarity, the typical approach has been to ignore the 
-issue\footnote{Deacon et al., 2015 are an exception to the rule.}.
-For instance, in a laudable study by Burke et al. (2015), they
-\begin{quote}
-``treat the pipeline completeness as having no uncertainty 
-due to the incomplete understanding of the stellar parameters, eccentricity, 
-and stellar binarity.''
-\end{quote}
-Though they do ``explore sensitivity in the derived planet occurrence rates 
-to alternative assumptions for the stellar parameters and non-zero 
-eccentricities'', they do not for binarity.
-
-Of course, on a system-by-system level stellar multiplicity affects the 
-interpretation of planet candidates. High resolution imaging 
-campaigns have consequently been undertaken to measure the multiplicity of 
-almost all {\it Kepler}\ Objects of Interest 
-(Howell et al. 2011; Adams et al. 2012, 2013; Horch et al. 2012, 2014; 
-Lillo-Box et al. 2012, 2014; Dressing et al. 2014; Law et al. 2014; Cartier et 
-al. 2015; Everett et al. 2015; Gilliland et al. 2015; Wang et al. 2015a, 
-2015b; Baranec et al. 2016).
-The results of these programs have been collected by Furlan et al. 
-(2017), and they represent an important advance in understanding the KOI 
-sample's multiplicity statistics.
-In particular, they can be immediately applied to rectify binarity's influence 
-on the mass-radius diagram (Furlan \& Howell 2017).
-
-For {\it Kepler}-derived occurrence rates, the high resolution imaging 
-campaigns have not yet come full circle to observe a comparison sample of 
-non-KOI host stars.
-The most recent rate studies have thus used Furlan et al. (2017)'s 
-catalog to test the effects of removing KOI hosts with known companions, which 
-is an important step towards reducing contamination in the ``numerator'' of 
-the occurrence rate (Fulton et al. 2017).
-However, without an understanding of the multiplicity statistics of the 
-non-KOI hosting stars that are assumed to be searchable, the 
-true completeness, the true number of searchable stars, and thus the true 
-occurrence rates will remain uncertain.
+The approach towards binarity is to ignore it.
+(There are some exceptions, {\it e.g.}, Deacon et al. 2015).
+%Burke+ 2015: "In this study, we treat the pipeline completeness as having no 
+%uncertainty due to the incomplete understanding of the stellar parameters, 
+%eccentricity, and stellar binarity."
+Of course, at a system-by-system level that this is wrong, and so for
 
 \paragraph{The Hot Jupiter Rate Discrepancy}
-There is at least one context in which measurement of absolute 
-occurrence rates may already be showing the signatures of binarity.
+
 Hot Jupiter occurrence rates measured by transit surveys ($\approx 0.5\%$) are 
@@
 Though the discrepancy is of weak statistical significance ($<3\sigma$),
-one plausible explanation for the difference is that the populations have 
-distinct metallicities.
+one reason to expect a difference in the rates is that the populations have 
+different metallicities.
 As originally argued by Gould et al. (2006), the RV sample is biased towards 
@@
 
-Aside from surveying stars of varying metallicities, radial velocity and 
+Aside from surveying stars of different metallicities, radial velocity and 
 transit surveys differ in how they treat binarity.
@@
 Ignoring these complications, in this work we focus on whether
-binarity might intrinsically bias transit survey occurrence rates, simply 
+binarity might intrinsically bias occurrence rate measurements, simply 
 through its effects on the number of searchable stars and the apparent radii 

### 2017-10-28 paper/figures.tex
paper: add q distribn figure

@@
     \label{fig:errcases_model_1}
+\end{figure}
+
+\begin{figure}[!tb]
+    \centering
+    \includegraphics[width=\textwidth]{figures/q_distribn_mag_limited.pdf}
+    \caption{
+    The distribution of the mass ratio for a magnitude limited sample of 
+    binary stars. The underlying mass ratios are drawn from a uniform 
+    distribution in a volume-limited sample, quite similar to that of Raghavan 
+    et al. (2010)'s Fig. 13.
+    The entire bias can be understood analytically (Eq.~\ref{eq:model_2_p_2},
+    with $\gamma=0$).
+    The numerical comparison uses a realistic empirical mass-luminosity 
+    relation rather than $L=M^{3.5}$, found by fitting main sequence dwarf 
+    mass luminosity data compiled from Torres et al. [2009] and Benedict et 
+    al. [2016].
+    }
+    \label{fig:q_distribn_mag_limited}
 \end{figure}

### 2017-10-28 paper/model_1.tex
paper: add q distribn figure

@@
 \end{equation}
-for $L_{\rm sys}=L_1(1+\gamma_R)$ the system luminosity, $T_{\rm dur}$ the 
+for $L_{\rm sys}=L_1(1+\ell)$ the system luminosity, $T_{\rm dur}$ the 
 transit duration, $A$ the detector area, and $N_{\rm tra}$ 
@@
 1 & \text{for } i=0,\ {\rm single} \\
-L_1 / L_{\rm sys} = (1+\gamma_R)^{-1}, & \text{for } i=1,\ {\rm primary} \\
-\gamma_R L_1 / L_{\rm sys} = (1 + \gamma_R^{-1})^{-1}, & 
+L_1 / L_{\rm sys} = (1+\ell)^{-1}, & \text{for } i=1,\ {\rm primary} \\
+\ell L_1 / L_{\rm sys} = (1 + \ell^{-1})^{-1}, & 
     \text{for } i=2,\ {\rm secondary},
@@
 \end{align}
-where the light ratio $\gamma_R$ of a given binary is defined as the ratio of 
+where the light ratio $\ell$ of a given binary is defined as the ratio of 
 the luminosity of the secondary to the primary.
@@
 \frac{n_b}{n_s} \left(\frac{d_{\rm sel,b}}{d_{\rm sel,s}}\right)^3 = 
-\frac{{\rm BF}}{1-{\rm BF}} (1+\gamma_R)^{3/2},
+\frac{{\rm BF}}{1-{\rm BF}} (1+\ell)^{3/2},
 \label{eq:mu_definition}

### 2017-10-28 paper/model_2.tex
paper: add q distribn figure

@@
 The main use of our binary-twin model is to help develop intuition.
-We now let the light ratio $\gamma_R = L_2/L_1$ vary across the binary 
+We now let the light ratio $\ell = L_2/L_1$ vary across the binary 
 population.
@@
 Malmquist-like bias, $(1+q^\alpha)^{3/2}$.
-We note that various authors (ourselves included) have incorrectly 
-used volume-limited binary distributions in Monte Carlo simulations of 
-transit surveys (Sullivan et al 2015, Hartman and Bakos XX, Guenther et al 
+We note that various authors (probably ourselves included) have incorrectly 
+drawn from volume-limited binary distributions in Monte Carlo simulations of 
+transit surveys (Sullivan et al 2015, Bakos et al 2012, Guenther et al 
 2017, Bouma et al 2017).
+The correct distribution for a magnitude-limited selection is shown in 
+Fig.~\ref{fig:q_distribn_mag_limited}.
 

### 2017-10-28 paper/conclusion.tex
paper: add q distribn figure
wordsmithing

--- 2017-10-28 paper/conclusion.tex
@@
 
-This work presented three simple models for the effects of binarity on 
+This study presented three simple models for the effects of binarity on 
 occurrence rates measured by transit surveys.
+The most realistic of these models (Model \#3) suggests that binarity leads to 
+underestimates of occurrence rates at less than $30\%$ relative 
+error.
+The model indicates that hot Jupiter rates measured by transit surveys are 
+biased to infer $\approx 2$ fewer hot Jupiters per thousand single stars than 
+surveys that only measure occurrence rates about single stars ({\it i.e..,} 
+radial velocity surveys).
+Our third model also indicates that binarity's effects on the measured 
+occurrence rates of Earth-sized planets are far smaller than current 
+systematic uncertainties.
+Though our models are simplistic, their agreement with Barclay \& 
+Collaborator's recent detailed simulations indicate that they capture the 
+essential ingredients.

### 2017-11-05 paper/main.tex
paper: add q distribn figure
wordsmith

@@
 
-\shortauthors{LGB, JNW}
+\shortauthors{Bouma, Winn, Masuda(?)}
 \shorttitle{Binarity and Occurrence Rates}
@@
 \author{
-L. G. Bouma, J. N. Winn
+L. G. Bouma, J. N. Winn, K. Masuda(?)
 }
@@
 For the occurrence of Earth-sized planets, in our most realistic model the 
-underestimate is by $\approx 15\%$~--~at present far smaller than other 
-systematic uncertainties.
+underestimate is by $\approx 15\%$~--~at present far smaller than 
+systematic uncertainties on real measurements of this number.
 For hot Jupiters, we find that the inferred occurrence rate is 

### 2017-11-06 src/numerical_models.py
changed some src files to be symlinks

@@ -1,534 +1,1 @@
-# -*- coding: utf-8 -*-
-'''
-See README.md for full description.
-
-Example usage:
->>> python numerical_models.py --quickrun --modelnumber 3 --LambdaTwo 0.5
-
-Alternatively, write a wrapper like `run_numerical_models.py`.
-'''
-
-from __future__ import division, print_function
-
-import numpy as np, pandas as pd, matplotlib.pyplot as plt
-from astropy import units as u, constants as c
-from scipy.integrate import trapz
-from scipy.interpolate import interp1d
-from math import pi as π
-import os, argparse, decimal
-
-def numerical_transit_survey(
-        quickrun,
-        model_number,
-        Λ_2,
-        BF=0.44,
-        α=3.5,
-        β=0,
-        γ=0,
-        δ=-2.92,
-        Λ_0=0.5,
-        Λ_1=0.5
-        ):
-    '''
-    Inputs:
-        quickrun = False
-        model_number = 3
-
-        BF = 0.44   # binary fraction. BF = n_d / (n_s+n_d). Raghavan+ 2010 solar.
-
-        α = 3.5     # exponent in L~M^α
-        β = 0       # exponent in p_vol_limited(q) ~ q^β
-        γ = 0       # exponent in p(has planet|secondary w/ q) ~ q^γ
-        δ = -2.92   # exponent in p(r) ~ r^δ, from Howard+ 2012
-
-        Λ_0 = 0.5 # fraction of selected singles with planet
-        Λ_1 = 0.5 # fraction of selected primaries with planet
-        Λ_2 = 0.5 # fraction of selected secondaries with planet.
-
-    Outputs:
-        Four data files (in ../data/):
-
-        'results_model_{:d}_Lambda2_{:.2f}'.format( model_number, Λ_2)
-
-        'results_model_{:d}_error_case_{:d}_Lambda2_{:.2f}.out'.format(
-            model_number, error_case_number, Λ_2)
-
-        The output data files contain binned occurrence rates as a function of
-        true and apparent radius.
-
-    '''
-
-    slowrun = not quickrun
-
-    print('Running Model #{:d}.'.format(model_number))
-
-    if γ != 0:
-        raise NotImplementedError
-
-    if abs(decimal.Decimal(str(Λ_2)).as_tuple().exponent) > 2:
-        # File names assume 2 digits of precision for Λ_2.
-        raise NotImplementedError
-
-    ######################
-    # STELLAR POPULATION #
-    ######################
[…]

### 2017-11-06 src/plot_HJ_correction_vs_LambdaTwo.py

@@ -1,94 +1,1 @@
-'''
-Given that hot Jupiters are thought to be less common around less-massive
-stars, it would be more sensible to consider Λ_2 < Λ_0 , while letting single
-stars and primaries host planets at the same rate.
-
-This plots the rates as a function of Λ_2 /Λ_0.
-'''
-# -*- coding: utf-8 -*-
-from __future__ import division, print_function
-import matplotlib as mpl
-mpl.use('pgf')
-pgf_with_custom_preamble = {
-    'pgf.texsystem': 'pdflatex', # xelatex is default; i don't have it
-    'font.family': 'serif', # use serif/main font for text elements
-    'text.usetex': True,    # use inline math for ticks
-    'pgf.rcfonts': False,   # don't setup fonts from rc parameters
-    'pgf.preamble': [
-        '\\usepackage{amsmath}',
-        '\\usepackage{amssymb}'
-        ]
-    }
-mpl.rcParams.update(pgf_with_custom_preamble)
-
-import matplotlib.pyplot as plt
-import numpy as np, pandas as pd
-
-if __name__ == '__main__':
-
-    model_number = 3
-
-    Λ_2_arr = np.arange(0, 0.5+0.05, 0.05)
-
-    Λ_HJ_true_arr, Λ_HJ_inferred_arr, Λ_HJ_true_primary_arr = [], [], []
-
-    # compute the HJ correction factors for each Λ_2
-    for Λ_2 in Λ_2_arr:
-
-        savdir = '../data/numerics/'
-        fname = 'results_model_{:d}_Lambda2_{:.2f}.out'.format(
-                model_number, Λ_2)
-        df = pd.read_csv(savdir+fname)
-
-        lower_bound = 8
-        inds = df['bin_left'] > lower_bound
-        Λ_HJ_true = np.sum(df[inds]['true_Γ'])
-        Λ_HJ_inferred = np.sum(df[inds]['inferred_Γ'])
-        Λ_HJ_true_primary = np.sum(df[inds]['true_primary_Γ'])
-
-        Λ_HJ_true_arr.append(float(Λ_HJ_true))
-        Λ_HJ_inferred_arr.append(float(Λ_HJ_inferred))
-        Λ_HJ_true_primary_arr.append(float(Λ_HJ_true_primary))
-
-    Λ_HJ_true_arr = np.array(Λ_HJ_true_arr)
-    Λ_HJ_inferred_arr = np.array(Λ_HJ_inferred_arr)
[…]

### 2017-11-06 src/plot_errcases_rate_densities_vs_radius.py

@@ -1,223 +1,1 @@
-'''
-first, for models 1,2, 3:
->>> python numerical_models.py
-then:
->>> python plot_absolute_rate_densities_vs_radius.py
-
-makes plots for paper
-'''
-# -*- coding: utf-8 -*-
-from __future__ import division, print_function
-import matplotlib as mpl
-mpl.use('pgf')
-pgf_with_custom_preamble = {
-    'pgf.texsystem': 'pdflatex', # xelatex is default; i don't have it
-    'font.family': 'serif', # use serif/main font for text elements
-    'text.usetex': True,    # use inline math for ticks
-    'pgf.rcfonts': False,   # don't setup fonts from rc parameters
-    'pgf.preamble': [
-        '\\usepackage{amsmath}',
-        '\\usepackage{amssymb}'
-        ]
-    }
-mpl.rcParams.update(pgf_with_custom_preamble)
-
-import matplotlib.pyplot as plt
-import pandas as pd
-import numpy as np
-
-from brokenaxes import brokenaxes
-
-def make_plot(model_number, logx=None, logy=None, withtext=None,
-        stdout=False, brokenx=False, Λ_2=None):
-
-    # Make summary plot
-    if not brokenx:
-        f, ax = plt.subplots(figsize=(4,4))
-    else:
-        #cf https://github.com/bendichter/brokenaxes/blob/master/brokenaxes.py
-        f = plt.figure(figsize=(4,4))
-        bigax = brokenaxes(
-                xlims=((0.695, .715), (.985, 1.005)),
-                d=0.02,
-                tilt=85,
-                hspace=.05,
-                despine=True)
-        ax=bigax
-
-    if isinstance(Λ_2,float) or isinstance(Λ_2,int):
-        fname = '../data/numerics/results_model_{:d}_Lambda2_{:.1f}.out'.format(
-                model_number, Λ_2)
-    else:
-        fname = '../data/results_model_'+repr(model_number)+'.out'
-
-    df = pd.read_csv(fname)
-
-    if model_number == 3:
-        xvals = np.append(0, df['bin_left'])
-        ytrue = np.append(0, df['true_Γ'])
-        yinferred = np.append(0, df['inferred_Γ'])
-    elif model_number == 1 or model_number == 2:
-        xvals = np.append(df['bin_left'],[1,1.1])
-        ytrue = np.append(df['true_Γ'],[0,0])
-        yinferred = np.append(df['inferred_Γ'],[0,0])
-
-    ax.step(xvals, ytrue, where='post', label='true')
-
[…]

### 2017-11-06 src/plot_print_earth_rate_analog.py

@@ -1,97 +1,1 @@
-'''
-What is the rate correction for earth analogs in Model #3?
-
-Print the result.
-
-Make the plot of rate as a function of Λ_2/Λ_0 as well.
-'''
-# -*- coding: utf-8 -*-
-from __future__ import division, print_function
-import matplotlib as mpl
-mpl.use('pgf')
-pgf_with_custom_preamble = {
-    'pgf.texsystem': 'pdflatex', # xelatex is default; i don't have it
-    'font.family': 'serif', # use serif/main font for text elements
-    'text.usetex': True,    # use inline math for ticks
-    'pgf.rcfonts': False,   # don't setup fonts from rc parameters
-    'pgf.preamble': [
-        '\\usepackage{amsmath}',
-        '\\usepackage{amssymb}'
-        ]
-    }
-mpl.rcParams.update(pgf_with_custom_preamble)
-
-import matplotlib.pyplot as plt
-
-import numpy as np, pandas as pd
-
-if __name__ == '__main__':
-
-    model_number = 3
-
-    Λ_0 = 0.5
-    Λ_2_arr = np.arange(0, 0.5+0.05, 0.05)
-
-    Λ_earth_true_arr, Λ_earth_inferred_arr, Λ_earth_true_primary_arr = [], [], []
-
-    # compute the rate for each Λ_2
-    for Λ_2 in Λ_2_arr:
-
-        savdir = '../data/numerics/'
-        fname = 'results_model_{:d}_Lambda2_{:.2f}.out'.format(
-                model_number, Λ_2)
-        df = pd.read_csv(savdir+fname)
-
-        for ix, bl in enumerate([0.5, 1.0]):
-
-            inds = (df['bin_left'] == bl)
-
-            Λ_earth_true = df[inds]['true_Γ']
-            Λ_earth_inferred = df[inds]['inferred_Γ']
-            Λ_earth_true_primary = df[inds]['true_primary_Γ']
-
-            Λ_earth_true_arr.append(float(Λ_earth_true))
-            Λ_earth_inferred_arr.append(float(Λ_earth_inferred))
-            Λ_earth_true_primary_arr.append(float(Λ_earth_true_primary))
-
-    Λ_earth_true_arr = np.array(Λ_earth_true_arr)
-    Λ_earth_inferred_arr = np.array(Λ_earth_inferred_arr)
[…]

### 2017-11-06 src/plot_rate_densities_vs_radius.py

@@ -1,164 +1,1 @@
-'''
-first, for models 1,2, 3:
->>> python numerical_models.py
-then:
->>> python plot_rate_densities_vs_radius.py
-
-makes plots for paper
-'''
-# -*- coding: utf-8 -*-
-from __future__ import division, print_function
-import matplotlib as mpl
-mpl.use('pgf')
-pgf_with_custom_preamble = {
-    'pgf.texsystem': 'pdflatex', # xelatex is default; i don't have it
-    'font.family': 'serif', # use serif/main font for text elements
-    'text.usetex': True,    # use inline math for ticks
-    'pgf.rcfonts': False,   # don't setup fonts from rc parameters
-    'pgf.preamble': [
-        '\\usepackage{amsmath}',
-        '\\usepackage{amssymb}'
-        ]
-    }
-mpl.rcParams.update(pgf_with_custom_preamble)
-
-import matplotlib.pyplot as plt
-import pandas as pd
-import numpy as np
-
-from brokenaxes import brokenaxes
-
-def make_plot(model_number, logx=None, logy=None, withtext=None,
-        stdout=False, brokenx=None, Λ_2=None):
-
-    # Make summary plot
-    if isinstance(Λ_2,float) or isinstance(Λ_2,int):
-        fname = '../data/numerics/results_model_{:d}_Lambda2_{:.1f}.out'.format(
-                model_number, Λ_2)
-    else:
-        fname = '../data/results_model_'+repr(model_number)+'.out'
-
-    df = pd.read_csv(fname)
-
-    if not brokenx:
-        f, ax = plt.subplots(figsize=(4,4))
-    else:
-        #cf https://github.com/bendichter/brokenaxes/blob/master/brokenaxes.py
-        f = plt.figure(figsize=(4,4))
-        bigax = brokenaxes(
-                xlims=((0.695, .715), (.985, 1.005)),
-                d=0.02,
-                tilt=85,
-                hspace=.05,
-                despine=True)
-        ax=bigax
-
-    if model_number == 3:
-        xvals = np.append(0, df['bin_left'])
-        ytrue = np.append(0, df['true_Γ'])
-        yinferred = np.append(0, df['inferred_Γ'])
-    elif model_number == 1 or model_number == 2:
-        xvals = np.append(df['bin_left'],[1,1.1])
-        ytrue = np.append(df['true_Γ'],[0,0])
[…]

### 2017-11-06 src/run_numerical_models.py

@@ -1,24 +1,1 @@
-'''
-Wrapper to `numerical_models.py`.
-
-Useful for running Model #3 with varying Λ_2 in parallel.
-'''
-
-import numpy as np
-from multiprocessing import Pool
-
-from numerical_models import numerical_transit_survey
-
-if __name__ == '__main__':
-
-    model_number = 3
-    quickrun = False
-    Λ_2_arr = np.arange(0., 0.5+0.05, 0.05)
-
-    paramlist = [(quickrun, model_number, Λ_2) for Λ_2 in Λ_2_arr]
-
-    with Pool(6) as p:
-        p.starmap(
-            numerical_transit_survey,
-            paramlist
-            )
+../../binary_biases/src/run_numerical_models.py
